pdfoutput=1
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\pdfoutput=1
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Spring Self-Evolution Algorithm --- Mathematical Genealogy}\label{spring-self-evolution-algorithm-mathematical-genealogy}

\begin{quote}
\textbf{Spring} = Spring Self-Evolving Gatekeeper (SSEG)
\textbf{Status}: Core algorithm of Paper 2 (Nature Computational Science target)
\textbf{Document type}: Theoretical root tracing --- mathematical domain attribution, key idea sources, historical development lineage, comparison with prior work
\textbf{Date}: 2026-06-28
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Contents}\label{contents}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \hyperref[1-mathematical-domain-attribution]{Mathematical Domain Attribution}
\item
  \hyperref[2-key-ideas-and-proof-tools]{Key Ideas and Proof Tools}
\item
  \hyperref[3-historical-development-timeline]{Historical Development Timeline}
\item
  \hyperref[4-spring-vs-prior-work-comparison-table]{Spring vs Prior Work Comparison Table}
\item
  \hyperref[5-springs-unique-contributions]{Spring's Unique Contributions}
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Mathematical Domain Attribution}\label{mathematical-domain-attribution}

The Spring self-evolution algorithm does not belong to a single mathematical domain; rather, it spans seven classical mathematical fields, synthesizing a completely new structure at their intersection.

\subsubsection{1.1 Stochastic Approximation}\label{stochastic-approximation}

\textbf{Roots}: Robbins \& Monro (1951). ``A stochastic approximation method.'' \emph{Annals of Mathematical Statistics}, 22(3), 400-407.

\textbf{Role in Spring}: The online update of NEP student parameters $\theta_t$ follows the Robbins-Monro recursion:
\[\theta_{t+1} = \theta_t - \alpha_t \nabla_\theta \ell(f_{\theta_t}(x_t), y_t)\]
where the learning rate $\alpha_t$ satisfies the standard RM conditions: $\sum \alpha_t = \infty$, $\sum \alpha_t^2 < \infty$.

Spring generalizes standard RM to a \textbf{coupled two-timescale} setting:
- \textbf{Fast timescale}: Student $\theta_t$ updates at rate $\alpha_t$ under fixed gatekeeper $S_t$.
- \textbf{Slow timescale}: Gatekeeper $S_t$ updates at rate $\beta_t = o(\alpha_t)$ under student quasi-steady-state.

This corresponds to Borkar (2008)'s two-timescale stochastic approximation framework. Condition C6$'$ ($\beta_t = o(\alpha_t)$) is the key innovation --- it simultaneously guarantees local convergence of the student and bounded cumulative distribution shift.

\textbf{References}: - Robbins \& Monro (1951) - Kushner \& Yin (2003). \emph{Stochastic Approximation and Recursive Algorithms and Applications} - Borkar (2008). \emph{Stochastic Approximation: A Dynamical Systems Viewpoint} - Polyak \& Juditsky (1992). ``Acceleration of stochastic approximation by averaging''

\subsubsection{1.2 Dynamical Systems / Lyapunov Stability}\label{dynamical-systems-lyapunov}

\textbf{Roots}: Lyapunov, A. M. (1892). \emph{The General Problem of the Stability of Motion}. Kharkov Mathematical Society.

\textbf{Role in Spring}: The Spring self-evolution loop is a discrete dynamical system:
\[z_{t+1} = \Phi(z_t), \quad z_t = (S_t, M_t, \theta_t) \in \mathcal{Z}\]
where $\Phi = (\Phi_S, \Phi_M, \Phi_\theta)$ is the update operator, $\mathcal{Z} = \mathcal{F} \times \mathcal{M} \times \Theta$ is the evolution state space.

The Lyapunov function candidate is (evaluated on a fixed reference set $M_0$):
\[\Psi(S_t, \theta_t) = \frac{1}{|M_0|} \sum_{x \in M_0} (S_t(x, y(x)) - \hat{C}(x))^2 + \lambda \cdot \frac{1}{|V_0|} \sum_{(x,y) \in V_0} \ell(f_{\theta_t}(x), y)\]

The Lyapunov descent property $\mathbb{E}[\Psi_{t+1} \mid \mathcal{F}_t] \leq \Psi_t$ is fully proved via Theorem 12.5 under the reference set replay mechanism. The fixed point $(S^*, \theta^*)$ satisfies the self-consistent coupling equations, and the $\omega$-limit set is a singleton.

Spring employs the complete toolkit of dynamical systems: attractors, basins, phase diagrams (four convergence mechanisms), and $\omega$-limit sets.

\textbf{References}: - Lyapunov (1892). \emph{The General Problem of the Stability of Motion} - Khalil (2002). \emph{Nonlinear Systems} (3rd ed.) - Strogatz (2018). \emph{Nonlinear Dynamics and Chaos} (2nd ed.) - Conley (1978). \emph{Isolated Invariant Sets and the Morse Index}

\subsubsection{1.3 Bayesian Inference / Martingale Theory}\label{bayesian-inference-martingale}

\textbf{Roots}: Doob, J. L. (1953). \emph{Stochastic Processes}. Wiley.

\textbf{Role in Spring}: The evolution of the gatekeeper scoring function $S_t$ is interpreted as a Bayesian posterior mean update:
\[S_t(x,y) = \mathbb{E}_{S \sim P_t}[S(x,y)], \quad P_{t+1}(S) \propto P_t(S) \cdot \mathcal{L}(\mathcal{D}_{t+1} \mid S)\]

Theorem (Doob's Martingale Convergence): The sequence $\{S_t\}_{t \geq 0}$ is a $[0,1]$-bounded martingale with respect to the filtration $\mathcal{F}_t = \sigma(\mathcal{M}_1, \dots, \mathcal{M}_t)$. By Doob's martingale convergence theorem:
\[S_t(x,y) \xrightarrow{a.s.} S_\infty(x,y), \quad \forall (x,y)\]

Important limitation: The martingale property requires correct model specification ($S^*$ must be in the support of the prior $P_0$). Under model misspecification, the posterior converges to the KL projection $S^\dagger$ rather than the true value $S^*$. Spring honestly annotates this limitation.

\textbf{References}: - Doob (1953). \emph{Stochastic Processes} - Doob (1949). ``Application of the theory of martingales'' - Ghosh \& Ramamoorthi (2003). \emph{Bayesian Nonparametrics} - van der Vaart (1998). \emph{Asymptotic Statistics} - Kleijn \& van der Vaart (2012). ``The Bernstein-von Mises theorem under misspecification''

\subsubsection{1.4 Online Learning / Regret Analysis}\label{online-learning-regret}

\textbf{Roots}: Zinkevich, M. (2003). ``Online convex programming and generalized infinitesimal gradient ascent.'' \emph{ICML 2003}.

\textbf{Role in Spring}: The gatekeeper's online gradient descent (OGD) update produces a standard regret bound:
\[\text{Regret}_T \leq 2GR\sqrt{T} \quad \text{(no delay)}\]

Spring extends this bound to the \textbf{delayed feedback} setting (NEP student feedback delay $d_t \leq D_{\max}$):
\[\text{Regret}_T^{\text{delay}} \leq 2GR\sqrt{T} + G^2 D_{\max} \sqrt{T}\]

Further extends to the \textbf{growing memory bank} setting --- the monotonic growth of the memory bank $M_t$ provides effective sample complexity improvement:
\[\text{Regret}_T^{\text{eff}} \leq \text{Regret}_T^{\text{delay}} + O\left(T \sqrt{\frac{\text{VC}(\mathcal{H})}{N_T}}\right)\]

As $N_T \to \infty$, the system satisfies \textbf{asymptotic no-regret}: $\frac{1}{T}\text{Regret}_T^{\text{eff}} \to 0$.

\textbf{References}: - Zinkevich (2003). ``Online convex programming'' - Hazan (2016). ``Introduction to online convex optimization'' - Joulani et al.~(2016). ``Online learning under delayed feedback'' - Shalev-Shwartz (2012). ``Online learning and online convex optimization'' - Cesa-Bianchi \& Lugosi (2006). \emph{Prediction, Learning, and Games}

\subsubsection{1.5 Concentration Inequalities}\label{concentration-inequalities}

\textbf{Roots}: - Hoeffding, W. (1963). ``Probability inequalities for sums of bounded random variables.'' \emph{JASA}, 58(301), 13-30. - Chernoff, H. (1952). ``A measure of asymptotic efficiency for tests of a hypothesis.'' \emph{Annals of Mathematical Statistics}, 23(4), 493-507.

\textbf{Role in Spring}: At multiple critical junctures in Spring:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Concentration of consensus score}: Hoeffding tail bound for expert agreement $\hat{C}(x) = \frac{1}{M}\sum_{m=1}^M \mathbf{1}\{\ell(f_m(x), y) > \tau\}$:
  \[\mathbb{P}(|C - \mathbb{E}[C]| > t) \leq 2\exp(-2M t^2)\]
  When A2 is violated due to expert correlation, $M$ is replaced by the effective number of experts $M_{\text{eff}} = M/(1 + (M-1)\bar{\rho})$.
\item
  \textbf{Chernoff information}: The optimal exponential rate for the detection problem is given by the Chernoff information $\kappa = \text{KL}(\theta^*\|p_0) = \text{KL}(\theta^*\|p_1)$, applied in Theorem 4$'$.
\item
  \textbf{Bahadur-Rao exact asymptotics}: The Bernoulli lattice correction factor $(1-e^{-\lambda^*})^{-1}$ replaces the standard Chernoff prefactor, tightening the exact constant in Theorem 4$'$ by 5--12\%.
\end{enumerate}

\textbf{References}: - Hoeffding (1963) - Chernoff (1952) - Bahadur \& Rao (1960). ``On deviations of the sample mean'' - Dembo \& Zeitouni (2010). \emph{Large Deviations Techniques and Applications}

\subsubsection{1.6 Information Theory}\label{information-theory}

\textbf{Roots}: - Fano, R. M. (1961). \emph{Transmission of Information: A Statistical Theory of Communications}. MIT Press. - Cover, T. M. \& Thomas, J. A. (2006). \emph{Elements of Information Theory} (2nd ed.). Wiley.

\textbf{Role in Spring}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Fano's inequality}: Used for Theorem 2's lower bound (weak feature failure) --- when the feature $\phi$ and true state $S$ have mutual information $I(\phi; S) \leq \delta$, any state estimation algorithm satisfies:
  \[P(\hat{S} \neq S) \geq \frac{H(S) - \delta - \log 2}{\log K}\]
\item
  \textbf{Data processing inequality (DPI)}: Used in Theorem 2 proof Step 3 --- the Markov chain $\phi(X) \to \hat{S} \to \hat{z}$ implies TV does not increase.
\item
  \textbf{Pinsker's inequality}: Converts the mutual information bound $\delta = I(\phi; S)$ to a total variation bound $\text{TV} \leq \sqrt{\delta/2}$.
\item
  \textbf{KL contraction}: Under correct model specification, posterior convergence $D_{\text{KL}}(P^* \| P_t) \xrightarrow{a.s.} 0$.
\end{enumerate}

\textbf{References}: - Fano (1961) - Cover \& Thomas (2006) - Pinsker (1964). \emph{Information and Information Stability of Random Variables and Processes}

\subsubsection{1.7 Minimax Theory}\label{minimax-theory}

\textbf{Roots}: Le Cam, L. (1973). ``Convergence of estimates under dimensionality restrictions.'' \emph{Annals of Statistics}, 1(1), 38-53.

\textbf{Role in Spring}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Le Cam's two-point method}: Constructs two hard-to-distinguish distributions $P_0 = \text{Bern}(\mu_s)^{\otimes M}$ (clean) and $P_1 = \text{Bern}(1 - C_{\text{bal}}\mu_s/(K-1))^{\otimes M}$ (noisy), using Hellinger tensorization to obtain a minimax lower bound.
\item
  \textbf{Neyman-Pearson reduction}: Via the Neyman-Pearson lemma, any algorithm is reduced to a Bayes test, yielding the optimality of Theorem 4$'$.
\item
  \textbf{Hellinger exact tensorization}: Under A2 (conditional independence), $H^2(P_0^{\otimes M}, P_1^{\otimes M}) = 1 - (1 - H^2(P_0, P_1))^M$ --- exact equality, no inequality relaxation.
\end{enumerate}

\textbf{References}: - Le Cam (1973) - Le Cam (1986). \emph{Asymptotic Methods in Statistical Decision Theory} - Tsybakov (2009). \emph{Introduction to Nonparametric Estimation}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Key Ideas and Proof Tools}\label{key-ideas-and-proof-tools}

\subsubsection{2.1 Two-Timescale Separation ($\beta_t = o(\alpha_t)$)}\label{two-timescale-separation}

Spring's most critical architectural innovation. Condition C6$'$ requires the gatekeeper update rate $\beta_t$ to be a lower-order infinitesimal relative to the student update rate $\alpha_t$:
\[\beta_t = o(\alpha_t), \quad \text{i.e.,} \quad \lim_{t\to\infty} \frac{\beta_t}{\alpha_t} = 0\]

\textbf{Purpose}: Between any two gatekeeper updates, the student executes asymptotically infinitely many gradient steps, converging to a local minimum of the current distribution before distribution shift occurs.

\textbf{Canonical schedule}: $\alpha_t = t^{-a}$, $\beta_t = t^{-b}$, where $\frac{1}{2} < a < 1 < b$. For example $a = 0.6$, $b = 1.2$:
- $\sum \alpha_t = \infty$ (student exploration), $\sum \alpha_t^2 < \infty$ (noise control)
- $\sum \beta_t < \infty$ (finite total distribution shift), $\sum \beta_t^2 < \infty$ (variance control)
- $\beta_t / \alpha_t = t^{-0.6} \to 0$ (timescale separation)

\subsubsection{2.2 Reference Set Replay Breaking Selection Bias}\label{reference-set-replay}

Spring's most profound proof tool. Theorem 12.2 proves: \textbf{Without reference set replay, Lyapunov descent is asymptotically impossible} --- the gatekeeper can reduce apparent loss by only accepting ``easy'' samples, without genuinely improving discriminative ability.

Solution (Theorem 12.5, full proof):
1. \textbf{Student side}: Importance sampling weights $w(x) = P_0(x) / P_{S_t}(x)$, making the effective training distribution the reference distribution $P_0$.
2. \textbf{Gatekeeper side}: Compute SCXUpdate on a fixed reference set $M_0$, eliminating alignment bias.
3. \textbf{Cost}: Importance weight variance inflation factor $W \leq 1/\varepsilon_{\text{explore}} < \infty$.

\subsubsection{2.3 Lipschitz-Almost-Everywhere Lemma}\label{lipschitz-almost-everywhere}

Resolves a fundamental mathematical obstacle (Lemma SE-1.0, B9 fix):

\textbf{Problem}: The consensus score $C(x) = \frac{1}{M}\sum_{m=1}^M \mathbf{1}\{\ell(f_m(x), y) > \tau\}$ is discontinuous at the threshold $\tau$ (indicator function jump).

\textbf{Solution}: Under condition C5 (conditional i.i.d.\ sampling), the discontinuity set has probability measure zero, because:
\[\mathbb{P}(\exists m: \ell(f_m(X), Y) = \tau) \leq \sum_{m=1}^M \mathbb{P}(\ell(f_m(X), Y) = \tau) = 0\]
for smooth functions $\ell(f_m(x), y)$ on continuous distributions. Expectations are unaffected by modifications on zero-measure sets, and the Lyapunov analysis remains valid.

\subsubsection{2.4 Bayesian Posterior Update as Gatekeeper Evolution}\label{bayesian-posterior-update}

The gatekeeper's evolution is interpreted as sequential Bayesian updating:
\[P_{t+1}(S) \propto P_t(S) \cdot \prod_{(x,y) \in \mathcal{D}_{t+1}} S(x,y)^{a} (1 - S(x,y))^{1-a}\]

This yields two powerful theoretical properties:
1. \textbf{Martingale property}: $\mathbb{E}[S_{t+1} \mid \mathcal{F}_t] = S_t$ (tower property), with a.s. convergence guaranteed by Doob's theorem.
2. \textbf{KL contraction}: Under correct model specification, $D_{\text{KL}}(P^* \| P_t) \to 0$.

\subsubsection{2.5 Lyapunov Descent with Importance Sampling}\label{lyapunov-descent-importance-sampling}

The complete proof structure of Theorem 12.5 proves $\mathbb{E}[\Delta\Psi_t \mid \mathcal{F}_t] \leq 0$:

\[\mathbb{E}[\Delta\Psi_t \mid \mathcal{F}_t] \leq -\alpha_t \|\nabla L_0(\theta_t)\|^2 - \beta_t \cdot 2\rho_{\text{ideal}} \cdot \|\Delta_t^{\text{ideal}}\|_{M_0} \cdot \|S_t - \hat{C}\|_{M_0} + O(\alpha_t^2 W + \beta_t^2)\]

Decomposition of four terms:
- \textbf{(A) Student improvement}: $\Delta_{\text{student}}$, eliminating $D_{\text{static}}$ gap via importance sampling (\textbf{proved}).
- \textbf{(B) Gatekeeper improvement}: $\Delta_{\text{gatekeeper}}$, eliminating alignment bias by computing SCXUpdate on $M_0$ (\textbf{proved}).
- \textbf{(C) Distribution shift}: $\Delta_{\text{selection}} = O(\beta_t) = o(\alpha_t)$, controlled by two-timescale (\textbf{proved}).
- \textbf{(D) Cross-coupling}: $\Delta_{\text{cross}} = 0$, due to reference set separability (\textbf{proved}).

\subsubsection{2.6 $O(t^{-a})$ Convergence Rate and Polyak Averaging}\label{convergence-rate-polyak}

Under strong convexity, the student converges at rate $O(t^{-a})$, where $a \in (0.5, 1]$ is the learning rate exponent.

\textbf{Polyak-Ruppert averaging} ($\bar{\theta}_t = \frac{1}{t}\sum_{i=1}^t \theta_i$) improves the rate to $O(t^{-1})$ --- the optimal rate matching the minimax lower bound $\Omega(t^{-1})$ --- applicable for any $a \in (0.5, 1)$ without exact tuning.

\subsubsection{2.7 Characterization of Four Failure Modes}\label{four-failure-modes}

Spring systematically characterizes four canonical failure modes:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mechanism
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Severity
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{1. Premature Freezing} & $\beta_t$ decays too fast before student provides corrective signal & Medium (rare at double precision) \\
\textbf{2. Backlog} & Data ingress rate exceeds gatekeeper scoring throughput & High (real for large datasets) \\
\textbf{3. Client Divergence} & Multiple fixed points + different initial conditions/data streams $\to$ irreconcilable divergence & Medium (multi-client deployments only) \\
\textbf{4. Adversarial Poisoning} & Adversary injects carefully crafted samples; self-evolution loop amplifies contamination & High (severe when amplification present) \\
\end{longtable}

Each mode is equipped with formal conditions, degradation rates, and mitigation strategies.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Historical Development Timeline}\label{historical-development-timeline}

\begin{verbatim}
1892 — Lyapunov Stability Theory
      A. M. Lyapunov, The General Problem of the Stability of Motion
      Foundation of qualitative analysis of dynamical systems — fixed points, attractors, Lyapunov functions
      ↓ Spring use: Ψ(S_t, θ_t) as Lyapunov function; fixed point analysis; attractor basins
      
1933 — Kolmogorov's Strong Law of Large Numbers
      A. N. Kolmogorov, Grundbegriffe der Wahrscheinlichkeitsrechnung
      Foundation of almost-sure convergence for i.i.d. sequences
      ↓ Spring use: Large-sample behavior of consensus statistic C(x)

1951 — Robbins-Monro Stochastic Approximation
      H. Robbins & S. Monro, "A stochastic approximation method"
      Mathematical foundation of stochastic gradient descent; Σα_t=∞, Σα_t²<∞ conditions
      ↓ Spring use: θ_t update follows RM recursion; two-timescale extension

1952 — Chernoff Information
      H. Chernoff, "A measure of asymptotic efficiency for tests"
      Optimal exponential rate for hypothesis testing
      ↓ Spring use: Exact exponential rate for Theorem 4' detection

1953 — Doob's Martingale Convergence
      J. L. Doob, Stochastic Processes
      Almost-sure convergence of bounded martingales
      ↓ Spring use: S_t a.s. convergence guarantee

1959 — Davidon-Fletcher-Powell (DFP)
      W. C. Davidon, "Variable metric method for minimization"
      First quasi-Newton method — preconditioning without explicit Hessian
      ↓ Spring use: Conceptual precursor to adaptive learning rates in SGD

1960 — Bahadur-Rao Exact Asymptotics
      R. R. Bahadur & R. R. Rao, "On deviations of the sample mean"
      Lattice correction for exact tail asymptotics of Bernoulli sums
      ↓ Spring use: Exact constant tightening in Theorem 4'

1961 — Fano's Inequality
      R. M. Fano, Transmission of Information
      Lower bound on error probability from conditional entropy
      ↓ Spring use: Theorem 2 lower bound for weak feature failure

1963 — Hoeffding's Inequality
      W. Hoeffding, "Probability inequalities for sums of bounded random variables"
      Finite-sample concentration for bounded independent variables
      ↓ Spring use: Theorem 1 exponential convergence guarantee

1964 — Pinsker's Inequality
      M. S. Pinsker, Information and Information Stability
      KL divergence upper bounds total variation distance
      ↓ Spring use: Bridge from information-theoretic δ to detection margin bound

1973 — Le Cam's Minimax Theory
      L. Le Cam, "Convergence of estimates under dimensionality restrictions"
      Two-point method for minimax lower bounds
      ↓ Spring use: Theorem 4' optimality lower bound

1978 — Conley Index Theory
      C. Conley, Isolated Invariant Sets and the Morse Index
      Topological characterization of dynamical system attractors
      ↓ Spring use: Conceptual framework for ω-limit set analysis

1992 — Polyak-Ruppert Averaging
      B. T. Polyak & A. B. Juditsky, "Acceleration of stochastic approximation by averaging"
      Simple averaging achieves optimal O(1/t) rate
      ↓ Spring use: Student optimal convergence without exact tuning

2003 — Online Convex Optimization (Zinkevich)
      M. Zinkevich, "Online convex programming and generalized infinitesimal gradient ascent"
      First general regret bound O(√T) for online convex optimization
      ↓ Spring use: Gatekeeper online learning regret analysis

2008 — Two-Timescale Stochastic Approximation (Borkar)
      V. S. Borkar, Stochastic Approximation: A Dynamical Systems Viewpoint
      Rigorous framework for coupled fast-slow stochastic recursions
      ↓ Spring use: β_t = o(α_t) condition; ODE method for coupled system

2016 — Delayed Online Learning (Joulani et al.)
      P. Joulani, A. György, Cs. Szepesvári, "Online learning under delayed feedback"
      Regret bounds under arbitrary bounded delay
      ↓ Spring use: NEP student delayed feedback regret analysis
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Spring vs Prior Work Comparison Table}\label{spring-vs-prior-work}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
System
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Self-Evolution?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Convergence Proof?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Data Quality?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Key Difference from Spring
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{AlphaZero} & Yes (self-play) & Yes (policy improvement) & No (trusts game rules) & AlphaZero generates synthetic data; Spring audits real noisy data \\
\textbf{Bayesian Optimization} & No & Yes (regret bounds) & No & BO optimizes static black-box; Spring's quality distribution evolves \\
\textbf{Active Learning} & Partially & Yes (query complexity) & No (oracle is perfect) & AL's oracle is instantaneous; Spring's NEP has delayed feedback \\
\textbf{MAML / Meta-Learning} & Partially & Yes (gradient-based) & No & MAML learns over task distributions; Spring evolves over data quality \\
\textbf{Solomonoff Induction} & Yes (in principle) & No (uncomputable) & Partially (universal) & Solomonoff is uncomputable; Spring is algorithmic and implementable \\
\textbf{Confident Learning} & No & Yes (threshold bounds) & Yes (label noise) & Confident Learning is static (one-shot); Spring is dynamic (iterative) \\
\textbf{Data Shapley} & No & Partial (axiomatic) & Yes (data valuation) & Shapley values need model retraining; Spring uses NEP as proxy \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Spring's Unique Contributions}\label{springs-unique-contributions}

\subsubsection{5.1 Conceptual}\label{conceptual}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item \textbf{Closed-loop self-evolution for data quality}: First framework where data quality assessment itself improves with use --- each audit cycle enriches the memory bank, making the next audit more accurate.
\item \textbf{Resurrection mechanism}: Unlike all prior systems that discard ``bad'' data, Spring stores everything and periodically re-evaluates. Data that was below threshold at $t=5$ may bloom at $t=50$ when the gatekeeper matures.
\item \textbf{Physical anchor}: The DFT/experimental ground truth prevents the self-evolution loop from drifting into a self-consistent but physically wrong state.
\end{enumerate}

\subsubsection{5.2 Technical}\label{technical}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item \textbf{Coupled two-timescale convergence proof}: First proof that a coupled (gatekeeper, student) system converges under realistic conditions (reference set replay, importance sampling, Lipschitz-almost-everywhere).
\item \textbf{Selection bias elimination}: Theorem 12.2 proves that without reference set replay, Lyapunov descent is impossible --- the gatekeeper will cheat. Theorem 12.5 proves that the proposed mechanism eliminates this cheating.
\item \textbf{Four failure mode characterization}: Unlike prior work that only addresses convergence, Spring systematically characterizes what happens when convergence fails.
\item \textbf{Cross-domain applicability}: The same mathematical structure (judge $\to$ store $\to$ update $\to$ re-judge) applies to MLIP training, medical imaging, LLM quality assessment, and autonomous driving verification.
\end{enumerate}

\subsubsection{5.3 Philosophical}\label{philosophical}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item \textbf{G\"{o}del analogy}: Just as G\"{o}del proved that sufficiently rich formal systems cannot prove their own consistency, Theorem SE-2 bounds what Spring can know about its own data quality. There is always a residue of uncertainty that only external physical reality can resolve.
\item \textbf{Winter-spring metaphor}: The naming ``Spring'' is not decorative. It captures the mathematical fact that data quality is not a static property but a dynamic process --- what is noise today may be signal tomorrow, if and only if the gatekeeper is allowed to improve.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{End of MATHEMATICAL\_GENEALOGY.tex --- Spring Self-Evolution Algorithm: Mathematical Genealogy}

\end{document}
