% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX 8-Theorem Papers Mathematical Completeness Review Report}\label{scx-8-theorem-papers-mathematical-completeness-review-report}

\subsection{0. SCX Axiom System Baseline}\label{scx-axiom-system-baseline}

Clarify the core axioms before review:

\begin{itemize}
\tightlist
\item
  \textbf{Theorem 1} (S1): Multi-expert consensus noise detection --- Under A1-A6 assumptions, F1 ≥ 1 - (1/η)Σρ\_s·exp(-2MΔ\_s²), exponential convergence.
\item
  \textbf{Theorem 2} (S2): Weak feature information-theoretic bound --- F1\_SCX ≤ F1\_base + C\_F·√(δ/2), δ=I(φ(X);S).
\item
  \textbf{Theorem 3} (S3): Noise-difficulty unidentifiability --- From observed data, label noise and intrinsic difficulty cannot be distinguished. This is SCX's ``uncertainty principle''.
\item
  \textbf{Theorem 4} (S4): Exact constant minimax optimality --- Bahadur-Rao large deviation exact constants.
\item
  \textbf{Cercis Score}: S(x) = Q(x) + η·N(x), quality+novelty decomposition.
\item
  \textbf{Situs Operator}: Encodes distributions as metric measure spaces (X, d\_P, μ\_P).
\item
  \textbf{Spring SE-1}: Self-evolving gating, Robbins-Monro convergence guarantee.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Per-Paper Review}\label{per-paper-review}

\subsubsection{1.1 Theorem 5 --- Active Learning (Cercis Optimal Sampling)}\label{theorem-5-active-learning-cercis-optimal-sampling}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - Theorem 5.1 (Optimal Distribution): Complete. Gibbs form p*(x) ∝ π(x)exp(αS(x)/λ), with uniqueness proof. - Theorem 5.2 (Optimal Rate): Statement complete, closed form ρ* = η/(1+η)·Σw\_k·V{[}Γ\_k{]}/Σw\_k·Γ̄\_k. But key derivation steps have \textbf{serious logical gaps} (see below). - Theorem 5.3 (Consistency): Complete. √n-consistency, Delta method.

\textbf{Proof Framework Logical Gaps}: ★★★☆☆ (3/5) --- \textbf{Serious issues exist} - Lemma 1 (Gibbs solution) derivation is standard variational inference, no problem. - \textbf{Fatal gap in Theorem 5.2 proof lines 414-421}: Jumping from ρ* = η·Σw\_kΓ̄\_k/(η·Σw\_kΓ̄\_k + Σw\_kV{[}Γ\_k{]}) to the final form ρ* = η/(1+η)·Σw\_kV{[}Γ\_k{]}/Σw\_kΓ̄\_k, the claimed identity Σw\_kΓ̄\_k = ((1+η)/η)·Σw\_kV{[}Γ\_k{]} is used. This identity is labeled ``holds at the optimum of the variational objective, from the equality of marginal returns and the marginal cost of η-scaled novelty penalty.'' \textbf{This is circular reasoning} — using ρ*'s definition to derive ρ*'s closed form. Algebraically: - From Φ'(ρ)=0 we get ρ = Σw\_kΓ̄\_k / (Σw\_kΓ̄\_k + η\^{}\{-1\}Σw\_kV{[}Γ\_k{]}) - This form is already a closed-form solution, no additional identity is needed - The paper's forced conversion to the η/(1+η) form introduces a hypothetical structural relation that breaks closure - \textbf{Assumption 1 (Cercis-Gain Affinity)}: Γ̄(x) = α·S(x) + β. This is the cornerstone of the entire derivation but is only justified by ``intuition'' and ``Taylor expansion'' — lacking rigorous proof. This is an unverified empirical assumption.

\textbf{Consistency with SCX Axiom System}: ★★★★☆ (4/5) - Correctly uses Cercis Score S=Q+ηN definition - η→0 recovers pure uncertainty sampling, η→∞ recovers pure diversity sampling, consistent with SCX framework - But the η/(1+η) saturation factor does not appear in the original SCX framework — this is a newly introduced structure

\textbf{Rigorously Derivable from Existing Axioms}: - Theorem 5.1 (Gibbs distribution): ✓ Rigorously derived from variational principle (no SCX axiom needed) - Theorem 5.3 (Consistency): ✓ Rigorously derived from law of large numbers + Delta method

\textbf{Requires New Assumptions}: - Theorem 5.2 (Sampling rate): ✗ Requires Assumption 1 (Cercis-Gain Affinity) + additional variational structural relation - The risk-regulation function Φ(ρ) with its three-term decomposition (expected return, marginal diminishing, novelty cost × variance) is an \textbf{ad hoc construction}, not derived from Cercis axioms

\textbf{Relationship to Other Theorems}: Independent. Cited by AR-Theorem's detectability result (``T5 optimal active learning strategy''), but AR only uses T5 as a black-box tool.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.2 Theorem 6 --- Protocol Game (Audit Protocol Adoption)}\label{theorem-6-protocol-game-audit-protocol-adoption}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - Theorem 6.1 (Regret bound): Complete. - Theorem 6.2 (Nash existence): Complete, standard Glicksberg fixed-point argument. - Theorem 6.3 (Uniqueness): Complete, explicit γ\_crit closed form. - Proposition 6.4 (Distortion): Complete.

\textbf{Proof Framework Logical Gaps}: ★★★★☆ (4/5) - Existence proof is standard — Glicksberg's theorem directly applies to mixed-strategy games with continuous payoffs. - Uniqueness proof's contraction mapping argument is correct, but one \textbf{gap} exists: the sensitivity of influence function φ\_P to early adopter count is bounded as √(2logN₁/(N₀+1)) + σ\_ξ/√N₀, assuming the derivative of the UCB confidence radius can be approximated by finite differences. This approximation holds asymptotically for large N₀ but requires additional justification in finite samples. - Lemma ``Influence Monotonicity'' relies only on intuitive reasoning (``increasing n\_P reduces standard error''), lacking formal treatment of UCB index randomness.

\textbf{Consistency with SCX Axiom System}: ★★★☆☆ (3/5) - The paper claims ``Quality(Q) corresponds to protocol quality μ\_P'' and ``Novelty(N) corresponds to UCB exploration reward.'' This mapping is \textbf{analogical rather than derivational} — the protocol adoption game is not a natural extension of the Cercis framework but an independently constructed game-theoretic model, then post-hoc analogized with SCX. - γ\_crit includes the UCB exploration term √(2logN₁/(N₀+1)), which has no direct mathematical connection to Cercis's η parameter, only conceptual parallelism.

\textbf{Rigorously Derivable from Existing Axioms}: - Theorem 6.1 (Regret): ✓ Standard UCB analysis - Theorem 6.2 (Existence): ✓ Standard game theory

\textbf{Requires New Assumptions}: - Theorem 6.3 (Uniqueness): Requires Gumbel noise assumption + influence sensitivity bound (UCB finite-difference approximation) - The entire model setup (sequential game, Gumbel preference shocks, Gaussian prior) is independent of SCX axioms

\textbf{Relationship to Other Theorems}: Indirectly connected to CD-Theorem (information externalities) and parallel to TS-Theorem's anchor concept. But \textbf{essentially an independent paper}, piggybacking on the SCX brand rather than derived from it.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.3 Theorem 7 --- Cross-Domain Partition Preservation}\label{theorem-7-cross-domain-partition-preservation}

\textbf{Theorem Statement Completeness}: ★★★★★ (5/5) - Theorem 7.1 (Main bound): Complete, three components clearly stated. - Theorem 7.2 (Tightness): Complete, constructed a tightness sequence. - Corollary 7.3 (Optimal granularity): Complete.

\textbf{Proof Framework Logical Gaps}: ★★★★☆ (4/5) - Five-step proof structure is sound. - \textbf{Step 3 has a potential issue}: The claim ``conditional densities are L\_k-Lipschitz ⇒ the entropy functional is L\_k-Lipschitz with respect to the Wasserstein metric.'' Entropy is not generally a Lipschitz function; additional conditions are needed (e.g., logarithmic Sobolev inequality or bounded conditional densities on bounded support). Lemma S2.2 (Entropy Lipschitz Property) provides an argument that H(Y|X=x) is (L/σ)-Lipschitz, but this assumes bounded conditional variance and approximately Gaussian conditional distributions — a \textbf{significant but unstated assumption}. - The ``conditioning does not increase average transport cost'' argument (Step 4) is a correct standard result.

\textbf{Consistency with SCX Axiom System}: ★★★★★ (5/5) - Directly departs from Situs conditional mutual information I(Y;P|S) - Uses Situs encoding and Wasserstein distance, fully within the SCX framework - Explicitly cites T7 cross-domain transfer theorem

\textbf{Rigorously Derivable from Existing Axioms}: - Main bound: Partially derivable (requires Assumption 1 + entropy Lipschitz condition) - Asymptotic tightness: ✓ Constructive proof

\textbf{Requires New Assumptions}: - Assumption 1 (Bounded Conditional Densities with Lipschitz property) - Lemma S2.2's entropy Lipschitz argument requires Gaussianity or equivalent conditions

\textbf{Relationship to Other Theorems}: Cited by CD-Theorem's causal relay theorem as ``T7 cross-domain fidelity.'' One of the \textbf{mathematically most self-consistent} papers among the 8.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.4 CD-Theorem --- Causal Discovery (Noise vs Confounding)}\label{cd-theorem-causal-discovery-noise-vs-confounding}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - CD-Theorem 1 (Strong Inseparability): Complete. - CD-Theorem 2 (Weak Separability): Statement complete but \textbf{extremely conditional}. - CD-Theorem 3 (Causal Relay): Complete.

\textbf{Proof Framework Logical Gaps}: ★★★☆☆ (3/5) - \textbf{Strong Inseparability proof gap}: Step 1 claims S(x) = S₀(x) + η·δ\_ε(x) + β·δ\_z(x), then asserts that when η⊥z, the joint distribution factorization makes δ\_z indistinguishable from effective noise. This is a direct analogical generalization of Theorem 3, but the \textbf{argument is overly sketchy} — it needs to prove that the joint distribution of (S, ∇S, H\_S) is indeed identical in both worlds, not merely the marginal distribution of S. Theorem 3's original proof explicitly constructs equivalent joint distributions for two worlds; CD-Theorem does not provide such an explicit construction. - \textbf{Problem with the Curl statistic}: τ = (1/|X|)∫‖∇×∇̂S‖²dx. This statistic computationally depends heavily on the definition of curl on Situs manifolds. The paper does not give a concrete construction of the curl operator on Situs manifolds, only vaguely referencing ``Situs manifolds in the SCX framework.'' Geometrically, curl requires a connection structure, while Situs encoding only provides a metric — \textbf{from metric to curl requires additional mathematical structure} (Levi-Civita connection may not exist on general metric spaces). - \textbf{Causal Relay}: Chain derivation is reasonable, but the interaction term C·ε·δ in g(ε,δ) gives no definition or bounding method for C.

\textbf{Consistency with SCX Axiom System}: ★★★★☆ (4/5) - Directly extends Theorem 3 to causal discovery settings - Situs skeleton preservation assumption is a major extension of the Situs operator but is honestly labeled as an open problem - Causal Relay correctly cites T7

\textbf{Rigorously Derivable from Existing Axioms}: - Strong Inseparability: Claimed as a direct corollary of Theorem 3, but proof is overly sketchy

\textbf{Requires New Assumptions}: - Situs skeleton preservation assumption (core open problem) - Existence of curl operator on Situs manifolds - All three theorems have honest annotations ({[}Rigorous{]}/{[}Conditionally Rigorous{]}/{[}Open{]})

\textbf{Relationship to Other Theorems}: Forms a triangular cross-reference with AR-Theorem (gradient anomalies masquerading as causal signals) and HC-Theorem (human cognition breaking symmetry).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.5 FA-Theorem --- Federated Audit (ZK Multi-Party)}\label{fa-theorem-federated-audit-zk-multi-party}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - FA-Theorem 1 (ε-Equivalence): Complete, four error sources. - FA-Theorem 2 (Info Lower Bound): Complete. - ZK construction sketch: Complete but only outline level.

\textbf{Proof Framework Logical Gaps}: ★★★☆☆ (3/5) - \textbf{Core unsolved problem}: Existence of the Situs homomorphic composition operator ⊕. Without it, the ε-Equivalence theorem has only conditional meaning. The paper honestly acknowledges this as an open problem. - \textbf{I-MMSE gap derivation} (Step 2 of Lower Bound): Uses the Guo-Shamai-Verdú I-MMSE relation, but that relation applies to Gaussian channels, while Cercis Score here is non-Gaussian. The argument requires generalization or additional defense. - \textbf{``Worst-case construction''} (Step 3): Claims ``the adversary can adjust S̃\_k to maximize bias'' but does not give a concrete form for the adjustment strategy. It must be shown that a tampering exists such that the deviation of E{[}T\_fed{]} from E{[}T\_full{]} achieves the claimed lower bound.

\textbf{Consistency with SCX Axiom System}: ★★★★☆ (4/5) - Correctly uses Cercis Score and Situs encoding - Audit Sword principle extends the SCX framework - Lower bound directly derived from Theorem 3

\textbf{Rigorously Derivable from Existing Axioms}: - Info Lower Bound (Theorem 2): ✓ Derived from Theorem 3 + data processing inequality

\textbf{Requires New Assumptions}: - Situs homomorphic composition operator ⊕ (core open problem) - Feasibility of ZK circuit construction (cryptographic engineering, not mathematical)

\textbf{Relationship to Other Theorems}: Cited by CD-Theorem and AR-Theorem as the federated extension of the ``Audit Sword'' principle. \textbf{The most ambitious extension paper}, but also the most dependent on unsolved mathematical problems (Situs homomorphism).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.6 AR-Theorem --- Adversarial Robustness (Cercis Taxonomy)}\label{ar-theorem-adversarial-robustness-cercis-taxonomy}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - AR-Theorem 1 (Reduction): Complete, conditions clear. - AR-Theorem 2 (Detectability): Complete, sample complexity lower bound. - Conjecture (Phase Transition): Statement complete but only a conjecture.

\textbf{Proof Framework Logical Gaps}: ★★★☆☆ (3/5) - \textbf{Key gap in Reduction theorem}: The proof claims ||∇S||/S ≤ η (``from Cercis operator definition'') yields A(x) ≤ η·ε. This inequality \textbf{is not a standard result in the SCX framework} — Cercis operator defines S=Q+ηN and does not directly constrain ||∇S||/S. This requires the additional assumption that ∇Q and ∇N are bounded in proportion to Q and N. - \textbf{Hessian bound in Taylor expansion}: ||H\_S||\_op ≤ L(x) is correct (Lipschitz gradient ⇒ Hessian bounded), but only holds when using the spectral norm and the function is twice differentiable. - \textbf{Detectability theorem lower bound}: Uses Le Cam two-point method to derive χ² divergence = Θ(γ²), a derivation that is \textbf{extremely sketchy} — requires explicitly computing the χ² divergence between gradient field distributions under H₀ and H₁, which depends on concrete assumptions about the distribution of A(x). The paper does not give this computation. - \textbf{Upper bound (T5 achievability)}: Claims Theorem 5 provides a matching upper bound, but T5's optimality concerns maximizing I(S;query), not rate-optimality for detecting adversarial structure. This is \textbf{different senses of optimality} — T5 optimizes sample efficiency for obtaining labels, while AR requires sample efficiency for detecting anomalous gradient structure.

\textbf{Consistency with SCX Axiom System}: ★★★☆☆ (3/5) - The ternary decomposition S=Q+ηN+γA is a major extension of the original Cercis binary decomposition - Whether introducing a third primitive (A) into the SCX framework is an \textbf{axiom-level question} - The gradient field diagnostic D\_adv construction is novel but lacks clear definition in the SCX framework

\textbf{Rigorously Derivable from Existing Axioms}: - Reduction's Lipschitz sufficient condition: Derivable, but uses an inequality (||∇S||/S ≤ η) not in SCX axioms

\textbf{Requires New Assumptions}: - Reasonableness of ternary decomposition S=Q+ηN+γA - ||∇S||/S ≤ η (Cercis gradient-to-value ratio bounded) - Incomplete correlation between adversarial exposure A and novelty N (I(A;N|S)>0) - Curl operator on Situs manifolds (shares this problem with CD-Theorem)

\textbf{Relationship to Other Theorems}: Explicitly cites CD (adversarial signals confused with causal signals), HC (humans detect adversarial examples), TS (S\_crit drift). But among the 8 papers, it has the \textbf{weakest mathematical foundation} — two theorems heavily depend on unverified assumptions, and the core conjecture is entirely open.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.7 TS-Theorem --- Temporal SCX (Drift \& Self-Audit)}\label{ts-theorem-temporal-scx-drift-self-audit}

\textbf{Theorem Statement Completeness}: ★★★★☆ (4/5) - TS-Theorem 1 (Drift Detectability): Complete, sample size formula. - TS-Theorem 2 (Drift Attribution): Complete, with anchorless/anchored cases. - TS-Theorem 3 (Sprint Self-Audit Limit): Complete, Lyapunov condition.

\textbf{Proof Framework Logical Gaps}: ★★★★☆ (4/5) - \textbf{Drift Detectability proof is reasonable}, but has an implicit assumption: the Lipschitz relation |η\_{t+1} - η\_t| ≤ κ·d\_Situs + |ξ\_t| between Situs distance and η drift. This relation is reasonably derived from T7's Lipschitz property. - \textbf{The variance formula (Equation 14)} includes a σ²\_est term but does not define or derive it. Under H₀, the estimation variance of η should be derivable from the Cercis Score distribution, but the paper directly asserts a form. - \textbf{Attribution unidentifiability construction} (Config A vs Config B) is convincing but only covers two configurations; it does not prove \textbf{all} possible decompositions are unidentifiable. A more general unidentifiability proof is needed. - \textbf{Lyapunov convergence theorem} is standard Foster-Lyapunov argument, no problem. The divergence condition derivation is also natural.

\textbf{Consistency with SCX Axiom System}: ★★★★☆ (4/5) - Explicitly temporalizes T7 cross-domain transfer theorem (Situs distance becomes inter-time-slice Situs distance) - Sprint self-audit is a natural temporal extension of the SCX framework - Anchor concept parallels HC-Theorem's anchor knowledge

\textbf{Rigorously Derivable from Existing Axioms}: - Drift Detectability: ✓ Derived from T7 + standard hypothesis testing - Attribution unidentifiability: ✓ Via counterexample construction - Anchored estimation: ✓ Directly derived from anchor definitions

\textbf{Requires New Assumptions}: - Lipschitz constant κ from Situs distance to η - Existence of anchor sets (practical but not theoretical requirement) - Contraction rate ρ in Lyapunov condition (needs derivation from Sprint update rule)

\textbf{Relationship to Other Theorems}: Shares anchor concept with T7 and HC. Convergence/divergence analysis parallels Spring SE-1's Robbins-Monro convergence but is different — Spring is model evolution, TS is distribution/audit parameter evolution.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.8 HC-Theorem --- Human-AI Collaborative Audit}\label{hc-theorem-human-ai-collaborative-audit}

\textbf{Theorem Statement Completeness}: ★★★★★ (5/5) - HC-Theorem 1 (Collaborative Gain): Complete. - HC-Theorem 2 (Budget Decay): Complete, √(B\_H/n) scaling law. - HC-Theorem 3 (Absolute Boundary): Complete.

\textbf{Proof Framework Logical Gaps}: ★★★★☆ (4/5) - \textbf{Collaborative Gain logic}: ``If human judgments carry super-SCX information, then collaborative auditing strictly outperforms pure AI.'' This is a conditional implication, logically correct, but the \textbf{condition itself (I(J\_H;ε|S)>0) is an empirical hypothesis}. The paper honestly acknowledges this as an open problem. - \textbf{Budget Decay derivation has a subtle issue}: Step 1 claims Ω\_collab = Ω\_AI + (B\_H/n)·Δ\_I·η\_eff, but Step 2 argues via Cramér-Rao that η\_eff ≤ c/√(B\_H/n)·1/Δ\_I, giving the √(B\_H/n) scaling. This compound derivation requires proving that optimal selection based on Cercis Score indeed produces the √(B\_H/n) efficiency loss. The paper's argument is overly sketchy — ``optimal selection introduces dependence among samples, reducing effective sample size from B\_H to √(B\_H·n)'' needs more rigorous treatment. - \textbf{Absolute Boundary theorem}: This is a direct logical extension of Theorem 3; the argument is robust. The concept of ``perfect noise'' is clear: I(ε; all observables|X)=0 ⇒ absolutely indistinguishable.

\textbf{Consistency with SCX Axiom System}: ★★★★★ (5/5) - Directly extends Theorem 3 to arbitrary cognitive systems - The ``multi-lens'' perspective is a philosophical deepening of the SCX framework - Establishes a continuum from Theorem 3 (AI-lens) to arbitrary lenses

\textbf{Rigorously Derivable from Existing Axioms}: - Absolute Boundary (Theorem 3): ✓ Direct corollary of Theorem 3 - Budget Decay: Derivable from Theorem 3 + information theory, but the √(B\_H/n) scaling rigor needs more work

\textbf{Requires New Assumptions}: - Information asymmetry condition I(J\_H;ε|S)>0 (empirical hypothesis) - Calibration condition (human uncertainty monotonic in Cercis Score) - Efficiency factor η\_eff upper bound derivation needs stronger regularity conditions

\textbf{Relationship to Other Theorems}: Connected to CD (intersection of human cognition and SCX signals), AR (humans detecting adversarial examples), TS (anchors). Among the 8 papers, it has the \textbf{deepest philosophy and most thorough understanding of Theorem 3}.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Cross-Paper Analysis}\label{cross-paper-analysis}

\subsubsection{2.1 Theorem Duplication or Contradiction}\label{theorem-duplication-or-contradiction}

\begin{itemize}
\tightlist
\item
  \textbf{Duplication}: No direct duplication. CD-Theorem's Strong Inseparability overlaps in spirit with Theorem 3 but addresses different problems (noise vs difficulty vs noise vs confounding).
\item
  \textbf{Potential contradictions}:

  \begin{itemize}
  \tightlist
  \item
    AR-Theorem's Detectability claims to break Theorem 3 at the ∇S level (``Theorem 3's barrier applies at the S level, not at the ∇S level''), while Theorem 3 originally claims to cover ``the Cercis Score and its derivatives.'' If Theorem 3's original proof indeed covers ∇S, then AR's claim contradicts Theorem 3. \textbf{This requires checking whether Theorem 3's proof genuinely covers gradient information.} Theorem 3's original text says ``observables accessible to an SCX auditor: the Cercis Score, its gradient, its Hessian,'' with the conclusion ``No algorithm operating on these observables can tell the worlds apart.'' AR-Theorem claims ∇S contains higher-order structure (curl) that Theorem 3 did not exhaust. \textbf{This is a potential internal contradiction — either Theorem 3's claim is weakened by AR, or AR's claim is invalid.}
  \item
    T5's η/(1+η) factor and the original SCX η parameter semantics are not fully consistent: in the original framework, η is a continuous exploration-exploitation tradeoff, while in T5 η affects the sampling rate through the η/(1+η) saturation factor. Should these two η's be considered the same parameter?
  \end{itemize}
\end{itemize}

\subsubsection{\texorpdfstring{2.2 Weakest Paper: \textbf{AR-Theorem (Adversarial Robustness)}}{2.2 Weakest Paper: AR-Theorem (Adversarial Robustness)}}\label{weakest-paper-ar-theorem-adversarial-robustness}

\begin{itemize}
\tightlist
\item
  \textbf{Reasons}:

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    One of the core results (Reduction) relies on the inequality ||∇S||/S ≤ η, which is not in SCX axioms
  \item
    Detectability's χ² divergence computation is not explicitly given
  \item
    The core problem (Phase Transition Conjecture) is entirely an open conjecture
  \item
    The ternary decomposition S=Q+ηN+γA is a major modification of SCX axioms, but the paper does not justify the legitimacy of this modification
  \item
    Potential contradiction with Theorem 3
  \item
    Le Cam lower bound derivation is overly sketchy and may contain hidden assumptions
  \end{enumerate}
\item
  \textbf{Honest critique}: This paper has a good idea (adversarial examples may constitute a third category), but the mathematical execution is insufficient. The conditions within the two ``theorems'' are excessively strong; the core conclusions are mostly ``if... then...'' conditional statements where the conditions themselves are neither satisfied nor verified. It is essentially a \textbf{position paper + research program}, not a settled theorem.
\end{itemize}

\subsubsection{\texorpdfstring{2.3 Strongest Paper: \textbf{Theorem 7 (Cross-Domain Partition Preservation)}}{2.3 Strongest Paper: Theorem 7 (Cross-Domain Partition Preservation)}}\label{strongest-paper-theorem-7-cross-domain-partition-preservation}

\begin{itemize}
\tightlist
\item
  \textbf{Reasons}:

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Most complete theorem statement; every symbol has a clear definition
  \item
    Five-step proof decomposition is clear, with well-developed lemmas
  \item
    Constructed asymptotic tightness, the highest standard of theoretical proof
  \item
    Highest consistency with the SCX framework — directly departs from Situs conditional mutual information
  \item
    Component decomposition (geometric × statistical × irreducible) provides actionable practical guidance
  \item
    Extensions (p-Wasserstein, multi-target domain, adaptive refinement) demonstrate the richness of the theory
  \item
    \textbf{The only significant weakness} (entropy Lipschitz requiring Gaussian assumption) is discussed in the appendix
  \end{enumerate}
\item
  \textbf{Second-strongest candidate}: HC-Theorem — Its philosophical deepening of Theorem 3 is outstanding, and the three-layer structure (AI-only → human-augmented → absolute limit) is logically perfect. Budget Decay's √(B\_H/n) scaling is a beautiful information-theoretic result. The only shortcoming is that the core assumption depends on empirical validation.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Global Assessment Matrix}\label{global-assessment-matrix}

\begin{longtable}[]{@{}llllll@{}}
\toprule\noalign{}
Paper & Completeness & Logical Rigor & SCX Consistency & New Assumption Dependence & Overall Score \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
T5-Active & ★★★★☆ & ★★★☆☆ & ★★★★☆ & High & \textbf{B+} \\
T6-Protocol & ★★★★☆ & ★★★★☆ & ★★★☆☆ & High & \textbf{B} \\
T7-CrossDomain & ★★★★★ & ★★★★☆ & ★★★★★ & Medium & \textbf{A} \\
CD-Causal & ★★★★☆ & ★★★☆☆ & ★★★★☆ & High & \textbf{B-} \\
FA-Federated & ★★★★☆ & ★★★☆☆ & ★★★★☆ & Very High & \textbf{B-} \\
AR-Adversarial & ★★★★☆ & ★★☆☆☆ & ★★★☆☆ & Very High & \textbf{C+} \\
TS-Temporal & ★★★★☆ & ★★★★☆ & ★★★★☆ & Medium & \textbf{B+} \\
HC-Human & ★★★★★ & ★★★★☆ & ★★★★★ & Medium & \textbf{A-} \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Summary of Key Findings}\label{summary-of-key-findings}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{T5's Theorem 5.2 proof has an algebraic/logical gap} — the final step of the closed form introduces circular reasoning. But this is fixable (directly use ρ = ΣΓ̄/(ΣΓ̄+η⁻¹ΣV) as the closed form without forcing conversion to the η/(1+η) form).
\item
  \textbf{AR and Theorem 3 have a potential contradiction} — AR claims the gradient field contains information not covered by Theorem 3, but Theorem 3 explicitly claims to cover gradients. Needs CLARIFICATION.
\item
  \textbf{The Situs homomorphic composition operator (⊕) is the core bottleneck for FA-Theorem and the blocking point for the entire federated SCX program}. Without ⊕, FA-Theorem degrades to a conditional statement.
\item
  \textbf{CD-Theorem's curl statistic depends on the connection structure on Situs manifolds}, which does not naturally exist in pure metric spaces. This is a geometric problem shared by CD and AR.
\item
  \textbf{Among the 8 papers, 4 have core results that are in some sense ``conditional theorems''} (if X holds, then Y holds), where X is often an open problem or empirical hypothesis. This is not a bad thing — it correctly positions the frontier of SCX research — but readers need to know which parts are settled and which are open.
\item
  \textbf{The strongest mathematics is in T7} (information-theoretic bounds + tightness construction), \textbf{the deepest philosophy is in HC} (multi-lens theory + absolute boundary of perfect noise), \textbf{the most practical is in T5} (closed-form sampling rate), \textbf{the most ambitious are in FA and AR} (but with the weakest mathematical foundation).
\end{enumerate}

\end{document}
