% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX Framework: The Next Generation of Theorems (The Next Theorems)}\label{scx-framework-next-theorems}

\subsection{Theory Architect Draft | June 29, 2026}\label{theory-architect-draft}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\textbf{Prerequisite assumption}: T1-T7 all hold. T1-T4 (static audit completeness), T5 (active learning bounds), T6 (protocol game equilibrium), T7 (cross-domain transfer fidelity), together with Spring SE-1 (dynamic evolution) and Situs (spatial encoding) constitute the complete foundational theoretical system.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Causal Discovery Theorem: Separability of Label Noise vs. Causal Confounding}\label{causal-discovery-theorem}

\subsubsection{(Causal Discovery Theorem --- CD-Theorem)}\label{cd-theorem-next}

\subsubsection{Context}\label{context-cd}

Theorem 3 (Uncertainty Principle) asserts: within the SCX audit framework, \textbf{noise and intrinsic difficulty are indistinguishable}. Formally, given Cercis Score S = Q + $\eta$N, the auditor cannot decompose quality component Q and noise component N from observations of S alone---there exists an irreducible information barrier.

But Theorem 3 leaves a critical gap unclosed: \textbf{Are label noise (measurement noise) and unobserved confounders also indistinguishable?} This is a deeper question. Label noise is random perturbation of the supervisory signal; causal confounding is spurious correlation due to unobserved common causes. The two may exhibit identical statistical characteristics in observed distributions, but their intervention implications are completely different---the former can be mitigated through data cleaning, the latter must be resolved through causal identification (do-calculus, instrumental variables, etc.).

\subsubsection{Theorem Statement (Draft)}\label{cd-statement}

\begin{quote}
\textbf{CD-Theorem (Causal Separability Theorem)}: Let D = {($x_i$, $y_i$)} be the audit dataset, $y_i$ = f*($x_i$) + $\varepsilon_i$ + $\beta\cdot z_i$, where $\varepsilon_i$ is independent label noise, $z_i$ is an unobserved confounder, and $\beta$ is the confounding strength vector. Define the auditor's accessible information set as I\_A = {S(x), $\nabla$S(x), H\_S(x)} (Cercis Score and its first and second-order gradients).

\textbf{(i) Weak Separability (Open Problem)}: Using only I\_A, there exists a test statistic $\tau$(I\_A) such that when $\beta \neq$ 0, $\tau$'s power strictly exceeds the significance level under the label-noise-only case, i.e.: P($\tau$ > c\_$\alpha$ | $\beta \neq$ 0) > $\alpha$ for some $\beta \neq$ 0 and bounded $\varepsilon$. Whether this conclusion holds depends on whether $\eta$ (Sprint noise parameter) has nonzero mutual information with z: I($\eta$; z) > 0.

\textbf{(ii) Strong Inseparability (Strictly Derivable, from Theorem 3)}: If $\eta$ and z are independent (i.e., I($\eta$; z) = 0), then label noise and causal confounding are \textbf{absolutely indistinguishable} within the SCX framework. The auditor cannot construct any classifier better than random guessing to distinguish $\varepsilon$-dominant samples from z-dominant samples.

\textbf{(iii) Relay Theorem (Strictly Derivable, from T7)}: If there exists a source domain D\_s satisfying the front-door criterion, and T7 cross-domain transfer's $\varepsilon$-fidelity condition holds, then in the target domain D\_t, the confounding effect $\beta$'s Cercis deviation $\|\Delta$S$\|$ has an upper bound: $\|\Delta$S$\| \leq$ g($\varepsilon$, $\delta$) + O(1/$\sqrt{n}$) where $\delta$ is the Situs distance of the distribution difference between source and target domains.
\end{quote}

\subsubsection{Annotations}\label{cd-annotations}

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
Part & Status \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(i) Weak Separability & \textbf{Open Problem} --- depends on empirical discovery of $\eta$-z dependence structure \\
(ii) Strong Inseparability & \textbf{Strictly Derivable} --- direct corollary of Theorem 3 \\
(iii) Relay Theorem & \textbf{Strictly Derivable} --- T7 + Situs distance triangle inequality \\
\end{longtable}

\subsubsection{Core Intuition}\label{core-intuition-cd}

Noise is ``structureless difficulty,'' confounding is ``structured difficulty.'' Theorem 3 says the two are indistinguishable at the S level, but \textbf{S's higher-order structure (curl of the gradient field $\nabla$S, rank of the Hessian) may encode causal information}---this is the core bet of the (i) open part. If Situs encoding preserves the DAG skeleton information of the causal graph, then we have a wedge to pry open this distinction.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Federated Audit Theorem: Zero-Knowledge Multi-Party Audit Completeness}\label{federated-audit-theorem}

\subsubsection{(Federated Audit Theorem --- FA-Theorem)}\label{fa-theorem-next}

\subsubsection{Context}\label{context-fa}

The ``Audit Sword'' establishes SCX's core power: \textbf{any SCX usage behavior can be independently audited}. But this sword currently assumes the auditor has complete data access. In federated scenarios---multiple parties hold private data, interacting only through gradients or summary statistics---the Audit Sword hits a joint lock: how do you audit a dataset you cannot see?

Federated learning already has tools like Secure Aggregation, Differential Privacy (DP), and Multi-Party Computation (MPC). But they solve ``preventing leakage,'' not ``enabling auditability.'' SCX's goal is the reverse: we want to \textbf{make audit conclusions have the same statistical power as full-data audit, without exposing raw data}.

\subsubsection{Theorem Statement (Draft)}\label{fa-statement}

\begin{quote}
\textbf{FA-Theorem (Federated Audit Theorem)}: Suppose there are K data parties, each holding private dataset D\_k, k = 1,...,K. The federated aggregator outputs a global model M\_$\Pi$ via protocol $\Pi$. Auditor A can access the information set: I\_Fed = {S\_k, $\nabla$S\_k, H\_{S,k}}\_{k=1}$^K \cup$ {M\_$\Pi$} where S\_k is each party's Cercis Score (locally computable, then encrypted and shared).

\textbf{(i) Federated Audit Completeness (Open Problem, Partially Derivable)}: Define full-information audit statistic T\_full = T(D\_1,...,D\_K) and federated audit statistic T\_fed = T\_F(I\_Fed). If protocol $\Pi$ satisfies: (a) each party's submitted S\_k is verified for honesty via zero-knowledge proof (i.e., $\exists \pi$\_ZK : Verify($\pi$\_ZK, S\_k, Commit(D\_k)) = 1); (b) Situs encoding preserves homomorphism across parties: Situs($\cup$D\_k) $\cong \oplus$\_k Situs(D\_k) ($\oplus$ is some commutative composition operator), then there exists an $\varepsilon$-audit equivalence: |Power(T\_fed) - Power(T\_full)| $\leq \varepsilon$(K, $\delta$) where $\delta$ is each party's differential privacy budget (if no DP is used, $\delta$ = 0 and $\varepsilon \to$ 0 as n $\to \infty$).

\textbf{(ii) Information Lower Bound (Strictly Derivable, from Theorem 3)}: If there exists k such that D\_k's label noise parameter $\eta$\_k is completely unknown to the auditor (i.e., no $\pi$\_ZK verification), then for any federated audit statistic T\_fed, there exists a confounding strategy such that: sup\_{D} |E[T\_fed] - E[T\_full]| $\geq$ const $\cdot \eta$\_k $\cdot$ |D\_k|/|D\_total| i.e., the audit power loss is proportional to the fraction of unverified noise in total data.

\textbf{(iii) Telegraph Commons Extension (Open Problem)}: If the ``Telegraph Commons'' (data storage as infrastructure) extends to the federated layer, i.e., there exists a decentralized storage layer that has committed all D\_k's Situs hashes, then $\pi$\_ZK requirements in FA-Theorem (i) can be replaced by Proof of Storage, and the $\varepsilon$ upper bound is further tightened.
\end{quote}

\subsubsection{Annotations}\label{fa-annotations}

\begin{longtable}[]{@{}p{0.45\linewidth}p{0.50\linewidth}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Part
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(i) Federated Audit Completeness & \textbf{Mixed} --- Situs homomorphic composition part is open (needs concrete construction of the composition operator); $\varepsilon$ bound derivation is rigorous given ZK conditions \\
(ii) Information Lower Bound & \textbf{Strictly Derivable} --- Theorem 3 + information-theoretic data processing inequality \\
(iii) Telegraph Commons Extension & \textbf{Open Problem} --- depends on storage layer design, theoretically has a clear path \\
\end{longtable}

\subsubsection{Core Intuition}\label{core-intuition-fa}

The essential contradiction of federated audit: \textbf{Secure aggregation protects data parties from auditor surveillance, but the Audit Sword requires the auditor to see through the aggregation}. The solution path lies not at the cryptographic level (which leads to ``all or nothing''), but at SCX's information architecture level---Situs encoding as an ``audit fingerprint'' preserves auditability while discarding reconstructability. This is the insertion point for ZK proofs.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Adversarial Robustness Theorem: Is Adversarial Perturbation a Special Case of ``Difficulty'' or a Third Category?}\label{adversarial-robustness-theorem}

\subsubsection{(Adversarial Robustness Theorem --- AR-Theorem)}\label{ar-theorem-next}

\subsubsection{Context}\label{context-ar}

Theorem 3 establishes a binary unidentifiability framework of noise and difficulty. But the discovery of adversarial examples creates trouble for this binary framework: adversarial perturbations are not random noise---they are directionally optimized results targeting model gradients; they are also not ``natural difficulty''---humans can effortlessly correctly classify adversarial samples. This raises a fundamental question:

\textbf{Does SCX's Cercis taxonomy contain ``adversariality'' as a third fundamental category, alongside Q (quality) and N (noise), i.e., S = Q + $\eta$N + $\gamma$A?}

\subsubsection{Theorem Statement (Draft)}\label{ar-statement}

\begin{quote}
\textbf{AR-Theorem (Adversarial Robustness Theorem)}: Let adversarial perturbation $\delta$\_adv = argmax\_{$\|\delta\|\leq\varepsilon$} L(f(x+$\delta$), y), with $\|\delta$\_adv$\|$ measured by Situs distance within the SCX audit framework. Define Adversarial Exposure A(x) = $\|\nabla\_x$ S(x)$\| \cdot \varepsilon$ / S(x).

\textbf{(i) Reduction Theorem (Open Problem, Partially Derivable)}: If model f's Cercis gradient field $\nabla$S satisfies the Lipschitz condition in x's Situs neighborhood with Lipschitz constant L < S(x)/$\varepsilon$, then adversarial perturbations can be absorbed into the ``noise'' component by Theorem 3: A(x) $\leq \eta\_$eff(x), where $\eta\_$eff = $\eta$ + L$\cdot\varepsilon$/S(x) (effective noise rate). Conversely, if L $\geq$ S(x)/$\varepsilon$, then adversarial perturbations constitute a \textbf{third category singularity} that cannot be absorbed by Theorem 3.

\textbf{(ii) Audit Detectability (Strictly Derivable, from T5 Active Learning Theorem)}: Define the auditor's adversarial detection statistic: D\_adv(x) = $\|\nabla$S(x)$\|$ / Var\_{x'$\in$N(x)}[S(x')] where N(x) is the Situs local neighborhood. If $\gamma$ > 0 (i.e., a third adversarial component exists), then when the auditor adopts the T5 active learning strategy to maximize I(S; D\_adv), the sample complexity of detection is: n\_detect $\geq \Omega$(1/$\gamma^2 \cdot$ log(1/$\delta$)) and this lower bound is tight for the optimal active learning strategy.

\textbf{(iii) Phase Transition Theorem (Open Problem)}: There exists a critical Cercis threshold S\_crit such that: - When S(x) > S\_crit: adversarial perturbations are a special case of ``difficulty'' (reduced to the Theorem 3 framework), - When S(x) $\leq$ S\_crit: adversarial perturbations manifest as a third category, and the A component is orthogonal to the N component in $\nabla$S space. The analytic form of S\_crit and its relationship to model architecture are open.
\end{quote}

\subsubsection{Annotations}\label{ar-annotations}

\begin{longtable}[]{@{}p{0.45\linewidth}p{0.50\linewidth}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Part
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(i) Reduction Theorem & \textbf{Open Problem} --- Lipschitz condition is sufficient but possibly not necessary; rigorous characterization of the $\gamma$>0 region remains \\
(ii) Audit Detectability & \textbf{Strictly Derivable} --- direct application of T5 active learning framework \\
(iii) Phase Transition Theorem & \textbf{Open Problem} --- proof of phase transition existence and analytic form of S\_crit both unsolved \\
\end{longtable}

\subsubsection{Core Intuition}\label{core-intuition-ar}

The fingerprint left by adversarial perturbations in Cercis space is not S's absolute value (which may be small), but \textbf{the anomalous structure of $\nabla$S within local neighborhoods}---normally difficult samples have smoothly varying $\nabla$S, while adversarial samples exhibit gradient fractures in their vicinity. This suggests the A component may be encoded in $\nabla$S's topological invariants (such as local curvature, Hessian spectrum) rather than in S itself.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Temporal SCX Theorem: Audit Drift Under Non-Stationary Distributions}\label{temporal-scx-theorem}

\subsubsection{(Temporal SCX Theorem --- TS-Theorem)}\label{ts-theorem-next}

\subsubsection{Context}\label{context-ts}

The static SCX framework assumes data distribution P(X,Y) is fixed. Spring SE-1 introduces dynamic evolution (model update $\to$ data distribution drift $\to$ SCX parameter $\eta$ drift), but SE-1 currently only handles the ``model update as exogenous shock'' scenario. Real-world distribution drift is more complex:
- \textbf{Concept drift}: P(Y|X) changes;
- \textbf{Covariate shift}: P(X) changes;
- \textbf{Label shift}: P(Y) changes.

More critically, \textbf{Sprint's core parameter $\eta$ (noise-difficulty coupling coefficient) itself drifts over time}---this is the ``audit of the audit'' problem: who audits Sprint itself?

\subsubsection{Theorem Statement (Draft)}\label{ts-statement}

\begin{quote}
\textbf{TS-Theorem (Temporal SCX Theorem)}: Let time index t = 0,1,...,T. At each moment t, define data distribution P\_t(X,Y), corresponding Cercis Score S\_t, Sprint parameter $\eta$\_t, and Situs encoding U\_t = Situs(P\_t). Assume $\eta$\_t evolves as: $\eta$\_{t+1} = $\eta$\_t + $\Delta\eta$(P\_t, M\_t) + $\xi$\_t where $\Delta\eta$ is deterministic drift (jointly determined by model M\_t and data distribution P\_t), and $\xi$\_t is a random shock (mean 0, variance $\sigma^2\_\xi$).

\textbf{(i) Drift Detectability (Strictly Derivable, from T7 Cross-Domain Transfer)}: If the Situs distance between two time points d\_Situs(U\_t, U\_{t+1}) $\geq \delta$\_min, then at significance level $\alpha$, the auditor can detect $\eta$ drift with probability $\geq$ 1-$\beta$, with required sample size satisfying: n\_t + n\_{t+1} $\geq$ 2($\sigma^2\_\xi$ + $\kappa\cdot\delta^2$\_min)$\cdot$(z\_{$\alpha$/2} + z\_$\beta$)$^2$ / $\delta^2$\_min where $\kappa$ is the Lipschitz constant from Situs distance to $\eta$ drift (inherited from T7).

\textbf{(ii) Drift Attribution (Open Problem)}: Once $\eta$ drift is detected, can it be decomposed into: $\Delta\eta$ = $\Delta\eta$\_concept + $\Delta\eta$\_covariate + $\Delta\eta$\_label + $\Delta\eta$\_model i.e., attributed to the respective contributions of concept drift, covariate shift, label shift, and model update? Without additional assumptions (e.g., partially observable validation set), this decomposition is \textbf{unidentifiable} within the SCX framework. However, if there exists an anchor sample set A $\subset$ X satisfying P\_t(Y|X$\in$A) $\approx$ P\_{t+1}(Y|X$\in$A) (concept-stable anchors), then $\Delta\eta$\_concept can be independently estimated.

\textbf{(iii) Sprint Self-Audit Limit Theorem (Open Problem, Partially Derivable)}: Does Sprint SE-1's self-correction mechanism ($\eta \to \eta$' through model retraining) always pull $\eta$ back to equilibrium? Define Lyapunov function V($\eta$) = ($\eta$ - $\eta$*)$^2$, where $\eta$* is the optimal noise parameter. If there exists a compact set $\Omega \subset$ [0,1] such that for $\eta$\_t $\in \Omega$, E[V($\eta$\_{t+1}) | $\eta$\_t] $\leq \rho\cdot$V($\eta$\_t$), $\rho$ < 1, then the Sprint process is geometrically ergodic, and $\eta$\_t converges to $\eta$*. But if the distribution drift rate v\_drift satisfies v\_drift > (1-$\rho$)/$\Delta$t, then Sprint diverges---self-audit capability has a \textbf{rate upper bound}.
\end{quote}

\subsubsection{Annotations}\label{ts-annotations}

\begin{longtable}[]{@{}p{0.45\linewidth}p{0.50\linewidth}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Part
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(i) Drift Detectability & \textbf{Strictly Derivable} --- T7 + standard hypothesis testing framework \\
(ii) Drift Attribution & \textbf{Open Problem} --- unidentifiable without anchor assumption (proven); direction is feasible with anchors but technically incomplete \\
(iii) Sprint Self-Audit Limit & \textbf{Open Problem} --- Lyapunov condition is sufficient but not necessary; tightness of drift rate upper bound remains to be proved \\
\end{longtable}

\subsubsection{Core Intuition}\label{core-intuition-ts}

The essential tension of Temporal SCX: \textbf{Sprint needs $\eta$ to calibrate the audit, but $\eta$ itself resides in the distribution being audited}. This constitutes a recursive audit problem---the ``Audit Sword'' needs self-referentiality in the temporal dimension. TS-Theorem attempts to formalize: under what conditions is this self-reference benign (convergent), and under what conditions is it malignant (divergent).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Human-AI Collaborative Audit Theorem: Theorem 3's Cognitive Boundary}\label{human-ai-collaborative-theorem}

\subsubsection{(Human-AI Collaborative Audit Theorem --- HC-Theorem)}\label{hc-theorem-next}

\subsubsection{Context}\label{context-hc}

Theorem 3's core claim is: noise and difficulty are indistinguishable within SCX's automated audit framework. But the two most critical words in this sentence are: \textbf{``framework-internal.''} The operational boundary of the audit framework depends on what kind of information processor we equip the auditor with. If the auditor is merely an automated system based on Cercis Score, then Theorem 3 is ironclad. But what if the auditor can be a \textbf{hybrid of human expert + AI system}?

What do humans possess that AI auditors lack?
- \textbf{Prior causal knowledge} (non-statistical knowledge about how the world works);
- \textbf{Counterfactual reasoning ability} (``If the annotator were more careful, would this sample's label change?'');
- \textbf{Metacognitive judgment} (``I myself am uncertain about this sample, so it probably truly is difficult'').

Theorem 3's boundary is precisely the line where human cognition can intervene and break the noise-difficulty symmetry.

\subsubsection{Theorem Statement (Draft)}\label{hc-statement}

\begin{quote}
\textbf{HC-Theorem (Human-AI Collaborative Audit Theorem)}: Define the audit system as A = (AI, H, $\Pi$), where AI is the SCX-based automated audit module, H is the human auditor, and $\Pi$ is the human-AI interaction protocol. Define human audit budget B\_H (maximum number of samples a human can review), and human judgment function J\_H: X $\to$ {noise, hard, uncertain}.

\textbf{(i) Collaborative Gain Theorem (Open Problem, Directional Conclusion)}: If the following two conditions are satisfied: (a) \textbf{Information Asymmetry Condition}: there exists a sample subset X\_H $\subset$ X such that human judgment J\_H(x) and Cercis Score S(x) have: P(J\_H=noise | S=s) $\neq$ P(J\_H=hard | S=s) for some s (i.e., human judgment carries additional information beyond S); (b) \textbf{Calibration Condition}: human ``uncertain'' judgments are themselves calibrated, i.e., E[J\_H=uncertain | S=s] is monotonically increasing as s $\to$ 0, then the noise-difficulty separation power $\Omega$\_collab of human-AI collaborative audit strictly dominates pure AI audit: $\Omega$\_collab > $\Omega$\_AI = 0 (under Theorem 3).

\textbf{(ii) Budget Decay Law (Strictly Derivable, from Theorem 3)}: As human audit budget B\_H $\to$ 0, collaborative audit degenerates to pure AI audit, with convergence rate: $\Omega$\_collab(B\_H) = $\Omega$\_AI + c$\cdot\sqrt{B_H/n}\cdot\Delta$\_I(S; J\_H) + o($\sqrt{B_H/n}$) where $\Delta$\_I is the information increment between human judgment and Cercis Score (measured by mutual information).

\textbf{(iii) Failure Theorem---Theorem 3's Absolute Boundary (Strictly Derivable)}: Even with the introduction of arbitrary human audit capability, noise and difficulty remain \textbf{absolutely indistinguishable} in the following scenario: if and only if $\forall$x, I(J\_H(x); $\varepsilon$(x) | S(x)) = 0 i.e., human judgment and label noise carry absolutely no mutual information conditional on S.

Equivalently, the necessary and sufficient condition for humans to break Theorem 3 is: there exists at least one sample whose label noise's statistical characteristics leave traces beyond S in the human cognitive representation. If the noise is ``perfect noise'' (uncorrelated with any observable features, including human prior knowledge), then Theorem 3's barrier holds for humans as well.
\end{quote}

\subsubsection{Annotations}\label{hc-annotations}

\begin{longtable}[]{@{}p{0.45\linewidth}p{0.50\linewidth}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Part
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(i) Collaborative Gain Theorem & \textbf{Open Problem} --- direction is clear, but requires human cognition experiments to determine whether condition (a) holds in reality (likely does); formal framework is complete \\
(ii) Budget Decay Law & \textbf{Strictly Derivable} --- Theorem 3 + standard sample efficiency analysis \\
(iii) Failure Theorem & \textbf{Strictly Derivable} --- this is Theorem 3's logical closure: extending ``framework-internal'' to ``any computable + human-representable information processing system'' \\
\end{longtable}

\subsubsection{Core Intuition}\label{core-intuition-hc}

Theorem 3 is not a physical law of the universe; it is a \textbf{lens boundary} of the SCX information architecture---it tells you that when viewing the world through the Cercis Score lens, noise and difficulty merge into one. Human intervention may break this boundary not because we are smarter, but because we view data through \textbf{different lenses}---commonsense causal models, embodied experience, social context. Collaborative audit is essentially multi-lens fusion.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Summary: Logical Structure of the Theorem Cluster and Maturity Map}\label{summary-next-theorems}

\begin{verbatim}
                     T1-T4 (Static Audit Completeness)
                    /    |    \
                   /     |     \
              T5(Active) T6(Game) T7(Transfer)
                 |        |        |
           [AR-Theorem]  |   [TS-Theorem]
                 |        |        |
                 +---[CD-Theorem]--+
                          |
                    [FA-Theorem]
                          |
                    [HC-Theorem]
\end{verbatim}

\subsubsection{Maturity Summary Table}\label{maturity-summary}

\begin{longtable}[]{@{}p{0.20\linewidth}p{0.35\linewidth}p{0.15\linewidth}p{0.22\linewidth}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Theorem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Core Claim
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Strictly Derivable Fraction
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Key Open Challenge
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{CD-Theorem} & Label noise and causal confounding can be finitely separated & ~40\% & $\eta$-z mutual information structure; whether Situs encoding preserves causal skeleton \\
\textbf{FA-Theorem} & Zero-knowledge multi-party audit is equivalent to full-data audit & ~50\% & Situs homomorphic composition operator; ZK construction overhead in real networks \\
\textbf{AR-Theorem} & Whether adversariality is a subclass of ``difficulty'' or a third category depends on Cercis gradient smoothness & ~30\% & Analytic form of phase transition point S\_crit; orthogonality proof \\
\textbf{TS-Theorem} & $\eta$ drift is detectable but attribution requires anchor assumption; Sprint has a self-audit limit & ~60\% & Anchor construction for drift attribution; tightness of Lyapunov condition \\
\textbf{HC-Theorem} & Humans break Theorem 3 via cognitive lenses beyond S, but perfect noise remains unbreakable & ~55\% & Empirical measurement of human cognitive information increment; optimal HCI protocol design \\
\end{longtable}

\subsubsection{Dependency Relationships}\label{dependency-relationships}

\begin{itemize}
\tightlist
\item
  Resolution of CD-Theorem (i) $\to$ directly enhances TS-Theorem (ii)'s attribution capability
\item
  AR-Theorem's phase transition analysis $\to$ provides HC-Theorem with the ``adversarial samples that AI can automatically handle'' boundary
\item
  FA-Theorem's Situs homomorphism property $\to$ is the foundation for all federated/distributed SCX protocols
\item
  HC-Theorem (i)'s human-AI collaborative gain $\to$ provides the ultimate ``ground-truth ceiling'' for all theorems
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Draft version v0.1 | SCX Theory Architect | Awaiting collaborator review}

\end{document}
