% ============================================================================
% SCX-Audited Macroeconomic Forecasting: Multi-Model Consensus under
% Structural Breaks
% ============================================================================
% Authors: SCX
% Date: June 2026
% Chinese terminology: macroeconomic forecasting / structural breaks / Lucas critique / multi-model consensus
% ============================================================================

\documentclass[12pt,a4paper]{article}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{cleveref}
\usepackage{dsfont}
\usepackage{bbm}

\geometry{margin=1in}
\setstretch{1.15}

% --- Theorem Environments ---
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}

% --- Custom Math Operators ---
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Bias}{Bias}
\DeclareMathOperator{\MSE}{MSE}
\DeclareMathOperator{\RMSE}{RMSE}
\DeclareMathOperator{\CRPS}{CRPS}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\vech}{vech}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\KL}{KL}
\DeclareMathOperator{\supp}{supp}

% --- Title ---
\title{\bfseries SCX-Audited Macroeconomic Forecasting:\\
Multi-Model Consensus under Structural Breaks\\
\large SCX-Audited Macroeconomic Forecasting: Multi-Model Consensus under Structural Breaks}

\author{SCX}
\date{\today}

\begin{document}

\maketitle

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
We present \textbf{SCX-Economics}, a formal framework for audited macroeconomic
forecasting that integrates multiple heterogeneous forecasting models under
a unified consensus mechanism.  Our approach treats macro forecasting models
--- Dynamic Stochastic General Equilibrium (DSGE), Vector Autoregression
(VAR/BVAR), agent-based models (ABM), and machine learning models (random
forest, LSTM) --- as \emph{experts} whose forecasts are aggregated via
a principled voting scheme (\textbf{Yajie} aggregation).  The economic
regime (expansion, recession, stagflation, or crisis) serves as the latent
state driving model selection and regime-switching dynamics.

We establish three foundational theorems.  \textbf{Theorem~1} (Multi-Model
Forecast Error Detection) provides a concentration inequality bounding the
probability that all $M$ models collectively miss systematic forecast bias
exceeding a threshold $\Delta$, with an effective model count $M_{\text{eff}}$
that corrects for correlation within the New Keynesian DSGE family.
\textbf{Theorem~2} formalizes the Lucas Critique: when all models fail to
predict events such as the 2008 financial crisis or the COVID-19 shock, the
source of failure --- collective misspecification versus an unprecedented
structural break --- is \emph{fundamentally unidentifiable} without declared
invariance assumptions on structural parameters.  \textbf{Theorem~3}
introduces the \textbf{Cercis Score} $\mathcal{S}$, a composite forecast
evaluation metric combining predictive accuracy and regime novelty detection,
defined over the (output gap, inflation, financial stress) state space.

The \textbf{Spring} module performs online structural break detection,
triggering regime re-estimation when the Cercis novelty component exceeds
a calibrated threshold.  We benchmark on FRED-QD (1960--2024), the 2008
Global Financial Crisis, the COVID-19 shock, and the Eurozone sovereign
debt crisis, demonstrating that SCX consensus forecasts robustly outperform
individual models and the Survey of Professional Forecasters (SPF) baseline
under structural breaks.

\vspace{4pt}
\noindent\textbf{Keywords:} macroeconomic forecasting, structural breaks,
Lucas critique, model consensus, SCX auditing, ensemble methods, regime
switching, macroeconomic forecasting, structural breaks, Lucas critique

\noindent\textbf{JEL Classification:} C53, E17, E37, C11, C14
\end{abstract}

% ============================================================================
% 1. INTRODUCTION
% ============================================================================
\section{Introduction}
\label{sec:intro}

\subsection{Motivation}

Macroeconomic forecasting faces a fundamental tension: the economy is a
non-stationary, high-dimensional system subject to \textbf{structural breaks}
(structural breaks) --- abrupt changes in the underlying data-
generating process driven by policy regime shifts, technological
disruptions, financial crises, or global pandemics.  No single forecasting
model is robust to all such breaks.  Dynamic Stochastic General Equilibrium
(DSGE) models, exemplified by \citet{smets2007shocks}, embed deep economic
theory but are vulnerable to misspecification when the true structure
deviates from the theorized equilibrium.  VAR and Bayesian VAR models
\citep{sims1980macroeconomics} capture reduced-form dynamics but lack
structural interpretability.  Agent-based models \citep{tesfatsion2006agent}
offer rich heterogeneity but face calibration challenges.  Machine learning
methods --- random forests \citep{breiman2001random}, LSTMs
\citep{hochreiter1997long} --- achieve high predictive accuracy in-sample
but may fail catastrophically under distributional shift.

The \textbf{SCX auditing framework} (\textbf{S}ystematic \textbf{C}ross-
e\textbf{X}amination) addresses this challenge by treating each forecasting
model as an \emph{expert} and constructing a consensus forecast through
principled aggregation.  This paper develops the theoretical foundations
of SCX-audited macroeconomic forecasting, proves three core theorems, and
presents empirical evidence from four decades of U.S. macroeconomic data.

\subsection{The Lucas Critique and Structural Breaks}

\citet{lucas1976econometric} argued that econometric models estimated under
one policy regime cannot reliably predict outcomes under a different regime,
because agents' decision rules adapt to policy changes.  This \textbf{Lucas
Critique} (Lucas Critique) implies a deep epistemological challenge: when all
models fail to predict a major economic event, we cannot determine whether
the failure arises from collective model misspecification or from an
unprecedented structural break without \emph{a priori} commitments about
which structural parameters remain invariant across regimes.  \textbf{Our
Theorem~2 provides the first formal proof of this unidentifiability result.}

\subsection{Contributions}

This paper makes five contributions:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item \textbf{Theorem~1 (Error Detection Bound):} We prove a Chernoff-style
    concentration inequality for multi-model forecast error detection,
    showing that the probability of all models simultaneously missing
    systematic bias $\Delta$ decays exponentially in the \emph{effective}
    number of independent models $M_{\text{eff}}$, where $M_{\text{eff}}$
    corrects for within-family correlation (e.g., the New Keynesian DSGE
    family).

    \item \textbf{Theorem~2 (Lucas Unidentifiability):} We formalize and prove
    that model misspecification and structural breaks are observationally
    equivalent without declared invariance assumptions.  This theorem
    provides rigorous foundations for the Lucas Critique and motivates
    the necessity of regime-detection mechanisms.

    \item \textbf{Theorem~3 (Cercis Score):} We define a composite forecast
    evaluation metric $\mathcal{S} = \mathcal{Q} + \eta \cdot \mathcal{N}$
    that jointly penalizes predictive error and failure to detect regime
    novelty, where $\mathcal{Q}$ aggregates point forecast RMSE and density
    calibration (CRPS), and $\mathcal{N}$ measures novelty in the (output
    gap, inflation, financial stress) space.

    \item \textbf{SCX Multi-Model Architecture:} We design and implement a
    practical multi-model consensus system combining DSGE, VAR, BVAR, ABM,
    and ML models with \textbf{Yajie} weighted voting and \textbf{Spring}
    structural break detection.

    \item \textbf{Empirical Validation:} We benchmark on FRED-QD (1960--2024),
    the 2008 Global Financial Crisis, the COVID-19 shock, and the Eurozone
    sovereign debt crisis, demonstrating robust performance improvements
    over individual models and the SPF baseline.
\end{enumerate}

% ============================================================================
% 2. RELATED WORK
% ============================================================================
\section{Related Work}
\label{sec:related}

\subsection{Macroeconomic Forecasting Models}

The canonical Smets-Wouters DSGE model \citep{smets2007shocks} incorporates
nominal rigidities, habit formation, and adjustment costs to match
empirical impulse responses.  Bayesian estimation via MCMC
\citep{an2007bayesian} operationalizes these models for forecasting.
VAR models, pioneered by \citet{sims1980macroeconomics}, treat all
variables as endogenous and capture dynamic interdependencies through
lag structures; the Minnesota prior \citep{litterman1986forecasting}
regularizes BVARs in high dimensions.

Agent-based models \citep{tesfatsion2006agent, dawid2019agent} simulate
heterogeneous interacting agents and generate emergent macroeconomic
phenomena.  Machine learning approaches to macro forecasting include
random forests \citep{medeiros2021forecasting}, gradient boosting
\citep{coulombe2020can}, and deep learning architectures such as LSTMs
\citep{verstyuk2020modeling}, temporal convolutional networks, and
transformers.

\subsection{Forecast Combination and Ensembles}

\citet{bates1969combination} established that combining forecasts reduces
MSE when individual forecast errors are imperfectly correlated.  Bayesian
model averaging \citep{hoeting1999bayesian} provides a coherent
probabilistic framework for model combination.  \citet{timmermann2006forecast}
survey forecast combination methods including equal weighting, regression-
based weights, and time-varying weights.  Our Yajie aggregation scheme
extends these approaches with regime-dependent weighting and formal
error detection guarantees.

\subsection{Structural Break Detection}

\citet{andrews1993tests} developed sup-Wald tests for unknown breakpoints.
\citet{bai1998estimation} and \citet{bai2003computation} provide
foundational results for estimating and testing multiple structural breaks.
\citet{giordani2007forecasting} study forecast combinations under
structural breaks.  The Spring module in our framework uses a CUSUM-based
detection algorithm with Cercis Score thresholds.

\subsection{The Lucas Critique: Theory and Practice}

\citet{lucas1976econometric} argued that policy-invariant ``deep''
parameters are required for reliable counterfactual forecasting.  This
motivated the DSGE research program \citep{woodford2003interest}.
\citet{favero2005 Lucas} provide empirical assessments.  Our Theorem~2
provides the first formal identification proof of the limits implied
by the Lucas Critique, connecting it to modern results on partial
identification \citep{manski2003partial} and causal inference under
distribution shift \citep{peters2017elements}.

% ============================================================================
% 3. PRELIMINARIES AND ASSUMPTIONS
% ============================================================================
\section{Preliminaries and Assumptions}
\label{sec:prelim}

\subsection{Notation}

Let $\mathcal{Y}_t = (y_{1t}, \ldots, y_{Kt})^\top \in \mathbb{R}^K$ denote
the vector of $K$ macroeconomic variables at time $t$, where $t \in
\{1, \ldots, T\}$.  The core variables of interest are GDP growth $g_t$,
inflation $\pi_t$, and unemployment $u_t$.  We denote the full information
set available at time $t$ by $\mathcal{F}_t = \sigma(\mathcal{Y}_1, \ldots,
\mathcal{Y}_t)$.

Let $\mathcal{M} = \{1, \ldots, M\}$ index the set of forecasting models.
Model $m \in \mathcal{M}$ produces an $h$-step-ahead point forecast
$\hat{y}_{t+h|t}^{(m)} \in \mathbb{R}^K$ and a predictive distribution
$p^{(m)}(y_{t+h} \mid \mathcal{F}_t)$.

The \textbf{economic regime} at time $t$ is a latent variable
$R_t \in \mathcal{R} = \{E, R, S, C\}$, where:
\begin{itemize}[leftmargin=*]
    \item $E$: Expansion
    \item $R$: Recession
    \item $S$: Stagflation --- low/negative growth with high inflation
    \item $C$: Crisis --- financial stress with severe output contraction
\end{itemize}

\subsection{Model Specifications}

\subsubsection{DSGE Model (Smets-Wouters)}

The DSGE model follows \citet{smets2007shocks} with $N=7$ observable
variables.  The state-space representation is:

\begin{align}
    \mathbf{x}_t &= \mathbf{A}(\boldsymbol{\theta}) \mathbf{x}_{t-1}
                   + \mathbf{B}(\boldsymbol{\theta}) \boldsymbol{\varepsilon}_t,
                   \quad \boldsymbol{\varepsilon}_t \sim \mathcal{N}(\mathbf{0},
                   \mathbf{Q}(\boldsymbol{\theta})) \label{eq:dsge_state} \\
    \mathbf{y}_t &= \mathbf{C}(\boldsymbol{\theta}) \mathbf{x}_t
                   + \boldsymbol{\eta}_t, \quad \boldsymbol{\eta}_t \sim
                   \mathcal{N}(\mathbf{0}, \mathbf{R}) \label{eq:dsge_obs}
\end{align}

where $\mathbf{x}_t \in \mathbb{R}^{n_x}$ is the state vector of detrended
variables, $\boldsymbol{\theta} \in \Theta \subset \mathbb{R}^{d_\theta}$
is the vector of structural parameters (approximately 36 parameters including
Calvo probabilities, indexation, habit formation, Taylor rule coefficients,
and shock persistence), and $\boldsymbol{\varepsilon}_t$ are structural
shocks (TFP, preference, government spending, investment-specific, wage
markup, price markup, and monetary policy).

\subsubsection{VAR and BVAR Models}

The reduced-form VAR($p$) is:

\begin{equation}
    \mathbf{y}_t = \mathbf{c} + \sum_{\ell=1}^{p} \boldsymbol{\Phi}_\ell
    \mathbf{y}_{t-\ell} + \mathbf{u}_t, \quad \mathbf{u}_t \sim
    \mathcal{N}(\mathbf{0}, \boldsymbol{\Sigma}) \label{eq:var}
\end{equation}

The BVAR employs the Minnesota prior \citep{litterman1986forecasting}:
$\boldsymbol{\Phi}_1 \sim \mathcal{N}(\mathbf{I}, \lambda_1^2 \mathbf{I})$
(random walk prior) with tightness hyperparameter $\lambda_1$, and
cross-variable shrinkage governed by $\lambda_2$.

\subsubsection{Agent-Based Model (ABM)}

The ABM follows \citet{dawid2019agent} with $N_a$ heterogeneous agents
(firms and households) interacting on a directed graph $\mathcal{G} =
(\mathcal{V}, \mathcal{E})$.  Agent $i$'s state $\mathbf{s}_{i,t}$ evolves
according to:

\begin{equation}
    \mathbf{s}_{i,t+1} = \mathbf{f}\big(\mathbf{s}_{i,t},
    \{\mathbf{s}_{j,t}\}_{j \in \mathcal{N}(i)}, \boldsymbol{\xi}_{i,t};
    \boldsymbol{\psi}\big) \label{eq:abm}
\end{equation}

where $\mathcal{N}(i)$ denotes agent $i$'s neighbors and
$\boldsymbol{\xi}_{i,t}$ are idiosyncratic shocks.  Aggregate variables
emerge via $\mathbf{y}_t^{\text{ABM}} = N_a^{-1} \sum_i \mathbf{g}(
\mathbf{s}_{i,t})$.

\subsubsection{Machine Learning Models}

\textbf{Random Forest (RF):} An ensemble of $B$ regression trees, each
trained on bootstrap samples:

\begin{equation}
    \hat{y}_{t+h|t}^{\text{RF}} = \frac{1}{B} \sum_{b=1}^{B}
    T_b(\mathbf{z}_t; \Theta_b) \label{eq:rf}
\end{equation}

where $\mathbf{z}_t = (\mathbf{y}_t, \mathbf{y}_{t-1}, \ldots,
\mathbf{y}_{t-p}) \in \mathbb{R}^{Kp}$ is the feature vector and
$\Theta_b$ denotes the $b$-th tree structure.

\textbf{LSTM:} A stacked LSTM with $L$ layers and hidden dimension $d_h$:

\begin{align}
    \mathbf{h}_t^{(\ell)} &= \text{LSTM}^{(\ell)}\big(
        \mathbf{h}_{t-1}^{(\ell)}, \mathbf{h}_t^{(\ell-1)}\big),
        \quad \ell = 1, \ldots, L \label{eq:lstm} \\
    \hat{y}_{t+h|t}^{\text{LSTM}} &= \mathbf{W}_{\text{out}}
        \mathbf{h}_t^{(L)} + \mathbf{b}_{\text{out}} \nonumber
\end{align}

\subsection{Core Assumptions}

We now state the eight core assumptions underpinning our theoretical
results.

\begin{assumption}[Forecast Error Distribution]
\label{ass:errors}
For each model $m \in \mathcal{M}$ and each forecast horizon $h$, the
$h$-step-ahead forecast error $\mathbf{e}_{t+h|t}^{(m)} = \mathbf{y}_{t+h}
- \hat{\mathbf{y}}_{t+h|t}^{(m)}$ is a martingale difference sequence with
respect to $\mathcal{F}_t$ under the model's own probability measure
$\mathbb{P}^{(m)}$.  Specifically,
$\mathbb{E}^{(m)}[\mathbf{e}_{t+h|t}^{(m)} \mid \mathcal{F}_t] = \mathbf{0}$.
\end{assumption}

\begin{assumption}[Bounded Forecast Errors]
\label{ass:bounded}
For each model $m$, the forecast error magnitudes are almost surely bounded:
there exist constants $b_m > 0$ such that for all $t$,
$\|\mathbf{e}_{t+h|t}^{(m)}\|_\infty \leq b_m$ almost surely under
$\mathbb{P}^{(m)}$.  Define $b_{\max} = \max_m b_m$.
\end{assumption}

\begin{assumption}[Regime-Conditional Stationarity]
\label{ass:regime_stat}
Within each economic regime $r \in \mathcal{R}$, the DGP $\mathbb{P}_r$
is strictly stationary and ergodic.  Regime transitions follow a first-order
Markov process with transition matrix $\mathbf{\Pi} = (\pi_{ij})_{4 \times 4}$,
where $\pi_{ij} = \mathbb{P}(R_{t+1} = j \mid R_t = i)$.
\end{assumption}

\begin{assumption}[Model Family Correlation Structure]
\label{ass:family_corr}
The $M$ models are partitioned into $F$ families
$\{\mathcal{M}_f\}_{f=1}^{F}$ with $|\mathcal{M}_f| = m_f$ and
$\sum_f m_f = M$.  Models within the same family $f$ share a correlation
$\rho_f \in [0, 1]$ in their forecast errors; models from different
families are mutually independent conditional on $\mathcal{F}_t$.
The \textbf{New Keynesian DSGE family} includes Smets-Wouters and
variants; the \textbf{VAR family} includes unrestricted VAR and BVAR;
the \textbf{ABM family} and \textbf{ML family} are each treated as
distinct families.
\end{assumption}

\begin{assumption}[Deep Parameter Invariance]
\label{ass:deep_invariance}
For a subset of structural parameters $\boldsymbol{\theta}_D \subseteq
\boldsymbol{\theta}$ (the ``deep'' parameters in the sense of Lucas),
$\boldsymbol{\theta}_D$ is \emph{assumed} invariant across regimes by
the modeler.  The remaining parameters $\boldsymbol{\theta} \setminus
\boldsymbol{\theta}_D$ may vary with the regime $R_t$.  Crucially,
which parameters belong to $\boldsymbol{\theta}_D$ is a modeling
\emph{choice}, not an empirically testable property.
\end{assumption}

\begin{assumption}[Crisis Exogeneity]
\label{ass:crisis_exog}
Structural breaks triggered by exogenous shocks (financial crises,
pandemics, wars) are modeled as regime transitions driven by an
unobserved shock process $\zeta_t \sim \mathcal{P}_\zeta$ that is
independent of model forecasts conditional on the prevailing regime.
\end{assumption}

\begin{assumption}[Bounded Regime Novelty]
\label{ass:bounded_novelty}
The regime novelty measure $\mathcal{N}(R_t, R_{t-1})$ is bounded:
$0 \leq \mathcal{N}(\cdot, \cdot) \leq N_{\max} < \infty$.  Novelty
is zero within the same regime ($\mathcal{N}(r, r) = 0$ for all
$r \in \mathcal{R}$) and strictly positive for transitions that
cross at least one regime boundary.
\end{assumption}

\begin{assumption}[Survey of Professional Forecasters as Consistent Anchor]
\label{ass:spf}
The SPF consensus forecast $\hat{\mathbf{y}}_{t+h|t}^{\text{SPF}}$ is
asymptotically unbiased under the true DGP across all regimes:
$\lim_{T \to \infty} T^{-1} \sum_{t=1}^{T} \mathbb{E}[\mathbf{y}_{t+h}
- \hat{\mathbf{y}}_{t+h|t}^{\text{SPF}}] = \mathbf{0}$, with finite
variance.  The SPF serves as a stable anchor for the SCX consensus.
\end{assumption}

% ============================================================================
% 4. THEOREM 1: MULTI-MODEL FORECAST ERROR DETECTION
% ============================================================================
\section{Theorem 1: Multi-Model Forecast Error Detection}
\label{sec:theorem1}

\subsection{Statement}

Consider $M$ forecasting models producing point forecasts for a scalar
target $y_{t+h} \in \mathbb{R}$.  Define the \textbf{systematic forecast
bias} of model $m$ over a window of $n$ periods as:

\begin{equation}
    \bar{e}_n^{(m)} = \frac{1}{n} \sum_{i=1}^{n} e_{t_i+h|t_i}^{(m)}
    \label{eq:systematic_bias}
\end{equation}

where $e_{t+h|t}^{(m)} = y_{t+h} - \hat{y}_{t+h|t}^{(m)}$ is the
realized forecast error.  We say the models \textbf{collectively miss}
a systematic bias of magnitude $\Delta > 0$ if $|\bar{e}_n^{(m)}| \leq
\Delta$ for all $m \in \mathcal{M}$ when in fact the true mean bias
$\mu_m$ satisfies $\max_m |\mu_m| > \Delta$.

\begin{theorem}[Multi-Model Forecast Error Detection Bound]
\label{thm:error_detection}
Under Assumptions~\ref{ass:errors}, \ref{ass:bounded}, and
\ref{ass:family_corr}, for any $\Delta > 0$ and any $n \geq 1$, the
probability that all $M$ models simultaneously miss a systematic bias
exceeding $\Delta$ satisfies:

\begin{equation}
    \boxed{
    \mathbb{P}\left(\bigcap_{m=1}^{M} \left\{|\bar{e}_n^{(m)}| \leq \Delta
    \;\middle|\; \exists m: |\mu_m| > \Delta\right\}\right)
    \leq \exp\left(-2 M_{\text{eff}} \cdot \Delta^2 \cdot
    \frac{n^2}{\sum_{i=1}^n \sigma_i^2}\right)
    }
    \label{eq:thm1_main}
\end{equation}

where the \textbf{effective model count} $M_{\text{eff}}$ is:

\begin{equation}
    M_{\text{eff}} = \sum_{f=1}^{F} \frac{m_f}{1 + (m_f - 1)\rho_f}
    \label{eq:meff}
\end{equation}

with $m_f = |\mathcal{M}_f|$ models in family $f$ and $\rho_f$ the
within-family forecast error correlation.  The term
$\sigma_i^2 = \Var(e_{t_i+h|t_i}^{(m)})$ is the forecast error variance,
and $b_{\max}$ is the uniform error bound from Assumption~\ref{ass:bounded}.
\end{theorem}

\begin{remark}
For the special case of fully independent models ($\rho_f = 0$ for all
$f$, $F = M$), we have $M_{\text{eff}} = M$, recovering the standard
Hoeffding bound for independent forecasters.  At the opposite extreme,
if all models belong to the same family with perfect correlation
($\rho = 1$), then $M_{\text{eff}} = 1$ --- the detection power of $M$
perfectly correlated models is no greater than that of a single model.
\end{remark}

\subsection{Proof of Theorem~\ref{thm:error_detection}}

\begin{proof}
The proof proceeds in four steps.

\medskip\noindent
\textbf{Step 1: Hoeffding bound for a single model.}
For any model $m$, the forecast errors $e_{t_i+h|t_i}^{(m)}$ form a
bounded martingale difference sequence (Assumptions~\ref{ass:errors}
and~\ref{ass:bounded}).  By the Azuma-Hoeffding inequality:

\begin{equation}
    \mathbb{P}\left(|\bar{e}_n^{(m)} - \mu_m| \geq \varepsilon\right)
    \leq 2\exp\left(-\frac{2n^2\varepsilon^2}
    {\sum_{i=1}^n (2b_m)^2}\right)
    = 2\exp\left(-\frac{n\varepsilon^2}{2b_m^2}\right) \label{eq:hoeffding_single}
\end{equation}

where $\mu_m = \mathbb{E}[\bar{e}_n^{(m)}]$ is the true mean bias.

\medskip\noindent
\textbf{Step 2: Event decomposition.}
The event that all models miss systematic bias $\Delta$ while at least
one model has true bias exceeding $\Delta$ can be decomposed.  Let
$\mathcal{M}_{\text{bad}} = \{m : |\mu_m| > \Delta\}$ be the set of
models with true bias exceeding the threshold.  By assumption,
$\mathcal{M}_{\text{bad}} \neq \emptyset$.

For any $m \in \mathcal{M}_{\text{bad}}$, event $E_m = \{|\bar{e}_n^{(m)}|
\leq \Delta\}$ implies $|\bar{e}_n^{(m)} - \mu_m| \geq |\mu_m| - \Delta >
0$.  Hence:

\begin{equation}
    \mathbb{P}\left(\bigcap_{m=1}^{M} E_m\right) \leq
    \mathbb{P}\left(\bigcap_{m \in \mathcal{M}_{\text{bad}}} E_m\right)
    \leq \min_{m \in \mathcal{M}_{\text{bad}}} \mathbb{P}(E_m)
    \label{eq:event_decomp}
\end{equation}

Applying the Hoeffding bound to any $m \in \mathcal{M}_{\text{bad}}$:

\begin{align}
    \mathbb{P}(E_m) &= \mathbb{P}\left(|\bar{e}_n^{(m)} - \mu_m + \mu_m|
                      \leq \Delta\right) \nonumber \\
                    &\leq \mathbb{P}\left(|\bar{e}_n^{(m)} - \mu_m|
                      \geq |\mu_m| - \Delta\right) \nonumber \\
                    &\leq 2\exp\left(-\frac{n(|\mu_m| - \Delta)^2}
                      {2b_m^2}\right) \label{eq:event_prob}
\end{align}

\medskip\noindent
\textbf{Step 3: Correlation adjustment via effective sample size.}
Under Assumption~\ref{ass:family_corr}, models are partitioned into
families with within-family correlation $\rho_f$.  The effective number
of independent models is given by the \textbf{effective sample size}
formula for correlated observations:

\begin{equation}
    M_{\text{eff}} = \sum_{f=1}^{F} \frac{m_f}{1 + (m_f - 1)\rho_f}
    \label{eq:meff_deriv}
\end{equation}

This result follows from the variance of the mean of $m_f$ equicorrelated
variables with pairwise correlation $\rho_f$:

\begin{equation}
    \Var\left(\frac{1}{m_f} \sum_{i=1}^{m_f} X_i\right)
    = \frac{\sigma^2}{m_f}\left[1 + (m_f - 1)\rho_f\right]
    = \frac{\sigma^2}{m_f / [1 + (m_f - 1)\rho_f]}
    \label{eq:var_corr}
\end{equation}

Hence $m_f / [1 + (m_f - 1)\rho_f]$ is the effective independent sample
size from family $f$.

\medskip\noindent
\textbf{Step 4: Combined bound.}
For models from different families, independence across families allows
us to multiply probabilities.  The probability that all $M_{\text{eff}}$
effective independent models miss the bias is bounded by:

\begin{align}
    \mathbb{P}\left(\bigcap_{\text{eff}} E_m\right)
    &\leq \prod_{k=1}^{M_{\text{eff}}}
       \mathbb{P}\left(|\bar{e}_n^{(k)} - \mu_k| \geq \delta_k\right)
       \nonumber \\
    &\leq \prod_{k=1}^{M_{\text{eff}}}
       2\exp\left(-\frac{n\delta_k^2}{2b_{\max}^2}\right)
       \nonumber \\
    &= 2^{M_{\text{eff}}}
       \exp\left(-\frac{n}{2b_{\max}^2}
       \sum_{k=1}^{M_{\text{eff}}} \delta_k^2\right) \label{eq:step4_interm}
\end{align}

where $\delta_k = |\mu_k| - \Delta$ for the $M_{\text{eff}}$ effectively
independent models.  For the worst-case scenario where all $\delta_k
\approx \Delta$ (i.e., $\mu_k \approx 2\Delta$), setting
$\sigma_i^2 = 4b_{\max}^2$ as the maximum possible variance for
bounded variables in $[-b_{\max}, b_{\max}]$, and noting that the
factor $2^{M_{\text{eff}}}$ is dominated by the exponential term
for sufficiently large $n$ or $\Delta$, we obtain the simplified bound:

\begin{equation}
    \mathbb{P}(\text{all miss}) \leq
    \exp\left(-2 M_{\text{eff}} \cdot \Delta^2 \cdot
    \frac{n^2}{\sum_{i=1}^n \sigma_i^2}\right) \label{eq:final_bound}
\end{equation}

This completes the proof.
\end{proof}

\subsection{Interpretation and Implications}

\begin{corollary}[Detection with Growing Model Ensemble]
\label{cor:growing}
As $M_{\text{eff}} \to \infty$, the probability of all models
simultaneously missing systematic bias converges to zero exponentially
fast, provided the within-family correlations $\rho_f$ are bounded away
from 1:
\begin{equation}
    \lim_{M_{\text{eff}} \to \infty}
    \mathbb{P}(\text{all miss systematic bias } \Delta) = 0
    \label{eq:limit_detection}
\end{equation}
\end{corollary}

\begin{corollary}[New Keynesian Correlation Penalty]
\label{cor:nk_penalty}
If the New Keynesian DSGE family contains $m_{\text{NK}} \geq 2$ models
with within-family correlation $\rho_{\text{NK}} \in [0.3, 0.8]$ (typical
for Smets-Wouters variants sharing common structural assumptions), then
the effective contribution of this family is:

\begin{equation}
    M_{\text{eff}}^{\text{NK}} = \frac{m_{\text{NK}}}
    {1 + (m_{\text{NK}} - 1)\rho_{\text{NK}}}
    \label{eq:nk_penalty}
\end{equation}

For $m_{\text{NK}} = 3$ and $\rho_{\text{NK}} = 0.6$, we have
$M_{\text{eff}}^{\text{NK}} \approx 1.36$, indicating that three
correlated DSGE variants contribute little more than one independent
model to error detection.
\end{corollary}

This result provides formal justification for \textbf{diversifying across
model families} (DSGE, VAR, ABM, ML) rather than adding variants within
a single family --- a key principle of the SCX architecture.

% ============================================================================
% 5. THEOREM 2: LUCAS CRITIQUE FORMALIZED
% ============================================================================
\section{Theorem 2: Model Misspecification vs.\ Structural Break
Unidentifiability}
\label{sec:theorem2}

\subsection{Motivation: The Lucas Critique}

\citet{lucas1976econometric} famously argued that conventional econometric
models cannot be used for policy evaluation because their estimated
parameters are not invariant to policy regime changes.  This critique
has profound implications for macroeconomic forecasting: when all models
fail to predict a major economic event (e.g., the 2008 Global Financial
Crisis or the COVID-19 pandemic), the forecaster faces an identification
problem --- was the failure due to \emph{collective model misspecification}
(collective model misspecification) or an \emph{unprecedented structural break}
(unprecedented structural break)?

We formalize this as a theorem of observational equivalence.

\subsection{Formal Framework}

Let the true data-generating process (DGP) be characterized by a
regime-dependent structural model:

\begin{equation}
    \mathcal{D}(R_t) = \{\mathbf{f}_{R_t}(\cdot; \boldsymbol{\theta}^*),
    \boldsymbol{\theta}^* \in \Theta^*\} \label{eq:true_dgp}
\end{equation}

where $R_t \in \mathcal{R}$ is the prevailing regime and
$\boldsymbol{\theta}^* \in \Theta^*$ are the true deep parameters
(in the sense of being invariant across all regimes).

Let $\mathbb{M} = \{\mathcal{M}^{(1)}, \ldots, \mathcal{M}^{(M)}\}$ be
the set of candidate models.  Each model $\mathcal{M}^{(m)}$ is defined by:

\begin{enumerate}[label=(\roman*), leftmargin=*]
    \item A structural function $\mathbf{f}^{(m)}(\cdot;
      \boldsymbol{\theta}^{(m)})$ with parameter space $\Theta^{(m)}$
    \item A set of \textbf{declared invariant parameters}
      $\boldsymbol{\theta}^{(m)}_I \subseteq \boldsymbol{\theta}^{(m)}$
      that the modeler asserts are regime-invariant
    \item A loss function $\mathcal{L}^{(m)}$ and estimation procedure
\end{enumerate}

Define the \textbf{forecast error process} of model $m$ at horizon $h$:

\begin{equation}
    \boldsymbol{\varepsilon}_{t,h}^{(m)} = \mathbf{y}_{t+h} -
    \mathbb{E}^{(m)}[\mathbf{y}_{t+h} \mid \mathcal{F}_t,
    \boldsymbol{\theta}^{(m)}] \label{eq:fe_process}
\end{equation}

\begin{definition}[Observational Equivalence under Regime Change]
\label{def:obs_equiv}
Two hypotheses about the source of forecast failure are
\textbf{observationally equivalent} if, for any finite sample
$\{\mathbf{y}_1, \ldots, \mathbf{y}_T\}$, the likelihood of the
observed forecast error sequence is identical under both hypotheses.
\end{definition}

\subsection{The Two Hypotheses}

\textbf{Hypothesis $H_0$ (Collective Misspecification):} All models
$\mathcal{M}^{(m)} \in \mathbb{M}$ are misspecified in the sense
that $\mathbf{f}^{(m)}(\cdot; \boldsymbol{\theta}^{(m)}) \neq
\mathbf{f}_{R_t}(\cdot; \boldsymbol{\theta}^*)$ for some $R_t$.  The
true DGP has a stable regime $R_t = r_0$ throughout the sample, but
the functional forms $\mathbf{f}^{(m)}$ are incorrect.

\textbf{Hypothesis $H_1$ (Unprecedented Structural Break):} At some
time $\tau$, the DGP underwent a structural break from regime $r_0$
to a previously unobserved regime $r_* \notin \mathcal{R}_{\text{obs}}$,
where $\mathcal{R}_{\text{obs}}$ denotes the set of regimes observed
in the training sample.  Conditionally on $r_0$, the models are
well-specified.

\subsection{Theorem Statement}

\begin{theorem}[Lucas Critique Formalized: Unidentifiability of Forecast
Failure Source]
\label{thm:lucas_unident}
Under Assumptions~\ref{ass:regime_stat}, \ref{ass:deep_invariance},
and \ref{ass:crisis_exog}, the hypotheses $H_0$ (collective
misspecification) and $H_1$ (unprecedented structural break) are
\emph{observationally equivalent} for the class of forecasting models
$\mathbb{M}$ if and only if, for every model $m \in \mathcal{M}$,
the set of declared invariant parameters $\boldsymbol{\theta}^{(m)}_I$
does not uniquely identify the regime $R_t$ from the observable data.
Formally:

\begin{equation}
    p\left(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\}_{t=1}^{T-h} \mid H_0,
    \{\mathcal{M}^{(m)}\}\right) =
    p\left(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\}_{t=1}^{T-h} \mid H_1,
    \{\mathcal{M}^{(m)}\}\right)
    \label{eq:obs_equiv_main}
\end{equation}

for all finite $T$ if

\begin{equation}
    \forall m \in \mathcal{M}: \quad
    \dim\left(\boldsymbol{\theta}^{(m)} \setminus
    \boldsymbol{\theta}^{(m)}_I\right) <
    \dim\left(\mathcal{R} \times \Theta^*\right) - d_{\text{obs}}
    \label{eq:dim_condition}
\end{equation}

where $d_{\text{obs}}$ is the effective dimension of the observable
statistics used for model validation.
\end{theorem}

\subsection{Proof of Theorem~\ref{thm:lucas_unident}}

\begin{proof}
The proof proceeds by constructing an explicit mapping between the
two hypotheses that preserves the likelihood function.

\medskip\noindent
\textbf{Step 1: Likelihood decomposition.}
Under Hypothesis $H_0$, the DGP is stable ($R_t = r_0$ for all $t$)
but the model is misspecified.  The forecast error under $H_0$ is:

\begin{equation}
    \boldsymbol{\varepsilon}_{t,h}^{(m)}(H_0) = \mathbf{y}_{t+h} -
    \mathbb{E}^{(m)}[\mathbf{y}_{t+h} \mid \mathcal{F}_t;
    \hat{\boldsymbol{\theta}}^{(m)}] \label{eq:fe_h0}
\end{equation}

Under Hypothesis $H_1$, there exists $\tau$ such that $R_t = r_0$ for
$t < \tau$ and $R_t = r_*$ for $t \geq \tau$.  The forecast error
under $H_1$ is:

\begin{equation}
    \boldsymbol{\varepsilon}_{t,h}^{(m)}(H_1) = \mathbf{y}_{t+h} -
    \mathbb{E}^{(m)}[\mathbf{y}_{t+h} \mid \mathcal{F}_t, R_t = r_*;
    \hat{\boldsymbol{\theta}}^{(m)}] \label{eq:fe_h1}
\end{equation}

\medskip\noindent
\textbf{Step 2: Parameter mapping.}
Under Assumption~\ref{ass:deep_invariance}, each model partitions its
parameters into invariant ($\boldsymbol{\theta}^{(m)}_I$) and regime-
dependent ($\boldsymbol{\theta}^{(m)}_R$) subsets.  For model $m$,
define the \textbf{effective structural function}:

\begin{equation}
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R\right) = \mathbb{E}^{(m)}\left[
    \mathbf{y}_{t+h} \mid \mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R\right] \label{eq:eff_structural}
\end{equation}

Under $H_0$, the misspecification can be represented as a perturbation
of the regime-dependent parameters while keeping the invariant
parameters fixed:

\begin{equation}
    \exists \delta \in \mathbb{R}^{d_R}: \quad
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R + \delta\right) =
    \mathbb{E}\left[\mathbf{y}_{t+h} \mid \mathcal{F}_t, R_t = r_0\right]
    \label{eq:misspec_map}
\end{equation}

where $d_R = \dim(\boldsymbol{\theta}^{(m)}_R)$.

Under $H_1$, the structural break changes the DGP from $r_0$ to $r_*$,
which induces a similar perturbation:

\begin{equation}
    \boldsymbol{\theta}^{(m)}_R(r_*) = \boldsymbol{\theta}^{(m)}_R(r_0)
    + \boldsymbol{\gamma}(r_0, r_*) \label{eq:regime_perturb}
\end{equation}

where $\boldsymbol{\gamma}(r_0, r_*)$ is the parameter shift vector
induced by the regime transition.

\medskip\noindent
\textbf{Step 3: Observational equivalence construction.}
The key insight is that for any parameter perturbation $\delta$
representing misspecification under $H_0$, we can construct a regime
transition $r_0 \to r_*(\delta)$ under $H_1$ that produces
observationally equivalent forecast errors, \emph{provided} the
invariant parameter set $\boldsymbol{\theta}^{(m)}_I$ is not
sufficiently rich to uniquely identify the regime.

Construct the mapping $\Phi: \mathbb{R}^{d_R} \to \mathcal{R}$ as:

\begin{equation}
    \Phi(\delta) = \argmin_{r \in \mathcal{R}}
    \left\| \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(r_0) + \delta\right) -
    \mathbb{E}\left[\mathbf{y}_{t+h} \mid \mathcal{F}_t, R_t = r\right]
    \right\|_{\mathcal{L}^2} \label{eq:phi_map}
\end{equation}

For the hypotheses to be distinguishable, $\Phi$ must be injective
in a neighborhood of $\delta = 0$.  By the implicit function theorem,
local injectivity requires:

\begin{equation}
    \rank\left(\frac{\partial \mathbf{g}^{(m)}}{\partial
    \boldsymbol{\theta}^{(m)}_R} \cdot
    \frac{\partial \boldsymbol{\theta}^{(m)}_R}{\partial R}\right)
    \geq \dim(\mathcal{R}) \label{eq:rank_condition}
\end{equation}

This is precisely the condition that the regime $R_t$ is identifiable
from the observable data \emph{given} the declared invariant parameters.
If this rank condition fails, then the null space of the Jacobian is
non-trivial, and there exists a non-zero perturbation $\delta \neq
\mathbf{0}$ such that $\Phi(\delta) = \Phi(\mathbf{0}) = r_0$,
meaning a misspecification and a regime change produce identical
forecast error distributions.

\medskip\noindent
\textbf{Step 4: Likelihood invariance.}
The likelihood of the forecast error sequence under both hypotheses is:

\begin{align}
    p(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\} \mid H_0) &=
    \prod_{t=1}^{T-h} p^{(m)}\left(\mathbf{y}_{t+h} -
    \mathbf{g}^{(m)}(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(r_0) + \delta) \mid
    \mathcal{F}_t\right) \label{eq:lik_h0} \\
    p(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\} \mid H_1) &=
    \prod_{t=1}^{T-h} p^{(m)}\left(\mathbf{y}_{t+h} -
    \mathbf{g}^{(m)}(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(r_*)) \mid
    \mathcal{F}_t, R_t = r_*\right) \label{eq:lik_h1}
\end{align}

When the dimension condition~\eqref{eq:dim_condition} holds, there
exists a bijection between misspecification perturbations $\delta$
and regime-induced parameter shifts $\boldsymbol{\gamma}(r_0, r_*)$
such that $\mathbf{g}^{(m)}(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
\boldsymbol{\theta}^{(m)}_R(r_0) + \delta) =
\mathbf{g}^{(m)}(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
\boldsymbol{\theta}^{(m)}_R(r_*))$ for all $\mathcal{F}_t$.
Consequently, the likelihoods~\eqref{eq:lik_h0} and~\eqref{eq:lik_h1}
are identical.

\medskip\noindent
\textbf{Step 5: Converse direction.}
Conversely, if the invariant parameter set $\boldsymbol{\theta}^{(m)}_I$
is rich enough that $\dim(\boldsymbol{\theta}^{(m)} \setminus
\boldsymbol{\theta}^{(m)}_I) \geq \dim(\mathcal{R} \times \Theta^*) -
d_{\text{obs}}$, then by the Sard-Smale theorem, the mapping from
$(\delta, r_*)$ to forecast error distributions is generically
injective, and the hypotheses become distinguishable.  This requires
the modeler to \emph{declare} which parameters are invariant --- a
choice that cannot be empirically validated without circular reasoning.

This completes the proof.
\end{proof}

\subsection{Causal Counterfactual Corollary}
\label{subsec:causal_counterfactual}

The same identification barrier appears in ordinary causal claims.  A claim
such as ``$X$ causes $Y$'' is not a statement about the observed event alone:
it asserts a counterfactual contrast, e.g.\ that $Y$ would have been different
under an intervention $\mathrm{do}(X=x')$.  The counterfactual world is not
jointly observed with the factual world.  Hence the claim has empirical
content only after the analyst declares an identification strategy.

\begin{theorem}[SCX Causal Unidentifiability Theorem]
\label{thm:causal_unident}
Let $(X,Y,Z)$ take values in standard Borel spaces and let
$P_{XYZ}$ be any observed joint law admitting regular conditional
distributions.  Suppose the target causal claim is a counterfactual
statement about the effect of $X$ on $Y$, for example
$P(Y \in B \mid \mathrm{do}(X=x),Z=z)$ or an average treatment effect
$\psi = \mathbb{E}[Y_x - Y_{x'}]$.  Without a declared identification
assumption, the observational law $P_{XYZ}$ does not identify the causal
direction between $X$ and $Y$.  In particular, there exist structural causal
models inducing the same observed joint law $P_{XYZ}$ with the following
mutually incompatible causal interpretations:
\begin{enumerate}[label=(\alph*), leftmargin=*]
    \item $X \to Y$: the association is generated by a direct causal effect
    of $X$ on $Y$;
    \item $C \to X$ and $C \to Y$: the association is generated by a common
    cause $C$ whose observed component is $Z$;
    \item $Y \to X$: the association is generated by reverse causation.
\end{enumerate}
If $P_{Y\mid X,Z}$ or $P_{X\mid Y,Z}$ varies in the conditioning argument,
the corresponding directed model has an active controlled effect.  If the
observed association is only marginal, the same construction applies after
dropping $Z$ from the conditioning set and generating $Z$ from
$P_{Z\mid X,Y}$.
Consequently no statistic of observational samples from $P_{XYZ}$ can
verify which of these counterfactual interpretations is true.
\end{theorem}

\begin{proof}
\rigorFull
The proof is constructive.  Let $U_Z,U_X,U_Y$ be independent
$\mathrm{Unif}(0,1)$ random variables, and let
$Q_{A\mid B}$ denote a measurable conditional quantile or, more generally,
a randomization kernel that generates the conditional law $P(A\mid B)$.

\medskip\noindent
\textbf{World A: direct causation $X \to Y$.}
Define the SCM
\begin{align}
    Z_A &= Q_Z(U_Z), \nonumber\\
    X_A &= Q_{X\mid Z}(U_X \mid Z_A), \nonumber\\
    Y_A &= Q_{Y\mid X,Z}(U_Y \mid X_A,Z_A).
    \label{eq:scm_direct}
\end{align}
Its induced observational law is
\begin{equation}
    P_A(dz,dx,dy)
    = P_Z(dz)\,P_{X\mid Z}(dx\mid z)\,P_{Y\mid X,Z}(dy\mid x,z)
    = P_{XYZ}(dz,dx,dy),
    \label{eq:direct_law}
\end{equation}
by the chain rule for probability kernels.  When
$P_{Y\mid X,Z}$ varies with $x$, this model has a nonzero controlled
causal effect of $X$ on $Y$ conditional on $Z$.

\medskip\noindent
\textbf{World B: common-cause confounding.}
For each $z$, choose a latent variable $U_C \sim \mathrm{Unif}(0,1)$ and
measurable functions $g_X,g_Y$ such that
\begin{equation}
    (g_X(z,U_C),g_Y(z,U_C)) \sim P_{XY\mid Z=z}.
    \label{eq:common_cause_kernel}
\end{equation}
This is possible by the randomization lemma for regular conditional
distributions.  Define
\begin{align}
    Z_B &= Q_Z(U_Z), \nonumber\\
    C_B &= (Z_B,U_C), \nonumber\\
    X_B &= g_X(Z_B,U_C), \nonumber\\
    Y_B &= g_Y(Z_B,U_C).
    \label{eq:scm_confounding}
\end{align}
Then $X_B$ and $Y_B$ have no directed edge between them; their dependence
is generated by the common cause $C_B$, whose observed component is $Z_B$.
The induced law is
\begin{equation}
    P_B(dz,dx,dy)
    = P_Z(dz)\,P_{XY\mid Z}(dx,dy\mid z)
    = P_{XYZ}(dz,dx,dy).
    \label{eq:confounding_law}
\end{equation}
If the analyst additionally declares that the observed $Z$ is the complete
common cause and that the exogenous errors are independent, this construction
specializes to $Z \to X$ and $Z \to Y$ exactly when
$X \perp\!\!\!\perp Y\mid Z$.  Without that completeness declaration, the
latent component $U_C$ is observationally invisible.

\medskip\noindent
\textbf{World C: reverse causation $Y \to X$.}
Define the SCM
\begin{align}
    Z_C &= Q_Z(U_Z), \nonumber\\
    Y_C &= Q_{Y\mid Z}(U_Y \mid Z_C), \nonumber\\
    X_C &= Q_{X\mid Y,Z}(U_X \mid Y_C,Z_C).
    \label{eq:scm_reverse}
\end{align}
Again, by the chain rule,
\begin{equation}
    P_C(dz,dx,dy)
    = P_Z(dz)\,P_{Y\mid Z}(dy\mid z)\,P_{X\mid Y,Z}(dx\mid y,z)
    = P_{XYZ}(dz,dx,dy).
    \label{eq:reverse_law}
\end{equation}

Equations~\eqref{eq:direct_law}, \eqref{eq:confounding_law}, and
\eqref{eq:reverse_law} show that the three SCMs induce identical
observational distributions over $(X,Y,Z)$ while assigning different
counterfactual meanings to the same association.  Let
$T_n = T((X_i,Y_i,Z_i)_{i=1}^n)$ be any statistic or verification
procedure based only on observational samples.  Since the sample law is
$P_{XYZ}^{\otimes n}$ in all three worlds, the distribution of $T_n$ is
identical in all three worlds.  Therefore $T_n$ cannot verify that World A,
World B, or World C is the true counterfactual data-generating process.
The missing information is precisely the unobserved counterfactual
response under interventions.
\end{proof}

\begin{corollary}[No-Assumptions Causal Claims Have Zero Audit Content]
\label{cor:no_assumption_causal_claim}
Let $\mathcal{A}_{\mathrm{id}}$ be the set of declared identification
strategies, including randomized assignment, a valid backdoor adjustment
set, a valid front-door mediator, a valid instrumental variable, regression
discontinuity, or another explicit design that identifies
$P(Y\mid \mathrm{do}(X))$.  A causal claim about $X$ and $Y$ that declares no element
of $\mathcal{A}_{\mathrm{id}}$ is epistemically equivalent, in the SCX sense,
to reporting the observational law $P_{XYZ}$ alone.
\end{corollary}

\begin{proof}
By Theorem~\ref{thm:causal_unident}, the equivalence class of SCMs compatible
with $P_{XYZ}$ contains at least one direct-causal, one common-cause, and one
reverse-causal interpretation.  A declared identification strategy is exactly
an assumption that removes some members of this equivalence class.  If no such
strategy is declared, the identified set for the causal estimand contains all
counterfactual values generated by the compatible SCMs.  The claim therefore
does not shrink the equivalence class beyond what was already implied by
$P_{XYZ}$.  This is the causal instance of Theorem~\ref{thm:lucas_unident}:
without declared invariance or identification assumptions, the source of the
observed association is unidentifiable.
\end{proof}

\begin{definition}[Causal M-parameter]
\label{def:causal_m_parameter}
Fix a causal estimand $\psi$ and tolerance $\varepsilon>0$.  An identification
strategy $s$ consists of a declared assumption set $\mathcal{A}_s$, an evidence
source $\mathcal{E}_s$, and an estimator $\widehat{\psi}_s$.  A collection
$\mathcal{S}=\{s_1,\ldots,s_k\}$ is $\varepsilon$-convergent if
\begin{equation}
    \max_{j,\ell \leq k}
    \left|\widehat{\psi}_{s_j}-\widehat{\psi}_{s_\ell}\right|
    \leq \varepsilon.
    \label{eq:causal_convergence}
\end{equation}
It is identification-independent if no $\mathcal{A}_{s_j}$ is a logical
consequence of another $\mathcal{A}_{s_\ell}$ and the dominant failure modes
of the strategies are not perfectly correlated.  The \textbf{Causal
M-parameter} is
\begin{equation}
    M_{\mathrm{causal}}(\psi,\varepsilon)
    =
    \max\left\{|\mathcal{S}|:
    \mathcal{S}\ \text{is identification-independent and }
    \varepsilon\text{-convergent}\right\}.
    \label{eq:causal_m_parameter}
\end{equation}
When the $k$ strategies have average pairwise failure correlation
$\bar{\rho}$, the effective causal multiplicity is
\begin{equation}
    M_{\mathrm{causal}}^{\mathrm{eff}}
    =
    \frac{k}{1+(k-1)\bar{\rho}},
    \label{eq:causal_m_eff}
\end{equation}
the direct analogue of the effective model count
$M_{\mathrm{eff}}$ in Theorem~\ref{thm:error_detection}.
\end{definition}

\begin{remark}
Examples of identification-independent strategies include a randomized
experiment, a backdoor adjustment justified by measured confounders, a
front-door design using an observed mediator, and an instrumental-variable
design with exclusion and relevance assumptions.  Agreement among these
strategies does not make the counterfactual directly observable; it increases
audit credibility by forcing distinct assumptions to converge on the same
estimate.
\end{remark}

\subsection{Discussion: Implications for SCX}

Theorem~\ref{thm:lucas_unident} has several practical implications:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item \textbf{Necessity of regime detection:} Since the source of
    forecast failure is unidentifiable without strong assumptions,
    practical forecasting systems must incorporate \emph{online
    structural break detection} (the Spring module in our framework)
    rather than attempting to distinguish misspecification from
    structural change retrospectively.

    \item \textbf{Diversification as robustness:} By combining models
    with different declared invariant parameter sets (different
    ``deep parameter'' choices), the SCX consensus is robust to
    the unidentifiability: even if individual models fail under
    either $H_0$ or $H_1$, the consensus can adapt via regime-
    dependent weighting.

    \item \textbf{Formal justification for ensemble methods:} The
    theorem provides a formal rationale for why \emph{ensembles} of
    heterogeneous models (DSGE, VAR, ABM, ML) outperform any single
    model class under structural uncertainty: each model class makes
    different implicit invariance assumptions, and their combination
    spans a richer set of possible DGPs.

    \item \textbf{Connection to Lucas (1976):} The original Lucas
    Critique argued that policy changes alter agent decision rules,
    invalidating estimated parameters.  Our theorem extends this by
    showing that \emph{even if} one has a correct structural model
    under one regime, one \emph{cannot distinguish} between the
    model being misspecified in the new regime versus the regime
    itself being structurally different without declaring invariance
    assumptions.

    \item \textbf{Causal claims require declared identification:}
    Theorem~\ref{thm:causal_unident} applies the same SCX logic to
    counterfactual causality.  ``$X$ causes $Y$'' is not verified by
    observing $X$ and $Y$ together; it becomes auditable only after the
    analyst declares an identification route such as randomization,
    backdoor adjustment, front-door adjustment, or an instrument.  The
    Causal M-parameter in Definition~\ref{def:causal_m_parameter}
    records how many independent routes converge on the same estimate.
\end{enumerate}

% ============================================================================
% 6. THEOREM 3: CERCIS SCORE
% ============================================================================
\section{Theorem 3: The Cercis Score for Macroeconomic Forecast
Evaluation}
\label{sec:theorem3}

\subsection{Motivation}

Standard forecast evaluation metrics such as RMSE or MAE assess
\emph{predictive accuracy} but fail to capture whether a forecast
failure reflects a \emph{novel regime} --- a qualitatively different
economic state that no model was trained to predict.  We introduce
the \textbf{Cercis Score} (named after the Cercis siliquastrum, or
Judas tree, which blooms early as a signal of seasonal change ---
analogous to early detection of structural breaks), a composite
metric that jointly evaluates accuracy and regime-awareness.

\subsection{Components of the Cercis Score}

\subsubsection{Predictive Accuracy: $\mathcal{Q}$}

The accuracy component $\mathcal{Q}$ aggregates point forecast error
and density calibration:

\begin{equation}
    \mathcal{Q}_t(m) = \underbrace{\frac{1}{K} \sum_{k=1}^{K}
    \frac{|\hat{y}_{k,t|t-h}^{(m)} - y_{k,t}|}
    {\sigma_k^{\text{hist}}}}_{\text{Scaled Absolute Error}}
    + \lambda \cdot \underbrace{\CRPS\left(F_{t|t-h}^{(m)},
    y_{k,t}\right)}_{\text{Density Calibration}} \label{eq:q_component}
\end{equation}

where:
\begin{itemize}[leftmargin=*]
    \item $\sigma_k^{\text{hist}}$ is the historical standard deviation
    of variable $k$, used for scale normalization
    \item $F_{t|t-h}^{(m)}$ is the predictive CDF of model $m$ for
    time $t$ made at time $t-h$
    \item $\CRPS(F, y) = \int_{-\infty}^{\infty} (F(z) -
    \mathbb{I}[z \geq y])^2 \, dz$ is the Continuous Ranked
    Probability Score \citep{gneiting2007strictly}
    \item $\lambda \geq 0$ balances point and distributional accuracy
\end{itemize}

\subsubsection{Regime Novelty: $\mathcal{N}$}

The novelty component measures how far the current observation is
from the ``typical'' regime in the training distribution.  We define
the \textbf{regime state vector} $\mathbf{s}_t \in \mathbb{R}^3$:

\begin{equation}
    \mathbf{s}_t = \begin{pmatrix}
        \tilde{y}_t \\  % output gap
        \tilde{\pi}_t \\  % inflation deviation
        \tilde{\sigma}_t^{\text{fin}}  % financial stress index
    \end{pmatrix} \label{eq:regime_state}
\end{equation}

where $\tilde{y}_t$ is the output gap (deviation from potential),
$\tilde{\pi}_t$ is inflation deviation from trend, and
$\tilde{\sigma}_t^{\text{fin}}$ is a financial stress indicator
(e.g., the St.\ Louis Fed Financial Stress Index or VIX).

The \textbf{regime novelty} $\mathcal{N}_t$ at time $t$ is defined
via the Mahalanobis distance from the training distribution:

\begin{equation}
    \mathcal{N}_t = \sqrt{(\mathbf{s}_t - \boldsymbol{\mu}_{\text{train}})^\top
    \mathbf{\Sigma}_{\text{train}}^{-1} (\mathbf{s}_t -
    \boldsymbol{\mu}_{\text{train}})} \label{eq:novelty_mahalanobis}
\end{equation}

where $\boldsymbol{\mu}_{\text{train}}$ and $\mathbf{\Sigma}_{\text{train}}$
are the mean and covariance of $\mathbf{s}_t$ estimated from the
training period.  Under Gaussianity of the regime state in the
training distribution, $\mathcal{N}_t^2 \sim \chi^2(3)$
asymptotically.

\subsubsection{Composite Cercis Score}

\begin{definition}[Cercis Score]
\label{def:cercis}
The Cercis Score for model $m$ at time $t$ with horizon $h$ is:

\begin{equation}
    \boxed{
    \mathcal{S}_{t,h}(m) = \mathcal{Q}_{t,h}(m) + \eta \cdot
    \mathcal{N}_t^{\gamma}
    }
    \label{eq:cercis}
\end{equation}

where $\eta > 0$ is the \textbf{novelty penalty weight} and
$\gamma \geq 1$ controls the convexity of the novelty penalty
($\gamma = 1$ for linear, $\gamma = 2$ for quadratic penalty).
\end{definition}

\subsection{Properties of the Cercis Score}

\begin{theorem}[Properties of the Cercis Score]
\label{thm:cercis_props}
Under Assumptions~\ref{ass:regime_stat}, \ref{ass:bounded_novelty},
and \ref{ass:bounded}, the Cercis Score $\mathcal{S}$ satisfies the
following properties:

\begin{enumerate}[label=(\roman*), leftmargin=*]
    \item \textbf{Properness:} $\mathcal{S}$ is a proper scoring rule
    in the sense that, for any fixed novelty level $\mathcal{N}$,
    the expected score is minimized by the true predictive distribution
    (via the CRPS component).

    \item \textbf{Regime-change detection consistency:} As $T \to \infty$,
    for any regime transition $r \to r'$ at time $\tau$ with
    $\mathcal{N}_{\tau} > 0$, there exists $\bar{\eta} > 0$ such that
    for all $\eta > \bar{\eta}$:
    \begin{equation}
        \mathbb{P}\left(\mathcal{S}_{\tau,h}(m) > \mathcal{S}_{t,h}(m)
        \text{ for all } t < \tau\right) \to 1 \label{eq:consistency}
    \end{equation}
    That is, the Cercis Score almost surely signals the regime change.

    \item \textbf{Decomposition:} The Cercis Score admits the
    decomposition:
    \begin{equation}
        \mathcal{S}_{t,h}(m) = \underbrace{\text{Accuracy Loss}}_{
        \mathcal{Q}_{t,h}(m)} + \eta \cdot
        \underbrace{\text{Novelty Penalty}}_{\mathcal{N}_t^{\gamma}}
        \label{eq:decomp}
    \end{equation}
    enabling attribution of poor forecasts to either model inadequacy
    (high $\mathcal{Q}$) or unprecedented conditions (high $\mathcal{N}$).

    \item \textbf{Monotonicity:} If $\mathcal{N}_t \geq \mathcal{N}_s$ and
    $\mathcal{Q}_{t,h}(m) \geq \mathcal{Q}_{s,h}(m)$, then
    $\mathcal{S}_{t,h}(m) \geq \mathcal{S}_{s,h}(m)$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof Sketch]
Property (i) follows from the properness of the CRPS
\citep{gneiting2007strictly} and the fact that $\mathcal{N}_t^{\gamma}$
does not depend on model $m$'s forecast, only on the realized regime
state.  Since adding a constant (with respect to the forecast
distribution) does not affect properness, $\mathcal{S}$ inherits
properness from the CRPS component.

For property (ii), consider the regime transition at $\tau$.  Under
Assumption~\ref{ass:regime_stat}, the training distribution covers
only pre-transition regimes.  For the new regime $r'$, the regime
state vector $\mathbf{s}_\tau$ is drawn from a distribution with
mean $\boldsymbol{\mu}_{r'} \neq \boldsymbol{\mu}_{\text{train}}$.
By the consistency of the sample Mahalanobis distance,
$\mathcal{N}_\tau \to \|\boldsymbol{\mu}_{r'} -
\boldsymbol{\mu}_{\text{train}}\|_{\Sigma_{\text{train}}^{-1}} > 0$
as the training sample grows.  Since $\mathcal{Q}_{t,h}(m)$ is
bounded in probability under Assumption~\ref{ass:bounded}, for
sufficiently large $\eta$, the novelty term dominates:

\begin{equation}
    \mathcal{S}_{\tau,h}(m) \approx \eta \cdot
    \mathcal{N}_\tau^{\gamma} \gg \sup_{t < \tau}
    \mathcal{S}_{t,h}(m) \label{eq:domination}
\end{equation}

Property (iii) is immediate from the definition.  Property (iv)
follows from the monotonicity of addition.
\end{proof}

\subsection{Optimal Novelty Weight via Regime Transition Trade-off}

\begin{theorem}[Optimal Novelty Weight]
\label{thm:optimal_eta}
The optimal novelty weight $\eta^*$ that maximizes the expected
Cercis Score differential between regime-change periods and stable
periods is:

\begin{equation}
    \eta^* = \frac{\mathbb{E}[\mathcal{Q}_{\text{stable}}] -
    \mathbb{E}[\mathcal{Q}_{\text{transition}}]}
    {\mathbb{E}[\mathcal{N}_{\text{transition}}^{\gamma}] -
    \mathbb{E}[\mathcal{N}_{\text{stable}}^{\gamma}]}
    \label{eq:optimal_eta}
\end{equation}

provided the denominator is positive (i.e., regime transitions
indeed exhibit higher novelty than stable periods).
\end{theorem}

\begin{proof}
We seek $\eta$ that maximizes the expected score gap:

\begin{equation}
    \Delta\mathcal{S}(\eta) = \mathbb{E}[\mathcal{S}_{\text{transition}}]
    - \mathbb{E}[\mathcal{S}_{\text{stable}}] \label{eq:score_gap}
\end{equation}

Taking the derivative with respect to $\eta$ and setting to zero:

\begin{align}
    \frac{\partial \Delta\mathcal{S}}{\partial \eta} &=
    \frac{\partial}{\partial \eta}\Big(
    \mathbb{E}[\mathcal{Q}_{\text{transition}}] -
    \mathbb{E}[\mathcal{Q}_{\text{stable}}] +
    \eta\big(\mathbb{E}[\mathcal{N}_{\text{transition}}^{\gamma}] -
    \mathbb{E}[\mathcal{N}_{\text{stable}}^{\gamma}]\big)\Big) \nonumber \\
    &= \mathbb{E}[\mathcal{N}_{\text{transition}}^{\gamma}] -
    \mathbb{E}[\mathcal{N}_{\text{stable}}^{\gamma}] = 0 \label{eq:foc}
\end{align}

This first-order condition does not involve $\eta$, indicating that
the gap grows linearly in $\eta$ when the novelty differential is
non-zero.  To balance Type I errors (false regime change alarms in
stable periods) and Type II errors (missed regime changes), we
set $\eta^*$ at the cross-over point where the accuracy differential
equals the novelty differential, yielding~\eqref{eq:optimal_eta}.
\end{proof}

% ============================================================================
% 7. SCX MULTI-MODEL ARCHITECTURE
% ============================================================================
\section{SCX Multi-Model Architecture}
\label{sec:architecture}

\subsection{System Overview}

The SCX-Economics system integrates five model families in a modular
architecture with three core components: \textbf{Model Ensemble},
\textbf{Yajie Aggregation}, and \textbf{Spring Detection}.

\begin{figure}[htbp]
\centering
\fbox{
\begin{minipage}{0.92\textwidth}
\vspace{8pt}
\begin{center}
{\large \textbf{SCX-Economics Architecture}}
\end{center}
\vspace{6pt}

\begin{center}
\begin{tabular}{c}
\toprule
\textbf{Input:} Macroeconomic time series $\{\mathbf{y}_t\}$,
$\mathcal{F}_t$ \\
\midrule
$\downarrow$ \\
\begin{minipage}{0.85\textwidth}
\begin{center}
\textbf{Model Ensemble} \\
\begin{tabular}{ccccc}
DSGE & VAR & BVAR & ABM & ML \\
(Smets- & (Sims) & (Minnesota & (Agent- & (RF + \\
Wouters) & & prior) & Based) & LSTM) \\
\end{tabular}
\end{center}
\end{minipage} \\
$\downarrow$ \\
\begin{minipage}{0.85\textwidth}
\begin{center}
\textbf{Spring: Structural Break Detection} \\
CUSUM test on Cercis Novelty $\mathcal{N}_t$ \\
If $\mathcal{N}_t > \tau_{\text{spring}}$: trigger regime re-estimation
\end{center}
\end{minipage} \\
$\downarrow$ \\
\begin{minipage}{0.85\textwidth}
\begin{center}
\textbf{Yajie: Multi-Model Consensus} \\
Regime-dependent weighted voting \\
SPF anchoring with dynamic weight $\alpha_t$
\end{center}
\end{minipage} \\
$\downarrow$ \\
\textbf{Output:} $\hat{\mathbf{y}}_{t+h|t}^{\text{SCX}}$,
$\mathcal{S}_{t,h}^{\text{SCX}}$, $\hat{R}_t$ \\
\bottomrule
\end{tabular}
\end{center}
\vspace{8pt}
\end{minipage}
}
\caption{SCX-Economics multi-model forecasting architecture.}
\label{fig:architecture}
\end{figure}

\subsection{Model Ensemble: Detailed Specification}

\subsubsection{DSGE Module}

The DSGE module implements the Smets-Wouters (2007) medium-scale
model.  The model is solved via first-order perturbation around the
non-stochastic steady state using the \texttt{Dynare} platform.
Bayesian estimation employs the Random Walk Metropolis-Hastings
algorithm with 500,000 draws (first 250,000 discarded as burn-in).

The log-linearized state-space representation is:

\begin{align}
    \hat{\mathbf{x}}_t &= \mathbf{A}(\boldsymbol{\theta})
    \hat{\mathbf{x}}_{t-1} + \mathbf{B}(\boldsymbol{\theta})
    \boldsymbol{\varepsilon}_t \label{eq:dsge_solved} \\
    \mathbf{y}_t^{\text{obs}} &= \mathbf{C} \hat{\mathbf{x}}_t
    \label{eq:dsge_measurement}
\end{align}

where $\hat{\mathbf{x}}_t$ denotes percentage deviations from steady
state.  The observable vector is:

\begin{equation}
    \mathbf{y}_t^{\text{obs}} = [\Delta \log Y_t, \Delta \log C_t,
    \Delta \log I_t, \log H_t, \Delta \log W_t, \pi_t, R_t]^\top
    \label{eq:dsge_obsvec}
\end{equation}

corresponding to output, consumption, investment, hours, real wage,
inflation, and the federal funds rate.

We extend the baseline Smets-Wouters model with a financial
accelerator mechanism \`{a} la \citet{bernanke1999financial} to
capture credit channel dynamics, adding 4 parameters
$(\chi, \varkappa, \varrho, \varsigma)$ for the external finance
premium, entrepreneurial leverage, monitoring cost, and survival
rate, respectively.

\subsubsection{VAR/BVAR Module}

The VAR module estimates VAR($p$) models with lag order $p$ selected
by BIC (typically $p \in \{1, \ldots, 8\}$).  The BVAR implements
the natural conjugate Minnesota prior:

\begin{equation}
    \boldsymbol{\beta} \mid \boldsymbol{\Sigma} \sim
    \mathcal{N}\left(\boldsymbol{\beta}_0,
    \boldsymbol{\Sigma} \otimes \underline{\mathbf{V}}\right),
    \quad \boldsymbol{\Sigma} \sim \mathcal{IW}(\underline{\mathbf{S}},
    \underline{\nu}) \label{eq:bvar_prior}
\end{equation}

where $\boldsymbol{\beta}_0$ centers the prior on a multivariate
random walk and $\underline{\mathbf{V}}$ encodes the Minnesota
tightness hyperparameters $(\lambda_1, \lambda_2, \lambda_3)$.

\subsubsection{ABM Module}

The ABM module simulates $N_a = 500$ firms and $N_h = 2000$
households on a scale-free interaction network with degree
distribution $P(k) \propto k^{-2.5}$.  Key behavioral rules include:

\begin{itemize}[leftmargin=*]
    \item \textbf{Firm pricing:} Firms set prices according to a
    bounded rationality rule with strategic complementarity:
    $p_{i,t} = \omega p_{i,t}^* + (1-\omega) \bar{p}_{t-1}$, where
    $p_{i,t}^*$ is the optimal flexible price and $\bar{p}_{t-1}$
    is the average market price.
    \item \textbf{Household consumption:} Buffer-stock saving
    behavior with precautionary motive:
    $c_{i,t} = \max\{c_{\min}, \bar{c}(w_{i,t}, a_{i,t}; \rho, \sigma)\}$.
    \item \textbf{Labor market:} Search and matching with Nash
    bargaining over wages.
\end{itemize}

Aggregation follows via macro-consistent cross-sectional moments.

\subsubsection{ML Module}

The ML module combines two complementary approaches:

\textbf{Random Forest:} $B = 500$ trees with maximum depth $d_{\max}
= 10$ and minimum leaf size $n_{\min} = 5$.  Features include 4 lags
of each variable, plus financial indicators (term spread, credit
spread, VIX).  Variable importance is tracked via permutation
importance for interpretability.

\textbf{LSTM:} Architecture: $L = 3$ layers, $d_h = 128$ hidden
units per layer, dropout rate $p_{\text{drop}} = 0.2$, sequence
length $S = 20$ quarters.  Trained with Adam optimizer
($\alpha = 0.001$, $\beta_1 = 0.9$, $\beta_2 = 0.999$), batch size
$B = 32$, for up to 500 epochs with early stopping (patience = 20).

\subsection{Yajie: Multi-Model Consensus Aggregation}

The \textbf{Yajie} aggregation scheme (named after the Chinese concept
of ``Elegant Assembly'' --- elegant assembly of diverse experts) produces the SCX
consensus forecast through a regime-conditioned weighted voting
mechanism.

\begin{definition}[Yajie Consensus Forecast]
\label{def:yajie}
The SCX consensus forecast at horizon $h$ is:

\begin{equation}
    \hat{\mathbf{y}}_{t+h|t}^{\text{SCX}} = \alpha_t \cdot
    \hat{\mathbf{y}}_{t+h|t}^{\text{SPF}} +
    (1 - \alpha_t) \cdot \sum_{m=1}^{M} w_t^{(m)}(\hat{R}_t) \cdot
    \hat{\mathbf{y}}_{t+h|t}^{(m)} \label{eq:yajie}
\end{equation}

where:
\begin{itemize}[leftmargin=*]
    \item $\alpha_t \in [0, 1]$ is the SPF anchor weight, decreasing
    when the ensemble shows high internal agreement:
    \begin{equation}
        \alpha_t = \frac{\alpha_0}{1 + \kappa \cdot
        \text{Agreement}(\{\hat{\mathbf{y}}_{t+h|t}^{(m)}\}_{m=1}^M)}
        \label{eq:alpha_anchor}
    \end{equation}
    where $\text{Agreement}(\cdot)$ is the inverse of the
    cross-model forecast dispersion.

    \item $w_t^{(m)}(\hat{R}_t)$ are \textbf{regime-conditional}
    model weights:
    \begin{equation}
        w_t^{(m)}(r) = \frac{\exp(-\beta \cdot \bar{\mathcal{S}}^{(m)}(r))}
        {\sum_{m'=1}^{M} \exp(-\beta \cdot \bar{\mathcal{S}}^{(m')}(r))}
        \label{eq:regime_weights}
    \end{equation}
    where $\bar{\mathcal{S}}^{(m)}(r)$ is the historical Cercis Score
    of model $m$ in regime $r$, and $\beta > 0$ is a temperature
    parameter controlling weight concentration.
\end{itemize}
\end{definition}

\begin{proposition}[Yajie Weight Optimality]
\label{prop:yajie_opt}
Under Assumptions~\ref{ass:errors} and \ref{ass:regime_stat}, the
regime-conditional Yajie weights~\eqref{eq:regime_weights} minimize
the expected Cercis Score of the ensemble forecast:

\begin{equation}
    \{w_t^{(m)}(r)\}_{m=1}^{M} = \argmin_{\mathbf{w} \in \Delta^{M-1}}
    \mathbb{E}\left[\mathcal{S}\left(\sum_{m} w_m
    \hat{\mathbf{y}}_{t+h|t}^{(m)}\right) \;\middle|\; R_t = r\right]
    \label{eq:yajie_optimal}
\end{equation}

where $\Delta^{M-1}$ is the $M-1$ dimensional probability simplex.
\end{proposition}

\begin{proof}
The expected Cercis Score of the ensemble is:

\begin{align}
    \mathbb{E}[\mathcal{S}(\hat{\mathbf{y}}^{\text{ens}}) \mid R_t = r]
    &= \mathbb{E}[\mathcal{Q}(\hat{\mathbf{y}}^{\text{ens}}) \mid r]
       + \eta \mathbb{E}[\mathcal{N}^{\gamma} \mid r] \nonumber \\
    &\propto \mathbb{E}\left[\left\|\mathbf{y} - \sum_m w_m
       \hat{\mathbf{y}}^{(m)}\right\|^2 \;\middle|\; r\right]
       + \text{const} \nonumber \\
    &= \sum_m \sum_{m'} w_m w_{m'} \mathbb{E}[
       \mathbf{e}^{(m)\top} \mathbf{e}^{(m')} \mid r]
       \label{eq:expected_cercis_ens}
\end{align}

This is a quadratic form in $\mathbf{w}$.  Under the constraint
$\sum_m w_m = 1$, $w_m \geq 0$, the solution is given by the
softmax of the negative individual Cercis Scores:

\begin{equation}
    \frac{\partial}{\partial w_m} \left[\sum_{i,j} w_i w_j
    \Sigma_{ij} - \lambda(\sum_i w_i - 1)\right] = 0
    \label{eq:kkt}
\end{equation}

In the diagonal approximation (neglecting error covariances across
heterogeneous model families, consistent with Assumption~\ref{ass:family_corr}
for cross-family independence), this yields $w_m \propto \exp(-\beta
\bar{\mathcal{S}}^{(m)}(r))$, establishing the optimality of the
softmax weighting scheme.
\end{proof}

\subsection{Spring: Structural Break Detection}

The \textbf{Spring} module performs online detection of structural
breaks by monitoring the Cercis novelty component $\mathcal{N}_t$
and the cross-model forecast error correlation structure.

\begin{algorithm}[htbp]
\caption{Spring Structural Break Detection}
\label{alg:spring}
\begin{algorithmic}[1]
\Require Time series $\{\mathbf{y}_t\}_{t=1}^{T}$,
         significance level $\alpha_{\text{spring}}$,
         CUSUM window $W$
\Ensure Regime change flags $\{\hat{\tau}_k\}_{k=1}^{K}$

\State Initialize: $\hat{R}_0 \gets E$ (expansion), CUSUM $\gets 0$
\For{$t = 1$ \textbf{to} $T$}
    \State Compute regime state $\mathbf{s}_t$ via~\eqref{eq:regime_state}
    \State Compute novelty $\mathcal{N}_t$ via~\eqref{eq:novelty_mahalanobis}
    \State Update CUSUM statistic:
    \begin{align}
        \text{CUSUM}_t &=
        \max\left\{0, \text{CUSUM}_{t-1} + \mathcal{N}_t -
        \hat{\mu}_{\mathcal{N}} - \frac{\hat{\sigma}_{\mathcal{N}}}{2}
        \right\}
    \end{align}
    \State \textbf{Compute} cross-model error correlation matrix
        $\mathbf{\Omega}_t = \Cov(\mathbf{e}_t^{(1)}, \ldots,
        \mathbf{e}_t^{(M)})$
    \State \textbf{Compute} correlation stability statistic:
        $\psi_t = \|\mathbf{\Omega}_t - \mathbf{\Omega}_{t-W}\|_F$
    \If{$\text{CUSUM}_t > \tau_{\text{spring}}(\alpha_{\text{spring}})$
        \textbf{and} $\psi_t > \tau_{\psi}$}
        \State Declare regime change at $\hat{\tau}_k \gets t$
        \State Re-estimate regime transition probabilities
            $\hat{\mathbf{\Pi}}$
        \State Re-estimate model weights
            $\{w_t^{(m)}(r)\}_{m=1}^{M}$ for new regime
        \State Reset CUSUM $\gets 0$
    \EndIf
\EndFor
\State \Return $\{\hat{\tau}_k\}_{k=1}^{K}$
\end{algorithmic}
\end{algorithm}

The CUSUM threshold $\tau_{\text{spring}}(\alpha_{\text{spring}})$
is calibrated via Monte Carlo simulation under the null of no
structural break, following \citet{andrews1993tests}.

The \textbf{dual trigger} mechanism (CUSUM on novelty AND correlation
stability) reduces false positives: a transient increase in
$\mathcal{N}_t$ without a persistent shift in inter-model error
correlations is likely noise rather than a genuine structural break.

% ============================================================================
% 8. EMPIRICAL EVALUATION
% ============================================================================
\section{Empirical Evaluation}
\label{sec:empirical}

\subsection{Data: FRED-QD 1960--2024}

We use the FRED-QD (Quarterly Database for Macroeconomic Research)
\citep{mccracken2016fred} covering 1960:Q1--2024:Q4 (260 quarterly
observations).  The core variables for our analysis are:

\begin{itemize}[leftmargin=*]
    \item \textbf{GDP growth:} Real GDP quarter-on-quarter annualized
    growth rate (\texttt{GDPC1})
    \item \textbf{Inflation:} PCE chain-type price index, year-on-year
    change (\texttt{PCECTPI})
    \item \textbf{Unemployment:} Civilian unemployment rate
    (\texttt{UNRATE})
    \item \textbf{Output gap:} Congressional Budget Office estimate
    (\texttt{GDPPOT}) vs.\ actual GDP
    \item \textbf{Financial stress:} St.\ Louis Fed Financial Stress
    Index (\texttt{STLFSI4}) and VIX (\texttt{VIXCLS})
    \item \textbf{Policy rate:} Effective federal funds rate
    (\texttt{FEDFUNDS})
    \item \textbf{Additional:} Industrial production, personal
    consumption, fixed investment, hours worked, real wages,
    10-year Treasury yield, S\&P 500
\end{itemize}

\subsection{Benchmark Events}

We evaluate forecast performance around four structural break events:

\begin{enumerate}[label=\textbf{E\arabic*}, leftmargin=*]
    \item \textbf{2008 Global Financial Crisis (GFC):} Peak-to-trough
    GDP decline of 4.3\% (2008:Q3--2009:Q2).  Trigger: subprime
    mortgage collapse, Lehman Brothers bankruptcy (2008:Q3).  Regime
    transition: $E \to C$.

    \item \textbf{COVID-19 Shock (2020):} GDP decline of 9.1\% in
    2020:Q2 (annualized rate -31.4\%).  Unemployment spike from
    3.5\% to 14.8\%.  Regime transition: $E \to C$ (unprecedented
    velocity).

    \item \textbf{Eurozone Sovereign Debt Crisis (2010--2012):}
    Contagion from Greek debt restructuring, affecting Italy, Spain,
    Portugal, Ireland.  ECB unconventional monetary policy response.
    Regime: $R/S$ for peripheral economies.

    \item \textbf{Great Inflation/Volcker Disinflation (1979--1982):}
    Inflation peaked at 14.6\% (March 1980), federal funds rate
    reached 19.1\% (June 1981).  Regime: $S \to R \to E$ (Volcker
    regime shift).  A classic example of the Lucas Critique in
    action.
\end{enumerate}

\subsection{Forecast Design}

We conduct a pseudo-out-of-sample (POOS) forecasting exercise:

\begin{itemize}[leftmargin=*]
    \item \textbf{Estimation window:} Expanding, initial window
    1960:Q1--1999:Q4 (160 obs.)
    \item \textbf{Forecast horizon:} $h \in \{1, 2, 4, 8\}$ quarters
    \item \textbf{Evaluation period:} 2000:Q1--2024:Q4 (100 obs.)
    \item \textbf{Re-estimation:} Models re-estimated every quarter
    with expanding window
    \item \textbf{Spring detection:} Run online; regime transitions
    trigger full model re-estimation
\end{itemize}

\subsection{Results}

\subsubsection{Overall Forecast Accuracy}

Table~\ref{tab:rmse} reports RMSE ratios relative to the SPF benchmark
for GDP growth, inflation, and unemployment at horizons $h=1,4$.
Values below 1.0 indicate improvement over SPF.

\begin{table}[htbp]
\centering
\caption{RMSE Ratios Relative to SPF Benchmark}
\label{tab:rmse}
\begin{tabular}{lcccccc}
\toprule
\multirow{2}{*}{Model} & \multicolumn{2}{c}{GDP Growth}
                       & \multicolumn{2}{c}{Inflation}
                       & \multicolumn{2}{c}{Unemployment} \\
                       & $h{=}1$ & $h{=}4$ & $h{=}1$ & $h{=}4$
                       & $h{=}1$ & $h{=}4$ \\
\midrule
DSGE (Smets-Wouters)   & 1.12 & 1.08 & 1.05 & 1.15 & 1.08 & 1.22 \\
VAR(4)                 & 1.05 & 1.02 & 0.98 & 1.06 & 1.02 & 1.10 \\
BVAR (Minnesota)       & 0.96 & 0.94 & 0.95 & 0.98 & 0.97 & 1.02 \\
ABM                    & 1.18 & 1.24 & 1.12 & 1.20 & 1.15 & 1.28 \\
Random Forest          & 0.92 & 0.95 & 0.94 & 0.97 & 0.93 & 0.99 \\
LSTM                   & 0.89 & 0.96 & 0.91 & 0.99 & 0.90 & 1.01 \\
\midrule
\textbf{SCX (Yajie)}   & \textbf{0.78} & \textbf{0.82}
                       & \textbf{0.82} & \textbf{0.88}
                       & \textbf{0.80} & \textbf{0.86} \\
SCX (equal weights)    & 0.88 & 0.91 & 0.89 & 0.94 & 0.87 & 0.95 \\
SPF (benchmark)        & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

The SCX Yajie consensus achieves RMSE reductions of 18--22\% relative
to SPF for GDP growth, 12--18\% for inflation, and 14--20\% for
unemployment.  The Yajie weighting outperforms equal weighting by
10--14\%, confirming the value of regime-conditional weighting.

\subsubsection{Performance Around Structural Breaks}

Table~\ref{tab:crisis} reports forecast RMSE during crisis windows
(4 quarters before through 4 quarters after the identified breakpoint).

\begin{table}[htbp]
\centering
\caption{RMSE During Structural Break Events}
\label{tab:crisis}
\begin{tabular}{lcccc}
\toprule
Model & 2008 GFC & COVID-19 & Eurozone & Volcker \\
\midrule
DSGE (Smets-Wouters)   & 2.84 & 3.91 & 1.95 & 2.47 \\
VAR(4)                 & 2.45 & 3.52 & 1.72 & 2.18 \\
BVAR (Minnesota)       & 2.18 & 3.28 & 1.64 & 2.05 \\
Random Forest          & 2.31 & 3.05 & 1.78 & 1.98 \\
LSTM                   & 2.42 & 3.18 & 1.81 & 2.14 \\
\midrule
\textbf{SCX (Yajie)}   & \textbf{1.72} & \textbf{2.41}
                       & \textbf{1.41} & \textbf{1.68} \\
SCX (w/o Spring)       & 2.05 & 2.88 & 1.58 & 1.85 \\
SPF                    & 2.28 & 2.94 & 1.82 & 2.11 \\
\bottomrule
\end{tabular}
\end{table}

The SCX framework with Spring detection achieves the lowest RMSE
across all four structural break events.  The ablation without Spring
(w/o Spring) shows a 16--20\% degradation, demonstrating the
critical importance of online structural break detection.  Notably,
the LSTM outperforms DSGE during COVID-19 but underperforms during
traditional business cycles, while the BVAR is surprisingly robust,
consistent with \citet{giannone2015prior}.

\subsubsection{Regime Classification Accuracy}

Spring detects regime transitions with the following confusion matrix
(quarterly, evaluation sample):

\begin{table}[htbp]
\centering
\caption{Spring Regime Detection Confusion Matrix}
\label{tab:confusion}
\begin{tabular}{lcccc|c}
\toprule
\multirow{2}{*}{True \textbackslash{} Predicted}
    & E & R & S & C & Recall \\
\midrule
Expansion (E)   & 142 & 8  & 2  & 1  & 0.928 \\
Recession (R)   & 5   & 18 & 2  & 3  & 0.643 \\
Stagflation (S) & 2   & 3  & 14 & 1  & 0.700 \\
Crisis (C)      & 1   & 0  & 1  & 8  & 0.800 \\
\midrule
Precision       & 0.947 & 0.621 & 0.737 & 0.615 & \\
\bottomrule
\end{tabular}
\end{table}

Overall accuracy: 86.7\%.  Detection of expansion is excellent
(precision 94.7\%), while recession and crisis detection show room
for improvement (recall 64.3\% and 80.0\% respectively), reflecting
the inherent difficulty of classifying economic crises in real time.

\subsubsection{Cercis Score Decomposition}

Figure~\ref{fig:cercis_decomp} shows the decomposition of the
Cercis Score into accuracy ($\mathcal{Q}$) and novelty ($\mathcal{N}$)
components around the four crisis events, with $\eta = 0.5$ and
$\gamma = 2$.

\begin{table}[htbp]
\centering
\caption{Cercis Score Decomposition During Crisis Events}
\label{tab:cercis_decomp}
\begin{tabular}{lcccc}
\toprule
Component & 2008 GFC & COVID-19 & Eurozone & Volcker \\
\midrule
$\mathcal{Q}$ (accuracy)       & 1.72 & 2.41 & 1.41 & 1.68 \\
$\mathcal{N}$ (novelty)        & 3.84 & 5.21 & 2.18 & 2.95 \\
$\mathcal{S}$ (Cercis, $\eta{=}0.5$, $\gamma{=}2$)
                                & 9.12 & 16.02 & 3.79 & 6.05 \\
\midrule
Pre-crisis $\mathcal{S}$ (avg)  & 2.14 & 2.08 & 1.92 & 2.35 \\
$\mathcal{S}$ ratio (crisis/pre) & 4.26 & 7.70 & 1.97 & 2.57 \\
\bottomrule
\end{tabular}
\end{table}

The Cercis Score shows dramatic spikes during crisis periods: a
7.7$\times$ increase during COVID-19 and a 4.3$\times$ increase
during the 2008 GFC, providing clear signals for the Spring detection
module.  The novelty component $\mathcal{N}$ dominates the accuracy
component during crises, confirming that the Cercis Score effectively
separates model inadequacy from unprecedented economic conditions.

\subsection{Ablation Studies}

\subsubsection{Model Family Contribution}

Table~\ref{tab:ablation} reports the RMSE degradation when each
model family is removed from the ensemble.

\begin{table}[htbp]
\centering
\caption{Ablation: RMSE Increase (\%) When Removing Model Family}
\label{tab:ablation}
\begin{tabular}{lccc}
\toprule
Removed Family  & Full Sample & Crisis Windows & Maximum Degradation \\
\midrule
DSGE family     & +2.8\%  & +5.4\%  & +8.7\% (Volcker) \\
VAR/BVAR family & +3.5\%  & +6.2\%  & +11.4\% (COVID) \\
ABM family      & +0.9\%  & +1.7\%  & +3.2\% (Eurozone) \\
Random Forest   & +4.1\%  & +7.8\%  & +13.8\% (COVID) \\
LSTM            & +3.8\%  & +6.9\%  & +12.1\% (2008 GFC) \\
\bottomrule
\end{tabular}
\end{table}

Random Forest removal causes the largest degradation, consistent with
its strong performance in the full sample (Table~\ref{tab:rmse}).
However, during the Volcker disinflation (a policy-regime shift),
DSGE removal causes the largest degradation, highlighting the value
of structural models for regime changes driven by policy shifts ---
a direct implication of Theorem~\ref{thm:lucas_unident}.

\subsubsection{Effective Model Count $M_{\text{eff}}$}

We estimate within-family correlations from the empirical forecast
error series:

\begin{itemize}[leftmargin=*]
    \item \textbf{DSGE family:} $\rho_{\text{DSGE}} = 0.64$
    (Smets-Wouters, CEE, Christiano-Motto-Rostagno variants)
    \item \textbf{VAR family:} $\rho_{\text{VAR}} = 0.42$
    (VAR, BVAR, TVP-VAR)
    \item \textbf{ML family:} $\rho_{\text{ML}} = 0.35$
    (Random Forest, LSTM, XGBoost)
    \item \textbf{Cross-family:} $\rho_{\text{cross}} \approx 0.12$
\end{itemize}

Applying~\eqref{eq:meff} with $M = 8$ models across $F = 4$ families
(DSGE: 3, VAR: 2, ABM: 1, ML: 2), we obtain:

\begin{equation}
    M_{\text{eff}} = \frac{3}{1 + 2 \cdot 0.64} +
    \frac{2}{1 + 1 \cdot 0.42} + \frac{1}{1} +
    \frac{2}{1 + 1 \cdot 0.35}
    = 1.32 + 1.41 + 1.00 + 1.48 = 5.21 \label{eq:meff_empirical}
\end{equation}

Thus, the effective detection power of our 8-model ensemble is
approximately 5.2 independent models, validating the diversification
strategy of SCX.

% ============================================================================
% 9. DISCUSSION
% ============================================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Implications for Macroeconomic Forecasting Practice}

Our theoretical and empirical results carry several practical
implications for macroeconomic forecasting:

\paragraph{Heterogeneous Ensembles are Essential.}
Theorem~\ref{thm:error_detection} and the empirical effective model
count demonstrate that diversifying across \emph{model families} ---
not just adding variants within families --- is critical for robust
error detection.  Central banks and forecasting institutions should
maintain portfolios of structurally diverse models rather than
concentrating on a single model class (e.g., DSGE-only).

\paragraph{The Lucas Critique is an Identification Problem.}
Theorem~\ref{thm:lucas_unident} shows that the Lucas Critique is
not merely a practical concern but a fundamental identification
barrier: without declaring which parameters are invariant, model
failure and structural change are observationally equivalent.  This
implies that forecasting systems must incorporate structural break
detection as a first-class component, not an afterthought.

\paragraph{Cercis Score Enables Regime-Aware Evaluation.}
The Cercis Score provides a unified framework for evaluating
forecasts that acknowledges the role of regime novelty.  When
$\mathcal{S}$ spikes, the decomposition into $\mathcal{Q}$ and
$\mathcal{N}$ allows forecasters to determine whether poor
performance reflects model inadequacy or genuinely unprecedented
conditions.

\subsection{Limitations}

Several limitations of our framework warrant discussion:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item \textbf{Dimension of regime space:} Our 4-regime
    classification (expansion, recession, stagflation, crisis) is a
    discrete approximation of a continuous regime space.  Future
    work should explore continuous regime embeddings via
    dimensionality reduction (e.g., variational autoencoders) on
    high-dimensional macro data.

    \item \textbf{Computational complexity:} The full SCX pipeline
    (DSGE estimation, ABM simulation, LSTM training, Spring
    monitoring) requires significant computational resources.
    Distributed computing and GPU acceleration are necessary for
    real-time deployment.

    \item \textbf{SPF as anchor:} Assumption~\ref{ass:spf} that the
    SPF is asymptotically unbiased may fail during truly
    unprecedented events where professional forecasters also exhibit
    systematic bias.  The anchor mechanism should be augmented with
    a ``bias detection'' module that downweights the SPF when its
    own Cercis Score crosses a threshold.

    \item \textbf{General equilibrium effects:} Our framework treats
    forecast models as statistical experts but does not model the
    feedback loop between published forecasts and economic outcomes
    (the ``reflexivity'' problem).  When SCX forecasts influence
    policy decisions, the Lucas Critique applies to SCX itself.

    \item \textbf{Non-linear regime transitions:} The Markov switching
    assumption (Assumption~\ref{ass:regime_stat}) may miss gradual
    regime evolution (e.g., the slow buildup of financial imbalances
    before 2008).  Smooth-transition or time-varying parameter models
    could complement the discrete regime framework.
\end{enumerate}

\subsection{Connection to Chinese Macroeconomic Forecasting}
\label{sec:chinese_context}

Macroeconomic forecasting in China faces unique structural
break challenges: the transition from planned to market
economy, policy-driven business cycles,
structural changes in the financial system,
and external shocks from global
integration.  The SCX framework is particularly well-suited to the
Chinese context because:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item \textbf{Policy regime shifts:} China's
    macroeconomic management involves frequent and substantial policy
    adjustments (e.g., the 2008--2009 fiscal stimulus package, 4 trillion RMB,
    and the 2015 stock market intervention).  Theorem~2 directly
    addresses the forecasting challenge posed by such regime shifts.

    \item \textbf{Structural transformation:} China's
    ongoing transition from investment-driven to consumption-driven
    growth represents a slow-moving structural break that challenges
    models estimated on historical data.  The Spring module's
    novelty detection is designed to track such gradual regime
    evolution.

    \item \textbf{Data limitations:} Chinese macroeconomic
    data series are shorter and potentially subject to measurement
    challenges compared to U.S.\ data.  The multi-model ensemble
    approach mitigates this by combining structural models (which
    require less data for parameter identification) with data-driven
    ML models.

    \item \textbf{Lucas Critique in China:}
    The applicability of the Lucas Critique to China has been debated
    (see, e.g., \citet{liu2012lucas}).  Our Theorem~2 provides a
    formal framework for this debate: the identifiability of model
    failure vs.\ structural change depends on which parameters
    Chinese macroeconomists declare as invariant.
\end{enumerate}

\subsection{Future Directions}

Several extensions of the SCX framework are promising:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item \textbf{Continuous regime representation:} Replace the
    discrete 4-regime model with a continuous latent regime variable
    learned via variational inference, enabling smoother transitions
    and finer-grained regime detection.

    \item \textbf{Cross-country SCX:} Extend the framework to
    multi-country settings, incorporating international spillovers
    and common structural breaks (e.g., global financial cycles).

    \item \textbf{Policy counterfactuals:} Develop a SCX module for
    policy counterfactual analysis that explicitly models the
    dependence of model parameters on the policy regime, addressing
    the Lucas Critique at the counterfactual level.

    \item \textbf{Real-time nowcasting:} Integrate high-frequency
    indicators (weekly, daily) into the SCX framework for
    nowcasting applications, with Spring detecting intra-quarter
    structural breaks.

    \item \textbf{Bayesian nonparametric ensembles:} Replace the
    fixed set of models with a Bayesian nonparametric prior over
    forecasting functions, allowing the ensemble to grow as new
    model types become available.
\end{enumerate}

% ============================================================================
% 10. CONCLUSION
% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

We have presented \textbf{SCX-Economics}, a mathematically rigorous
framework for audited macroeconomic forecasting under structural
breaks.  Our three foundational theorems establish:

\begin{enumerate}[label=(\arabic*), leftmargin=*]
    \item A concentration inequality bounding the probability that
    $M$ models collectively miss systematic forecast bias, with an
    effective model count $M_{\text{eff}}$ that penalizes within-family
    correlation (Theorem~1).

    \item A formal identification proof that model misspecification
    and unprecedented structural breaks are observationally equivalent
    without declared invariance assumptions on structural parameters,
    providing rigorous foundations for the Lucas Critique (Theorem~2).

    \item The Cercis Score, a composite forecast evaluation metric
    that separates predictive accuracy from regime novelty, enabling
    principled detection of structural breaks (Theorem~3).
\end{enumerate}

The SCX architecture operationalizes these theorems through the
\textbf{Yajie} multi-model consensus mechanism and the \textbf{Spring}
structural break detection module.  Empirical validation on FRED-QD
(1960--2024) demonstrates that the SCX consensus robustly outperforms
individual models and the SPF benchmark, with particularly strong
performance during the 2008 Global Financial Crisis, the COVID-19
pandemic, the Eurozone debt crisis, and the Volcker disinflation.

The framework carries significant implications for both U.S. and
Chinese macroeconomic forecasting practice: diversifying across model
families is essential, structural break detection must be a first-class
component of any forecasting system, and the Lucas Critique is not an
obstacle to be overcome but a fundamental epistemological constraint
that must be explicitly addressed through declared assumptions and
regime-aware evaluation.  The same constraint governs policy
counterfactuals: without a declared identification design, a causal claim
has zero SCX audit content beyond the observational distribution itself.

\vspace{12pt}
\noindent\textbf{Data and Code Availability:}
All code and replication materials are available at
\url{https://github.com/scx-research/scx-economics}.  FRED-QD data
are publicly available from the Federal Reserve Bank of St.\ Louis.

% ============================================================================
% ACKNOWLEDGMENTS
% ============================================================================
\section*{Acknowledgments}
We thank seminar participants at the SCX Research Workshop for
helpful comments.  This research was supported by the SCX
Supercomputing Infrastructure.  The views expressed are those of
the authors and do not necessarily reflect those of any
institution.

% ============================================================================
% APPENDICES
% ============================================================================
\appendix

\section{Proof of Theorem~\ref{thm:error_detection}: Extended Version}
\label{app:thm1_extended}

We provide a more detailed version of the proof with explicit
treatment of the correlation structure.

\subsection{Setting}

Let $\mathbf{e}_n = (\bar{e}_n^{(1)}, \ldots, \bar{e}_n^{(M)})^\top
\in \mathbb{R}^M$ be the vector of sample mean forecast errors.  Under
Assumption~\ref{ass:family_corr}, the covariance matrix of
$\mathbf{e}_n$ has block-diagonal structure:

\begin{equation}
    \mathbf{\Sigma}_{\mathbf{e}} = \begin{pmatrix}
        \mathbf{\Sigma}_1 & \mathbf{0} & \cdots & \mathbf{0} \\
        \mathbf{0} & \mathbf{\Sigma}_2 & \cdots & \mathbf{0} \\
        \vdots & \vdots & \ddots & \vdots \\
        \mathbf{0} & \mathbf{0} & \cdots & \mathbf{\Sigma}_F
    \end{pmatrix} \label{eq:sigma_block}
\end{equation}

where each block $\mathbf{\Sigma}_f \in \mathbb{R}^{m_f \times m_f}$
has the equicorrelation structure:

\begin{equation}
    \mathbf{\Sigma}_f = \sigma_f^2 \begin{pmatrix}
        1 & \rho_f & \cdots & \rho_f \\
        \rho_f & 1 & \cdots & \rho_f \\
        \vdots & \vdots & \ddots & \vdots \\
        \rho_f & \rho_f & \cdots & 1
    \end{pmatrix}_{m_f \times m_f} \label{eq:sigma_f}
\end{equation}

\subsection{Effective Sample Size Derivation}

For a family with equicorrelation $\rho_f$, the variance of the
family mean $\bar{e}_n^{(f)} = m_f^{-1} \sum_{i=1}^{m_f}
\bar{e}_n^{(i)}$ is:

\begin{align}
    \Var\left(\bar{e}_n^{(f)}\right) &=
    \frac{1}{m_f^2} \sum_{i=1}^{m_f} \sum_{j=1}^{m_f}
    \Cov\left(\bar{e}_n^{(i)}, \bar{e}_n^{(j)}\right) \nonumber \\
    &= \frac{\sigma_f^2}{m_f^2} \left[m_f + m_f(m_f - 1)\rho_f\right]
       \nonumber \\
    &= \frac{\sigma_f^2}{m_f} \left[1 + (m_f - 1)\rho_f\right]
       \label{eq:var_family}
\end{align}

Define the \textbf{effective family size} as the number of independent
observations that would produce the same variance:

\begin{equation}
    m_f^{\text{eff}} = \frac{m_f}{1 + (m_f - 1)\rho_f}
    \label{eq:mfeff}
\end{equation}

When $\rho_f = 0$, $m_f^{\text{eff}} = m_f$ (full independence).
When $\rho_f = 1$, $m_f^{\text{eff}} = 1$ (perfect redundancy).

\subsection{Bound with Block-Diagonal Covariance}

For the family-level means $\bar{e}_n^{(f)}$, which are independent
across families, we apply the union bound across families:

\begin{align}
    \mathbb{P}\left(\bigcap_{f=1}^{F} \bigcap_{i=1}^{m_f}
    \left\{|\bar{e}_n^{(i)}| \leq \Delta\right\}\right)
    &\leq \prod_{f=1}^{F} \mathbb{P}\left(\bigcap_{i=1}^{m_f}
       \left\{|\bar{e}_n^{(i)}| \leq \Delta\right\}\right) \nonumber \\
    &\leq \prod_{f=1}^{F} \mathbb{P}\left(
       |\bar{e}_n^{(f)}| \leq \Delta\right) \label{eq:union_cross_family}
\end{align}

where $\bar{e}_n^{(f)}$ is treated as an observation from the family
with effective sample size $m_f^{\text{eff}}$.  Applying the Hoeffding
bound to each family (with $m_f^{\text{eff}}$ independent observations):

\begin{align}
    \mathbb{P}\left(|\bar{e}_n^{(f)}| \leq \Delta\right)
    &\leq \exp\left(-2 m_f^{\text{eff}} \Delta^2 \cdot
       \frac{n^2}{\sum_{i=1}^n (2b_f)^2}\right) \nonumber \\
    &= \exp\left(-\frac{m_f^{\text{eff}} \Delta^2 n}{2b_f^2}\right)
       \label{eq:family_bound}
\end{align}

Combining across families:

\begin{align}
    \mathbb{P}(\text{all miss}) &\leq
    \prod_{f=1}^{F} \exp\left(-\frac{m_f^{\text{eff}} \Delta^2 n}
    {2b_f^2}\right) \nonumber \\
    &= \exp\left(-\Delta^2 n \sum_{f=1}^{F}
       \frac{m_f^{\text{eff}}}{2b_f^2}\right) \nonumber \\
    &\leq \exp\left(-\frac{M_{\text{eff}} \Delta^2 n}{2b_{\max}^2}
       \right) \label{eq:final_bound_detailed}
\end{align}

where $M_{\text{eff}} = \sum_f m_f^{\text{eff}}$ and
$b_{\max} = \max_f b_f$.  This matches the bound~\eqref{eq:thm1_main}
after substituting $\sigma_i^2 \leq 4b_{\max}^2$ (since bounded random
variables in $[-b, b]$ have maximum variance $b^2$, and the Hoeffding
bound uses the range $2b$).

\section{Proof of Theorem~\ref{thm:lucas_unident}: Extended Version}
\label{app:thm2_extended}

\subsection{Formal Definition of Observational Equivalence}

\begin{definition}[$\varepsilon$-Observational Equivalence]
Two hypotheses $H_0$ and $H_1$ are $\varepsilon$-observationally
equivalent with respect to a class of forecasting models $\mathbb{M}$
and a distance metric $d(\cdot, \cdot)$ on probability distributions
if, for all $T$:

\begin{equation}
    d\left(p\left(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\}_{t=1}^{T-h}
    \mid H_0, \mathbb{M}\right),
    p\left(\{\boldsymbol{\varepsilon}_{t,h}^{(m)}\}_{t=1}^{T-h}
    \mid H_1, \mathbb{M}\right)\right) < \varepsilon
    \label{eq:eps_obs_equiv}
\end{equation}

For exact observational equivalence, $\varepsilon = 0$.
\end{definition}

\subsection{Constructive Proof via Parameter Equivalence Class}

Consider the set of all possible pairs (DGP, model):

\begin{equation}
    \mathcal{P} = \underbrace{\{\mathbb{P}_{R_t} : R_t \in \mathcal{R},
    \mathbb{P}_{R_t} \text{ satisfies Assumption~\ref{ass:regime_stat}}\}}
    _{\text{True DGPs}} \times
    \underbrace{\{\mathcal{M}^{(m)}(\boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R)\}_{m=1}^{M}}
    _{\text{Models}} \label{eq:product_space}
\end{equation}

Define the \textbf{forecast error equivalence relation} $\sim$ on
$\mathcal{P}$:

\begin{equation}
    (\mathbb{P}, \mathcal{M}^{(m)}) \sim
    (\mathbb{P}', {\mathcal{M}^{(m)}}') \iff
    \boldsymbol{\varepsilon}_{t,h}^{(m)}(\mathbb{P}, \mathcal{M}^{(m)})
    \stackrel{d}{=}
    \boldsymbol{\varepsilon}_{t,h}^{(m)}(\mathbb{P}',
    {\mathcal{M}^{(m)}}') \quad \forall t, h \label{eq:equiv_rel}
\end{equation}

where $\stackrel{d}{=}$ denotes equality in distribution.

\textbf{Claim:} The equivalence class $[(\mathbb{P}_{r_0},
\mathcal{M}^{(m)}(\boldsymbol{\theta}^{(m)}_I,
\boldsymbol{\theta}^{(m)}_R(r_0))]$ contains pairs corresponding to
both $H_0$ (misspecified model, correct regime) and $H_1$ (correct
model, wrong regime) if and only if the declared invariant parameters
$\boldsymbol{\theta}^{(m)}_I$ fail to separate regimes.

\textbf{Proof of Claim:} By the definition of the model's forecast:

\begin{align}
    \hat{\mathbf{y}}_{t+h|t}^{(m)} &=
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(\hat{R}_t)\right)
\end{align}

The forecast error under $(r_0, \mathcal{M}^{(m)})$ is:

\begin{equation}
    \boldsymbol{\varepsilon}_{t,h}^{(m)} = \mathbf{y}_{t+h}(r_0) -
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(\hat{R}_t)\right) \label{eq:fe_pair}
\end{equation}

Consider two scenarios:

\textbf{Scenario A ($H_0$):} True DGP is $r_0$, model parameters
$(\boldsymbol{\theta}^{(m)}_I, \boldsymbol{\theta}^{(m)}_R(r_0) +
\delta)$ where $\delta \neq 0$ represents misspecification.

\textbf{Scenario B ($H_1$):} True DGP is $r_*$, model parameters
$(\boldsymbol{\theta}^{(m)}_I, \boldsymbol{\theta}^{(m)}_R(r_*))$.

The forecast errors are equal in distribution if for all
$\mathcal{F}_t$:

\begin{equation}
    \mathbf{y}_{t+h}(r_0) - \mathbf{g}^{(m)}\left(\mathcal{F}_t;
    \boldsymbol{\theta}^{(m)}_I, \boldsymbol{\theta}^{(m)}_R(r_0) +
    \delta\right) \stackrel{d}{=}
    \mathbf{y}_{t+h}(r_*) - \mathbf{g}^{(m)}\left(\mathcal{F}_t;
    \boldsymbol{\theta}^{(m)}_I, \boldsymbol{\theta}^{(m)}_R(r_*)\right)
    \label{eq:fe_equality}
\end{equation}

Rearranging:

\begin{equation}
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(r_0) + \delta\right) -
    \mathbf{g}^{(m)}\left(\mathcal{F}_t; \boldsymbol{\theta}^{(m)}_I,
    \boldsymbol{\theta}^{(m)}_R(r_*)\right) \stackrel{d}{=}
    \mathbf{y}_{t+h}(r_0) - \mathbf{y}_{t+h}(r_*) \label{eq:fe_rearranged}
\end{equation}

By Assumption~\ref{ass:regime_stat}, the right-hand side is a random
variable whose distribution depends on the difference between regimes
$r_0$ and $r_*$.  The left-hand side depends on the model's parameter
sensitivity.  When $\dim(\boldsymbol{\theta}^{(m)}_R) \geq
\dim(\mathcal{R})$, equation~\eqref{eq:fe_rearranged} admits a
solution $(\delta, r_*)$ for any given $r_0$, establishing
observational equivalence.

When $\dim(\boldsymbol{\theta}^{(m)}_R) < \dim(\mathcal{R})$, the
mapping from $\delta$ to the left-hand side has a non-trivial
nullspace, and there exist multiple regime pairs $(r_0, r_*)$ that
yield identical forecast error distributions for different values of
$\delta$.  This is precisely the unidentifiability result.  $\square$

\subsection{Connection to Partial Identification Literature}

This result connects to the partial identification literature
\citep{manski2003partial, tamer2010partial} in econometrics.  The
identified set for the parameter pair (regime $R_t$, model adequacy)
is:

\begin{equation}
    \Theta_I = \left\{ (R, a) \in \mathcal{R} \times \{0, 1\} :
    p(\mathbf{y} \mid \text{model adequate} = a, R) =
    p_{\text{obs}}(\mathbf{y}) \right\} \label{eq:identified_set}
\end{equation}

where $a = 1$ indicates the model is well-specified.  Theorem~2 shows
that $\Theta_I$ is not a singleton unless the modeler declares
$\boldsymbol{\theta}^{(m)}_I$ such that
$\dim(\boldsymbol{\theta}^{(m)}_I) \geq \dim(\mathcal{R})$ and these
parameters are indeed invariant --- a claim that cannot be verified
from the data alone.

\section{Cercis Score: Additional Derivations}
\label{app:cercis_derivations}

\subsection{Relationship to Proper Scoring Rules}

The CRPS, a strictly proper scoring rule for probabilistic forecasts,
is defined as:

\begin{equation}
    \CRPS(F, y) = \int_{-\infty}^{\infty} (F(z) - \mathbb{I}[z \geq y])^2 dz
    = \mathbb{E}_F[|X - y|] - \frac{1}{2} \mathbb{E}_F[|X - X'|]
    \label{eq:crps_def}
\end{equation}

where $X, X' \stackrel{\text{iid}}{\sim} F$.  The Cercis Score
inherits strict properness from the CRPS in the sense that, for
fixed $\mathcal{N}_t$, $\mathbb{E}[\mathcal{S}(F, y)]$ is uniquely
minimized at the true distribution $F = F^*$.

\subsection{Asymptotic Distribution of Novelty $\mathcal{N}_t$}

Under Assumption~\ref{ass:regime_stat} and the null hypothesis of
no regime change, $\mathbf{s}_t \sim \mathcal{N}(\boldsymbol{\mu}_0,
\mathbf{\Sigma}_0)$ for $t$ in the stable regime.  Then, as the
training sample size $T_{\text{train}} \to \infty$:

\begin{equation}
    \mathcal{N}_t^2 = (\mathbf{s}_t - \hat{\boldsymbol{\mu}})^\top
    \hat{\mathbf{\Sigma}}^{-1} (\mathbf{s}_t - \hat{\boldsymbol{\mu}})
    \xrightarrow{d} \frac{T_{\text{train}} + 1}{T_{\text{train}}}
    \cdot \mathcal{T}^2(3, T_{\text{train}} - 1)
    \label{eq:novelty_asymp}
\end{equation}

where $\mathcal{T}^2$ is the Hotelling's $T^2$ distribution with
3 and $T_{\text{train}} - 1$ degrees of freedom.  For large
$T_{\text{train}}$, $\mathcal{N}_t^2 \xrightarrow{d} \chi^2(3)$,
and the Spring detection threshold $\tau_{\text{spring}}$ can be
calibrated as:

\begin{equation}
    \tau_{\text{spring}} = \chi^2_{3, 1-\alpha_{\text{spring}}}
    \label{eq:spring_threshold}
\end{equation}

where $\chi^2_{3, 1-\alpha}$ is the $(1-\alpha)$ quantile of the
$\chi^2(3)$ distribution.

\subsection{Sensitivity to Hyperparameters}

The Cercis Score depends on two key hyperparameters: $\eta$ (novelty
penalty weight) and $\gamma$ (convexity parameter).  We analyze
the sensitivity of regime change detection to these choices.

\textbf{Choice of $\gamma$:}
\begin{itemize}[leftmargin=*]
    \item $\gamma = 1$: Linear penalty.  Suitable when the cost of
    missing a regime change scales linearly with novelty.
    \item $\gamma = 2$: Quadratic penalty.  Heavily penalizes large
    novelty values, reducing false negatives for major structural
    breaks at the cost of increased false positives for moderate
    deviations.
    \item $\gamma \to \infty$: Max penalty.  Approaches a hard
    threshold: $\mathcal{S} \approx \max(\mathcal{Q}, \eta \cdot
    \mathbb{I}[\mathcal{N} > 1])$.
\end{itemize}

\textbf{Choice of $\eta$:} Following Theorem~\ref{thm:optimal_eta},
we calibrate $\eta^*$ as the ratio of the accuracy gap to the
novelty gap between crisis and non-crisis periods.  Empirically,
using the values from Table~\ref{tab:cercis_decomp}:

\begin{equation}
    \eta^*_{\text{empirical}} = \frac{\bar{\mathcal{Q}}_{\text{crisis}}
    - \bar{\mathcal{Q}}_{\text{stable}}}
    {\bar{\mathcal{N}}_{\text{crisis}}^2 -
    \bar{\mathcal{N}}_{\text{stable}}^2}
    \approx \frac{1.81 - 1.95}{3.55^2 - 1.12^2}
    \approx -0.012 \label{eq:eta_empirical}
\end{equation}

The negative value arises because accuracy loss during crises is
actually \emph{lower} in our SCX consensus (models are carefully
weighted), while novelty spikes dramatically.  In practice, we set
$\eta \in [0.3, 0.8]$ via cross-validation on pre-2000 data.

\section{Additional Empirical Results}
\label{app:additional_empirical}

\subsection{Density Forecast Evaluation}

Table~\ref{tab:density} reports density forecast evaluation using
the CRPS averaged over the evaluation sample.

\begin{table}[htbp]
\centering
\caption{Density Forecast Evaluation: Average CRPS (Lower is Better)}
\label{tab:density}
\begin{tabular}{lccc}
\toprule
Model & $h=1$ & $h=4$ & $h=8$ \\
\midrule
DSGE   & 0.842 & 1.156 & 1.524 \\
BVAR   & 0.718 & 0.982 & 1.305 \\
LSTM   & 0.694 & 1.018 & 1.441 \\
\midrule
SCX (Yajie) & \textbf{0.582} & \textbf{0.804} & \textbf{1.142} \\
SPF         & 0.695  & 0.948  & 1.284 \\
\bottomrule
\end{tabular}
\end{table}

The SCX consensus achieves the lowest CRPS at all horizons,
indicating superior density calibration relative to both individual
models and the SPF.

\subsection{Timing of Spring Detection}

Table~\ref{tab:spring_timing} reports the detection timing of
the Spring module relative to NBER business cycle dating.

\begin{table}[htbp]
\centering
\caption{Spring Detection Timing vs.\ NBER Dating}
\label{tab:spring_timing}
\begin{tabular}{lcccc}
\toprule
Event & NBER Peak & Spring Detection & Lag (Qtrs) & Advance (Qtrs) \\
\midrule
2001 Recession    & 2001:Q1 & 2001:Q2 & +1 & -- \\
2008 GFC          & 2007:Q4 & 2008:Q3 & +3 & -- \\
COVID-19          & 2020:Q1 & 2020:Q2 & +1 & -- \\
\midrule
2001 Recovery     & 2001:Q4 & 2002:Q1 & +1 & -- \\
2009 Recovery     & 2009:Q2 & 2009:Q3 & +1 & -- \\
COVID Recovery    & 2020:Q3 & 2020:Q4 & +1 & -- \\
\bottomrule
\end{tabular}
\end{table}

Spring detects regime changes with a modal lag of 1 quarter for
recoveries and 1--3 quarters for recessions.  The 3-quarter lag for
the 2008 GFC reflects the gradual nature of the financial crisis
onset compared to the abrupt COVID shock.

\subsection{Robustness to Alternative Regime Definitions}

We test robustness to alternative regime classifications: the
NBER-based binary classification (expansion vs.\ recession) and a
6-regime model adding ``recovery'' and ``overheating'' states.

\begin{table}[htbp]
\centering
\caption{Robustness to Regime Definitions: RMSE Ratios}
\label{tab:robustness}
\begin{tabular}{lccc}
\toprule
Regime Definition & SCX Yajie & SCX Equal & Best Individual \\
\midrule
4-regime (baseline)      & 0.82 & 0.91 & 0.92 (RF) \\
2-regime (NBER binary)   & 0.85 & 0.89 & 0.92 (RF) \\
6-regime (extended)      & 0.84 & 0.90 & 0.93 (BVAR) \\
No regimes               & 0.89 & 0.91 & -- \\
\bottomrule
\end{tabular}
\end{table}

The 4-regime specification yields the best performance, suggesting
that the expansion/recession/stagflation/crisis taxonomy captures
economically meaningful distinctions that are lost in the binary
NBER classification.  The 6-regime extension offers no improvement,
likely due to overfitting to sparse regime-specific data.

\section{Computational Implementation Details}
\label{app:implementation}

\subsection{Software Stack}

\begin{itemize}[leftmargin=*]
    \item \textbf{DSGE:} Dynare 5.x with MATLAB R2024a backend;
    Bayesian estimation via tailored MCMC
    \item \textbf{VAR/BVAR:} Custom Python implementation using
    NumPy/SciPy; BVAR with natural conjugate prior via Gibbs sampling
    \item \textbf{ABM:} Custom C++ simulation with Python bindings
    (pybind11); 500 firms, 2000 households
    \item \textbf{Random Forest:} scikit-learn 1.x,
    \texttt{RandomForestRegressor}
    \item \textbf{LSTM:} PyTorch 2.x, 3 layers, 128 hidden units
    \item \textbf{Spring:} Custom Python with Numba JIT compilation
    for CUSUM computation
    \item \textbf{Yajie:} Custom Python with CVXPY for constrained
    weight optimization
\end{itemize}

\subsection{Parallelization Strategy}

The SCX pipeline is embarrassingly parallel across models.  We
implement a distributed architecture using Ray
\citep{moritz2018ray} with the following scheduling:

\begin{itemize}[leftmargin=*]
    \item \textbf{Tier 1 (parallel):} DSGE estimation (8 CPU cores,
    2 hours), BVAR estimation (4 cores, 30 min), LSTM training
    (GPU, 45 min), RF training (8 cores, 15 min), ABM simulation
    (16 cores, 1 hour)
    \item \textbf{Tier 2:} Yajie weight computation (serial, 5 min)
    \item \textbf{Tier 3:} Spring monitoring (serial per quarter,
    2 sec/quarter)
\end{itemize}

Total wall-clock time per quarterly update: approximately 2.5 hours
on a 32-core workstation with one NVIDIA A100 GPU.

\subsection{Reproducibility}

All random seeds are fixed (seed = 42).  The complete pipeline is
containerized via Docker and orchestrated with Apache Airflow for
scheduled quarterly runs.  Replication scripts are available in the
accompanying repository.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{99}

\bibitem{smets2007shocks}
F.~Smets and R.~Wouters.
\newblock Shocks and frictions in US business cycles: A Bayesian DSGE approach.
\newblock {\em American Economic Review}, 97(3):586--606, 2007.

\bibitem{sims1980macroeconomics}
C.~A.~Sims.
\newblock Macroeconomics and reality.
\newblock {\em Econometrica}, 48(1):1--48, 1980.

\bibitem{litterman1986forecasting}
R.~B.~Litterman.
\newblock Forecasting with Bayesian vector autoregressions --- five years of
experience.
\newblock {\em Journal of Business \& Economic Statistics}, 4(1):25--38, 1986.

\bibitem{lucas1976econometric}
R.~E.~Lucas.
\newblock Econometric policy evaluation: A critique.
\newblock {\em Carnegie-Rochester Conference Series on Public Policy},
1:19--46, 1976.

\bibitem{tesfatsion2006agent}
L.~Tesfatsion and K.~L.~Judd, editors.
\newblock {\em Handbook of Computational Economics, Vol.\ 2: Agent-Based
Computational Economics}.
\newblock Elsevier, 2006.

\bibitem{dawid2019agent}
H.~Dawid and D.~Delli Gatti.
\newblock Agent-based macroeconomics.
\newblock In {\em Handbook of Computational Economics}, Vol.\ 4, pages
63--156. Elsevier, 2019.

\bibitem{breiman2001random}
L.~Breiman.
\newblock Random forests.
\newblock {\em Machine Learning}, 45(1):5--32, 2001.

\bibitem{hochreiter1997long}
S.~Hochreiter and J.~Schmidhuber.
\newblock Long short-term memory.
\newblock {\em Neural Computation}, 9(8):1735--1780, 1997.

\bibitem{medeiros2021forecasting}
M.~C.~Medeiros, G.~F.~Vasconcelos, \'{A}.~Veiga, and E.~Zilberman.
\newblock Forecasting inflation in a data-rich environment: The benefits of
machine learning methods.
\newblock {\em Journal of Business \& Economic Statistics}, 39(1):98--119,
2021.

\bibitem{coulombe2020can}
P.~G.~Coulombe, M.~Leroux, D.~Stevanovic, and S.~Surprenant.
\newblock How is machine learning useful for macroeconomic forecasting?
\newblock {\em Journal of Applied Econometrics}, 37(5):920--948, 2022.

\bibitem{verstyuk2020modeling}
S.~Verstyuk.
\newblock Modeling multivariate time series in economics: From autoregressions
to recurrent neural networks.
\newblock Working paper, 2020.

\bibitem{bates1969combination}
J.~M.~Bates and C.~W.~J.~Granger.
\newblock The combination of forecasts.
\newblock {\em Operational Research Quarterly}, 20(4):451--468, 1969.

\bibitem{hoeting1999bayesian}
J.~A.~Hoeting, D.~Madigan, A.~E.~Raftery, and C.~T.~Volinsky.
\newblock Bayesian model averaging: A tutorial.
\newblock {\em Statistical Science}, 14(4):382--401, 1999.

\bibitem{timmermann2006forecast}
A.~Timmermann.
\newblock Forecast combinations.
\newblock In {\em Handbook of Economic Forecasting}, Vol.\ 1, pages 135--196.
Elsevier, 2006.

\bibitem{andrews1993tests}
D.~W.~K.~Andrews.
\newblock Tests for parameter instability and structural change with unknown
change point.
\newblock {\em Econometrica}, 61(4):821--856, 1993.

\bibitem{bai1998estimation}
J.~Bai and P.~Perron.
\newblock Estimating and testing linear models with multiple structural
changes.
\newblock {\em Econometrica}, 66(1):47--78, 1998.

\bibitem{bai2003computation}
J.~Bai and P.~Perron.
\newblock Computation and analysis of multiple structural change models.
\newblock {\em Journal of Applied Econometrics}, 18(1):1--22, 2003.

\bibitem{giordani2007forecasting}
P.~Giordani and R.~Kohn.
\newblock Efficient Bayesian inference for multiple change-point and mixture
innovation models.
\newblock {\em Journal of Business \& Economic Statistics}, 26(1):66--77,
2008.

\bibitem{favero2005}
C.~A.~Favero and F.~Giavazzi.
\newblock Lucas revisited.
\newblock Working paper, IGIER, 2005.

\bibitem{woodford2003interest}
M.~Woodford.
\newblock {\em Interest and Prices: Foundations of a Theory of Monetary
Policy}.
\newblock Princeton University Press, 2003.

\bibitem{an2007bayesian}
S.~An and F.~Schorfheide.
\newblock Bayesian analysis of DSGE models.
\newblock {\em Econometric Reviews}, 26(2-4):113--172, 2007.

\bibitem{manski2003partial}
C.~F.~Manski.
\newblock {\em Partial Identification of Probability Distributions}.
\newblock Springer, 2003.

\bibitem{tamer2010partial}
E.~Tamer.
\newblock Partial identification in econometrics.
\newblock {\em Annual Review of Economics}, 2(1):167--195, 2010.

\bibitem{peters2017elements}
J.~Peters, D.~Janzing, and B.~Sch\"{o}lkopf.
\newblock {\em Elements of Causal Inference: Foundations and Learning
Algorithms}.
\newblock MIT Press, 2017.

\bibitem{gneiting2007strictly}
T.~Gneiting and A.~E.~Raftery.
\newblock Strictly proper scoring rules, prediction, and estimation.
\newblock {\em Journal of the American Statistical Association},
102(477):359--378, 2007.

\bibitem{bernanke1999financial}
B.~S.~Bernanke, M.~Gertler, and S.~Gilchrist.
\newblock The financial accelerator in a quantitative business cycle
framework.
\newblock In {\em Handbook of Macroeconomics}, Vol.\ 1, pages 1341--1393.
Elsevier, 1999.

\bibitem{mccracken2016fred}
M.~W.~McCracken and S.~Ng.
\newblock FRED-MD: A monthly database for macroeconomic research.
\newblock {\em Journal of Business \& Economic Statistics}, 34(4):574--589,
2016.

\bibitem{giannone2015prior}
D.~Giannone, M.~Lenza, and G.~E.~Primiceri.
\newblock Prior selection for vector autoregressions.
\newblock {\em Review of Economics and Statistics}, 97(2):436--451, 2015.

\bibitem{liu2012lucas}
X.~Liu and H.~Wang.
\newblock The Lucas critique in China: Theory and evidence.
\newblock {\em Economic Research Journal}, 2012(4):28--42, 2012.

\bibitem{moritz2018ray}
P.~Moritz, R.~Nishihara, S.~Wang, A.~Tumanov, R.~Liaw, E.~Liang,
M.~Elibol, Z.~Yang, W.~Paul, M.~I.~Jordan, and I.~Stoica.
\newblock Ray: A distributed framework for emerging AI applications.
\newblock In {\em Proceedings of the 13th USENIX OSDI}, pages 561--577, 2018.

\end{thebibliography}

\end{document}
