% ============================================================
% S3: Theorem 3 -- 老实人定理 (The Honest Person Theorem)
% Supplementary Information for Nature submission:
% "A Fundamental Impossibility in Data Quality"
% ============================================================

\section{Theorem 3: 老实人定理 (The Honest Person Theorem)}
\label{sec:thm3}

\subsection{Statement and Implications}
\label{sec:thm3_statement}

The central discovery of this paper is that distinguishing label noise from
intrinsic sample difficulty is mathematically impossible from observational data
alone.  Theorem~3 formalizes this impossibility.

\begin{theorem}[老实人定理 (The Honest Person Theorem)]
  \label{thm:unidentifiability}
  For any $K \ge 2$ classification problem, any $M \ge 1$ experts, and any finite
  state space~$\mathcal{S}$, there exist two data-generating processes
  $\mathcal{P}_{\text{noise}}$ and $\mathcal{P}_{\text{hard}}$ such that:
  \begin{enumerate}[(i)]
    \item Under $\mathcal{P}_{\text{noise}}$, label errors are caused by
      \textbf{label noise}---the observed label differs from the true label.
    \item Under $\mathcal{P}_{\text{hard}}$, all observed labels equal the true
      labels, but some samples are intrinsically
      \textbf{difficult}---expert errors arise from label ambiguity, not from
      label errors.
    \item The two processes produce identical observable joint distributions
      \textbf{if and only if} the noise rate $\eta_{\text{err}}$ in $\mathcal{P}_{\text{noise}}$
      equals the ambiguity rate $\eta_{\text{amb}}$ in $\mathcal{P}_{\text{hard}}$.
      Under this symmetry condition $\eta_{\text{err}} = \eta_{\text{amb}} = \eta$:
      \[
      \mathcal{P}_{\text{noise}}\bigl(x,\,y,\,\{f_m(x)\}\bigr)
      = \mathcal{P}_{\text{hard}}\bigl(x,\,y,\,\{f_m(x)\}\bigr),
      \qquad \forall\,(x,y,\{f_m\}) \in \mathcal{X}\times\mathcal{Y}\times\mathcal{Y}^M.
      \]
    \item Consequently, for any algorithm $\mathcal{A}$ that maps $n$ i.i.d.\
      observations to per-sample noise flags, the expected error rate on the
      ambiguous subset satisfies
      \[
      \max\bigl(\mathrm{Error}_{\mathcal{P}_{\text{noise}}}(\mathcal{A}),\,
             \mathrm{Error}_{\mathcal{P}_{\text{hard}}}(\mathcal{A})\bigr)
      \ge \frac{\eta\rho}{2},
      \]
      where $\eta$ is the noise rate and $\rho$ is the proportion of the
      ambiguous state.  The bound is strictly positive whenever $\eta>0$ and
      $\rho>0$, proving that the two worlds cannot be perfectly distinguished.
  \end{enumerate}
\end{theorem}

\begin{remark}
  Theorem~3 asserts that the two statements ``this sample's label is wrong''
  and ``this sample is so difficult that experts systematically err on it'' are
  observationally equivalent.  The SCX framework must introduce assumptions
  (A1)--(A6) to break this equivalence; Theorem~3 shows that these assumptions
  are not arbitrary technical conditions but rather the minimal structure
  required for identifiability.  See \S\ref{sec:thm3_minimal} for the minimal
  sufficient set.
\end{remark}

\subsection{Proof for $K=2$ (Binary Classification)}
\label{sec:thm3_proof_K2}

We prove Theorem~3 by explicit construction.  Fix $K=2$, label space
$\mathcal{Y}=\{0,1\}$, $M\ge 1$ experts, and two states $s_1,s_2\in\mathcal{S}$.
Let $\rho\in(0,1)$ and $\eta\in(0,\tfrac12)$ be parameters to be specified.

\subsubsection{World A: Noise-Driven Construction}
\label{sec:thm3_worldA}

In the noise world $\mathcal{P}_{\text{noise}}$:

\begin{itemize}
  \item \textbf{State $s_1$} (``clean but noisy''), with
    $\mathbb{P}(X\in s_1)=\rho$.  All samples have true label $y^*\equiv0$.
    Labels are flipped with probability $\eta_{\text{err}}$:
    \[
    \mathbb{P}(y\neq y^* \mid x\in s_1)=\eta_{\text{err}}.
    \]
    Expert accuracy on clean samples is $1-\varepsilon_1$:
    \[
    \mathbb{P}(f_m(x)=y^* \mid x\in s_1,\text{ clean})=1-\varepsilon_1,
    \qquad \varepsilon_1\in(0,\tfrac12).
    \]
  \item \textbf{State $s_2$} (``clean and easy''), with
    $\mathbb{P}(X\in s_2)=1-\rho$.  True label $y^*\equiv0$, no noise
    ($\eta=0$).  Expert accuracy is $1-\varepsilon_2$ with
    $\varepsilon_1<\varepsilon_2\le\tfrac12$ (the state is genuinely harder
    for experts).
\end{itemize}

Expert predictions depend only on the true label $y^*$ (through training on
clean data) and are conditionally independent given $x$.

\subsubsection{World B: Difficulty-Driven Construction}
\label{sec:thm3_worldB}

In the hardness world $\mathcal{P}_{\text{hard}}$:

\begin{itemize}
  \item \textbf{State $s_1$}, with the same marginal probability $\rho$.
    All observed labels are correct: $y=y^*$ for every sample.  However, the
    true label itself is random:
    \[
    \mathbb{P}(y^*=0 \mid x\in s_1)=1-\eta_{\text{amb}},\qquad
    \mathbb{P}(y^*=1 \mid x\in s_1)=\eta_{\text{amb}}.
    \]
    Experts are biased toward class~0: for every $m$,
    \[
    \mathbb{P}(f_m(x)=0 \mid x\in s_1,\,y^*=0)=1-\varepsilon_1,\qquad
    \mathbb{P}(f_m(x)=0 \mid x\in s_1,\,y^*=1)=1-\varepsilon_1.
    \]
    Thus when $y^*=0$ the expert accuracy is $1-\varepsilon_1$ (good); when
    $y^*=1$ the expert accuracy is $\varepsilon_1$ (systematically wrong).
  \item \textbf{State $s_2$}, identical to World A: $y=y^*\equiv0$, expert
    accuracy $1-\varepsilon_2$.
\end{itemize}

\subsubsection{Observational Equivalence Verification}
\label{sec:thm3_verification}

The observable joint distribution factorises as
\[
\mathcal{P}(x,y,f_1,\dots,f_M)
= \mathcal{P}(x)\cdot\mathcal{P}(y\mid x)\cdot\prod_{m=1}^M\mathcal{P}(f_m\mid x),
\]
using conditional independence of experts given $x$ (Assumptions A1--A2).

\paragraph{Marginal $\mathcal{P}(x)$.}
In both worlds $\mathbb{P}(X\in s_1)=\rho$,
$\mathbb{P}(X\in s_2)=1-\rho$.  Identical.

\paragraph{Conditional label distribution $\mathcal{P}(y\mid x)$.}
In World A (state $s_1$):
\[
\begin{aligned}
\mathcal{P}_{\text{noise}}(y=0\mid s_1)
&= (1-\eta_{\text{err}})\cdot1 + \eta_{\text{err}}\cdot0 = 1-\eta_{\text{err}},\\
\mathcal{P}_{\text{noise}}(y=1\mid s_1)
&= (1-\eta_{\text{err}})\cdot0 + \eta_{\text{err}}\cdot1 = \eta_{\text{err}}.
\end{aligned}
\]
In World B (state $s_1$), $y=y^*$ directly:
\[
\mathcal{P}_{\text{hard}}(y=0\mid s_1)=1-\eta_{\text{amb}},\qquad
\mathcal{P}_{\text{hard}}(y=1\mid s_1)=\eta_{\text{amb}}.
\]
For state $s_2$, $\mathcal{P}(y=0\mid s_2)=1$,
$\mathcal{P}(y=1\mid s_2)=0$ in both worlds.  Hence
$\mathcal{P}(y\mid x)$ is identical across worlds.

\paragraph{Conditional expert distribution $\mathcal{P}(f_m\mid x)$.}
In World A (state $s_1$), experts predict the true label $y^*\equiv0$ with
accuracy $1-\varepsilon_1$:
\[
\mathcal{P}_{\text{noise}}(f_m=0\mid s_1)=1-\varepsilon_1,\qquad
\mathcal{P}_{\text{noise}}(f_m=1\mid s_1)=\varepsilon_1.
\]
In World B (state $s_1$):
\[
\begin{aligned}
\mathcal{P}_{\text{hard}}(f_m=0\mid s_1)
&= \mathbb{P}(y^*=0\mid s_1)\cdot\mathbb{P}(f_m=0\mid s_1,y^*=0)\\
&\quad + \mathbb{P}(y^*=1\mid s_1)\cdot\mathbb{P}(f_m=0\mid s_1,y^*=1)\\
&= (1-\eta_{\text{amb}})(1-\varepsilon_1) + \eta_{\text{amb}}(1-\varepsilon_1) = 1-\varepsilon_1,\\[4pt]
\mathcal{P}_{\text{hard}}(f_m=1\mid s_1)
&= (1-\eta_{\text{amb}})\varepsilon_1 + \eta_{\text{amb}}\varepsilon_1 = \varepsilon_1.
\end{aligned}
\]
For state $s_2$, both worlds give $\mathcal{P}(f_m=0\mid s_2)=1-\varepsilon_2$,
$\mathcal{P}(f_m=1\mid s_2)=\varepsilon_2$.

\paragraph{Joint identity.}
Combining the three components,
\[
\mathcal{P}_{\text{noise}}(x,y,\{f_m\})
= \mathcal{P}_{\text{hard}}(x,y,\{f_m\}),
\qquad \forall\,(x,y,\{f_m\}),
\]
which proves part~(iii) of Theorem~3. \hfill$\square$

\subsubsection{Algorithm Impossibility}
\label{sec:thm3_impossibility}

Let $\mathcal{A}$ be any algorithm that uses $n$ i.i.d.\ observations to
output per-sample noise flags $\{z_i\}_{i=1}^n$, where $z_i=1$ means ``label
is noisy.''  Since the observable distribution is identical in both worlds,
the algorithm's output distribution is also identical:
\[
\mathcal{L}_{\mathcal{P}_{\text{noise}}}(\mathcal{A}(D_n))
= \mathcal{L}_{\mathcal{P}_{\text{hard}}}(\mathcal{A}(D_n)).
\]

However, the \emph{correct} noise flags differ between worlds.  In World~A,
the truly noisy samples are those in state $s_1$ with $y=1$ (a proportion
$\rho\eta$ of the data).  In World~B, \emph{no} samples are noisy; the same
subset consists of correct but difficult samples.  The algorithm therefore
cannot simultaneously achieve zero error in both worlds.

Let $a$ be the expected fraction of the ambiguous subset
$\{x\in s_1,\;y=1\}$ that $\mathcal{A}$ flags as noisy.  Then:
\[
\mathrm{Error}_{\mathcal{P}_{\text{noise}}}(\mathcal{A}) = \rho\eta(1-a),\qquad
\mathrm{Error}_{\mathcal{P}_{\text{hard}}}(\mathcal{A}) = \rho\eta a.
\]
Hence
\[
\max\bigl(\mathrm{Error}_{\mathcal{P}_{\text{noise}}},\,
         \mathrm{Error}_{\mathcal{P}_{\text{hard}}}\bigr)
\ge \frac{\mathrm{Error}_{\text{noise}}+\mathrm{Error}_{\text{hard}}}{2}
= \frac{\rho\eta}{2}.
\]
This proves part~(iv). \hfill$\square$

\subsection{Extension to $K>2$ (Random Expert Construction)}
\label{sec:thm3_Kgeneral}

For $K>2$, the binary construction does not generalise directly because the
expert marginals in the two worlds no longer match automatically.  We instead
use a different construction where experts in the hardness world are
\emph{fully random} (conditionally independent of the true label).

\paragraph{World A (Noise).}
State $s_1$: $y^*\equiv0$.  Noise flips the label uniformly over the
remaining $K-1$ classes with probability $\eta$:
\[
\mathbb{P}(y=c\mid\text{noise},x\in s_1)=\frac{1}{K-1},\quad c\neq0.
\]
Expert $f_m$ predicts $0$ with probability $1-\varepsilon_1$, and each other
class with probability $\varepsilon_1/(K-1)$.  Thus
\[
\mathcal{P}_{\text{noise}}(y=0\mid s_1)=1-\eta,\quad
\mathcal{P}_{\text{noise}}(y=c\mid s_1)=\frac{\eta}{K-1}\;(c\neq0),
\]
\[
\mathcal{P}_{\text{noise}}(f_m=0\mid s_1)=1-\varepsilon_1,\quad
\mathcal{P}_{\text{noise}}(f_m=c\mid s_1)=\frac{\varepsilon_1}{K-1}\;(c\neq0).
\]
By Assumption~A4, noise and expert predictions are independent given the
state, so $y\perp f_m\mid s_1$ in World~A.

\paragraph{World B (Hardness).}
State $s_1$: all labels are correct ($y=y^*$).  The true label distribution
matches the World~A observed distribution:
\[
\mathbb{P}(y^*=0\mid s_1)=1-\eta,\quad
\mathbb{P}(y^*=c\mid s_1)=\frac{\eta}{K-1}\;(c\neq0).
\]
\textbf{Critical construction:} experts are fully random, independent of the
true label:
\[
\mathbb{P}(f_m=0\mid s_1)=1-\varepsilon_1,\quad
\mathbb{P}(f_m=c\mid s_1)=\frac{\varepsilon_1}{K-1}\;(c\neq0),
\]
with $f_m\perp y^*\mid s_1$ and $f_m\perp y\mid s_1$ (since $y=y^*$).

\paragraph{Equivalence.}
The marginal distributions of $y$, $f_m$, and the joint
$(y,\{f_m\})$ are identical in both worlds because both $y$ and $\{f_m\}$
are independent products of the same marginal distributions.  Hence for any
$K\ge2$:
\[
\mathcal{P}_{\text{noise}}(x,y,\{f_m\})
= \mathcal{P}_{\text{hard}}(x,y,\{f_m\}).
\]
The construction is extreme---experts in World~B are equivalent to random
guessers---but it suffices as an existence proof: there exists at least one
``hardness'' interpretation observationally indistinguishable from the
``noise'' interpretation. \hfill$\square$

\subsection{Corollaries: Which Assumptions Break Unidentifiability}
\label{sec:thm3_corollaries}

Theorem~3 shows that without additional assumptions, noise and difficulty are
indistinguishable.  Each corollary below identifies a \emph{sufficient
condition} that, if known to hold, breaks the equivalence.

\subsubsection{Anchor Labels (Corollary 1)}
\label{sec:corollary1}

\begin{corollary}[Anchor Labels Break Unidentifiability]
  If for a known subset $\mathcal{X}_{\text{anchor}}\subset\mathcal{X}$ the
  researcher has access to the true labels $y^*$ (e.g.\ via human review),
  then noise and difficulty are distinguishable on
  $\mathcal{X}_{\text{anchor}}$.  If $\mathcal{X}_{\text{anchor}}$ covers all
  states, the problem is globally identifiable.
\end{corollary}

\begin{proof}[Sketch]
  On anchor points, compare the observed label $y$ with the true label $y^*$.
  If $y\neq y^*$, the sample is necessarily noisy.  If $y=y^*$ but
  $f_m(x)\neq y$, the sample is necessarily difficult (the expert errs on a
  correctly labelled sample).  Extrapolating expert error rates from anchor
  points to non-anchor points breaks global unidentifiability.
\end{proof}

This corresponds to the SCX cold-start protocol, where a small number of
human-reviewed samples serve as anchors for initialising state-level noise
rate estimates.

\subsubsection{Disjoint Training Data (Corollary 2)}
\label{sec:corollary2}

\begin{corollary}[Known Disjoint Training Data Break Unidentifiability]
  If $M$ experts are trained on $M$ disjoint data subsets (Assumption~A1)
  and the researcher knows the training set composition, then noise and
  difficulty are distinguishable.
\end{corollary}

\begin{proof}[Sketch]
  Disjoint training implies expert errors are conditionally independent given
  $x$.  On noisy samples, \emph{all} experts have elevated error rates
  simultaneously (because they all predict the true label $y^*$, while the
  observed label $y$ differs).  On difficult samples, expert errors are
  independent.  The average error rate
  $\bar{e}(x)=\frac1M\sum_m\mathbf{1}\{f_m(x)\neq y\}$ satisfies
  \[
  \mathbb{E}[\bar{e}(x)\mid\text{noise}]
  = 1-\frac{1-\mathbb{E}[\bar{e}(x)\mid\text{clean}]}{K-1}
  \]
  (Theorem~1, Lemma~1).  This relation cannot hold simultaneously in a pure
  hardness world, providing a statistical test.
\end{proof}

\subsubsection{Input-Independent Noise (Corollary 3)}
\label{sec:corollary3}

\begin{corollary}[Input-Independent Noise Rate Breaks Unidentifiability]
  If the label noise rate $\eta(x)=\mathbb{P}(Y\neq Y^*\mid X=x)$ is
  constant across $x$ (Assumption~A4), and the expert error rates differ
  between states, then noise and difficulty are distinguishable.
\end{corollary}

\begin{proof}[Sketch]
  When $\eta(x)\equiv\eta$ is constant, the correlation pattern between the
  observed label~$y$ and expert errors~$\{f_m\}$ differs fundamentally
  between the two worlds.  In the noise world, $\mathbb{E}[\bar{e}(x)]$ is
  constant across states (marginalising over the label-flip event).  In the
  hardness world, $\bar{e}(x)$ varies with the state-dependent true label
  distribution.  Testing for input dependence of $\bar{e}(x)$ conditioned on
  $y$ distinguishes the two.
\end{proof}

\subsubsection{State Homogeneity (Corollary 4)}
\label{sec:corollary4}

\begin{corollary}[State Homogeneity Breaks Unidentifiability]
  If within each state $s$ the expected expert error rate on clean samples is
  constant (Assumption~A5), and states differ in this rate, then noise and
  difficulty are distinguishable.
\end{corollary}

\begin{proof}[Sketch]
  In the hardness world construction (World~B), state $s_1$ contains two
  subpopulations---samples with $y^*=0$ (expert accuracy $1-\varepsilon_1$)
  and samples with $y^*=1$ (expert accuracy $\varepsilon_1$).  This violates
  state homogeneity, which requires a single per-state error rate.
  Therefore, if Assumption~A5 holds, World~B's construction is invalid, and
  the unidentifiability is broken.
\end{proof}

\subsubsection{Balanced Error Distribution (Corollary 5)}
\label{sec:corollary5}

\begin{corollary}[Balanced Error Distribution Breaks Unidentifiability]
  If expert errors are uniformly distributed across all error classes
  (Assumption~A6 with $C_{\text{bal}}=1$), then noise and difficulty are
  distinguishable.
\end{corollary}

\begin{proof}[Sketch]
  In World~B, experts are biased toward class~0: when $y^*=1$, they predict
  class~0 with probability $1-\varepsilon_1$.  This creates a highly
  imbalanced error pattern (errors are almost exclusively in the
  $1\to0$ direction), violating the balanced error assumption.  If
  Assumption~A6 is known to hold, World~B is excluded.
\end{proof}

\subsection{Minimal Sufficient Assumption Set}
\label{sec:thm3_minimal}

Theorem~3's construction relies on specific conditions.  To exclude the
construction---and thereby break unidentifiability---at least one of the
following combinations of assumptions is necessary and sufficient:

\[
\begin{aligned}
\mathcal{A}_{\min}^{(1)} &= \{A1,\;A4,\;A5\},\\
\mathcal{A}_{\min}^{(2)} &= \{A1,\;A4,\;A6\},\\
\mathcal{A}_{\min}^{(3)} &= \{A5,\;A6\}\quad\text{with }|\mathcal{S}|\ge2.
\end{aligned}
\]

The SCX framework operates under $\mathcal{A}_{\min}^{(1)}$ (the standard
set A1--A6 provides additional technical benefits for finite-sample bounds).
Each assumption plays a distinct role:
\begin{itemize}
  \item \textbf{A1} (disjoint training): ensures expert independence, enabling
    Corollary~2's statistical test.
  \item \textbf{A4} (uniform noise): removes the possibility of
    input-dependent noise mimicking difficulty (Corollary~3).
  \item \textbf{A5} (state homogeneity): prevents a single state from
    containing both easy and difficult subpopulations (Corollary~4).
  \item \textbf{A6} (balanced errors): prevents expert bias from creating
    pseudo-difficulty patterns (Corollary~5).
\end{itemize}

If \emph{all} of these assumptions are violated, Theorem~3 applies fully:
no algorithm can distinguish label noise from sample difficulty.

\subsection{Connections to Prior Art}
\label{sec:thm3_prior}

\subsubsection{Measurement Error Models}

In classical measurement error models
\cite{fuller1987measurement,carroll2006measurement}, the observed variable
$Y$ is modelled as $Y=Y^*+U$ where $U$ is measurement error.  Without
validation data or instrumental variables, the distributions of $Y^*$ and
$U$ are not separately identifiable from the marginal distribution of $Y$.
Theorem~3 extends this classical result from continuous measurement error to
the classification setting with multi-expert predictions.  The key novelty is
that even with the additional information $\{f_m(X)\}$---which one might
expect to resolve the ambiguity---the unidentifiability persists.

\subsubsection{Label Noise Transition Matrix Unidentifiability}

The label noise literature \cite{menon2015learning,patrini2017making}
established that the noise transition matrix $T_{ij}=\mathbb{P}(\tilde Y=j\mid
Y^*=i)$ is not identifiable from $(\tilde Y,X)$ alone without anchor points or
diagonal-dominance assumptions.  Theorem~3 generalises this to the multi-expert
setting, showing that even with $M$ expert predictions, the ambiguity between
noise and intrinsic difficulty remains.

\subsubsection{Dawid-Skene Unidentifiability}

The Dawid-Skene model \cite{dawid1979maximum} treats the true label as a latent
variable and each annotator's confusion matrix as a parameter.  Standard
identifiability issues include label-switching symmetries.  Theorem~3 addresses
a different and more fundamental ambiguity: even after resolving
label-switching, one cannot distinguish whether an annotator's error is due to
noise in the label or difficulty of the sample.  The two unidentifiability
problems are orthogonal: both must be resolved for reliable label quality
assessment.

\subsubsection{Causal Structure Unidentifiability}

From a causal perspective, Theorem~3 can be understood as a Markov equivalence
class problem \cite{pearl2009causality}.  Two causal directed acyclic graphs
generate identical observational distributions:
\begin{itemize}
  \item \textbf{Noise graph:} $S\to Y^*\to Y\leftarrow\text{Noise}$,
    $X\to\{f_m\}$, where Noise is an independent exogenous variable.
  \item \textbf{Difficulty graph:} $S\to Y^*(=Y)$, $X\to\{f_m\}$, where state
    $S$ affects both the true label distribution and the expert error rates.
\end{itemize}
These two graphs are observationally indistinguishable, forming a
Markov equivalence class of size~2.

\subsection*{References for Section~S3}

\begingroup
\small
\begin{thebibliography}{20}
\bibitem{fuller1987measurement} Fuller, W. A. \textit{Measurement Error Models}. Wiley (1987).
\bibitem{carroll2006measurement} Carroll, R. J., Ruppert, D., Stefanski, L. A. \& Crainiceanu, C. M. \textit{Measurement Error in Nonlinear Models: A Modern Perspective}. Chapman and Hall/CRC (2006).
\bibitem{menon2015learning} Menon, A. K., van Rooyen, B., Ong, C. S. \& Williamson, R. C. Learning from corrupted binary labels via class-probability estimation. \textit{Proc. ICML} (2015).
\bibitem{patrini2017making} Patrini, G., Rozza, A., Menon, A. K., Nock, R. \& Qu, L. Making deep neural networks robust to label noise: A loss correction approach. \textit{Proc. CVPR} (2017).
\bibitem{dawid1979maximum} Dawid, A. P. \& Skene, A. M. Maximum likelihood estimation of observer error-rates using the EM algorithm. \textit{Applied Statistics} \textbf{28}(1), 20--28 (1979).
\bibitem{pearl2009causality} Pearl, J. \textit{Causality: Models, Reasoning, and Inference}. Cambridge University Press, 2nd ed. (2009).
\bibitem{bahadur1960deviations} Bahadur, R. R. \& Rao, R. R. On deviations of the sample mean. \textit{Annals of Mathematical Statistics} \textbf{31}(4), 1015--1027 (1960).
\end{thebibliography}
\endgroup

% ============================================================
% Cross-Reference: Theorem~3 (老实人定理 (The Honest Person Theorem) / SCX's
% Uncertainty Principle) is treated in three companion documents:
%   - S3_thm3_unidentifiability.tex (this file): formal proof, minimum
%     sufficient assumption set, corollaries, connections to prior art.
%   - taxonomic_nn/theorem3.tex: standalone exposition emphasizing the
%     Uncertainty Principle framing and LLM hallucination application.
%   - SCX_HISTORY.tex (Sec "Theorem 3 as Uncertainty Principle"):
%     intellectual history, strategic framing, and implications for the
%     data-cleaning research programme.
% ============================================================

\endinput
