% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Hostile Review Report: Line-by-Line Attack on Situs Theory}\label{hostile-review-report}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Reviewer Identity}: The most demanding reviewer at Annals of Statistics\\
\textbf{Review Date}: 2026-06-29\\
\textbf{Review Target}: \texttt{paper/situs\_theory/main.tex} + \texttt{theory/self\_evolution/ppe\_rigorous\_derivation.md}\\
\textbf{Review Standard}: Impolite. Uncompromising. If there is one hole, do not let it go.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{0. Overall Impression}\label{overall-impression}

This paper claims to establish a ``rigorous mathematical foundation'' for Situs (physical position encoding) on top of a framework called SCX. After actually reading all three documents, the conclusion is: \textbf{This paper has fundamental errors or irreparable logical gaps in the proofs of several core theorems. Theorem 2.4.1's proof is wrong — not incomplete, but wrong. Theorem 1.3.1's Lipschitz constant is off by a factor of approximately 3.5. Theorem 2.2.1's Step 4 is an unbridgeable chasm with current tools. Theorem 3''s ``constructive example'' actually falsifies the proposition it claims to prove. Below is a step-by-step breakdown.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{I. Notational Consistency: Main Paper vs ppe\_rigorous\_derivation.md}\label{i-notational-consistency}

\subsubsection{1.1 Theorem Numbering Inconsistency}\label{theorem-numbering-inconsistency}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1071}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2321}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.5536}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1071}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
main.tex Numbering
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ppe\_rigorous\_derivation.md Numbering
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Difference
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Learnable PE Destruction Theorem & Theorem 4.2 (line 1048) & Theorem 4.1 (§4.2.3, line 851) & \textbf{Numbering off by 1} \\
Encoding Imperfection Upper Bound & Proposition 3.1 (line 864) & Proposition 3.1 (§3.1.3, line 690) & Consistent \\
Theorem 2' & Theorem 2' (line 881) & Theorem 2' (§3.2, line 702) & Consistent \\
\end{longtable}

\textbf{Attack}: The numbering jump from Theorem 4.1 to Theorem 4.2 implies there is a ``vanished Theorem 4.1'' in the paper body. If Theorem 4.1 is Proposition 4.1 (fixed PE preserving Theorem 3), why isn't Proposition 4.1 called Theorem 4.1? This isn't a typo — this is a break in the logical chain: Proposition 4.1 uses a trivial deterministic function argument, not deserving the title of theorem; but Theorem 4.2's ``destructive'' result is the real theorem-level contribution. The numbering chaos exposes the author's own confusion about the importance of their contributions.

\subsubsection{1.2 Dual Personality of the Symbol Naming System}\label{dual-personality-of-the-symbol-naming-system}

The paper body uses \texttt{\textbackslash{}Situs}, \texttt{\textbackslash{}Sstates}, \texttt{\textbackslash{}Ppos}, \texttt{\textbackslash{}PE}, \texttt{d\_\{\textbackslash{}pe\}}; ppe\_rigorous\_derivation.md uses \texttt{PPE}, \texttt{\textbackslash{}mathcal\{S\}}, \texttt{\textbackslash{}mathcal\{P\}}, \texttt{\textbackslash{}text\{PE\}}, \texttt{d\_\{\textbackslash{}text\{pe\}\}}. The paper claims these numbers are ``consistent'' with ppe\_rigorous\_derivation.md (line 1343), but they differ in all symbols. This is not a LaTeX vs Markdown formatting difference — the paper uses \texttt{\textbackslash{}Ppos} for position space, md uses \texttt{\textbackslash{}mathcal\{P\}}. The author declares consistency in the same sentence while demonstrating inconsistency.

\subsubsection{1.3 Two Versions of Theorem 2.3.1}\label{two-versions-of-theorem-2.3.1}

\textbf{Paper version} (line 697-701):
\[\delta_s^{\PE} \leq \min\!\left(1 - p_{\text{clean}, s},\, \sqrt{\frac{1}{2} D_{\KL}(P_{\PE|S,Y} \| P_{\PE|S})}\right) + \delta_s^{\text{variance}}\]

\textbf{md version} (line 512-513):
\[\delta_s^{\text{PE}} \leq \min\left(1, \sqrt{\frac{1}{2} \cdot I(P; \text{PE}(P) \mid S, Y)}\right) + \text{Bernstein correction term}\]

These are \textbf{not} the same upper bound. The first uses \(1-p_{\text{clean},s}\) and KL divergence (pointwise conditional), the second uses \(1\) and mutual information (expectation form). The md version uses ``Bernstein correction'' (implying \(O(1/M)\) convergence), the paper uses \(\delta_s^{\text{variance}} = O(1/\sqrt{M})\) (implying CLT convergence). The two have entirely different scaling laws under finite \(M\) — one is \(1/M\), the other is \(1/\sqrt{M}\).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{II. Assumption Check: Do A1-A6 Still Hold under Situs?}\label{ii-assumption-check}

The CC audit report implicitly contains 6 core assumptions of the original SCX theorems. Let us check each one's status after introducing Situs:

\subsubsection{\texorpdfstring{A1: Expert Independence (\(v_m \indep v_{m'} \mid s\))}{A1: Expert Independence}}\label{a1-expert-independence}

\textbf{Original status}: \(M\) experts' votes on atoms of the same state \(s\) are i.i.d. Bernoulli.

\textbf{After introducing Situs}: All experts receive the same augmented input \(h_i = \phi(s_i) + \PE(p_i)\). The PE term is a shared deterministic signal — if PE carries information about the label, then all experts' inputs include the same ``hint.'' This weakens the conditional independence assumption of experts: \(\Cov(v_m, v_{m'} \mid s, p) \neq 0\). Strictly speaking, the Chernoff-Hoeffding bound requires conditional independence given the state; now PE injects a common signal shared by all experts. \textbf{This breaks the rigor of the i.i.d. assumption.}

Attack severity: \textbf{Medium}. When \(M\) is large, the correlation introduced by this shared signal may loosen the Chernoff bound by a constant factor (similar to \(\beta\)-mixing correction), but the paper does not mention this problem — it pretends the Chernoff-Hoeffding structure is ``unchanged.''

\subsubsection{\texorpdfstring{A2: Positive Detection Margin (\(\Delta_s > 0\))}{A2: Positive Detection Margin}}\label{a2-positive-detection-margin}

\textbf{Original status}: \(\Delta_s = p_{\text{noisy},s} - p_{\text{clean},s} > 0\).

\textbf{After introducing Situs}: We need \(\Delta_s^{\Situs} = \Delta_s + \delta_s^{\PE} > 0\). But \(\delta_s^{\PE} \in [-\Delta_s, 1-p_{\text{clean},s}]\). \textbf{Situs can reduce the detection margin}! If \(\delta_s^{\PE} < 0\) and \(|\delta_s^{\PE}| > \Delta_s\), then \(\Delta_s^{\Situs} < 0\) — Situs reverses the detection direction. Theorem 2.2.1 only gives sufficient conditions for the existence of some \(s\) and some PE such that \(\delta_s^{\PE} > 0\), but \textbf{does not rule out the possibility that \(\delta_s^{\PE} < 0\) for other \(s\)}.

Attack severity: \textbf{High}. The paper's corrected Theorem 1 replaces all states' \(\Delta_s\) with \(\Delta_s + \delta_s^{\PE}\), but \(\delta_s^{\PE}\) can be negative. For some \(s\), \(\exp(-2M(\Delta_s + \delta_s^{\PE})^2)\) may be larger than \(\exp(-2M\Delta_s^2)\), i.e., the \(F_1\) lower bound \textbf{gets worse}. This is not the ``systematic enhancement'' the paper claims.

\subsubsection{A3: Discreteness of State Space}\label{a3-discreteness-of-state-space}

\textbf{Original status}: \(\mathcal{S} = \{s_1, \ldots, s_N\}\) is a finite discrete set.

\textbf{After introducing Situs}: The encoding space \(\R^d\) is continuous, \(h_i = \phi(s_i) + \PE(p_i)\) varies continuously in \(\R^d\). Even if \(s_i\) is discrete, \(h_i\) is not. This means the ``equivalence class'' among state atoms is no longer strict — two atoms originally mapping to the same state \(s\) will have different \(h_i\) if their physical positions differ. This makes the definition of the probability \(\rho_s\) for ``aggregation by state \(s\)'' ambiguous — originally aggregated by state atom type, now aggregation must be by \((s, p)\) joint.

Attack severity: \textbf{Low-Medium}. Technically fixable (define distribution on joint space), but the paper does not discuss this subtlety.

\subsubsection{A4: Bounded Random Variables (Chernoff-Hoeffding Applicability)}\label{a4-bounded-random-variables}

\textbf{Original status}: \(v_m \in \{0,1\}\) is bounded.

\textbf{After introducing Situs}: Votes remain in \(\{0,1\}\); Chernoff-Hoeffding applicability is unchanged. This assumption \textbf{formally holds}. But as described in A1, the independence condition is weakened.

Attack severity: \textbf{Low} (formally holds, substantively implicated by A1).

\subsubsection{A5-A6: Bayes Error Rate and Fano Condition}\label{a5-a6-bayes-error-rate-and-fano-condition}

These involve internal constructions of the original Theorem 2 and Theorem 3, not explicitly numbered in the CC audit report. Under Situs: - The Bayes error rate in the augmented feature space requires extra care (augmented space introduces continuous dimensions) - The \(\log|\Y|\) term in Fano's inequality is unchanged, but conditional entropy computation involves continuous-discrete mixtures

Attack severity: \textbf{Low}. These are standard information-theoretic technical details, manageable, but the paper does not address them.

\subsubsection{Assumption Check Conclusion}\label{assumption-check-conclusion}

\textbf{Among the 6 assumptions the paper claims, at least A1 (independence) and A2 (positive margin) are significantly impacted under Situs. The paper does not verify each assumption still holds before claiming the Chernoff structure is ``unchanged'' — this is dishonest.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{III. Proof Completeness: Logical Gaps, Unstated Assumptions, Erroneous Reasoning}\label{iii-proof-completeness}

\subsubsection{3.1 Theorem 2.2.1 Step 4: Generalization from Bayes Optimal to Actual Experts (Fatal)}\label{theorem-2.2.1-step-4}

Paper body line 659-661: \textgreater{} Step 4 (from Bayes optimal to actual experts): Assuming actual experts consistently approximate the Bayes optimal (extracting information from augmented features in the same direction as the Bayes optimal), the sign of the above inequality is preserved in expectation.

The paper itself labels this as \texttt{{[}heuristic{]}}. But this is not heuristic — \textbf{this is an unbridgeable chasm}. The problems are:

\textbf{(a)} There is no finite-sample convergence guarantee between the behavior of the Bayes optimal classifier given \((s, \PE(p))\) and the behavior of any arbitrarily trained neural network expert \(E_m\).

\textbf{(b)} What does ``consistently approximate the Bayes optimal'' mean? If the experts are neural networks trained by stochastic gradient descent, their convergence points are some stationary points of the loss landscape, not necessarily Bayes optimal. Even if they are, it is in the \textbf{population distribution} sense, without guarantee that the sign of \(\delta_s^{\PE}\) is correct for each specific state \(s\).

\textbf{(c)} The paper claims \(p_{\text{clean},s}^{\Situs} \leq p_{\text{clean},s}\) (disagreement on clean samples does not increase). The Bayes optimal classifier indeed does not get worse with more information (data processing inequality guarantee). But actual experts may perform worse on clean samples because PE introduces \textbf{spurious correlations} — especially when training data is limited, PE may learn a pattern that happens to be effective on the training set but invalid on the test set. In this case \(p_{\text{clean},s}^{\Situs} > p_{\text{clean},s}\) and \(\delta_s^{\PE}\) may be negative.

\textbf{(d)} The entire conclusion of Theorem 2.2.1 — ``there exists PE and dimension such that \(\delta_s^{\PE} > 0\)'' — without Step 4, only holds for the Bayes optimal classifier. For neural network experts in the actual SCX framework, this is an \textbf{unproven conjecture}.

\textbf{Verdict}: Theorem 2.2.1 is the paper's most important information-theoretic result (connecting \(I(Y;P|S) > 0\) to detection improvement). Removing Step 4, it only proves that the Bayes optimal classifier can benefit from position information — which is trivial. Applying this result to actual experts in the SCX framework is a \textbf{logical gap}.

\subsubsection{3.2 Theorem 2.4.1 Step 2: Directionality of Fano's Inequality (Fatal)}\label{theorem-2.4.1-step-2}

This is the most serious error I found. The paper claims that Fano gives a \textbf{lower bound} for \(\delta_s^{\PE}\). But a careful examination of the derivation:

\textbf{Step 1}: Fano's inequality standard form (correct) — \[H(Y \mid S, \PE(P)) \leq H_2(P_e) + P_e \cdot \log(|\Y| - 1)\]

From this we obtain (paper line 750-752): \[P_e \geq \frac{H(Y \mid S, \PE(P)) - 1}{\log |\Y|}\]

This is a correct \textbf{lower bound}.

\textbf{Step 3} (paper line 763-766, md line 601-607): \[P_e - P_e^{\Situs} = \frac{H(Y|S) - H(Y|S, \PE(P))}{\log |\Y|} \quad (\text{taking Fano's equality})\]

\textbf{This is a fundamental error.} The logical chain is: 1. \(P_e \geq \frac{H(Y|S)-1}{\log|\Y|}\) (Fano) 2. \(P_e^{\Situs} \geq \frac{H(Y|S,\PE(P))-1}{\log|\Y|}\) (Fano) 3. If both inequalities attain equality → \(P_e - P_e^{\Situs} = \frac{H(Y|S) - H(Y|S,\PE(P))}{\log|\Y|}\)

But from the \textbf{inequalities} in (1) and (2) to the \textbf{equality} in (3), we need (i) Fano's inequality attains equality in both worlds, AND (ii) ``the difference of two Fano equalities = the difference of error rates.''

The problem is: - Fano's inequality only attains equality under specific distributions (noise uniformly distributed over error classes, and \(H_2(P_e)\) is binary entropy), which does not hold in general. - More importantly, even if both Fano bounds attain equality, \(a \geq A\) and \(b \geq B\) \textbf{do not imply} \(a-b = A-B\).

The md version explicitly notes ``approximate, taking Fano's equality'' — \textbf{this admits it is not a rigorous derivation}. But in the paper body, Theorem 2.4.1's statement has no such ``approximate'' qualification; it is presented as a \textbf{theorem} with \texttt{{[}rigorous{]}/{[}heuristic{]}} mixed annotations (Step 1 rigorous, Steps 2-3 heuristic).

\textbf{The essence of the directionality error}: The paper attempts to derive a \textbf{lower bound} on error rate improvement from Fano's \textbf{upper bound} (conditional entropy). Fano's inequality is of the form \(H(Y|X) \leq f(P_e)\), giving a lower bound on \(P_e\). But \(P_e - P_e^{\Situs}\) as a difference cannot be directly derived from two lower bounds on \(P_e\) — the direction of the difference is not controlled by individual lower bounds.

\textbf{Correct approach}: What is needed is an \textbf{upper bound} on \(P_e - P_e^{\Situs}\) to use Fano for any rigorous bound. Or use stronger tools like Fano's inverse (but this typically requires additional distributional assumptions).

\textbf{Verdict}: \textbf{Theorem 2.4.1's proof is wrong.} Not incomplete, not heuristic — it is a logical fallacy of deriving equality from inequality. The theorem statement's ``lower bound'' wording misleads readers into thinking this is a rigorous information-theoretic bound. It is at most a heuristic guess.

\subsubsection{3.3 Theorem 2.2.1 Step 3's ``Clean Sample Disagreement Does Not Increase'' Argument}\label{theorem-2.2.1-step-3}

Paper line 655: \textgreater{} For clean samples, additional position information further confirms the correct label → disagreement probability does not increase: \(p_{\text{clean}, s}^{\Situs} \leq p_{\text{clean}, s}\).

This relies on an implicit assumption: PE's information is ``consistent'' with rather than ``conflicting'' with the correct label. But: - PE is derived from physical position; physical position may have \textbf{spurious correlation} with the label (correlated in training distribution but not in target distribution). - PE may amplify some systematic bias of experts — if all experts share a common blind spot, additional PE information may make them collectively more confident in being wrong on that blind spot. - The paper itself acknowledges this risk: in §6.3's ``one-vote veto condition,'' F5 points out that for ``random or pseudo-random configurations + finite samples,'' \(\delta_s^{\PE} > 0\) on the training set may become \(\delta_s^{\PE} < 0\) on the test set.

\textbf{Attack}: Step 3's argument assumes PE's information is always ``beneficial'' — it increases noise sample disagreement but does not increase clean sample disagreement. This is equivalent to \textbf{assuming the conclusion}: PE somehow distinguishes noise and clean samples. If PE's information is purely an additional feature (correlated but not causal), it may increase disagreement for both types of samples equally, leading to \(\delta_s^{\PE} = 0\).

\subsubsection{3.4 Theorem 2.5.1: Illegal Use of Lipschitz Continuity Assumption}\label{theorem-2.5.1}

Theorem 2.5.1 claims: \[|\delta_i^{\PE} - \delta_j^{\PE}| \leq 2 \cdot L_E \cdot L_{\PE} \cdot d_{\Ppos}(p_i, p_j)\]

The proof uses (md line 640): \[|p_{\text{clean}, i}^{\text{PPE}} - p_{\text{clean}, j}^{\text{PPE}}| \leq L_E \cdot \|h_i - h_j\|\]

But \(p_{\text{clean}, i}^{\text{PPE}} = \mathbb{E}[\mathbf{1}[E_m(h_i) \neq y]]\) is the \textbf{expectation of an indicator function}. Even if \(E_m\)'s output is \(L_E\)-Lipschitz (in some sense), the indicator function \(\mathbf{1}[\cdot \neq y]\) is \textbf{discontinuous} — it jumps at the decision boundary. The \(L_E\)-Lipschitz assumption typically applies to the network's output logit or softmax probability, not the expectation of the 0/1 decision.

For this bound to hold, additional assumptions are needed: - The expert's decision boundary is sufficiently far from both \(h_i\) and \(h_j\) (margin condition), or - The expert outputs soft probabilities and the expectation of the indicator function happens to be Lipschitz (requiring the expert's predicted probability to be Lipschitz in the input, and the threshold crossing probability to be measure-zero)

The paper does not provide these conditions. \textbf{This is not a technical detail — it is key to whether the bound is non-vacuous.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{IV. Sign Errors: More Than One}\label{iv-sign-errors}

\subsubsection{4.1 Sign Error in CC Audit Report (Corrected, Confirmed)}\label{sign-error-in-cc-audit-report}

The sign error in Proposition 2.1 of the CC audit report is indeed wrong. Its \(\delta_s^{\PE} = (p_{\text{clean}}^{\text{PPE}} - p_{\text{clean}}) - (p_{\text{noisy}}^{\text{PPE}} - p_{\text{noisy}})\) has the physical meaning: encoding is only helpful when the consistency gain on clean samples exceeds the ``confusion gain'' on noise samples. This is counterintuitive to the notion that ``encoding should increase the noise detection margin.'' The paper's correction (swapping the order of the two terms) is correct.

\textbf{However}: The CC audit report internally has inconsistent notation. §0.2 defines \(p_{\text{clean}}\) as disagreement probability (\(<0.5\)), but §2.3's proof contains formulas like \(|p_{\text{clean}} + p_{\text{noisy}} - 1|\) that treat \(p_{\text{clean}}\) as a consensus probability. So the CC report's ``sign error'' may not simply be a flipped sign — it is a deeper \textbf{internal schism of the symbol system}: §0.2 and §2.3 use completely different semantics for \(p\).

The paper corrected the sign error but \textbf{did not point out the internal schism within the CC report} — it simply flipped the sign and claimed correction. In reality, the CC report's problem is more serious.

\subsubsection{4.2 Theorem 1.3.1: Massive Overestimation of Lipschitz Constant (New Derivation Error)}\label{theorem-1.3.1-lipschitz-constant}

This is not a sign error — it is a magnitude error. The paper claims: \[L_{\PE}^{\text{rot}} = 2\sqrt{\alpha^2 + \beta^2 + \gamma^2}\]

The actual Lipschitz constant is \(\max(\alpha, \beta, \gamma)\). Difference factor: approximately \(2\sqrt{3} \approx 3.46\) (when \(\alpha=\beta=\gamma\)).

\textbf{Derivation}: For \(\mathbf{e}_0 = (1,0,0,1,0,0,\ldots)^T\) (first dimension of each rotation plane is 1), \[\|\PE_{\text{rot}}(\mathbf{p}) - \PE_{\text{rot}}(\mathbf{q})\|^2 = 4\sin^2(\alpha\Delta x/2) + 4\sin^2(\beta\Delta y/2) + 4\sin^2(\gamma\Delta z/2)\]

As \(\|\Delta\mathbf{p}\| \to 0\): \[\|\PE_{\text{rot}}(\mathbf{p}) - \PE_{\text{rot}}(\mathbf{q})\| \approx \sqrt{\alpha^2\Delta x^2 + \beta^2\Delta y^2 + \gamma^2\Delta z^2}\]

By Cauchy-Schwarz: \[\sqrt{\alpha^2\Delta x^2 + \beta^2\Delta y^2 + \gamma^2\Delta z^2} \leq \max(\alpha, \beta, \gamma) \cdot \|\Delta\mathbf{p}\|_2\]

\textbf{The true Lipschitz constant is \(\max(\alpha, \beta, \gamma)\)}, and this bound is tight (take \(\Delta\mathbf{p}\) along the direction of the largest frequency parameter).

The error in the paper's proof lies in Step 2's triangle inequality decomposition (md line 300-317). The author correctly obtained: \[\|\mathbf{R}(\mathbf{p}) - \mathbf{R}(\mathbf{q})\|_F \leq 2\alpha|\Delta x| + 2\beta|\Delta y| + 2\gamma|\Delta z|\]

But the three difference matrices act on \textbf{mutually disjoint subspaces} (dimension pairs (0,1), (2,3), (4,5)). For matrix sum \(D = D_x + D_y + D_z\) on mutually disjoint subspaces: \[\|D\|_F^2 = \|D_x\|_F^2 + \|D_y\|_F^2 + \|D_z\|_F^2\]

The triangle inequality \(\|D\|_F \leq \|D_x\|_F + \|D_y\|_F + \|D_z\|_F\) can be loose by a factor of \(\sqrt{3}\) in the worst case (three equal-norm terms). Then the author introduces another \(\sqrt{3}\) relaxation via Cauchy-Schwarz (\(\ell_1 \to \ell_2\) conversion).

\textbf{The two relaxations multiply to produce a constant approximately 3.46 times too high.} Worse, the author noticed this problem (Note 1.3.1/md note 1.3.1 acknowledges that the effective constant in a single direction is only \(\alpha\) not \(2\alpha\)), yet still writes the loose bound as the ``exact constant'' in the theorem statement.

\textbf{Verdict}: The ``exact Lipschitz constant'' claimed in Theorem 1.3.1 is neither exact nor tight. The true exact Lipschitz constant (taking \(\|\cdot\|_2\) vector norm and \(\|\cdot\|_2\) position norm) is \(\max(\alpha, \beta, \gamma)\).

\subsubsection{4.3 Missing Normalization Factor in Theorem 1.2.1}\label{missing-normalization-factor}

The encoding kernel definition (Definition 1.2.1) is: \[\langle \PE_{\text{scalar}}(p), \PE_{\text{scalar}}(q) \rangle = \sum_{j=0}^{d/2-1} \cos(2\pi(p-q)/\lambda_j) = K_{\PE}(\Delta)\]

At \(\Delta = 0\), \(K_{\PE}(0) = d/2\). The target kernel \(k(0) = 1\).

As \(d \to \infty\), the encoding kernel \textbf{diverges} to infinity rather than converging to the target kernel. For the encoding kernel to approximate the target kernel, normalization is required: \[\frac{2}{d} K_{\PE}(\Delta) \to \int_0^\infty S(\omega)\cos(\omega\Delta)d\omega = k(\Delta)\]

But Theorem 1.2.1's statement and proof \textbf{completely omit} this \(2/d\) factor. The theorem's statement that ``the encoding kernel optimally approximates \(k(\Delta)\) in \(L^2([0,L])\)'' is mathematically wrong — \(K_{\PE}\) and \(k\) have different magnitudes (\(O(d)\) vs \(O(1)\)).

\subsubsection{4.4 Legitimacy of Pinsker Application in Theorem 2.3.1}\label{pinsker-application}

Theorem 2.3.1 uses Pinsker's inequality to relate \(\delta_s^{\PE}\) to KL divergence. But the application of Pinsker's inequality has a subtle issue:

Pinsker bound is: \(|P(A) - Q(A)| \leq \sqrt{\frac{1}{2}D_{\KL}(P\|Q)}\), where \(P\) and \(Q\) are \textbf{probability measures on the same measurable space} and \(A\) is an event in that space.

In the paper's application, \(P\) and \(Q\) are conditional distributions of \(\PE(P)\) given \(S\) (clean vs mixed), but the bounded quantity \(p_{\text{clean},s}^{\Situs} - p_{\text{clean},s}\) involves \textbf{the expert's output event}, which is not in the space of \(\PE(P)\).

Strictly, the correct argument is: \[p_{\text{clean},s}^{\Situs} = \int P(E_m(\phi(s)+z) \neq y) \cdot dP_{\PE(P)|S=s,\text{clean}}(z)\] \[p_{\text{clean},s} = P(E_m(\phi(s)) \neq y)\]

The second quantity does not even involve the distribution of \(\PE(P)\). The paper compares \(p_{\text{clean},s}^{\Situs}\) and \(p_{\text{clean},s}\), but Pinsker only constrains the distributional change of \(\PE(P)\), not directly the change in output probability — unless one additionally assumes the expert's output dependence on PE is Lipschitz (returning to the problem of Theorem 2.5.1).

\textbf{This may be fixable}: If we compare \(p_{\text{clean},s}^{\Situs}\) and \(p_{\text{clean},s}^{\text{avg}} = \int P(E_m(\phi(s)+z) \neq y) \cdot dP_{\PE(P)|S=s}(z)\) (i.e., clean disagreement rate averaged over PE), Pinsker can be legitimately applied. But the paper did not do this decomposition.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{V. Tightness: Claimed Tight but None Are Tight}\label{v-tightness}

\subsubsection{5.1 Theorem 1.2.3: Scalar Sine Lipschitz Constant}\label{theorem-1.2.3}

Claimed ``tight'' (line 414, md line 203). The derivative of each \((\sin,\cos)\) pair indeed satisfies \(\cos^2 + \sin^2 = 1\) identically for any \(p\), so for any two nearby points, the local rate of change of this pair indeed reaches \(\omega_j\). For the entire vector, all pairs act independently in orthogonal subspaces, so the Lipschitz constant is indeed \(\sqrt{\sum_j \omega_j^2 \cdot 2}\) = the claimed value. \textbf{This bound is indeed tight.} ✓

\subsubsection{5.2 Theorem 1.3.1: 3D Rotation Lipschitz Constant}\label{theorem-1.3.1}

Claimed ``worst-case'' tight (line 519, md line 337). As shown in §4.2, \textbf{not tight} — off by approximately 3.46x. The true tight Lipschitz constant is \(\max(\alpha, \beta, \gamma)\). ✗

\subsubsection{5.3 Theorem 1.2.2: Approximation Error Bound}\label{theorem-1.2.2}

Claims \(O(1/d)\) with constant \(\frac{2\pi^2\xi}{d} \cdot \frac{L}{\xi} = \frac{2\pi^2 L}{d}\). \textbf{No proof.} Moreover, \(\xi\) appears in both numerator and denominator and cancels, indicating that the constant \(\frac{2\pi^2 L}{d}\) is independent of the physical correlation length \(\xi\) — this is extremely suspicious. How can the approximation quality of a kernel be independent of the kernel's width?

Actual derivation (my own): For Riemann-Stieltjes sum approximation of \(\cos(\omega\Delta)\) on \([0,L]\) using quantile sampling, the error is controlled jointly by truncated high-frequency components and the coarsest low-frequency resolution. When \(d\) is fixed, the error should vary with \(\xi\) — small \(\xi\) (short-range correlation) means more high-frequency components, harder to approximate. The paper's claimed constant being independent of \(\xi\) is physically absurd.

\subsubsection{5.4 Theorem 2''s Bound}\label{theorem-2-bound}

Claims \(F_{1,\text{SCX}}^{\Situs} \leq F_{1,\text{base}} + C_F \cdot \sqrt{(\delta + 2\varepsilon_{\PE}/C_F^2)/2}\).

Note that when \(\varepsilon_{\PE} \to I(Y;P|X)\) (completely failed encoding): \[\delta_{\text{PPE}} = \delta + \frac{2I(Y;P|X)}{C_F^2}\]

The upper bound is looser than the original Theorem 2 by a factor of \(\sqrt{\delta + 2I/C_F^2} / \sqrt{\delta}\). When \(I(Y;P|X)\) is large (position has much information but encoding completely loses it), the upper bound can become arbitrarily loose. This is certainly not a ``tight'' bound — it is a \textbf{non-vacuous but potentially very loose} upper bound. The paper does not claim it is tight (this point is honest), but the abstract stating ``fixed the weak feature failure upper bound'' implies some precision, when in reality it merely added a positive term to an already loose bound.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{VI. Theorem 3''s Constructive Example: Shooting Itself in the Foot}\label{vi-theorem-3-constructive-example}

\subsubsection{6.1 Contradiction Between the Example's Claim and Reality}\label{contradiction-between-claim-and-reality}

§4.3 (paper line 1096-1137) is titled ``Constructive Minimal Example.'' What does this ``constructive minimal example'' actually demonstrate?

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 Constructs two worlds \(W_A, W_B\), verifies \(P_{W_A}(X,Y) = P_{W_B}(X,Y)\) ✓
\item
 Introduces learnable PE \(\PE_\theta(p) = \theta \cdot p\)
\item
 Computes \(\mathbb{E}_{W_A}[Y|X] = \mathbb{E}_{W_B}[Y|X] = 0.6 \cdot \text{sign}(x)\) — \textbf{numerically identical}
\item
 Conclusion: \textbf{Even learnable PE cannot distinguish the two worlds} (paper line 1131-1132 original: ``therefore even learnable PE cannot distinguish the two worlds. This precisely confirms the essence of Theorem 3'')
\end{enumerate}

\textbf{This is a ``constructive counterexample'' — it constructs an example where learnable PE cannot distinguish, exactly proving Theorem 3's robustness rather than its fragility!}

\subsubsection{6.2 Hand-Waved ``Truly Distinguishable Conditions''}\label{hand-waved-conditions}

After proving that learnable PE cannot distinguish, the paper (line 1134-1136) hand-waves: \textgreater{} ``However, when the auxiliary loss involves unobserved physical quantity \(Z\) — e.g., in \(W_A\) noisy labels cause large \(|Z_{\text{pred}} - Z_{\text{DFT}}|\) (physically inconsistent), while in \(W_B\) the difference is small (difficult but physically consistent) — then the optimal parameters of learnable PE differ between the two worlds, and unidentifiability is broken.''

This is \textbf{not a constructive proof} — it verbally describes an imagined scenario. No concrete definition of \(Z\) is given, no concrete form of \(\mathcal{L}_{\text{aux}}\) is given, and no verification that the two worlds indeed yield different optima is provided. The entire §4.3 promises a ``constructive minimal example'' but delivers only an \textbf{empty claim}.

\subsubsection{6.3 Final Verdict on Theorem 3'}\label{final-verdict-on-theorem-3}

Theorem 3' itself is a reasonably conditionalized theorem statement (if preconditions (A)(B) hold, then unidentifiability is preserved), but: - The ``distinguishability'' conclusion when precondition (A) fails \textbf{has no constructive proof support} - The ``constructive minimal example'' is actually the theorem's \textbf{counterexample's counterexample} — i.e., it demonstrates the theorem's robustness - The actual mechanism for distinguishing noise and difficult samples (multi-task learning + physical consistency) is not formalized, not proven, not constructed

\textbf{The ``breaking'' part of Theorem 3' is an unproven conjecture.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{VII. Summary of Fatal Defects}\label{vii-summary-of-fatal-defects}

\subsubsection{Fatal Defect \#1: Fano Derivation in Theorem 2.4.1}\label{fatal-defect-1}

\textbf{Nature}: Logical error (misuse of inequality direction)\\
\textbf{Location}: main.tex §3.4 (line 746-776), md §2.4 (line 557-617)\\
\textbf{Specific Error}: Deriving a lower bound on error rate difference from two Fano lower bounds. \(a \geq A\) and \(b \geq B\) do not imply \(a-b = A-B\) or \(a-b \geq A-B\).\\
\textbf{Fixability}: Not fixable. Requires a completely different proof strategy. Possible directions are strong Fano or Fano's inverse, but these require concrete distributional assumptions.\\
\textbf{Consequence}: Theorem 2.4.1 (Fano lower bound) cannot be called a theorem — it is at most a heuristic estimate.

\subsubsection{Fatal Defect \#2: Lipschitz Constant in Theorem 1.3.1}\label{fatal-defect-2}

\textbf{Nature}: Constant factor overestimated by approximately 3.46x\\
\textbf{Location}: main.tex §2.3 (line 483-492), md §1.3.4 (line 269-276)\\
\textbf{Specific Error}: Double relaxation of triangle inequality + Cauchy-Schwarz should simplify to direct Frobenius norm computation since subspaces are orthogonal.\\
\textbf{Fixability}: Fixable. The correct Lipschitz constant is \(\max(\alpha, \beta, \gamma)\).\\
\textbf{Consequence}: Theorem 1.4.1 (Unified Lipschitz) and Theorem 2.5.1 (Spatial smoothness) both use this overestimated constant. The spatial smoothness bound is actually approximately 3.5 times tighter than claimed — this is good news, but indicates the paper's analysis is imprecise.

\subsubsection{Fatal Defect \#3: Bayes→Actual Expert Chasm in Theorem 2.2.1}\label{fatal-defect-3}

\textbf{Nature}: Unbridgeable logical gap\\
\textbf{Location}: main.tex line 659-661, md line 484\\
\textbf{Specific Error}: Behavior of Bayes optimal classifier cannot be generalized to actually trained neural network experts. No finite-sample guarantee.\\
\textbf{Fixability}: Possibly patchable by assuming the expert model class has some uniform approximation property, but this would essentially change the theorem's nature (from information-theoretic to learning-theoretic).\\
\textbf{Consequence}: Theorem 2.2.1 (the paper's most important connection theorem), upon removing Step 4, degenerates to: \(I(Y;P|S) > 0\) implies the Bayes optimal classifier performs better when using \(P\) information. This is trivial, not a contribution.

\subsubsection{Fatal Defect \#4: Missing Normalization Factor in Theorem 1.2.1}\label{fatal-defect-4}

\textbf{Nature}: Derivation omission\\
\textbf{Location}: main.tex §2.2.2, md §1.2.2\\
\textbf{Specific Error}: \(K_{\PE}(0) = d/2\) while \(k(0) = 1\). Without normalization, convergence is impossible.\\
\textbf{Fixability}: Fixable. Add \(1/\sqrt{d/2}\) factor in encoding definition or \(2/d\) normalization in inner product. But all subsequent constants (Lipschitz, error bounds) need corresponding adjustment.\\
\textbf{Consequence}: Theorem 1.2.1's ``optimal approximation'' statement and Theorem 1.2.2's approximation error bound are wrong without normalization.

\subsubsection{Semi-Fatal Defect \#5: No Constructive Proof for Theorem 3''s ``Breaking'' Direction}\label{semi-fatal-defect-5}

\textbf{Nature}: Unsupported claim\\
\textbf{Location}: main.tex §4.3, md §4.3\\
\textbf{Specific Error}: The ``constructive minimal example'' instead proves Theorem 3's robustness.\\
\textbf{Fixability}: Requires a true example showing learnable PE can produce different parameters on observationally equivalent data. This requires the auxiliary loss to involve unobserved variables — such an example is possible but \textbf{has not been constructed}.\\
\textbf{Consequence}: The ``if precondition (A) fails'' part of Theorem 3' is currently an unproven conjecture.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{VIII. Additional Issues: Small but Real}\label{viii-additional-issues}

\subsubsection{\texorpdfstring{8.1 Unclear Origin of \(\min(1-p_{\text{clean},s}, \ldots)\) in Theorem 2.3.1}{8.1 Unclear Origin}}\label{unclear-origin}

Why does the upper bound have a \(1-p_{\text{clean},s}\) term? The derivation (line 708-719) from Pinsker + triangle inequality gives: \[\delta_s^{\PE} \leq \sqrt{\frac{1}{2}D_{\KL}^{(1)}} + \sqrt{\frac{1}{2}D_{\KL}^{(2)}}\] Then suddenly becomes \(\min(1-p_{\text{clean},s}, \ldots)\). Where did \(1-p_{\text{clean},s}\) come from? It looks like the range upper bound of \(\delta_s^{\PE}\) (from \(\delta_s^{\PE} \in [-\Delta_s, 1-p_{\text{clean},s}]\)), but what is the mathematical basis for taking \(\min\) of the range upper bound and the Pinsker upper bound? If Pinsker gives 0.3 and the range upper bound gives 0.7, why is the upper bound 0.3 not 0.7? One needs to argue that the Pinsker bound \(\leq\) range upper bound always holds, or explain why both Pinsker and range upper bounds constitute valid upper bounds. Currently this \(\min\) is introduced without logic.

\subsubsection{\texorpdfstring{8.2 Unit Ambiguity of \(\log 2\) in Theorem 2.4.1}{8.2 Unit Ambiguity}}\label{unit-ambiguity}

The paper's Fano lower bound contains \(\frac{2\rho I - \log 2}{\log|\Y|}\). \(I(\cdot;\cdot)\) is typically in nats (if using natural log), what unit is \(\log 2\)? If unified in nats, \(H_2(P_e) \leq \ln 2\) nats; if unified in bits, \(H_2(P_e) \leq 1\) bit, no \(\log 2\). The appearance of \(\log 2\) indicates the paper switches between nats and bits without declaration. This is unacceptably sloppy for an information theory paper.

\subsubsection{8.3 Theorem 1.4.1 is Pure Packaging}\label{theorem-1.4.1-pure-packaging}

Theorem 1.4.1 aggregates two Lipschitz constants in a table and claims it is a ``Unified Lipschitz Continuity Theorem.'' It contains no new mathematical content — it is merely the union of two constant definitions. This is not a theorem; it is a definition or summary table.

\subsubsection{8.4 Exponential Decay Claim in Proposition 3.1 (Dimension-Information Tradeoff)}\label{exponential-decay-claim}

\[\varepsilon_{\PE} \leq I(Y; P \mid X) \cdot \exp(-d/d_0)\]

Labeled \texttt{{[}heuristic{]}}. Exponential decay holds for Fourier series approximation of analytic kernels (Cauchy) under certain smoothness conditions, but what is being approximated here is the kernel \(k(\Delta)\), and the information loss \(\varepsilon_{\PE}\) involves conditional distributions \(P_{Y|X,P}\) and the KL divergence caused by encoding. From kernel approximation error to information loss, at least two Pinsker applications and assumptions about the smoothness of \(P_{Y|X,P}\) are needed. These assumptions are not stated. The exponential decay functional form \(\exp(-d/d_0)\) is at the level of a guess.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{IX. Bonus: Additional Issues within the CC Audit Report}\label{ix-bonus}

Although this review primarily targets the paper body and ppe\_rigorous\_derivation.md, there is one issue in the CC audit report worth pointing out that the paper did not correct:

\subsubsection{9.1 Internal Schism in Detection Margin Definition in CC Audit Report Proposition 2.1}\label{internal-schism}

CC report §0.2 defines \(p_{\text{clean}} < 0.5\) as disagreement probability. But §2.3's proof suddenly contains: \[\Delta_s^{\text{PPE}} = |p_{\text{clean},s}^{\text{PPE}} - (1 - p_{\text{noisy},s}^{\text{PPE}})| = |p_{\text{clean},s}^{\text{PPE}} + p_{\text{noisy},s}^{\text{PPE}} - 1|\]

If \(p_{\text{noisy}}\) is disagreement probability (\(>0.5\)), \(1-p_{\text{noisy}}\) is consensus probability (\(<0.5\)), then \(\Delta = |\text{disagreement}_{\text{clean}} - \text{consensus}_{\text{noisy}}|\) — this is not the standard definition of detection margin. The standard definition should be \(|\text{disagreement}_{\text{noisy}} - \text{disagreement}_{\text{clean}}|\). This indicates the CC report switched the semantics of \(p_{\text{noisy}}\) (disagreement → consensus) in §2.3 without declaration. The paper's ``correction'' merely flipped the sign of \(\delta_s^{\PE}\) and did not touch this deeper notational schism.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{X. Final Score}\label{x-final-score}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Score
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Comment
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Problem Importance & 8/10 & Distinguishing physical PE and statistical PE is a genuine and important problem \\
Theoretical Ambition & 9/10 & Establishing rigorous foundations on four levels is ambitious \\
Theorem 1.2.1-1.4.1 & 5/10 & Core ideas correct, but 1.3.1 constant error, 1.2.1 missing normalization \\
Theorem 2.2.1 & 4/10 & Steps 1-3 correct but trivial, Step 4 unbridgeable \\
Theorem 2.3.1 & 5/10 & Direction correct but details problematic (Pinsker application, unclear min term origin) \\
Theorem 2.4.1 & \textbf{1/10} & \textbf{Proof is logically erroneous, cannot be called a theorem} \\
Theorem 2.5.1 & 4/10 & Lipschitz assumption invalid on indicator functions \\
Theorem 2' & 7/10 & Information-theoretic definition + algebraic derivation correct; best part of the paper \\
Theorem 3' & 3/10 & Conclusion under valid preconditions is correct but trivial; breaking-direction conclusion has no proof \\
Honesty & 6/10 & Heuristic vs rigorous annotations present, but 2.4.1 annotation insufficiently honest (labeled partially rigorous when it is wholly wrong) \\
\textbf{Overall} & \textbf{4/10} & \textbf{Contains one fatal logical error (2.4.1), one magnitude error (1.3.1), one unbridgeable generalization (2.2.1 Step 4). Unsuitable for publication before these are fixed.} \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{XI. Minimum Required Fixes for Revision}\label{xi-minimum-required-fixes}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{Theorem 2.4.1}: Remove the ``lower bound'' claim, re-label as ``heuristic estimate'' or delete entirely. If retained, must be reproved using a different method. The Fano equality assumption must be replaced with an explicit construction under distributional assumptions.
\item
 \textbf{Theorem 1.3.1}: Correct the Lipschitz constant to \(\max(\alpha, \beta, \gamma)\), and correct all subsequent results depending on this constant (Theorem 1.4.1, Theorem 2.5.1).
\item
 \textbf{Theorem 1.2.1}: Add normalization factor \(1/\sqrt{d/2}\) in the encoding definition (or equivalently, \(2/d\) in the inner product), and correspondingly adjust the error bound constants in Theorem 1.2.2.
\item
 \textbf{Theorem 2.2.1}: Explicitly restrict the conclusion ``there exists \(\PE\) and \(d\) such that \(\delta_s^{\PE} > 0\)'' to ``holds for Bayes optimal classifier.'' Generalization to actual experts requires a separate learning-theoretic analysis (possibly involving VC dimension, Rademacher complexity, etc.), or explicitly label as a conjecture.
\item
 \textbf{Theorem 3'}: Either provide a genuine constructive example (demonstrating learnable PE leading to distinguishable parameters on observationally equivalent data), or change the ``breaking direction'' conclusion to an open problem.
\item
 \textbf{Theorem 2.3.1}: Clarify the origin of the \(\min\) term, fix the Pinsker application conditions.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Reviewer signature: No need to sign. These problems are objective.}

\emph{P.S. This paper is not without value. The encoding function theory (§2) and information-theoretic definitions (§3.1-3.2) are solid. But Theorem 2.4.1's Fano derivation is a fatal flaw — you cannot treat two inequality signs as equality signs and then claim to give a lower bound. Theorem 1.3.1's 3.5x overestimation indicates the author did not fully verify each step of their own derivation. These issues must be corrected before submission to Annals of Statistics. The current state is more suitable as a technical report.}

\end{document}
