% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Spring Self-Evolution Algorithm ---
Mathematical Genealogy}\label{spring-ux81eaux8fdbux5316ux7b97ux6cd5-ux6570ux5b66ux8c31ux7cfbux5b66}

\begin{quote}
\textbf{Spring} = Spring Self-Evolving Gatekeeper (SSEG)
\textbf{Status}: Paper 2 (Nature Computational Science title) core algorithm
\textbf{Document Type}: Theoretical Genealogy ---
Mathematical domain classification, key idea origins, historical development timeline, and comparison with prior work
\textbf{Date}: 2026-06-28
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Table of Contents}\label{ux76eeux5f55}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \hyperref[1-ux6570ux5b66ux9886ux57dfux5f52ux5c5e]{Mathematical Domain Classification}
\item
  \hyperref[2-ux5173ux952eux601dux60f3ux4e0eux8bc1ux660eux5de5ux5177]{Key Ideas and Proof Tools}
\item
  \hyperref[3-ux5386ux53f2ux53d1ux5c55ux65f6ux95f4ux7ebf]{Historical Development Timeline}
\item
  \hyperref[4-spring-ux4e0eux524dux4ebaux7684ux5bf9ux6bd4ux8868]{Spring
  Comparison with Prior Work}
\item
  \hyperref[5-spring-ux72ecux6709ux7684ux8d21ux732e]{Spring's Unique Contributions}
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Mathematical Domain Classification}\label{ux6570ux5b66ux9886ux57dfux5f52ux5c5e}

Spring
The Spring self-evolution algorithm does not belong to a single mathematical domain, but spans seven classical mathematical domains, synthesizing a completely new structure at their intersection.

\subsubsection{1.1 Stochastic Approximation (Stochastic
Approximation)}\label{ux968fux673aux903cux8fd1-stochastic-approximation}

\textbf{Origin}: Robbins \& Monro (1951). ``A stochastic approximation
method.'' \emph{Annals of Mathematical Statistics}, 22(3), 400-407.

\textbf{Role in Spring}: NEP student parameter \(\theta_t\) 'sonline update follows
Robbins-Monro recursion:

\[\theta_{t+1} = \theta_t - \alpha_t \nabla_\theta \ell(f_{\theta_t}(x_t), y_t)\]

where the learning rate \(\alpha_t\) satisfies the standard RM
conditions:\(\sum \alpha_t = \infty\), \(\sum \alpha_t^2 < \infty\). 

Spring generalizes the standard RM to\textbf{coupled two-timescale}setting:  -
\textbf{Fast timescale}: student  \(\theta_t\) at fixed  Gatekeeper \(S_t\)
under with Rate \(\alpha_t\) Update - \textbf{Slow timescale}: Gatekeeper \(S_t\)
under the student quasi-steady-state at rate \(\beta_t = o(\alpha_t)\) Update

This corresponds to Borkar (2008)'s two-timescale stochastic approximation framework, where condition C6'
(\(\beta_t = o(\alpha_t)\))
is the key innovation---it simultaneously guarantees local convergence of the student and bounded cumulative distribution shift.

\textbf{References}: - Robbins \& Monro (1951) - Kushner \& Yin (2003).
\emph{Stochastic Approximation and Recursive Algorithms and
Applications} - Borkar (2008). \emph{Stochastic Approximation: A
Dynamical Systems Viewpoint} - Polyak \& Juditsky (1992). ``Acceleration
of stochastic approximation by averaging''

\subsubsection{1.2 Dynamical Systems / Lyapunov
Stability}\label{ux52a8ux529bux7cfbux7edf-lyapunov-ux7a33ux5b9aux6027}

\textbf{Origin}: Lyapunov, A. M. (1892). \emph{The General Problem of the
Stability of Motion}. Kharkov Mathematical Society.

\textbf{Role in Spring}: Spring self-evolution loopis adiscrete dynamical system: 

\[z_{t+1} = \Phi(z_t), \quad z_t = (S_t, M_t, \theta_t) \in \mathcal{Z}\]

where \(\Phi = (\Phi_S, \Phi_M, \Phi_\theta)\)
Yesupdate operator, \(\mathcal{Z} = \mathcal{F} \times \mathcal{M} \times \Theta\)
Yesevolution state space. 

Lyapunov Functioncandidate as (On a fixed reference set \(M_0\) on evaluation):

\[\Psi(S_t, \theta_t) = \frac{1}{|M_0|} \sum_{x \in M_0} (S_t(x, y(x)) - \hat{C}(x))^2 + \lambda \cdot \frac{1}{|V_0|} \sum_{(x,y) \in V_0} \ell(f_{\theta_t}(x), y)\]

Lyapunov under descent property
\(\mathbb{E}[\Psi_{t+1} \mid \mathcal{F}_t] \leq \Psi_t\) via  Theorem
12.5 at reference set replayMechanismunder is CompleteProof. non-point  \((S^*, \theta^*)\)
satisfies consistent coupling condition, \(\omega\)-limitset as point . 

Spring
make using the dynamicalSystem'sCompletetoolchain: attractors, basins, phase diagram (four convergencemechanisms), \(\omega\)-limit set.

\textbf{References}: - Lyapunov (1892). \emph{The General Problem of the
Stability of Motion} - Khalil (2002). \emph{Nonlinear Systems} (3rd ed.)
- Strogatz (2018). \emph{Nonlinear Dynamics and Chaos} (2nd ed.) -
Conley (1978). \emph{Isolated Invariant Sets and the Morse Index}

\subsubsection{1.3 Bayesian Inference /
Martingale Theory}\label{ux8d1dux53f6ux65afux63a8ux65ad-ux9785ux7406ux8bba}

\textbf{Origin}: Doob, J. L. (1953). \emph{Stochastic Processes}. Wiley.

\textbf{Role in Spring}: Gatekeeper scoring function \(S_t\)
'sevolutionis explain as Bayesian posterior update: 

\[S_t(x,y) = \mathbb{E}_{S \sim P_t}[S(x,y)], \quad P_{t+1}(S) \propto P_t(S) \cdot \mathcal{L}(\mathcal{D}_{t+1} \mid S)\]

theorem  (Doob Martingale Convergence): order sequence  \(\{S_t\}_{t \geq 0}\) Yesregarding factor 
\(\mathcal{F}_t = \sigma(\mathcal{M}_1, \dots, \mathcal{M}_t)\) 's
\([0,1]\)-bounded martingale. by  Doob Martingale Convergencetheorem : 

\[S_t(x,y) \xrightarrow{a.s.} S_\infty(x,y), \quad \forall (x,y)\]

Important limitation : Martingale Propertymust model is correct(\(S^*\) in the prior \(P_0\)
'ssupport). at model misspecificationunder , Convergenceto  KL projection \(S^\dagger\) rather than the true value
\(S^*\). Spring honestly acknowledges this limitation.

\textbf{References}: - Doob (1953). \emph{Stochastic Processes} - Doob
(1949). ``Application of the theory of martingales'' - Ghosh \&
Ramamoorthi (2003). \emph{Bayesian Nonparametrics} - van der Vaart
(1998). \emph{Asymptotic Statistics} - Kleijn \& van der Vaart (2012).
``The Bernstein-von Mises theorem under misspecification''

\subsubsection{1.4 at Online Learning /
Regret Analysis}\label{ux5728ux7ebfux5b66ux4e60-ux9057ux61beux5206ux6790}

\textbf{Origin}: Zinkevich, M. (2003). ``Online convex programming and
generalized infinitesimal gradient ascent.'' \emph{ICML 2003}.

\textbf{Role in Spring}: Gatekeeper 'sat Online under  (OGD)
update yieldsstandard regret bound: 

\[\text{Regret}_T \leq 2GR\sqrt{T} \quad \text{(No Delay)}\]

Spring generalizes this bound to the\textbf{delayed feedback}setting (the NEP student's feedback delay
\(d_t \leq D_{\max}\)):

\[\text{Regret}_T^{\text{delay}} \leq 2GR\sqrt{T} + G^2 D_{\max} \sqrt{T}\]

Further generalized to the\textbf{growing memory bank}setting------memory bank \(M_t\)
'smonotonically growingprovide has sample complexityimprovement: 

\[\text{Regret}_T^{\text{eff}} \leq \text{Regret}_T^{\text{delay}} + O\left(T \sqrt{\frac{\text{VC}(\mathcal{H})}{N_T}}\right)\]

As
\(N_T \to \infty\), Systemsatisfies \textbf{Asymptotic No-Regret}: \(\frac{1}{T}\text{Regret}_T^{\text{eff}} \to 0\). 

\textbf{References}: - Zinkevich (2003). ``Online convex programming'' -
Hazan (2016). ``Introduction to online convex optimization'' - Joulani
et al.~(2016). ``Online learning under delayed feedback'' -
Shalev-Shwartz (2012). ``Online learning and online convex
optimization'' - Cesa-Bianchi \& Lugosi (2006). \emph{Prediction,
Learning, and Games}

\subsubsection{1.5 concentration inequalities}\label{ux96c6ux4e2dux4e0dux7b49ux5f0f}

\textbf{Origin}: - Hoeffding, W. (1963). ``Probability inequalities for
sums of bounded random variables.'' \emph{JASA}, 58(301), 13-30. -
Chernoff, H. (1952). ``A measure of asymptotic efficiency for tests of a
hypothesis.'' \emph{Annals of Mathematical Statistics}, 23(4), 493-507.

\textbf{Role in Spring}: at  Spring 'smultiple critical locations: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Concentration of the Consensus Score}: Expert consensus
  \(\hat{C}(x) = \frac{1}{M}\sum_{m=1}^M \mathbf{1}\{\ell(f_m(x), y) > \tau\}\)
  Hoeffding tail bound:
  \[\mathbb{P}(|C - \mathbb{E}[C]| > t) \leq 2\exp(-2M t^2)\] When A2
  is violated due to expert correlation,\(M\) is replaced by the effective number of experts
  \(M_{\text{eff}} = M/(1 + (M-1)\bar{\rho})\). 
\item
  \textbf{Chernoff information}: detection problem'soptimal exponentialRateby  Chernoff information
  \(\kappa = \text{KL}(\theta^*\|p_0) = \text{KL}(\theta^*\|p_1)\)
  , applied to Theorem 4'.
\item
  \textbf{Bahadur-Rao Exact Asymptotics}: Bernoulli lattice correctionbecause factor 
  \((1-e^{-\lambda^*})^{-1}\) replaces the standard Chernoff prefactor, tightening Theorem 4'
  's exact constant by 5--12\%.
\end{enumerate}

\textbf{References}: - Hoeffding (1963) - Chernoff (1952) - Bahadur \& Rao
(1960). ``On deviations of the sample mean'' - Dembo \& Zeitouni (2010).
\emph{Large Deviations Techniques and Applications}

\subsubsection{1.6 Information Theory}\label{ux4fe1ux606fux8bba}

\textbf{Origin}: - Fano, R. M. (1961). \emph{Transmission of Information:
A Statistical Theory of Communications}. MIT Press. - Cover, T. M. \&
Thomas, J. A. (2006). \emph{Elements of Information Theory} (2nd ed.).
Wiley.

\textbf{Role in Spring}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Fano non-etc.formula }: using  Theorem 2(Weak Feature Failure)'slower bound------when feature
  \(\phi\) and true state \(S\) have mutual information \(I(\phi; S) \leq \delta\)
  , any state estimationAlgorithmhas : 
  \[P(\hat{S} \neq S) \geq \frac{H(S) - \delta - \log 2}{\log K}\]
\item
  \textbf{Data Processingnon-etc.formula  (DPI)}: using  Theorem 2 Proof Step
  3------the Markov chain \(\phi(X) \to \hat{S} \to \hat{z}\) means  TV
  does not increase.
\item
  \textbf{Pinsker non-etc.formula }: will mutual information bound \(\delta = I(\phi; S)\)
  to a total variation bound \(\text{TV} \leq \sqrt{\delta/2}\). 
\item
  \textbf{KL contraction}: at correct modelunder , Convergence
  \(D_{\text{KL}}(P^* \| P_t) \xrightarrow{a.s.} 0\). 
\end{enumerate}

\textbf{References}: - Fano (1961) - Cover \& Thomas (2006) - Pinsker
(1964). \emph{Information and Information Stability of Random Variables
and Processes}

\subsubsection{1.7 Minimax Theory}\label{minimax-ux7406ux8bba}

\textbf{Origin}: Le Cam, L. (1973). ``Convergence of estimates under
dimensionality restrictions.'' \emph{Annals of Statistics}, 1(1), 38-53.

\textbf{Role in Spring}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Le Cam two-point method}: construct with distinguishable distributions
  \(P_0 = \text{Bern}(\mu_s)^{\otimes M}\)(Clean)and 
  \(P_1 = \text{Bern}(1 - C_{\text{bal}}\mu_s/(K-1))^{\otimes M}\)(Noise), using using 
  Hellinger tensorization to obtain  minimax lower bound. 
\item
  \textbf{Neyman-Pearson reduction}: through  Neyman-Pearson lemma will ArbitraryAlgorithmreducboundas 
  Bayes checktest, give  Theorem 4' 'sOptimalproperty . 
\item
  \textbf{Hellinger Exact Tensorization:}: at 
  A2(Conditional Independence)under , \(H^2(P_0^{\otimes M}, P_1^{\otimes M}) = 1 - (1 - H^2(P_0, P_1))^M\)------precise etc.formula , Not an inequality relaxation. 
\end{enumerate}

\textbf{References}: - Le Cam (1973) - Le Cam (1986). \emph{Asymptotic
Methods in Statistical Decision Theory} - Tsybakov (2009).
\emph{Introduction to Nonparametric Estimation}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.
Key Ideas and Proof Tools}\label{ux5173ux952eux601dux60f3ux4e0eux8bc1ux660eux5de5ux5177}

\subsubsection{\texorpdfstring{2.1 Two-Timescale Separation
(\(\beta_t = o(\alpha_t)\))}{2.1 Two-Timescale Separation (\textbackslash beta\_t = o(\textbackslash alpha\_t))}}\label{ux53ccux65f6ux95f4ux5c3aux5ea6ux5206ux79bb-beta_t-oalpha_t}

Spring key 'sarchitectural innovation. Condition C6' must  Gatekeeper update rate \(\beta_t\)
Yesstudent update rate \(\alpha_t\) 'slower-order infinitesimal:

\[\beta_t = o(\alpha_t), \quad \text{i.e.} \quad \lim_{t\to\infty} \frac{\beta_t}{\alpha_t} = 0\]

\textbf{Purpose: 's}: Between any two Gatekeeper
updates, student performs asymptotically infinitely many gradient steps, at before distribution shift, convergingto when Distribution'slocal minimum value . 

\textbf{Canonical Schedule}: \(\alpha_t = t^{-a}\), \(\beta_t = t^{-b}\), where
\(\frac{1}{2} < a < 1 < b\). e.g.such as  \(a = 0.6\), \(b = 1.2\):  -
\(\sum \alpha_t = \infty\)(student exploration), \(\sum \alpha_t^2 < \infty\)(noise control)
-
\(\sum \beta_t < \infty\)(has finite totalDistribution Shift), \(\sum \beta_t^2 < \infty\)(variance control)
- \(\beta_t / \alpha_t = t^{-0.6} \to 0\)(Timescale Separation)

\subsubsection{2.2
Reference Set Replay BreaksSelection Bias}\label{ux53c2ux8003ux96c6ux91cdux653eux6253ux7834ux9009ux62e9ux504fux5dee}

Spring mostdeepcharacterize'sProoftool. Theorem 12.2
Proof: \textbf{Without reference set replay, Lyapunov
under at asymptotically cannot be 's}------Gatekeeper
can reduce its apparent loss by only accepting''easy''samples to reduce apparent loss, while not truly improving discrimination ability . 

Resolution (Theorem 12.5, CompleteProof): 1. \textbf{Student Side}: importance sampling weights
\(w(x) = P_0(x) / P_{S_t}(x)\), make has training distributionas reference distribution \(P_0\) 2.
\textbf{Gatekeeper Side}: On a fixed reference set \(M_0\) on Computation
SCXUpdate, eliminating alignment bias 3. \textbf{Cost}: Importance weight variance inflation factor
\(W \leq 1/\varepsilon_{\text{explore}} < \infty\)

\subsubsection{2.3 Lipschitz-Almost-Everywhere
Lemma}\label{lipschitz-almost-everywhere-ux5f15ux7406}

Resolves a fundamental mathematical obstacle (Lemma SE-1.0, B9 fix):

\textbf{Problem}: consensus score
\(C(x) = \frac{1}{M}\sum_{m=1}^M \mathbf{1}\{\ell(f_m(x), y) > \tau\}\)
at the threshold \(\tau\) non-continuous (indicator function jump). 

\textbf{Resolution}: at conditions C5(conditions i.i.d.
sampling), the discontinuity set has probability measure zero, because:
\[\mathbb{P}(\exists m: \ell(f_m(X), Y) = \tau) \leq \sum_{m=1}^M \mathbb{P}(\ell(f_m(X), Y) = \tau) = 0\]
for continuous Distributionon 'ssmoothFunction
\(\ell(f_m(x), y)\). expectationnot affected byzero-measure set modification'sinfluence , Lyapunov Analysisconservretainhas effect. 

\subsubsection{2.4 BayesiantestUpdateas  Gatekeeper
Evolution}\label{ux8d1dux53f6ux65afux540eux9a8cux66f4ux65b0ux4f5cux4e3a-Gatekeeper-ux8fdbux5316}

Gatekeeper 'sEvolutionis explain as order Bayesian update: 

\[P_{t+1}(S) \propto P_t(S) \cdot \prod_{(x,y) \in \mathcal{D}_{t+1}} S(x,y)^{a} (1 - S(x,y))^{1-a}\]

this reinforce'stheoretical properties:  1. \textbf{Martingale Property}:
\(\mathbb{E}[S_{t+1} \mid \mathcal{F}_t] = S_t\)(Tower Property), by  Doob
theorem guarantees a.s. Convergence 2. \textbf{KL contraction}:
at correct modelunder , \(D_{\text{KL}}(P^* \| P_t) \to 0\)

\subsubsection{2.5 Lyapunov
Descent with Importance Sampling}\label{lyapunov-ux4e0bux964dux4e0eux91cdux8981ux6027ux91c7ux6837}

Theorem 12.5 'scomplete proof structure establishes
\(\mathbb{E}[\Delta\Psi_t \mid \mathcal{F}_t] \leq 0\): 

\[\mathbb{E}[\Delta\Psi_t \mid \mathcal{F}_t] \leq -\alpha_t \|\nabla L_0(\theta_t)\|^2 - \beta_t \cdot 2\rho_{\text{ideal}} \cdot \|\Delta_t^{\text{ideal}}\|_{M_0} \cdot \|S_t - \hat{C}\|_{M_0} + O(\alpha_t^2 W + \beta_t^2)\]

Decomposition into four terms: - \textbf{(A) Student Improvement}:
\(\Delta_{\text{student}}\), eliminating via importance sampling \(D_{\text{static}}\)
gap (\textbf{already Proof}) - \textbf{(B) Gatekeeper Improvement}:
\(\Delta_{\text{Gatekeeper}}\), by computing \(M_0\) on Computation SCXUpdate
eliminating alignment bias(\textbf{already Proof}) - \textbf{(C) Distribution Shift}:
\(\Delta_{\text{selection}} = O(\beta_t) = o(\alpha_t)\), controlled by two-timescale (\textbf{already Proof})
- \textbf{(D) Cross-Coupling}:
\(\Delta_{\text{cross}} = 0\), due to reference set separability (\textbf{already Proof})

\subsubsection{\texorpdfstring{2.6 \(O(t^{-a})\) Convergencerate and  Polyak
averageall }{2.6 O(t\^{}\{-a\}) Convergencerate and  Polyak all }}\label{ot-a-ux6536ux655bux7387ux4e0e-polyak-ux5e73ux5747}

at reinforceconvexconditionsunder , student 'sConvergencerate as  \(O(t^{-a})\), where \(a \in (0.5, 1]\)
as learning rate exponent. 

\textbf{Polyak-Ruppert Averaging}
(\(\bar{\theta}_t = \frac{1}{t}\sum_{i=1}^t \theta_i\)) will Rateboost to 
\(O(t^{-1})\)------and  minimax lower bound \(\Omega(t^{-1})\)
matching'sOptimalRate------for any \(a \in (0.5, 1)\) and Norequires precise tuneharmonic. 

\subsubsection{2.7
Characterization of Four Failure Modes}\label{ux56dbux79cdux5931ux6548ux6a21ux5f0fux523bux753b}

Spring systematically characterizesfour canonical failure modes: 

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mechanism
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Severity
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{1. Premature Freezing} & \(\beta_t\) decays too quickly before the student provides corrective signals &
in etc.(doubleAccuracyunder rare ) \\
\textbf{2. Backlog} & Data arrivalRate Gatekeeper evaluatethroughput quantity  &
(datasetunder actualexistat ) \\
\textbf{3. Client Divergence} & multiple non-point  + non-initial conditions/data streams →
irreconcilable divergence & Medium (multi-client deployment only) \\
\textbf{4. Adversarial Poisoning} & for adversaryNote carefulcarefullyconstruct 'sSample; self-evolution loopamplifycontamination &
High (severe with amplification) \\
\end{longtable}

each Modehas formula ization conditions, ization Rateand mitigationsolutionstrategy . 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3.
Historical Development Timeline}\label{ux5386ux53f2ux53d1ux5c55ux65f6ux95f4ux7ebf}

\begin{verbatim}
1892 ─ Lyapunov StabilityTheory
      A. M. Lyapunov, The General Problem of the Stability of Motion
      Systemproperty Analysis'sFoundations---non-point , attractdirectfactor , Lyapunov Function
      ↓ Spring make using : Ψ(S_t, θ_t) as  Lyapunov Function; non-point Analysis; attractdirectfactor basin
      
1933 ─ Kolmogorov Strong Law of Large Numbers
      A. N. Kolmogorov, Grundbegriffe der Wahrscheinlichkeitsrechnung
      independent Distributionorder sequence 'salmostnecessarilyConvergenceFoundations
      ↓ Spring make using : consensus statistic C(x) 'slarge-samplelinesas 
      
1951 ─ Robbins-Monro Stochastic Approximation
      H. Robbins & S. Monro, "A stochastic approximation method"
      Stochastic Gradient Descent'snumber learning Foundations; ∑α_t=∞, ∑α_t²<∞ conditions
      ↓ Spring make using : θ_t update follows RM recursion; time scale extension
      
1952 ─ Chernoff information
      H. Chernoff, "A measure of asymptotic efficiency for tests"
      Assumptionschecktest'soptimal exponentialRate
      ↓ Spring make using : Theorem 4' Detection'sprecise number Rate

1953 ─ Doob Martingale Convergence
      J. L. Doob, Stochastic Processes
      bounded martingale'salmostnecessarilyConvergence
      ↓ Spring Used for: S_t as a [0,1]-bounded martingale → S_t → S_∞ a.s.

1960 ─ Bahadur-Rao Exact Asymptotics
      R. R. Bahadur & R. R. Rao, "On deviations of the sample mean"
      Sampleall value tail rate 'sprecise orderasymptoticnear(containlattice correction)
      ↓ Spring Used for: exact constant for Theorem 4'; Bernoulli lattice correction
      
1961 ─ Fano non-etc.formula 
      R. M. Fano, Transmission of Information
      Information Theorylower bound---interactioninformationand error probability'srelationship 
      ↓ Spring make using : Theorem 2 featurelower bound
      
1963 ─ Hoeffding non-etc.formula 
      W. Hoeffding, "Probability inequalities for sums of bounded random variables"
      Exponential concentration for sums of bounded independent variables
      ↓ Spring Used for: concentration bound for consensus score; main tool for Theorem 1
      
1964 ─ Pinsker non-etc.formula 
      M. S. Pinsker, Information and Information Stability
      TV ≤ √(KL/2) --- bridge between information divergence and statistical distance
      ↓ Spring Used for: δ → TV conversion in Theorem 2
      
1973 ─ Le Cam Minimax
      L. Le Cam, "Convergence of estimates under dimensionality restrictions"
      two-point methodlower bound, convex hull argument, minimax optimality framework
      ↓ Spring make using : Theorem 4' lower bound; Hellinger tensorization
      
1981 ─ Pollard k-means one property 
      D. Pollard, "Strong consistency of k-means clustering"
      Almost sure consistency of k-means
      ↓ Spring make using : Theorem 5 state discovery'sTheoryFoundations

1985 ─ QWERTY Lock-In (David)
      P. A. David, "Clio and the Economics of QWERTY"
      Pathdependenceand superiortechnologtechniqueLock-In'svia learning Theory
      ↓ Conceptual analogy: Spring 'snon-point Structure---Gatekeeper can be can Lock-Insuperiorfitmatchplace
      
1989 ─ Increasing Returns (Arthur)
      W. B. Arthur, "Competing technologies, increasing returns, and lock-in by historical events"
      Multiple equilibria in positive feedback loops
      ↓ Conceptual analogy: Spring Gatekeeper 'sSelection Biascycleenvironment---initial bias is amplified throughSelf-Evolutionis amplify

2003 ─ Online Convex Programming (Zinkevich)
      M. Zinkevich, "Online convex programming and generalized infinitesimal gradient ascent"
      Regret bound O(√T) for online gradient descent
      ↓ Spring make using : Gatekeeper OGD update, delayed feedbackextension, growing memory bankextension

2010 ─ Domainshould Theory
      Ben-David et al., "A theory of learning from different domains"
      Trainingand TestDistributionnon-matching'sization bound 
      ↓ Spring make using : D_static betweengap'sDomainshould bound  (Lemma 12.1)

2017 ─ AlphaZero Self-Play
      D. Silver et al., "Mastering Chess and Shogi by Self-Play..."
      Achieves superhuman level through self-play, policy iteration, Monte Carlo tree search
      ↓ Spring class like: self-evolution loopStructure, monotone improvement guarantee, memory replay
      ↓ Spring non-: process physicalData(non-), not yet knowenvironment dynamic , formal convergence proof

2026 ─ Spring Self-Evolution Gatekeeper (This Work)
      SCX Research Group, "Spring: A Self-Evolving Gatekeeper with Provable Convergence"
      Self-Evolutionevaluation criteria(non-only strategy ), monotonically growing memory, 
      resurrection mechanism---"Winter does not kill --- it waits for spring"
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Spring
Comparison with Prior Work}\label{spring-ux4e0eux524dux4ebaux7684ux5bf9ux6bd4ux8868}

\subsubsection{4.1 Core Comparison}\label{ux6838ux5fc3ux5bf9ux6bd4}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1154}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1731}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.2115}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1346}}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Algorithm / Framework
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Year
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Self-Evolving?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Convergence Proof?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Memory?
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Domain
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Robbins-Monro} & 1951 & No & Yes(asymptoticnear a.s.) & No &
throughusing Stochastic Approximation \\
\textbf{Hedge / expert Recommendation} & 1997 & No & Yes(regret  O(√T)) & No &
at Online decision \\
\textbf{Learning} & 2009 & No(Fixed Query Strategy) & Yes(Label Complexity) &
Label pool & class (unitModel) \\
\textbf{Bayesian Optimization} & 2016 & No(Fixed Prior) & Yes(regret ) & GP posterior &
Black-box optimization \\
\textbf{AlphaZero} & 2017 & Yes(Self-Play) & No(via property )* &
Replay buffer (with replacement) & Games (known environment) \\
\textbf{Solomonoff reducinduct} & 1964 & N/A(Theorylimit) & Yes(for can be Computationenvironment ) &
All history & using reducinduct(cannot be Computation) \\
\textbf{Spring (Cercis)} & 2026 & \textbf{Yes}(Gatekeeper co-evolution) &
\textbf{Yes}(Theorem SE-1/2, Theorem 12.5) & \textbf{Yes}(M\_t
monotonically growing, no deletion) & \textbf{Arbitrary}(design using physicalScience) \\
\end{longtable}

*AlphaZero at tabular caseunder has strategy iteration'sConvergenceProof; for via network as via property . 

\subsubsection{4.2 Spring vs AlphaZero
Detailed Comparison}\label{spring-vs-alphazero-ux8be6ux7ec6ux5bf9ux6bd4}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3200}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
AlphaZero
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Spring
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Data Generation} & Synthetic (self-play)------environment fully known & Real (NEP
simulation)------physicalrulelawnot yet  \\
\textbf{Feedback} & Instant (game outcome) & Delayed (NEP batch training) \\
\textbf{Convergence Mechanism} & Policy iteration + MCTS & Lyapunov under  + Bayesian update \\
\textbf{Memory} & Replay buffer (with replacement) & monotonicgrowing memory bank(non-deleteremov) \\
\textbf{Convergence Guarantee} & monotonicimprovement(tabular casealready Proof) & Finite-time fixed point (Theorem
SE-2, already Proof) \\
\textbf{Final State} & then on Optimalstrategy  & Self-consistent fixed point (possibly suboptimal) \\
\textbf{Exploration} & Dirichlet noise + visit count & Random exploration fraction ε +
annealed acceptance threshold \\
\end{longtable}

\subsubsection{4.3 Spring vs
Bayesian Optimization}\label{spring-vs-ux8d1dux53f6ux65afux4f18ux5316}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3200}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Bayesian Optimization
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Spring
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Objective} & Optimize static black-box function & Evolve evaluation criteria + optimize model \\
\textbf{Surrogate Model} & Gaussian process & NEP student (neural network) \\
\textbf{Acquisition Function} & EI / UCB / Thompson & state valueFunction V(s) =
R̂(s)·ρ(s)·(1-C(s)) \\
\textbf{Granularity} & Pointwise & Status(aggregatemultiple Sample) \\
\textbf{Number of Objectives} & Single-objective (usually) &
Multi-objective (noise detection, expert disagreement, data diversity) \\
\textbf{Cost Structure} & Homogeneous & differenquality(DFT Computationbecause Statuswhile differen) \\
\textbf{ConvergenceRate} & O(√(T·γ\_T))(Cumulative regret) & O(t\^{}\{-a\}) +
has time stoppterminatGuarantee \\
\end{longtable}

\subsubsection{4.4 Spring vs
Learning}\label{spring-vs-ux4e3bux52a8ux5b66ux4e60}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2609}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3913}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3478}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Learning
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Spring
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Query Strategy} & non-determine property  / QBC / Expected model change &
State Certification------at state-levelaggregatedecision \\
\textbf{Model} & Single discriminative model & Multi-expert ensemble + student Model \\
\textbf{Query Decision} & Pointwise & Hierarchical------firststate-level, again sample-level \\
\textbf{Noise Handling} & Oracle assumes clean (most) &
\textbf{must design Objective}------explicitly detect label noise \\
\textbf{Label Complexity} & O(θ·d·log(1/ε)) & O(K\_S·(1/Δ\_min²)·log(1/ε)) \\
\textbf{Certification} & No & each linesby state-levelguarantee certification \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Spring
Unique Contributions}\label{spring-ux72ecux6709ux7684ux8d21ux732e}

with under Yes Spring while \textbf{no prior work has simultaneously achieved}:

\subsubsection{5.1
Self-Evolutionevaluation criteria(non-only strategy )}\label{ux81eaux8fdbux5316ux8bc4ux4f30ux6807ux51c6ux4e0dux4ec5ux81eaux6539ux8fdbux7b56ux7565}

\textbf{Prior Work Limitations}: AlphaZero
improvementits strategy (such as under chess), but evaluation criteria(gamegamerulethen )Yesfixed 'strue. Learningand Bayesian Optimizationhas Fixed Query Strategy/Acquisition Function. 

\textbf{Spring's Innovation}: Gatekeeper \(S_t\)
\textbf{itself}YesLearningfor object------it defines''what Yes'sData''. Systemis notat fixed evaluation criteriaunder ization strategy , while Yes\textbf{simultaneously evolvingevaluation criteriaitself}. this is aLearningProblem: Learningsuch as LearningJudgmentDataquantity . 

\subsubsection{5.2
monotonically growing memory(non-replacementreplacimportantamplify)}\label{ux5355ux8c03ux589eux957fux8bb0ux5fc6ux4e0dux66ffux6362ux91cdux653e}

\textbf{Prior Work Limitations}:
which has importantamplifymitigationconflictscheme(AlphaZero, DQN, via testimportantamplify)at someprocesson replacreplacoldMemory------mitigationconflicthas bound , mostold/mostnon-has using 'sSampleis discarddiscard. 

\textbf{Spring's Innovation}:
\(M_t \subseteq M_{t+1}\)------Memoryonly increaselength. no deletionstrategy means :  1.
Nocatastrophidifficultproperty regretforget------earlyperiodLearning'sStructurenon-is discarddiscard 2.
Memorymonotonicproperty provide by 'sstructural property(Lemma SE-1.2 requires must ) 3.
resurrection mechanismas can be can ------number Structures below threshold'sStructureconservretainat memory bankin , etc.wait
Gatekeeper to mature and re-score them

\subsubsection{5.3 resurrection mechanism}\label{ux590dux6d3bux673aux5236}

\textbf{Prior Work Limitations}:
NoAlgorithmcontain''duplicatresurre'''s  . is reject'sSamplemust is discarddiscard(Learning), must non-is importantrevisit. 

\textbf{Spring's Innovation}: ``Winter does not kill------it waits for spring''. Structures below threshold
\(\tau_{\text{keep}}\) 'sStructurenon-is deleteremov------ dormancy. when  Gatekeeper
at subsequeniterationin Evolution(\(S_{t+1}\)
can be can for Structuregive number ), dormancyStructureis duplicatresurreandintroduce Training: 

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{for}\NormalTok{ (x, y, old\_score) }\KeywordTok{in}\NormalTok{ M\_t where old\_score }\OperatorTok{\textless{}}\NormalTok{ τ\_keep:}
\NormalTok{    new\_score }\OperatorTok{=}\NormalTok{ S\_\{t}\OperatorTok{+}\DecValTok{1}\NormalTok{\}(x, y)}
    \ControlFlowTok{if}\NormalTok{ new\_score }\OperatorTok{\textgreater{}=}\NormalTok{ τ\_keep:}
\NormalTok{        RESURRECT(x, y)  }\CommentTok{\# Spring resurrects it}
\end{Highlighting}
\end{Shaded}

this at number learning on code as : Gatekeeper
Updatecan be with at any time will Sample'sevaluatededucthresholdvalue , tionrejectas nearsubject to------this at Learning(one oncerejecti.e.permanentdiscarddiscard)or Bayesian Optimization(one Evaluationi.e.consumeepre-compute)in Yescannot be can 's. 

\subsubsection{5.4 Coupled Dynamical System Analysis(Gatekeeper +
student co-evolution)}\label{ux8026ux5408ux52a8ux529bux7cfbux7edfux5206ux6790Gatekeeper-ux5b66ux751fux534fux540cux8fdbux5316}

\textbf{Prior Work Limitations}: Robbins-Monro analyzes a single recursion. AlphaZero
'sstrategy iterationcommutreplacfixed one item update the other(strategy Evaluationand strategy improvementbetweenNoCoupled Dynamical System Analysis). 

\textbf{Spring's Innovation}: Complete'sSystemformula ization :  - state space
\(\mathcal{Z} = \mathcal{F} \times \mathcal{M} \times \Theta\)(Product Space)
- update operator
\(\Phi = (\Phi_S, \Phi_M, \Phi_\theta)\)(because resultpreviousfeedStructure, block-triangular
Jacobian) - time scale  ODE Analysis(fastslowprocessgroup) - Lyapunov
Functionat Product Spaceon 'sformula under  - \(\omega\)--limit set, basin of attraction, phase diagram

has Algorithmat formula Systemframework in AnalysisLearningitem 'sco-evolution. 

\subsubsection{5.5 precise ConvergenceRateand  Polyak
averageall }\label{ux7cbeux786eux6536ux655bux901fux7387ux4e0e-polyak-ux5e73ux5747}

\textbf{Prior Work Limitations}: Bayesian Optimizationgive cumulative regret bounds
\(O(\sqrt{T \cdot \gamma_T})\)
but non-give parameter convergence rates. Learninggive Label Complexitybut non-give LearningRate. 

\textbf{Spring's Innovation}: - reinforceconvexunder student Rateas  \(O(t^{-a})\)(Theorem 11.1,
already Proof) - receiveshrinkunder  Gatekeeper Rateas  \(O(t^{-b})\)(Theorem 11.2,
already Proof) - Polyak-Ruppert Averagingwill Rateboost to  \(O(t^{-1})\)------matching
minimax lower bound \(\Omega(t^{-1})\) - Rateby student bottleneck  \(O(t^{-a})\)
lead to(Theorem 11.3) - has time (non-asymptoticnear)bound containexplicitformula number (Theorem 11.6,
already Proof)

\subsubsection{5.6
Formalization Selection BiascycleenvironmentAnalysis}\label{ux5f62ux5f0fux5316ux9009ux62e9ux504fux5deeux5faaux73afux5206ux6790}

\textbf{Prior Work Limitations}:
NoAlgorithmformula ization Analysisits itselfDataSelectionorder contributestablishreinforceization bias cycleenvironment'sMechanism. 

\textbf{Spring's Innovation}: Theorem 12.2
Proof------\textbf{Without reference set replay, Lyapunov
under at asymptotically cannot be 's}------Gatekeeper
can reduce its apparent loss by only accepting''easy''Samplereduce its tableobserveloss, while non-true improvement. this Yesfor any self-selecting learning system'sdeepcharacterizelimitation , at 
Spring not yet is formula ization Proof. 

Specifically, Selection Biascycleenvironment: 

\[\text{Gatekeeper } S_t \xrightarrow{\text{Selection}} \text{Memory } M_{t+1} \xrightarrow{\text{Training}} \text{NEP } f_{\theta_{t+1}} \xrightarrow{\text{Feedback}} \text{Gatekeeper } S_{t+1}\]

is solution, quantity ization Proofas  Lyapunov under downgrade'smust obstacle ------then resolved through the combined mechanism of reference set replay +
importantmust property sampleSimilarly'sMechanismis Resolution(Theorem 12.5, CompleteProof). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Summary}\label{ux603bux7ed3}

Spring
Self-evolution algorithmat seven via number learning Domain'scommutconvergeatoperation: Stochastic Approximation, System, Bayesian Inference, at Online Learning, concentration inequalities, Information Theoryand 
minimax
Theory. its core architectureconfigurinnovation ------Two-Timescale Separation, Reference Set Replay BreaksSelection Bias, monotonically growing memory, resurrection mechanism, Coupled Dynamical System Analysis------by this Domain'sCompleteProofchainsupport . and which has previous'sproperty intervalat : Spring
\textbf{simultaneously}Self-Evolutionits evaluation criteriaand is EvaluationModel, formally proves convergence, maintain monotonically growing'sMemory(permanennon-regretforget). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{spectrumsystemlearning by  SCX Theorygroupencode, 2026-06-28}

\end{document}
