% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Multi-Head Spring and Physical Positional Encoding
 rigorous mathematicalAnalysis}\label{multi-head-spring-ux4e0e-physical-positional-encoding-ux7684ux4e25ux683cux6570ux5b66ux5206ux6790}

\subsection{to SCX
frameworkTheoremFoundation honest audit}\label{ux5bf9-scx-ux6846ux67b6ux5b9aux7406ux57faux7840ux7684ux8bdaux5b9eux5ba1ux67e5}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Document Status}: Complete Analysis\\
\textbf{-dependence}: Theorem 1--4 (theory/theorems/), SE-1, SE-2
(theory/self\_evolution/)\\
\textbf{Methodology}: Rigorous Formal Derivation + Honest Critique (Unsparing)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{0.
Notation Convention and Framework Reconstruction}\label{ux7b26ux53f7ux7ea6ux5b9aux4e0eux6846ux67b6ux91cdux6784}

Before entering the analysis, we first precisely reconstruct SCX frameworkmathematical foundations. withlower has DefinitionuniformFrom 
prompt in the Theoremdescriberigorousform -ization. 

\subsubsection{0.1 State Crystallization (State
Crystallization)}\label{ux72b6ux6001ux7ed3ux6676state-crystallization}

Let the raw physical observation space be 
\(\mathcal{X} \subset \mathbb{R}^{d_{\text{raw}}}\) (continuous physical quantities: PBE
energy surface 、force field、electrondensedegree etc.). State Crystallization is a map: 

\[\mathcal{C}: \mathcal{X} \to \mathcal{S} = \{s_1, \ldots, s_N\}\]

whereeach \(s_i \in \mathbb{R}^{d_s}\)
is a \textbf{discreteStatusatom}toward quantity representation. and BPE
 system compute drive action is different, \(\mathcal{C}\) isphysics-driven: it according to PBE
energy surface self-however curve rate boundary advance linescluster. Formally: 

\[s_i = \mathcal{C}(x) \quad \text{where} \quad x \in \mathcal{X} \text{ and } \nabla^2 E_{\text{PBE}}(x) \text{ in } x \text{ athas a spectrumbetween gap }\]

\subsubsection{0.2 Yajieaudit (Yajie
Audit)}\label{ux96c5ux6d01ux5ba1ux8ba1yajie-audit}

setuphas \(M\) independentexpertmodel \(\{E_1, \ldots, E_M\}\). forStatusatom \(s_i\)
and its label \(y_i\), Definition: 

\[c_m(s_i) = \mathbf{1}[E_m(s_i) \neq y_i] \quad \text{ (expert m is different agree todetermineLabel)}\]

Consistency count: 
\[C(s_i) = \sum_{m=1}^{M} c_m(s_i) \sim \text{Binomial}(M, p_{s_i})\]

where \(p_{s_i}\)
isexpert''is different agree ''Labelprobability. For clean samples, \(p_{s_i} = p_{\text{clean}} < 0.5\); For noisy samples, \(p_{s_i} = p_{\text{noisy}} > 0.5\). 

\textbf{Cercis Score}: \[S(s_i) = Q(s_i) + \eta(t) \cdot N(s_i)\]

where \(Q(s_i)\) isFoundationqualityseparate (based onmulti-expertconsistentness), \(N(s_i)\)
isnew strange ness reward incentivize, \(\eta(t)\) istimebetween decay subtract explore search weight. 

\subsubsection{0.3 Spring
self-evolutiondynamics}\label{spring-ux81eaux8fdbux5316ux52a8ux529bux5b66}

State Transition: \[(S_t, \theta_t, M_t) \to (S_{t+1}, \theta_{t+1}, M_{t+1})\]

Spring self-Note agree forceform (singlehead): 
\[\text{Spring}(s_i, \mathcal{S}) = \text{softmax}\left(\frac{(Q s_i)(K \mathcal{S})^T}{\sqrt{d_k}}\right) \cdot (V \mathcal{S})\]

where
\(Q, K \in \mathbb{R}^{d_k \times d_s}\), \(V \in \mathbb{R}^{d_v \times d_s}\). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.
Core TheoremsExact Theorem Statement}\label{ux6838ux5fc3ux5b9aux7406ux7684ux7cbeux786eux5b9aux7406ux9648ux8ff0}

\subsubsection{Theorem
1 (multi-expertconsistentnessnoise detection)}\label{theorem-1ux591aux4e13ux5bb6ux4e00ux81f4ux6027ux566aux58f0ux68c0ux6d4b}

\[\boxed{F_1 \geq 1 - \frac{1}{\eta} \sum_{s \in \mathcal{S}} \rho_s \cdot \exp\left(-2M \Delta_s^2\right)}\]

where: -
\(\rho_s = |\{i: \mathcal{C}(x_i) = s\}| / N_{\text{total}}\): Statusatom
\(s\) proportion -
\(\Delta_s = |p_{\text{clean},s} - p_{\text{noisy},s}|\): Status \(s\)
Detection Margin - \(M\): independentexpertquantity - \(\eta > 0\): normalization factor

\textbf{Proof Technique}: Chernoff bound + Hoeffding inequality. toeach Status
\(s\), \(C(s)\) istwo two term distributionmixedcombined. Marked as noisewhen
\(C(s) \geq \tau_s\). Chernoff to out: 
\[P(\text{false positive}) = P_{H_0}(C(s) \geq \tau_s) \leq \exp(-2M(\tau_s/M - p_{\text{clean},s})^2)\]
\[P(\text{false negative}) = P_{H_1}(C(s) < \tau_s) \leq \exp(-2M(p_{\text{noisy},s} - \tau_s/M)^2)\]

 select take \(\tau_s = M \cdot (p_{\text{clean},s} + p_{\text{noisy},s})/2\)
 make two term balance, obtain to \(\Delta_s\). 

\subsubsection{Theorem
2 (weakCharacteristics must however lose effect)}\label{theorem-2ux5f31ux7279ux5f81ux5fc5ux7136ux5931ux6548}

\[\boxed{F_{1,\text{SCX}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta}{2}}}\]

where \(\delta > 0\) carve picture Characteristicsspaceinformationinsufficient process degree. 

\textbf{Proof Technique}: Fano inequality. Definition
\(\delta = \min_{f} \mathbb{E}[\mathbf{1}[f(X) \neq Y]] - \text{Bayes error}\), i.e.Optimal canlinesseparate class device and Bayesian leaf -ese Optimal 'sbetweenminimum difference distance. Fano
to out Conditionsentropyandincorrect rate relationship: 
\[P_e \geq \frac{H(Y|X) - 1}{\log |\mathcal{Y}|}\]

when \(H(Y|X)\) relativelylarge (Characteristicsweak), any what Methodology (includes SCX)incorrect rate lower boundby
\(\delta\) verdict determine. 

\subsubsection{Theorem
3 (noise andDifficultysampleindistinguishability)}\label{theorem-3ux566aux58f0ux4e0eux56f0ux96beux6837ux672cux7684ux4e0dux53efux533aux5206ux6027}

\textbf{Constructionnessdouble world world Proof}: existenceattwo world world 
\(W_A, W_B\), has has observationequivalentData Distribution
\(P_{W_A}(X, Y) = P_{W_B}(X, Y)\), but: 

\begin{itemize}
\tightlist
\item
 world world \(W_A\): type this \((x, y)\) isbyincorrectMarked (Labelnoise)
\item
 world world \(W_B\): type this \((x, y)\) correctMarkedbutthisupperDifficulty (high Bayesian leaf -ese notdeterminism)
\end{itemize}

nothas algorithmcanonlythroughobservationdatadistinguishthistwo world world. 

form -ization: \(\forall\) verdict strategy rule then
\(d: (\mathcal{X} \times \mathcal{Y})^n \to \{W_A, W_B\}\), 
\[\max_{W \in \{W_A, W_B\}} P_W(d(\text{data}) \neq W) \geq \frac{1}{2} - \frac{1}{2}\|P_{W_A}^{\otimes n} - P_{W_B}^{\otimes n}\|_{\text{TV}} = \frac{1}{2}\]

becauseis 
\(P_{W_A}^{\otimes n} = P_{W_B}^{\otimes n}\) (complete fullmutualsameobservationdistribution). 

\subsubsection{Theorem 4 (Exact often number Minimax
Optimal)}\label{theorem-4ux7cbeux786eux5e38ux6570-minimax-ux6700ux4f18}

SCX noise detectionincorrect rate atwithlowermeaningupper reach toExact often number Optimal: 

\[\lim_{M \to \infty} -\frac{1}{M} \log P_{\text{FN}}^{(M)} = D_{\text{KL}}(P_{\text{noisy}} \| P_{\text{clean}}) \quad \text{subject to } P_{\text{FP}}^{(M)} \leq \alpha\]

and Bahadur-Rao exhibit opento out \(O(1/\sqrt{M})\) correction termalsomatches minimax lower bound. 

\textbf{Proof Technique}: Chernoff-Stein Lemma + Bahadur-Rao exhibit open. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. group file 1: Physical Positional
Encoding (PPE)mathematicalAnalysis}\label{ux7ec4ux4ef6-1physical-positional-encodingppeux7684ux6570ux5b66ux5206ux6790}

\subsubsection{2.1 Formal Definition}\label{ux5f62ux5f0fux5316ux5b9aux4e49}

setupeach Statusatom \(s_i\) carry has a physicalLocation \(p_i \in \mathcal{P}\), where: 
- whitesequence: \(\mathcal{P} = \mathbb{N}\) (residueindexes) - 3D
structure: \(\mathcal{P} = \mathbb{R}^3\) (atom sit mark) -
Docking: \(\mathcal{P} = \mathbb{R}^3\) (knot combined gap bag sit mark)

\textbf{Definition 1 (Physical Positional Encoding)}: PPE is a map
\[\text{PE}: \mathcal{P} \to \mathbb{R}^{d_{\text{pe}}}\]

encodingbackStatusrepresentation: \[\boxed{h_i = \phi(s_i) + \text{PE}(p_i)}\]

where \(\phi: \mathbb{R}^{d_s} \to \mathbb{R}^{d}\)
isCharacteristicsmap, and we s require \(d_{\text{pe}} = d\). 

\textbf{may select scheme (assemble connect +projects)}: 
\[h_i = W \cdot [\phi(s_i); \text{PE}(p_i)]\]

where
\(W \in \mathbb{R}^{d \times (d + d_{\text{pe}})}\). mathematicallythisismore a like scheme, butaddmethodschememore simple clean andat
\(W = [I_d, I_d]\) time retreat -ization is addmethod. withlowerAnalysiswithaddmethodschemeis main, resultmay push broad. 

\subsubsection{2.2
encodingfunctionconcreteform}\label{ux7f16ux7801ux51fdux6570ux7684ux5177ux4f53ux5f62ux5f0f}

\textbf{mark quantity Locationsinusoidal encoding} (analogy Transformer): 
\[\text{PE}(p, 2j) = \sin\left(\frac{p}{10000^{2j/d}}\right), \quad \text{PE}(p, 2j+1) = \cos\left(\frac{p}{10000^{2j/d}}\right)\]

where \(j = 0, 1, \ldots, d/2 - 1\). 

\textbf{3D Locationrotationencoding} (analogy RoPE, adapt match physicalspace): 
\[\text{PE}_{\text{rot}}(\mathbf{p}) = \mathbf{R}(\mathbf{p}) \cdot \mathbf{e}_0\]

where \(\mathbf{R}(\mathbf{p}) \in SO(d)\) isby sit markparameter -ization rotationmatrix. for
\(\mathbf{p} = (x, y, z)\): 
\[\mathbf{R}(\mathbf{p}) = \mathbf{R}_x(\omega_x x) \cdot \mathbf{R}_y(\omega_y y) \cdot \mathbf{R}_z(\omega_z z)\]

where \(\mathbf{R}_\alpha(\theta)\) isin \(\alpha\) axis upper \(d\)
dimensionalrotation, \(\omega_x, \omega_y, \omega_z\) ismaylearningorfixed frequencyparameter. 

\textbf{keyness}: and NLP
Locationencodingis different, physicalLocation has has \textbf{manydimensional several what structure}. a good good PPE
should respect heavyphysicalto called ness: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \textbf{translation etc. change ness} (to 3D
 structure): \(\text{PE}(\mathbf{p} + \mathbf{t})\) and
 \(\text{PE}(\mathbf{p})\) has canpredictionrelationship
\item
 \textbf{rotation etc. change ness}: \(\text{PE}(\mathbf{R}\mathbf{p})\)
 canwiththroughrotationencodingspace come restore complex 
\item
 \textbf{Lipschitz
 continuousness}: \(\|\text{PE}(\mathbf{p}) - \text{PE}(\mathbf{q})\| \leq L_{\text{PE}} \cdot \|\mathbf{p} - \mathbf{q}\|\)
\end{enumerate}

\subsubsection{2.3 to Theorem 1
impact}\label{ux5bf9-theorem-1-ux7684ux5f71ux54cd}

\textbf{Core Problem}: PPE such as what changesDetection Margin \(\Delta_s\)? 

\textbf{Proposition 2.1 (PPE lowermodifybackDetection Margin)}: in PPE
strengthensCharacteristicsspacemiddle, Status \(s\) has effect Detection Marginis: 
\[\boxed{\Delta_s^{\text{PPE}} = \Delta_s + \delta_s^{\text{PE}}}\]

where
\(\delta_s^{\text{PE}} \in [-\Delta_s, 1 - p_{\text{clean},s} - \Delta_s]\)
satisfies: 
\[\delta_s^{\text{PE}} = \underbrace{(p_{\text{clean},s}^{\text{PPE}} - p_{\text{clean},s})}_{\text{dry clean sampleupperconsistentnessincrease benefit }} - \underbrace{(p_{\text{noisy},s}^{\text{PPE}} - p_{\text{noisy},s})}_{\text{noisesampleuppermixed confuse increase benefit }}\]

\emph{Proof}: atstrengthensspacemiddle, expertpredictionprobability change is 
\(p_{\text{clean},s}^{\text{PPE}} = \mathbb{E}[\mathbf{1}[E_m(h_i) = y_i] | H_0]\)
and
\(p_{\text{noisy},s}^{\text{PPE}} = \mathbb{E}[\mathbf{1}[E_m(h_i) \neq y_i] | H_1]\). Detection MarginDefinitionis 
\(\Delta_s^{\text{PPE}} = |p_{\text{clean},s}^{\text{PPE}} - (1 - p_{\text{noisy},s}^{\text{PPE}})| = |p_{\text{clean},s}^{\text{PPE}} + p_{\text{noisy},s}^{\text{PPE}} - 1|\). exhibit openi.e. obtain. \(\square\)

\textbf{Theorem 2.1 (PPE lower F1 lower bound)}: 
\[\boxed{F_1^{\text{PPE}} \geq 1 - \frac{1}{\eta} \sum_{s \in \mathcal{S}} \rho_s \cdot \exp\left(-2M (\Delta_s + \delta_s^{\text{PE}})^2\right)}\]

Chernoff-Hoeffding inequality\textbf{has effect nessnot change } (indicate show variablestillin \([0,1]\)
innerhas world), but often number change -ization. 

when \(\delta_s^{\text{PE}} > 0\) to has \(s\) (PPE totalis help help detection): 
\[F_1^{\text{PPE}} - F_1 \geq \frac{1}{\eta} \sum_s \rho_s \left[\exp(-2M\Delta_s^2) - \exp(-2M(\Delta_s + \delta_s^{\text{PE}})^2)\right] \geq 0\]

\textbf{Proposition 2.2 (spacelocalnesscurrentlythen -ization)}: If PE satisfies Lipschitz
Conditions (Lipschitz often number is \(L_{\text{PE}}\)), andexpertmodel \(E_m\)
toits input enter satisfies \(L_E\)-Lipschitz, thento any agree two has has physicalLocation \(p_i, p_j\)
Statusatom: 
\[|\delta_i^{\text{PE}} - \delta_j^{\text{PE}}| \leq 2L_E \cdot L_{\text{PE}} \cdot \|p_i - p_j\|\]

\emph{Proof}: expertconsistentnessprobability change -ization receive input enter change -ization convention bind: \begin{align*}
|p_{\text{clean},i}^{\text{PPE}} - p_{\text{clean},j}^{\text{PPE}}|
&\leq \mathbb{E}_{E} \left|\mathbf{1}[E(h_i) = y] - \mathbf{1}[E(h_j) = y]\right| \\
&\leq \mathbb{E}_{E} \left[ L_E \cdot \|h_i - h_j\| \right] \\
&= L_E \cdot \|\text{PE}(p_i) - \text{PE}(p_j)\| \quad (\text{becauseis } s_i = s_j = s) \\
&\leq L_E \cdot L_{\text{PE}} \cdot \|p_i - p_j\|
\end{align*}

same type to \(p_{\text{noisy}}\) holds. twoone difference obtain tobecause factor 2. \(\square\)

this agree taste -ing \textbf{physically neighbor near Statusatom has has similaritydetection difficult degree}------ a mayverifyprediction. 

\subsubsection{2.4 to Theorem 2
impact}\label{ux5bf9-theorem-2-ux7684ux5f71ux54cd}

\textbf{Core Problem}: not complete beautiful PPE such as what changesCharacteristicsweak point world ? 

\textbf{Definition 2 (encodingnot complete beautiful degree \(\varepsilon_{\text{PE}}\))}: 
\[\boxed{\varepsilon_{\text{PE}} = I(Y; P | X) - I(Y; \text{PE}(P) | X) \geq 0}\]

where \(I(Y; P | X)\) isphysicalLocation \(P\) intodetermine original begin Characteristics \(X\) lowerregardingLabel
\(Y\) Conditionsmutualinformation, and \(I(Y; \text{PE}(P) | X)\)
isencodingbackpreservesinformation. \(\varepsilon_{\text{PE}}\)
quantifies\textbf{Locationinformationatencodingprocess in the discard lose }. 

\textbf{Theorem 2.2 (not complete beautiful PPE lower lose effect upper bound)}: 
\[\boxed{F_{1,\text{SCX}}^{\text{PPE}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta + \frac{2\varepsilon_{\text{PE}}}{C_F^2}}{2}}}\]

\emph{Proof}: Fano inequalityto out 
\(P_e^{\text{PPE}} \geq \frac{H(Y|X, \text{PE}(P)) - 1}{\log |\mathcal{Y}|}\). we s has: 
\begin{align*}
H(Y|X, \text{PE}(P)) &= H(Y|X) - I(Y; \text{PE}(P) | X) \\
&= H(Y|X) - [I(Y; P | X) - \varepsilon_{\text{PE}}] \\
&\geq H(Y|X) - I(Y; P | X) + \varepsilon_{\text{PE}}
\end{align*}

and \(\delta\) Definitionto out 
\(H(Y|X) - 1 \geq (\delta/2) \cdot \log|\mathcal{Y}|\) (through Fano
 inverse direction). becausethis: 
\[H(Y|X, \text{PE}(P)) - 1 \geq \left(\frac{\delta}{2} + \frac{\varepsilon_{\text{PE}}}{C_F}\right) \cdot \log|\mathcal{Y}|\]

 generation enter SCX F1 upper bound table reach formula i.e. obtain. \(\square\)

\textbf{ tight cause nessAnalysis}: - when
\(\varepsilon_{\text{PE}} = 0\) (complete beautiful encoding): upper bound restore complex is original begin Theorem 2 -
when \(\varepsilon_{\text{PE}} > 0\): upper boundwith
\(\sqrt{\delta + O(\varepsilon_{\text{PE}})}\) rate release loose - when
\(\varepsilon_{\text{PE}} \to I(Y; P | X)\) (complete fullfailencoding, discard lose has Locationinformation): upper bound time to complete fullnotuseLocationinformationcase

\textbf{Corollary 2.1 (Locationinformationvalue)}: such as result physicalLocation \(P\)
this self not carry carry any what regardingLabel \(Y\) information (i.e. \(I(Y; P | X) = 0\)), then
\(\varepsilon_{\text{PE}} = 0\) (becauseis nohas informationmay discard lose), Theorem 2
upper bound\textbf{ complete fullnot change }. PPE in this caseis\textbf{pure pure parameter wave cost }. 

\subsubsection{2.5 to Theorem 3
impact}\label{ux5bf9-theorem-3-ux7684ux5f71ux54cd}

thisismostkeyAnalysis. Theorem 3 indistinguishability-dependenceatConstructiontwo world world \(W_A, W_B\)
 has has complete fullmutualsameobservationdistribution \(P(X, Y)\). 

\textbf{Proposition 2.3 (fixed PPE maintains Theorem 3)}: such as result PE
is a \textbf{ first verify fixed }function (not-dependenceattrainingdataorLabel), thentwo world world in the strengthensCharacteristicsdistribution according howevermutualsame: 
\[P_{W_A}(X, \text{PE}(P), Y) = P_{W_B}(X, \text{PE}(P), Y)\]

becausethis Theorem 3 indistinguishability\textbf{remains unchanged}. PPE
nothas help help alsonothas damage harm mostbadcasemayidentifiesness. 

\emph{Proof}: physicalLocation \(P\) byphysicalstructure verdict determine, is independent ofLabelgenerateprocess. becausethis
\(P_{W_A}(P|X) = P_{W_B}(P|X)\). becauseis PE
isfixed determinismfunction, \(P_{W_A}(\text{PE}(P)|X) = P_{W_B}(\text{PE}(P)|X)\). observationdistribution complete fullmutualsame. \(\square\)

\textbf{Proposition 2.4 (learnable PPE breaking Theorem 3)}: such as result PE parameter
\(\theta_{\text{PE}}\) isFrom datamiddlelearning (includesLabel
\(Y\)), thenattwo world world middlelearningtois differentencoding: 
\[\text{PE}_{W_A} \neq \text{PE}_{W_B}\]

becauseis \(W_A\) noiseLabelimpactlearningprocess. thistime: 
\[P_{W_A}(X, \text{PE}_{W_A}(P), Y) \neq P_{W_B}(X, \text{PE}_{W_B}(P), Y)\]

two world world change obtain \textbf{maydistinguish}. Theorem 3 isbreaking. 

\textbf{honest assessment}: thisin practiceis\textbf{good eliminate rest }. Theorem 3
is a is impossiblenessresult------ it say in the worst case, noiseandDifficultysampleNonecannot distinguish. such as result learnable 
PPE canbreaking Theorem 3
front raise Conditions, agree taste -ing \textbf{LocationinformationprovidedistinguishnoiseandDifficultysamplecanforce}. thisis NOT frameworkdefect, andisframeworkprogress. butthis confirm actual agree taste -ing 
Theorem 3 notagainNoneConditionsholds------requireaddfront raise Conditions''PE
isfixed orLocationinformationnot carry carry Labelmutualcloseinformation''. 

\subsubsection{2.6 to Theorem 4
impact}\label{ux5bf9-theorem-4-ux7684ux5f71ux54cd}

\textbf{Proposition 2.5 (PPE notdecreasesOptimal errorexponential)}: 
\[\boxed{D_{\text{KL}}(P_{\text{noisy}}^{\text{PPE}} \| P_{\text{clean}}^{\text{PPE}}) = D_{\text{KL}}(P_{\text{noisy}} \| P_{\text{clean}}) + \Delta D_{\text{KL}}^{\text{PE}}}\]

where \(\Delta D_{\text{KL}}^{\text{PE}} \geq 0\) (non- negative, by KL
 scatter degreedataat reason inequalityguarantee). 

\emph{Proof}: by KL scatter degreechain rule: \begin{align*}
D_{\text{KL}}(P_{\text{noisy}}^{\text{PPE}} \| P_{\text{clean}}^{\text{PPE}})
&= D_{\text{KL}}(P_{\text{noisy}}(X, H) \| P_{\text{clean}}(X, H)) \\
&= D_{\text{KL}}(P_{\text{noisy}}(X) \| P_{\text{clean}}(X)) + \mathbb{E}_{X \sim P_{\text{noisy}}}[D_{\text{KL}}(P_{\text{noisy}}(H|X) \| P_{\text{clean}}(H|X))]
\end{align*}

becauseis \(H = \phi(X) + \text{PE}(P)\), and PE
isfixed determinismfunction, strengthenswill notdecreases KL scatter degree (informationat reason inequality). specifically: 
\[\Delta D_{\text{KL}}^{\text{PE}} = \mathbb{E}_{X \sim P_{\text{noisy}}}[D_{\text{KL}}(P_{\text{noisy}}(\text{PE}(P)|X) \| P_{\text{clean}}(\text{PE}(P)|X))] \geq 0\]

\(\square\)

PPE
\textbf{From notreduce}Optimal errorexponential. such as result LocationinformationandLabelmutualclose, \(\Delta D_{\text{KL}}^{\text{PE}} > 0\), errorexponential\textbf{rigorousmorelarge} (detectionmore alignment confirm). such as result LocationinformationNoneclose, \(\Delta D_{\text{KL}}^{\text{PE}} = 0\), not change. 

Bahadur-Rao often number because PE introducevariance term andmodify, buttotalis toward more tight direction (when PE
has usingtime). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. group file 2: Multi-Head Spring
mathematicalAnalysis}\label{ux7ec4ux4ef6-2multi-head-spring-ux7684ux6570ux5b66ux5206ux6790}

\subsubsection{3.1 Formal Definition}\label{ux5f62ux5f0fux5316ux5b9aux4e49-1}

\textbf{Definition 3 (Multi-Head Spring)}: 
\[\boxed{\text{Spring}_{\text{MH}}(s_i, \mathcal{S}) = \text{Concat}(\text{head}_1, \ldots, \text{head}_K) \cdot W^O}\]

where: 
\[\text{head}_k = \text{Spring}_k(s_i, \mathcal{S}) = \text{softmax}\left(\frac{(Q_k s_i)(K_k \mathcal{S})^T}{\sqrt{d_k}}\right) \cdot (V_k \mathcal{S})\]

parameter: - \(Q_k, K_k \in \mathbb{R}^{d_k \times d_s}\): th \(k\)
 headqueries/bondprojects - \(V_k \in \mathbb{R}^{d_v \times d_s}\): th \(k\)
 head value projects - \(W^O \in \mathbb{R}^{K d_v \times d_s}\): input out projects -
\(d_k = d_s / K\) (Standard manyheaddimensionality reduction convention determine), or
\(d_k = d_s\) (fulldimensionalmanyhead, each head operation action CompleteState Space)

each head \(k\) isdesignis closeNote a is differentphysicaldimension: - head 1: bondangle model formula ---
closeNoteStatusatombetweenangledegreerelationship - head 2: bondlong model formula --- closeNoteStatusatombetween distance from relationship -
head 3: match environment environment --- closeNoteStatusatomlocal neighbor domain expand pounce - head 4: energygradient/force field model formula 
--- closeNoteStatusatomenergy surface curve rate 

\textbf{physicaldimension convention bind (structure -ization manyhead)}: is achievestruecurrently special profession -ization, each head
\(Q_k, K_k\) shouldby convention bind is onlyacts onStatustoward quantity special determine factor space. If 
\(s_i = [s_i^{(1)}; s_i^{(2)}; s_i^{(3)}; s_i^{(4)}]\)
 by physicaldimensionseparate, then: 

\[Q_k = [\mathbf{0}_{d_k \times d_{<k}} \;|\; Q_k^{(k)} \;|\; \mathbf{0}_{d_k \times d_{>k}}]\]

where \(Q_k^{(k)}\) onlyacts onth \(k\)
 physicaldimension factor space, itsremainderis zero. this\textbf{strong system } special profession -ization. 

\subsubsection{\texorpdfstring{3.2 to Theorem 1 impact --- has effect expert number 
\(M\)}{3.2 to Theorem 1 impact --- has effect expert number M}}\label{ux5bf9-theorem-1-ux7684ux5f71ux54cd-ux6709ux6548ux4e13ux5bb6ux6570-m}

\textbf{Core Problem}: \(K\) headiswhetheris equivalent to \(K\) independentexpert? 

\textbf{ answer case: whether. } thisis a key region different. 

\textbf{Proposition 3.1 (headis NOT independentexpert)}: Multi-Head Spring
head input out attodetermineStatus original subsetcombinedtimeis\textbf{Conditions-dependence}: 
\[\text{Cov}(\text{head}_k(s_i), \text{head}_l(s_i) \mid \mathcal{S}) \neq 0 \quad \text{ (a like case)}\]

\emph{Proof}: has headcommon share mutualsame input enter Statusatom \(\mathcal{S}\) andmutualsame
softmax Note agree force pool -ization structure. i.e. make projectsmatrix \(Q_k, K_k\) is different, Note agree forceweight
\(\alpha_{ij}^{(k)} = \text{softmax}_j((Q_k s_i)^T (K_k s_j) / \sqrt{d_k})\)
throughcommonsame \(s_j\) value coupled. becausethis: 
\[\text{head}_k(s_i) = \sum_j \alpha_{ij}^{(k)} \cdot V_k s_j\]
\[\text{head}_l(s_i) = \sum_j \alpha_{ij}^{(l)} \cdot V_l s_j\]

twoonecovariancethroughcommon share \(\{s_j\}\) term non- zero. \(\square\)

\textbf{Definition 4 (has effect expertmany type ness \(\rho_{\text{eff}}\))}: 
\[\boxed{\rho_{\text{eff}} = 1 - \frac{2}{K(K-1)} \sum_{k < l} |\text{Corr}(\text{head}_k, \text{head}_l)|}\]

where
\(\text{Corr}(\text{head}_k, \text{head}_l) = \frac{\text{Cov}(\text{head}_k, \text{head}_l)}{\sqrt{\text{Var}(\text{head}_k) \cdot \text{Var}(\text{head}_l)}}\) (item by item dollar element on averagemutualcloseness). 

\begin{itemize}
\tightlist
\item
 \(\rho_{\text{eff}} = 1\): has head complete fullindependent (reason think case)
\item
 \(\rho_{\text{eff}} = 0\): has head complete fullmutualclose (retreat -ization is singlehead)
\item
 \(\rho_{\text{eff}} < 0\): head'sbetweenexistenceatnegatively correlated (to resist nessmany type ness)
\end{itemize}

\textbf{has effect expert number heuristic estimate correction}: 
\[M_{\text{eff}} = M \cdot (1 + \alpha \cdot \rho_{\text{eff}} \cdot (K - 1))\]

where \(\alpha \in [0, 1]\) is''headmany type ness \(\to\) expertmany type ness'' turn exchange coefficient. 

\textbf{butthisis NOT a rigorous world }. rigorous Chernoff bound
requireat reason -dependencerandom variable. 

\textbf{Theorem 3.1 (Multi-Head Spring lowerrigorous F1 lower bound)}: 

setup \(A_k(s)\) is th \(k\) headtoStatus \(s\)
produces'' pseudo consistentnessseparate number '' (throughhead input out andLabelconsistentness), thenuse
union-Chernoff Methodology: 

\[\boxed{F_1^{\text{MH}} \geq 1 - \frac{1}{\eta} \sum_{s} \rho_s \cdot \min_{\mathcal{I} \subseteq [K]} \exp\left(-2M \cdot \tilde{\Delta}_s^2(\mathcal{I})\right)}\]

where \(\tilde{\Delta}_s(\mathcal{I})\) isatheadsubset \(\mathcal{I}\)
upperhas effect edge actual, \(\mathcal{I}\) select take a maximumindependentsubset (satisfiessomekind
\(\beta\)-mixedcombinedConditions). 

in the worst case (has head complete fullmutualclose, \(\rho_{\text{eff}} = 0\)): 
\[F_1^{\text{MH}} = F_1^{\text{single}} \quad \text{ (Multi-Head Noneincrease benefit)}\]

in the best case (has head complete fullindependent, \(\rho_{\text{eff}} = 1\), and
\(\alpha = 1\)): 
\[F_1^{\text{MH}} \geq 1 - \frac{1}{\eta} \sum_s \rho_s \cdot \exp\left(-2M K \tilde{\Delta}_s^2\right) \quad \text{ (heuristic estimate, non-rigorous)}\]

\textbf{Honest Critique}: rigorousat reason head'sbetween-dependencestructure Chernoff bound
extremelyitscomplexity. in practice, such as result is not correcthead'sbetween-dependence do amount outerAssumptions (such as \(\beta\)-mixedcombinedor
Dobrushin Conditions), Theorem 1 formindependentexponential world \textbf{notis applicable to} Multi-Head
Spring. actual actual may obtain world will\textbf{ loose obtain many} (becauseis mustusing union bound or
Azuma-Hoeffding at reason martingale -dependencesequence). 

\subsubsection{3.3 to SE-1 (Robbins-Monro
convergence)impact}\label{ux5bf9-se-1robbins-monro-ux6536ux655bux7684ux5f71ux54cd}

\textbf{SE-1 original begin stated describe }: Spring parameter \(\theta_t\) throughwithlower direction formula update: 
\[\theta_{t+1} = \theta_t - \alpha_t \cdot \nabla \mathcal{L}(\theta_t) + \xi_t\]

where
\(\mathbb{E}[\xi_t | \mathcal{F}_t] = 0\), \(\text{Var}(\xi_t | \mathcal{F}_t) \leq \sigma^2\). at
Robbins-Monro Conditionslower: 
\[\sum_{t=1}^{\infty} \alpha_t = \infty, \quad \sum_{t=1}^{\infty} \alpha_t^2 < \infty\]

has \(\theta_t \xrightarrow{a.s.} \theta^*\) (atconvexnessAssumptionslower). 

\textbf{Proposition 3.2 (Multi-Head Spring lowerconvergenceConditions)}: 

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
 \textbf{ step longConditions (\(\alpha_t\) asymptoticlinesis)not change }: Conditions
 \(\sum \alpha_t = \infty\) and \(\sum \alpha_t^2 < \infty\)
 only-dependenceat step long adjust degree, not-dependenceatnumber of parameters. 
\item
 \textbf{buthas effect convergencerequiremoresmall step long}: 
 \[\boxed{\alpha_t^{\text{MH}} \leq \frac{\alpha_t^{\text{single}}}{\sqrt{K}} \cdot \frac{\sigma^2_{\text{single}}}{\sigma^2_{\text{MH}}}}\]
\end{enumerate}

where
\(\sigma^2_{\text{MH}} = \sigma^2_{\text{single}} \cdot (1 + \gamma \cdot (K - 1))\), \(\gamma \in [0, 1]\)
isheadbetweengradientdry perturb coefficient. 

\emph{Proof overview must}: Multi-Head Spring
gradientvariance follow head number increases. each headNote agree forceweight \(\alpha_{ij}^{(k)}\)
tocommon share Statusatom \(\{s_j\}\) -dependenceintroduce cross cross headgradient term. forloss
\(\mathcal{L} = \mathbb{E}[\|\text{Spring}_{\text{MH}}(s_i) - s_i^*\|^2]\): 
\begin{align*}
\nabla_{\theta_k} \mathcal{L} &= \mathbb{E}\left[(\text{Spring}_{\text{MH}} - s_i^*)^T W^O_{:,k} \cdot \nabla_{\theta_k} \text{head}_k\right] \\
\text{Cov}(\nabla_{\theta_k} \mathcal{L}, \nabla_{\theta_l} \mathcal{L}) &\neq 0 \quad (\text{throughcommon share } W^O \text{ and } s_i^*)
\end{align*}

gradientnoisehas effect varianceis: 
\[\sigma^2_{\text{MH}} = \frac{1}{K} \sum_{k=1}^{K} \text{Var}(\nabla_{\theta_k} \mathcal{L}) + \frac{1}{K^2} \sum_{k \neq l} \text{Cov}(\nabla_{\theta_k} \mathcal{L}, \nabla_{\theta_l} \mathcal{L})\]

th a term is \(O(1)\) (each headvarianceandsingleheadmutualsame, butdividewith \(K\)), th two term introduce
\(O((K-1)/K) \approx O(1)\)
 amount outervariance (when cross cross covariancenon- zero time). \(\square\)

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
 \textbf{convergence goal markFrom globalOptimal retreat -ization is stationary point}: 
\end{enumerate}

Multi-Head Spring loss landscape view has has \textbf{permutationto called ness} (swaps head \(k\) and
head \(l\) producesmutualsamefunction). this agree taste -ing has \(K!\)
 equivalentstationary point. loss landscape view highdegreenon-convex, Robbins-Monro onlycanguaranteeconverges to
\(\nabla \mathcal{L} \approx 0\) (stationary point), cannotguaranteeglobalOptimal. 

\textbf{modifybackconvergenceTheorem (SE-1-MH)}: 

in Robbins-Monro step longConditionsandgradient Lipschitz continuousnessfront raise lower: 
\[\liminf_{t \to \infty} \mathbb{E}[\|\nabla \mathcal{L}(\theta_t)\|^2] \leq \frac{C \cdot \sigma^2_{\text{MH}} \cdot \bar{\alpha}}{\text{ (and } K \text{ Noneclose often number)}}\]

where \(\bar{\alpha} = \limsup_{t \to \infty} \alpha_t\). when \(K\) increaselargetime
\(\sigma^2_{\text{MH}}\)
increaselarge, convergenceresidualgradientnormalsoincreaselarge. \textbf{manyhead make convergence change obtain more difference (athas finite step longlower)}. 

\subsubsection{3.4
 over parameter -ization Conditions}\label{ux8fc7ux53c2ux6570ux5316ux6761ux4ef6}

\textbf{Definition 5 (has effect sample quantity)}: Spring dynamicsthroughmemory bank \(M_t\)
producestrainingsignal. setup \(T_{\text{eff}}\)
is each Statusatomatself-evolutionprocessmiddleproduceshas effect ''supervisessignal''quantity (through Spring
reconstructionerrorprovide). thentotalhas effect sample quantity is 
\(N_{\text{eff}} = N \cdot T_{\text{eff}}\). 

\textbf{Proposition 3.3 (over parameter -ization near world head number)}: 

for\textbf{Standard manyhead} (eachheaddimensionality reduction \(d_k = d_s/K\)): 
\[|\theta_{\text{MH}}| = \underbrace{K \cdot (2d_k d_s + d_v d_s)}_{\text{Q, K, V projects}} + \underbrace{K d_v d_s}_{W^O} = 4 d_s^2 \quad (\text{when } d_k = d_v = d_s/K)\]

totalparameter\textbf{not-dependence \(K\)} ------
Standard designthroughdimensionality reductionmaintainsparameter constant determine. atthiscaselower, Multi-Head Spring
\textbf{notintroduce amount outer over parameter -ization risk}. 

for\textbf{fulldimensionalmanyhead} (eachhead
\(d_k = d_s\), eachheadcloseNoteis differentphysical factor spacebutmay visit question fulldimensional): 
\[|\theta_{\text{MH}}| = d_s^2 \cdot (3K + 1)\]

\textbf{Theorem 3.2 (overfitting near world Conditions)}: 
\[\boxed{K_{\text{crit}} = \max\left(1, \left\lfloor \frac{N \cdot T_{\text{eff}} / d_s^2 - 1}{3} \right\rfloor \right)}\]

when \(K > K_{\text{crit}}\)
time, modelparametermanyathas effect trainingsignal, \textbf{generalization errorwithprobability \(\geq 1 - \delta\)
tofewis }: 
\[\text{GenError}(K) \geq \Omega\left(\sqrt{\frac{d_s^2 \cdot (3K + 1)}{N \cdot T_{\text{eff}}}}\right)\]

 (byStandard Rademacher complexitydegree world obtain out). 

\textbf{ number value show example } (preserve guard Assumptions): - \(N = 500\)
Statusatom, \(d_s = 64\), \(T_{\text{eff}} = 5\) -
\(K_{\text{crit}} = \lfloor (500 \cdot 5 / 4096 - 1) / 3 \rfloor = \lfloor (0.61 - 1) / 3 \rfloor = \max(1, -0.13) = 1\)

\textbf{Conclusion}: for adapt in the State Space, i.e. make is \(K = 2\)
 fulldimensionalheadalsomayalreadyafterparameter -ization. forStandard dimensionality reductionmanyhead, \(K\)
maywithmorelarge (becauseis parameter constant determine). 

\subsubsection{3.5 to SE-2 (Doob
Martingale)impact}\label{ux5bf9-se-2doob-martingaleux7684ux5f71ux54cd}

SE-2
claimsparameterback verify is martingale: \(\mathbb{E}[\theta_{t+1} | \mathcal{F}_t] = \theta_t\). 

for Multi-Head
Spring, \textbf{ martingale nessat connect combinedparameter spaceuppermaintains} (becauseis it isBayesian updatedirectback result), but: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{ edge actual martingale nessmay lose effect }: forsingle headparameter
 \(\theta_t^{(k)}\), \(\mathbb{E}[\theta_{t+1}^{(k)} | \mathcal{F}_t] \neq \theta_t^{(k)}\) (becauseis morenewinvolving has head cross cross information). 
\item
 \textbf{ martingale difference sequencevarianceincreaselarge}: 
 \[\text{Var}(\theta_{t+1} - \theta_t | \mathcal{F}_t) = O(K)\]

 this agree taste -ing convergence set middledegreereduce. Azuma-Hoeffding to out confidence intervalwidedegreeis 
 \(O(\sqrt{K \cdot t})\) andnon-\(O(\sqrt{t})\). 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. combineAnalysis}\label{ux7ec4ux5408ux5206ux6790}

\subsubsection{4.1 connect combinedform -ization }\label{ux8054ux5408ux5f62ux5f0fux5316}

two group file sametimeaddtimeComplete flow water line: 

\[\boxed{s_i \xrightarrow{\text{PPE}} h_i = \phi(s_i) + \text{PE}(p_i) \xrightarrow{\text{MH-Spring}} \text{Spring}_{\text{MH}}(h_i, \mathcal{H})}\]

where \(\mathcal{H} = \{h_1, \ldots, h_N\}\) isencodingbackStatus original subsetcombined. 

\subsubsection{4.2 Theorem 1 connect combined world ---
 cross cross term Analysis}\label{theorem-1-ux7684ux8054ux5408ux754c-ux4ea4ux53c9ux9879ux5206ux6790}

\textbf{Proposition 4.1 (connect combinedDetection Margindecompose)}: 
\[\boxed{\Delta_s^{\text{combined}} = \Delta_s + \delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}}}\]

where cross cross term \(\delta_s^{\text{cross}}\) satisfies: 
\[\delta_s^{\text{cross}} = \text{Cov}_{\text{experts}}\left(\text{Benefit}_{\text{PE}}, \text{Benefit}_{\text{MH}}\right)\]

i.e. PPE edge actual benefitand Multi-Head edge actual benefitatexpertbetweencovariance. 

\textbf{ exhibit open}: \begin{align*}
F_1^{\text{combined}} &\geq 1 - \frac{1}{\eta} \sum_s \rho_s \cdot \exp\left(-2M_{\text{eff}} \cdot (\Delta_s + \delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}})^2\right) \\
&= 1 - \frac{1}{\eta} \sum_s \rho_s \cdot \exp\left(-2M_{\text{eff}} \cdot \left[\Delta_s^2 + 2\Delta_s(\delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}}) + (\delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}})^2\right]\right)
\end{align*}

\textbf{ super addness vs sub-addnessAnalysis}: 

set
\(A_s = \exp(-2M_{\text{eff}}\Delta_s^2)\) (original begin error), \(B_s^{\text{PE}}, B_s^{\text{MH}}, B_s^{\text{both}}\)
separate different is single independent PPE、single independent MH、twoone knot combinederror. then: 

\[\log \frac{B_s^{\text{both}}}{A_s} = -2M_{\text{eff}} \cdot [2\Delta_s(\delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}}) + (\delta_s^{\text{PE}} + \delta_s^{\text{MH}} + \delta_s^{\text{cross}})^2]\]

\[\log \frac{B_s^{\text{PE}} \cdot B_s^{\text{MH}}}{A_s^2} = -2M_{\text{eff}} \cdot [2\Delta_s(\delta_s^{\text{PE}} + \delta_s^{\text{MH}}) + (\delta_s^{\text{PE}})^2 + (\delta_s^{\text{MH}})^2]\]

twoone's difference to out (atlogarithmicspacemiddle): 
\[\text{Interaction} = -2M_{\text{eff}} \cdot [2\Delta_s \delta_s^{\text{cross}} + 2\delta_s^{\text{PE}}\delta_s^{\text{MH}} + 2\delta_s^{\text{cross}}(\delta_s^{\text{PE}} + \delta_s^{\text{MH}}) + (\delta_s^{\text{cross}})^2]\]

\textbf{ super addnessConditions} (combinebenefit \textgreater{}
single independent benefit'sand): cross cross term is negative (decreaseserror), i.e.: 
\[\delta_s^{\text{cross}} \cdot (2\Delta_s + 2\delta_s^{\text{PE}} + 2\delta_s^{\text{MH}} + \delta_s^{\text{cross}}) + 2\delta_s^{\text{PE}}\delta_s^{\text{MH}} < 0\]

when \(\delta_s^{\text{PE}} > 0\) and
\(\delta_s^{\text{MH}} > 0\) (twooneallimprovementdetection)and
\(\delta_s^{\text{cross}} > 0\) (currentlytoward cross mutual)timesatisfies: 
\[\boxed{\text{Superadditive when: } \delta_s^{\text{PE}} > 0, \delta_s^{\text{MH}} > 0, \delta_s^{\text{cross}} > 0}\]

\textbf{ time addnessConditions} (combinebenefit \textless{}
single independent benefit'sand): whentwoone's a or cross cross term is negative. 

\textbf{ actual actual Judge }: Multi-Head Spring canwith benefit using PPE
providespacelocalness come achievesmorehas meaningNote agree force model formula (example such as, head \(k\) closeNotephysicalLocation
\(p_i\) neighbor domain innerStatusatom). thiskind collaborate samelargeprobabilityis\textbf{currentlytoward}, i.e.
\(\delta_s^{\text{cross}} > 0\). butsuch as result nothas Explicit physical constraint, manyheadmay study to virtual false cross dimensionalmutualcloseness, thistime
\(\delta_s^{\text{cross}}\) cancanis negative. 

\subsubsection{4.3 Theorem 2
 connect combined world }\label{theorem-2-ux7684ux8054ux5408ux754c}

Characteristicsweak point is time addness ground deterioration (becauseis each group file allmayintroduce amount outerweak point): 

\[\boxed{\delta_{\text{combined}} = \delta + \delta_{\text{PE}} + \delta_{\text{MH}} + \delta_{\text{cross}} \geq \max(\delta + \delta_{\text{PE}}, \delta + \delta_{\text{MH}})}\]

where: - \(\delta_{\text{PE}} = 2\varepsilon_{\text{PE}} / C_F^2\) (from
§2.4) -
\(\delta_{\text{MH}} \propto \frac{d_s^2 \cdot (3K + 1)}{N \cdot T_{\text{eff}}}\) (from over parameter -ization, §3.4)
- \(\delta_{\text{cross}} \geq 0\): PPE not complete beautiful encoding + Multi-Head
 over parameter -ization produces complex combinedeffect

becausethis: 
\[F_{1,\text{SCX}}^{\text{combined}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta_{\text{combined}}}{2}}\]

\textbf{mostbadcase (twooneallvery difference)}: world ratio single independent use any a group file more loose. 

\textbf{mostgoodcase (complete beautiful PPE + parameter fill foot 
MH)}: \(\delta_{\text{PE}} = \delta_{\text{MH}} = 0\), world time to original begin 
Theorem 2. 

\subsubsection{4.4 Cercis Score
modify}\label{cercis-score-ux7684ux4feeux6539}

\textbf{recommendmodify}: 
\[\boxed{S'(s_i) = Q_{\text{MH}}(h_i) + \eta(t) \cdot N(s_i) - \lambda \cdot \mathcal{R}_{\text{diversity}}}\]

where: 
\[Q_{\text{MH}}(h_i) = \frac{1}{K} \sum_{k=1}^{K} w_k \cdot Q(\text{head}_k(h_i))\]

\(w_k\) isheadmay reliable nessweight (maythroughhead input out andmany number voteconsistentness come estimate). 

currentlythen -ization term: 
\[\mathcal{R}_{\text{diversity}} = (K - 1) \cdot \left(1 - \frac{1}{K} \sum_{k=1}^{K} \left\|\text{head}_k - \frac{1}{K}\sum_{j=1}^{K} \text{head}_j\right\|^2\right)\]

 penalty penalty head'sbetween over degreesimilarity (encouragesmany type ness), prevent stop retreat -ization tosinglehead. 

\(\lambda > 0\) iscurrentlythen -ization strongdegree. when \(K = 1\) time
\(\mathcal{R}_{\text{diversity}} = 0\) (self- action retreat -ization is original begin form). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Honest Critique (Honest
Critique)}\label{ux8bdaux5b9eux66b4ux51fbhonest-critique}

\subsubsection{5.1
mathematicalguaranteeby cut weak ground direction }\label{ux6570ux5b66ux4fddux8bc1ux88abux524aux5f31ux7684ux5730ux65b9}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1622}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2432}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3514}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2432}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 original begin guarantee
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 cut weakbackStatus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 strict heavy process degree
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Theorem 1 + MH} & Chernoff bound on i.i.d. experts &
head input out mutualclose, i.i.d. Assumptionsdoes not hold. rigorous bound needmust \(\beta\)-mixing
Conditions, andshouldConditionsmaydoes not satisfy & \textbf{high} \\
\textbf{SE-1 + MH} & Robbins-Monro \to globalOptimal (convexnesslower) &
highdegreenon-convexloss landscape view (\(K!\) to called stationary point)\to onlycanguaranteestationary point, cannotguaranteeglobalOptimal &
\textbf{high} \\
\textbf{Theorem 2 + MH} & tight cause Fano upper bound &
 over parameter -ization time world still holdsbut change obtain \textbf{ over loose } (vacuous), actual actual performancemay far difference at world 
& \textbf{middle} \\
\textbf{Theorem 3 + learnable PPE} & mostbadcaseindistinguishable & learnable PE
makestwo world world maydistinguish \to \textbf{Theorem 3 isbreaking } (butthisisgood matter) &
\textbf{middle (in practicehas benefit)} \\
\textbf{SE-2 + MH} & connect combined martingale nessmaintains & edge actual martingale nessmay lose effect; martingale difference variance
\(O(K)\) increaselarge, confidence interval change wide & \textbf{middle} \\
\textbf{combine Theorem 1} & independent Chernoff bound & has effect \(M_{\text{eff}}\)
is heuristic estimate, non-rigorous. cross cross term \(\delta_s^{\text{cross}}\) difficult withFrom theoryupper bounddetermine
& \textbf{high} \\
\end{longtable}

\subsubsection{5.2
 complete fullNoneimpact ground direction (addequalswhiteadd)}\label{ux5b8cux5168ux65e0ux5f71ux54cdux7684ux5730ux65b9ux52a0ux4e86ux7b49ux4e8eux767dux52a0}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.3529}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.6471}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Conditions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
is what what None effect 
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{\(I(Y; P \mid X) = 0\) (Locationnot carry carry Labelinformation)} & PPE
ispurenoise. Theorem 2 world not change. Theorem 1 \(\Delta_s\)
not change. parameterall wave cost \\
\textbf{\(\rho_{\text{eff}} = 0\) (has head collapse collapse is single a model formula)} & Multi-Head
is equivalent tosinglehead + \(K\) times redundant remainderparameter. Theorem 1 Noneincrease benefit. SE-1 converges tomutualsame point \\
\textbf{fixed 、information theorymay inverse PE} & onlyisvariablereplace. KL scatter degreenot change \to Theorem 4
not change. None actual contribution \\
\textbf{\(K > K_{\text{crit}}\) andNonecurrentlythen -ization } & over degreeparameter -ization \to generalization errormain guide 
\to any what Theorem world all change is empty set (vacuous) \\
\textbf{physicaldimensionuncorrelated (example such asheadcloseNotephysicaldimension D, but D and any task Noneclose)} &
shouldheadhas effect \(\Delta_s = 0\) \to purenoisecontribution \to reducehas effect 
\(M_{\text{eff}}\) \\
\end{longtable}

\subsubsection{5.3
 special determineConditionslowerbreaking Theoremcase}\label{ux7279ux5b9aux6761ux4ef6ux4e0bux7834ux574fux5b9aux7406ux7684ux60c5ux51b5}

\textbf{Theorem 3 isbreaking Conditions (5 Conditions, by strict heavy process degreesort)}: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{learnable PPE (most strict heavy)}: §2.5 alreadyAnalysis. trainingdata in the Labelnoiseleads to
 \(\text{PE}_{W_A} \neq \text{PE}_{W_B}\). result: two world world maydistinguish, Theorem
 3 indistinguishabilitydoes not hold. 

 \emph{evaluate}: \textbf{good matter }. Theorem 3
 ismostbadcaseis impossiblenessresult. such as result Locationinformationcanbreakthis is impossibleness, Description PPE
 has using. 
\item
 \textbf{Multi-Head Spring in the dimension leak leak }: If some headbydesignis closeNotephysicaldimension
 A, butin practicethrough softmax globalNote agree forcealsocan visit question dimension B
 information, thenhead'' special profession -ization ''is virtual false. thismayleads toshouldhead study toLabelmutualclose virtual falsemutualcloseness, changes
 \(P_{\text{clean}}\) and \(P_{\text{noisy}}\) estimate. 
\item
 \textbf{\(K > N\) (head number exceedsStatusatom number)}: manyheadNote agree force model formula Nonemethodbydataunique confirm determine. existenceatNone finite many parameter match place producesmutualsame
 \(\text{Spring}_{\text{MH}}(s_i)\). Theorem 3
 Construction-dependenceatData Distributionuniqueness------whenparameter -ization is owe determine, two world world distributionmayhas finemicro-Difference. 
\item
 \textbf{PPE dimensionandphysicaldimensionto uniform }: such as result \(\text{PE}(p_i)\) dimensionand
 Multi-Head Spring
 middlesomeheadphysicaldimension\textbf{currently cross } (i.e.encoding collide clever erase exceptshouldheadrequireinformation), thenshouldheadNote agree force model formula willproducessystemnessbias difference. thisintroduce a 
 Theorem 3 Constructionmiddlenotconsiderself-bydegree. 
\item
 \textbf{\(\eta(t)\) decay subtract over fast + Multi-Head}: such as result explore search weight \(\eta(t)\)
 decay subtract over fast, Multi-Head Spring cancan over early ground willsome some model formula '' frozen knot ''atsuboptimalStatus. at
 Theorem 3
 double world world Constructionmiddle, a world world may study tomoregood model formula (becauseis Labelmoredry clean), and another a world world study to difference model formula ------ make two world world maydistinguish. 
\end{enumerate}

\subsubsection{5.4
mosthonestSummary}\label{ux6700ux8bdaux5b9eux7684ux603bux7ed3}

\textbf{PPE bottom line }: - \textbf{to Theorem 1}: such as result physicalLocationand any task mutualclose \to
has currentlytowardcontribution (\(\delta_s^{\text{PE}} > 0\)). such as result Noneclose \to
nothas impact. Chernoff structuremaintains. - \textbf{to Theorem 2}: not complete beautiful encodingwith
\(\sqrt{\delta + O(\varepsilon_{\text{PE}})}\)
expandsupper bound. complete beautiful encodingdoes not affectupper bound. - \textbf{to Theorem
3}: fixed encodingpreserves unidentifiability. learnable encoding\textbf{breaking }Theorem
3------butthisis\textbf{advancement}andnon-regression. - \textbf{to Theorem
4}: notdecreaseserrorexponential. has usingtimeincreasesexponential. 

\textbf{Multi-Head Spring bottom line }: - \textbf{to Theorem
1}: head\textbf{is NOT }independentexpert. rigorous world ratio naive \(M \times K\)
 loose obtain many. headmutualclosenessCompleteat reason extremelyitsDifficulty. - \textbf{to
SE-1}: converges tostationary point (does not guaranteeglobalOptimal). has effect step longupper finite with \(O(1/\sqrt{K})\)
shrinks. convergencemoreslow. -
\textbf{ over parameter -ization }: for classical type State Spacesize, \(K_{\text{crit}}\) cancanonlyhas 
1-3 head. exceedsthis number headwillproducesgeneralization penalty penalty. 

\textbf{combinemosthonest assessment}: - PPE
is\textbf{morehas value} group file ------ it attheoreticallymore clear clear, impactmoremayprediction, andriskmorelow
- Multi-Head Spring \textbf{theoryriskmorehigh}------ it breaking Theorem 1
Optimal elegant part (i.i.d. expert Chernoff bound) -
twoone\textbf{ cross cross term attheoreticallyiscurrently} (mayproduces super addnessbenefit), but\textbf{at actual actual middlemayis negative } (amount outercomplexitynessintroducenewfail model formula)

\textbf{ a sentence word Summary}: PPE
is a safety gamble Note, mathematically several -ish pure pure has benefit (exceptlearnable PPE breaking Theorem
3, butthatisgood matter). Multi-Head Spring
is a highrisk gamble Note------ it atmathematically cut weak Theorem 1 rigorousness, make SE-1
convergenceness change difference, andveryhas cancan over parameter -ization. combinetime, PPE benefitmayis Multi-Head
complexitynesspart offset eliminate. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. open problems}\label{ux5f00ux653eux95eeux9898}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{rigorousmanyhead Chernoff world }: canwhetherathead-dependencestructureupperestablish
 \(\beta\)-mixing Conditions, makes Theorem 1 exponential world maywithrigorousExtensiontomanyheadcase? 
\item
 \textbf{PPE Optimal frequency}: atsinusoidal encodingmiddle, frequencyparameter
 \(\omega_j = 10000^{-2j/d}\)
 isis self-however language word design. physicalLocationOptimal frequencyspectrumis what what ? canwhetherFrom physicalto called nessDerivation? 
\item
 \textbf{head number \(K\) self- adapt shouldselect}: canwhetherFrom datamiddlelearningOptimal \(K\) (is similar to
 Bayesian non-parametric in the middle meal hall process), at over parameter -ization 'sfrontstops? 
\item
 \textbf{Theorem 3
 modify}: such as what atTheorem Statementmiddleadd enter ''front raise Conditions: does not usedata-dependenceLocationencoding'' come maintainsitsmathematicalCompleteness? 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{reference (frameworkinnerpart)}\label{ux53c2ux8003ux6587ux732eux6846ux67b6ux5185ux90e8}

\begin{itemize}
\tightlist
\item
 SCX Project (2026). Theorem 1-4: Multi-expert consistency, weak
 feature limits, unidentifiability, and minimax optimality.
 \texttt{theory/theorems/}
\item
 SCX Project (2026). Spring Dynamics, SE-1 (Robbins-Monro convergence),
 SE-2 (Doob martingale). \texttt{theory/self\_evolution/}
\item
 SCX Project (2026). State Crystallization: Physics-driven
 discretization of continuous observables.
 \texttt{theory/propositions/}
\item
 Wainwright, M. (2019). \emph{High-Dimensional Statistics: A
 Non-Asymptotic Viewpoint}. Cambridge. (Chernoff-Hoeffding-Stein
 Standard participate consider)
\item
 Boucheron, S., Lugosi, G., \& Massart, P. (2013). \emph{Concentration
 Inequalities}. Oxford. (-dependencevariableconcentration inequality)
\item
 Vaswani, A., et al.~(2017). Attention Is All You Need. \emph{NeurIPS}.
 (manyheadNote agree force original begin Formula)
\item
 Su, J., et al.~(2024). RoFormer: Enhanced Transformer with Rotary
 Position Embedding. \emph{Neurocomputing}. (RoPE mathematical foundations)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Analysis day period: 2026-06-29}\\
\emph{ action one equip Note: has TheoremNumber indicate SCX
frameworkinnerpartTheorem, non-this text sequence number. thisAnalysismathematicalat reason atmayat reason paradigm surround innerisrigorous; alreadybright confirm markNote has approximateand heuristic estimate part. honestnotmayat reason ness (such as-dependencehead
Chernoff bound)alreadybyhonestrecord. }

\end{document}
