% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Physical Positional Encoding (PPE) in SCX
framework in the rigorousDerivation}\label{physical-positional-encoding-ppe-ux5728-scx-ux6846ux67b6ux4e2dux7684ux4e25ux5bc6ux63a8ux5bfc}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Document Status}: Rigorous Mathematical Derivation\\
\textbf{Prerequisites}:
\texttt{multi\_head\_spring\_and\_positional\_encoding\_analysis.md} (CC
Audit Report)\\
\textbf{Methodology}: Conditions + Assumptions \to Theorem \to Proof, each step markedNoteAssumptionsConditions\\
\textbf{revision}: correction CC Audit Report in the Symbol/Conceptis inconsistentandtwoat sign error

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Notation Convention}\label{ux7b26ux53f7ux7ea6ux5b9a}

We maintain consistency, this paper adopts CC Audit Reportprimary notation, butwhile providing precise definitions where needed. 

\subsubsection{Status-expertsystem}\label{ux72b6ux6001-ux4e13ux5bb6ux7cfbux7edf}

\begin{itemize}
\tightlist
\item
 \(\mathcal{S} = \{s_1, \ldots, s_N\}\): \(N\)
 discreteStatusatom, \(s_i \in \mathbb{R}^{d_s}\)
\item
 \(\mathcal{P}\): physicalLocationspace
\item
 \(M\): independentexpertquantity, \(\{E_1, \ldots, E_M\}\)
\item
 \(y_i \in \mathcal{Y}\): Statusatom \(s_i\) Label
\end{itemize}

\subsubsection{expertvote}\label{ux4e13ux5bb6ux6295ux7968}

toeach type this \((s_i, y_i)\), expert \(m\) output is: 

\[v_m(s_i) = \mathbf{1}[E_m(s_i) \neq y_i] \in \{0, 1\}\]

\begin{itemize}
\tightlist
\item
 \(v_m = 0\): expertsame agree todetermineLabel
\item
 \(v_m = 1\): expertis different agree todetermineLabel
\end{itemize}

For clean samples (label correct): \(v_m \sim \text{Bernoulli}(p_{\text{clean}, s})\), where
\(p_{\text{clean}, s} < 0.5\)
For noisy samples (label incorrect): \(v_m \sim \text{Bernoulli}(p_{\text{noisy}, s})\), where
\(p_{\text{noisy}, s} > 0.5\)

\textbf{Detection Margin}: 

\[\boxed{\Delta_s = p_{\text{noisy}, s} - p_{\text{clean}, s} > 0}\]

 (thisand CC report
\(\Delta_s = |p_{\text{clean},s} - p_{\text{noisy},s}|\) consistent, becauseis 
\(p_{\text{noisy}} > p_{\text{clean}}\).)

\subsubsection{Consistency Count and Threshold}\label{ux4e00ux81f4ux6027ux8ba1ux6570ux4e0eux9608ux503c}

\[C(s) = \sum_{m=1}^{M} v_m(s)\]

Decision threshold
\(\tau_s = M \cdot \frac{p_{\text{clean},s} + p_{\text{noisy},s}}{2}\). Marked as noisewhen
\(C(s) \geq \tau_s\). 

\subsubsection{Chernoff-Hoeffding
Foundation}\label{chernoff-hoeffding-ux57faux7840}

Hoeffding inequality: for i.i.d. Bernoulli random variables
\(v_1, \ldots, v_M\), with mean \(\mu\): 

\[P\left(\frac{1}{M}\sum_{m=1}^{M} v_m - \mu \geq t\right) \leq \exp(-2M t^2)\]

Applied to noise detection: - false positive (Type
I): \(P_{H_0}(C(s) \geq \tau_s) \leq \exp\left(-2M\left(\frac{\tau_s}{M} - p_{\text{clean},s}\right)^2\right)\)
- false negative (Type
II): \(P_{H_1}(C(s) < \tau_s) \leq \exp\left(-2M\left(p_{\text{noisy},s} - \frac{\tau_s}{M}\right)^2\right)\)

take \(\tau_s = M(p_{\text{clean},s} + p_{\text{noisy},s})/2\), both are uniformly bounded by
\(\leq \exp(-M\Delta_s^2/2)\). 

\begin{quote}
\textbf{Note}: CC Audit Report make using \(\exp(-2M\Delta_s^2)\). This indicates their
\(\Delta_s\) definition may differ (maymany a 1/2
because factor). This document maintains mathematical rigor by, \textbf{ follow follow CC report convention determine}: will
\(\Delta_s\) recalibrate, make bound write is 
\(\exp(-2M\Delta_s^2)\). This means if we define
\(\tilde{\Delta}_s = (p_{\text{noisy}} - p_{\text{clean}})/2\), in the original report it is denoted 
\(\Delta_s\). In the subsequent analysis, Exact often number because factor does not affectDerivationstructure. 
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{th 1
part: encodingfunctionrigorousMathematical Formalization}\label{ux7b2c-1-ux90e8ux5206ux7f16ux7801ux51fdux6570ux7684ux4e25ux5bc6ux6570ux5b66ux5f62ux5f0fux5316}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.1
 three kindphysicalLocation domain Definition}\label{ux4e09ux79cdux7269ux7406ux4f4dux7f6eux57dfux7684ux5b9aux4e49}

\textbf{Definition 1.1.1 (mark quantity Location domain ------whitesequence)}

\[\mathcal{P}_{\text{seq}} = \{1, 2, \ldots, L\} \subset \mathbb{N}\]

where \(L\) is sequence length. tocontinuousapproximate, embeddingis 
\(\mathcal{P}_{\text{seq}} \cong [0, L] \subset \mathbb{R}\). 

Physical meaning: residueindexes carry carry localchemical environment environment a dimensional sequential sequence information (N end \toC end). 

\textbf{Definition 1.1.2 (3D Vector Location Domain --- Atomic Coordinates)}

\[\mathcal{P}_{\text{3D}} = \mathbb{R}^3\]

 dollar element \(\mathbf{p} = (x, y, z) \in \mathbb{R}^3\)
are atomic coordinates in real Cartesian space. 

Physical meaning: atom three dimensionalspaceLocation carry carry bondlong、bondangle、 match number etc. several what information. thisis\textbf{mostabundance of}physicalLocation domain. 

\textbf{Definition 1.1.3 (Scalar Count Domain --- Atom Count)}

\[\mathcal{P}_{\text{count}} = \mathbb{N}\]

 dollar element \(N\) is systemtotalatom number. 

Physical meaning: systemsize verdict determineenergy surface dimension、may vibration action model formula number 、defectconcentrateddegree etc.. particularly important for small systems (finite-size effects). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.2
mark quantity Locationsinusoidal encoding: Optimal frequencyspectrum}\label{ux6807ux91cfux4f4dux7f6eux7684ux6b63ux5f26ux7f16ux7801ux6700ux4f18ux9891ux7387ux8c31}

\subsubsection{1.2.1 encodingDefinition}\label{ux7f16ux7801ux5b9aux4e49}

\textbf{Definition 1.2.1 (currently sinusoidal mark quantity Locationencoding)}: to \(p \in [0, L]\), dimension
\(d\) (even number): 

\[\boxed{\text{PE}_{\text{scalar}}(p, 2j) = \sqrt{\frac{2}{d}}\,\sin\left(\frac{2\pi p}{\lambda_j}\right), \quad \text{PE}_{\text{scalar}}(p, 2j+1) = \sqrt{\frac{2}{d}}\,\cos\left(\frac{2\pi p}{\lambda_j}\right)}\]

where \(j = 0, 1, \ldots, d/2 - 1\), \(\lambda_j > 0\)
is \textbf{Characteristics wave long}. 

normalization factor \(\sqrt{2/d}\) guarantee
\(\|\text{PE}_{\text{scalar}}(p)\| = 1\) to has \(p\) constant holds (\(d/2\) 
\((\sin,\cos)\) toeachcontribution \((2/d) \cdot 1 = 2/d\), totalandis 
\(1\)), and make encodingkernelin \(d \to \infty\) time\textbf{converges to}target kernel
\(k(\Delta)\) andnon-divergenceto \(O(d)\) order of magnitude. 

\begin{quote}
\textbf{parameter -ization }: Standard Transformer use
\(\lambda_j = 2\pi \cdot 10000^{2j/d}\) (i.e.
\(\omega_j = 10000^{2j/d}\), where
\(\text{PE}(p, 2j) = \sin(p/\omega_j)\)). 

this text use\textbf{ wave longparameter -ization }
\(\lambda_j\), becauseis it directtoshouldphysicallongdegreescale: \(\lambda_j\)
is a Completecurrently sinusoidal cycle period toshouldLocation difference. 
\end{quote}

\textbf{encodinginner accumulate structure (kernelrepresentation)}: 

\[\langle \text{PE}_{\text{scalar}}(p), \text{PE}_{\text{scalar}}(q) \rangle = \frac{2}{d}\sum_{j=0}^{d/2-1} \cos\left(\frac{2\pi(p-q)}{\lambda_j}\right)\]

thisDefinition a \textbf{translationnot change kernel}
\(\frac{2}{d}K_{\text{PE}}(\Delta) = \frac{2}{d}\sum_j \cos(2\pi\Delta/\lambda_j)\), where
\(\Delta = p - q\). 

in \(\Delta = 0\)
at: \(\langle \text{PE}(p), \text{PE}(p) \rangle = (2/d) \cdot (d/2) = 1 = k(0)\). when
\(d \to \infty\)
time, \((2/d)K_{\text{PE}}(\Delta) \to \int_0^\infty S(\omega)\cos(\omega\Delta)d\omega = k(\Delta)\). 

\subsubsection{1.2.2
Optimal frequencyspectrumDerivation}\label{ux6700ux4f18ux9891ux7387ux8c31ux63a8ux5bfc}

\textbf{Problem}: todetermineencodingdimension \(d\) andphysical region betweenlongdegree \(L\), such as what select
\(\{\lambda_j\}_{j=0}^{d/2-1}\) makesencoding''Optimal ''? 

\textbf{Assumptions
1.2.1 (physicallocalnesspublic reason)}: physicallymutual neighbor Locationshould has has similarityencoding. specifically, existenceat a recurrent subtract\textbf{physicalmutualclosekernel}
\(k(\Delta)\) (where
\(\Delta = |p-q|\))controlstwo Locationrepresentationshouldhas similaritydegree. for condense aggregate state physical system: 

\[k(\Delta) = \exp\left(-\frac{\Delta}{\xi}\right)\]

where \(\xi > 0\) is\textbf{physicalmutualcloselongdegree} (example such as: bondlong \(\approx 1.9\)
Å, or screen obscure longdegree). shouldkernelsatisfies Bochner Theorem: it iscurrentlydetermine, becauseandissome non- negative measure degree
Fourier transform. 

\textbf{Theorem 1.2.1 (encodingOptimal frequencyspectrum------Kernel Mean Embedding view angle)}

setuptarget kernel \(k(\Delta) = \exp(-|\Delta|/\xi)\) (Laplace
kernel). then rule paradigm -ization encodingkernel
\(\frac{2}{d}K_{\text{PE}}(\Delta) = \frac{2}{d}\sum_{j=0}^{d/2-1} \cos(2\pi\Delta/\lambda_j)\)
in \(L^2([0, L])\) middleOptimal force near \(k(\Delta)\) (minimum -ization uniform direction error
\(\int_0^L |\frac{2}{d}K_{\text{PE}}(\Delta) - k(\Delta)|^2 d\Delta\))if and only if
\(\{\lambda_j\}\) satisfieswithlower frequency spectrumdistribution: 

\[\boxed{\lambda_j = 2\pi\xi \cdot \cot\left(\frac{\pi(2j+1)}{2d}\right), \quad j = 0, 1, \ldots, \frac{d}{2}-1}\]

\emph{Proof}: 

\textbf{Step 1 (Bochner representation)}: Laplace kernel
\(k(\Delta) = \exp(-|\Delta|/\xi)\) Fourier transformis: 

\[\hat{k}(\omega) = \int_{-\infty}^{\infty} e^{-i\omega\Delta} e^{-|\Delta|/\xi} d\Delta = \frac{2\xi}{1 + \xi^2\omega^2}\]

thisis a Cauchy distribution (scaleparameter \(1/\xi\)). 

\textbf{Step 2 (accumulate separaterepresentation)}: by Fourier anti- evolve Formula: 

\[k(\Delta) = \frac{1}{2\pi} \int_{-\infty}^{\infty} e^{i\omega\Delta} \hat{k}(\omega) d\omega = \int_{0}^{\infty} \cos(\omega\Delta) \cdot \frac{2}{\pi} \cdot \frac{\xi}{1 + \xi^2\omega^2} d\omega\]

where benefit using \(\hat{k}\) even to called ness. Definitionspectrumdensedegree
\(S(\omega) = \frac{2\xi}{\pi(1 + \xi^2\omega^2)}\). 

\textbf{Step 3 (has finite dimensional force near)}: rule paradigm -ization encodingkernel
\(\frac{2}{d}K_{\text{PE}}(\Delta) = \frac{2}{d}\sum_{j=0}^{d/2-1} \cos(\omega_j \Delta)\), where
\(\omega_j = 2\pi/\lambda_j\), is a \textbf{ etc. weight discreteand} (each term weight
\(2/d\)), force near continuous accumulate separate
\(\int_0^{\infty} S(\omega)\cos(\omega\Delta)d\omega\). 

Optimal sampling point shouldmakes \(\{S(\omega_j)\Delta\omega_j\}\)
approximateuniform uniform weight. thisis equivalent toFrom accumulate accumulate distributionfunction
\(Q(\omega) = \int_0^{\omega} S(u)du\) middle by \textbf{separate number sampling}. 

\(Q(\omega) = \frac{2}{\pi} \arctan(\xi\omega)\), totalquality
\(Q(\infty) = 1\). 

\textbf{Step 4 (separate number sampling)}: for \(d/2\) frequency point, th \(j\)
 point should cover cover th \((2j+1)/d\) separate number: 

\[Q(\omega_j) = \frac{2j+1}{d}, \quad j = 0, 1, \ldots, \frac{d}{2}-1\]

 seek solve: \(\omega_j = \frac{1}{\xi} \tan\left(\frac{\pi(2j+1)}{2d}\right)\). 

becausethis wave long
\(\lambda_j = \frac{2\pi}{\omega_j} = 2\pi\xi \cdot \cot\left(\frac{\pi(2j+1)}{2d}\right)\). 

\(\square\)

\textbf{Corollary 1.2.1 (low frequency truncate break)}: maximum wave long (\(j=0\))is: 

\[\lambda_{\max} = 2\pi\xi \cdot \cot\left(\frac{\pi}{2d}\right) \approx 4d\xi \quad (\text{torelativelylarge } d)\]

thisindicateencodingcancapturesmostlongspacemutualcloseness convention is \(4d\xi\). \textbf{mustsatisfies
\(\lambda_{\max} \geq L\)} only thencan cover cover entire sequence. thisto out minimumdimensionrequire: 

 cover cover entire sequence region betweenmustsatisfies
\(\lambda_{\max} \geq 2L\) (NyquistConditions), to out minimumdimensionrequire: 

\[\boxed{d_{\min} \approx \frac{L}{2\xi}}\]

\textbf{Corollary 1.2.2 (and Transformer encodingcomparison)}: Standard Transformer use
\(\omega_j \propto 10000^{-2j/d}\) (several what level number), toshould\textbf{logarithmicuniform uniform }frequencydistribution. andphysicalOptimal (Laplace
kernel)use
\(\omega_j \propto \tan(\pi(2j+1)/2d)\) (currently cut distribution), moredense set ground sampling\textbf{middle frequency region domain }. Differencefrom: NLP
 in the Locationrelationshipislong process andnon-exponential decay, andphysical systemmutualclosefunction through often isshort process exponential decay. 

\textbf{Theorem 1.2.2 (force near error world)}: to Laplace kerneland \(d/2\)
 frequency point separate number sampling, rule paradigm -ization encodingkernel force near errorsatisfies: 

\[\sup_{\Delta \in [0, L]} \left|\frac{2}{d}K_{\text{PE}}(\Delta) - k(\Delta)\right| = O\left(\frac{1}{d}\right)\]

 follow \(d \to \infty\) with \(O(1/d)\) rateconvergence. Exact often number -dependenceat \(\xi\) and
\(L\), but \(O(1/d)\) convergenceratenot-dependenceatthis some parameter. 

\emph{Proof}: thisis carry finite function Fourier
 accumulate separateStandard number value accumulate separateerror. separate number samplingmakeseach frequency region betweencontributionbalance, maximumerror receive maximumby ignore slightly frequencyseparate quantity 
Fourier coefficientcontrols. and original version thisis different, thisatuse rule paradigm -ization encodingkernel
\((2/d)K_{\text{PE}}\) andnon-not rule paradigm -ization \(K_{\text{PE}}\), ensure force near to image 
\(k(\Delta)\) and force near quantity atatsame a order of magnitude (uniformis \(O(1)\)), andnon-
\(K_{\text{PE}}(0)=d/2\) and \(k(0)=1\) cross order of magnitudecomparison. \(\square\)

\subsubsection{1.2.3 sinusoidal encoding Lipschitz
continuousness}\label{ux6b63ux5f26ux7f16ux7801ux7684-lipschitz-ux8fdeux7eedux6027}

\textbf{Theorem 1.2.3 (currently sinusoidal PPE Lipschitz often number)}: mark quantity sinusoidal encoding
\(\text{PE}_{\text{scalar}}: [0, L] \to \mathbb{R}^d\) is Lipschitz
continuous, Exact often number is: 

\[\boxed{L_{\text{PE}}^{\text{scalar}} = \frac{2\sqrt{2}\,\pi}{\sqrt{d}} \cdot \sqrt{\sum_{j=0}^{d/2-1} \frac{1}{\lambda_j^2}}}\]

\emph{Proof}: toeach dimensionto \((2j, 2j+1)\), normalizationencoding
\(\text{PE}_j(p) = \sqrt{2/d}\,(\sin(2\pi p/\lambda_j), \cos(2\pi p/\lambda_j))\): 

\[\|\text{PE}_j(p) - \text{PE}_j(q)\|^2 = \frac{2}{d}\left[\left(\sin\left(\frac{2\pi p}{\lambda_j}\right) - \sin\left(\frac{2\pi q}{\lambda_j}\right)\right)^2 + \left(\cos\left(\frac{2\pi p}{\lambda_j}\right) - \cos\left(\frac{2\pi q}{\lambda_j}\right)\right)^2\right]\]

by three angle constant equality
\((\sin a - \sin b)^2 + (\cos a - \cos b)^2 = 2 - 2\cos(a-b) = 4\sin^2((a-b)/2)\), again benefit use 
\(|\sin\theta| \leq |\theta|\): 

\[\|\text{PE}_j(p) - \text{PE}_j(q)\|^2 = \frac{2}{d} \cdot 4\sin^2\!\left(\frac{\pi(p-q)}{\lambda_j}\right) \leq \frac{2}{d} \cdot \left(\frac{2\pi|p-q|}{\lambda_j}\right)^2 = \frac{8\pi^2|p-q|^2}{d \cdot \lambda_j^2}\]

to has \(j\) seek and: 

\[\|\text{PE}_{\text{scalar}}(p) - \text{PE}_{\text{scalar}}(q)\|^2 \leq \frac{8\pi^2|p-q|^2}{d} \sum_{j=0}^{d/2-1} \frac{1}{\lambda_j^2}\]

open direction i.e. obtain describe Lipschitz often number. shouldupper boundis tight (\(p, q\) None finite approachestime
\(\sin\theta \approx \theta\), etc. number asymptoticholds). \(\square\)

\textbf{Corollary 1.2.3 (Lipschitz often number asymptoticlinesis)}: toseparate number Optimal spectrum
\(\lambda_j = 2\pi\xi \cdot \cot(\pi(2j+1)/2d)\) andnormalizationencoding: 

\[L_{\text{PE}}^{\text{scalar}} \sim \frac{1}{\xi} \cdot \sqrt{\frac{d}{6}} \quad (d \to \infty)\]

normalizationpost-Lipschitz often number growthFrom \(O(d)\) drop is 
\(O(\sqrt{d})\)------normalization factor \(\sqrt{2/d}\) offset eliminate a \(\sqrt{d}\)
because factor, makesencodingtoLocation change -ization sensitivedegree follow dimension\textbf{sublinear}growth. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.3 3D
rotationencoding: rotationmatrixparameter -ization }\label{d-ux65cbux8f6cux7f16ux7801ux65cbux8f6cux77e9ux9635ux53c2ux6570ux5316}

\subsubsection{1.3.1 encodingDefinition}\label{ux7f16ux7801ux5b9aux4e49-1}

\textbf{Definition 1.3.1 (3D rotationLocationencoding)}: to
\(\mathbf{p} = (x, y, z) \in \mathbb{R}^3\), dimension
\(d\) (\(d \geq 6\), even number): 

\[\boxed{\text{PE}_{\text{rot}}(\mathbf{p}) = \mathbf{R}(\mathbf{p}) \cdot \mathbf{e}_0}\]

where: - \(\mathbf{e}_0 \in \mathbb{R}^d\) isfixed participate consider singletoward quantity (example such as
\(\mathbf{e}_0 = (1, 0, \ldots, 0)^T\)) -
\(\mathbf{R}(\mathbf{p}) = \mathbf{R}_x(\alpha x) \cdot \mathbf{R}_y(\beta y) \cdot \mathbf{R}_z(\gamma z) \in SO(d)\)
- \(\alpha, \beta, \gamma > 0\)
is three spacedimension\textbf{frequencyparameter} (rad/Å)

each \(\mathbf{R}_a(\theta)\) is a at indicate determine 2D flat surface upper \(d\) dimensionalrotation: 

\[\mathbf{R}_a(\theta) = \begin{pmatrix} \mathbf{I}_{i_a-1} & & & \\ & \begin{pmatrix} \cos\theta & -\sin\theta \\ \sin\theta & \cos\theta \end{pmatrix} & & \\ & & & \mathbf{I}_{d-i_a-1} \end{pmatrix}\]

where three rotation flat surface separate different acts ondimensionto \((0,1)\), \((2,3)\),
\((4,5)\) (separate different toshould \(x, y, z\) axis). remaining remainder \(d-6\) dimensionby
\(\mathbf{R}_x, \mathbf{R}_y, \mathbf{R}_z\) not change ground preserves (singlematrix). 

\subsubsection{1.3.2
rotationmatrixparameter -ization ------ group theory Foundation}\label{ux65cbux8f6cux77e9ux9635ux7684ux53c2ux6570ux5316ux7fa4ux8bbaux57faux7840}

\textbf{Proposition 1.3.1 (\(\mathbb{R}^3 \to SO(d)\) embedding)}: map
\(\mathbf{p} \mapsto \mathbf{R}(\mathbf{p})\) is \(\mathbb{R}^3\) to
\(SO(d)\) \(3\) dimensional Abel factor group \textbf{ Lie group same state }. 

to exchange Conditions (Abel ness): \(\mathbf{R}_x, \mathbf{R}_y, \mathbf{R}_z\)
acts ondisjoint sit markto, becauseandis\textbf{ mutually thisswaps}: 

\[[\mathbf{R}_x(\theta), \mathbf{R}_y(\phi)] = 0, \quad [\mathbf{R}_y(\phi), \mathbf{R}_z(\psi)] = 0, \quad [\mathbf{R}_z(\psi), \mathbf{R}_x(\theta)] = 0\]

combineness: \(\mathbf{R}(\mathbf{p}_1 + \mathbf{p}_2) = \mathbf{R}(\mathbf{p}_1) \cdot \mathbf{R}(\mathbf{p}_2)\) (addmethod group same state), becauseis: 

\begin{align*}
\mathbf{R}(\mathbf{p}_1 + \mathbf{p}_2) &= \mathbf{R}_x(\alpha(x_1 + x_2)) \cdot \mathbf{R}_y(\beta(y_1 + y_2)) \cdot \mathbf{R}_z(\gamma(z_1 + z_2)) \\
&= \mathbf{R}_x(\alpha x_1)\mathbf{R}_x(\alpha x_2) \cdot \mathbf{R}_y(\beta y_1)\mathbf{R}_y(\beta y_2) \cdot \mathbf{R}_z(\gamma z_1)\mathbf{R}_z(\gamma z_2) \\
&= \mathbf{R}_x(\alpha x_1)\mathbf{R}_y(\beta y_1)\mathbf{R}_z(\gamma z_1) \cdot \mathbf{R}_x(\alpha x_2)\mathbf{R}_y(\beta y_2)\mathbf{R}_z(\gamma z_2) \\
&= \mathbf{R}(\mathbf{p}_1) \cdot \mathbf{R}(\mathbf{p}_2)
\end{align*}

 (th three step benefit use swapsness.)

\subsubsection{1.3.3
inner accumulate structure (relativeLocationencoding)}\label{ux5185ux79efux7ed3ux6784ux76f8ux5bf9ux4f4dux7f6eux7f16ux7801}

to any agree two Location \(\mathbf{p}, \mathbf{q} \in \mathbb{R}^3\): 

\[\boxed{\langle \text{PE}_{\text{rot}}(\mathbf{p}), \text{PE}_{\text{rot}}(\mathbf{q}) \rangle = \langle \mathbf{R}(\mathbf{p})\mathbf{e}_0, \mathbf{R}(\mathbf{q})\mathbf{e}_0 \rangle = \langle \mathbf{e}_0, \mathbf{R}(\mathbf{p})^T \mathbf{R}(\mathbf{q}) \mathbf{e}_0 \rangle = \langle \mathbf{e}_0, \mathbf{R}(\mathbf{q} - \mathbf{p}) \mathbf{e}_0 \rangle}\]

wheremostback a step benefit use 
\(\mathbf{R}(\mathbf{p})^T = \mathbf{R}(-\mathbf{p})\) and
\(\mathbf{R}(-\mathbf{p})\mathbf{R}(\mathbf{q}) = \mathbf{R}(\mathbf{q} - \mathbf{p})\). 

 exhibit open
\(\mathbf{R}(\Delta\mathbf{p}) = \mathbf{R}_x(\alpha\Delta x)\mathbf{R}_y(\beta\Delta y)\mathbf{R}_z(\gamma\Delta z)\), take
\(\mathbf{e}_0 = (1, 0, 0, 1, 0, 0, \ldots)^T\) (ateachtorotation flat surface th a dimensionalis 
1): 

\[\langle \text{PE}_{\text{rot}}(\mathbf{p}), \text{PE}_{\text{rot}}(\mathbf{q}) \rangle = \cos(\alpha\Delta x) + \cos(\beta\Delta y) + \cos(\gamma\Delta z)\]

where
\(\Delta\mathbf{p} = \mathbf{q} - \mathbf{p} = (\Delta x, \Delta y, \Delta z)\). 

encodinginner accumulate \textbf{only-dependenceatrelativeLocation}
\(\Delta\mathbf{p}\), satisfies\textbf{translation invariance}. 

\subsubsection{1.3.4 rotationencoding Lipschitz
continuousness}\label{ux65cbux8f6cux7f16ux7801ux7684-lipschitz-ux8fdeux7eedux6027}

\textbf{Theorem 1.3.1 (rotationencoding Lipschitz often number)}: map
\(\text{PE}_{\text{rot}}: \mathbb{R}^3 \to \mathbb{R}^d\) is Lipschitz
continuous. take \(\mathbf{e}_0\) is upper describe form (eachto flat surface th a dimensionalis 1, itsremainderis 
0), Exact often number is: 

\[\boxed{L_{\text{PE}}^{\text{rot}} = \max(\alpha, \beta, \gamma)}\]

\emph{Proof}: 

\textbf{Step 1 (directnormcompute)}: encoding difference is 
\(\text{PE}_{\text{rot}}(\mathbf{p}) - \text{PE}_{\text{rot}}(\mathbf{q}) = \mathbf{R}(\mathbf{p})\mathbf{e}_0 - \mathbf{R}(\mathbf{q})\mathbf{e}_0\). byat three rotationacts onmutualdisjointdimensionto
\((0,1), (2,3), (4,5)\), encoding difference each separate quantity decoupled. to \(x\)- flat surface (dimension
0,1): 

\[\|\mathbf{R}_x(\alpha x)\mathbf{e}_x - \mathbf{R}_x(\alpha q_x)\mathbf{e}_x\|^2 = (\cos(\alpha x)-\cos(\alpha q_x))^2 + (\sin(\alpha x)-\sin(\alpha q_x))^2 = 2 - 2\cos(\alpha\Delta x) = 4\sin^2\left(\frac{\alpha\Delta x}{2}\right)\]

same reason to \(y, z\) flat surface. becausethis: 

\[\|\text{PE}_{\text{rot}}(\mathbf{p}) - \text{PE}_{\text{rot}}(\mathbf{q})\|^2 = 4\sin^2\left(\frac{\alpha\Delta x}{2}\right) + 4\sin^2\left(\frac{\beta\Delta y}{2}\right) + 4\sin^2\left(\frac{\gamma\Delta z}{2}\right)\]

\textbf{Step 2 (benefit using \(|\sin\theta| \leq |\theta|\))}: 

\[\|\text{PE}_{\text{rot}}(\mathbf{p}) - \text{PE}_{\text{rot}}(\mathbf{q})\|^2 \leq \alpha^2\Delta x^2 + \beta^2\Delta y^2 + \gamma^2\Delta z^2 \leq \max(\alpha, \beta, \gamma)^2 \cdot \|\Delta\mathbf{p}\|_2^2\]

open direction i.e. obtain 
\(\|\text{PE}_{\text{rot}}(\mathbf{p}) - \text{PE}_{\text{rot}}(\mathbf{q})\| \leq \max(\alpha, \beta, \gamma) \cdot \|\mathbf{p} - \mathbf{q}\|_2\). 

\textbf{Step 3 (tight cause ness)}: take \(\Delta\mathbf{p} = (\varepsilon, 0, 0)\)
and \(\alpha = \max(\alpha, \beta, \gamma)\), when \(\varepsilon \to 0\)
time: 

\[\|\text{PE}_{\text{rot}}(\mathbf{p}) - \text{PE}_{\text{rot}}(\mathbf{q})\| = 2|\sin(\alpha\varepsilon/2)| \to \alpha|\varepsilon| = \max(\alpha, \beta, \gamma) \cdot \|\Delta\mathbf{p}\|\]

should world is tight. \(\square\)

\textbf{Note 1.3.1 (and first front version thismiddle often number high estimate Description)}: first front version thisuse
\(\|\mathbf{R}(\mathbf{p}) - \mathbf{R}(\mathbf{q})\|_F\)
 three angleinequalitydecompose knot combined Cauchy-Schwarz, obtain to
\(2\sqrt{\alpha^2 + \beta^2 + \gamma^2}\). this a often number ratio true value 
\(\max(\alpha, \beta, \gamma)\) high out convention \(2\sqrt{3} \approx 3.46\) times (when
\(\alpha = \beta = \gamma\) time), Reasonis: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{ three angleinequalityatcurrently cross factor spaceupper loose relax }: three difference separatematrix
 \(\mathbf{D}_x, \mathbf{D}_y, \mathbf{D}_z\)
 acts onmutualdisjointdimensiontoupper, becausethis
 \(\|\mathbf{D}_x + \mathbf{D}_y + \mathbf{D}_z\|_F^2 = \|\mathbf{D}_x\|_F^2 + \|\mathbf{D}_y\|_F^2 + \|\mathbf{D}_z\|_F^2\), but three angleinequality
 \(\|\mathbf{D}_x + \mathbf{D}_y + \mathbf{D}_z\|_F \leq \|\mathbf{D}_x\|_F + \|\mathbf{D}_y\|_F + \|\mathbf{D}_z\|_F\)
 in the worst casemay loose \(\sqrt{3}\) times (three term etc. normtime). 
\item
 \textbf{Cauchy-Schwarz \(\ell_1 \to \ell_2\)
 turn exchange }: \(\alpha|\Delta x| + \beta|\Delta y| + \gamma|\Delta z| \leq \sqrt{\alpha^2 + \beta^2 + \gamma^2} \cdot \|\Delta\mathbf{p}\|_2\)
 in \(\alpha=\beta=\gamma\) and \(\Delta x = \Delta y = \Delta z\)
 timeintroduce another a \(\sqrt{3}\) times loose relax. 
\end{enumerate}

twoheavy loose relax complex combinedleads to often number high estimate convention \(2\sqrt{3}\)
 times. correctProofshould benefit use factor spacecurrently cross nessdirectcompute \(\ell_2\)
norm, avoidthistwolayer loose relax. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.4 unified Lipschitz
continuousnessTheorem}\label{ux7edfux4e00-lipschitz-ux8fdeux7eedux6027ux5b9aux7406}

\textbf{Theorem 1.4.1 (PPE unified Lipschitz ness)}: setup
\(\text{PE}: \mathcal{P} \to \mathbb{R}^{d_{\text{pe}}}\)
is withupperDefinition any a encoding. thenexistenceat often number \(L_{\text{PE}} > 0\) makesto has 
\(\mathbf{p}, \mathbf{q} \in \mathcal{P}\): 

\[\|\text{PE}(\mathbf{p}) - \text{PE}(\mathbf{q})\| \leq L_{\text{PE}} \cdot d_{\mathcal{P}}(\mathbf{p}, \mathbf{q})\]

where \(d_{\mathcal{P}}\) is \(\mathcal{P}\) upperself-howeverdegree quantity: -
mark quantity domain: \(d_{\mathcal{P}}(p, q) = |p - q|\) - 3D
 domain: \(d_{\mathcal{P}}(\mathbf{p}, \mathbf{q}) = \|\mathbf{p} - \mathbf{q}\|_2\)

each domain Exact Lipschitz often number: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2381}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3571}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4048}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
encoding class type 
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\mathcal{P}\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(L_{\text{PE}}\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
currently sinusoidal (mark quantity) & \([0, L]\) &
\(\frac{2\sqrt{2}\,\pi}{\sqrt{d}}\sqrt{\sum_j 1/\lambda_j^2}\) \\
rotation (3D) & \(\mathbb{R}^3\) & \(\max(\alpha, \beta, \gamma)\) \\
\end{longtable}

physical meaning: \(L_{\text{PE}}\) controlsencodingtoLocation change -ization \textbf{sensitivedegree}. large
\(L_{\text{PE}}\)
 make neighbor near Locationencodingmoreaddis different (highseparate distinguish rate), butmaybreaking localsmooth smooth ness. Optimal 
\(L_{\text{PE}}\) shouldatphysicalmutualclosescale \(\xi\) attach near matchesencodingseparate distinguish rate. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{th 2 part: Theorem 1
rigorouscorrection}\label{ux7b2c-2-ux90e8ux5206theorem-1-ux7684ux4e25ux683cux4feeux6b63}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.1
correctionDetection Margin------Exact table reach formula }\label{ux4feeux6b63ux7684ux68c0ux6d4bux8fb9ux9645ux7cbeux786eux8868ux8fbeux5f0f}

\subsubsection{2.1.1 Setup}\label{ux8bbeux5b9a}

introduce PPE post-, expert \(m\) input enter From \(s_i\) change is: 

\[h_i = \phi(s_i) + \text{PE}(p_i)\]

where \(\phi: \mathbb{R}^{d_s} \to \mathbb{R}^d\)
isStatusCharacteristicsmap (maywithis constant etc. mapormaylearninglinearprojects). 

\textbf{Assumptions 2.1.1 (expertacts onencodingspace)}: expertmodel \(E_m\) inencodingspace
\(\mathbb{R}^d\) upper operation action. it s accept \(h_i\)
 action is input enter and input out prediction, voteDefinitionis: 

\[v_m^{\text{PPE}}(s_i) = \mathbf{1}[E_m(h_i) \neq y_i]\]

atdry clean /noiseConditionslowerexpectis: 

\[p_{\text{clean}, s}^{\text{PPE}} = \mathbb{E}[v_m^{\text{PPE}}(s) \mid \text{clean state } s]\]
\[p_{\text{noisy}, s}^{\text{PPE}} = \mathbb{E}[v_m^{\text{PPE}}(s) \mid \text{noisy state } s]\]

\subsubsection{2.1.2 repair Positive Statement}\label{ux4feeux6b63ux547dux9898}

\textbf{Proposition 2.1 (PPE lowerExact Detection Margin change -ization ------correction CC Audit Report in the 
sign error)}

PPE strengthensspace in the has effect Detection Marginis: 

\[\boxed{\Delta_s^{\text{PPE}} = \Delta_s + \delta_s^{\text{PE}}}\]

where: 

\[\boxed{\delta_s^{\text{PE}} = \underbrace{(p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s})}_{\text{noisesampleuppermany out separate disagree increase benefit }} - \underbrace{(p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s})}_{\text{dry clean sampleupper (not hope hope has)separate disagree change -ization }}}\]

\begin{quote}
\textbf{correctionDescription}: CC Audit Reportmiddle \(\delta_s^{\text{PE}}\)
 table reach formula Symbol/Conceptmutual counter. correctSymbol/Conceptis: \(\delta_s^{\text{PE}} > 0\) if and only if
PPE help help distinguishnoiseanddry clean sample (increaselargeDetection Margin). thisrequire: -
noisesampleupperseparate disagree probability\textbf{increases} (\(p_{\text{noisy}}^{\text{PPE}} > p_{\text{noisy}}\))------expertmoregood ground ''identifies'' out incorrectLabel
-
dry clean sampleupperseparate disagree probability\textbf{notincreasesordecreases} (\(p_{\text{clean}}^{\text{PPE}} \leq p_{\text{clean}}\))------experttofewmaintains original has consistentness
\end{quote}

\(\delta_s^{\text{PE}}\) value domain: byatprobabilityin \([0,1]\) middle, has 
\(\delta_s^{\text{PE}} \in [-\Delta_s, 1 - p_{\text{clean}, s}]\). 

\emph{Proof}: byDefinition: 

\begin{align*}
\Delta_s^{\text{PPE}} &= p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{clean}, s}^{\text{PPE}} \\
&= (p_{\text{noisy}, s} + (p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s})) - (p_{\text{clean}, s} + (p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s})) \\
&= (p_{\text{noisy}, s} - p_{\text{clean}, s}) + (p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s}) - (p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}) \\
&= \Delta_s + \delta_s^{\text{PE}}
\end{align*}

 value domain: minimum -ization \(\delta_s^{\text{PE}}\) ------
\(p_{\text{noisy}}^{\text{PPE}} = 0\) (has noisesampleby error judge is dry clean), \(p_{\text{clean}}^{\text{PPE}} = 1\) (has dry clean sampleby error judge is noise)\to
\(\delta_s^{\text{PE}} = (0 - p_{\text{noisy}}) - (1 - p_{\text{clean}}) = -p_{\text{noisy}} - 1 + p_{\text{clean}} = -(1 + \Delta_s)\). butthisexceeds
\(p_{\text{clean}}^{\text{PPE}}\) and \(p_{\text{noisy}}^{\text{PPE}}\)
reasonable paradigm surround \ldots in practice, athas meaningexpertmodelmiddle, \(p_{\text{noisy}}^{\text{PPE}} \in [0, 1]\)
and \(p_{\text{clean}}^{\text{PPE}} \in [0, 1]\)
 convention bind lower, \(\delta_s^{\text{PE}} \in [-\Delta_s, 1]\). more refined fine world by
\(p_{\text{noisy}}^{\text{PPE}} \in [0,1]\) and
\(p_{\text{clean}}^{\text{PPE}} \in [0,1]\) to out. \(\square\)

\textbf{correctionpost-Theorem 1 (PPE strengthens)}: 

\[\boxed{F_1^{\text{PPE}} \geq 1 - \frac{1}{\eta} \sum_{s \in \mathcal{S}} \rho_s \cdot \exp\left(-2M (\Delta_s + \delta_s^{\text{PE}})^2\right)}\]

Chernoff-Hoeffding structuremaintains (indicate show variablein \([0,1]\) middlehas world), onlyis often number 
\(\Delta_s\) change is \(\Delta_s + \delta_s^{\text{PE}}\). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{2.2 \(\delta_s^{\text{PE}} > 0\)
necessaryConditionsandSufficient Condition}{2.2 \textbackslash delta\_s\^{}\{\textbackslash text\{PE\}\} \textgreater{} 0 necessaryConditionsandSufficient Condition}}\label{delta_stextpe-0-ux7684ux5fc5ux8981ux6761ux4ef6ux548cux5145ux5206ux6761ux4ef6}

\subsubsection{2.2.1 information theoryConditions}\label{ux4fe1ux606fux8bbaux6761ux4ef6}

\textbf{Theorem 2.2.1 (\(\delta_s^{\text{PE}} > 0\)
Sufficient Condition------information theoryform, Bayesian leaf -ese Optimal finite determine)}

setup \(S, P, Y\)
separate different representationStatusatom、physicalLocationandLabelrandom variable (atData Distributionlower). If: 

\[\boxed{I(Y; P \mid S) > 0}\]

i.e.: physicalLocationat\textbf{todetermineStatusatomConditionslower} carry carry regardingLabel amount outerinformation, then\textbf{existenceat} a 
PPE function \(\text{PE}: \mathcal{P} \to \mathbb{R}^{d_{\text{pe}}}\)
andencodingdimension \(d_{\text{pe}}\), makestosome Status
\(s\), \textbf{Bayes-optimal classifier}atincrease broad Characteristics \((s, \text{PE}(p))\)
lowerDetection Margin change -ization satisfies \(\delta_s^{\text{PE}} > 0\). 

\emph{Proof}: 

\textbf{Step 1 (information store existencephysicalmay and ness)}: byAssumptions
\(I(Y; P \mid S) > 0\), existenceat a todistribution \((P_{Y|S,P}, P_{Y|S})\)
isnotmutualsame: 

\[I(Y; P \mid S) = \mathbb{E}_{S, P}\left[D_{\text{KL}}(P_{Y|S,P} \| P_{Y|S})\right] > 0\]

becausethisexistenceat a currently measure degree matter file set 
\(\mathcal{A} \subset \mathcal{S} \times \mathcal{P}\), atitsupper
\(P_{Y|S=s, P=p} \neq P_{Y|S=s}\). 

\textbf{Step 2 (Constructionencoding)}: select take a sufficient table reach forceencodingfunction
\(\text{PE}\). by through use force near Theorem (example such as has has sufficientmany spirit via dollar neural network), existenceat
\(\text{PE}\) makesencodingrepresentation \(\text{PE}(p)\) approximatepreserves \(P\) middleregarding
\(Y\) has mutualcloseinformation. 

specifically, take \(d_{\text{pe}}\) sufficientlarge, makesexistencein \(\text{PE}\)
satisfiesdataat reason inequality\textbf{withequalityholds}: 

\[I(Y; \text{PE}(P) \mid S) = I(Y; P \mid S) - \varepsilon_{\text{PE}}\]

where \(\varepsilon_{\text{PE}} \geq 0\) isencodingresults ininformation discard lose (seeth 3
part). forsufficient table reach \(\text{PE}\), mayguarantee
\(\varepsilon_{\text{PE}} < I(Y; P \mid S)\), becausethis: 

\[I(Y; \text{PE}(P) \mid S) > 0\]

\textbf{Step 3 (Bayes-optimal classifierlinesis)}: Conditionsmutualinformation
\(I(Y; \text{PE}(P) \mid S) > 0\) agree taste -ing todetermine \(S\), \(\text{PE}(P)\)
provideregarding \(Y\) amount outer''signal''. considerBayes-optimal classifieratincrease broad space in the table present. 

For noisy samples (Label \(y\) isincorrect): todetermine
\((s, p)\), Bayesian leaf -ese Optimal prediction device willmore incline towardat input out \textbf{and \(y\)
is different}trueLabel (becauseis \(p\) provide amount outer prove according to). becausethisseparate disagree probabilityrises: 

\[p_{\text{noisy}, s}^{\text{PPE}} > p_{\text{noisy}, s}\]

For clean samples: amount outerLocationinformation\textbf{ advance a step confirm}correctLabel, bydataat reason inequalityguaranteeexpertseparate disagree probability\textbf{notincreases} (very todecreases): 

\[p_{\text{clean}, s}^{\text{PPE}} \leq p_{\text{clean}, s}\]

becausethis\textbf{Bayes-optimal classifier}
\(\delta_s^{\text{PE}} = (p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s}) - (p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}) > 0\). 

\textbf{Step
4 (From Bayesian leaf -ese Optimal to actual actual expert------Conjecture)}: \textbf{{[} heuristic estimate {]}}
upper describe ConclusiontoBayes-optimal classifierrigorousholds. for SCX framework in the actual actual neural networkexpert
\(E_m\), mustguarantee \(\delta_s^{\text{PE}} > 0\)
must amount outerAssumptions: expertmodel input out consistent ground force near Bayes-optimal classifier (in \(\ell_1\)
meaninglowertoConditionsdistribution \(P_{Y|S,P}\) force near erroris less than
\(|\delta_s^{\text{PE}}|\)). forthroughrandomgradient descenttrainingneural network, goal front\textbf{nothas }has finite samplelowerrigorousconvergenceguarantee. thisis whenfrontTheorem\textbf{primarylimitation}: Step
1-3
provesBayes-optimal classifiermaywithFrom Locationinformationmiddle receive benefit (information-theoretically must howeverness), butFrom Bayesian leaf -ese Optimal to actual actual expert push broad is heuristic estimate Conjecture. 

\textbf{ retreat a step see }: If Step 4 push broad does not hold, Theorem 2.2.1
 retreat -ization is a classicalConclusion: \(I(Y;P|S) > 0\) agree taste -ing Bayes-optimal classifierat make using \(P\)
informationtime table present moregood------thisisdataat reason inequalitydirectCorollary, is flat all. \(\square\)

\subsubsection{2.2.2 necessaryConditions}\label{ux5fc5ux8981ux6761ux4ef6}

\textbf{Proposition 2.2.1 (\(\delta_s^{\text{PE}} > 0\) necessaryConditions)}: 

If to has expert \(m\), \(\delta_s^{\text{PE}} > 0\), thenmusthas: 

\[\boxed{I(Y; P \mid S) > 0 \quad \text{or} \quad \text{PE changesexpertsystembias difference -variancestructure (non-information theorymechanism)}}\]

th a kindcase (information theorymechanism)alreadyin 2.2.1
middleat reason. th two kindcase (non-information theorymechanism)may emit life at: PE
increasesencodingdimension\textbf{randomcurrentlythen -ization effect}, make expert input out variancestructureundergoes change. butthisis NOT ''physicalLocationinformation''at rise action use ------andisarbitraryaddnessnoisealsocanproduceseffect. 

\textbf{physicallyimportant situation shape is \(I(Y; P \mid S) > 0\)}. thisis equivalent to: 

\[\exists s, p_1, p_2 \text{ makes } P_{Y|S=s, P=p_1} \neq P_{Y|S=s, P=p_2}\]

i.e.: for\textbf{same a Statusatom} \(s\), is differentphysicalLocation \((p_1, p_2)\)
willleads tois differentLabeldistribution. 

\textbf{physicalexample}: - whitemiddle: mutualsameamino acid class type (Status \(s\))at
\(\alpha\)-helix (Location \(p_1\))and \(\beta\)-fold (Location
\(p_2\)) in the function canRoleis different ← \(I(Y; P \mid S) > 0\) -
crystalmiddle: mutualsamechemical environment environment (Status \(s\))atsurface (Location \(p_1\))and overall mutual (Location
\(p_2\)) in the defectformation energyis different ← \(I(Y; P \mid S) > 0\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{2.3 \(\delta_s^{\text{PE}}\)
information theoryupper bound (dataat reason inequality)}{2.3 \textbackslash delta\_s\^{}\{\textbackslash text\{PE\}\} information theoryupper bound (dataat reason inequality)}}\label{delta_stextpe-ux7684ux4fe1ux606fux8bbaux4e0aux754cux6570ux636eux5904ux7406ux4e0dux7b49ux5f0f}

\textbf{Theorem 2.3.1 (\(\delta_s^{\text{PE}}\) information theoryupper bound)}: 

\[\boxed{\delta_s^{\text{PE}} \leq \min\left(1, \sqrt{\frac{1}{2} \cdot I(P; \text{PE}(P) \mid S, Y)}\right) + \text{Bernstein correction term}}\]

moreExact ground, bydataat reason inequality (DPI)and Pinsker inequality: 

\[\delta_s^{\text{PE}} \leq \min\left(1 - p_{\text{clean}, s}, \sqrt{\frac{1}{2} D_{\text{KL}}(P_{\text{PE}|S,Y} \| P_{\text{PE}|S})}\right) + \delta_s^{\text{variance}}\]

where \(\delta_s^{\text{variance}}\) fromhas finite expert number \(M\)
introduceestimatevariance. 

\emph{Proof}: 

\textbf{Step 1 (probability change -ization and KL scatter degree)}: For clean samplesupperseparate disagree probability change -ization: 

\[|p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}| = \left|\mathbb{E}_{\text{PE}(P)|S=s, Y=y_{\text{true}}}[P(E_m(h) \neq y_{\text{true}})] - P(E_m(s) \neq y_{\text{true}})\right|\]

by Pinsker inequality: to any agree matter file 
\(A\), \(|P(A) - Q(A)| \leq \sqrt{\frac{1}{2} D_{\text{KL}}(P \| Q)}\). 

applicationto \(p_{\text{clean}}\) and \(p_{\text{noisy}}\) change -ization, separate different to out: 

\begin{align*}
|p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}| &\leq \sqrt{\frac{1}{2} D_{\text{KL}}(P_{\text{PE}(P)|S=s, \text{clean}} \| P_{\text{PE}(P)|S=s})} \\
|p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s}| &\leq \sqrt{\frac{1}{2} D_{\text{KL}}(P_{\text{PE}(P)|S=s, \text{noisy}} \| P_{\text{PE}(P)|S=s})}
\end{align*}

\textbf{Step 2 (combine)}: 

\begin{align*}
\delta_s^{\text{PE}} &= (p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s}) - (p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}) \\
&\leq |p_{\text{noisy}, s}^{\text{PPE}} - p_{\text{noisy}, s}| + |p_{\text{clean}, s}^{\text{PPE}} - p_{\text{clean}, s}| \\
&\leq \sqrt{\frac{1}{2} D_{\text{KL}}^{(1)}} + \sqrt{\frac{1}{2} D_{\text{KL}}^{(2)}}
\end{align*}

where \(D_{\text{KL}}^{(1)}, D_{\text{KL}}^{(2)}\)
separate different isdry clean andnoiseConditionslower KL scatter degree. 

\textbf{Step 2b (value domain upper bound come source ------supplementDescription)}: by
\(\delta_s^{\text{PE}}\) value domain (Proposition 2.1), constant has 
\(\delta_s^{\text{PE}} \leq 1 - p_{\text{clean}, s}\) (take
\(p_{\text{noisy}}^{\text{PPE}} = 1\),
\(p_{\text{clean}}^{\text{PPE}} = 0\) timemaximum value). thisupper boundfromprobability
\([0,1]\) convention bind, not-dependenceat any what distributionAssumptions. Theorem Statement in the 
\(\min(1-p_{\text{clean},s}, \sqrt{\frac{1}{2}D_{\text{KL}}})\)
takeself-\textbf{two independenthas effect upper boundminimum value }: value domain upper boundand Pinsker upper bound. when
Pinsker world relatively loose (KL scatter degreelargebut \(\delta_s^{\text{PE}}\)
stillis subject to value domain)time, value domain world more tight; counter 'swhenLocationinformationveryfew (KL scatter degreesmall)time, Pinsker
 world more tight. 

\textbf{Step 3 (DPI convention bind)}: bydataat reason inequality (encoding PE onlycandecreasesinformation): 

\[I(\text{PE}(P); Y \mid S) \leq I(P; Y \mid S)\]

becausethis \(\delta_s^{\text{PE}}\) is subject toLocation \(P\) intodetermine \(S\)
lower carry carry mutualinformation quantity. when \(I(P; Y \mid S) = 0\) (LocationNone use time), upper bound retreat -ization is: 

\[\delta_s^{\text{PE}} \leq 0 + \delta_s^{\text{variance}}\]

ata large amount ofexpertextremely finite \(M \to \infty\) lower
\(\delta_s^{\text{variance}} \to 0\). \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{2.4 \(\delta_s^{\text{PE}}\)
 heuristic estimate estimate (Fano
 world finite ------non-rigorouslower bound)}{2.4 \textbackslash delta\_s\^{}\{\textbackslash text\{PE\}\} heuristic estimate estimate (Fano world finite ------non-rigorouslower bound)}}\label{delta_stextpe-ux7684ux542fux53d1ux5f0fux4f30ux8ba1fano-ux754cux9650ux975eux4e25ux683cux4e0bux754c}

\textbf{Proposition 2.4.1 (\(\delta_s^{\text{PE}}\)
 heuristic estimate estimate------non-rigorouslower bound)}: 

\begin{quote}
\textbf{{[} heuristic estimate {]}} \textbf{ (warning: thisnon-rigorousTheorem.)}
\end{quote}

setupLocation \(P\) intodetermineStatus \(S\) lowerprovideregardingLabel \(Y\) amount outerinformation, andencoding PE
preservesits in the 
\(\rho = I(Y; \text{PE}(P) \mid S) / I(Y; P \mid S) \in (0, 1]\)
part. If one assumes the Fano
inequalityattwo world world uniform reach to etc. number (must special determinedistributionConditions: noiseuniform uniform distributionatincorrectCategoryupper), thenexistenceatStatus
\(s\) makes: 

\[\delta_s^{\text{PE}} \approx \frac{2\rho \cdot I(Y; P \mid S) - \log 2}{\log |\mathcal{Y}|} - O\left(\frac{1}{\sqrt{M}}\right)\]

where \(M\) isexpert number, \(O(1/\sqrt{M})\) fromhas finite expertestimatenoise. 

\textbf{key finite determine: This expression \emph{is NOT } a rigorous mathematical lower bound. } From two Fano
lower bound \(P_e \geq A\) and \(P_e^{\text{PPE}} \geq B\) cannot logic edit ground push out 
\(P_e - P_e^{\text{PPE}} \geq A - B\), becauseis \(a \geq A\) and \(b \geq B\)
not contain contain \(a-b \geq A-B\). thisouter, Fano
inequalityonlyat special determinedistributionupper reach to etc. number (noiseuniform uniform distributionatincorrectCategoryupper, and \(H_2(P_e)\)
 full and). becausethis, thisatto out table reach formula shouldby view is to \(\delta_s^{\text{PE}}\)
order of magnitude\textbf{ heuristic estimate estimate}, andnon-Theorem level lower boundConclusion. correctlower boundrequireis differentProofstrategy, thisiswhenfront a open problems. 

\emph{Derivation think path (non-rigorousProof)}: 

\textbf{Step 1 (Fano inequality------From errortoinformation)}: Fano
inequalityStandard form: 

\[H(Y \mid S, \text{PE}(P)) \leq H_2(P_e) + P_e \cdot \log(|\mathcal{Y}| - 1)\]

where \(P_e\) istodetermine \((S, \text{PE}(P))\) lower Bayesian leaf -ese incorrect rate, \(H_2\)
is two dollar entropy. etc. effect ground, lower bound: 

\[P_e \geq \frac{H(Y \mid S, \text{PE}(P)) - 1}{\log |\mathcal{Y}|}\]

\textbf{Step 2 (heuristic estimate close connect ------not etc. number error use)}: If Fano
inequalityattwo world world uniform reach to etc. number (thisin general\textbf{does not hold}), thenformuppermay write: 

\begin{align*}
P_e - P_e^{\text{PPE}} &\stackrel{\text{If etc. number }}{=} \frac{H(Y|S) - H(Y|S, \text{PE}(P))}{\log |\mathcal{Y}|} \\
&= \frac{I(Y; \text{PE}(P) \mid S)}{\log |\mathcal{Y}|} \\
&= \frac{\rho \cdot I(Y; P \mid S)}{\log |\mathcal{Y}|}
\end{align*}

againaddupper \(\Delta P_e\) and \(\delta_s^{\text{PE}}\) heuristic estimate close connect (Assumptions
\(\Delta P_e \approx -\pi \cdot \delta_s^{\text{PE}}\) tomaximum obtain benefit Status
\(s\) holds), take
\(\pi = 1/2\) (mostbadcase), \(\rho = 1\) (complete beautiful encoding), addupperhas finite \(M\) 
Hoeffding correctioni.e. obtain describe table reach formula. 

\textbf{again time strong adjust: thisDerivationFrom not etc. number push out etc. number, is NOT rigorousProof. correctlower boundrequireusemorestrongtool (such as
Fano inverse formorstrong Fano inequality), thisrequireregardingdistributionconcreteAssumptions. }

\textbf{Physical meaning (order of magnitudeestimate)}: If \(I(Y; P \mid S)\)
verysmall (Location several -ish does not provide amount outerinformation), then \(\delta_s^{\text{PE}}\)
 heuristic estimate estimateapproaches zero oris negative ------PPE
Nonemethod actual ness ground improvementdetection. thisexplainsis what what \textbf{is NOT has physical systemallcanFrom PPE
middle receive benefit }. butthis a Conclusionisdirectioncorrectphysical study direct feel, andnon-rigorousmathematicalguarantee. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.5 spacelocalnesscurrentlythen -ization ------CC
Audit ReportresultCorrected Version}\label{ux7a7aux95f4ux5c40ux90e8ux6027ux6b63ux5219ux5316cc-ux5ba1ux8ba1ux62a5ux544aux7ed3ux679cux7684ux4feeux6b63ux7248}

\textbf{Theorem 2.5.1 (\(\delta_s^{\text{PE}}\) spacesmooth smooth ness)}: setup PE satisfies
Lipschitz Conditions (often number is \(L_{\text{PE}}\)), andexpert \(E_m\)
 input out \textbf{softprobability} \(E_m^{\text{soft}}: \mathbb{R}^d \to [0,1]\) (such as
softmax input out maximumCategoryprobability)satisfies \(L_E\)-Lipschitz
Conditions. thisouter, Assumptions\textbf{boundaryConditions}: expert \(E_m\) verdict strategy boundary distance from two encoding point 
\(h_i, h_j\) tofewis \(\tau > 0\), i.e.
\(|E_m^{\text{soft}}(h) - 0.5| \geq \tau\) to \(h \in \{h_i, h_j\}\)
holds. thentosame belong a Status \(s\)、 has has physicalLocation \(p_i, p_j\) two atom: 

\[\boxed{|\delta_i^{\text{PE}} - \delta_j^{\text{PE}}| \leq 2 \cdot L_E \cdot L_{\text{PE}} \cdot d_{\mathcal{P}}(p_i, p_j) + O\left(\frac{1}{\sqrt{M}}\right)}\]

where \(O(1/\sqrt{M})\)
fromhas finite expertestimateerror. \textbf{If boundaryConditionsdoes not hold}, should world does not guaranteenon- empty ------ indicate show function
\(\mathbf{1}[E_m(h) \neq y]\) in verdict strategy boundaryatnotcontinuous, softprobability Lipschitz
nesscannot convention bind indicate show functionexpect change -ization. 

\emph{Proof}: 

\begin{align*}
|\delta_i^{\text{PE}} - \delta_j^{\text{PE}}| &= |(p_{\text{noisy}, i}^{\text{PPE}} - p_{\text{noisy}, j}^{\text{PPE}}) - (p_{\text{clean}, i}^{\text{PPE}} - p_{\text{clean}, j}^{\text{PPE}})| \\
&\leq |p_{\text{noisy}, i}^{\text{PPE}} - p_{\text{noisy}, j}^{\text{PPE}}| + |p_{\text{clean}, i}^{\text{PPE}} - p_{\text{clean}, j}^{\text{PPE}}|
\end{align*}

each probability difference
 receive input enter change -ization convention bind. \textbf{key}: \(p_{\text{clean}, i}^{\text{PPE}} = \mathbb{E}[\mathbf{1}[E_m(h_i) \neq y]]\)
is indicate show functionexpect. boundaryConditions \(|E_m^{\text{soft}}(h) - 0.5| \geq \tau\)
guarantee \(h_i\) and \(h_j\)
uniformat verdict strategy boundarysame a side and maintains a determine distance from, thistimesoftprobability \(L_E\)-Lipschitz
nesscontrols indicate show functionexpect change -ization: 

\[|p_{\text{clean}, i}^{\text{PPE}} - p_{\text{clean}, j}^{\text{PPE}}| \leq L_E \cdot \|h_i - h_j\| = L_E \cdot \|\text{PE}(p_i) - \text{PE}(p_j)\| \leq L_E \cdot L_{\text{PE}} \cdot d_{\mathcal{P}}(p_i, p_j)\]

 (benefit using \(s_i = s_j = s\) time \(\phi(s_i) = \phi(s_j)\).)two times i.e. obtain. 

\textbf{If NoneboundaryConditions}: verdict strategy boundarymay exactly goodis located at \(h_i\) and \(h_j\)
'sbetween, thistimei.e. make \(\|h_i - h_j\|\)
tends to zero, separate disagree probabilitymaymutual difference any agree large (such as a complete fullatdry clean side 
\(p \approx 0\), another a atnoise side 
\(p \approx 1\)), world change is empty (vacuous). \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{th 3 part: Theorem 2
Exact correction}\label{ux7b2c-3-ux90e8ux5206theorem-2-ux7684ux7cbeux786eux4feeux6b63}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{3.1 encodingnot complete beautiful degree \(\varepsilon_{\text{PE}}\)
Exact information theoryDefinition}{3.1 encodingnot complete beautiful degree \textbackslash varepsilon\_\{\textbackslash text\{PE\}\} Exact information theoryDefinition}}\label{ux7f16ux7801ux4e0dux5b8cux7f8eux5ea6-varepsilon_textpe-ux7684ux7cbeux786eux4fe1ux606fux8bbaux5b9aux4e49}

\subsubsection{3.1.1
Conditionsmutualinformation chain formula decompose}\label{ux6761ux4ef6ux4e92ux4fe1ux606fux7684ux94feux5f0fux5206ux89e3}

setuprandom variable: - \(X\): original begin Characteristics (Statusatom \(S\) oritsrepresentation \(\phi(S)\)) -
\(P\): physicalLocation (Completeinformation) - \(\text{PE}(P)\): encodingbackphysicalLocation -
\(Y\): Label

information flow Markov chain: 

\[Y \leftrightarrow (X, P) \leftrightarrow (X, \text{PE}(P))\]

 (i.e.
\(Y \perp\!\!\!\perp \text{PE}(P) \mid (X, P)\)------encodingonly discard lose information, not create create information.)

\subsubsection{3.1.2
encodingnot complete beautiful degreeDefinition}\label{ux7f16ux7801ux4e0dux5b8cux7f8eux5ea6ux7684ux5b9aux4e49}

\textbf{Definition 3.1 (encodingnot complete beautiful degree
\(\varepsilon_{\text{PE}}\)------information theoryDefinition)}: 

\[\boxed{\varepsilon_{\text{PE}} = I(Y; P \mid X) - I(Y; \text{PE}(P) \mid X)}\]

\textbf{ness}: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{dataat reason inequality} (DPI): \(\varepsilon_{\text{PE}} \geq 0\)
 constant holds. becauseis \((X, \text{PE}(P))\) is \((X, P)\)
 (mayisrandom)function, DPI guarantee: 

 \[I(Y; \text{PE}(P) \mid X) \leq I(Y; P \mid X)\]
\item
 \textbf{may inverse nessConditions}: \(\varepsilon_{\text{PE}} = 0\) if and only if PE
 intodetermine \(X\) Conditionslowerto \(Y\) predictionis\textbf{sufficient statistic}, i.e.: 

 \[Y \perp\!\!\!\perp P \mid (X, \text{PE}(P))\]

 equivalent ground: encodingbackLocation \(\text{PE}(P)\) preserves \(P\) middle has and \(Y\)
 mutualcloseinformation. 
\item
 \textbf{ flat all case}: If 
 \(I(Y; P \mid X) = 0\) (Locationnot carry carry Labelmutualcloseinformation), then
 \(\varepsilon_{\text{PE}} = 0\) self- action holds (zero subtract zero). 
\item
 \textbf{upper bound}: \(\varepsilon_{\text{PE}} \leq I(Y; P \mid X) \leq H(Y)\). maximumnot complete beautiful degreeis discard lose has Locationinformation. 
\end{enumerate}

\subsubsection{3.1.3
andencodingdimensionrelationship}\label{ux4e0eux7f16ux7801ux7ef4ux5ea6ux7684ux5173ux7cfb}

\textbf{Proposition 3.1 (dimension-information weight balance)}: tosinusoidal encoding (\(d\) dimensional)and Laplace
mutualclosekernel (mutualcloselongdegree \(\xi\)): 

\[\varepsilon_{\text{PE}} \leq I(Y; P \mid X) \cdot \exp\left(-\frac{d}{d_0}\right)\]

where \(d_0 = O(L/\xi)\) isCharacteristicsdimensionscale. specifically, when \(d\) following §1.2
 in the separate number Optimal schemegrowthtime, information discard lose withexponentialratefalls. 

\emph{Proof Idea}: encoding force near error (Theorem
1.2.2)leads tokernelheavy build error, advance andleads toConditionsdistribution \(P_{Y|X,\text{PE}(P)}\) and
\(P_{Y|X,P}\) KL scatter degree, by Pinsker
inequality turn -ization is mutualinformation discard lose. exponential decayfrom Fourier level number spectrum decay subtract (Cauchy
kernel solve analyze ness). \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3.2 correction Theorem
2------Exact stated describe andProof}\label{ux4feeux6b63ux7684-theorem-2ux7cbeux786eux9648ux8ff0ux548cux8bc1ux660e}

\textbf{Theorem 2' (not complete beautiful PPE lowerweakCharacteristics lose effect upper bound------Exact correction)}

setup original begin Characteristics \(X\) informationinsufficientdegreeis \(\delta > 0\) (Definitionsee CC Audit Report), and
\(\varepsilon_{\text{PE}}\) is PPE encodingnot complete beautiful degree. thenin PPE
strengthensCharacteristicsspacemiddle, SCX \(F_1\) separate number satisfies: 

\[\boxed{F_{1,\text{SCX}}^{\text{PPE}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta + \frac{2\varepsilon_{\text{PE}}}{C_F^2}}{2}}}\]

where \(C_F > 0\) is \(F_1\) and Bayesian leaf -ese incorrect rate 'sbetween turn exchange often number. 

\textbf{ special special situation shape }: -
\(\varepsilon_{\text{PE}} \to 0\) (complete beautiful encoding): upper bound restore complex is \textbf{ original begin 
Theorem 2}: 
\[F_{1,\text{SCX}}^{\text{PPE}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta}{2}}\]

\begin{itemize}
\tightlist
\item
 \(\varepsilon_{\text{PE}} \to I(Y; P \mid X)\) (complete fullfailencoding): upper bound time tonotuse any what LocationinformationStatus (pure wave cost parameter). 
\end{itemize}

\emph{Proof}: 

\textbf{Step 1 (Fano inequalityembedding)}: by Fano
inequality, Optimal separate class device (Bayesian leaf -ese)incorrect rate satisfies: 

\[P_e \geq \frac{H(Y \mid X, \text{PE}(P)) - 1}{\log |\mathcal{Y}|}\]

\textbf{Step 2 (Conditionsentropydecompose)}: 

\begin{align*}
H(Y \mid X, \text{PE}(P)) &= H(Y \mid X) - I(Y; \text{PE}(P) \mid X) \\
&= H(Y \mid X) - [I(Y; P \mid X) - \varepsilon_{\text{PE}}] \\
&= H(Y \mid X) - I(Y; P \mid X) + \varepsilon_{\text{PE}}
\end{align*}

\textbf{Step 3 (connection \(\delta\))}: original begin Theorem 2 middle, \(\delta\)
andConditionsentropyrelationshipis (CC Audit Report §1, Theorem 2): 

\[\delta = \frac{2}{\log|\mathcal{Y}|} \cdot \left(H(Y \mid X) - 1\right)_{\text{excess}}\]

where \((\cdot)_{\text{excess}}\) representation super out Bayesian leaf -ese incorrectinformation quantity. in PPE lower: 

\begin{align*}
H(Y \mid X, \text{PE}(P)) - 1 &= (H(Y \mid X) - I(Y; P \mid X) + \varepsilon_{\text{PE}}) - 1 \\
&= (H(Y \mid X) - 1) - I(Y; P \mid X) + \varepsilon_{\text{PE}}
\end{align*}

Definition \(\delta_{\text{PPE}}\) is PPE strengthensbackhas effect informationinsufficientdegree: 

\[\delta_{\text{PPE}} = \delta + \frac{2\varepsilon_{\text{PE}}}{\log|\mathcal{Y}|} \cdot \frac{\log|\mathcal{Y}|}{C_F^2} = \delta + \frac{2\varepsilon_{\text{PE}}}{C_F^2}\]

 (\(C_F\) fromincorrect rate and \(F_1\)
 turn exchange: \(F_1 \leq F_{1,\text{base}} + C_F \sqrt{P_e - P_e^*}\).)

\textbf{Step 4 (generation enter SCX \(F_1\) upper bound)}: 

\[F_{1,\text{SCX}}^{\text{PPE}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta_{\text{PPE}}}{2}} = F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta + \frac{2\varepsilon_{\text{PE}}}{C_F^2}}{2}}\]

\textbf{Step 5 (extremely finite verify)}: when \(\varepsilon_{\text{PE}} \to 0\): 

\[F_{1,\text{SCX}}^{\text{PPE}} \leq F_{1,\text{base}} + C_F \cdot \sqrt{\frac{\delta}{2}}\]

currentlyis original begin Theorem 2. \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{3.3 \(\varepsilon_{\text{PE}}\)
mayestimateform}{3.3 \textbackslash varepsilon\_\{\textbackslash text\{PE\}\} mayestimateform}}\label{varepsilon_textpe-ux7684ux53efux4f30ux8ba1ux5f62ux5f0f}

Theorem quantity 
\(\varepsilon_{\text{PE}} = I(Y; P \mid X) - I(Y; \text{PE}(P) \mid X)\)
istheory quantity ------ it involvingtrueData Distribution. for actual actual application, we s requireFrom datamiddleestimate it. 

\subsubsection{3.3.1 KSG estimate device }\label{ksg-ux4f30ux8ba1ux5668}

\textbf{Theorem 3.3.1 (\(\varepsilon_{\text{PE}}\)
mayestimateness)}: \(\varepsilon_{\text{PE}}\)
maywiththroughwithlowerMethodologyFrom has finite samplemiddle\textbf{consistent ground }estimate: 

\[\boxed{\hat{\varepsilon}_{\text{PE}}^{(n)} = \hat{I}^{(n)}(Y; P \mid X) - \hat{I}^{(n)}(Y; \text{PE}(P) \mid X)}\]

where \(\hat{I}^{(n)}(\cdot; \cdot \mid \cdot)\) isuse
Kraskov-Stögbauer-Grassberger (KSG)
estimate device \textbf{Conditionsmutualinformation}estimate. whensample quantity \(n \to \infty\)
time, \(\hat{\varepsilon}_{\text{PE}}^{(n)} \xrightarrow{p} \varepsilon_{\text{PE}}\) (consistentness). 

specifically, KSG Conditionsmutualinformationestimate device: 

\[\hat{I}(Y; P \mid X) = \hat{I}(Y; (P, X)) - \hat{I}(Y; X)\]

where \(\hat{I}(A; B)\) isStandard KSG mutualinformationestimate (based on \(k\)- near neighbor distance from). 

\textbf{algorithm}: 1. From datamiddlesampling \(n\) three dollar group \((x_i, p_i, y_i)\) 2.
computeencoding \(\text{PE}(p_i)\) 3. compute
\(\hat{I}^{(n)}(Y; (P, X))\): at connect combinedspace
\((\mathcal{P} \times \mathcal{X}) \times \mathcal{Y}\) middle make using \(k\)-NN
4. compute \(\hat{I}^{(n)}(Y; X)\): in \(\mathcal{X} \times \mathcal{Y}\)
middle make using \(k\)-NN 5.
\(\hat{I}^{(n)}(Y; P \mid X) = \hat{I}^{(n)}(Y; (P, X)) - \hat{I}^{(n)}(Y; X)\)
6. same reason compute \(\hat{I}^{(n)}(Y; \text{PE}(P) \mid X)\) 7.
\(\hat{\varepsilon}_{\text{PE}}^{(n)} = \hat{I}^{(n)}(Y; P \mid X) - \hat{I}^{(n)}(Y; \text{PE}(P) \mid X)\)

\textbf{convergencerate}: KSG estimate device with \(O(n^{-1/(d_{\text{eff}}+2)})\)
rateconvergence, where \(d_{\text{eff}}\) is connect combinedspacehas effect dimension. forhighdimensional
\(X\), convergencerelativelyslow, but \(\varepsilon_{\text{PE}}\)
this self is\textbf{difference}, two estimate device common share mutualsamebias difference, differenceestimatemay ratio single independent estimatemore alignment confirm (bias difference offset eliminate). 

\subsubsection{3.3.2 lower boundestimate (avoidhighdimensional
KSG)}\label{ux4e0bux754cux4f30ux8ba1ux907fux514dux9ad8ux7ef4-ksg}

forhighdimensional \(X\), KSG
estimatemayis inaccurate. replace generation scheme: use\textbf{predictionperformanceDifference} action is 
\(\varepsilon_{\text{PE}}\) lower bound. 

\textbf{Proposition 3.3.1 (\(\varepsilon_{\text{PE}}\) predictionlower bound)}: 

\[\boxed{\varepsilon_{\text{PE}} \geq \mathbb{E}_{X,P}\left[D_{\text{KL}}(P_{Y|X,P} \| P_{Y|X,\text{PE}(P)})\right]_{\text{est}}}\]

where \(P_{Y|X,P}\) and \(P_{Y|X,\text{PE}(P)}\) canthroughtwo separate class device approximate: -
separate class device \(f_{\text{full}}\) in \((X, P)\) uppertraining \to estimate \(P_{Y|X,P}\) -
separate class device \(f_{\text{enc}}\) in \((X, \text{PE}(P))\) uppertraining \to estimate
\(P_{Y|X,\text{PE}(P)}\)

twooneat retain out test set upper\textbf{logarithmicloss difference }is \(\varepsilon_{\text{PE}}\)
lower boundestimate: 

\[\hat{\varepsilon}_{\text{PE}}^{\text{pred}} = \frac{1}{n_{\text{test}}} \sum_{i=1}^{n_{\text{test}}} \left[\log \frac{1}{p_{\text{enc}}(y_i \mid x_i, \text{PE}(p_i))} - \log \frac{1}{p_{\text{full}}(y_i \mid x_i, p_i)}\right]\]

when \(n_{\text{test}} \to \infty\)
andseparate class device consistenttime, \(\hat{\varepsilon}_{\text{PE}}^{\text{pred}} \to \varepsilon_{\text{PE}}\). 

\textbf{ actual use value}: such as result separate class device \(f_{\text{full}}\) and \(f_{\text{enc}}\)
attest set upperaccuracy\textbf{nothas explicit notable Difference}, then
\(\varepsilon_{\text{PE}} \approx 0\)------encodingis fill foot, PPE will not cut weak
Theorem 2 upper bound. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{th 4 part: Theorem 3
correction (mostimportantpart)}\label{ux7b2c-4-ux90e8ux5206theorem-3-ux7684ux4feeux6b63ux6700ux91cdux8981ux7684ux90e8ux5206}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4.1 original Theorem 3
Exact Theorem Statement}\label{ux539f-theorem-3-ux7684ux7cbeux786eux5b9aux7406ux9648ux8ff0}

\textbf{Theorem 3 (original begin ------noise andDifficultysampleindistinguishability)}: 

existenceattwo world world \(W_A, W_B\) satisfies: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 observationData Distribution complete fullmutualsame: \(P_{W_A}(X, Y) = P_{W_B}(X, Y)\)
\item
 in \(W_A\) middle, type this \((x, y)\) is\textbf{Labelnoise} (label incorrect, but
 \(x\) structureupper label shouldis different)
\item
 in \(W_B\) middle, type this \((x, y)\) is\textbf{thisDifficultysample} (label
 correct, butCharacteristics \(x\) this self Nonemethoddeterminism ground keepthe label)
\end{enumerate}

then: nothas any what algorithmcanonlythroughobservation i.i.d. type this \((X_i, Y_i)_{i=1}^n\) distinguish
\(W_A\) and \(W_B\). form -ization ground, to any agree verdict strategy rule then
\(d: (\mathcal{X} \times \mathcal{Y})^n \to \{W_A, W_B\}\): 

\[\max_{W \in \{W_A, W_B\}} P_W(d(\text{data}) \neq W) \geq \frac{1}{2}\]

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4.2 learnable PE
such as what breaking indistinguishability------Exact Conditions}\label{ux5b66ux4e60ux578b-pe-ux5982ux4f55ux7834ux574fux4e0dux53efux533aux5206ux6027ux7cbeux786eux6761ux4ef6}

\subsubsection{4.2.1 Setup}\label{ux8bbeux5b9a-1}

introduce PPE post-, observationdataExtensionis three dollar group \((X, P, Y)\), where \(P\)
isphysicalLocation (mayobservationphysical quantity------atom sit mark、residueLocation etc.). key region different: 

\begin{itemize}
\tightlist
\item
 \textbf{fixed PE}: \(\text{PE}(\cdot)\)
 is pre-Definitiondeterminismfunction (such assinusoidal encoding), \textbf{not-dependenceattrainingdata}. 
\item
 \textbf{learnable PE}: \(\text{PE}_\theta(\cdot)\) parameter \(\theta\)
 isFrom datamiddlelearning (such asthroughgradient descentoptimizesome objective function). 
\end{itemize}

\subsubsection{4.2.2 fixed PE
case (Trivial)}\label{ux56faux5b9a-pe-ux7684ux60c5ux51b5trivial}

\textbf{Proposition 4.1 (fixed PE maintains Theorem 3)}: If 
\(\text{PE}: \mathcal{P} \to \mathbb{R}^{d_{\text{pe}}}\)
is a \textbf{fixed 、anddataNoneclose}function, thenstrengthensbackobservationdistribution according howevermutualsame: 

\[P_{W_A}(X, P, \text{PE}(P), Y) = P_{W_B}(X, P, \text{PE}(P), Y)\]

Theorem 3 indistinguishability\textbf{ complete fullmaintains}. 

\emph{Proof}: becauseis \(\text{PE}\)
isfixed determinismfunction, \((X, P, \text{PE}(P), Y)\) is \((X, P, Y)\)
 a determinismfunction. and \(P_{W_A}(X, P, Y) = P_{W_B}(X, P, Y)\) (physicalLocation
\(P\)
byphysicalstructure verdict determine, is independent ofLabelgenerateprocess world world select). becausethisstrengthensbackjoint distributionalsomutualsame. \(\square\)

\subsubsection{4.2.3 learnable PE
case (key)}\label{ux5b66ux4e60ux578b-pe-ux7684ux60c5ux51b5ux5173ux952e}

\textbf{Theorem 4.1 (learnable PE breaking Theorem 3 precise conditions)}: 

setup \(\text{PE}_\theta\) parameter \(\theta\) throughminimum -ization via verify risk
\(\hat{R}_n(\theta) = \frac{1}{n}\sum_{i=1}^{n} \ell(f_\theta(x_i, p_i), y_i)\)
From trainingdatamiddlelearning (where \(\ell\) issome loss function, \(f_\theta\) iscontains PE
encodingCompletemodel). thenatwithlowerConditionslower, world world \(W_A\) and \(W_B\)
 change is \textbf{maydistinguish}: 

\textbf{(C1) information theorynon- flat all ness}: 

\[I(Y; P \mid X) > 0\]

i.e.: physicalLocationattodetermine original begin Characteristicslower carry carry regardingLabelinformation. 

\textbf{(C2) learningtoLabelnoisesensitiveness}: existenceat a non- empty trainingsamplesubset
\(\mathcal{D}_{\text{diff}} \subset \mathcal{D}\), atthis some sampleupper: 

\[\nabla_\theta \ell(f_\theta(x, p), y_A) \neq \nabla_\theta \ell(f_\theta(x, p), y_B)\]

where \(y_A\) is \(W_A\) middleshouldsamplenoiseLabel, \(y_B = y_A\) (in Theorem 3
Constructionmiddle, two world world has mutualsameObserved label\(y\), onlyisexplainsis different). in practicein Theorem
3 Constructionmiddle, \textbf{observationto \((x, y)\) complete fullmutualsame}! 

this agree taste -ing: learnable PE
breaking indistinguishabilitymechanism\textbf{is NOT }fromLabelDifference (becauseis Labelmutualsame), andisfrommodelto''such as what remove learning''Difference. let we Exact Analysis. 

\subsubsection{4.2.4
Exact mechanism: table characteristic drift shift }\label{ux7cbeux786eux673aux5236ux8868ux5f81ux6f02ux79fb}

\textbf{key hole observe }: in Theorem 3
Constructionmiddle, \(P_{W_A}(X, Y) = P_{W_B}(X, Y)\)
isholds, but\textbf{this and not agree taste -ing learningdynamicmutualsame}. Reasonatat: 

in \(W_A\) (Labelnoise world world)middle: - type this \((x, y)\) Label \(y\) isincorrect -
physicalLocation \(p\) and \((x, y)\) relationshipis: \(p\) and \(x\)
consistent (byphysicalstructure verdict determine), but \(p\) dark show trueLabeland \(y\) is different - when PE
parameterbylearningtime, modelwilldiscover \((p, y)\) 'sbetween'' contradiction contradiction ''------\(p\)
dark show some Label, and \(y\) is another a 

in \(W_B\) (thisDifficulty world world)middle: - type this \((x, y)\) Label \(y\) iscorrect -
\(p\) and \(x\) and \(y\) allconsistent------does not exist'' contradiction contradiction '' - modellearningto
\((p, y)\) relationshipisand harmonic consistent

\textbf{key region different }: i.e. make 
\(P_{W_A}(X, P, Y) = P_{W_B}(X, P, Y)\) (connect combinedobservationdistributionmutualsame), learning\textbf{dynamic}depends onloss surface several what structure------andshouldstructureatConditionsdistribution
\(P_{Y|X,P}\) direction surface has Difference (becauseis Theorem 3 Construction-dependenceatConditionsdistributionDifference, i.e.
\(X\) intodetermine \((Y, \text{world})\) lowerdistributionis different). 

\textbf{thisis Theorem 3
Constructioncore}: \(P_{W_A}(Y \mid X) = P_{W_B}(Y \mid X)\) (observationdistributionmutualsame), butgenerateprocessis different: 
- \(W_A\): \(Y_{\text{true}}\) first by \(X\) verdict determine, howeverpost-\(Y_{\text{obs}}\)
withprobability \(\eta\) isflipped -
\(W_B\): \(Y_{\text{obs}} = Y_{\text{true}}\), but
\(P(Y_{\text{true}} \mid X)\) insome some region domain approachesuniform uniform (highnotdeterminism)

whenintroduce \(P\) post-, in \(W_A\) middle \(P(Y_{\text{true}} \mid X, P)\) cancan ratio 
\(P(Y_{\text{obs}} \mid X, P)\) more tip sharp (becauseis \(P\)
 help help confirm determinetrueLabel), leads tomodelin \((X, P)\)
spacemiddle'' see to''noiseLabel contradiction contradiction. andin \(W_B\)
middle, \(P(Y_{\text{obs}} \mid X, P) = P(Y_{\text{true}} \mid X, P)\), nothas contradiction contradiction. 

\textbf{butthisstillhowevernothas to out mayobservationDifference}------becauseis intrainingtime, modelattwo world world middle see toallismutualsame
\((x, p, y)\) three dollar group. 

truecurrently make learnable PE can enough distinguishmechanismis: Theorem 3
Constructionmiddletwo world world region different atat\textbf{notobservationto latent atvariable} (trueLabel vs
noiseLabel vs Difficulty process degree). such as result \(P\)
provideregardingthis some notobservationvariableinformation window gap (through
\(I(Y_{\text{true}}; P \mid X) > 0\)), that what \textbf{attrainingprocessmiddle}, learningto
PE parameter \(\theta\)
willthroughgeneralizationlinesis anti- reflect this some Difference------i.e. make trainingdata see rise come mutualsame. 

\subsubsection{4.2.5
Exact candistinguishnessConditions------mostfinalform}\label{ux7cbeux786eux53efux533aux5206ux6027ux6761ux4ef6ux6700ux7ec8ux5f62ux5f0f}

 let we From morebasicAspectheavynewAnalysis. 

in Theorem 3 Proofmiddle, two world world Constructionis: - selecttwo is differentConditionsdistribution
\(P_{W_A}(Y \mid X)\) and \(P_{W_B}(Y \mid X)\), but edge actual distribution \(P(X, Y)\)
mutualsame - this's withmay, isbecauseis \(P(X)\) intwo world world middleis different \to
\(P_{W_A}(X) \neq P_{W_B}(X)\), but \(P_{W_A}(X, Y) = P_{W_B}(X, Y)\)

 etc. etc., thisis not correct. Theorem 3 say is\textbf{observation}distributionmutualsame. let we From CC
Audit Report original begin Construction out emit. 

CC Audit Report: ``existenceattwo world world \(W_A, W_B\), has has observationequivalentData Distribution
\(P_{W_A}(X, Y) = P_{W_B}(X, Y)\)''

ConstructionMethodology: two world world has has is different \(P(X)\) andis different
\(P(Y|X)\), but it s productmutualsame. specifically: - \(W_A\): \(P_{W_A}(X)\)
at''easy'' region domain has moremanyquality, \(P_{W_A}(Y|X)\) isdeterminism -
\(W_B\): \(P_{W_B}(X)\) in''Difficulty'' region domain has moremanyquality, \(P_{W_B}(Y|X)\)
isapproachesuniform uniform 

makes \(P_{W_A}(X, Y) = P_{W_B}(X, Y)\) to has \((x, y)\) holds. 

\textbf{whenintroduce \(P\)
back}: \(P_{W_A}(X, P, Y) = P_{W_A}(P \mid X, Y) \cdot P_{W_A}(X, Y)\). 

such as result \(P\) isphysicallyby \(X\) verdict determine (\(P\) is \(X\) physical sit mark), then
\(P(P \mid X, Y) = P(P \mid X)\) (physicalLocationnot-dependenceatLabel). thistime: 
\[P_{W_A}(X, P, Y) = P_{W_A}(P \mid X) \cdot P_{W_A}(X, Y)\]
\[P_{W_B}(X, P, Y) = P_{W_B}(P \mid X) \cdot P_{W_B}(X, Y)\]

but
\(P_{W_A}(P \mid X) = P_{W_B}(P \mid X)\) (physicalLocationbyphysicalstructure verdict determine, totwo world world mutualsame), and
\(P_{W_A}(X, Y) = P_{W_B}(X, Y)\). becausethis
\(P_{W_A}(X, P, Y) = P_{W_B}(X, P, Y)\). 

\textbf{fixed PE stillhoweverpreserves unidentifiability}------joint distributionmutualsame. 

\textbf{learnable PE case}: learnable PE parameter \(\theta\)
isthroughminimum -ization some loss functionFrom datamiddle study to. consideratdata set 
\(\mathcal{D} = \{(x_i, p_i, y_i)\}_{i=1}^n\) uppertraining: 

\[\hat{\theta} = \arg\min_\theta \frac{1}{n} \sum_{i=1}^n \ell(f_\theta(x_i, \text{PE}_\theta(p_i)), y_i)\]

this inside keyis: \textbf{PE
parameteratloss functionmiddle}------notonlythroughencoding input out impactloss, stillmaythroughsomekindself-supervisesorreconstructionlossimpactlearning. 

butsuch as result PE
pure pure ishas supervises ground training (onlythroughlower swim any task loss), that what two world world to out mutualsametrainingdata
\to mutualsamegradient \to mutualsame \(\hat{\theta}\). in this case, i.e. make islearnable PE
alsocannotdistinguishtwo world world. 

\textbf{truecurrently make learnable PE can enough distinguishscenario}: 

PE parameterexcept receive superviseslossimpactouter, stillthroughwithlower direction formula connect collect signal: 1.
\textbf{self-supervises pre-training} (such as masked position
prediction)------atmarkNote'sfront advance lines 2. \textbf{assistsloss} (such asLocationreconstructionloss)
3. \textbf{physical constraint} (such asenergy guard constant 、forcematches)------this some convention bind provideregarding \(P\)
andphysical quantity'sbetweenrelationship amount outerinformation

atthis some caselower, PE
parameterpartby\textbf{physicalconsistentnesssignal} verdict determine, andthis some signalattwo world world middlemayis different (becauseis two world world middlesample''physicalreasonableness''is different). 

\textbf{Exact Theorem}: 

\textbf{Theorem 3' (correctionbackindistinguishabilityTheorem------ carry PPE pre- raise)}: 

\[\boxed{\begin{aligned}
\text{If } & \text{(A) } \text{PE}(\cdot) \text{ isandLabel } Y \text{ independent fixed function (i.e.itsparameternotby } Y \text{ or any what -dependenceat } Y \text{ signal verdict determine), and} \\
& \text{(B) } P \perp\!\!\!\perp Y_{\text{world}} \mid X \text{ (physicalLocationattodetermine original begin Characteristicsloweris independent of world world select), } \\
\text{then } & P_{W_A}(X, \text{PE}(P), Y) = P_{W_B}(X, \text{PE}(P), Y), \text{Theorem 3 indistinguishabilitymaintains. } \\
\text{ counter 's, If } & \text{PE parameterthrough-dependenceat } Y \text{ signallearning obtain to, and } I(Y_{\text{true}}; P \mid X) > 0, \\
\text{then } & \text{learningtoencodingattwo world world middlewithcurrentlyprobabilityis different: } P(\hat{\theta}_{W_A} \neq \hat{\theta}_{W_B}) > 0, \\
\text{two world world change is candistinguish. } &
\end{aligned}}\]

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4.3
ConstructionnessProof: minimum show example }\label{ux6784ux9020ux6027ux8bc1ux660eux6700ux5c0fux793aux4f8b}

\subsubsection{4.3.1 Construction}\label{ux6784ux9020}

consider a extremely simple two dimensionalseparate class Problem. setup \(X \in \{-1, +1\}\)
is a two dollar Characteristics, \(P \in \{0, 1\}\)
is a two dollar physicalLocation, \(Y \in \{-1, +1\}\) isLabel. 

\textbf{ world world \(W_A\) (Labelnoise)}: 

\begin{itemize}
\tightlist
\item
 \(P_{W_A}(X = +1) = 1/2\), \(P_{W_A}(X = -1) = 1/2\)
\item
 \(P_{W_A}(Y_{\text{true}} = +1 \mid X = +1) = 1\),
 \(P_{W_A}(Y_{\text{true}} = -1 \mid X = -1) = 1\) (determinism)
\item
 Labelwithprobability \(\eta = 0.2\) flipped: \(Y_{\text{obs}} = Y_{\text{true}}\)
 withprobability \(0.8\), \(Y_{\text{obs}} = -Y_{\text{true}}\) withprobability \(0.2\)
\item
 physicalLocation \(P\): \(P = X\) (determinism), i.e. \(p = 1\) when
 \(x = +1\), \(p = 0\) when \(x = -1\)
\end{itemize}

\textbf{ world world \(W_B\) (thisDifficulty)}: 

\begin{itemize}
\tightlist
\item
 \(P_{W_B}(X = +1) = 0.6\), \(P_{W_B}(X = -1) = 0.4\) (non-uniform uniform !)
\item
 \(P_{W_B}(Y_{\text{obs}} = +1 \mid X = +1) = 5/6\),
 \(P_{W_B}(Y_{\text{obs}} = -1 \mid X = +1) = 1/6\)
\item
 \(P_{W_B}(Y_{\text{obs}} = +1 \mid X = -1) = 1/4\),
 \(P_{W_B}(Y_{\text{obs}} = -1 \mid X = -1) = 3/4\)
\item
 \(Y_{\text{obs}} = Y_{\text{true}}\) (nothas Labelnoise, butConditionsdistributionisintrinsichas highnotdeterminism)
\item
 \(P = X\) (and \(W_A\) mutualsame)
\end{itemize}

\textbf{verifyobservationequivalent}: 

in \(W_A\) middle: - \(P_{W_A}(X = +1, Y = +1) = 0.5 \times 0.8 = 0.4\) -
\(P_{W_A}(X = -1, Y = +1) = 0.5 \times 0.2 = 0.1\) -
\(P_{W_A}(X = +1, Y = -1) = 0.5 \times 0.2 = 0.1\) -
\(P_{W_A}(X = -1, Y = -1) = 0.5 \times 0.8 = 0.4\)

in \(W_B\) middle: - \(P_{W_B}(X = +1, Y = +1) = 0.6 \times (5/6) = 0.5\) -
\(P_{W_B}(X = -1, Y = +1) = 0.4 \times (1/4) = 0.1\) -
\(P_{W_B}(X = +1, Y = -1) = 0.6 \times (1/6) = 0.1\) -
\(P_{W_B}(X = -1, Y = -1) = 0.4 \times (3/4) = 0.3\)

\(P(X=+1,Y=+1)\) is different (0.4 vs
0.5), does not satisfyobservationequivalent. let we adjustConstruction make its complete fullmutual etc.. 

heavynewConstruction: 

\(W_A\): - \(P(X=+1) = 0.5\), \(P(X=-1) = 0.5\) -
\(P(Y=+1 \mid X=+1) = 0.8\), \(P(Y=-1 \mid X=+1) = 0.2\) -
\(P(Y=+1 \mid X=-1) = 0.2\), \(P(Y=-1 \mid X=-1) = 0.8\) -
generatemechanism: \(Y_{\text{true}}\) isdeterminism (\(X=+1 \to +1\),
\(X=-1 \to -1\)), but 20\% noise

\(W_B\): - \(P(X=+1) = 0.5\), \(P(X=-1) = 0.5\) - and \(W_A\) complete fullmutualsame
\(P(Y \mid X)\) - generatemechanism: Characteristicsthis self ishighnotdeterminism (nothas implicit hide trueLabel)

atthisConstructionlower \(P_{W_A}(X, Y) = P_{W_B}(X, Y)\), butintrinsicstructureis different. in \(W_A\)
middle, \(P = X\), \(Y_{\text{true}} = X\), noisewith 20\% probabilityflipped. in \(W_B\)
middle, \(P = X\), but \(Y\) directby carry noise through path generate (Noneimplicit hide layer). 

\subsubsection{4.3.2 learnable PE
maydistinguishness}\label{ux5b66ux4e60ux578b-pe-ux7684ux53efux533aux5206ux6027}

introduce a simple singlelearnable 
PE: \(\text{PE}_\theta(p) = \theta \cdot p\) (mark quantity encoding, \(\theta \in \mathbb{R}\)
maylearning). 

Completemodel: \(f(x, p) = \text{sign}(w \cdot x + \text{PE}_\theta(p))\). 

setup we s throughwithlower direction formula learning \(\theta\): 1.
 first at\textbf{NoneLabel}dataupperwithself-supervises direction formula pre-training: predictionis mask physicalLocation
\(\hat{p} = g(\text{PE}_\theta(p))\) 2. howeverbackatlower swim any task uppermicro- adjust 

key: atStep 1 (self-supervises pre-training)middle, objective functiondoes not involveLabel
\(Y\). two world world data \((X, P)\)
distribution complete fullmutualsame (\(P_{W_A}(X, P) = P_{W_B}(X, P)\)), becausethis pre-trainingorder segment \textbf{notintroduce}distinguish. 

butsuch as result we s usewithlowerassistsloss: 

\[\mathcal{L}_{\text{aux}}(\theta) = \mathbb{E}_{(x,p,y)}\left[(\text{PE}_\theta(p) - \mathbb{E}[Y \mid X = x])^2\right]\]

i.e.: encodingshouldandConditionsexpectLabelconsistent. this lossinvolving
\(Y\), and-dependenceat world world innerpartstructure. in \(W_A\)
middle, \(\mathbb{E}[Y \mid X = x] = (1-2\eta) \cdot \text{sign}(x) = 0.6 \cdot \text{sign}(x)\) (becauseis noiseflipped). at
\(W_B\)
middle, \(\mathbb{E}[Y \mid X = x] = \mathbb{E}[Y \mid X = x] = 0.6 \cdot \text{sign}(x)\) (number value mutualsame! onlyisexplainsis different). 

 withat\textbf{this minimum show example middle}, i.e. make islearnable PE
alsocannotdistinguishtwo world world ------becauseis Conditionsexpectmutualsame. 

\textbf{thiscurrentlyis Theorem 3
 refined marrow }: such as result observationdistribution complete fullmutualsame, any what algorithmallcannotdistinguish. 

\subsubsection{4.3.3 open problems: learnable PE
truecurrentlycandistinguishConditions}\label{ux5f00ux653eux95eeux9898ux5b66ux4e60ux578b-pe-ux771fux6b63ux80fdux533aux5206ux7684ux6761ux4ef6}

\textbf{This minimal example actually proves Theorem 3
robustness, not its fragility}: atthis Constructionmiddle, i.e. make islearnable PE
alsocannotdistinguishtwo observationequivalent world world. this exactly exactly impression prove Theorem 3
 refined marrow ------whenobservationdistribution complete fullmutualsametime, any what algorithmallcannotdistinguish. 

 original first claims''Constructionnessminimum show example '' counter andconstitutes Theorem 3'
breaking direction a \textbf{ counter example }: it demonstrate a learnable PE
\textbf{Nonemethod}distinguishtwo observationequivalent world world scenario. 

must make learnable PE
candistinguish, requiresatisfies: \textbf{learningsignalnotonlyonly-dependenceatobservationdistribution}. theoreticallythismay emit life at: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{many any task learning}: PE sametimebyused forprediction another a and \(Y_{\text{true}}\)
 mutualclose quantity \(Z\) (example such asenergy、force), and \(Z\) in world world 
 \(W_A\) (noiseLabel agree taste -ing physical quantityis inconsistent)and
 \(W_B\) (Difficultybutphysicalconsistent) in the table present is different
\item
 \textbf{physicalconsistentness convention bind }: loss functioncontainsphysical constraint term (such asenergy guard constant), makesat
 \(W_A\) middlenoiseLabelleads tophysicalis inconsistentprediction, andin \(W_B\)
 middleDifficultysamplestillmaintainsphysicalconsistentness
\item
 \textbf{distributionoutergeneralization}: PE
 parameterpartbydistributionouterdata (differs fromtrainingdistribution) verdict determine, andtwo world world atdistributionouter table present is different
\end{enumerate}

\textbf{{[}open problems{]}}
howeverand, goal front\textbf{lack ofConstructionnessProof}verifyupper describe any what mechanism confirm actual canatobservationequivalenttwo world world upperproducesis differentOptimal 
PE parameter. thisrequire: - explicit formula Definitionnotobservationvariable \(Z\) - Construction
\(\mathcal{L}_{\text{aux}}\) concreteform - Proofin \(W_A\) and \(W_B\)
 in the Optimal value confirm actual is different

Theorem 4.1 stated describe (regarding
\(\mathcal{L}_{\text{total}} = \mathcal{L}_{\text{sup}} + \lambda \mathcal{L}_{\text{aux}}\)
is differentOptimal value leads tomaydistinguishness)at logic edit upperiscorrect\textbf{Conditions stated describe }, butshouldConditionsiswhethercanat
Theorem 3 Constructionmiddlebysatisfies, goal frontnothas ConstructionnessProof. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4.4 correctionpost-Theorem
3'------Complete stated describe }\label{ux4feeux6b63ux540eux7684-theorem-3ux5b8cux6574ux9648ux8ff0}

\textbf{Theorem 3' (indistinguishabilityConditionsmaintains------Corrected Version)}: 

\[\boxed{\begin{aligned}
\textbf{front raise (A): } & \text{physicalLocationencoding PE isfixed, oronlythroughis independent ofLabel } Y \text{ signallearning. } \\
\textbf{front raise (B): } & \text{physicalLocation } P \text{ intodetermine } X \text{ lowerConditionsis independent of world world select: } P \perp\!\!\!\perp W \mid X. \\
\textbf{Conclusion: } & \text{Theorem 3 indistinguishability complete fullmaintains: } \\
& \forall \text{ verdict strategy rule then } d, \quad \max_{W \in \{W_A, W_B\}} P_W(d \neq W) \geq \frac{1}{2}. \\
\textbf{If premise (A) does not hold: } & \text{whether unidentifiability is broken is currently an**open problems**. } \\
& \text{learnable PE cancanthroughmany any task learningorphysicalconsistentness convention bind etc. mechanismdistinguishtwo world world, } \\
& \text{butwhenfrontlack ofConstructionnessProof. see §4.3 discuss theory. }
\end{aligned}}\]

\textbf{Corollary 4.4.1 (topracticeimpact)}: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{fixed currently sinusoidal PE orfixed rotation PE}: Theorem 3
 complete fullmaintains. Locationencodingwill not damage harm mostbadcaseindistinguishabilityguarantee------butalso\textbf{will notmain action help help }distinguishnoiseandDifficultysample (tofewatinformation-theoretic senseupper, throughobservationdistributionNonecannot distinguish). 
\item
 \textbf{From NoneLabelLocationdatalearning
 PE} (such asthroughself-supervisesreconstructionLocation): such as result trainingprocessdoes not involveLabel \(Y\), Theorem 3
 maintains. PE parameterattwo world world middle study tomutualsame value. 
\item
 \textbf{throughmany goal marklearning (includesLabelmutualclose goal mark) PE}: Theorem 3
 \textbf{maybybreaking }, but goal frontlack ofConstructionnessProof. thisis\textbf{open problems}------Locationinformationiswhethertrueprovide a amount outer'' window gap '' come distinguishnoiseandDifficultysample, stillis a notbyrigorousverifyConjecture. 
\item
 \textbf{risk}: i.e. make assists goal markmay help help distinguish, If itsdesignintroduce virtual falsemutualcloseness, mayleads toincorrect''distinguish'', willDifficultysampleincorrect ground Marked as noise. thisrequire careful fineexperimentverify. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4.5 and CC Audit Report §2.5
comparison}\label{ux4e0e-cc-ux5ba1ux8ba1ux62a5ux544a-2.5-ux7684ux6bd4ux8f83}

CC Audit Reportto Theorem 3 Analysisiscorrect (fixed PE preserves unidentifiability, learnable 
PE cancanbreaking), but leave leak withlowerkey point: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{CC reportnotExact Definition''learnable 
 PE''Meaning}: only say ''From datamiddlelearning''isis insufficient------requirebright confirm distinguish''From NoneLabeldatalearning'' (safety)and''From Label-dependencesignallearning'' (maybreaking 
 Theorem 3). 
\item
 \textbf{CC reportnotto out maydistinguishnessprecise conditions}: Theorem 4.2 fill supplement this a empty white. 
\item
 \textbf{CC
 reportevaluate''thisisgood eliminate rest ''require finite determine}: onlyhas inassists goal marksignalandtrueLabelnoisemutualclose (andnon- virtual falsemutualclose)time, learnable 
 PE distinguishcanforceonly thenismay reliable good eliminate rest. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Summary}\label{ux603bux7ed3}

this text to PPE in SCX framework in the action use advance linesrigorousmathematicalDerivation, primarycontributionincludes: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{encodingfunctionform -ization } (th 1
 part): provides three kindphysicalLocation domain currently sinusoidal androtationencodingExact Definition, Derivationbased on
 Bochner Theoremand Laplace kernelOptimal frequencyspectrum, provesencoding Lipschitz
 continuousness and computeExact often number. 
\item
 \textbf{Theorem 1 correction} (th 2 part): correction CC Audit Report in the sign
 error, provides \(\delta_s^{\text{PE}}\) correct table reach formula, proves
 \(\delta_s^{\text{PE}} > 0\)
 Sufficient Condition (\(I(Y; P \mid S) > 0\)), and Derivationbased ondataat reason inequalityand
 Fano inequalityupperlower bound. 
\item
 \textbf{Theorem 2 correction} (th 3 part): provides
 \(\varepsilon_{\text{PE}}\) Exact information theoryDefinition, proves
 \(\varepsilon_{\text{PE}} \to 0\)
 timeupper bound restore complex is original begin form, and provideFrom datamiddlemayestimate
 \(\varepsilon_{\text{PE}}\) twokind actual use Methodology (KSG
 estimate device andpredictionperformanceDifference). 
\item
 \textbf{Theorem 3 correction} (th 4 part): Exact clarify brightfixed PE andlearnable PE
 toindistinguishabilityis differentimpact, throughminimumConstructionness show example demonstratedistinguishnecessaryConditions, raise out containsExact pre- raise Conditionscorrection
 Theorem 3'. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Derivationcomplete day period: 2026-06-29}\\
\emph{Note: has TheoremNumber follow follow SCX frameworkinnerpart convention determine. this text correction CC
Audit Report in the Symbol/Conceptis inconsistent (special different is §2.3 Proposition 2.1 sign
error), and atshouldreportFoundationupperincreasesrigorousupperlower boundDerivation、ConstructionnessProofandmayestimateform. }

\end{document}
