% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX Self-Evolution Theory ---
Mathematical Formalization}\label{scx-self-evolution-theory-mathematical-formalization}

\begin{quote}
Formalize SCX's self-evolution closed loop (judge $\to$ store $\to$ update $\to$ re-judge $\to$ re-update $\to$
\ldots) as a rigorous mathematical theory, and establish connections with known theoretical frameworks.
\end{quote}

\textbf{Last Updated}: 2026-06-28

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Problem}\label{problem}

The SCX system forms the following closed loop during operation:

\begin{verbatim}
Judge (S_t scoring) → Store (M_t memory bank) → Update (S_{t+1} scoring update) → Re-judge → ...
                                ↘ Update (θ_{t+1} NEP student update) ↗
\end{verbatim}

\textbf{Core Question}: Does this closed loop converge? To what? At what rate? Under what conditions does it diverge?

\textbf{Difficulty}: The Gatekeeper \(S_t\) and NEP student \(f_{\theta_t}\)
form a coupled dynamical system---\(S_t\)'s selection determines \(f_{\theta_t}\)'s
training distribution, while \(f_{\theta_t}\)'s output provides feedback to \(S_t\).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Core Insight}\label{core-insight}

\begin{verbatim}
The correct mathematical object for SCX self-evolution is the coupled dynamical system (S_t, θ_t, M_t),
whose global behavior is dominated by the monotone descent property of a Lyapunov function V(S, θ).
\end{verbatim}

where \(S_t: \mathcal{X} \to [0,1]\) is the gatekeeper scoring function, \(\theta_t\)
is the NEP student parameter, and \(M_t\) is the monotonically growing empirical memory bank.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Nine Core Files}\label{nine-core-files}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3571}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2143}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
File
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Content
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Core Result
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{01\_symbol\_system.md} & Symbol System and Problem Setup & Structure space
\(\mathbb{X}\), state space \(\mathbb{Z}\), update operator \(\Phi\), new assumptions SE-A1
to SE-A6 & \textbf{Complete} \\
\texttt{02\_dynamical\_system.md} & Discrete Dynamical System Formalization & Lyapunov function
\(V(z_t)\), fixed-point existence (Theorems 4--5), monotone descent (Theorem 6), phase diagram of four regions &
\textbf{Complete} \\
\texttt{03\_online\_learning\_regret.md} & Online Learning Regret Analysis & Regret
\(\leq 2GR\sqrt{T}\) (OGD), delayed feedback Regret
\(\leq 2GR\sqrt{T} + G^2 D_{\max}\sqrt{T}\) (Theorem 8) & \textbf{Complete} \\
\texttt{04\_bayesian\_update.md} & Bayesian Update Interpretation & \(S_t\)
as posterior mean, Bayesian martingale \(\mathbb{E}[S_{t+1}|\mathcal{F}_t]=S_t\), Doob
convergence \(S_t \to S_\infty\) a.s. & \textbf{Complete} \\
\texttt{05\_stochastic\_approximation.md} & Stochastic Approximation Analysis & Robbins-Monro
form \(\theta_{t+1} = \theta_t - \alpha_t \nabla \ell + \xi_t\), ODE
method, two-timescale analysis & \textbf{Complete} \\
\texttt{06\_fixed\_point\_convergence.md} &
\textbf{Central Theorem: Fixed Point and Convergence} & \textbf{Theorem SE-1}: finite structure space +
Lipschitz + sufficient annealing $\to$ \((S_t, \theta_t) \to (S^*, \theta^*)\)
a.s.; four convergence paths & \textbf{Complete} \\
\texttt{07\_completeness.md} & Completeness Analysis & \textbf{Theorem
SE-2}: under physical constraints, $\exists$ finite \(T^*\) such that for \(t \geq T^*\) the system reaches
\(\varepsilon\)-approximate fixed point; G\"odel analogy & \textbf{Complete} \\
\texttt{08\_theory\_connections.md} & Connections to Known Theories & Formalized comparison table with AlphaZero
self-play, Bayesian optimization, active learning, Solomonoff induction &
\textbf{Complete} \\
\texttt{09\_verification\_report.md} & Verification Report &
Consistency checks, cross-theorem cross-references, symbol conflicts, proof gaps (10), open problems (5) & \textbf{Complete} \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Core Theorems}\label{core-theorems}

\subsubsection{Theorem SE-1: Convergence of SCX
Self-Evolution}\label{theorem-se-1-convergence-of-scx-self-evolution}

\textbf{Conditions (C1--C7)}: 
1. Structure space \(\mathbb{X}\) finite (finite data, finite precision)
2. NEP student \(f_\theta\) is \(L_f\)-Lipschitz in \(\theta\)
3. Gatekeeper \(S_t\) is \(L_S\)-Lipschitz in its parameters
4. Learning rate satisfies Robbins-Monro conditions: \(\sum \alpha_t = \infty, \sum \alpha_t^2 < \infty\)
5. Given \(S_t\), memory bank sampling is i.i.d.
6. Sufficient annealing: \(\alpha_t \to 0\), and gatekeeper update frequency \(\to 0\)
7. Gatekeeper update step is bounded: \(\|S_{t+1} - S_t\|_\infty \leq \eta_t\), \(\sum \eta_t < \infty\)

\textbf{Conclusion}: The sequence \((S_t, \theta_t)\) converges almost surely to a joint fixed point
\((S^*, \theta^*)\), where:
- \(S^*\) is self-consistent: \(S^*(x) = \mathbb{P}(\text{correct} \mid x, \mathcal{M}_\infty)\)
- \(\theta^*\) is a local minimum of the NEP expected loss under the distribution induced by \(S^*\)

\textbf{Proof Path}: Lemma SE-1.1 (Lyapunov supermartingale) + Lemma SE-1.2
(finite types) + Lemma SE-1.3 (vanishing step) + Lemma SE-1.4 (limit point is fixed point).

\subsubsection{Theorem SE-2: Completeness
Bound}\label{theorem-se-2-completeness-bound}

Under physical constraints (finite data \(N_{\max}\), finite precision
\(\varepsilon_{\text{mach}}\), finite compute), there exists a finite time \(T^*\)
such that for all \(t \geq T^*\), the system is at an \(\varepsilon\)-approximate fixed point.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Four Convergence Paths}\label{four-convergence-paths}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Path
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Characteristics
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Conditions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{I. Classical Convergence} & \((S_t, \theta_t) \to (S^*, \theta^*)\), monotone improvement
& C1--C7 all satisfied, sufficient annealing \\
\textbf{II. Limit Cycle} & System oscillates periodically among finite configurations & Insufficient annealing, excessive coupling \\
\textbf{III. Perpetual Discovery} & New structures continuously discovered, \(\mathcal{M}_t\) grows without bound &
Open physical world, infinite exploration budget \\
\textbf{IV. Divergent Collapse} & \(S_t\) degenerates, quality declines & Feedback loop broken, NEP
feedback noise excessive \\
\end{longtable}

The precise boundary conditions for paths II--IV are \textbf{open problems} (see
\texttt{09\_verification\_report.md}).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Relationship to Existing Theoretical Frameworks}\label{relationship-to-existing-theories}

\subsubsection{Dependency Graph}\label{dependency-graph}

\begin{verbatim}
Theorem 3 (Unidentifiability) ─── provides necessity argument for self-evolution ──→ self-evolution loop must exist
        │
        ▼
Theorem 1 (Noise Detection) ──── provides guarantee for S_t initialization ──→ initial condition for Theorem SE-1
        │
        ▼
Theorem SE-1 (Self-Evolution Convergence) ─── central result ──→ loop converges asymptotically to self-consistent point
        │
        ├──→ Theorem 2 (Weak Feature Bound) ── constrains improvement rate
        ├──→ Theorem 4' (Minimax Optimal) ── provides asymptotic optimality of S^*
        └──→ Theorem SE-2 (Completeness Bound) ── physical finiteness enforces finite-time convergence
\end{verbatim}

\subsubsection{Newly Introduced Assumptions (SE-A1 to
SE-A6)}\label{new-assumptions-se-a1-to-se-a6}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Assumption & Content & Corresponding Existing Assumption \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{SE-A1} & Lyapunov descent & \emph{Newly introduced, core assumption} \\
\textbf{SE-A2} & Finite structure space & Physical constraint (not statistical) \\
\textbf{SE-A3} & Lipschitz NEP & Smoothness assumption \\
\textbf{SE-A4} & Conditional i.i.d.\ sampling & Extension of A1--A2 \\
\textbf{SE-A5} & Robbins-Monro learning rate & Standard SA condition \\
\textbf{SE-A6} & Gatekeeper update bounded & Annealing condition \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Connection Summary with Known Theories}\label{connection-summary}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2273}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2273}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.5455}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Framework
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Core Mechanism
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Key Differences from SCX Self-Evolution
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{AlphaZero} & Policy iteration + MCTS & AlphaZero generates synthetic games; SCX relies on real NEP data \\
\textbf{Bayesian Optimization} & Acquisition + Surrogate & BO optimizes a static black box; SCX's target distribution evolves \\
\textbf{Active Learning} & Query strategy + Oracle & AL's oracle gives instant feedback; SCX's NEP gives delayed feedback \\
\textbf{Solomonoff Induction} & Universal prior & Solomonoff is uncomputable; SCX is restricted to finite hypothesis classes \\
\end{longtable}

See \texttt{08\_theory\_connections.md} for the full comparison table.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Theoretical Status Assessment}\label{theoretical-status}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Proportion
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Explanation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Rigorous Theory} & \textasciitilde40\% & Finite-state version of Theorem SE-1, Bayesian martingale convergence, Robbins-Monro standard convergence, Regret upper bound \\
\textbf{Formalized Conjecture} & \textasciitilde20\% & Continuous state-space extension, two-timescale convergence, KL contraction rate \\
\textbf{Reasonable Hypothesis} & \textasciitilde40\% & Exact Lyapunov function form, limit cycle conditions, G\"odel analogy, optimal annealing strategy \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Open Problems (Priority-Sorted)}\label{open-problems}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Priority & Problem & File \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{P0} & Explicit Lyapunov function \(\Phi\) definition and descent proof &
\texttt{06}, \texttt{09} \\
\textbf{P1} & Precise boundaries for the four convergence paths & \texttt{06} \\
\textbf{P1} & Tight convergence rate for the coupled \((S_t, \theta_t)\) system & \texttt{05},
\texttt{06} \\
\textbf{P2} & Optimal gatekeeper update frequency & \texttt{02}, \texttt{03} \\
\textbf{P2} & Minimum memory bank size for \(\varepsilon\)-convergence & \texttt{07} \\
\textbf{P3} & Non-asymptotic (finite-time) guarantees & \texttt{03}, \texttt{06} \\
\textbf{P4} & Robustness characterization under distribution shift & \texttt{05} \\
\textbf{P5} & Formalization of the G\"odel analogy & \texttt{07} \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{File Structure}\label{file-structure}

\begin{verbatim}
theory/self_evolution/
├── README.md                              ← This file
├── 01_symbol_system.md                    # Symbol System and Problem Setup
├── 02_dynamical_system.md                 # Discrete Dynamical System Formalization
├── 03_online_learning_regret.md           # Online Learning Regret Analysis
├── 04_bayesian_update.md                  # Bayesian Update Interpretation
├── 05_stochastic_approximation.md         # Stochastic Approximation Analysis
├── 06_fixed_point_convergence.md          # Fixed Point and Convergence Theorems (Central)
├── 07_completeness.md                     # Completeness Analysis
├── 08_theory_connections.md               # Connections to Known Theories
└── 09_verification_report.md              # Verification Report
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Cross-Domain Applicability}\label{cross-domain}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1053}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1930}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3684}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Domain
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(S_t\) Role
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(f_{\theta_t}\) Role
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\mathcal{M}_t\) Role
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{MLIP} & Judging correctness of DFT labels & NEP potential energy surface model & Accumulated DFT computation trajectory \\
\textbf{Medical Imaging} & Judging diagnostic label quality & Auxiliary diagnostic model &
Accumulated expert review records \\
\textbf{LLM} & Judging annotation/generation quality & Downstream task model & Accumulated human review logs \\
\textbf{Autonomous Driving} & Judging detection box/label quality & Perception model &
Accumulated road-test validation data \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References (Newly Added)}\label{references}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Robbins, H., \& Monro, S. (1951). A stochastic approximation method.
  \emph{Annals of Mathematical Statistics}, 22(3), 400--407.
\item
  Borkar, V. S. (2008). \emph{Stochastic Approximation: A Dynamical
  Systems Viewpoint}. Cambridge University Press.
\item
  Kushner, H. J., \& Yin, G. G. (2003). \emph{Stochastic Approximation
  and Recursive Algorithms and Applications} (2nd ed.). Springer.
\item
  Doob, J. L. (1953). \emph{Stochastic Processes}. Wiley.
\item
  Bernardo, J. M., \& Smith, A. F. M. (1994). \emph{Bayesian Theory}.
  Wiley.
\item
  van der Vaart, A. W. (1998). \emph{Asymptotic Statistics}. Cambridge
  University Press.
\item
  Silver, D., et al.~(2017). Mastering the game of Go without human
  knowledge. \emph{Nature}, 550, 354--359.
\item
  Shahriari, B., et al.~(2016). Taking the human out of the loop: A
  review of Bayesian optimization. \emph{Proceedings of the IEEE},
  104(1), 148--175.
\item
  Settles, B. (2009). Active learning literature survey.
  \emph{University of Wisconsin-Madison Technical Report}.
\item
  Solomonoff, R. J. (1964). A formal theory of inductive inference.
  \emph{Information and Control}, 7(1), 1--22, 224--254.
\item
  Li, M., \& Vit\'anyi, P. (2008). \emph{An Introduction to Kolmogorov
  Complexity and Its Applications} (3rd ed.). Springer.
\item
  Zinkevich, M. (2003). Online convex programming and generalized
  infinitesimal gradient ascent. \emph{ICML}.
\item
  Cesa-Bianchi, N., \& Lugosi, G. (2006). \emph{Prediction, Learning,
  and Games}. Cambridge University Press.
\item
  SCX Theorem 1--3. \texttt{../theorems/}.
\item
  SCX Unified Theorem Document. \texttt{../THEOREMS\_UNIFIED.md}.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{End of README.md --- SCX Self-Evolution Theory}

\end{document}
