% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Situs (Physical Positional Encoding)
physical realism verificationandapplicable scenarioAnalysis}\label{situs-physical-positional-encoding-ux7269ux7406ux771fux5b9eux6027ux9a8cux8bc1ux4e0eux9002ux7528ux573aux666fux5206ux6790}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Document Status}: Honest Critique --- Unsparing\\
\textbf{Prerequisites}:
\texttt{ppe\_rigorous\_derivation.md} ( rigorous mathematicalDerivation), \texttt{multi\_head\_spring\_and\_positional\_encoding\_analysis.md} (CC
Audit Report)\\
\textbf{Methodology}: real physicaldata + mathematicalTheorem physical correspondence + one by onescenarioargumentation\\
\textbf{corecriterion}: Theorem 2.2.1 ---
\(\delta_s^{\text{PE}} > 0 \iff I(Y; P \mid S) > 0\), i.e.physicalLocationattodetermineStatusatomConditionslowermust携带regardingLabel额outerinformation

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{第 1
part: physical realismcheck}\label{ux7b2c-1-ux90e8ux5206ux7269ux7406ux771fux5b9eux6027ux68c0ux67e5}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.1
whitesequenceLocationencoding}\label{ux86cbux767dux8d28ux5e8fux5217ux4f4dux7f6eux7f16ux7801}

\subsubsection{1.1.1 mutualsameamino acid, is differentLocation: Lys-37 (活ness point )vs
Lys-289 (surface
loop)}\label{ux76f8ux540cux6c28ux57faux9178ux4e0dux540cux4f4dux7f6elys-37ux6d3bux6027ux4f4dux70b9vs-lys-289ux8868ux9762-loop}

\textbf{biological学事实 (Nonemay争议)}: 

same一 lysineresidueatwhiteis differentLocation具has 完fullis different功canRole: -
\textbf{活ness point Lys-37}: 典型catalyticresidue, 侧链 pKa canbias离溶液值 2--4
 single (From \textasciitilde10.5 降to
6.0--8.0), 参and酸碱catalyticorcommon价catalytic. 例such as: acetoacetate decarboxylase middle
Lys-115 pKa ≈ 6.0 {[}Bartlett et al., \emph{J. Mol. Biol.} 324, 105
(2002){]}; ribonuclease A middle Lys-41 pKa ≈ 8.8 {[}Raines, \emph{Chem.
Rev.} 98, 1045 (1998){]}. highdegree保守, tomutationextremelydegreesensitive (ΔΔG \textgreater{} 3
kcal/mol often see). - \textbf{surface loop Lys-289}: solventexposed, pKa
approaches溶液值 (10.4--10.6), 功canuppermany参andnon-特differentness静电mutualmutual作 use orwhite-whiteidentifiesinterface, 保守nesslow, tomutation容忍degreehigh (ΔΔG
通 often \textless{} 1 kcal/mol). 

\textbf{physicalJudge }: 

\[I(Y; P \mid S) > 0 \quad \text{——attodetermine"amino acid类型 = Lys"Conditionslower, residueLocation携带a large amount of功caninformation}\]

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1311}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3443}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3607}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1639}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
physical quantity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Lys-37 (活ness point )
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Lys-289 (surface loop)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Difference来源
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
侧链 pKa & 6.0--8.0 & 10.4--10.6 & local静电环境 (Buried vs solventexposed) \\
SASA (solventmay及surface积) & \textless{} 10 Å² & \textgreater{} 100 Å² & 3D
foldLocation \\
进化保守ness (BLOSUM62 Locationentropy) & \textless{} 0.2 bits & \textgreater{}
0.8 bits & 功can约束strongdegree \\
mutation ΔΔG & \textgreater{} 3 kcal/mol & \textless{} 1 kcal/mol &
foldstablenesscontribution \\
\end{longtable}

\textbf{BUT------keyrestricted}: 

Situs sinusoidal encodingencodingis\textbf{sequenceLocation} (1D residueNumber 37 vs
289), andnon-\textbf{3D spaceLocation} (活ness point 口袋 vs surface). thisis Situs
atwhiteshouldusing in the rootthisdefect: 

\begin{itemize}
\tightlist
\item
 sequenceLocation → 3D Locationmap\textbf{is NOT functional relationship}. same源whitemiddle, sequenceLocation 37
 in homolog A middlemayis活ness point, in homolog B middlemayis
 loop (insert/deleteleads tosequence移). 
\item
 fornothas 3D structurecase (mostwhiteprediction any 务), Situs
 onlycanusesequenceLocation. sequenceLocationis 3D 功canLocation\textbf{weak代理variable}. 
\item
 onlyhas when我们has 3D structureandwill Situs applied to 3D
 坐markencodingtime, upper述功canDifferenceonly thencanbyhas 效captures. 
\end{itemize}

\textbf{honestConclusion}: 1D sequencesinusoidal encodingtodistinguish Lys-37 活ness point vs Lys-289 surface
loop \textbf{has informationbutsignalweak}. sequenceLocationand 3D 功canLocationmutualclosenessis subject to fold
保守ness. \(I(Y; P_{\text{1D}} \mid S) > 0\) butis much smaller than
\(I(Y; P_{\text{3D}} \mid S)\). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.1.2 α-helix 3.6
residue/圈periodicity}\label{ux3b1-ux87baux65cbux7684-3.6-ux6b8bux57faux5708ux5468ux671fux6027}

\textbf{physical事实}: 

Pauling classical α-helix {[}Pauling et al., \emph{PNAS} 37, 205 (1951){]}: -
each圈 3.6 residue (Exact 值 = 3.6, or 18 residue/5 圈) - 螺距 = 5.4
Å (eachresiduerises 1.5 Å) - hydrogenbond模式: residue \(i\) C=O ··· H--N residue
\(i+4\) - this意味着residue \(i\) and \(i+4\) inspaceuppermutual邻 (\textasciitilde3.1
Å), and \(i\) and \(i+3\) (or \(i+5\))athelixrelative面

\textbf{sinusoidal encodingcanwhethercaptures? }

forsinusoidal encoding \(\sin(2\pi p / \lambda_j)\): - If 波long
\(\lambda_j = 3.6\): \(\sin(2\pi p/3.6)\) 确实具has 3.6 residueExact 周期 -
residue \(i\) and \(i+3.6\) obtainmutualsameencoding值 → residue \(i\) and \(i+4\) (距离
4, 周期差 0.4/3.6 ≈ 11\%)encodingmutual近butnotmutualsame

\textbf{butis------Standard Transformer encoding做notto}: 

Standard sinusoidal encoding {[}Vaswani et al., \emph{NeurIPS} 2017{]} 波long集combined: 
\[\lambda_j = 2\pi \cdot 10000^{2j/d}\]

when \(d = 512\) time, 波long范围约is \(2\pi\) to \(2\pi \cdot 10000\) (≈ 6.28
to ≈ 62832). for典型whitelongdegree \(L \sim 300\): -
mostshort波long (\(j = d/2-1 = 255\)): \(\lambda_{255} = 2\pi \cdot 10000^{510/512} \approx 2\pi \cdot 10000^{0.996} \approx 6.28 \times 9550 \approx 60,000\)
------is much larger than 3.6! - nohas any what 一 \(\lambda_j\) approaches 3.6

\textbf{physicalOptimal spectrum(Theorem 1.2.1)canwhether解决? }

according toTheorem 1.2.1, Optimal frequencyspectrumrequire知道physicalmutualcloselongdegree \(\xi\). for
α-helix, \(\xi \approx 3.6\) (helix周期). If 我们with \(\xi = 3.6\)
is 目markdesignencoding: -
\(\lambda_{\min} = 2\pi\xi \cdot \cot(\pi(2(d/2-1)+1)/2d) \approx 2\pi \cdot 3.6 \cdot \cot(\pi/2 - \pi/2d) \approx 22.6 \cdot (2d/\pi) \approx 14.4 d\)
- for \(d = 64\), \(\lambda_{\min} \approx 920\) ------stillis much larger than 3.6! 

\textbf{rootthisProblem}: sinusoidal encoding''separate辨率''is subject to \(d\). mustin \(L = 300\)
sequenceupperseparate辨出 3.6 residue周期, require
\(\lambda_{\min} \lesssim 3.6\), i.e.require
\(d \gtrsim 300/3.6 \times 2 = 167\) dimensional (each 周期tofew 2
 sampling point, Nyquist). for典型 \(d = 64\) or \(d = 128\), 3.6
residueperiodicity\textbf{theoreticallyNonemethodbycaptures}. 

\textbf{honestConclusion}: for实际useencodingdimension (\(d \leq 128\)), sinusoidal encoding\textbf{Nonemethod}may靠captures
α-helix 3.6
residue/圈periodicity. thisis NOT encodingformdefect, andisdimension-separate辨率basic权衡 (Corollary
1.2.1). such as果requireatsequenceLocationencodingmiddlecaptureshelixperiodicity, shouldexplicit式add一 
\(\lambda \approx 3.6\) frequencyseparate量, oroneusemaylearningperiodicityCharacteristics. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.1.3 intrinsichas None序区域 (IDR)middle Situs
退化}\label{ux56faux6709ux65e0ux5e8fux533aux57dfidrux4e2d-situs-ux7684ux9000ux5316}

\textbf{physical事实}: 

intrinsichas None序区域 (IDR)占truekernelwhite组 30--40\% {[}Dunker et al.,
\emph{FEBS J.} 272, 5129 (2005){]}. keyCharacteristics: -
at生理Conditionslower\textbf{notfoldis single一stable 3D structure} -
功canfrom\textbf{amino acidcomposes} (such ascharge模式、脯氨酸含量)andnon-Exact fold -
sequenceLocationand功canrelationship\textbf{extremelyweak}: IDR 功can (such as linker、entropic
spring、PTM point 集群)primarybycomposes决determine, Locationpermutationhas highdegree容忍degree - same一 IDR
insert/delete 5--20 residue通 often does not affect功can {[}Zibaee et al., \emph{Protein
Sci.} 16, 906 (2007){]}

\textbf{Situs in IDR upperlinesis }: 

\[I(Y; P_{\text{seq}} \mid S = \text{"IDR residue"}) \approx 0\]

byTheorem 2.2.1 and 2.3.1: 
\[\delta_s^{\text{PE}} \leq \sqrt{\frac{1}{2} I(Y; P \mid S)} + \delta_s^{\text{variance}} \approx 0 + O(1/\sqrt{M})\]

i.e.: \textbf{Situs 退化is dimensionis \(d\) addmethodnoise}. 

specifically, in IDR 区域 \(L\) continuousLocationupper: 
\[\text{PE}(p) = [\sin(2\pi p/\lambda_1), \cos(2\pi p/\lambda_1), \ldots]\]

this some encodingtoward量随 \(p\) smooth变化 (Lipschitz
continuousness), but它们携带变differentnessandLabel \(Y\) \textbf{Noneclose}. attrainingmiddle: -
dry净sample: \(\text{PE}(p)\) addandLabelNoneclosewave →
may slightly micro-reduceexpertconsistentness →
\(p_{\text{clean}}^{\text{PPE}} \geq p_{\text{clean}}\) -
noisesample: \(\text{PE}(p)\) same type does not providedistinguishnoisecanforce →
\(p_{\text{noisy}}^{\text{PPE}} \approx p_{\text{noisy}}\) -
result: \(\delta_s^{\text{PE}} \leq 0\) (负or零)

\textbf{honestConclusion}: IDR is Situs
\textbf{mostbadscenario}------Locationencoding退化is None害butNone use dimension膨胀. 它will notmain动breaking model (Lipschitz
continuousnessguarantee \(h_i\) stillhoweversmooth滑), but消耗 \(d\)
dimensionalcomputeandexistence储, andmaythroughgradientnoise slightly micro-reduceconvergencespeed. such as果data集middle IDR
proportionhigh (\textgreater30\% such astruekernelwhite), Situs 净benefit\textbf{mayis 负}. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.2 material 3D
structureLocationencoding}\label{ux6750ux6599-3d-ux7ed3ux6784ux4f4dux7f6eux7f16ux7801}

\subsubsection{1.2.1 V\_N 空formation energy: grain boundary vs 体mutual vs
surface}\label{v_n-ux7a7aux4f4dux5f62ux6210ux80fdux6676ux754c-vs-ux4f53ux76f8-vs-ux8868ux9762}

\textbf{physical事实 (DFT computedata)}: 

AlN middlenitrogen空 \(V_N\) formation energystrong烈-dependenceatspaceLocation {[}Freysoldt et al.,
\emph{Rev.~Mod. Phys.} 86, 253 (2014); Stampfl \& Van de Walle,
\emph{Phys. Rev.~B} 65, 155212 (2002){]}: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.1714}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.5429}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2857}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
formation energy \(E_f\) (eV)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
physicalReason
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{体mutual} (bulk, 配数 4) & 3.2--4.0 & Complete四面体配, 断open 4 
Al--N bond \\
\textbf{surface} ((0001) surface, 配数 3) & 1.8--2.5 &
surface弛豫releasesshould变can, 配数reduce \\
\textbf{grain boundary} (Σ7 to称倾斜grain boundary) & 1.0--1.8 & grain boundarythis征shouldforce field + manyremainder体积
+ 悬挂bond预existencein \\
\end{longtable}

Difference幅degree: \(\Delta E_f \approx 1.5\)--\(2.5\) eV, 远super-\(k_B T\) (室温lower
≈ 0.025 eV), at\textbf{ any what }温degreelowerall具has 决determinenessimpact. 

\textbf{physicalbecause果链} (notmaysimplifies): 

\[\text{3D Location} \xrightarrow{\text{决determine}} \text{local配环境} \xrightarrow{\text{决determine}} \text{defectformation energy}\]

thisis一 清晰because果箭head------Locationisself-variable, formation energyisbecausevariable. Locationencodingcapturesisthis because果链\textbf{Reason侧}. 

\[I(Y = E_f; P_{\text{3D}} \mid S = V_N) > 0 \quad \text{——information-theoretic senseupperrigorouscurrentlyand量large}\]

according to Fano lower bound(Theorem 2.4.1), formany类separate类 (is differentdefectcharge态): 
\[\delta_s^{\text{PE}} \geq \frac{2 \cdot I(Y; P \mid S) - \log 2}{\log |\mathcal{Y}|}\]

take \(I(Y; P \mid S) \approx H(Y|S) \approx \log 3 \approx 1.6\) bits (3
kindformation energy区between: high/middle/low), \(\log |\mathcal{Y}| \approx \log 3\): 
\[\delta_s^{\text{PE}} \gtrsim 0.5 \text{–}0.8\]

thisis一 \textbf{may观Detection Marginincrease益}, toshould Theorem 1 exponential界explicit著tightens. 

\textbf{honestConclusion}: formaterialdefectdetection, 3D Situs
is\textbf{physicallymosthas valueapplicationscenario}. Locationdirectbecause果决determineLabel. thisis Situs
''home turf''. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.2.2 3D
rotationencodingphysicalto称nesscheck}\label{d-ux65cbux8f6cux7f16ux7801ux7684ux7269ux7406ux5bf9ux79f0ux6027ux68c0ux67e5}

\textbf{Situs rotationencodingDefinition} (Definition 1.3.1): 
\[\text{PE}_{\text{rot}}(\mathbf{p}) = \mathbf{R}_x(\alpha x) \cdot \mathbf{R}_y(\beta y) \cdot \mathbf{R}_z(\gamma z) \cdot \mathbf{e}_0\]

\textbf{check 1: translation invariance} ✓

inner积: 
\[\langle \text{PE}(\mathbf{p}), \text{PE}(\mathbf{q}) \rangle = \langle \mathbf{e}_0, \mathbf{R}(\mathbf{q} - \mathbf{p}) \mathbf{e}_0 \rangle = \cos(\alpha\Delta x) + \cos(\beta\Delta y) + \cos(\gamma\Delta z)\]

only-dependenceat
\(\Delta\mathbf{p} = \mathbf{q} - \mathbf{p}\). \textbf{translation invariancerigorousholds}. thissatisfiescrystal学basicrequire------完美crystalmiddle, two Locationequivalentif and only if它们relative移islatticetoward量. 

\textbf{check 2: rotation等变ness} ✗ \textbf{does not hold}

considerglobalrotation \(R_{\text{phys}} \in SO(3)\)
acts on坐mark: \(\mathbf{p} \to R_{\text{phys}} \mathbf{p}\). 

for \$R\_\{\text{phys}\} = \$ 绕 z 轴rotation 90° (will \((x, y, z)\) 变is 
\((-y, x, z)\)): 
\[\text{PE}_{\text{rot}}(R_{\text{phys}}\mathbf{p}) = \mathbf{R}_x(-\alpha y) \mathbf{R}_y(\beta x) \mathbf{R}_z(\gamma z) \mathbf{e}_0\]

and
\(\text{PE}_{\text{rot}}(\mathbf{p}) = \mathbf{R}_x(\alpha x) \mathbf{R}_y(\beta y) \mathbf{R}_z(\gamma z) \mathbf{e}_0\). 

thistwo encoding'sbetween\textbf{does not exist}一 and \(\mathbf{p}\) Noneclosetransform \(T\) makes
\(T \cdot \text{PE}(\mathbf{p}) = \text{PE}(R_{\text{phys}}\mathbf{p})\). Reason: encodingis 三 坐mark轴separate配\textbf{disjointdimensionto}------\((0,1)\)
toshould \(x\), \((2,3)\) toshould \(y\), \((4,5)\) toshould \(z\). physicalrotationmixedcombined
\(x\) and \(y\) separate量, butencodingspacemiddlenothas toshould''mixedcombined''mechanism. 

\textbf{breaking back果}: 

atwithlowerscenariomiddlerotation等变nessmissingisFatal: 1.
\textbf{many晶/grain boundarysystem}: is different晶粒taketowardis different, mutualsame类型defectencodingbecausetaketowardis differentandis different
→ \(h_i\) introduceandLabelNoneclose变differentness 2.
\textbf{separate子dynamics轨迹}: separate子wholerotationleads to has atomencoding变化 →
modelmay学to''rotationStatus''andnon-''physicalStatus'' 3. \textbf{surface slab
compute}: same一surfaceis different切面 (such as AlN (0001) vs (11̄00))具has is different坐mark轴taketoward

\textbf{maywith修复吗? }

theoreticallymaywith. 替代scheme: -
\textbf{球谐functionencoding}: \(\text{PE}(\mathbf{p}) = [Y_{\ell m}(\theta, \phi)]\), 具has rigorous
\(SO(3)\) rotation等变ness - \textbf{E(3) 等变网络}: make using Tensor Field
Networks {[}Thomas et al., \emph{NeurIPS} 2018{]} or SE(3)-Transformers
{[}Fuchs et al., \emph{NeurIPS} 2020{]}, 天however具备translation+rotation等变ness -
\textbf{cost}: 球谐functionencodingdimension随angle动量 \(L\) growthis 
\((L+1)^2\), \(L=3\) i.e.need 16 dimensional (vs 6 dimensionalrotationencoding), and Lipschitz
linesis morecomplexity

\textbf{honestConclusion}: Situs 3D
rotationencoding\textbf{only具has translation invariance, not具rotation等变ness}. formostmaterialapplication (requireat理 any 意taketowardcrystal), thisis\textbf{严heavydefect}. butat受控scenario (single晶、fixed taketoward
slab、taketowardalreadyto齐data集)middle, thisnotconstitutesProblem. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1.3 atomtotal数 N: 32 vs 256 supercell DFT
accuracy}\label{ux539fux5b50ux603bux6570-n32-vs-256-ux8d85ux80deux7684-dft-ux7cbeux5ea6}

\subsubsection{1.3.1
systemnessDifferencephysical起源}\label{ux7cfbux7edfux6027ux5deeux5f02ux7684ux7269ux7406ux8d77ux6e90}

\textbf{Makov-Payne has 限尺寸markdegree} {[}Makov \& Payne, \emph{Phys. Rev.~B}
51, 4014 (1995){]}: 

带电defectformation energy随supercell尺寸 \(L\) markdegreelinesis: 
\[E_f(L) = E_f(\infty) + \frac{\alpha q^2}{2\varepsilon L} + \frac{2\pi q Q}{3\varepsilon L^3} + O(L^{-5})\]

where \(q\) isdefectcharge, \(\varepsilon\) is介电 often 数, \(Q\) is弹ness四extremely矩. 

\textbf{concrete数值estimate (\(V_N^{+1}\) in AlN
middle, \(\varepsilon \approx 9\))}: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1538}}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1231}}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.2308}}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1692}}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.2000}}
 >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1231}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
supercell尺寸
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
atom数
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
periodicity图像between距
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(1/L\) correction
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(1/L^3\) correction
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
totalcorrection
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
2×2×2 & 32 & \textasciitilde6.2 Å & \textasciitilde0.35 eV &
\textasciitilde0.12 eV & \textbf{\textasciitilde0.5 eV} \\
2×2×3 & 48 & \textasciitilde6.2--9.3 Å & \textasciitilde0.25 eV &
\textasciitilde0.05 eV & \textbf{\textasciitilde0.3 eV} \\
3×3×3 & 108 & \textasciitilde9.3 Å & \textasciitilde0.15 eV &
\textasciitilde0.02 eV & \textbf{\textasciitilde0.17 eV} \\
4×4×4 & 256 & \textasciitilde12.4 Å & \textasciitilde0.08 eV &
\textasciitilde0.005 eV & \textbf{\textasciitilde0.09 eV} \\
\end{longtable}

\textbf{physicalroot源} (三 independent has 限尺寸effect, uniformnotmay忽 slightly ): 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \textbf{静电像chargemutualmutual作 use } (\(1/L\) main导项): 

 \begin{itemize}
 \tightlist
 \item
 带电defectanditsperiodicity镜像'sbetween Coulomb mutualmutual作 use 
 \item
 32 atomsupercellmiddlemutual邻镜像only隔 6.2 Å → mutualmutual作 use can \textasciitilde0.3--0.5
 eV
 \item
 physicallythisis人is introduce------true孤立defectnothas periodicity镜像
 \end{itemize}
\item
 \textbf{弹ness像mutualmutual作 use } (\(1/L^3\)): 

 \begin{itemize}
 \tightlist
 \item
 defectcauseslocalshould变场anditsperiodicity镜像mutualmutual作 use 
 \item
 to \(V_N\) (removeatomproduces拉伸should变)thiseffectnotmay忽 slightly 
 \end{itemize}
\item
 \textbf{defect波functionheavy叠} (exponential decay \(\propto e^{-L/\xi}\)): 

 \begin{itemize}
 \tightlist
 \item
 todeepcan级defect (such as \(V_N\) in AlN middleproduces \textasciitilde1.5 eV
 deep施maincan级), 波function局域化half径 \(\xi \approx 2\)--\(3\) Å
 \item
 32 atomsupercellmiddlemutual邻波function尾heavy叠notmay忽 slightly →
 defectcan级separate裂成can带 (人造色散)
 \end{itemize}
\end{enumerate}

\textbf{thisand Situs \(N\) encodinghas 什么relationship? }

Situs 第三域 (Definition 1.1.3)willtotalatom数 \(N\)
作is mark量Locationencoding入. Problemis: 

\begin{itemize}
\tightlist
\item
 \(N=32\) and \(N=256\)
 data\textbf{is NOT same一physical systemis different''Location''}------它们is\textbf{is differentaccuracy级别}approximate
\item
 \(N\) encodingwill''computeaccuracy''and''physicalLabel''mixed淆. \(N=32\) \(E_f\)
 bias差iscompute artifact, is NOT physicalsignal
\item
 such as果training集mixedcombined 32 and 256 atomsupercelldata, \(N\)
 encodingmay学to''\(N=32 \to\) high \(E_f\)``虚falsemutualcloseness
\end{itemize}

\textbf{physicallycorrectat理}: - to \(E_f\) shouldusing Freysoldt-Neugebauer-Van
de Walle (FNV) has 限尺寸correction {[}Freysoldt et al., \emph{Phys. Rev.~Lett.}
102, 016402 (2009){]}, eliminate \(N\) systemnessbias差 -
\textbf{'sback}againconsideriswhether make using \(N\) encoding------thistime \(N\)
only then反映truesystem尺寸effect (such asquantum限域), andnon-compute artifact

\textbf{honestConclusion}: such as果trainingdatanotafterhas 限尺寸correction, \(N\)
encoding学tois\textbf{compute artifact andnon-physical}. atcorrection'sback, \(N\)
携带 real physicalinformation (quantum限域、振动模式数)fordefectphysical来说relativelysmall (\(\Delta E_f\)
 \(N\)-dependence \textless{} 0.05 eV for
\(N > 100\)). formostmaterial any 务, \(N\) encoding \(\delta_s^{\text{PE}}\)
\textbf{approaches零}------thisis一 physically动机weak、practiceupper危险 (easy学to
artifact)encodingdimension. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{第 2 part: Situs
at什么caselowerhas 害? }\label{ux7b2c-2-ux90e8ux5206situs-ux5728ux4ec0ux4e48ux60c5ux51b5ux4e0bux6709ux5bb3}

Honest Critique: Situs
is NOT 一 ''None害butmayNone use ''组件. at特determineConditionslower它is\textbf{main动has 害}. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.1 虚falsemutualcloseness (Spurious
Correlation)}\label{ux865aux5047ux76f8ux5173ux6027spurious-correlation}

\textbf{mechanism}: 

whenphysicalLocation \(P\) andLabel \(Y\)
attrainingdatamiddle具has 统计学mutualclose, butshouldmutualclose\textbf{not反映causal relationship}time: 

\[\text{Corr}_{\text{train}}(P, Y) \neq 0, \quad \text{but } P \perp\!\!\!\perp Y \text{ (because果meaningupperindependent)}\]

byat Situs will \(P\) encodingis \(\text{PE}(P)\) and Note入representation, model (expert
\(E_m\))willlearning利 use this 虚falsemutualclose. 

\textbf{concreteexample}: 

\textbf{whitedata集bias差}: - PDB
middle解析whitestructureexistenceat''易at结晶''bias差------倾towardatmorestable、more紧凑fold -
such as果training集middle has 酶活ness point 恰goodat结晶structuremiddleis located atwhitecore (low B-factor
区域 → lowLocationindexes附近by解析得more清楚) - Situs
学to: residueLocation靠front (\(p < 100\))→ 活ness point - test集middle出现 C
端活ness point white → systemness误判

\textbf{DFT computeparameterbias差}: - Materials Project
 in the computeuseis differentcomputeparameter (赝势、k
 point densedegree), andthis some parameterandmaterial类型has systemnessclose联 - such as果 Situs atom数encoding
\(N\) 学to: \(N=32\) (典型gold属compute)→
lowformation energy, \(N=256\) (典型half导volume defectcompute)→ highformation energy -
thiskind''知识''iscompute惯例 artifact, is NOT physical

\textbf{虚falsemutualclosenessdetection}: \(\delta_s^{\text{PE}}\) intraining集upper
\textgreater{} 0, butin i.i.d. verify集upper → 0 or \textless{} 0. Theorem
2' (correction Theorem 2)upper boundatdistributionouterdataupper\textbf{反andmore松}------becauseis 
\(\varepsilon_{\text{PE}}\) intrainingdistributionupperbylow估. 

\textbf{honestConclusion}: 虚falsemutualclosenessis Situs
\textbf{maximum实际risk}. physicist知道''position matters''buttrainingdata in the 
position-label
mutualclosehas manyfewfrombecause果、has manyfewfromdata收集bias差, mustitem by item案Analysis. \textbf{nothas 捷径}. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.2
manyheavycommonlinear (Multicollinearity)}\label{ux591aux91cdux5171ux7ebfux6027multicollinearity}

\textbf{mechanism}: 

whenStatusrepresentation \(\phi(s_i)\) \textbf{already经implicit式encodingLocationinformation}time, addmethodNote入
\(\text{PE}(p_i)\) produces冗remainderdimension: 

\[h_i = \underbrace{\phi(s_i)}_{\text{already含implicit式Locationinformation}} + \underbrace{\text{PE}(p_i)}_{\text{explicit式Locationinformation}}\]

thisleads toencodingspace in the has 效dimension膨胀, 伴随: 1.
\textbf{expertmodelgradientvarianceincreaselarge}: each expertrequireFrom 冗remainderdimensionmiddle提takesignal 2.
\textbf{has 效expertmany type ness \(\rho_{\text{eff}}\)
falls}: when has expertatsame一冗remainderspacemiddle操作, 它们输出mutualclosenessincreaselarge

\textbf{concreteexample}: 

\textbf{图neural network (GNN) at理crystal}: - GNN
alreadythrough消息传递聚combined邻域information, containsimplicit式Location/距离encoding -
such as果againtoeach 节 point add Situs 3D
rotationencoding, mutualwhenatadd\textbf{already经by图卷积layercapturesinformation} -
result: \(I(Y; \text{PE}(P) \mid \phi(S)) \approx 0\) (todetermine GNN
Characteristicsback, LocationNone额outerinformation) - but PE 额outerdimensionincreasesoptimize难degree, 等效at
\(\delta_s^{\text{variance}}\) increaselarge

\textbf{white语言model (pLM) at理sequence}: - ESM-2 {[}Lin et al.,
\emph{Science} 379, 1123 (2023){]} 等 pLM attention
mechanism天howevercontainsLocationinformation (attention bias) - to ESM embedding againadd Situs
sinusoidal encoding = addalreadyis attention at理冗remaindersignal -
thiskind''doubleheavyencoding''atstatisticallyintroducecommonlinear, butnotincreaseshas 效information

\textbf{determine量impact}: 

setup原始Characteristicsspacehas 效秩is 
\(r_{\text{eff}} = \text{rank}(\Sigma_\phi)\). add PE post-: 
\[\text{rank}(\Sigma_h) \leq r_{\text{eff}} + \min(d_{\text{pe}}, N)\]

butsuch as果 \(\text{span}(\text{PE}(P)) \subseteq \text{span}(\phi(S))\) (PE
in \(\phi\) 列spacemiddle), then \(\text{rank}(\Sigma_h) = r_{\text{eff}}\)
------dimensionincreasesbuthas 效秩not变. thisis\textbf{pure浪费}. 

\textbf{honestConclusion}: atalreadyhas powerfulstructureencodingmodel (GNN、Transformer、pLM)'supper叠add
Situs, largeprobabilityis\textbf{零increase益 + micro-损害}. Situs
mosthas usingtimeis作is \textbf{onlyhas }Locationinformation源 (when \(\phi(s_i)\)
完fullnothas spaceinformationtime). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2.3 physicalto称nessbreaking (Physical Symmetry
Breaking)}\label{ux7269ux7406ux5bf9ux79f0ux6027ux7834ux574fphysical-symmetry-breaking}

\textbf{mechanism}: 

Situs addmethodNote入 \(h_i = \phi(s_i) + \text{PE}(p_i)\)
\textbf{刻意breaking }physical systemsome some to称ness. whenthis some to称nesstoprediction any 务importanttime, thisishas 害. 

\textbf{三kindbreaking to称ness}: 

\subsubsection{2.3.1
translational symmetrybreaking (to体mutualnesspredictionisbad)}\label{ux5e73ux79fbux5bf9ux79f0ux6027ux7834ux574fux5bf9ux4f53ux76f8ux6027ux8d28ux9884ux6d4bux662fux574fux7684}

完美crystalmiddle, has primitive cellequivalent------体mutualness (带隙、弹ness often 数、声子frequency)should具has \textbf{完fulltranslation invariance}. 

 make using Situs post-: - two equivalentprimitive cell in the atomobtainis different \(h_i\) -
modelmust额outerlearning''忽 slightly Location''------increasesis unnecessarylearning负担 -
more糟: such as果training集has 限, modelmay学to''带隙-dependenceatatomindexes''虚false规律

\textbf{example}: prediction AlN 体mutual带隙 (\textasciitilde6.2 eV, {[}Yan et al.,
\emph{Phys. Rev.~B} 93, 014110 (2016){]}) - \(N=32\) supercellmiddleeach N
atom Situs encodingis different -
such as果model use this some encodingpredictionsame一 带隙Label, 它mustatinnerpartlearning''this some encodingshouldby忽 slightly ''
- tosmalldata集, thismayleads tooverfittingto特determineencoding值

\textbf{Judge }: such as果 any 务Label具has translation invariance (体mutualness), Situs
\textbf{notshoulduse}. such as果 any 务Labelnot具has translation invariance (defectformation energy、surface energy、局域density of states), Situs
\textbf{shouldwhenuse}. 

\subsubsection{2.3.2 rotational symmetrybreaking (3D rotationencodinghard伤------alreadyin 1.2.2
Analysis)}\label{ux65cbux8f6cux5bf9ux79f0ux6027ux7834ux574f3d-ux65cbux8f6cux7f16ux7801ux7684ux786cux4f24ux5df2ux5728-1.2.2-ux5206ux6790}

\subsubsection{2.3.3
permutationto称nessbreaking (to集combined型predictionisbad)}\label{ux6392ux5217ux5bf9ux79f0ux6027ux7834ux574fux5bf9ux96c6ux5408ux578bux9884ux6d4bux662fux574fux7684}

atomsystemthisupperis\textbf{粒子集combined}------atomNumber/顺序is人is, notshouldimpactphysicalprediction. 

Situs -dependence''第 \(i\) atomLocation \(p_i\)``, where \(i\)
isatomindexes. such as果dataadd载timeatomsortundergoes change, \(h_i\)
willis different------i.e. make physical system完fullmutualsame. 

\textbf{thisatwithlowerscenariomiddleresults inProblem}: - DFT
computemiddleatom列表sortNonephysical meaning (VASP/PWscf
输出 in the atom顺序is输入顺序function) -
separate子dynamics轨迹middle, atommay跨过periodicityboundary, changesitsat数组 in the Location -
is differentdata源atomsort约determineis different

\textbf{缓解措施}: use\textbf{等变}Locationencoding (such asbased on距离matrixoratom环境encoding), orrequiredatapreprocessingmiddleguaranteeconsistentatomsort (such as按坐marksort------butthisagainintroducesort人is ness). 

\textbf{honestConclusion}: Situs
breaking permutationto称nessis一 \textbf{实际工程Problem}andnon-theorydefect------maywiththroughconsistentpreprocessing缓解, butNonemethodroot除. such as果atomsortatis differentsamplebetweenis inconsistent, Situs
introducenoisemay淹notsignal. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{第 3
part: applicable scenarioAnalysis表}\label{ux7b2c-3-ux90e8ux5206ux9002ux7528ux573aux666fux5206ux6790ux8868}

to 6 concretescenario, evaluation \(\delta_s^{\text{PE}}\)
Symbol/Concept (currently/零/负)、order of magnitude (large/middle/small/零)and物theory证. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.0577}}
 >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1154}}
 >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.5192}}
 >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1923}}
 >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1154}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\#
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\delta_s^{\text{PE}}\) Symbol/Concept
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
order of magnitudeestimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
理by
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{1} & \textbf{AlN defectdetection: \(V_N\) 空, 534 帧} & \textbf{currently} &
\textbf{large} & 3D Locationbecause果决determinedefectformation energy (§1.2.1). 534
帧atmaterialdefectdatamiddlebelongs tomiddle等规模, 足withlearningLocation-energyrelationship. Locationinformationnot冗remainder (\(\phi(s_i)\)
is localchemicalCharacteristics, not含globalspaceupperlower文). 3D
rotationencodingtranslation invarianceatthisscenariomiddle恰goodcombined适 (defectmiddle心to齐backrelativeLocationhas meaning). \(I(Y; P \mid S)\)
estimate: 1.5--2.5 bits (throughformation energy三separate
binning), \(\delta_s^{\text{PE}} \gtrsim 0.4\). \textbf{butis}: such as果 534
帧containsmany 晶粒taketoward, rotation等变nessmissingwillintroducenoise (see §1.2.2
警告). Recommendation first checktaketowarddistribution, necessarytime做taketowardto齐preprocessing. \\
\textbf{2} & \textbf{white功canprediction: 酶活ness point separate类} & \textbf{currently} (weak)
& \textbf{middlesmall} & sequenceLocationand功can point has 统计mutualcloseness (N 端/C
端biasgood、structure域Location), buteffect量weak. \(I(Y; P_{\text{1D}} \mid S) > 0\)
holdsbut数值has 限 (\textless{} 0.3
bits, becauseis the majority of功caninformationalreadyatresidue类型andlocalsequenceupperlower文middle). \textbf{keyrestricted}: Situs
 make using 1D sequenceLocationandnon-3D spaceLocation------and活ness point 功canby 3D
Location决determine. sequence→3D mapbecause fold
Differenceandnon-single值. \textbf{Conclusion}: has micro-weakcurrently面contribution, but远notsuch asdirect make using 3D
structureinformation (such ashas AF2 predictionstructure). If alreadyhas pLM
embedding (含Locationinformation), benefitapproaches零 (commonlinearProblem, §2.2). \\
\textbf{3} & \textbf{Drug-target to接评separate: DrugBank 6784 } &
\textbf{currently} (has Conditions) & \textbf{middlesmall} & Conditions: make using 3D to接姿态坐mark (andnon-
1D sequenceLocation). 结combined亲andforce (\(K_d\), \(K_i\), \(\Delta G\))is 3D
姿态function------口袋Location、hydrogenbond several what 、sparse水接触面积allbyspaceLocation决determine. \(I(Y; P_{\text{3D}} \mid S)\)
\textgreater{} 0. \textbf{butis}: 6784 dataforlearningcomplexity 3D
Location-亲andforcemapbiasfew (特别isconsidertoto接姿态巨large构象space). Situs 
Lipschitz smooth滑ness(Theorem 
2.5.1)atthisishas 益------它currentlythen化邻近spaceLocationprediction. has 过degree-dependence结combined point Locationrisk (separate子骨架structurealsohas contribution). \textbf{Recommendation}: onlyatto接姿态qualityhigh (RMSD
\textless{} 2.0 Å to experimental)subsetupper make using 3D Situs. \\
\textbf{4} & \textbf{purechemicalcomposesseparate类: Nonespacestructure} & \textbf{零} (or slightly 负)
& \textbf{零} & \(\mathcal{P}\)
notDefinition. nothas physicalLocationinformationmayencoding. stronglinesuse (such as人工赋予虚拟indexes)→
encodingisdataadd载顺序 artifact. \(I(Y; P \mid S) \equiv 0\). byTheorem
2.2.1, \(\delta_s^{\text{PE}} = 0\). byTheorem
2.5.1, encoding随虚拟Locationsmooth变化------butthis变化not含 any what Labelinformation. uniquenon-零effectfrom
\(\delta_s^{\text{variance}}\) (has 限expertestimatenoise), 表现is 
\(O(1/\sqrt{M})\) \textbf{净负contribution}. \textbf{thisis Situs
绝to禁区}------add一determinemore差. \\
\textbf{5} & \textbf{randompermutationatom构型: None周期structure} & \textbf{负} &
\textbf{middle} & Locationexistenceat (has 3D
坐mark), butLocationandLabel'sbetween\textbf{nothas physical规律ness}------构型israndom. \(I(Y; P \mid S) = 0\) (randompermutationbreaking has Location-Labelinformation theoryconnection). buttraining集middlehas 限random涨落mayproduces\textbf{偶however}Location-Labelmutualclose, modelwill学tothis some 虚false模式
→ testtimegeneralization差. by Theorem 2' (not完美 PPE
lower bound), \(\varepsilon_{\text{PE}}\)
attrainingdistributionupperbylow估, leads toupper bound过松 (过degreeself-信). \(\delta_s^{\text{PE}}\)
attraining集upper \textgreater{} 0 (虚false), attest集upper \textless{}
0 (true). \textbf{thisis Situs
most危险scenario}------surfacebenefitis统计false象. usefrontmust检验 \(I(Y; P \mid S)\)
统计explicit著ness (through permutation test). \\
\textbf{6} & \textbf{carbon纳 meter 管手nessseparate类: (n, m) indexes} & \textbf{currently} &
\textbf{large} & 手ness \((n,m)\) 完fullby 3D
atompermutationglobal several what 决determine------thisisLocationencoding理想scenario. has carbonatom具has mutualsamelocal
\(sp^2\) chemical环境 (\(\phi(s_i) \approx\) often 数), becausethis
\(I(Y; S) \approx 0\) (purechemicalcomposesNonecannot distinguish手ness)and
\(I(Y; P) = H(Y) = \log_2 K\) (\(K\)
is 手nessCategory数). \(I(Y; P \mid S) \approx H(Y)\)------\textbf{ several 乎allinformationfromLocation}. Situs
成is information\textbf{primary}andnon-assists来源. currently弦/rotationencodingthrough捕捉atombetweenrelativeLocation模式, maywithencodinghelix缠绕periodicityand管径information. \textbf{butis}: requireNote意periodicityboundaryConditionsencodingconsistentness------CNT
at圆周directionhas periodicity, Situs 3D
rotationencodingwill notself-动沿圆周闭combined. forsmall直径管 (n+m \textless{}
15), 周longeffectbrightexplicit, require特殊at理. \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{第 4
part: each scenario一句话Recommendation}\label{ux7b2c-4-ux90e8ux5206ux6bcfux4e2aux573aux666fux4e00ux53e5ux8bddux5efaux8bae}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.0682}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.1364}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.7955}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\#
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Recommendation (addnotadd Situs + 一句话理by)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{1} & AlN \(V_N\) defect 534 帧 & \textbf{add}------3D
Locationisformation energydirectphysicalReason, 534
帧sufficientlearningspace模式, front提istaketowardalreadyto齐ormodelcan容忍rotationnon-等变ness. \\
\textbf{2} & 酶活ness point separate类 & \textbf{notadd} (or改using 3D Situs)------1D
sequenceLocationis 3D 功canLocationweak代理, 尤itswhenalreadyhas pLM
embedding (含Note意forceLocationbias差)time, 1D Situs 沦is commonlinear冗remainder. such ashas AlphaFold
predictionstructure, 改using 3D Situs. \\
\textbf{3} & DrugBank to接评separate & \textbf{has Conditionsadd}------onlywhenusehighquality
3D to接姿态 (RMSD \textless{} 2 Å)timeadd 3D Situs; 6784 datato 3D
encodingdimensionbiassmall, must配combinedstrongcurrentlythen化ordimensionality reductionencoding. \\
\textbf{4} & purecomposesseparate类Nonestructure & \textbf{绝tonotadd}------nothas Locationthennothas 
Situs; any what 人is 赋予伪Locationallis统计noise, mayproduces虚falsemutualclose. \\
\textbf{5} & randomatom构型 &
\textbf{notadd}------randomLocationandLabelinformation theoryconnectionis 零, Situs
atthisis虚falsemutualcloseness滋生器; If truemustuse, must first 做 permutation test verify
\(I(Y;P \mid S)\) 统计explicit著ness. \\
\textbf{6} & CNT 手ness (n,m) separate类 & \textbf{add}------手nessbyglobal 3D
 several what unique确determine, Situs
From onlyhas Locationinformationmiddle提takeCharacteristics (carbonatomlocalchemical环境完fullnot携带手nessinformation), thisis
Situs ''教科书级''applicationscenario. \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Summary: Situs
physical适 use boundary}\label{ux603bux7ed3situs-ux7684ux7269ux7406ux9002ux7528ux8fb9ux754c}

\subsection{它 what time工作 (按信心sort)}\label{ux5b83ux4f55ux65f6ux5de5ux4f5cux6309ux4fe1ux5fc3ux6392ux5e8f}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \textbf{3D Locationbecause果决determineLabel + None冗remainderstructureencoding} (CNT
 手ness、defectspaceseparate辨)→ \(\delta_s^{\text{PE}}\) largeandcurrently. thisis Situs
 design目markscenario. 
\item
 \textbf{3D Locationbecause果决determineLabel +
 has 冗remainderbuthas 额outerstructure} (to接评separate、materialdefect)→ \(\delta_s^{\text{PE}}\)
 middleandcurrently, butdimensionselectandcurrentlythen化key. 
\item
 \textbf{1D sequenceLocationweakly correlatedLabel + Nonestructureencoding} (简singlewhite功canprediction, None
 pLM)→ \(\delta_s^{\text{PE}}\) smallandcurrently, micro-weakbenefit. 
\end{enumerate}

\subsection{它 what timeNone use (addequalswhiteadd)}\label{ux5b83ux4f55ux65f6ux65e0ux7528ux52a0ux4e86ux7b49ux4e8eux767dux52a0}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
 \textbf{Locationnot携带todetermineStatus额outerLabelinformation} (IDR、purecomposesseparate类、体mutualness)→
 \(\delta_s^{\text{PE}} \approx 0\), puredimension浪费. 
\item
 \textbf{alreadyhas strongforcestructureencoding器} (GNN incrystalupper、pLM inwhiteupper)→
 commonlinear, \(\delta_s^{\text{PE}} \approx 0\) buthas micro-small负effect. 
\end{enumerate}

\subsection{它 what timehas 害 (addmore差)}\label{ux5b83ux4f55ux65f6ux6709ux5bb3ux52a0ux4e86ux66f4ux5dee}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
 \textbf{伪LocationorrandomLocation +
 has 限sample} (Nonestructuredata虚拟indexes、random构型)→ \(\delta_s^{\text{PE}}\)
 training \textgreater{} 0 buttest \textless{} 0, most危险虚falsemutualclose. 
\item
 \textbf{Label具has translation/rotational symmetrybut any 务not尊heavyshouldto称ness} (体mutualness + Situs
 encodingbreakto称ness)→ introduceandphysical矛盾归纳bias差. 
\item
 \textbf{notcorrectioncompute artifact
 isencodingis ''Location''} (mixedcombinedis differentsupercell尺寸notcorrection DFT data)→ 学tocompute
 artifact andnon-physical. 
\end{enumerate}

\subsection{theory洞察and工程现实鸿沟}\label{ux7406ux8bbaux6d1eux5bdfux4e0eux5de5ux7a0bux73b0ux5b9eux7684ux9e3fux6c9f}

Theorem 2.2.1 to出Sufficient Condition \(I(Y; P \mid S) > 0\)
is\textbf{necessaryandExact }, butatpracticemiddle: - estimate \(I(Y; P \mid S)\)
this身requirea large amount ofsample (KSG estimate器convergencedimension灾难, Theorem 3.3.1) - i.e. make 
\(I(Y; P \mid S) > 0\), Theorem 2.3.1 upper boundmaybecauseencodingnot完美degree
\(\varepsilon_{\text{PE}}\) andextremelydegree松弛 - Theorem 3' reveal''learnable PPE
maydistinguishnoiseandDifficultysample''atpracticemiddle\textbf{通 often not发生}------becauseis physical学 often use 
PE isfixed, andlearnable PE signaltooweakNonemethodmay靠breaking indistinguishability

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Analysis day 期: 2026-06-29}\\
\emph{physicaldata引 use: Bartlett (2002), Creighton (1993), Dunker (2005),
Freysoldt (2014), Makov \& Payne (1995), Pauling (1951), Stampfl \& Van
de Walle (2002), Van de Walle \& Neugebauer (2004), Yan (2016), 等}\\
\emph{Methodology声bright: has \(\delta_s^{\text{PE}}\)
estimatefromphysical推演andnon-数值experiment------mustmust DFT
computeandwhitestructureAnalysis来verify/校准this some estimate. }

\end{document}
