% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Spring
self-evolution convergenceAnalysis hostile review report}\label{spring-ux81eaux8fdbux5316ux6536ux655bux5206ux6790ux7684ux654cux5bf9ux5ba1ux7a3fux62a5ux544a}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{review orientation}: non-convex optimizationdirectionstrict reviewer\\
\textbf{review date}: 2026-06-29\\
\textbf{review target}:
\texttt{theory/self\_evolution/spring\_convergence\_analysis.md}\\
\textbf{reviewMethodology}: item by itemTheoremattack + AssumptionsComplete adversarial review + Symbol/Conceptconsistent adversarial review\\
\textbf{reviewStandard }: not polite polite. notaccept''partrigorous+Conjecture''. has leak hole then hit penetrate. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{〇、total overall impression image }\label{ux603bux4f53ux5370ux8c61}

this article documentclaimsto Spring
self-evolutiondynamicsconvergenceness advance lines''Rigorous Mathematical Derivation+Honest Critique''. actual actual read complete backConclusionis: \textbf{honestdegreeLabel has has deceive deceive ness. }
manyatmarkNoteis \texttt{{[}rigorousProof{]}}
result-dependencenot claim brightAssumptionsorexistenceat logic edit jump leap; Theorem 1.4
ConstructionnessProofmiddlehas a order of magnitudeincorrect; Theorem 2.1
 ignore slightly action action spacemay change nessthis a Core Problem; Theorem 3.1 counter example is hand select illusion number andnon-From 
Spring dynamicsmiddleDerivation out. withlowerisitem by itemTheorem solve dissect. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ a 、Theorem 1.2 O(log T/\sqrtT) convergence rate ------Ghadimi-Lan
Conditionsfull surface attack}\label{ux4e00theorem-1.2-ux7684-olog-tt-ux6536ux655bux7387ghadimi-lan-ux6761ux4ef6ux7684ux5168ux9762ux653bux51fb}

\subsubsection{1.1 Ghadimi-Lan
framework adapt use adversarial review}\label{ghadimi-lan-ux6846ux67b6ux7684ux9002ux7528ux6027ux5ba1ux67e5}

Theorem 1.2 claimsuse''Standard non-convex SGD Analysis''. text contribute middle cite using Ghadimi \& Lan
(2013), butnot careful fineverifyitsConditionsiswhetherin Spring setup place lowerholds. 

Ghadimi-Lan (2013, SIAM J. Optim.) coreAssumptions: - \textbf{(GL1)} \(f\)
is \(L\)-smooth smooth: \(\|\nabla f(x) - \nabla f(y)\| \leq L\|x - y\|\)
to has \(x, y\) - \textbf{(GL2)} stochastic gradient \(G(x, \xi)\) satisfies
\(\mathbb{E}[G(x,\xi)] = \nabla f(x)\) and
\(\mathbb{E}[\|G(x,\xi) - \nabla f(x)\|^2] \leq \sigma^2\) -
\textbf{(GL3)} \(f\) lower direction has world: \(\inf_x f(x) > -\infty\)

\subsubsection{\texorpdfstring{1.2 attack point \#1: global \(L\)-smooth smooth nessat
Spring
middledoes not hold (or often number explode explode)}{1.2 attack point \#1: global L-smooth smooth nessin Spring middledoes not hold (or often number explode explode)}}\label{ux653bux51fbux70b9-1ux5168ux5c40-l-ux5149ux6ed1ux6027ux5728-spring-ux4e2dux4e0dux6210ux7acbux6216ux5e38ux6570ux7206ux70b8}

Spring loss landscape view has has \(K!\)
 permutationto called ness (§1.2). considertwo is located atis differentpermutationbasin point \(\theta\) and
\(\pi \cdot \theta\). at it s 'sbetweenPathupper, loss functionmust penetrate the morebasinbetween ridge line. this some ridge line highdegreeand curve rate verdict determineglobal
Lipschitz often number. 

\textbf{ inquire }: Spring loss functionglobal \(L\)-smooth smooth often number ismanyfew? such as result \(L\) and
\(K!\) mutualclose (example such as \(L = \Omega(K!)\)
becauseis needmust cover cover has permutationbasinbetween ridge line), then: - \(\alpha_t \leq 1/L\)
Conditionsrequire step long \(\leq 1/K!\), andTheoremuse
\(\alpha_t = \alpha/\sqrt{t}\) in \(t < L^2\) time violate counter thisConditions -
convergence rate often number because factor \(L\alpha^2\sigma^2/2\) in the \(L\)
may make world change is \textbf{ empty set } (vacuous)

\textbf{documentmiddleNone any what regarding \(L\) order of magnitudeestimate. } thisis strict heavy leave leak ------nothas 
\(L\) estimate, \(O(\log T/\sqrt{T})\) often number mayisexponential level giant large. 

\subsubsection{\texorpdfstring{1.3 attack point \#2: \(\log T\)
because factor isself- find ------Ghadimi-Lan
 original thisnotrequire it }{1.3 attack point \#2: \textbackslash log T because factor isself- find ------Ghadimi-Lan original thisnotrequire it }}\label{ux653bux51fbux70b9-2log-t-ux56e0ux5b50ux662fux81eaux627eux7684ghadimi-lan-ux539fux672cux4e0dux9700ux8981ux5b83}

Ghadimi-Lan (2013) Standard result: take\textbf{ often number step long}
\(\alpha_t = \alpha = \sqrt{\frac{2(f(x_1)-f^*)}{L\sigma^2 T}}\), obtain to: 

\[\min_{t=1,\ldots,T} \mathbb{E}[\|\nabla f(x_t)\|^2] \leq \sqrt{\frac{2L\sigma^2(f(x_1)-f^*)}{T}} = O\left(\frac{1}{\sqrt{T}}\right)\]

\textbf{Note agree: None \(\log T\) because factor. }

document Theorem 1.2 with out present \(\log T/\sqrt{T}\), isbecauseis use
\(\alpha_t = \alpha/\sqrt{t}\)
 decay subtract step long------\(\sum \alpha_t^2 = \alpha^2 \sum 1/t = O(\log T)\)
take generation 
\(T\alpha^2\). but decay subtract step longatthisAnalysismiddle\textbf{nothas goodat}------ often number step longatnon-convex
SGD middlealready via to out \(O(1/\sqrt{T})\) rate (None
\(\log T\)). documentselect step long adjust degree\textbf{ inferior -ization }convergence rate. 

\textbf{must what documentnot know path Ghadimi-Lan
 often number step longresult (cite use butnotcorrectuse), must what therefore agree usemoreslow adjust degree come make result see rise come more'' present actual ''------twooneallnotmayaccept. }

\subsubsection{1.4 attack point 
\#3: has world varianceAssumptionsandpermutationto called nessforce}\label{ux653bux51fbux70b9-3ux6709ux754cux65b9ux5deeux5047ux8bbeux4e0eux6392ux5217ux5bf9ux79f0ux6027ux7684ux5f20ux529b}

\(\mathbb{E}[\|\zeta_t\|^2 | \mathcal{F}_t] \leq \sigma^2\)
Assumptionsgradientnoisehas uniform uniform upper bound. atpermutationto called basinboundaryat, forfromis differentbasinsample, mini-batch
gradientmay indicate toward complete fullis differentdirection------thistime \(\zeta_t\) (mini-batch and
full-batch gradient's difference)variancemay\textbf{atbasinboundaryatdivergence}. 

this is NOT theoretical nitpicking ------ this is aracticemiddlealready know phenomenon: whendatacontainsmany to called model formula time, SGD
gradientnoiseatto called nessboundaryatdifferent often large. Spring \(K!\)
heavypermutationto called nessamplifythis a Problem. 

\textbf{documentadmitsthis a point } (§1.2.2
''honestNote''admits quotient spaceuppernoisecovariancerequiremore careful fineAnalysis), but Theorem 1.2
 stated describe andProofmiddle and not overall present this a restricted. TheoremmarkNoteis \texttt{{[}rigorousProof{]}}
butnothas Notebrightnoisevarianceuniform uniform nessimplicit formula Assumptionsmaydoes not hold. 

\subsubsection{1.5 attack point \#4: gradientnormconvergence \neq
lossconvergence}\label{ux653bux51fbux70b9-4ux68afux5ea6ux8303ux6570ux6536ux655b-ux635fux5931ux6536ux655b}

documentself- self admitsthis a point (``Honest Critique''), butadmitsforcedegreeis insufficient. athighdegreenon-convex landscape view middle, \(\|\nabla f\| \approx 0\)
 point includes: - localextremelysmall value (mayveryshallow, lossveryhigh) - saddle point (lossmay any agree high) -
globalextremelysmall value (we s hope hope reaches ground direction)

for Spring, permutationto called ness agree taste -ing has \(K!\)
 equivalentglobalextremelysmall value ------butalsohas exponentialmanysaddle point connect connect it s. Theorem 1.2
 world cannotdistinguish''converges toglobalextremelysmall''and'' card atsaddle point''. document say thisis''gradientnorm world, not necessarily agree taste -ing loss function value convergence''------thisshould write advance Theorem Statementmiddle, andis NOT release atback surface ''Honest Critique'' inside. 

\subsubsection{1.6 to Theorem 1.2
 judge verdict }\label{ux5bf9-theorem-1.2-ux7684ux5224ux51b3}

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
Problem & strict heavydegree \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
global L-smooth smooth nessin Spring middleNoneestimate & \textbf{high}------ world maycan vacuous \\
log T because factor is step longselectresults inself- injury & \textbf{middle}------ often number step longmayavoid \\
has world varianceAssumptionsatbasinboundaryatmay doubt & \textbf{middle}------requireverify \\
gradientnormconvergence\neqlossconvergence cover cover & \textbf{middle}------honestmarkNoteinsufficient \\
\end{longtable}

\textbf{Summary}: \texttt{{[}rigorousProof{]}} markNoteattechniqueupperiscorrect (Standard non-convex
SGD Analysis confirm actual rigorous), butthisisto\textbf{some not necessarilyis Spring
function}rigorousProof------Spring iswhethersatisfiesshouldAnalysisallConditionsnotbyverify. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ two 、Theorem 1.4
information theorylower bound------Constructionness show example item by itemlinesattack}\label{ux4e8ctheorem-1.4-ux7684ux4fe1ux606fux8bbaux4e0bux754cux6784ux9020ux6027ux793aux4f8bux7684ux9010ux884cux653bux51fb}

\subsubsection{\texorpdfstring{2.1 attack point \#1: \(\sigma^2/\sqrt{T}\)
 in the \(T^{1/4}\)
order of magnitudeincorrect}{2.1 attack point \#1: \textbackslash sigma\^{}2/\textbackslash sqrt\{T\} in the T\^{}\{1/4\} order of magnitudeincorrect}}\label{ux653bux51fbux70b9-1sigma2sqrtt-ux4e2dux7684-t14-ux91cfux7ea7ux9519ux8bef}

documentclaims (§1.4.1 Proof): \textgreater{} ``for \(T\)
 step, noisecauses shift quantity is 
\(\sigma \sqrt{\sum \alpha_t^2} \approx \sigma \cdot O(T^{1/4})\) (when
\(\alpha_t = 1/\sqrt{t}\) time)''

\textbf{thisiswrong. }
\(\sum_{t=1}^T \alpha_t^2 = \sum_{t=1}^T 1/t = H_T = O(\log T)\), with: 

\[\sigma \sqrt{\sum_{t=1}^T \alpha_t^2} = \sigma \cdot O(\sqrt{\log T})\]

\textbf{andnon-} \(\sigma \cdot O(T^{1/4})\). 

\(O(T^{1/4})\) come source set person cost solve. such as result \(\alpha_t = 1/\sqrt{t}\), then
\(\alpha_t^2 = 1/t\), seek andis \(O(\log T)\), flat direction rootis 
\(O(\sqrt{\log T})\). must obtain to \(O(T^{1/4})\), require
\(\alpha_t = 1/t^{3/4}\) or class similar step long------butTheorem Statementmiddlebright confirm write is
\(\alpha_t = \alpha/\sqrt{t}\). 

\textbf{this order of magnitudeincorrect out present atTheoremProofcorecomputemiddle. }
entire noise drive action shift argumentation-dependenceatthis order of magnitude, becausethislower boundmiddle \(\sigma^2/\sqrt{T}\)
 term Derivationisnotmay reliable. 

\subsubsection{2.2 attack point \#2: 1D Constructioncannot generation table highdimensional Spring
 landscape view }\label{ux653bux51fbux70b9-21d-ux6784ux9020ux4e0dux80fdux4ee3ux8868ux9ad8ux7ef4-spring-ux666fux89c2}

documentConstruction \(f(x) = \sin^2(\pi K x / 2)\) in \([0,1]\)
upper a dimensionalfunction, claims'' model propose \(K\) headatpermutationlowerequivalentstationary point''. 

\textbf{ inquire }: - 1D functionhas \(K/2\)
 globalextremelysmall value (\(f=0\)), uniform uniform distribution. Spring highdimensional landscape view has \(K!\)
 permutationto called stationary point. \(K/2\) and \(K!\) scale complete fullis different------for
\(K=4\), Constructionmiddlehas 2 extremelysmall value, but Spring has 24. lower boundis what what nothas overall present 
\(K!\) because factor ? -
 a dimensionalfunctionNonemethod model propose highdimensional landscape view middlebasin'sbetween\textbf{dimension}and\textbf{ connect connect structure}. athighdimensionalmiddle, SGD
maywiththroughsaddle pointatis differentbasinbetween escape escape (thiscurrentlyis Proposition 1.3 Content!), but 1D
functionmiddlethiskind escape escape isis impossible (must flip the more potential barrier). -
willhighdimensionalProblemprojectsto a dimensional'' sequence participate quantity ''upper agree taste -ing togradientnoiseandgradientsignal advance lines a kindnot claim brighton average. thiskindon averagemaysystemness ground changesconvergence rate. 

\textbf{1D Construction in the extremelysmall value quantity (\(K/2\))and Spring
 in the extremelysmall value quantity (\(K!\))notatsame a order of magnitude. } becausethisshouldlower bound
\(K\)--dependencenessnot via verify. 

\subsubsection{2.3 attack point \#3: function family iswhetherin Spring
Assumptions class inner? }\label{ux653bux51fbux70b9-3ux51fdux6570ux65cfux662fux5426ux5728-spring-ux7684ux5047ux8bbeux7c7bux5185}

\(f(x) = \sin^2(\pi K x / 2)\) is: - smooth smooth (\(C^\infty\))✓ - cycle period 
✗ (Spring
loss landscape view is NOT cycle period ------althoughhoweverhas permutationto called ness, butis differentpermutationbasinatparameter space in the distributionis NOT uniform uniform lattice point, andisby
Weyl chambers describecomplexity several what) - lower direction has world ✓ -
 has extremelysmall value equivalent (uniformis globalextremelysmall)✓ (thisandpermutationto called nessconsistent)

but Spring loss functionis NOT \(\sin^2\) simple singlescale------ it has \(K!\)
 extremelysmall value distributionin \(O(K d_s^2)\)
dimensionalspacemiddle, andextremelysmall value 'sbetween connect connect structureinvolvingcomplexitysaddle point network network. use a dimensional cycle period function come ''lower bound''this type a complexity landscape view, itsargumentationrequire a \textbf{dimension convention simple Lemma} come form -ization ------documentnothas provide. 

\subsubsection{\texorpdfstring{2.4 attack point \#4: lower bound
\(\Omega(\min(\sigma^2/\sqrt{T}, 1/T))\) \(\min\)
structuremay doubt }{2.4 attack point \#4: lower bound \textbackslash Omega(\textbackslash min(\textbackslash sigma\^{}2/\textbackslash sqrt\{T\}, 1/T)) \textbackslash min structuremay doubt }}\label{ux653bux51fbux70b9-4ux4e0bux754c-omegaminsigma2sqrtt-1t-ux7684-min-ux7ed3ux6784ux53efux7591}

forhighnoise \((\sigma^2 \gg 1/\sqrt{T})\), lower boundis 
\(\Omega(\sigma^2/\sqrt{T})\). forlownoise, lower bound change is 
\(\Omega(1/T)\). this agree taste -ing atNonenoiseextremely finite lower (\(\sigma \to 0\)), lower boundis 
\(\Omega(1/T)\)------thisis a \textbf{ a order confirm determine type gradient descent}rate. 

but confirm determine type gradient descentatnon-convexsmooth smooth functionuppermaywith do to \(O(1/T)\)
gradientnormconvergence. withlownoisecaselowerlower boundis
\(\Omega(1/T)\), andupper boundmatches------thissurfaceupperis tight. butthis inside \(\min\)
 operation action in practiceisat say ''taketwoonemiddlerelativelysmallthat action is lower bound'', and
\(\min(\sigma^2/\sqrt{T}, 1/T)\) structure agree taste -ing: \textbf{when \(T\)
is insufficientlargetime, noise term main guide; when \(T\) sufficientlargetime, confirm determine type errormain guide. }
this structurecorrectness strict heavy-dependenceat §2.1
 in the order of magnitudeAnalysis------and we s already via discoverthat inside has wrong. 

\subsubsection{2.5 to Theorem 1.4
 judge verdict }\label{ux5bf9-theorem-1.4-ux7684ux5224ux51b3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.4286}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5714}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 strict heavydegree
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\sigma\sqrt{\sum\alpha_t^2} = O(T^{1/4})\) isorder of magnitudeincorrect &
\textbf{Fatal}------shouldis \(O(\sqrt{\log T})\) \\
1D Constructionmiddleextremelysmall value quantity (\(K/2\))and Spring (\(K!\))notmatches &
\textbf{high}------\(K\)-scalenotverify \\
From \(K!\) dimensionalto 1D projectsnothas argumentation & \textbf{high}------dimension convention simple notProof \\
function family (cycle period \(\sin^2\))notmatches Spring non- cycle period basin several what &
\textbf{middle}------requiremorestronganalogyargumentation \\
\end{longtable}

\textbf{Summary}: \(T^{1/4}\) order of magnitudeincorrectsingle independent then foot with make Theorem 1.4
ProofNone effect. i.e. make correctionorder of magnitudeincorrect, 1D Constructionand Spring
highdimensional landscape view 'sbetween grand trench alsonotby bridge connect. whenfrontform \texttt{{[}rigorousProof{]}}
markNote has has error guide ness. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ three 、Theorem 2.1 leave regret upper bound------Exp3
algorithm adapt match full surface review}\label{ux4e09theorem-2.1-ux9057ux61beux4e0aux754cexp3-ux7b97ux6cd5ux9002ux914dux7684ux5168ux9762ux5ba1ux67e5}

\subsubsection{3.1 attack point \#1: action action spacemay change ness------Exp3
Standard Analysisimplicit hide crack trace }\label{ux653bux51fbux70b9-1ux52a8ux4f5cux7a7aux95f4ux7684ux53efux53d8ux6027exp3-ux6807ux51c6ux5206ux6790ux7684ux9690ux85cfux88c2ux75d5}

Exp3 Standard regret bound (Auer et al., 2002)Assumptions\textbf{fixed action action space} (\(N\)
 arm, begin finalmay use). Spring memory bankdynamic carry come two Problem: 

\textbf{(a) Variable Action Space Size}: \(\mathcal{M}_t^{\text{dormant}}\)
size follow \(t\) change -ization. memory bankpruneremoveStatusatom; newStatusdiscoveraddatom. Exp3
weighttoward quantity \(w_t\) dimension follow 'schanges. documentimplicitly use 
\(|\mathcal{S}|\) (totalStatusatom number)take generation 
\(|\mathcal{M}_t^{\text{dormant}}|\)------butthisonlyat
\(|\mathcal{M}_t^{\text{dormant}}| \leq |\mathcal{S}|\)
timeto out upper bound. thisis a has effect loose relax, but loose relax order of magnitudeis 
\(|\mathcal{S}| / |\mathcal{M}_t^{\text{dormant}}|\), whenthe majority ofStatusatomatat''active'' (non-dormant)Statustime, world change obtain non- often loose. 

\textbf{(b) newincrease arm weightinitialization -ization }: whennewStatusatombydiscover and add enter 
\(\mathcal{M}_{t+1}^{\text{dormant}}\) time, Exp3
requireis it initialization -ization a weight. documentnothas indicate determineinitialization -ization rule then. such as result initializationweight over small (mutualforalreadyhas arm), new arm atverylongtimebetweeninnerwill notbyselect------thisincreases leave regret. such as result initializationweight over large, it may squeeze occupy alreadyhas arm selectprobability. 

\subsubsection{3.2 attack point \#2: to resist nessAssumptions------to hand iswhetheris Oblivious
? }\label{ux653bux51fbux70b9-2ux5bf9ux6297ux6027ux5047ux8bbeux5bf9ux624bux662fux5426ux662f-oblivious-ux7684}

documentclaimsmemory bankdynamicproduces\textbf{to resist ness} reward incentivize (§2.2.1): \textgreater{}
``byatmemory bankdynamicdepends on history history select, reward incentivize sequencemayisto resist ness (andnon-random
i.i.d.)''

Exp3 regret bound confirm actual is applicable to\textbf{non-random} (non-stochastic /
adversarial) reward incentivize sequence. buthas a finemicro-'sat: 

in Auer et al.~(2002)
frameworkmiddle, to hand maywithis\textbf{self- adapt should} (according to over remove select verdict determinewhenfront reward incentivize), butcannotaccording to\textbf{whenfront round time random -ization } verdict determine reward incentivize. Spring
memory bankdynamicbelongs tofrontone------atth \(t\) round select \(a_t\) post-, Memory Bank Update
\(\mathcal{M}_{t+1}\) willimpactth \(t+1\) round reward incentivize, butwill notimpactth \(t\)
 round (becauseis th \(t\) round reward incentivize byth \(t\) round Statusatomquality verdict determine, atselect
\(a_t\) post- confirm determine). 

\textbf{butthis inside has a cyclesbecause result Problem}: th \(t\) round select \(a_t\)
based onsome evaluationresult, andthis evaluationresult\textbf{this self -dependenceatwhenpre-Spring parameter
\(\theta_t\)}. \(\theta_t\) isbypre-\(t-1\) round memory bankStatusthrough SE-1 (SGD
update) verdict determine. becausethis reward incentivize sequencethrough a complexity counter feedback time path -dependenceat history history ------this confirm actual is
Exp3 canwithat reason to resist nessSetup. ✓

\textbf{moremicro- wonderful Problem}: Exp3 needmust reward incentivize 
\(r_t(a_t) \in [0,1]\). document say reward incentivize is''afterYajieaudit obtain toqualityseparate (normalizationback
Cercis separate number orDetection Margin \(\Delta_s\) estimate)``. but \(\Delta_s\)
estimaterequire\textbf{trueLabel} come confirm determine \(p_{\text{clean}}\) and
\(p_{\text{noisy}}\). at actual actual Spring systemmiddle, Yajieauditto out onlyis
\(C(s) = \sum v_m\)------expertvoteConsistency count, andnon-trueDetection Margin. use 
\(\hat{\Delta}_s\) replace generation \(\Delta_s\)
introduceestimateerror, thisatregret boundmiddlenothas is build model. 

\subsubsection{\texorpdfstring{3.3 attack point \#3: comparison base quasi-\(\mu^*\)
DefinitionProblem}{3.3 attack point \#3: comparison base quasi-\textbackslash mu\^{}* DefinitionProblem}}\label{ux653bux51fbux70b9-3ux6bd4ux8f83ux57faux51c6-mu-ux7684ux5b9aux4e49ux95eeux9898}

\[\mu^* = \max_{\mathcal{I} \subset \mathcal{S}, |\mathcal{I}| = B} \frac{1}{B} \sum_{s \in \mathcal{I}} \Delta_s\]

this comparison base alignment Assumptions: 1. \textbf{Optimal fixed set combined}attimebetweenuppernot change ------but
\(\Delta_s\) this self follow timebetween change -ization (\(\Delta_s(t)\)), becauseis Spring
parameterat advance -ization. documentuseis \(\Delta_s\) (Nonetimebetween indicate mark), thisdark show static
\(\Delta_s\)------and §3 middle \(\Delta_s(t)\) entire timebetween-dependenceAnalysis contradiction contradiction. 2.
\textbf{ has \(\Delta_s\) matter first already know }------Optimal fixed strategyrequire know path has 
\(\Delta_s\) true value. this some value i.e. make atbacksee'sbrightmiddlealsoisnot know (becauseis it s -dependenceat
counterfactual Spring advance -ization track trace). with \(\mu^*\)
is a \textbf{uncomputable}comparison base alignment. 

at actual actual middle, uniquemay use comparison base alignment is'' matter backOptimal via verify estimate'', andnon-\(\mu^*\). 

\subsubsection{3.4 attack point \#4: Exp3
importantnessadd weight atmemory bankdeletetimenotDefinitionlinesis }\label{ux653bux51fbux70b9-4exp3-ux7684ux91cdux8981ux6027ux52a0ux6743ux5728ux8bb0ux5fc6ux5e93ux5220ux9664ux65f6ux7684ux672aux5b9aux4e49ux884cux4e3a}

Exp3 update rule then (§2.2.1): 

\[w_{t+1}(s) = w_t(s) \cdot \exp\left(\frac{\gamma \cdot \hat{r}_t(s)}{|\mathcal{M}_t^{\text{dormant}}| \cdot p_t(s)}\right)\]

whenStatus \(s\) isFrom memory bankmiddle\textbf{pruneremove}time, \(w_{t+1}(s)\)
shouldhow type ? such as result delete, thenshould arm permanent far Nonemethodagainbyselect------butthis agree taste -ing prune operation action is\textbf{irreversible}, andsuch as result a highqualityStatusby error delete (false negative), algorithmpermanent far nothas machine will restore complex. 

document Corollary 2.1
 confirm actual discuss theory preserve guard prunewithavoid error delete ------butthiswillpruneFrom algorithma portion of change become a requiresatisfies special determineinequalityouterpart convention bind. such as result this some convention bind notbysatisfies (actual actual middlevery difficult guarantee), Ext3
regret boundrequire amount outer a term come penalty penalty error delete. 

\subsubsection{3.5 to Theorem 2.1
 judge verdict }\label{ux5bf9-theorem-2.1-ux7684ux5224ux51b3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.4286}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5714}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 strict heavydegree
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
 action action spacesizemay change ------Exp3 AnalysisAssumptionsfixed & \textbf{middle}------ use 
\(|\mathcal{S}|\) upper bound cover cover butleads to loose relax \\
 reward incentivize signal-dependenceunobservable \(\Delta_s\) &
\textbf{high}------ actual actual middle reward incentivize is carry noise estimate \\
comparison base quasi-\(\mu^*\) usestatic \(\Delta_s\) and §3 contradiction contradiction &
\textbf{middle}------Symbol/Conceptis inconsistent \\
newincrease/delete arm weightmanagesnotDefinition & \textbf{middle}------requirealgorithmfine -ization \\
\end{longtable}

\textbf{Summary}: \texttt{{[}rigorousProof{]}} markNotetoStandard Exp3 Analysisiscorrect, but
Spring
 special differentness (may change action action space、estimateerror、irreversibleprune)introduceStandard Analysisnot cover cover complexitybecause element. Theorem
2.1 shouldmarkNoteis 
\texttt{{[}rigorousProof (atfixed action action spaceandExact reward incentivize observationAssumptionslower){]}}. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ four 、Theorem 3.1
monotonicity counter example ------Constructioncombinedmethodnessattack}\label{ux56dbtheorem-3.1-ux7684ux5355ux8c03ux6027ux53cdux4f8bux6784ux9020ux7684ux5408ux6cd5ux6027ux653bux51fb}

\subsubsection{4.1 attack point 
\#1: counter example in the probability number character is hand select illusion number }\label{ux653bux51fbux70b9-1ux53cdux4f8bux4e2dux7684ux6982ux7387ux6570ux5b57ux662fux624bux9009ux7684ux5e7bux6570}

Theorem 3.1 Constructionprovidesconcrete number character: -
\(p_{\text{clean},a}(t) = 0.2 \to p_{\text{clean},a}(t+1) = 0.35\) -
\(p_{\text{noisy},a}(t) = 0.7 \to p_{\text{noisy},a}(t+1) = 0.65\)

this some number character From \textbf{ which inside } come ? documentclaimsbecause result chain is: \textgreater{}
``Spring self-Note agree forcemechanismatth \(t\) round updateparameter \(\theta_t\)
withminimum -ization self-reconstructionloss\ldots byatmanyheadNote agree forcecommon share representation, to
\(s_a^{\text{noisy}}\) overfittingimpactStatus \(s_a\) has examplerepresentation''

but\textbf{nothas any what Derivation}willself-reconstructionlossgradient descentandexpertseparate disagree probabilityconcrete number value change -ization connection rise come. claim bright
\(p_{\text{clean}}\) From 0.2 change is 0.35 onlyis\textbf{asserts}, nothas: - Spring
lossgradientcompute - expertmodeltorepresentation change -ization sound shouldfunction -
From parameter change -ization toseparate disagree probability change -ization map

\subsubsection{4.2 attack point 
\#2: counter example Core Mechanism------``overfittingtonoisesample''------and Spring
training goal markis inconsistent}\label{ux653bux51fbux70b9-2ux53cdux4f8bux7684ux6838ux5fc3ux673aux5236ux8fc7ux62dfux5408ux5230ux566aux58f0ux6837ux672cux4e0e-spring-ux7684ux8badux7ec3ux76eeux6807ux4e0dux4e00ux81f4}

\textbf{Spring training goal markisminimum -ization self-reconstructionerror}: 

\[\mathcal{L}_{\text{Spring}}(\theta) = \sum_{i=1}^{N} \|\text{Spring}_{\text{MH}}(s_i; \theta) - s_i\|^2\]

thisis a \textbf{Nonesupervises} autoencoding goal mark. Spring
bytrainingis From upperlower text heavy build Statusatom------ it \textbf{notdirect} make using Cercis
 evaluate separateorLabelinformation. 

Theorem 3.1 counter example claims Spring
``overfittingtomemory bankmiddle error marknoisesample''. but'' error mark''is Cercis
 evaluate separate (orYajieaudit)result------is NOT Spring training input enter. Spring
trainingtimeonly see toStatusatomthis self (\(s_i\)), andnon- it s qualityLabel. 

\textbf{key inquire }: is what what Spring self-reconstructiontrainingwill''overfitting''to a in Cercis
 evaluate separatemiddleby error markis dry clean Statusatom? self-reconstructionlossonlyclose core heavy build accuracy------such as result 
\(s_a^{\text{noisy}}\) is a classical type Statusatom (andits he \(s_a\)
exampleatCharacteristicsspacemiddlesimilarity), Spring heavy build it shouldandits he \(s_a\)
exampleNonedifferent. such as result it different often (ratio such ascontainsnoise model formula), Spring
reconstructionerrorwillverylarge------butthiswill drive make Spring
adjustparameterwith\textbf{moregood ground }reconstruction it, thismay counter andincreaselarge it andcurrently often \(s_a\)
examplerepresentation distance from (andnon- corrupt dye it s). 

\textbf{ counter example argumentation logic edit in practicemayis counter toward}: such as result noisesampleanddry clean sampleatCharacteristicsspacemiddleis different, Spring
maywilllearningwill it s separateopen (withminimum -ization eachself-reconstructionerror), this counter andwillincreases
\(\Delta_s\) (andnon-decreases). 

\subsubsection{4.3 attack point 
\#3: `` pseudo generation code ''is NOT model propose ------is hit impression language sentence }\label{ux653bux51fbux70b9-3ux4f2aux4ee3ux7801ux4e0dux662fux6a21ux62dfux662fux6253ux5370ux8bedux53e5}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p\_clean\_init }\OperatorTok{=} \FloatTok{0.2}
\NormalTok{p\_noisy\_init }\OperatorTok{=} \FloatTok{0.7}
\CommentTok{\#... }
\NormalTok{p\_clean\_after }\OperatorTok{=} \FloatTok{0.35} \CommentTok{\# hand select value }
\NormalTok{p\_noisy\_after }\OperatorTok{=} \FloatTok{0.65} \CommentTok{\# hand select value }
\ControlFlowTok{assert}\NormalTok{ Delta\_after }\OperatorTok{\textless{}}\NormalTok{ Delta\_init}
\end{Highlighting}
\end{Shaded}

this 12
lines generation code nothas do any what matter situation ------ it claim brightinitialization value 、 claim bright hand select mostfinal value 、howeverbackverify pre- first designgoodConclusion. \textbf{thisis NOT model propose, thisisheavy word formula } (tautology). it proves''such as result we s select
p\_clean\_after = 0.35 and p\_noisy\_after = 0.65, that what Δ
willsubtractsmall''------thisis\textbf{ a determine}is true, becauseis 0.30 \textless{} 0.50. 

must become is a combinedmethod counter example Construction, generation code requirecontains: - Spring
modelachieves (tofewisparameter -ization simplifies version) - trainingcycles (SGD Step) -
expertmodelachieves (tofewis model propose) - demonstrateparameterupdate
\(\theta_t \to \theta_{t+1}\) confirm actual leads to describe probability change -ization 

whenfrontform'' pseudo generation code model propose verify''is 0 information quantity. 

\subsubsection{4.4 attack point 
\#4: noiseNote enter mechanismiswhethercombinedmethod? }\label{ux653bux51fbux70b9-4ux566aux58f0ux6ce8ux5165ux673aux5236ux662fux5426ux5408ux6cd5}

 counter example claims ``Spring self-reconstructionloss\ldots will \(s_a^{\text{noisy}}\)
 model formula action is currently often model formula come learning''. thisimplicitlyAssumptions: 

\textbf{Assumptions}: memory bankmiddleexistenceat a by error marknoisesample, andthis sample''noise model formula ''maywiththroughself-reconstructionlossgradient pass broadcast toits he dry clean samplerepresentation. 

this Assumptionsis equivalent toclaims: \textbf{self-reconstructiontrainingmaywith pass broadcast Labelnoise}------i.e. make trainingsignalthis self does not involveLabel. thisis a notbykeepclaims. such as result self-reconstructiononlyislearningdata flow shape several what structure (Standard 
autoencoder
linesis), that what noisesampleat flow shape upperLocation (andnon- it Label) verdict determine it such as what impactrepresentation. such as result noisesampleatCharacteristicsspacemiddleatat flow shape lowdensedegree region domain, autoencoder
maywill ignore slightly it (highreconstructionerror \to
smallgradientcontribution); such as result it atathighdensedegree region domain, it andits he dry clean sample a rise Definition flow shape ------thistime it not compute ''noise'', andonlyis flow shape a portion of. 

\subsubsection{4.5 to Theorem 3.1
 judge verdict }\label{ux5bf9-theorem-3.1-ux7684ux5224ux51b3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.4286}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5714}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 strict heavydegree
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
probability change -ization number character is hand select illusion number, non-From dynamicsDerivation &
\textbf{Fatal}------ violate counter ConstructionnessProofbasicrequire \\
self-reconstructiontrainingandLabelincorrect'sbetweenbecause result chain notestablish &
\textbf{Fatal}------ counter example mechanismbased on a notverifyAssumptions \\
`` pseudo generation code ''is 0 information quantity heavy word formula & \textbf{high}------constitutes error guide \\
noise pass broadcast mechanismimplicit formula Assumptionsnot claim bright & \textbf{high}------self-reconstruction\neqLabelsupervises \\
\end{longtable}

\textbf{Summary}: Theorem 3.1 \texttt{{[} counter example Construction, rigorous{]}}
markNote strict heavynot actual. this\textbf{is NOT } a counter example Construction------ it is a group hand select number character attach upper a similar isandnon- narrate matter. Core Mechanism (self-reconstructiontrainingwill
Cercis
 error mark pass broadcast is representation corrupt dye)nothas isDerivation、nothas is model propose 、nothas isProof. it may collide clever at\textbf{some some }Spring
examplemiddleholds (direct feel upperis plausible
), butwhenfrontformlowernotshouldby called is ''rigorousProof''. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ five 、equationNumber、Symbol/Conceptconsistent adversarial review}\label{ux4e94ux65b9ux7a0bux7f16ux53f7ux7b26ux53f7ux4e00ux81f4ux6027ux5ba1ux67e5}

\subsubsection{\texorpdfstring{5.1 \(\Delta_s(t)\) vs \(\Delta_s\) vs
\(\Delta_{a_t}\)
mixed chaotic }{5.1 \textbackslash Delta\_s(t) vs \textbackslash Delta\_s vs \textbackslash Delta\_\{a\_t\} mixed chaotic }}\label{delta_st-vs-delta_s-vs-delta_a_t-ux7684ux6df7ux4e71}

\begin{itemize}
\tightlist
\item
 \textbf{§0.2}: Definition
 \(\Delta_s(t) = p_{\text{noisy},s}(t) - p_{\text{clean},s}(t)\) (carry timebetween indicate mark)
\item
 \textbf{Theorem 1.1-1.2}: not make using \(\Delta_s\) (this some Theoremregardinggradientconvergence)
\item
 \textbf{§2.1.2}: Definition
 \(\mu^* = \max_{\mathcal{I}} \frac{1}{B} \sum_{s \in \mathcal{I}} \Delta_s\) (\textbf{None}timebetween indicate mark!)
\item
 \textbf{§2.1.2}: Definition
 \(R_T = T \cdot \mu^* - \sum_{t=1}^{T} \Delta_{a_t}\) (mixedcombined: \(\mu^*\)
 usingstatic \(\Delta_s\), summation using \(\Delta_{a_t}\)------but
 \(\Delta_{a_t}\) istimebetween \(t\) edge actual stillissome fixed value ?)
\end{itemize}

\textbf{attack}: \(\Delta_s\) in §0.2
middleistimebetweenmutualclose (\(\Delta_s(t)\)), butin §2
 leave regret Analysismiddlebywhen action timebetweenNoneclose. such as result \(\Delta_s\) follow timebetween change -ization (thisis §3
allmain problem !), thencomparison base quasi-\(\mu^*\)
shouldistimebetweenmutualclose------Optimal fixed set combinedatis differenttimebetweenmayis different. document complete fullnothas at reason this a point. 

\subsubsection{\texorpdfstring{5.2 reward incentivize and leave regret Definition in the \(\Delta_{a_t}\)
ambiguity}{5.2 reward incentivize and leave regret Definition in the \textbackslash Delta\_\{a\_t\} ambiguity}}\label{ux5956ux52b1ux548cux9057ux61beux5b9aux4e49ux4e2dux7684-delta_a_t-ux6b67ux4e49}

 leave regret Definition
\(R_T = T \cdot \mu^* - \sum_{t=1}^{T} \Delta_{a_t}\) (§2.1.2). this inside 
\(\Delta_{a_t}\) is: - (a) selectstructure \(a_t\) Detection Margin? - (b)
observationto reward incentivize (afterYajieauditestimateDetection Margin)? 

such as result \(\Delta_{a_t}\)
istrueDetection Margin (unobservable), that what atpracticemiddleNonemethodcompute leave regret. such as result it isestimate value, that what 
\(R_T\) is via verify leave regret, and Theorem 2.1-2.3 middleAnalysistheory leave regret is different. 

\subsubsection{5.3 memory bankDefinition in the Cercis
 evaluate separate language meaning }\label{ux8bb0ux5fc6ux5e93ux5b9aux4e49ux4e2dux7684-cercis-ux8bc4ux5206ux8bedux4e49}

\textbf{§0.1}: \(\mathcal{M}_t = \{(s_i, \text{Cercis}_t(s_i), \text{audit}_t(s_i))\}_{i=1}^{N}\)

\textbf{§2.1.1}: action action space
\(\mathcal{A}_t = \mathcal{M}_t^{\text{dormant}}\) (dormantstructure)

but \(\mathcal{M}_t^{\text{dormant}}\) Definitionat what at? \(\mathcal{M}_t\)
Definitioncontains has \(N\) Statusatom and its Cercis/audit
 dollar data------ which partis''dormant''? ``dormant'' agree taste -ing what what ? document say ''notbysufficientevaluationStatusatom''------butthisandmemory bankDefinition (memory bankcontains has Statusatom evaluate separate dollar data)is inconsistent. 

\subsubsection{\texorpdfstring{5.4 §1.4.1 middle \(x\)
variablenot claim bright}{5.4 §1.4.1 middle x variablenot claim bright}}\label{ux4e2d-x-ux53d8ux91cfux7684ux672aux58f0ux660e}

Theorem 1.4 Proofintroduce \(f(x) = \sin^2(\pi K x / 2)\). conflict howeverFrom 
\(\theta\) (highdimensionalparameter) cut exchange to \(x\) (a dimensionalvariable), andnot claim bright \(x\) such as what and
\(\theta\) close connect. this cut exchange isnon-method jump leap ------ read onebyrequireself-lines brain supplement a From 
\(\Theta\) to \([0,1]\) '' sequence participate quantity ''map. 

\subsubsection{5.5 §3.2.2 in the Fisher
consistentnessSymbol/Concept}\label{ux4e2dux7684-fisher-ux4e00ux81f4ux6027ux7b26ux53f7}

Theorem 3.2 Assumptions (A2) claims
\(\mathbb{E}[E_m(h)] = P(Y_{\text{true}} = 1 | h)\). butexpert input out 
\(E_m(h)\) is \(\{0,1\}\)
 value vote (\(v_m = \mathbf{1}[E_m(s) \neq y]\)), andis NOT Conditionsprobability. \(E_m\)
in §0.2
middleDefinitionis producesdiscretevotemodel------itsexpecthow what equalsConditionsdistribution? this inside existenceat\textbf{ class type notmatches}: left edge isseparate class incorrectprobability, right edge isLabelConditionsdistribution. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ six 、not claim brightAssumptions------honestdegree bare character }\label{ux516dux672aux58f0ux660eux7684ux5047ux8bbeux8bdaux5b9eux5ea6ux8d64ux5b57}

\subsubsection{6.1 Proposition
1.3 (saddle point escape escape)------noiseeachtowardsamenessAssumptions}\label{proposition-1.3ux978dux70b9ux9003ux9038ux566aux58f0ux5404ux5411ux540cux6027ux5047ux8bbe}

Proposition 1.3 claims quotient spaceuppernoisemaintains''eachtowardsamenesscovariancestructure (becauseis 
\(S_K\) is tight group, Haar measure degreeisuniform uniform)``. this argumentationmixed confuse two: -
\(S_K\) Haar measure degreeisuniform uniform ------thisguaranteetorandompermutationon averageisuniform uniform -
gradientnoise \(\zeta_t\) in \(S_K\) action use lowerdistribution------thismustmust \(\zeta_t\)
distributionin \(S_K\) lower\textbf{not change }

\textbf{Haar measure degreeuniform uniform nessnot contain contain gradientnoise \(S_K\)-not change ness. } gradientnoisefrom
mini-batch sampling------such as result mini-batch
middleis differentpermutationstructureproportionnotuniform uniform (actual actual middleis exactly the case------some some permutationmay out present atmoremanysamplemiddle), noiseat quotient spaceupperprojectsthenis NOT eachtowardsameness. 

\textbf{this Assumptionsnothas inTheorem StatementorProofmiddle claim bright. }

\subsubsection{6.2 Theorem 3.2 (expectmonotonicity)------Fisher
consistentnesstoneural networkdoes not hold}\label{theorem-3.2ux671fux671bux5355ux8c03ux6027fisher-ux4e00ux81f4ux6027ux5bf9ux795eux7ecfux7f51ux7edcux4e0dux6210ux7acb}

Assumptions (A2) (expert Fisher
consistentness)claimseach expert input out probabilityatexpectmeaninglower force near Conditionsdistribution. \textbf{forthrough
SGD trainingneural network, thisnothas has finite sampleguarantee. } matter actual upper: -
neural networkis\textbf{incorrect indicate determine} (misspecified)------trueConditionsdistribution through often notat network network mayrepresentationAssumptions class inner
- SGD onlycanconverges toloss landscape view some stationary point, not necessarilyis Bayesian leaf -ese Optimal -
has finite samplelowerexistenceatgeneralization error

documentadmitsthisis''extremelyitsstrongAssumptions'', but\textbf{stillhoweverwill Theorem 3.2 markNoteis 
\texttt{{[}rigorousProof, atAssumptionsConditionslower{]}}}. whenAssumptionsat actual actual middle several -ish agree determinedoes not holdtime, markNote''rigorous''is has has error guide ness------Theoremhas effect ness complete full suspend float at
Fisher consistentness empty middle building pavilion upper. 

\subsubsection{6.3
not claim brightAssumptions converge total table }\label{ux672aux58f0ux660eux7684ux5047ux8bbeux6c47ux603bux8868}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2143}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4643}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3214}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
not claim brightAssumptions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
is what important
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Thm 1.2 & global \(L\)-smooth smooth often number is NOT \(O(K!)\) &
impactconvergence rate often number because factor, may make world vacuous \\
Thm 1.2 & \(\sigma^2\) inpermutationbasinboundaryathas world &
basinboundaryatgradientvariancemaydivergence \\
Prop 1.3 & gradientnoise \(S_K\)-not change ness & saddle point escape escape coreargumentation \\
Thm 1.4 & From \(\Theta\) to \([0,1]\) dimension convention simple maintainslower bound & 1D
Constructioncombinedmethodnessfront raise \\
Thm 2.1 & action action spacefixed (Standard Exp3 Assumptions) & Spring action action spacemay change \\
Thm 2.1 & reward incentivize \(\Delta_s\) canExact observation & actual actual middleonlycanobservation carry noise Cercis
estimate \\
Thm 2.2 & \(\Delta_s\) satisfies Lipschitz Conditions &
paperself- self admitsthisis actual nessAssumptions, treat verify \\
Thm 3.1 & self-reconstructiontraining pass broadcast Labelnoise & anti- example corebecause result chain notbyProof \\
Thm 3.2 & expert Fisher consistentness & toneural networkdoes not hold \\
Thm 3.4 & -dependence Thm 3.2 has Assumptions & aggregate combinedmonotonicity suspend float atth 3 layer empty middle building pavilion upper \\
Prop 3.1 & Lipschitz Assumptionsat indicate show functionexpectuppermay pass recurrent ness &
boundaryConditionsAssumptions amount outerrequire (documentpartadmits) \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ seven 、Fataldefect converge total}\label{ux4e03ux81f4ux547dux7f3aux9677ux6c47ux603b}

\subsubsection{\texorpdfstring{Fataldefect \#1: Theorem 1.4 \(T^{1/4}\)
order of magnitudeincorrect}{Fataldefect \#1: Theorem 1.4 T\^{}\{1/4\} order of magnitudeincorrect}}\label{ux81f4ux547dux7f3aux9677-1-theorem-1.4-ux7684-t14-ux91cfux7ea7ux9519ux8bef}

\textbf{ness}: mathematicalcomputeincorrect\\
\textbf{Location}: §1.4.1, Theorem 1.4 Proof\\
\textbf{concreteincorrect}:
\(\sigma\sqrt{\sum\alpha_t^2} = \sigma\sqrt{H_T} = O(\sqrt{\log T})\), andnon-claims
\(O(T^{1/4})\)\\
\textbf{back result }: noise shift order of magnitudebyhigh estimate convention \(T^{1/4}/\sqrt{\log T}\)
 times. entire lower boundnoisemutualclose term Derivationrequireheavy do. Theorem \texttt{{[}rigorousProof{]}}
Labelatwhenfrontformlowerdoes not hold. 

\subsubsection{Fataldefect \#2: Theorem 3.1 '' counter example ''is NOT From Spring
dynamicsmiddleDerivation}\label{ux81f4ux547dux7f3aux9677-2-theorem-3.1-ux7684ux53cdux4f8bux4e0dux662fux4ece-spring-ux52a8ux529bux5b66ux4e2dux63a8ux5bfcux7684}

\textbf{ness}: ConstructionnotComplete------ hand select illusion number andnon-Derivation out quantity \\
\textbf{Location}: §3.2.1-3.2.3\\
\textbf{concreteincorrect}: probability change -ization number character (0.2\to0.35,
0.7\to0.65)isfrontdetermineselect, is NOT From Spring
self-reconstructiontraininggradient descentmiddleDerivation out. pseudo generation code verifyiscycles (tautological). self-reconstructiontraining
\to Labelnoise pass broadcast because result chain notbyestablish. \\
\textbf{back result }: anti- example \texttt{{[}rigorousProof{]}} Labelnot actual. should drop level is 
\texttt{{[} heuristic estimate guess measure {]}} orsupplementFrom Spring dynamicstoprobability change -ization concreteDerivation. 

\subsubsection{Fataldefect \#3: Theorem 1.2 log T
because factor is step long adjust degreeselectresults in}\label{ux81f4ux547dux7f3aux9677-3-theorem-1.2-ux7684-log-t-ux56e0ux5b50ux662fux6b65ux957fux8c03ux5ea6ux9009ux62e9ux9020ux6210ux7684}

\textbf{ness}: usenon-Optimal step long adjust degree, obtain tonon-Optimal convergence rate \\
\textbf{Location}: §1.3.2\\
\textbf{concreteincorrect}: Ghadimi-Lan (2013) use often number step longprovides
\(O(1/\sqrt{T})\) (None log T). document make using \(\alpha_t = \alpha/\sqrt{t}\)
 decay subtract step longleads to \(\sum \alpha_t^2 = O(\log T)\), From and obtain to
\(O(\log T/\sqrt{T})\). thisis NOT convergence rate thisCharacteristics------ it isAnalysismiddleselect
artifact. \\
\textbf{back result }: documentto out convergence rate ratio text contribute in the Optimal resultslow a \(\log T\)
because factor. althoughhoweveris NOT mathematicalincorrect, butwill artifact present present is ''convergence rate ''isto text contribute error pass. 

\subsubsection{halfFataldefect \#4: Theorem 2.1 Exp3
Analysis ignore slightly action action spacemay change ness}\label{ux534aux81f4ux547dux7f3aux9677-4-theorem-2.1-ux7684-exp3-ux5206ux6790ux5ffdux7565ux52a8ux4f5cux7a7aux95f4ux53efux53d8ux6027}

\textbf{ness}: to Spring special differentness build model insufficient\\
\textbf{Location}: §2.2.1\\
\textbf{concreteincorrect}: Exp3 Standard AnalysisAssumptionsfixed action action space. Spring
memory bankadd/deletemakes \(|\mathcal{M}_t^{\text{dormant}}|\) can change. use 
\(|\mathcal{S}|\) upper bound cover cover introduce
\(|\mathcal{S}|/|\mathcal{M}_t^{\text{dormant}}|\)
 times loose relax. newincrease/delete arm weightmanagesnotDefinition. \\
\textbf{back result }: regret boundatformuppercorrect (action is upper bound), butmayat actual actual Spring
systemmiddlenon- often loose relax. requiremore refined fineAnalysis come at reason may change action action space. 

\subsubsection{halfFataldefect \#5:
not claim brightAssumptions accumulate accumulate effect}\label{ux534aux81f4ux547dux7f3aux9677-5-ux672aux58f0ux660eux5047ux8bbeux7684ux7d2fux79efux6548ux5e94}

\textbf{ness}: honestdegree bare character \\
\textbf{Location}: full text \\
\textbf{concreteincorrect}: tofew 11 not claim brightAssumptions (see §6.3
 converge total table), wheremany (Fisher consistentness、noiseeachtowardsameness、global
\(L\)-smooth smooth often quantity level 、self-reconstruction\toLabelnoise pass broadcast)toTheoremhas effect nesstocloseimportant. \\
\textbf{back result }: many \texttt{{[}rigorousProof{]}} Labelshould drop level is 
\texttt{{[}rigorousProof (atwithlowernot via verifyAssumptionslower){]}}. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{eight、 attach addProblem: smallbuttrue}\label{ux516bux9644ux52a0ux95eeux9898ux5c0fux4f46ux771fux5b9e}

\subsubsection{8.1 Theorem 2.2
 leave regret exponential}\label{theorem-2.2-ux7684ux9057ux61beux6307ux6570}

Theorem 2.2 claims leave regret upper boundis 
\(O(T^{\frac{d_{\text{eff}}}{d_{\text{eff}}+2}})\). documentwillitsdescribeis 
\(O(T^{(d+1)/(d+2)})\), butto out table reach formula is
\(O(T^{d_{\text{eff}}/(d_{\text{eff}}+2)})\). for
\(d_{\text{eff}}=1\): \(T^{1/3} \approx T^{0.333}\); for
\(d_{\text{eff}}=3\): \(T^{3/5} = T^{0.6}\). thisanddocumentIn the abstract, say ''exponential convention is 
\(4/5\)`` (\(=0.8\))not symbol ------\(T^{0.6} \neq T^{0.8}\). 

\subsubsection{8.2 equation (1) boxed
 table reach formula missing}\label{ux65b9ux7a0b-1-ux7684-boxed-ux8868ux8fbeux5f0fux7f3aux5931}

§1.3.1 middle PL
ConditionslowerrecursioninequalityDerivationmiddle, \(\Delta_t = \mathcal{L}(\theta_t) - \mathcal{L}(\theta^*)\)
atexpect recurrent push formula middle out present \(\Delta_{t+1}\) and
\(\Delta_t\), but solve recursionconcreteStepby omit slightly (`` solve recursioninequality obtain to describe rate''). for a claims''rigorousProof''result, tofewshouldto out recursion close combined solve or cite use Standard result. 

\subsubsection{\texorpdfstring{8.3 §1.2.1 middle \(H_T = \sum 1/t\)
 out present atTheorem Statementmiddle}{8.3 §1.2.1 middle H\_T = \textbackslash sum 1/t out present atTheorem Statementmiddle}}\label{ux4e2d-h_t-sum-1t-ux51faux73b0ux5728ux5b9aux7406ux9648ux8ff0ux4e2d}

\(H_T = \sum_{t=1}^T 1/t = O(\log T)\). thisis adjust and number, butTheorem Statementmiddlewill
\(H_T\) retain atseparate factor middle (th 153 lines), howeverbackatlower a (th 157 lines) turn is 
\(O(\log T)\)
asymptoticform. thiskind''Theorem Statementmiddlecontainsnotsimplifies table reach formula '' do methodatmathematical write action middlenot rule paradigm ------must what at stated describe middlethen write become 
\(O(\log T)\), must what willasymptoticform release atCorollarymiddle. 

\subsubsection{8.4 reference in the Ghadimi
 assemble write }\label{ux53c2ux8003ux6587ux732eux4e2dux7684-ghadimi-ux62fcux5199}

referencemiddle write is ``Ghadimi, S., \& Lan, G.''. correct character is
\textbf{Ghadimi} (None error), butth two action one character atfull text middleis \textbf{Guanghui
Lan} (through often shrink write is G. Lan). this invert is NOT incorrect, but cite use timeshouldensure action onecorrect. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ nine 、mostfinal evaluate separate}\label{ux4e5dux6700ux7ec8ux8bc4ux5206}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 evaluate separate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 evaluate language 
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Problemimportantness & 8/10 & Spring
convergencenessAnalysis confirm actual isnon-convex optimization+Online Learningfront along cross cross \\
theory wild core & 9/10 & three independentProblemunifiedat reason is male core vigorous vigorous \\
Theorem 1.1-1.4 (convergence rate) & 5/10 & 1.1-1.3 correctbut 1.4 has order of magnitudeincorrect, 1.2 
log T is artifact \\
Theorem 2.1-2.3 (regret bound) & 5/10 & Exp3 applicationcorrectbut ignore slightly Spring
 special differentness (may change space) \\
Theorem 3.1-3.4 (monotonicity) & \textbf{2/10} & 3.1
 counter example is NOT Derivation out (illusion number +cycles pseudo generation code), 3.2-3.4 suspend float in Fisher consistentnessupper \\
honestdegree & 4/10 & \texttt{{[}rigorousProof{]}} Label over degreeuse------tofew 3
 TheoremAssumptionsnotbyverify \\
Symbol/Conceptconsistentness & 5/10 & \(\Delta_s\) timebetween-dependencein Q2 vs Q3 'sbetween contradiction contradiction \\
\textbf{total overall } & \textbf{4/10} & \textbf{Theorem 1.4 has order of magnitudeincorrect, Theorem
3.1
 counter example does not matchConstructionnessProofStandard, many coreAssumptionsnot claim bright. atwhenfrontformlowernot adapt combined action is alreadycompleterigorousAnalysis. } \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{ ten 、correctionRecommendation (such as result mustpreserveswhenfrontframework)}\label{ux5341ux4feeux6b63ux5efaux8baeux5982ux679cux8981ux4fddux7559ux5f53ux524dux6846ux67b6}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{Theorem 1.4}: correction \(\sigma\sqrt{\sum\alpha_t^2}\)
 order of magnitudecompute (\(O(\sqrt{\log T})\) andnon-
 \(O(T^{1/4})\)), heavynewDerivationlower bound. orone change use often number step longAnalysiswithavoid
 \(\log T\) because factor. sametimesupplementFrom highdimensional \(\Theta\) to 1D \(x\)
 dimension convention simple Lemma. 
\item
 \textbf{Theorem 1.2}: change using Ghadimi-Lan (2013)
 often number step longAnalysis, willconvergence rate to \(O(1/\sqrt{T})\) (None
 \(\log T\)). supplement Spring loss functionglobal
 \(L\)-smooth smooth often number estimate (toa small amount of level). 
\item
 \textbf{Theorem 3.1}: must what (a) From Spring
 self-reconstructiontraininggradient descentmiddletruecurrentlyDerivation out probability change -ization number value, must what (b) willLabel drop level is 
 \texttt{{[} heuristic estimate guess measure {]}}------ counter example physical intuitionmayisto, butwhenfrontnothas mathematical support. deleteorheavy write pseudo generation code. 
\item
 \textbf{Theorem 2.1}: explicit formula at reason may change action action space------oroneProof
 \(|\mathcal{M}_t^{\text{dormant}}| \leq |\mathcal{S}|\)
 loose relax is tight, oroneuseat reason may change arm quantity bandit framework (such as Cesa-Bianchi et
 al.~ sleeping experts). 
\item
 \textbf{global}: add a ''notverifyAssumptions'' clear single, will-dependencestrongAssumptions
 \texttt{{[}rigorousProof{]}} resultbright confirm markNoteis 
 \texttt{{[}rigorousProof (atwithlowerAssumptionslower:...){]}}. willmany 
 \texttt{{[}partrigorous+Conjecture{]}}
 resultmiddle''rigorous''partand''Conjecture''part use explicit formula boundary line separate separate. 
\item
 \textbf{Symbol/Concept}: unified \(\Delta_s\) timebetween-dependenceness------must what atfull text use
 \(\Delta_s(t)\), must what in Q2 leave regret Analysismiddle change is 
 \(\bar{\Delta}_s = \frac{1}{T}\sum_t \Delta_s(t)\). 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{review person sign character: Nonemust sign character ------upper describe each attack point allismayindependentverify. }

\emph{P.S.
thisdocumentatstructureupper overall present out honest agree figure (``Honest Critique''Label、open problems list table), butconcrete execute linesmiddlehas manyathonestdegreeandmathematicalrigorousness'sbetween grand trench. most strict heavyis
Theorem 1.4 order of magnitudeincorrectand Theorem 3.1
cycles'' counter example ''. this some is NOT study technique not end ------ it s isreviewprocessmiddleshouldbydiscoverincorrect. correctionback, this copy Analysiswillisnon-convex optimization+Spring
framework cross cross Domainmiddle a has valuecontribution. }

\end{document}
