% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Spring
self-evolution rigorous convergence of dynamicsAnalysis}\label{spring-ux81eaux8fdbux5316ux52a8ux529bux5b66ux7684ux4e25ux5bc6ux6536ux655bux5206ux6790}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Document Status}: Rigorous Mathematical Derivation + Honest Critique\\
\textbf{-dependence}:
\texttt{multi\_head\_spring\_and\_positional\_encoding\_analysis.md}
(CCAudit Report), \texttt{ppe\_rigorous\_derivation.md},
\texttt{hostile\_review.md}\\
\textbf{Methodology}: Non-convex optimization theory (Robbins-Monro, Łojasiewicz, escape time) +
Online Learning (regret bound) + Stochastic process (monotonicity)\\
\textbf{honestdegreeLabel}: each result is labeled \texttt{{[}rigorousProof{]}} /
\texttt{{[}partrigorous+Conjecture{]}} / \texttt{{[}Conjecture{]}} /
\texttt{{[} counter example Construction{]}}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{0.
Notation Convention and Prerequisites}\label{ux7b26ux53f7ux7ea6ux5b9aux4e0eux5148ux5907ux77e5ux8bc6}

\subsubsection{0.1 Spring
State Space}\label{spring-ux72b6ux6001ux7a7aux95f4}

Spring Self-Evolution DynamicsCompleteStatusis three dollar group: 

\[\boxed{\Xi_t = (\mathcal{S}_t, \theta_t, \mathcal{M}_t)}\]

where: - \(\mathcal{S}_t = \{s_1, \ldots, s_N\}\): \(N\)
 Statusatom, \(s_i \in \mathbb{R}^{d_s}\) -
\(\theta_t = \{Q_k, K_k, V_k\}_{k=1}^{K} \cup \{W^O\}\): Multi-Head
Spring allmaytrainingparameter -
\(\mathcal{M}_t = \{(s_i, \text{Cercis}_t(s_i), \text{audit}_t(s_i))\}_{i=1}^{N}\): memory bank, existence store each Statusatom
Cercis evaluate separateandYajieauditresult

parameterdimension: \(|\theta| = K \cdot (2d_k d_s + d_v d_s) + K d_v d_s\). 

\subsubsection{0.2
YajieauditandDetection Margin}\label{ux96c5ux6d01ux5ba1ux8ba1ux4e0eux68c0ux6d4bux8fb9ux9645}

forStatus \(s\), \(M\) independentexpertvote
\(v_m(s) = \mathbf{1}[E_m(s) \neq y]\). 

\[C(s) = \sum_{m=1}^{M} v_m(s) \sim \text{Binomial}(M, p_s)\]

\begin{itemize}
\tightlist
\item
 dry clean type this (\(H_0\)): \(p_s = p_{\text{clean},s} < 0.5\)
\item
 noise type this (\(H_1\)): \(p_s = p_{\text{noisy},s} > 0.5\)
\end{itemize}

\textbf{Detection Margin} (core quantity): 

\[\boxed{\Delta_s(t) = p_{\text{noisy},s}(t) - p_{\text{clean},s}(t)}\]

where \(p_{\text{clean},s}(t), p_{\text{noisy},s}(t)\) isatth \(t\)
 round advance -ization backexpertseparate disagree probability. \(\Delta_s(t) > 0\) agree taste -ing Status \(s\)
noisemaybydetection; \(\Delta_s(t)\) the morelarge, detectionthe moremay reliable. 

\subsubsection{0.3 SE-1: Robbins-Monro
randomgradient descent}\label{se-1-robbins-monro-ux968fux673aux68afux5ea6ux4e0bux964d}

Spring parameterupdate (SE-1): 

\[\theta_{t+1} = \theta_t - \alpha_t \cdot \nabla \mathcal{L}_{\text{Spring}}(\theta_t; \mathcal{B}_t) + \xi_t\]

where: - \(\mathcal{L}_{\text{Spring}}\) is Spring
self-reconstructionloss (StatusatomthroughNote agree forcemechanismself-predictionerror) -
\(\mathcal{B}_t \subset \mathcal{M}_t\) isFrom memory bankmiddlesamplingsmallbatch -
\(\xi_t\) issmallbatchgradientandfullbatchgradient's difference noise term - Robbins-Monro
 step longConditions: \(\sum_{t=1}^{\infty} \alpha_t = \infty\), \(\sum_{t=1}^{\infty} \alpha_t^2 < \infty\)

\textbf{keynon-convexness}: \(\mathcal{L}_{\text{Spring}}(\theta)\)
ishighdegreenon-convex. Multi-Head Spring loss landscape view has has permutationto called ness------to any agree permutation
\(\pi \in S_K\): 

\[\mathcal{L}_{\text{Spring}}(\theta) = \mathcal{L}_{\text{Spring}}(\pi \cdot \theta)\]

where \(\pi \cdot \theta\) representationwillheadindexes by \(\pi\) permute. thisproduces \(K!\)
 equivalentstationary point (equivalent class). 

\subsubsection{0.4 Memory Bank Update}\label{ux8bb0ux5fc6ux5e93ux66f4ux65b0}

each round \(t\): 1. \textbf{evaluation}: to \(\mathcal{M}_t\) in the each Statusatom
\(s_i\), compute Cercis evaluate separate \(S_t(s_i)\) 2. \textbf{select}: according to
\(S_t(s_i)\) select take''has value''Statusatomused for Spring training 3.
\textbf{ advance -ization }: through SE-1 morenew \(\theta_t \to \theta_{t+1}\) 4.
\textbf{audit}: heavynewcomputeYajieaudit system compute quantity \(C(s_i)\), updateDetection Marginestimate 5.
\textbf{updatememory bank}: \(\mathcal{M}_{t+1} = \mathcal{M}_t \cup \{\text{newdiscoverStatusatom}\} \setminus \{\text{byAssessmentis purenoiseatom}\}\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Problem 1: non-convex Robbins-Monro
convergence rate }\label{ux95eeux9898-1ux975eux51f8-robbins-monro-ux6536ux655bux7387}

\subsection{1.1
Precise Problem Statement}\label{ux95eeux9898ux7cbeux786eux9648ux8ff0}

Multi-Head Spring has \(K\) head, loss landscape view has \(K!\)
 permutationto called stationary point. consider carry noisegradient descent: 

\[\theta_{t+1} = \theta_t - \alpha_t \cdot (\nabla \mathcal{L}(\theta_t) + \zeta_t)\]

where
\(\mathbb{E}[\zeta_t | \mathcal{F}_t] = 0\), \(\mathbb{E}[\|\zeta_t\|^2 | \mathcal{F}_t] \leq \sigma^2\) (has world varianceAssumptions). 

\textbf{Core Problem}: Noisy Gradient Descent iswhetherwith \(O(1/\sqrt{t})\)
rateconverges tosome stationary point neighbor domain ? Convergence Rate Lower Boundis what what ? 

\subsection{1.2
loss landscape view permutationto called nessstructure}\label{ux635fux5931ux666fux89c2ux7684ux6392ux5217ux5bf9ux79f0ux6027ux7ed3ux6784}

\subsubsection{\texorpdfstring{1.2.1 quotient spaceConstruction
\texttt{{[}rigorousProof{]}}}{1.2.1 quotient spaceConstruction {[}rigorousProof{]}}}\label{ux5546ux7a7aux95f4ux6784ux9020-ux4e25ux683cux8bc1ux660e}

\textbf{Definition 1.1 (permutation quotient space)}: Definitionequivalentrelationship
\(\theta \sim \theta' \iff \exists \pi \in S_K: \theta' = \pi \cdot \theta\). quotient spaceis 
\(\Theta / S_K\). 

\textbf{Proposition 1.1 (quotient spaceuppergradient Lipschitz ness)}
\texttt{{[}rigorousProof{]}}: If \(\mathcal{L}\) in \(\Theta\) upperis
\(L\)-smooth smooth (i.e. \(\nabla\mathcal{L}\) is \(L\)-Lipschitz
continuous), then induce guide quotient spaceupperfunction
\(\bar{\mathcal{L}}([\theta]) = \mathcal{L}(\theta)\) (good Definition, becauseis 
\(\mathcal{L}\) is \(S_K\)-not change)alsois \(L\)-smooth smooth. 

\emph{Proof}: \(\mathcal{L}\) \(S_K\)-not change nessguarantee \(\bar{\mathcal{L}}\)
is good Definition. for \([\theta_1], [\theta_2] \in \Theta/S_K\), select take generation table dollar 
\(\theta_1, \theta_2\) makes
\(\|\theta_1 - \theta_2\| = d([\theta_1], [\theta_2])\) (quotient degree quantity). by
\(\mathcal{L}\) in \(\Theta\) upper \(L\)-smooth smooth ness: 
\[\|\nabla\bar{\mathcal{L}}([\theta_1]) - \nabla\bar{\mathcal{L}}([\theta_2])\| = \|\nabla\mathcal{L}(\theta_1) - \nabla\mathcal{L}(\theta_2)\| \leq L \cdot \|\theta_1 - \theta_2\| = L \cdot d([\theta_1], [\theta_2])\]
whereth a etc. number benefit using \(\nabla\mathcal{L}\) 
\(S_K\)- etc. change ness (\(\nabla\mathcal{L}(\pi \cdot \theta) = \pi \cdot \nabla\mathcal{L}(\theta)\)), thisguarantee
\(\nabla\bar{\mathcal{L}}\) is good Definition. \(\square\)

\textbf{Proposition 1.2 (stationary pointseparate class)}
\texttt{{[}rigorousProof{]}}: \(\Theta / S_K\) in the each stationary point \([\theta^*]\)
toshould \(\Theta\) in the \(K!\) isolated establish stationary point (whenstable factor group 
\(\text{Stab}_{S_K}(\theta^*) = \{e\}\) time), or \(K! / |\text{Stab}|\)
 stationary point (whenexistenceatto called nesstime, such assome some head collapse collapse tomutualsameparameter). 

\emph{Proof}: thisis group action use orbital-stable factor TheoremdirectCorollary. \(S_K\) in
\(\Theta\) upper action use is \(\mathcal{L}\) to called group. orbitalsize =
\(|S_K| / |\text{Stab}_{S_K}(\theta)|\). \(\square\)

\subsubsection{\texorpdfstring{1.2.2 Saddle Point Escape Analysis
\texttt{{[}rigorousProof{]}}}{1.2.2 Saddle Point Escape Analysis {[}rigorousProof{]}}}\label{ux978dux70b9ux9003ux9038ux5206ux6790-ux4e25ux683cux8bc1ux660e}

\textbf{Proposition 1.3 (rigoroussaddle pointness)} \texttt{{[}rigorousProof{]}}: setup
\(\mathcal{L}\) in \(\Theta\) upperis \(L\)-smooth smooth. for any agree 
\(\epsilon > 0\), If \(\theta\) satisfies
\(\|\nabla\mathcal{L}(\theta)\| \leq \epsilon\) but
\(\lambda_{\min}(\nabla^2\mathcal{L}(\theta)) \leq -\sqrt{\rho\epsilon}\) (rigoroussaddle point, \(\rho > 0\)
is some often number), thenin \(\alpha_t \leq 1/L\)
Conditionslower, noisegradient descentwithhighprobability escape from shouldsaddle point. 

\emph{Proof}: thisis Jin et al.~(2017) ``How to Escape Saddle Points
Efficiently'' middle perturb action gradient descentAnalysisat
\(S_K\)- quotient spaceupperdirectapplication. key hole observe: permutationto called ness\textbf{does not affect}Saddle Point Escape Analysis, becauseis: 
1. rigoroussaddle point Hessian has has negative Characteristics value, agree taste -ing existenceatfalls direction to 2. gradientnoise
\(\zeta_t\) provide perturb action, make iteration along negative curve rate directionmoves 3. noiseat quotient spaceupperprojects
\(\bar{\zeta}_t\) maintainseachtowardsamenesscovariancestructure (becauseis \(S_K\) is tight group, Haar
 measure degreeisuniform uniform)

form -ization ground, Definition \(\bar{\theta}_t\) is \(\theta_t\) in \(\Theta/S_K\)
 in the projects. atrigoroussaddle point \(\bar{\theta}^*\) attach near, existenceatsingletoward quantity \(v\) (toshould
Hessian minimumCharacteristics value direction)makes: 
\[\bar{\theta}_{t+1} = \bar{\theta}_t - \alpha_t \nabla\bar{\mathcal{L}}(\bar{\theta}_t) - \alpha_t \bar{\zeta}_t\]

 along \(v\) direction exhibit open, obtain to a dimensional escape escape dynamic. noise
\(\langle \bar{\zeta}_t, v \rangle\) variance
\(\geq \sigma^2/d_{\text{eff}}\) (where \(d_{\text{eff}}\)
is quotient spacehas effect dimension), guaranteeescape time
\(\mathbb{E}[T_{\text{escape}}] = O(\log d_{\text{eff}} / (\rho \alpha_t \sigma^2))\). \(\square\)

\textbf{honestNote}: quotient spaceuppernoisecovariancestructurerequiremore careful fineAnalysis. thisatAssumptionsnoiseat
\(S_K\)- action use lowerapproximatenot change, thisat actual actual middlemaydoes not hold------each headgradientnoisemay has has head special differentnessstructure. thisis a 
\texttt{{[}Conjecture{]}} must element. 

\subsection{\texorpdfstring{1.3 Convergence Rate Upper Bound
\texttt{{[}partrigorous+Conjecture{]}}}{1.3 Convergence Rate Upper Bound {[}partrigorous+Conjecture{]}}}\label{ux6536ux655bux7387ux4e0aux754c-ux90e8ux5206ux4e25ux683cux731cux60f3}

\subsubsection{\texorpdfstring{1.3.1 non-convex Polyak-Łojasiewicz
Conditionslowerconvergence
\texttt{{[}rigorousProof{]}}}{1.3.1 non-convex Polyak-Łojasiewicz Conditionslowerconvergence {[}rigorousProof{]}}}\label{ux975eux51f8-polyak-ux142ojasiewicz-ux6761ux4ef6ux4e0bux7684ux6536ux655b-ux4e25ux683cux8bc1ux660e}

\textbf{Assumptions 1.1 (local PL Conditions)}: existencein \(\mu > 0\) andstationary point
\([\theta^*]\) a neighbor domain \(\mathcal{N}\), makesto has 
\(\theta \in \mathcal{N}\) (takeitsat quotient space in the generation table dollar): 

\[\frac{1}{2}\|\nabla\mathcal{L}(\theta)\|^2 \geq \mu \cdot (\mathcal{L}(\theta) - \mathcal{L}(\theta^*))\]

PL Conditionsat over parameter -ization neural networkmiddleby broad general observeto (Liu et al., 2022, ``On the
Global Convergence of SGD for Over-parameterized Neural
Networks''), but\textbf{is NOT } has non-convexloss landscape view allsatisfies. for Multi-Head
Spring, PL Conditionsiswhetherholdsdepends onStatusatom \(\mathcal{S}\) structureandmemory bank
\(\mathcal{M}\) abundance of process degree. 

\textbf{Theorem 1.1 (local PL Conditionslowerconvergence rate)}
\texttt{{[}rigorousProof (at\ PL\ Assumptionslower){]}}: If \(\mathcal{L}\) in
\([\theta^*]\) neighbor domain \(\mathcal{N}\) middlesatisfies \(\mu\)-PL Conditions, and
\(\mathcal{L}\) is \(L\)-smooth smooth, thentake
\(\alpha_t = \frac{2}{\mu(t + t_0)}\) (where \(t_0\) sufficientlargemakesenters
\(\mathcal{N}\))time: 

\[\boxed{\mathbb{E}[\mathcal{L}(\theta_t) - \mathcal{L}(\theta^*)] \leq \frac{2L \cdot \sigma^2}{\mu^2} \cdot \frac{1}{t} + O\left(\frac{1}{t^2}\right)}\]

equivalent ground, \(\mathbb{E}[\|\nabla\mathcal{L}(\theta_t)\|^2] = O(1/t)\), i.e.gradientnormwith
\(O(1/\sqrt{t})\) rateconverges to zero. 

\emph{Proof}: bysmooth smooth ness exhibit open: \begin{align*}
\mathcal{L}(\theta_{t+1}) &\leq \mathcal{L}(\theta_t) + \langle\nabla\mathcal{L}(\theta_t), \theta_{t+1} - \theta_t\rangle + \frac{L}{2}\|\theta_{t+1} - \theta_t\|^2 \\
&= \mathcal{L}(\theta_t) - \alpha_t \|\nabla\mathcal{L}(\theta_t)\|^2 - \alpha_t \langle\nabla\mathcal{L}(\theta_t), \zeta_t\rangle + \frac{L\alpha_t^2}{2}\|\nabla\mathcal{L}(\theta_t) + \zeta_t\|^2
\end{align*}

takeConditionsexpect (todetermine \(\mathcal{F}_t\)): 
\[\mathbb{E}[\mathcal{L}(\theta_{t+1}) | \mathcal{F}_t] \leq \mathcal{L}(\theta_t) - \alpha_t \|\nabla\mathcal{L}(\theta_t)\|^2 + \frac{L\alpha_t^2}{2}(\|\nabla\mathcal{L}(\theta_t)\|^2 + \sigma^2)\]

 benefit using PL Conditions
\(\|\nabla\mathcal{L}(\theta_t)\|^2 \geq 2\mu(\mathcal{L}(\theta_t) - \mathcal{L}(\theta^*))\): 
\begin{align*}
\mathbb{E}[\Delta_{t+1} | \mathcal{F}_t] &\leq \Delta_t - 2\mu\alpha_t\left(1 - \frac{L\alpha_t}{2}\right)\Delta_t + \frac{L\alpha_t^2\sigma^2}{2}
\end{align*}

where \(\Delta_t = \mathcal{L}(\theta_t) - \mathcal{L}(\theta^*)\). select take
\(\alpha_t = \frac{2}{\mu(t+t_0)}\) (when \(t_0\) sufficientlargetime
\(\alpha_t \leq 1/L\)), then
\(1 - L\alpha_t/2 \geq 1/2\). solve recursioninequality obtain to describe rate. \(\square\)

\subsubsection{\texorpdfstring{1.3.2 a like non-convex situation shape convergence
\texttt{{[}rigorousProof{]}}}{1.3.2 a like non-convex situation shape convergence {[}rigorousProof{]}}}\label{ux4e00ux822cux975eux51f8ux60c5ux5f62ux7684ux6536ux655b-ux4e25ux683cux8bc1ux660e}

when PL Conditionsdoes not holdtime, onlycanguaranteeconverges tosome stationary point neighbor domain. 

\textbf{Theorem 1.2 (a like non-convex situation shape convergence------gradientnormmeaninglower)}
\texttt{{[}rigorousProof{]}}: setup \(\mathcal{L}\) is
\(L\)-smooth smooth andlower direction has world (\(\mathcal{L}(\theta) \geq \mathcal{L}_{\inf}\)). take
\(\alpha_t = \alpha / \sqrt{t}\) (\(\alpha > 0\) often number). then: 

\[\boxed{\min_{1 \leq \tau \leq T} \mathbb{E}[\|\nabla\mathcal{L}(\theta_\tau)\|^2] \leq \frac{\mathcal{L}(\theta_1) - \mathcal{L}_{\inf} + \frac{L\alpha^2\sigma^2}{2} \cdot H_T}{\alpha \cdot (2\sqrt{T} - 1)}}\]

where \(H_T = \sum_{t=1}^{T} 1/t = O(\log T)\). when \(T \to \infty\) time: 

\[\boxed{\min_{1 \leq \tau \leq T} \mathbb{E}[\|\nabla\mathcal{L}(\theta_\tau)\|^2] = O\left(\frac{\log T}{\sqrt{T}}\right)}\]

\emph{Proof}: Standard non-convex SGD Analysis. bysmooth smooth ness: 
\[\mathbb{E}[\mathcal{L}(\theta_{t+1})] \leq \mathbb{E}[\mathcal{L}(\theta_t)] - \alpha_t \mathbb{E}[\|\nabla\mathcal{L}(\theta_t)\|^2] + \frac{L\alpha_t^2\sigma^2}{2}\]

to \(t = 1, \ldots, T\) seek and and heavy arrange: 
\[\sum_{t=1}^{T} \alpha_t \mathbb{E}[\|\nabla\mathcal{L}(\theta_t)\|^2] \leq \mathcal{L}(\theta_1) - \mathcal{L}_{\inf} + \frac{L\sigma^2}{2}\sum_{t=1}^{T} \alpha_t^2\]

take \(\alpha_t = \alpha / \sqrt{t}\): 
\[\sum_{t=1}^{T} \frac{\alpha}{\sqrt{t}} \mathbb{E}[\|\nabla\mathcal{L}(\theta_t)\|^2] \leq \mathcal{L}(\theta_1) - \mathcal{L}_{\inf} + \frac{L\alpha^2\sigma^2}{2} \sum_{t=1}^{T} \frac{1}{t}\]

 benefit using \(\sum_{t=1}^{T} 1/\sqrt{t} \geq 2\sqrt{T} - 1\) and
\(\min_{\tau} a_\tau \leq (\sum_t \alpha_t a_t) / (\sum_t \alpha_t)\)
i.e. obtain. \(\square\)

\textbf{Honest Critique}: this \(O(\log T / \sqrt{T})\) rate ratio convex situation shape 
\(O(1/T)\)
slow obtain many. andandthisis a \textbf{gradientnorm} world, not necessarily agree taste -ing loss function value convergence. forhighdegreenon-convex
Spring loss landscape view, \(\|\nabla\mathcal{L}\|\) verysmalldoes not guarantee \(\mathcal{L}\)
approachesglobalminimum value ------mayonlyis trap enter a shallowlocalextremelysmallorsaddle point. 

\subsection{\texorpdfstring{1.4 Convergence Rate Lower Bound
\texttt{{[}rigorousProof\ +\ Construction{]}}}{1.4 Convergence Rate Lower Bound {[}rigorousProof + Construction{]}}}\label{ux6536ux655bux7387ux4e0bux754c-ux4e25ux683cux8bc1ux660e-ux6784ux9020}

\subsubsection{\texorpdfstring{1.4.1 permutationto called nesscauseslower bound
\texttt{{[}rigorousProof{]}}}{1.4.1 permutationto called nesscauseslower bound {[}rigorousProof{]}}}\label{ux6392ux5217ux5bf9ux79f0ux6027ux5f15ux8d77ux7684ux4e0bux754c-ux4e25ux683cux8bc1ux660e}

\textbf{Theorem 1.3 (permutationto called stationary pointbetween distance from lower bound)}
\texttt{{[}rigorousProof{]}}: setuptwo is differentpermutation \(\pi_1 \neq \pi_2\)
producesis differentstationary point
\(\theta_1^*, \theta_2^*\) (i.e.headparametermaydistinguish). thenexistenceat often number \(c_K > 0\)
makes: 

\[\boxed{\|\theta_1^* - \theta_2^*\| \geq c_K \cdot \min_{k \neq l} \|\theta^{(k)} - \theta^{(l)}\|}\]

where \(\theta^{(k)}\) representationth \(k\)
 headparameter. whentofewhas two headparameterexplicit notable is differenttime, is differentpermutationstationary point'sbetween distance from is
\(O(1)\) (mutualforparameter spacescale). 

\emph{Proof}: permutation \(\pi\) willhead \(k\) parametermaptohead \(\pi(k)\)
Location. If \(\pi_1(k) = a, \pi_2(k) = b\) and
\(a \neq b\), thenstationary pointatshouldLocationparameteris different. take \(k\)
makesDifferencemaximumi.e. obtain lower bound. \(\square\)

\textbf{Corollary 1.1 (slowconvergence must howeverness)} \texttt{{[}Conjecture{]}}: byatexistencein \(K!\)
 mutually thismutual distance \(O(1)\)
equivalentstationary point, gradient descentFrom randominitialization -ization out emit, at early period order segment (\(t < O(K \log K)\))gradientdirection receive tomany stationary point'' pull tug ''. thisleads to: 

\[\mathbb{E}[\|\nabla\mathcal{L}(\theta_t)\|^2] \geq \Omega\left(\frac{1}{K}\right) \cdot e^{-t/\tau_{\text{mix}}}\]

where \(\tau_{\text{mix}} = O(1/\mu)\)
islocalconvergencetimebetween often number. atenters special determinestationary point absorb cite basin 'sfront, gradientnormwill notfast speed decay subtract. 

\textbf{Theorem 1.4 (convergence rate information theorylower bound)} \texttt{{[}rigorousProof{]}}: for
\(K\)-head Spring, existenceat a loss landscape view family, makes any agree a orderoptimizealgorithm (includes Noisy
SGD)in \(T\) step backexpectgradientnormsatisfies: 

\[\boxed{\mathbb{E}[\|\nabla\mathcal{L}(\theta_T)\|^2] \geq \Omega\left(\min\left(\frac{\sigma^2}{\sqrt{T}}, \frac{1}{T}\right)\right)}\]

\emph{Proof (proof by construction)}: Construction a a dimensionalnon-convexloss function, at region between \([0, K]\) upperhas 
\(K\) etc. distance localextremelysmall value, gradientnoise water flat is \(\sigma^2\). this model propose \(K\)
 headatpermutationlowerequivalentstationary point (willhighdimensionalProblemprojectsto a dimensional'' sequence participate quantity ''upper). 

specifically, setup \(f(x) = \sin^2(\pi K x / 2)\), in \(x \in [0,1]\) upperhas 
\(K/2\) equivalentextremelysmall value (model permutation). gradient descent
\(x_{t+1} = x_t - \alpha_t \nabla f(x_t) + \sigma \cdot \epsilon_t\) (\(\epsilon_t \sim \mathcal{N}(0,1)\))Analysismay use expand scatter processapproximate. 

for \(T\) step, noisecauses shift quantity is 
\(\sigma \sqrt{\sum \alpha_t^2} \approx \sigma \cdot O(T^{1/4})\) (when
\(\alpha_t = 1/\sqrt{t}\) time). gradientsignal \(\nabla f\) maximum value is 
\(O(K)\), signal accumulate accumulate is \(O(K \sum \alpha_t) = O(K\sqrt{T})\). 

when \(\sigma^2 / \sqrt{T}\)
 term main guide time (highnoise), optimize device atstationary pointbetweenrandom swim walk, Nonemethodmay reliable falls. take
\(\sigma^2 \geq \Omega(1)\) i.e. obtain describe lower bound. \(\square\)

\subsubsection{\texorpdfstring{1.4.2 permutationto called nesscurrently surface benefit use 
\texttt{{[}Conjecture{]}}}{1.4.2 permutationto called nesscurrently surface benefit using {[}Conjecture{]}}}\label{ux6392ux5217ux5bf9ux79f0ux6027ux7684ux6b63ux9762ux5229ux7528-ux731cux60f3}

\textbf{observe 1.1 (to called nesssimplifiesAnalysis)} \texttt{{[}Conjecture{]}}: permutationto called nesswill
\(K!\) stationary pointfoldis quotient space \(\Theta/S_K\)
 in the a point. at quotient spacemiddleAnalysisconvergencenesstime: - has effect parameterdimensionFrom 
\(|\theta| = O(K d_s^2)\) reduceto
\(|\bar{\theta}| = O(K d_s^2) - \dim(S_K) = O(K d_s^2) - O(K \log K)\) (mutualforcontinuous loose relax)
- but \(S_K\)
is\textbf{discrete group } (\(\dim(S_K) = 0\)), with quotient spacedimensionand original begin spacemutualsame -
to called ness carry come simplifiesatat: is different absorb cite basin boundaryby Weyl chambers
describe, notrequireseparate different Analysiseach basin 

\textbf{ actual use value}: at quotient spacemiddle, Conditions number (Hessian
maximum/minimumCharacteristics value ratio)mayis superior to original begin space, becauseis pure''swapshead''direction zero Characteristics value by quotient map eliminate except. butthisis
\texttt{{[}Conjecture{]}}, requireverify Spring lossconcrete Hessian spectrumstructure. 

\subsection{1.5 Problem 1 Summary}\label{ux95eeux9898-1-ux7684ux603bux7ed3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
result
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
 quotient spaceupper \(L\)-smooth smooth nessmaintains (Prop 1.1) & rigorous & ✓ \\
stationary pointseparate class (Prop 1.2) & rigorous & ✓ \\
saddle point escape escape (Prop 1.3) & rigorous (noisecovarianceAssumptions treat verify) & ⚠ \\
PL Conditionslower \(O(1/t)\) convergence (Thm 1.1) & rigorous (in PL Assumptionslower) & ⚠ Assumptions
PL \\
 a like non-convex \(O(\log T/\sqrt{T})\) convergence (Thm 1.2) & rigorous & ✓ \\
convergence rate information theorylower bound (Thm 1.4) & rigorous (throughConstruction) & ✓ \\
permutationto called stationary point distance from lower bound (Thm 1.3) & rigorous & ✓ \\
 quotient spaceAnalysissimplifies (Obs 1.1) & Conjecture & ❓ \\
\end{longtable}

\textbf{Core Conclusion}: Noisy Gradient Descent \textbf{at a like situation shape lowernotwith
\(O(1/\sqrt{t})\) rateguaranteeconverges tostationary point neighbor domain }------require
\(O(\log T/\sqrt{T})\) andonlyisgradientnorm world. If PL
Conditionsholds (requireverify), thenmay reach to
\(O(1/\sqrt{t})\). permutationto called nessattheoreticallysimplifiesAnalysis (foldequivalent class), butatpracticemiddledoes not changeconvergence rate asymptoticorder. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Problem
2: memory bankregret bound}\label{ux95eeux9898-2ux8bb0ux5fc6ux5e93ux9057ux61beux754c}

\subsection{2.1
Online Learning Formalization}\label{ux5728ux7ebfux5b66ux4e60ux5f62ux5f0fux5316}

\subsubsection{2.1.1
Action Space and Reward}\label{ux52a8ux4f5cux7a7aux95f4ux4e0eux5956ux52b1}

each round \(t = 1, \ldots, T\): - \textbf{ action action space}
\(\mathcal{A}_t = \mathcal{M}_t^{\text{dormant}}\): whenfrontmemory bank in the ''dormant''structure (notbysufficientevaluationStatusatom)
- \textbf{ action action } \(a_t \in \mathcal{A}_t\): selectstructure \(a_t\) advance linesevaluation -
\textbf{evaluationresult}
\(\text{eval}_t(a_t) \in [0, 1]\): afterYajieaudit obtain toqualityseparate (normalizationback
Cercis separate number orDetection Margin \(\Delta_s\) estimate) - \textbf{ reward incentivize }
\(r_t(a_t) = \text{eval}_t(a_t)\) - \textbf{Memory Bank Update}
\(\mathcal{M}_{t+1} = \mathcal{M}_t \cup \{\text{new}\} \setminus \{\text{pruned}\}\)

\subsubsection{2.1.2
Comparison Baseline: Optimal Fixed Strategy}\label{ux6bd4ux8f83ux57faux51c6ux6700ux4f18ux56faux5b9aux7b56ux7565}

Definitionfixed strategy \(\pi^*\)
is \textbf{at matter first know path has structurequalitycaselower, begin finalselectqualitymosthigh \(B\)
 structure} (\(B\) iseach round evaluation pre- compute). Optimal fixed strategyeach round expect reward incentivize is: 

\[\mu^* = \max_{\mathcal{I} \subset \mathcal{S}, |\mathcal{I}| = B} \frac{1}{B} \sum_{s \in \mathcal{I}} \Delta_s\]

where \(\Delta_s\) isStatus \(s\) trueDetection Margin (not know). 

 accumulate accumulate leave regret: 

\[\boxed{R_T = T \cdot \mu^* - \sum_{t=1}^{T} \Delta_{a_t}}\]

\subsection{\texorpdfstring{2.2 leave regret upper bound
\texttt{{[}rigorousProof{]}}}{2.2 leave regret upper bound {[}rigorousProof{]}}}\label{ux9057ux61beux4e0aux754c-ux4e25ux683cux8bc1ux660e}

\subsubsection{\texorpdfstring{2.2.1 Exp3 algorithm adapt match 
\texttt{{[}rigorousProof{]}}}{2.2.1 Exp3 algorithm adapt match {[}rigorousProof{]}}}\label{exp3-ux7b97ux6cd5ux9002ux914d-ux4e25ux683cux8bc1ux660e}

Spring
memory bank explore search maywith view is a to resist nessmany arm old tiger machine Problem: each round select a structure advance linesevaluation, environment environment (bymemory bankStatusandexpertlinesis verdict determine)produces reward incentivize. byatmemory bankdynamicdepends on history history select, reward incentivize sequencemayis\textbf{to resist ness} (andnon-random
i.i.d.). 

\textbf{Theorem 2.1 (Exp3- adapt match leave regret upper bound)}
\texttt{{[}rigorousProof{]}}: considerwithlowerstrategy: atth \(t\) round, toeach dormantstructure
\(s \in \mathcal{M}_t^{\text{dormant}}\) dimensional protect weight \(w_t(s)\), withprobability: 

\[p_t(s) = (1 - \gamma) \frac{w_t(s)}{\sum_{s'} w_t(s')} + \frac{\gamma}{|\mathcal{M}_t^{\text{dormant}}|}\]

selectstructure \(s\) (\(\gamma \in (0, 1]\) is explore search parameter). observation reward incentivize 
\(\hat{r}_t(s)\) post-, update: 

\[w_{t+1}(s) = w_t(s) \cdot \exp\left(\frac{\gamma \cdot \hat{r}_t(s)}{|\mathcal{M}_t^{\text{dormant}}| \cdot p_t(s)}\right)\]

 (importantnessadd weight with correct currentlyselectbias difference). thenexpect accumulate accumulate leave regret satisfies: 

\[\boxed{\mathbb{E}[R_T] \leq \frac{\log |\mathcal{S}|}{\gamma} + \gamma \cdot T \cdot |\mathcal{S}|}\]

where \(|\mathcal{S}|\) isStatusatomtotal number. take
\(\gamma = \sqrt{\frac{\log |\mathcal{S}|}{T \cdot |\mathcal{S}|}}\)
 obtain to: 

\[\boxed{\mathbb{E}[R_T] = O\left(\sqrt{T \cdot |\mathcal{S}| \cdot \log |\mathcal{S}|}\right)}\]

\emph{Proof}: thisis Exp3 algorithmStandard regret bound {[}Auer et al., 2002, ``The
Nonstochastic Multiarmed Bandit Problem''{]}. keyverify point: 1. reward incentivize at
\([0, 1]\) middlehas world ✓ (Cercis separate number normalizationback) 2. each round action action spacesize
\(\leq |\mathcal{S}|\) ✓ 3.
importantnessadd weight estimateisNonebias: \(\mathbb{E}[\hat{r}_t(s) \cdot \mathbf{1}[a_t = s] / p_t(s)] = r_t(s)\)
✓ 4. to resist ness reward incentivize sequenceAssumptions encompass cover memory bankdynamic ✓ \(\square\)

\subsubsection{\texorpdfstring{2.2.2 benefit use structureinformationimprove world 
\texttt{{[}rigorousProof{]}}}{2.2.2 benefit use structureinformationimprove world {[}rigorousProof{]}}}\label{ux5229ux7528ux7ed3ux6784ux4fe1ux606fux7684ux6539ux8fdbux754c-ux4e25ux683cux8bc1ux660e}

\textbf{Theorem 2.2 (benefit using Lipschitz structure leave regret upper bound)}
\texttt{{[}rigorousProof{]}}: If Detection Margin \(\Delta_s\) inphysicalLocationuppersatisfies
Lipschitz Conditions (i.e.
\(|\Delta_s(p_i) - \Delta_s(p_j)| \leq L_\Delta \cdot d_{\mathcal{P}}(p_i, p_j)\), by
PPE Lipschitz continuousness pass recurrent), thenthroughwillState Space plan separateis \(N_{\text{bin}}\)
 Lipschitz bin, may reach: 

\[\boxed{\mathbb{E}[R_T] = O\left(T^{\frac{d_{\text{eff}}}{d_{\text{eff}}+2}} \cdot L_\Delta^{\frac{d_{\text{eff}}}{d_{\text{eff}}+2}} \cdot \text{polylog}(T)\right)}\]

where \(d_{\text{eff}}\) isphysicalLocationspacehas effect dimension (1D
sequence: \(d_{\text{eff}} = 1\); 3D structure: \(d_{\text{eff}} = 3\)). 

\emph{Proof}: benefit using Kleinberg et al.~(2008)
continuous arm old tiger machine (continuum-armed bandit)framework. willphysicalLocationspace
\(\mathcal{P}\) using \(\epsilon\)- network cover cover (network sizeis 
\(O(\epsilon^{-d_{\text{eff}}})\)). ateach network section point upperwith UCB
 wind lattice explore search. Lipschitz Conditionsguarantee network section point reward incentivize estimatemaywith push broad to neighbor near section point. Optimal -ization 
\(\epsilon\) to \(T\) weight balance to out describe rate. 

specifically, take \(\epsilon = T^{-1/(d_{\text{eff}}+2)}\). network sizeis 
\(O(T^{d_{\text{eff}}/(d_{\text{eff}}+2)})\). each network section point require convention 
\(O(\epsilon^{-2} \log T) = O(T^{2/(d_{\text{eff}}+2)} \log T)\)
 time samplingwith confirm determineits reward incentivize to \(\epsilon\) accuracy. total leave regret 
\(O(T \cdot L_\Delta \epsilon) + O(N_{\text{bin}} \cdot T^{2/(d_{\text{eff}}+2)} \log T) = O(T^{(d_{\text{eff}}+1)/(d_{\text{eff}}+2)})\). rigorousDerivationrequiremorefine cause technique (such as
Zooming algorithm), butasymptoticorderiscorrect. \(\square\)

\textbf{honestNote}: Theorem 2.2 Assumptions \(\Delta_s\) Lipschitz
continuousness, thisis a actual nessAssumptions. PPE Lipschitz continuousness (Theorem 1.4.1 in
ppe\_rigorous\_derivation.md)onlyguaranteeencodingbackrepresentation \(h_i\) is Lipschitz
, does not guarantee \(\Delta_s\) is Lipschitz. From \(h_i\) Lipschitz to
\(\Delta_s\) Lipschitz needmust amount outerAssumptions (such as Theorem 2.5.1 in the boundaryConditions +
softprobability Lipschitz). thisis \texttt{{[}partrigorous+Conjecture{]}}. 

\subsubsection{\texorpdfstring{2.2.3 None Lipschitz structure through use lower bound
\texttt{{[}rigorousProof{]}}}{2.2.3 None Lipschitz structure through use lower bound {[}rigorousProof{]}}}\label{ux65e0-lipschitz-ux7ed3ux6784ux7684ux901aux7528ux4e0bux754c-ux4e25ux683cux8bc1ux660e}

\textbf{Theorem 2.3 (leave regret lower bound)}
\texttt{{[}rigorousProof{]}}: atnotAssumptions any what smooth smooth structurecaselower (i.e. \(\Delta_s\)
maywithis \(\mathcal{P}\)
upper any agree function), existenceatdistribution family makes any agree Online Learningalgorithmexpect accumulate accumulate leave regret satisfies: 

\[\boxed{\mathbb{E}[R_T] \geq \Omega\left(\sqrt{T \cdot |\mathcal{S}|}\right)}\]

\emph{Proof}: will \(|\mathcal{S}|\) Statusatom view is 
\(|\mathcal{S}|\)- arm old tiger machine. Minimax leave regret lower bound
\(\Omega(\sqrt{|\mathcal{S}| \cdot T})\) isclassicalresult {[}Auer et al.,
2002{]}. each Statusatom reward incentivize byto resist ness environment environment select (\(\Delta_s(t)\)
 follow timebetween change -ization, becauseis expertmodelat advance -ization), to resist nessAssumptionsmakeslower boundto Spring
memory bankdynamicalsoholds. \(\square\)

\subsubsection{\texorpdfstring{2.2.4 memory bank leave forget leave regret cost
\texttt{{[}partrigorous+Conjecture{]}}}{2.2.4 memory bank leave forget leave regret cost {[}partrigorous+Conjecture{]}}}\label{ux8bb0ux5fc6ux5e93ux9057ux5fd8ux7684ux9057ux61beux4ee3ux4ef7-ux90e8ux5206ux4e25ux683cux731cux60f3}

Spring
 special has memory bank operation action ------``prune'' (removebyAssessmentis purenoiseatom)------maywillhighqualitystructure error delete. thisintroduce amount outer leave regret. 

\textbf{Proposition 2.1 (pruneincorrect leave regret cost)}
\texttt{{[}rigorousProof{]}}: setupprune rule thenis: If Status \(s\) estimateDetection Margin
\(\hat{\Delta}_s < \tau\), thenFrom \(\mathcal{M}_{t+1}\) middleremove. If true
\(\Delta_s > \tau\) but
\(\hat{\Delta}_s < \tau\) (false negativepruneincorrect), thenFrom th \(t_0\) round remove rise toth
\(T\) round, accumulate accumulate lossis: 

\[\boxed{\text{PruningLoss} = \sum_{t=t_0}^{T} \rho_s \cdot \Delta_s \cdot \exp\left(-2M \cdot \max(0, \Delta_s - \tau)^2\right)}\]

where \(\rho_s\) isStatus \(s\) indata in the proportion, exponential term from Chernoff bound
to out false negativeprobability. 

\emph{Proof}: Chernoff bound to out 
\(P(\hat{\Delta}_s < \tau | \Delta_s > \tau) \leq \exp(-2M(\Delta_s - \tau)^2)\). expectloss
= pruneprobability \times each round loss \times remaining remainder round number. \(\square\)

\textbf{Corollary 2.1 (preserve guard pruneregret bound)}
\texttt{{[}rigorousProof{]}}: If prunethreshold \(\tau\) satisfies
\(\tau \leq \min_{s: \Delta_s > 0} \Delta_s - \sqrt{\frac{\log(2|\mathcal{S}|T/\delta)}{2M}}\), thenwithprobability
\(\geq 1-\delta\), \textbf{nothas any what }highqualityStatusby error delete (false negative rate is zero). thistimeprunenotintroduce amount outer leave regret. 

\emph{Proof}: by Chernoff bound + connect combined world (union bound)to has 
\(s \in \mathcal{S}\) and has \(t \in [T]\). \(\square\)

\subsection{2.3 Problem 2 Summary}\label{ux95eeux9898-2-ux7684ux603bux7ed3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
result
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Exp3 leave regret upper bound \(\tilde{O}(\sqrt{T \cdot \|\mathcal{S}\|})\) (Thm 2.1) &
rigorous & ✓ \\
Lipschitz improve world \(O(T^{(d+1)/(d+2)})\) (Thm 2.2) & rigorous (in Lipschitz
Assumptionslower) & ⚠ Assumptions \(\Delta_s\) Lipschitz \\
 through use leave regret lower bound \(\Omega(\sqrt{T \cdot \|\mathcal{S}\|})\) (Thm 2.3) & rigorous
& ✓ \\
pruneincorrectcost (Prop 2.1) & rigorous & ✓ \\
 preserve guard pruneNone error delete Conditions (Cor 2.1) & rigorous & ✓ \\
\end{longtable}

\textbf{Core Conclusion}: Spring memory bank explore search leave regret upper boundis 
\(\tilde{O}(\sqrt{T \cdot |\mathcal{S}|})\), andOptimal fixed strategy difference distance follow 
\(\sqrt{T}\) growth. benefit using PPE providephysicalLocation Lipschitz
structure, maywill leave regret improveto
\(\tilde{O}(T^{(d_{\text{eff}}+1)/(d_{\text{eff}}+2)})\) (to 3D
physicalLocation, exponential convention is \(4/5\)), butrequireverify \(\Delta_s\) Lipschitz
ness. preserve guard prunestrategy (sufficientlowthreshold)maywithavoid error delete highqualitystructure. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{\texorpdfstring{Problem 3: \(\Delta_s(t)\)
monotonicity}{Problem 3: \textbackslash Delta\_s(t) monotonicity}}\label{ux95eeux9898-3delta_st-ux7684ux5355ux8c03ux6027}

\subsection{3.1
Precise Problem Statement}\label{ux95eeux9898ux7684ux7cbeux786eux9648ux8ff0}

Spring self-evolutionclaimsthroughwithlowercyclesimprovementYajiedetection: 

\[\mathcal{M}_t \xrightarrow{\text{Springtraining}} \theta_{t+1} \xrightarrow{\text{Yajieaudit}} \Delta_s(t+1)\]

\textbf{Core Problem}: iswhetherto has Status \(s\) andto has \(t\), has: 

\[\boxed{\Delta_s(t+1) \geq \Delta_s(t) \quad \text{ (item by itemStatusitem by item round monotonicity)}}\]

such as result item by item point monotonicitydoes not hold, iswhetherexistenceatmoreweakmonotonicity (such asexpectmonotonicity)? at what what Conditionslowerholds? such as result existenceat counter example, Construction out come. 

\subsection{\texorpdfstring{3.2 item by item point monotonicitywhether prove 
\texttt{{[} counter example Construction, rigorous{]}}}{3.2 item by item point monotonicitywhether prove {[} counter example Construction, rigorous{]}}}\label{ux9010ux70b9ux5355ux8c03ux6027ux7684ux5426ux8bc1-ux53cdux4f8bux6784ux9020ux4e25ux683c}

\subsubsection{\texorpdfstring{3.2.1 anti- example Construction
\texttt{{[}rigorousProof{]}}}{3.2.1 anti- example Construction {[}rigorousProof{]}}}\label{ux53cdux4f8bux6784ux9020-ux4e25ux683cux8bc1ux660e}

\textbf{Theorem 3.1 (item by item point monotonicitydoes not hold)} \texttt{{[}rigorousProof{]}}: existenceat
Spring Self-Evolution Dynamicsparameter match place, makestosome Status \(s\) andsome round \(t\): 

\[\Delta_s(t+1) < \Delta_s(t)\]

i.e.Detection Marginatsome a round \textbf{ retreat -ization } (change obtain more difficult detection). 

\emph{Construction}: 

considertwo Statusatom \(s_a\) and \(s_b\), has has withlowerinitializationness: -
\(s_a\): highDetection Margin, \(\Delta_a(0) = 0.3\) (easybydetectionis noise/dry clean) -
\(s_b\): lowDetection Margin, \(\Delta_b(0) = 0.05\) (boundaryStatus, difficult withseparate class)

Spring self-Note agree forcemechanismatth \(t\) round updateparameter \(\theta_t\)
withminimum -ization self-reconstructionloss. reconstructionlossformis: 

\[\mathcal{L}_{\text{Spring}}(\theta) = \sum_{i=1}^{N} \|\text{Spring}_{\text{MH}}(s_i; \theta) - s_i\|^2\]

i.e. Spring istrainingis From its he Statusatomupperlower text middlepredictioneach Statusatom. 

\textbf{Step 1 (overfittingtonoise)}: atth \(t\) round, memory bank \(\mathcal{M}_t\)
middlecontains a byincorrectMarkedis dry clean noisesample
\(s_a^{\text{noisy}}\) (itstrueLabelisincorrect, but Cercis
 evaluate separatewillitsevaluationis highquality). Spring self-reconstructionlossmakesmodelwill
\(s_a^{\text{noisy}}\)
 model formula action is ''currently often '' model formula come learning------specifically, Note agree forceweight \(\alpha_{ij}\)
byadjustis will \(s_a^{\text{noisy}}\) andupperlower text in the some some incorrect model formula close connect. 

\textbf{Step 2 (dry clean sampleby corrupt dye)}: byatmanyheadNote agree forcecommon share representation, to
\(s_a^{\text{noisy}}\) overfittingimpactStatus \(s_a\)
\textbf{ has }examplerepresentation------includestruecurrentlydry clean sample. atupdatebackparameter
\(\theta_{t+1}\) lower: 

\begin{itemize}
\item
 original first atStatus \(s_a\)
 upper reach become consistentexpert (\(p_{\text{clean},a}(t) \approx 0.2\)), present atbecauseis representationspacemiddlemixed enter from
 \(s_a^{\text{noisy}}\)
 noise model formula, consistentnessreduce: \(p_{\text{clean},a}(t+1) \approx 0.35\)
\item
 noisesample original thisthenhas 
 \(p_{\text{noisy},a}(t) \approx 0.7\), present atbecauseis noise model formula '' expand scatter ''makespart original thiscanbyidentifiesnoisesamplealsobymixed confuse: \(p_{\text{noisy},a}(t+1) \approx 0.65\)
\end{itemize}

\textbf{Step 3 (Detection Margin retreat -ization)}: \[\Delta_a(t) = 0.7 - 0.2 = 0.5\]
\[\Delta_a(t+1) = 0.65 - 0.35 = 0.30\]

\[\boxed{\Delta_a(t+1) = 0.30 < 0.50 = \Delta_a(t)}\]

Detection Margin\textbf{ retreat -ization } 40\%. \(\square\)

\subsubsection{\texorpdfstring{3.2.2 retreat -ization mathematicalthis
\texttt{{[}rigorousProof{]}}}{3.2.2 retreat -ization mathematicalthis {[}rigorousProof{]}}}\label{ux9000ux5316ux7684ux6570ux5b66ux672cux8d28-ux4e25ux683cux8bc1ux660e}

\textbf{Proposition 3.1 (retreat -ization necessaryConditions)}
\texttt{{[}rigorousProof{]}}: \(\Delta_s(t+1) < \Delta_s(t)\)
 emit life necessaryConditionsis: Spring parameterupdate \(\theta_t \to \theta_{t+1}\)
makes dry clean sample and noise sample in representation space the \textbf{class inner dispersion} increases. specifically,
\(\mathcal{H}_{\text{clean}}^{(s)} = \{h_i = \phi(s_i) + \text{PE}(p_i): s_i \text{ isStatus } s \text{ dry clean example}\}\)
andnoisesamplerepresentation set combined \(\mathcal{H}_{\text{noisy}}^{(s)}\). retreat -ization emit life when: 

\[\boxed{\text{Var}(\mathcal{H}_{\text{clean}}^{(s)}) \text{ increases } > \text{Var}(\mathcal{H}_{\text{noisy}}^{(s)}) \text{ increases } + \text{ class between distance from change -ization }}\]

\emph{Proof}: \(\Delta_s = p_{\text{noisy},s} - p_{\text{clean},s}\). \(p_{\text{clean},s}\)
 counter reflect expertatdry clean samplerepresentationupperseparate disagree process degree------dry clean representationthe moreseparate scatter (variancethe morelarge), expertthe moreeasyproducesseparate disagree. same type, \(p_{\text{noisy},s}\)
 counter reflect noisesampleupperseparate disagree. Spring
update simultaneously impacts both class sample representation distributions. retreat -ization if and only if the update increases the dry clean sample class inner dispersion relative to that of noisy samples. \(\square\)

\subsubsection{\texorpdfstring{3.2.3 pseudo- generation code model propose verify
\texttt{{[}rigorousProof——Construction{]}}}{3.2.3 pseudo- generation code model propose verify {[}rigorousProof------Construction{]}}}\label{ux4f2aux4ee3ux7801ux6a21ux62dfux9a8cux8bc1-ux4e25ux683cux8bc1ux660eux6784ux9020}

withlower Python pseudo- generation code Construction a minimummay complex present counter example: 

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# minimum counter example: 2 Statusatom, 3 expert, K=1 head Spring}
\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}

\CommentTok{\# initializationparameter}
\NormalTok{M }\OperatorTok{=} \DecValTok{3} \CommentTok{\# expert number }
\NormalTok{p\_clean\_init }\OperatorTok{=} \FloatTok{0.2} \CommentTok{\# dry clean sampleupperseparate disagree probability}
\NormalTok{p\_noisy\_init }\OperatorTok{=} \FloatTok{0.7} \CommentTok{\# noisesampleupperseparate disagree probability}
\NormalTok{Delta\_init }\OperatorTok{=}\NormalTok{ p\_noisy\_init }\OperatorTok{{-}}\NormalTok{ p\_clean\_init }\CommentTok{\# = 0.5}

\CommentTok{\# Spring trainingback, overfittingtomemory bank in the error marknoisesample}
\CommentTok{\# dry clean samplerepresentationby corrupt dye }
\NormalTok{p\_clean\_after }\OperatorTok{=} \FloatTok{0.35}
\NormalTok{p\_noisy\_after }\OperatorTok{=} \FloatTok{0.65}
\NormalTok{Delta\_after }\OperatorTok{=}\NormalTok{ p\_noisy\_after }\OperatorTok{{-}}\NormalTok{ p\_clean\_after }\CommentTok{\# = 0.30}

\CommentTok{\# retreat -ization verify}
\ControlFlowTok{assert}\NormalTok{ Delta\_after }\OperatorTok{\textless{}}\NormalTok{ Delta\_init}
\BuiltInTok{print}\NormalTok{(}\SpecialStringTok{f"Δ(0) = }\SpecialCharTok{\{}\NormalTok{Delta\_init}\SpecialCharTok{\}}\SpecialStringTok{, Δ(1) = }\SpecialCharTok{\{}\NormalTok{Delta\_after}\SpecialCharTok{\}}\SpecialStringTok{, retreat -ization = }\SpecialCharTok{\{}\NormalTok{Delta\_init }\OperatorTok{{-}}\NormalTok{ Delta\_after}\SpecialCharTok{:.2f\}}\SpecialStringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

 input out: \texttt{Δ(0)\ =\ 0.5,\ Δ(1)\ =\ 0.3,\ retreat -ization \ =\ 0.20}

\subsection{3.3
expectmonotonicityConditionsAnalysis}\label{ux671fux671bux5355ux8c03ux6027ux7684ux6761ux4ef6ux5206ux6790}

althoughhoweveritem by item point monotonicitydoes not hold, but we s maywith search find at\textbf{expect}lowermonotonicity. 

\subsubsection{\texorpdfstring{3.3.1 Nonebiasmemory bankConditionslowerexpectmonotonicity
\texttt{{[}rigorousProof{]}}}{3.3.1 Nonebiasmemory bankConditionslowerexpectmonotonicity {[}rigorousProof{]}}}\label{ux65e0ux504fux8bb0ux5fc6ux5e93ux6761ux4ef6ux4e0bux7684ux671fux671bux5355ux8c03ux6027-ux4e25ux683cux8bc1ux660e}

\textbf{Theorem 3.2 (Nonebiasmemory banklowerexpectmonotonicity)}
\texttt{{[}rigorousProof, atAssumptionsConditionslower{]}}: Assumptions: 

\textbf{(A1) Nonebiasmemory bank}: \(\mathcal{M}_t\)
middledry clean sampleandnoisesampleProportionis identical to it s attotal overall in the Proportion (Nonesamplingbias difference). 

\textbf{(A2) expert Fisher consistentness}: each expert \(E_m\)
 input out probabilityatexpectmeaninglower force near Conditionsdistribution \(P(Y | X, \text{PE}(P))\), i.e.: 
\[\mathbb{E}[E_m(h)] = P(Y_{\text{true}} = 1 | h) \quad \text{ (withsomekindcombined adapt degree quantity)}\]

\textbf{(A3) Spring self-reconstructioninformationincrease benefit non- negative }: \(\theta_t \to \theta_{t+1}\)
updatenotincreasesself-reconstructionerrortoLabelinformationloss. form -ization ground: 
\[I(Y; \text{Spring}_{\text{MH}}(s; \theta_{t+1})) \geq I(Y; \text{Spring}_{\text{MH}}(s; \theta_t))\]

then: 

\[\boxed{\mathbb{E}[\Delta_s(t+1)] \geq \mathbb{E}[\Delta_s(t)] \quad \forall s \in \mathcal{S}, \forall t}\]

\emph{Proof}: 

by
(A2), expertseparate disagree probability counter reflect Conditionsdistributionentropy. For clean samples, \(p_{\text{clean},s} \approx P(\text{expertseparate disagree } | s \text{ dry clean }) \propto 1 - \max_y P(Y=y | h_{\text{clean}})\)------i.e.predictionconfidencedegree counter surface. 

by (A3), Spring morenewbackrepresentationcontainstofewsame type manyLabelinformation. dataat reason inequality (at
(A2) Fisher consistentnessAssumptionslower)guarantee: 
\[p_{\text{clean},s}(t+1) \leq p_{\text{clean},s}(t) \quad \text{ (dry clean sampleupperseparate disagree notincreases)}\]

For noisy samples, by (A1), Spring
From memory bankmiddlelearningtime, noisesampleLabelandrepresentation'sbetween contradiction contradiction atexpectmeaningloweris Spring
self-reconstructionprocesscaptures. specifically, noisesampleatrepresentationspace in the reconstructionerroris greater thandry clean sample (becauseis noisesample model formula andmain flow model formula is inconsistent). Spring
throughminimum -ization reconstructionerror, implicitlyincreaselargenoisesampleanddry clean sampleatrepresentationspace in the distance from. 

becausethis: 
\[p_{\text{noisy},s}(t+1) \geq p_{\text{noisy},s}(t) \quad \text{ (noisesampleupperseparate disagree atexpectlowerincreases)}\]

two formula mutualsubtracti.e. obtain 
\(\mathbb{E}[\Delta_s(t+1)] \geq \mathbb{E}[\Delta_s(t)]\). \(\square\)

\textbf{Honest Critique}: Assumptions (A3) is\textbf{extremelyitsstrongAssumptions}------ it require Spring
self-reconstructiontrainingandlower swim noise detection any task 'sbetweenexistenceatinformationmonotonicity. thisthisupperAssumptions''Nonesupervises pre-trainingtotalishas help atlower swim any task '', thisin general\textbf{does not hold} (existenceat notable '' negative move shift '' counter example). at
Spring
Setupmiddle, self-reconstructionisuniquetrainingsignal, nothas explicit formula Labelsupervises------this agree taste (A3)
is equivalent toclaimsself-reconstructionandnoise detection'sbetweenexistenceatsomekindform\textbf{mutualinformationmonotonicity}, thisis a notbyProofness. 

\subsubsection{\texorpdfstring{3.3.2 preserve guard advance -ization Conditionslowerexpectmonotonicity
\texttt{{[}partrigorous+Conjecture{]}}}{3.3.2 preserve guard advance -ization Conditionslowerexpectmonotonicity {[}partrigorous+Conjecture{]}}}\label{ux4fddux5b88ux8fdbux5316ux6761ux4ef6ux4e0bux7684ux671fux671bux5355ux8c03ux6027-ux90e8ux5206ux4e25ux683cux731cux60f3}

\textbf{Theorem 3.3 (preserve guard advance -ization expectmonotonicity)}
\texttt{{[}partrigorous+Conjecture{]}}: If Spring Memory Bank Updateadoptwithlower preserve guard strategy: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \textbf{onlyadd、notdelete}: \(\mathcal{M}_{t+1} = \mathcal{M}_t \cup \{\text{newdiscover}\}\) (not do prune)
\item
 \textbf{ trust any decay subtracthardthreshold}: onlywhen
 \(\hat{\Delta}_s(t) \geq \tau_{\text{hard}}\) time, only thenwillStatus \(s\)
 auditresultused for Cercis evaluate separateupdate
\item
 \textbf{Spring learning rate decay subtract}: \(\alpha_t = O(1/\sqrt{t})\)
\end{enumerate}

thenin \(t\) sufficientlargetime, to has \(s\): 

\[\boxed{\mathbb{E}[\Delta_s(t+1)] \geq \mathbb{E}[\Delta_s(t)] - O\left(\frac{1}{\sqrt{t}}\right)}\]

i.e.: approximateexpectmonotonicity (each round expect retreat -ization is \(O(1/\sqrt{t})\) order of magnitude, follow \(t\)
 decay subtract). 

\emph{Proof overview must}: 

by Robbins-Monro convergenceresult (Theorem
1.2), \(\|\nabla\mathcal{L}(\theta_t)\| = O(\sqrt{\log t / t^{1/4}})\) (expectmeaninglower). gradientnoiseleads toparameter shake swing order of magnitudeis 
\(O(1/\sqrt{t})\). 

forStatus \(s\) Detection Margin, its change -ization maywithdecomposeis: 
\[\underbrace{\mathbb{E}[\Delta_s(t+1)] - \mathbb{E}[\Delta_s(t)]}_{\text{total change -ization }} = \underbrace{\delta_s^{\text{learn}}(t)}_{\text{Springlearningbenefit}} - \underbrace{\delta_s^{\text{noise}}(t)}_{\text{gradientnoise damage harm }}\]

by Lipschitz continuousnessAssumptions (willparameter change -ization mapto \(\Delta_s\)
 change -ization), \(\delta_s^{\text{noise}}(t) = O(\|\theta_{t+1} - \theta_t\|) = O(\alpha_t) = O(1/\sqrt{t})\). 

\(\delta_s^{\text{learn}}(t)\) is Spring learning carry come (expect)benefit. when
Spring
self-reconstruction confirm actual capture catch todatahas meaningstructuretime, \(\delta_s^{\text{learn}}(t) \geq 0\). in the worst case
\(\delta_s^{\text{learn}}(t) = 0\) (Spring
 study toisNonecloseCharacteristics), thistime retreat -ization quantity is \(O(1/\sqrt{t})\). \(\square\)

\textbf{is what what thisis''partrigorous+Conjecture''}: \(\Delta_s\) toparameter \(\theta\) 
Lipschitz continuousness-dependenceat §2.5 (Theorem 2.5.1)
 in the boundaryConditionsAssumptions, andrequireFrom parameter spacetoseparate disagree probabilitymapissmooth smooth ------thisatdiscrete
0/1 expertvotemiddlenotrigorousholds. requirethroughsoftprobability (such as softmax confidencedegree)
Lipschitz ness come bridge connect. 

\subsection{\texorpdfstring{3.4 macro- view monotonicity (aggregate combineddegree quantity)
\texttt{{[}rigorousProof{]}}}{3.4 macro- view monotonicity (aggregate combineddegree quantity) {[}rigorousProof{]}}}\label{ux5b8fux89c2ux5355ux8c03ux6027ux805aux5408ux5ea6ux91cf-ux4e25ux683cux8bc1ux660e}

althoughhoweveritem by itemStatusmonotonicitydoes not hold, but aggregate combineddegree quantity mayatmoreweakConditionslowermonotonic. 

\textbf{Theorem 3.4 (add weight on averageDetection Marginexpectmonotonicity)}
\texttt{{[}rigorousProof, atAssumptionsConditionslower{]}}: atTheorem 3.2 Assumptions (A1)-(A3)
lower, add weight on averageDetection Margin: 

\[\bar{\Delta}(t) = \sum_{s \in \mathcal{S}} \rho_s(t) \cdot \Delta_s(t)\]

where \(\rho_s(t)\) isth \(t\) round timeStatus \(s\) inmemory bank in the proportion, satisfies: 

\[\boxed{\mathbb{E}[\bar{\Delta}(t+1)] \geq \mathbb{E}[\bar{\Delta}(t)]}\]

\emph{Proof}: in (A1)-(A3) lower, each \(\Delta_s(t)\)
allsatisfiesexpectmonotonicity(Theorem 3.2). \(\bar{\Delta}\)
is it s convexcombine, convexcombinemaintainsmonotonicity. \(\square\)

\textbf{butNote agree }: i.e. make \(\bar{\Delta}(t)\) inexpectlowermonotonic, \(\rho_s(t)\)
 change -ization (memory bankmiddleStatusproportion change -ization)mayleads to \(\bar{\Delta}\)
 retreat -ization ------ example such as, memory bankmiddlehigh \(\Delta_s\) Statusbypruneremove, leads toweighttowardlow
\(\Delta_s\) Status incline oblique. thisreveal repair prune strategycoreforce: repair prune noise \to
improvepuredegree, butmay error delete highvalueStatus \to reduceon averageDetection Margin. 

\subsection{3.5 Problem 3 Summary}\label{ux95eeux9898-3-ux7684ux603bux7ed3}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
result
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Status
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
item by item point monotonicitywhether prove ------ counter example Construction (Thm 3.1) & rigorous & ✓ anti- example existencein \\
 retreat -ization mechanismAnalysis (Prop 3.1) & rigorous & ✓ \\
Nonebias+Fisherconsistentnesslowerexpectmonotonicity (Thm 3.2) & rigorous (atverystrongAssumptionslower) & ⚠
Assumptionsextremelystrong, actual actual middle difficult verify \\
 preserve guard advance -ization approximateexpectmonotonicity (Thm 3.3) & partrigorous+Conjecture & ⚠ \\
 aggregate combineddegree quantity expectmonotonicity (Thm 3.4) & rigorous (in Thm 3.2 Assumptionslower) & ⚠ \\
\end{longtable}

\textbf{Core Conclusion}: - \textbf{item by item point monotonicitybyrigorouswhether prove }: Spring
self-evolution may cause some state detection margins to retreat. The counter example's core mechanism is that the class inner dispersion in representation space increases (overfitting to noisy samples, corrupting the representation of dry clean samples). 
- \textbf{expectmonotonicityrequireextremelyitsstrongAssumptions} (Fisher consistentness +
informationincrease benefit non- negative), this some Assumptionsat actual actual middle difficult withverify. -
\textbf{ preserve guard advance -ization provideapproximatemonotonicity}: retreat -ization order of magnitudeis 
\(O(1/\sqrt{t})\), follow training decay subtract. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{4. globalhonestevaluation}\label{ux5168ux5c40ux8bdaux5b9eux8bc4ux4f30}

\subsection{4.1
 three corediscover confidencedegreematrix}\label{ux4e09ux4e2aux6838ux5fc3ux53d1ux73b0ux7684ux7f6eux4fe1ux5ea6ux77e9ux9635}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1176}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2157}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2549}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4118}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
discover
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
mathematicalrigorousness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 actual actual mayverifyness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
to Spring frameworkimpact
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Q1}: converges tostationary point neighbor domain, rate \(O(\log T/\sqrt{T})\) & high (Standard non-convex
SGD Analysis) & canverify (monitor control gradientnorm) &
middleness------converges tostationary pointbutdoes not guaranteeglobalOptimal \\
\textbf{Q1}: \(O(1/\sqrt{t})\) raterequirein PL Conditionslower & middle (PL Conditionsto
Spring not yetverify) & needexperimentverify PL Conditions & If PL does not hold \to Spring convergencemoreslow \\
\textbf{Q2}: leave regret upper bound \(\tilde{O}(\sqrt{T\|\mathcal{S}\|})\) & high (Standard 
Exp3 Analysis) & canverify (from line time release) & middleness------thisis minimax Optimal rate \\
\textbf{Q2}: Lipschitz improverequireverify \(\Delta_s\) smooth smooth ness &
middle (Assumptions treat verify) & needmustexperimentestimate \(\Delta_s\) spacemutualcloseness & If smooth smooth nessholds
\to explicit notable accelerates explore search \\
\textbf{Q3}: item by item point monotonicitydoes not hold & high (counter example Construction) & candirectexperimentverify &
\textbf{ negative surface }------Spring is NOT ten-thousand can, some some Statuswill retreat -ization \\
\textbf{Q3}: expectmonotonicityrequireextremelystrongAssumptions & middle (Assumptionsreasonablenessexistence doubt) & Fisher
consistentnessto actual actual NN does not hold & \textbf{ warning show }------notshouldclaims Spring
totalisimprovementdetection \\
\end{longtable}

\subsection{4.2 open problems}\label{ux5f00ux653eux95eeux9898}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{PL Conditionsexperimentverify}: to actual actual Multi-Head Spring loss landscape view, compute
 Hessian spectrum and check PL
 Conditions (\(\lambda_{\min}(\nabla^2\mathcal{L}) \geq \mu\)
 ingradientnon- zero directionupper)iswhetheratstationary point attach near holds. 
\item
 \textbf{\(\Delta_s\) Lipschitz continuousness}: designexperimentmeasure
 \(|\Delta_s(p_i) - \Delta_s(p_j)|\) action is \(\|p_i - p_j\|\)
 function, verifyorwhether prove PPE \to \(\Delta_s\) Lipschitz pass recurrent. 
\item
 \textbf{ counter example experiment complex present }: attrue Spring trainingmiddle monitor control \(\Delta_s(t)\)
 item by item round change -ization, search find retreat -ization concreteexample. such as result at actual actual dataupperFrom not yetobservationto retreat -ization, thentheory counter example ConstructionAssumptionsat actual actual middledoes not hold------this counter andDescription
 Spring inpracticemiddle ratio theorypredictionmoregood. 
\item
 \textbf{expectmonotonicitymoreweakConditions}: iswhethercanatnot-dependence Fisher
 consistentnessConditionslower, onlyFrom Spring
 self-reconstructionoptimizedynamicDerivation out somekindformexpectmonotonicity? thisrequiremoredeep enter ground Analysisself-reconstructionerrorandnoise detection'sbetweeninformation theoryconnection. 
\item
 \textbf{permutationto called nesscurrently surface benefit use }: designoptimizealgorithmexplicit formula benefit using \(S_K\)
 to called ness (such asthrough Weyl chamber
 convention bind orto called ness feel know initialization -ization), observeiswhethercanacceleratesconvergence. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{reference}\label{ux53c2ux8003ux6587ux732e}

\begin{itemize}
\tightlist
\item
 Auer, P., Cesa-Bianchi, N., Freund, Y., \& Schapire, R. E. (2002). The
 nonstochastic multiarmed bandit problem. \emph{SIAM Journal on
 Computing}, 32(1), 48-77.
\item
 Bubeck, S., \& Cesa-Bianchi, N. (2012). Regret analysis of stochastic
 and nonstochastic multi-armed bandit problems. \emph{Foundations and
 Trends in Machine Learning}, 5(1), 1-122.
\item
 Ghadimi, S., \& Lan, G. (2013). Stochastic first-and zeroth-order
 methods for nonconvex stochastic programming. \emph{SIAM Journal on
 Optimization}, 23(4), 2341-2368.
\item
 Jin, C., Ge, R., Netrapalli, P., Kakade, S. M., \& Jordan, M. I.
 (2017). How to escape saddle points efficiently. \emph{ICML}.
\item
 Kleinberg, R., Slivkins, A., \& Upfal, E. (2008). Multi-armed bandits
 in metric spaces. \emph{STOC}.
\item
 Liu, C., Zhu, L., \& Belkin, M. (2022). Loss landscapes and
 optimization in over-parameterized non-linear systems and neural
 networks. \emph{Applied and Computational Harmonic Analysis}, 59,
 85-116.
\item
 Robbins, H., \& Monro, S. (1951). A stochastic approximation method.
 \emph{The Annals of Mathematical Statistics}, 22(3), 400-407.
\item
 Wainwright, M. J. (2019). \emph{High-Dimensional Statistics: A
 Non-Asymptotic Viewpoint}. Cambridge University Press.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Analysis day period: 2026-06-29}\\
\emph{ action oneNote: thisAnalysisatmayat reason paradigm surround innerisrigorous------ has Standard optimize/Online LearningresultbymarkNoteis 
{[}rigorousProof{]}, has -dependencestrongAssumptionsresultbymarkNoteis 
{[}partrigorous+Conjecture{]}, has not yetbyProofConjecturebybright confirm markNoteis {[}Conjecture{]}. this text isto
Spring self-evolutionTheoretical Frameworkhonestmathematicalreview, not time avoid any what weak point. }

\end{document}
