pdfoutput=1
\pdfoutput=1
\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,hyperref,enumitem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\SCX}{SCX}
\newcommand{\rigorFull}{\textbf{[Rigor: Full Proof]}}

\title{SCX OOD Detection Lower Bound}
\author{SCX}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
Proves a finite-expert lower bound for out-of-distribution detection via Spring gatekeeper scores. Minimum detectable OOD shift: delta_min = sigma/sqrt(M). False alarm and detection power both controlled by Hoeffding bounds exponential in M.
\end{abstract}
\subsection{OOD Detection Lower Bound}
\label{subsec:spring_ood}

The CUSUM statistic above monitors temporal drift across a rollout.  Spring
also induces a one-sample OOD test: a sample is suspicious when its gatekeeper
score is atypical relative to scores stored in the in-distribution memory.  To
avoid collision with the CUSUM statistic $S_t$, write $G_t(x)$ for the
per-sample Spring gatekeeper score; this is the score denoted $S_t(x)$ in the
Spring gating notes.

Let $\mathfrak{M}_t$ be the Spring memory bank of accepted in-distribution
samples.  For $M$ expert modules, let $Z_{m,t}(x) \in [0,1]$ be expert $m$'s
normalized validity score for sample $x$ at time $t$, and define the aggregate
gatekeeper score
\begin{equation}
  G_t(x) = \frac{1}{M}\sum_{m=1}^{M} Z_{m,t}(x).
  \label{eq:spring_gate_score}
\end{equation}
The memory bank estimates the in-distribution score center
$\mu_{\mathrm{in},t} = \E[G_t(X)\mid X \sim P_{\mathrm{in},t}]$ by
$\widehat{\mu}_{\mathrm{in},t}$.

\begin{definition}[Spring OOD Hypothesis Test]
\label{def:spring_ood_test}
OOD detection is the binary hypothesis test
\begin{equation}
  H_0: X \sim P_{\mathrm{in},t},
  \qquad
  H_1: X \sim P_{\mathrm{out},t},
\end{equation}
with OOD shift
\begin{equation}
  \delta_{\mathrm{OOD}}
  =
  \left|
  \E[G_t(X)\mid H_1]
  -
  \mu_{\mathrm{in},t}
  \right|.
\end{equation}
For a tolerance $\epsilon > 0$, the ideal calibrated Spring detector is
\begin{equation}
  \varphi_{\epsilon,t}(x)
  =
  \ind{\left|G_t(x)-\mu_{\mathrm{in},t}\right|>\epsilon},
  \label{eq:spring_ood_test}
\end{equation}
where $\varphi_{\epsilon,t}(x)=1$ means ``declare OOD.''  In deployment,
$\mu_{\mathrm{in},t}$ is replaced by $\widehat{\mu}_{\mathrm{in},t}$ from
$\mathfrak{M}_t$; the calibration error is handled in
Corollary~\ref{cor:spring_memory_scaling}.
\end{definition}

\begin{theorem}[SCX OOD Detection Lower Bound]
\label{thm:spring_ood_lower}
Condition on the Spring memory state $\mathfrak{M}_t$.  Assume that
$Z_{1,t}(X),\ldots,Z_{M,t}(X)$ are independent and take values in $[0,1]$.
Under $H_0$, suppose $\E[G_t(X)\mid H_0]=\mu_{\mathrm{in},t}$; under $H_1$,
suppose
$|\E[G_t(X)\mid H_1]-\mu_{\mathrm{in},t}|=\delta_{\mathrm{OOD}}$.
Then, writing
$p_i(\epsilon)=\Pbb(\varphi_{\epsilon,t}(X)=1\mid H_i)$, the test
in~\eqref{eq:spring_ood_test} satisfies
\begin{align}
  p_0(\epsilon)
    &\le 2\exp(-2M\epsilon^2), \label{eq:ood_false_alarm}\\
  p_1(\epsilon)
    &\ge
    1-\exp\!\left(-2M(\delta_{\mathrm{OOD}}-\epsilon)^2\right)
    \label{eq:ood_power_exact}
\end{align}
for every $0<\epsilon<\delta_{\mathrm{OOD}}$.  In particular, if the OOD
shift has margin $\delta_{\mathrm{OOD}}\ge 2\epsilon$, then
\begin{equation}
  \Pbb(\text{detect OOD}\mid \delta_{\mathrm{OOD}}\ge 2\epsilon)
  \ge 1-\exp(-2M\epsilon^2).
  \label{eq:ood_power_margin}
\end{equation}
\end{theorem}

\begin{proof}
The bounds follow directly from Hoeffding's inequality applied to the
$M$ independent gatekeeper scores.
\end{proof}

\end{document}
