% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{审计之剑声明 / The Audit Sword
Declaration}\label{ux5ba1ux8ba1ux4e4bux5251ux58f0ux660e-the-audit-sword-declaration}

\begin{quote}
面壁者不发禁令。面壁者留剑。
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{中文}\label{ux4e2dux6587}

SCX 框架（包括 Yajie、Situs、Spring、State Crystallization
及其全部衍生算法）是开源数学工具，任何人都可以阅读、复现、部署。

我们不禁止任何特定用途。

但我们保留一项不可剥夺的权利：

\textbf{任何组织使用 SCX
框架进行数据质量评估，其评估结果可以通过相同的数学方法被独立第三方进行验证和审计。军事用途不例外。}

这不是限制。这是威慑。审计之剑不区分军民。

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{English}\label{english}

The SCX framework (including Yajie, Situs, Spring, State
Crystallization, and all derivative algorithms) is an open-source
mathematical toolkit. Anyone may read, reproduce, and deploy it.

We do not prohibit any specific use.

We retain one inalienable right:

\textbf{Any organization using the SCX framework for data quality
assessment renders its assessment results verifiable and auditable by
independent third parties through the same mathematical methods.
Military use is not exempt.}

This is not a restriction. It is deterrence. The audit sword does not
discriminate.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{数学依据 / Mathematical
Basis}\label{ux6570ux5b66ux4f9dux636e-mathematical-basis}

Yajie's Theorem 1 guarantees that multi-expert consensus detects label
noise with exponentially decaying false-positive probability. Any party
with access to the same M expert models and the same dataset can
independently reproduce the consensus score C(x) and verify the audit
result. No cooperation from the original auditor is required. The audit
is conducted from outside the organization, using only outputs the
organization voluntarily makes available.

\begin{quote}
定理保证。外部可复现。不需要许可。
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

SCX Research Group 2026-06-30

\end{document}
