% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX 8-Theorem Papers: Mathematical Completeness Review Report}\label{scx-8-theorem-papers-mathematical-completeness-review-report}

\subsection{0. SCX Axiom System Baseline}\label{scx-axiom-system-baseline}

Clarify the core axioms before review:

\begin{itemize}
\tightlist
\item
  \textbf{Theorem 1} (S1): Multi-expert consensus noise detection --- Under assumptions A1-A6, F1 $\geq 1 - (1/\eta)\sum\rho_s\cdot\exp(-2M\Delta_s^2)$, exponential convergence.
\item
  \textbf{Theorem 2} (S2): Weak feature information-theoretic bound --- F1$_{\text{SCX}} \leq$ F1$_{\text{base}} + C_F\cdot\sqrt(\delta/2)$, $\delta=I(\phi(X);S)$.
\item
  \textbf{Theorem 3} (S3): Noise-difficulty indistinguishability --- Label noise and intrinsic difficulty are indistinguishable from observational data. This is SCX's ``uncertainty principle.''
\item
  \textbf{Theorem 4} (S4): Exact constant minimax optimality --- Bahadur-Rao large deviation exact constant.
\item
  \textbf{Cercis Score}: $S(x) = Q(x) + \eta\cdot N(x)$, quality + novelty decomposition.
\item
  \textbf{Situs Operator}: Encodes distributions as metric measure spaces $(X, d_P, \mu_P)$.
\item
  \textbf{Spring SE-1}: Self-evolving gatekeeper, Robbins-Monro convergence guarantee.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Paper-by-Paper Review}\label{paper-by-paper-review}

\subsubsection{1.1 Theorem 5 --- Active Learning (Cercis Optimal Sampling)}\label{theorem-5-active-learning-cercis-optimal-sampling}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Theorem 5.1 (Optimal Distribution): Complete. Gibbs form $p^*(x) \propto \pi(x)\exp(\alpha S(x)/\lambda)$, with uniqueness proof.
- Theorem 5.2 (Optimal Rate): Statement complete, closed form $\rho^* = \eta/(1+\eta)\cdot\sum w_k\cdot V[\Gamma_k]/\sum w_k\cdot\bar{\Gamma}_k$. But key derivation steps contain \textbf{serious logical leaps} (see below).
- Theorem 5.3 (Consistency): Complete. $\sqrt{n}$-consistency, Delta method.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5) --- \textbf{Serious problems exist}
- Lemma 1 (Gibbs solution) derivation is standard variational inference, no problem.
- \textbf{Fatal leap in Theorem 5.2 proof, lines 414--421}: Jumping from $\rho^* = \eta\cdot\sum w_k\bar{\Gamma}_k/(\eta\cdot\sum w_k\bar{\Gamma}_k + \sum w_k V[\Gamma_k])$ to the final form $\rho^* = \eta/(1+\eta)\cdot\sum w_k V[\Gamma_k]/\sum w_k\bar{\Gamma}_k$ claims to use the identity $\sum w_k\bar{\Gamma}_k = ((1+\eta)/\eta)\cdot\sum w_k V[\Gamma_k]$. This identity is labeled as ``holding at the optimum of the variational objective, from equality of marginal benefit and $\eta$-scaled novelty penalty marginal cost.'' \textbf{This is circular reasoning} --- using the definition of $\rho^*$ to derive the closed form of $\rho^*$.
- \textbf{Assumption 1 (Cercis-Gain Affinity)}: $\bar{\Gamma}(x) = \alpha\cdot S(x) + \beta$. This is the cornerstone of the entire derivation but is only justified by ``intuition'' and ``Taylor expansion,'' lacking rigorous proof. This is an unverified empirical assumption.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Correctly uses Cercis Score $S=Q+\eta N$ definition.
- $\eta\to 0$ recovers pure uncertainty sampling, $\eta\to\infty$ recovers pure diversity sampling, consistent with SCX framework.
- But the $\eta/(1+\eta)$ saturation factor does not appear in the original SCX framework --- this is a newly introduced structure.

\textbf{Rigorously Derivable from Existing Axioms}:
- Theorem 5.1 (Gibbs distribution): \checkmark\ Rigorously derived from variational principle (no SCX axioms needed).
- Theorem 5.3 (Consistency): \checkmark\ Rigorously derived from law of large numbers + Delta method.

\textbf{Requires New Assumptions}:
- Theorem 5.2 (Sampling rate): $\times$\ Needs Assumption 1 (Cercis-Gain Affinity) + additional variational structural relations.
- The three-term decomposition of risk-adjustment function $\Phi(\rho)$ (expected benefit, diminishing marginal return, novelty cost $\times$ variance) is an \textbf{ad hoc construction}, not derived from Cercis axioms.

\textbf{Relationship to Other Theorems}: Independent. Cited by AR-Theorem's detectability result (``T5 optimal active learning strategy''), but AR only uses T5 as a black-box tool.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.2 Theorem 6 --- Protocol Game (Audit Protocol Adoption)}\label{theorem-6-protocol-game-audit-protocol-adoption}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Theorem 6.1 (Regret bound): Complete.
- Theorem 6.2 (Nash existence): Complete, standard Glicksberg fixed-point argument.
- Theorem 6.3 (Uniqueness): Complete, explicit $\gamma_{\text{crit}}$ closed form.
- Proposition 6.4 (Distortion): Complete.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Existence proof is standard --- Glicksberg theorem applies directly to mixed-strategy games with continuous payoffs.
- Uniqueness proof's contraction mapping argument is correct, but one \textbf{leap} exists: the sensitivity of the influence function $\varphi_P$ to the number of early adopters is bounded by $\sqrt{2\log N_1/(N_0+1)} + \sigma_\xi/\sqrt{N_0}$, which assumes that the derivative of the UCB confidence radius can be approximated by finite differences. This approximation is asymptotically valid at large $N_0$ but requires additional justification in finite samples.
- The ``influence monotonicity'' lemma's proof relies solely on intuitive reasoning (``increasing $n_P$ decreases standard error''), lacking formal treatment of the randomness in UCB indices.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5)
- The paper claims ``Quality(Q) corresponds to protocol quality $\mu_P$,'' ``Novelty(N) corresponds to UCB exploration bonus.'' This mapping is \textbf{analogical rather than derivational} --- the protocol adoption game is not a natural extension of the Cercis framework but an independently constructed game-theoretic model, post-hoc analogized to SCX.
- $\gamma_{\text{crit}}$ includes the UCB exploration term $\sqrt{2\log N_1/(N_0+1)}$, which has no direct mathematical connection to Cercis's $\eta$ parameter, only conceptual parallelism.

\textbf{Rigorously Derivable from Existing Axioms}:
- Theorem 6.1 (Regret): \checkmark\ Standard UCB analysis.
- Theorem 6.2 (Existence): \checkmark\ Standard game theory.

\textbf{Requires New Assumptions}:
- Theorem 6.3 (Uniqueness): Requires Gumbel noise assumption + influence sensitivity bound (UCB finite-difference approximation).
- The entire model setup (sequential game, Gumbel preference shock, Gaussian prior) is unrelated to SCX axioms.

\textbf{Relationship to Other Theorems}: Indirectly connected to CD-Theorem (information externality), parallel to TS-Theorem's anchor concept. But \textbf{essentially an independent paper} riding the SCX brand rather than deriving from it.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.3 Theorem 7 --- Cross-Domain Partition Preservation}\label{theorem-7-cross-domain-partition-preservation}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar\bigstar$ (5/5)
- Theorem 7.1 (Main bound): Complete, three components clearly separated.
- Theorem 7.2 (Tightness): Complete, constructed tightness sequence.
- Corollary 7.3 (Optimal granularity): Complete.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Five-step proof structure is reasonable.
- \textbf{Step 3 has a potential problem}: The claim ``conditional density is $L_k$-Lipschitz $\Rightarrow$ the entropy functional is $L_k$-Lipschitz w.r.t.\ Wasserstein metric.'' Entropy is not generally a Lipschitz function; additional conditions are required (e.g., log-Sobolev inequality or bounded conditional density on bounded support). Lemma S2.2 (Entropy Lipschitz Property) gives the argument that $H(Y|X=x)$ is $(L/\sigma)$-Lipschitz, but this assumes bounded conditional variance and approximately Gaussian conditional distributions --- an \textbf{important but undeclared assumption}.
- The ``conditioning does not increase average transport cost'' argument (Step 4) is a correct standard result.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar\bigstar$ (5/5)
- Directly departs from Situs conditional mutual information $I(Y;P|S)$.
- Uses Situs encoding and Wasserstein distance, entirely within the SCX framework.
- Explicitly references T7 cross-domain transfer theorem.

\textbf{Rigorously Derivable from Existing Axioms}:
- Main bound: Partially derivable (needs Assumption 1 + entropy Lipschitz condition).
- Asymptotic tightness: \checkmark\ Constructive proof.

\textbf{Requires New Assumptions}:
- Assumption 1 (Bounded Conditional Densities with Lipschitz property).
- Lemma S2.2's entropy Lipschitz argument requires Gaussianity or equivalent conditions.

\textbf{Relationship to Other Theorems}: Cited by CD-Theorem's causal relay theorem as ``T7 cross-domain fidelity.'' One of the \textbf{mathematically most self-consistent} papers among all 8.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.4 CD-Theorem --- Causal Discovery (Noise vs Confounding)}\label{cd-theorem-causal-discovery-noise-vs-confounding}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- CD-Theorem 1 (Strong Inseparability): Complete.
- CD-Theorem 2 (Weak Separability): Statement complete but \textbf{extremely heavily conditional}.
- CD-Theorem 3 (Causal Relay): Complete.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5)
- \textbf{Strong Inseparability proof leap}: Step 1 claims $S(x) = S_0(x) + \eta\cdot\delta_\varepsilon(x) + \beta\cdot\delta_z(x)$, then asserts that when $\eta\perp z$, the joint distribution factorizes so that $\delta_z$ is indistinguishable from effective noise. This is a direct analogical generalization of Theorem 3, but the \textbf{argument is overly sketchy} --- one must prove that the joint distribution of $(S, \nabla S, H_S)$ is indeed identical in both worlds, not just the marginal distribution of $S$. Theorem 3's original proof explicitly constructs joint distribution equivalence for the two worlds; CD-Theorem does not perform this explicit construction.
- \textbf{Problems with the Curl statistic}: $\tau = (1/|\mathcal{X}|)\int\|\nabla\times\hat{\nabla}S\|^2 dx$. This statistic heavily depends on the definition of curl on the Situs manifold in computation. The paper does not provide a concrete construction of the curl operator on the Situs manifold, only vaguely referencing ``the Situs manifold in the SCX framework.'' From a geometric perspective, curl requires a connection structure, while Situs encoding only provides a metric --- \textbf{going from metric to curl requires additional mathematical structure} (the Levi-Civita connection does not necessarily exist on general metric spaces).
- \textbf{Causal Relay}: The chain derivation is reasonable, but the function $g(\varepsilon,\delta)$ with interaction term $C\cdot\varepsilon\cdot\delta$ gives no definition or bounding method for $C$.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Directly extends Theorem 3 to the causal discovery setting.
- The Situs skeleton preservation hypothesis is a significant extension of the Situs operator, but the paper honestly marks it as an open problem.
- Causal Relay correctly references T7.

\textbf{Rigorously Derivable from Existing Axioms}:
- Strong Inseparability: Claims to be a direct corollary of Theorem 3, but the proof is overly sketchy.

\textbf{Requires New Assumptions}:
- Situs skeleton preservation hypothesis (core open problem).
- Existence of curl operator on Situs manifold.
- All three theorems have honest annotations ([Rigorous]/[Conditionally Rigorous]/[Open]).

\textbf{Relationship to Other Theorems}: Forms triangular cross-references with AR-Theorem (gradient anomaly masquerading as causal signal) and HC-Theorem (human cognition breaking symmetry).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.5 FA-Theorem --- Federated Audit (ZK Multi-Party)}\label{fa-theorem-federated-audit-zk-multi-party}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- FA-Theorem 1 ($\varepsilon$-Equivalence): Complete, four error sources.
- FA-Theorem 2 (Info Lower Bound): Complete.
- ZK construction sketch: Complete but only an outline.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5)
- \textbf{Core unsolved problem}: Existence of the Situs homomorphic composition operator $\oplus$. Without it, the $\varepsilon$-Equivalence theorem has only conditional significance. The paper honestly acknowledges this as an open problem.
- \textbf{I-MMSE gap derivation} (Step 2 of Lower Bound): Uses the Guo-Shamai-Verd\'{u} I-MMSE relation, but that relation applies to Gaussian channels, while the Cercis Score here is non-Gaussian. The argument needs generalization or additional defense.
- \textbf{``Worst-case construction''} (Step 3): Claims ``the adversary can adjust $\tilde{S}_k$ to maximize deviation,'' but does not give the concrete form of the adjustment strategy. One must prove the existence of a tampering such that the deviation between $\mathbb{E}[T_{\text{fed}}]$ and $\mathbb{E}[T_{\text{full}}]$ achieves the claimed lower bound.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Correctly uses Cercis Score and Situs encoding.
- The Audit Sword principle is an extension of the SCX framework.
- The lower bound is directly derived from Theorem 3.

\textbf{Rigorously Derivable from Existing Axioms}:
- Info Lower Bound (Theorem 2): \checkmark\ Derived from Theorem 3 + data processing inequality.

\textbf{Requires New Assumptions}:
- Situs homomorphic composition operator $\oplus$ (core open problem).
- Feasibility of ZK circuit construction (cryptographic engineering problem, not a mathematical one).

\textbf{Relationship to Other Theorems}: Cited by CD-Theorem and AR-Theorem as the federated extension of the ``Audit Sword'' principle. \textbf{The most ambitious extension paper}, but also the one most dependent on unsolved mathematical problems (Situs homomorphicity).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.6 AR-Theorem --- Adversarial Robustness (Cercis Taxonomy)}\label{ar-theorem-adversarial-robustness-cercis-taxonomy}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- AR-Theorem 1 (Reduction): Complete, clear conditions.
- AR-Theorem 2 (Detectability): Complete, sample complexity lower bound.
- Conjecture (Phase Transition): Complete statement but merely a conjecture.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5)
- \textbf{Key leap in Reduction theorem}: The proof claims that $||\nabla S||/S \leq \eta$ (``from the definition of the Cercis operator'') yields $A(x) \leq \eta\cdot\varepsilon$. This inequality is \textbf{not a standard result in the SCX framework} --- the Cercis operator is defined as $S=Q+\eta N$ and does not directly constrain $||\nabla S||/S$. This requires the additional assumption that $\nabla Q$ and $\nabla N$ are bounded relative to $Q$ and $N$ in proportion.
- \textbf{Hessian bound in Taylor expansion}: $||H_S||_{\text{op}} \leq L(x)$ is correct (Lipschitz gradient $\Rightarrow$ Hessian bounded), but only when using the spectral norm and the function is twice differentiable.
- \textbf{Detectability theorem's lower bound}: Uses Le Cam's two-point method to derive $\chi^2$ divergence $= \Theta(\gamma^2)$. This derivation is \textbf{extremely sketchy} --- one must explicitly compute the $\chi^2$ divergence of the gradient field distribution under $H_0$ vs $H_1$, which depends on specific assumptions about the distribution of $A(x)$. The paper does not provide this computation.
- \textbf{Upper bound (T5 achievability)}: Claims Theorem 5 provides a matching upper bound, but T5's optimality is about maximizing $I(S;\text{query})$, not about rate optimality for detecting adversarial structure. These are \textbf{different senses of optimality} --- T5 optimizes sample efficiency for acquiring labels, while AR needs sample efficiency for detecting anomalous gradient structure.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ (3/5)
- The three-term decomposition $S=Q+\eta N+\gamma A$ is a major extension of the original Cercis binary decomposition.
- Whether introducing a third primitive ($A$) within the SCX framework is an \textbf{axiom-level question}.
- The construction of the gradient field diagnostic $D_{\text{adv}}$ is novel but lacks clear definition in the SCX framework.

\textbf{Rigorously Derivable from Existing Axioms}:
- Reduction's Lipschitz sufficient condition: Derivable, but uses an inequality ($||\nabla S||/S \leq \eta$) not in SCX axioms.

\textbf{Requires New Assumptions}:
- Validity of the three-term decomposition $S=Q+\eta N+\gamma A$.
- $||\nabla S||/S \leq \eta$ (Cercis gradient-to-value ratio bounded).
- Incomplete correlation between adversarial exposure $A$ and novelty $N$ ($I(A;N|S)>0$).
- Curl operator on Situs manifold (shared with CD-Theorem).

\textbf{Relationship to Other Theorems}: Explicitly cites CD (adversarial signals confused with causal signals), HC (humans detecting adversarial examples), TS ($S_{\text{crit}}$ drift). But \textbf{mathematically the weakest foundation} among all 8 --- both theorems heavily depend on unverified assumptions, and the core conjecture is completely open.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.7 TS-Theorem --- Temporal SCX (Drift \& Self-Audit)}\label{ts-theorem-temporal-scx-drift-self-audit}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- TS-Theorem 1 (Drift Detectability): Complete, sample size formula.
- TS-Theorem 2 (Drift Attribution): Complete, split into anchor-free and anchored cases.
- TS-Theorem 3 (Sprint Self-Audit Limit): Complete, Lyapunov condition.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- \textbf{Drift Detectability proof is reasonable}, but has one implicit assumption: the Lipschitz relation between Situs distance and $\eta$-drift $|\eta_{t+1} - \eta_t| \leq \kappa\cdot d_{\text{Situs}} + |\xi_t|$. Deriving this relation from T7's Lipschitz property is reasonable.
- \textbf{The variance formula (Eq. 14)} contains $\sigma^2_{\text{est}}$ but does not define or derive it. Under $H_0$, the estimation variance of $\eta$ should be derivable from the distribution of the Cercis Score, but the paper directly asserts a form.
- \textbf{The attribution unidentifiability construction} (Config A vs Config B) is convincing, but only for two configurations, not proving that \textbf{all} possible decompositions are unidentifiable. A more general unidentifiability proof is needed.
- \textbf{The Lyapunov convergence theorem} is a standard Foster-Lyapunov argument, no problem. The divergence condition derivation is also natural.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- Explicitly temporalizes T7 cross-domain transfer theorem (Situs distance becomes Situs distance between time slices).
- The Sprint self-audit problem is a natural temporal extension of the SCX framework.
- The anchor concept parallels HC-Theorem's anchor knowledge.

\textbf{Rigorously Derivable from Existing Axioms}:
- Drift Detectability: \checkmark\ Derived from T7 + standard hypothesis testing.
- Attribution unidentifiability: \checkmark\ Via counterexample construction.
- Anchored estimation: \checkmark\ Directly derived from anchor definition.

\textbf{Requires New Assumptions}:
- Lipschitz constant $\kappa$ from Situs distance to $\eta$.
- Existence of anchor set (practical but not theoretically required).
- Contraction rate $\rho$ in Lyapunov condition (needs derivation from Sprint update rule).

\textbf{Relationship to Other Theorems}: Shares anchor concept with T7, HC. The convergence/divergence analysis is parallel to but distinct from Spring SE-1's Robbins-Monro convergence --- Spring is about model evolution, TS is about distribution/audit parameter evolution.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.8 HC-Theorem --- Human-AI Collaborative Audit}\label{hc-theorem-human-ai-collaborative-audit}

\textbf{Theorem Statement Completeness}: $\bigstar\bigstar\bigstar\bigstar\bigstar$ (5/5)
- HC-Theorem 1 (Collaborative Gain): Complete.
- HC-Theorem 2 (Budget Decay): Complete, $\sqrt{B_H/n}$ scaling law.
- HC-Theorem 3 (Absolute Boundary): Complete.

\textbf{Proof Framework Logical Leaps}: $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ (4/5)
- \textbf{Collaborative Gain logic}: ``If human judgment carries super-SCX information, then collaborative audit strictly dominates pure AI.'' This is a conditional implication, logically correct, but \textbf{the condition itself ($I(J_H;\varepsilon|S)>0$) is an empirical hypothesis}. The paper honestly acknowledges this as an open problem.
- \textbf{Budget Decay derivation has a subtle issue}: Step 1 claims $\Omega_{\text{collab}} = \Omega_{\text{AI}} + (B_H/n)\cdot\Delta_I\cdot\eta_{\text{eff}}$, but Step 2 uses a Cram\'{e}r-Rao argument that $\eta_{\text{eff}} \leq c/\sqrt{B_H/n}\cdot 1/\Delta_I$, yielding the $\sqrt{B_H/n}$ scaling. This composite derivation must prove: the optimal selection based on Cercis Score indeed produces $\sqrt{B_H/n}$ efficiency loss. The paper's argument is too sketchy --- ``optimal selection introduces inter-sample dependence, reducing effective sample size from $B_H$ to $\sqrt{B_H\cdot n}$'' needs more rigorous treatment.
- \textbf{Absolute Boundary theorem}: This is a direct logical extension of Theorem 3, with solid argumentation. The concept of ``perfect noise'' is clear: $I(\varepsilon;\text{all observables}|X)=0 \Rightarrow$ absolutely indistinguishable.

\textbf{Consistency with SCX Axiom System}: $\bigstar\bigstar\bigstar\bigstar\bigstar$ (5/5)
- Directly extends Theorem 3 to arbitrary epistemic systems.
- The ``multi-lens'' perspective is a philosophical deepening of the SCX framework.
- Establishes a continuum from Theorem 3 (AI-lens) to arbitrary lenses.

\textbf{Rigorously Derivable from Existing Axioms}:
- Absolute Boundary (Theorem 3): \checkmark\ Direct corollary of Theorem 3.
- Budget Decay: Derivable from Theorem 3 + information theory, but the $\sqrt{B_H/n}$ scaling rigorousness needs more work.

\textbf{Requires New Assumptions}:
- Information asymmetry condition $I(J_H;\varepsilon|S)>0$ (empirical hypothesis).
- Calibration condition (human uncertainty monotonic in Cercis Score).
- The upper bound derivation of efficiency factor $\eta_{\text{eff}}$ requires stronger regularity conditions.

\textbf{Relationship to Other Theorems}: Connected to CD (intersection of human cognition and SCX signal), AR (human detection of adversarial examples), TS (anchors). \textbf{Philosophically the most profound and most thorough in understanding Theorem 3} among all 8 papers.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Cross-Paper Analysis}\label{cross-paper-analysis}

\subsubsection{2.1 Theorem Duplication or Contradiction}\label{theorem-duplication-or-contradiction}

\begin{itemize}
\tightlist
\item
  \textbf{Duplication}: No direct duplication. CD-Theorem's Strong Inseparability overlaps in spirit with Theorem 3 but addresses different problems (noise vs difficulty vs noise vs confounding).
\item
  \textbf{Potential Contradictions}:
  \begin{itemize}
  \tightlist
  \item
    AR-Theorem's Detectability claims to break Theorem 3 at the $\nabla S$ level (``Theorem 3's barrier applies at the $S$ level, not at the $\nabla S$ level''), while Theorem 3's original statement covers ``the Cercis Score and its derivatives.'' If Theorem 3's original proof indeed covers $\nabla S$, then AR's claim contradicts Theorem 3. \textbf{This requires checking whether Theorem 3's proof truly covers gradient information.} Theorem 3's text says ``observables accessible to an SCX auditor: the Cercis Score, its gradient, its Hessian,'' with the conclusion ``No algorithm operating on these observables can tell the worlds apart.'' AR-Theorem claims there is higher-order structure (curl) in $\nabla S$ that Theorem 3 has not exhausted. \textbf{This is a potential internal contradiction --- either Theorem 3's claim is weakened by AR, or AR's claim is invalid.}
  \item
    T5's $\eta/(1+\eta)$ factor is semantically not fully consistent with the original SCX $\eta$ parameter: in the original framework $\eta$ is a continuous exploration-exploitation tradeoff, while in T5 $\eta$ affects the sampling rate through an $\eta/(1+\eta)$ saturation factor. Should these two $\eta$'s be treated as the same parameter?
  \end{itemize}
\end{itemize}

\subsubsection{2.2 Weakest Paper: \textbf{AR-Theorem (Adversarial Robustness)}}\label{weakest-paper-ar-theorem-adversarial-robustness}

\begin{itemize}
\tightlist
\item
  \textbf{Reasons}:
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    One of the core results (Reduction) depends on the inequality $||\nabla S||/S \leq \eta$, which is not in the SCX axioms.
  \item
    Detectability's $\chi^2$ divergence calculation is not explicitly given.
  \item
    The core problem (Phase Transition Conjecture) is entirely an open conjecture.
  \item
    The three-term decomposition $S=Q+\eta N+\gamma A$ is a major modification of SCX axioms, but the paper does not justify the legitimacy of this modification.
  \item
    Potential contradiction with Theorem 3.
  \item
    Le Cam lower bound derivation is overly sketchy, may contain hidden assumptions.
  \end{enumerate}
\item
  \textbf{Honest critique}: This paper has a good idea (adversarial examples may constitute a third category), but the mathematical execution is insufficient. The two ``theorems'' are heavily conditional, and most core conclusions are ``if \ldots then \ldots'' conditional statements whose conditions are themselves unmet or unverified. It is essentially a \textbf{positioning paper + research program}, not a resolved theorem.
\end{itemize}

\subsubsection{2.3 Strongest Paper: \textbf{Theorem 7 (Cross-Domain Partition Preservation)}}\label{strongest-paper-theorem-7-cross-domain-partition-preservation}

\begin{itemize}
\tightlist
\item
  \textbf{Reasons}:
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Most complete theorem statements, every symbol has a clear definition.
  \item
    Five-step proof decomposition is clear, lemma support is thorough.
  \item
    Constructed asymptotic tightness, the highest standard of theoretical proof.
  \item
    Highest consistency with SCX framework --- directly departs from Situs conditional mutual information.
  \item
    Component decomposition (geometric $\times$ statistical $\times$ irreducible) provides actionable practical guidance.
  \item
    Extensions ($p$-Wasserstein, multi-target domain, adaptive refinement) demonstrate the theory's richness.
  \item
    \textbf{The only notable weakness} (entropy Lipschitz requiring Gaussian assumption) is discussed in the appendix.
  \end{enumerate}
\item
  \textbf{Second-strongest candidate}: HC-Theorem --- The philosophical deepening of Theorem 3 is exceptionally excellent; the three-layer structure (AI-only $\to$ human-augmented $\to$ absolute limit) is logically perfect. The Budget Decay $\sqrt{B_H/n}$ scaling is a beautiful information-theoretic result. The only shortcoming is that the core hypothesis depends on empirical verification.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Global Assessment Matrix}\label{global-assessment-matrix}

\begin{longtable}[]{@{}llllll@{}}
\toprule\noalign{}
Paper & Completeness & Logical Rigor & SCX Consistency & New Assumption Dependence & Overall Grade \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
T5-Active & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & High & \textbf{B+} \\
T6-Protocol & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & High & \textbf{B} \\
T7-CrossDomain & $\bigstar\bigstar\bigstar\bigstar\bigstar$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar\bigstar$ & Medium & \textbf{A} \\
CD-Causal & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & High & \textbf{B-} \\
FA-Federated & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & Very High & \textbf{B-} \\
AR-Adversarial & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar{\scriptstyle\star}{\scriptstyle\star}$ & Very High & \textbf{C+} \\
TS-Temporal & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & Medium & \textbf{B+} \\
HC-Human & $\bigstar\bigstar\bigstar\bigstar\bigstar$ & $\bigstar\bigstar\bigstar\bigstar{\scriptstyle\star}$ & $\bigstar\bigstar\bigstar\bigstar\bigstar$ & Medium & \textbf{A-} \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Summary: Key Issues Across 8 Papers}\label{summary-key-issues-across-8-papers}

\subsubsection{4.1 Inter-Paper Contradictions}\label{inter-paper-contradictions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item \textbf{AR vs Theorem 3: Gradient-level attack claims}. AR claims to break Theorem 3's indistinguishability at the $\nabla S$ level, while Theorem 3 explicitly covers derivatives. Requires resolution.
\item \textbf{T5 vs Core: $\eta$ semantics}. T5's $\eta/(1+\eta)$ factor is not in the original Cercis definition.
\end{enumerate}

\subsubsection{4.2 Shared Open Problems}\label{shared-open-problems}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item \textbf{Situs homomorphic composition} ($\oplus$): Needed by FA, CD, AR. Core open problem.
\item \textbf{Curl on Situs manifold}: Needed by CD, AR. Requires connection structure.
\item \textbf{Entropy Lipschitz}: Needed by T7. Requires Gaussian-like conditions.
\item \textbf{Three-term Cercis decomposition}: Needed by AR. Axiom-level question.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Recommendations}\label{recommendations}

\subsubsection{For Immediate Submission}: T7, HC --- mathematically most self-consistent, clearly connected to SCX axioms.

\subsubsection{For Revision}: T5, TS --- solid core but need assumption tightening.

\subsubsection{For Major Rework}: CD, FA, T6 --- significant gaps or dependence on unsolved math.

\subsubsection{For Complete Restructuring}: AR --- good idea, insufficient execution. Consider reframing as a research program rather than a theorem paper.

\end{document}
