pdfoutput=1
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\pdfoutput=1
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX/EGP MoE
}\label{scxegp-ux7adeux4e89ux5206ux6790ux62a5ux544amoe-ux67b6ux6784ux84b8ux998fux4e0eux4e13ux5bb6ux6cbbux7406}

\begin{quote}
2026-06-25

2023-2026  Mixture-of-Experts (MoE)Knowledge
Distillation (KD)MLIP gauge Expert
Governance

 SCX/EGP  SCX/EGP

\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{}\label{ux76eeux5f55}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \hyperref[a-mixture-of-experts-moe-in-mlip]{A. Mixture of Experts
  (MoE) in MLIP}
\item
  \hyperref[b-knowledge-distillation-for-mlip]{B. Knowledge Distillation
  for MLIP}
\item
  \hyperref[c-expert-merging--gauge--energy-alignment]{C. Expert Merging
  / Gauge / Energy Alignment}
\item
  \hyperref[d-expert-compilation--governancescx-ux6838ux5fc3ux521bux65b0ux533a]{D.
  Expert Compilation / GovernanceSCX }
\item
  \hyperref[e-ux7adeux4e89ux683cux5c40ux603bux7ed3ux4e0eux6218ux7565ux5efaux8bae]{E.
  }
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{A. Mixture of Experts (MoE) in
MLIP}\label{a.-mixture-of-experts-moe-in-mlip}

\subsubsection{A1. DeepMD MoE/MoLE --- ``Scaling MLIP with Mixtures of
Experts''}\label{a1.-deepmd-moemole-scaling-mlip-with-mixtures-of-experts}

\textbf{} Liu, Zhang, Peng, E, Zhang, Wang. ``Scaling Machine
Learning Interatomic Potentials with Mixtures of Experts.''
arXiv:2603.07977 (2026).

\textbf{}  MoE  MLIP  MoE 
NLP  CV  MoE


\textbf{} -
\textbf{MoLEMixture-of-Linear-Experts} 
\textbf{MoE MLP } -
\textbf{}element-wise routing
vs.~configuration-level routing vs.~ MoE  -
\textbf{shared expert}  -  DeepMD-kit
 DP 

\textbf{} - \textbf{E(3) equivariant
routing} MoE  - 
MoE +  MoLE -
\textbf{}
-  OMol25OMat24OC20M  SOTA

\textbf{ SCX/EGP } - \textbf{} DeepMD MoE 
SCX/EGP  -  MoE
DeepMD MoE \textbf{} MoE
 SCX/EGP
\textbf{}------ -
DeepMD MoE emergent expertise SCX/EGP
pre-defined expertise

\textbf{SCX } - \textbf{} SCX/EGP

DeepMD MoE  - \textbf{}
EGP MACE + NEP + SevenNetDeepMD MoE
 DP  - \textbf{} EGP
DeepMD MoE 

\textbf{} - DeepMD MoE
\textbf{}------SCX/EGP
'''' - DeepMD MoE
------SCX/EGP  MoE ''MoE  +
''

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{A2. Allegro/NequIP MoE --- Spatial Partition +
Co-training
Agreement}\label{a2.-allegronequip-moe-spatial-partition-co-training-agreement}

\textbf{} Nascimento et al.~``Mixture of Experts Framework in
Machine Learning Interatomic Potentials for Atomistic Simulations.''
arXiv:2604.26143 (2026).

\textbf{} \textbf{ MoE}
artificial stress
fields

\textbf{} -  E(3)-equivariant \textbf{Allegro}
 MoE  - \textbf{Spatial Partitioning}

- \textbf{Co-training}
\textbf{agreement
constraints}------
- Pt+CO 

\textbf{} - exact energy conservation maintained -
EOS -  -
\textbf{ 2 }

\textbf{ SCX/EGP } - \textbf{}
\textbf{}SCX/EGP
\textbf{} -
\textbf{}
- '''' SCX/EGP ''Gauge
''------

\textbf{SCX } - SCX/EGP
\textbf{}
MoE  - SCX/EGP


\textbf{} -
''''SCX/EGP
\textbf{} -
------

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{A3.  MoE --- Shazeer et al.~2017, Switch
Transformer}\label{a3.-ux901aux7528-moe-shazeer-et-al.-2017-switch-transformer}

\textbf{} - Shazeer et al.~``Outrageously Large Neural Networks:
The Sparsely-Gated Mixture-of-Experts Layer.'' ICLR 2017. - Fedus, Zoph
\& Shazeer. ``Switch Transformers: Scaling to Trillion Parameter Models
with Simple and Efficient Sparsity.'' JMLR 2022.

\textbf{} Shazeer 2017 \textbf{ MoE
} top-k
''''Switch
Transformer  top-1 

\textbf{} - \textbf{Noisy Top-K Gating}  logits
 top-k - \textbf{Switch
Transformer}
''''
- \textbf{Capacity Factor}  token
 - \textbf{} router 
float32 dropout

\textbf{}  Mixtral 8x7BDeepSeek-V2/V3Llama 4 
MoE LLM

\textbf{ SCX/EGP } - \textbf{} 
MoE SCX/EGP
'''' -
NLP MoE  token - SCX/EGP
 - NLP MoE  FFN
SCX/EGP 

\textbf{SCX } - SCX/EGP
------\textbf{``''}------
MoE  1. \textbf{}  MoE
collapse 2.
\textbf{}  - SCX/EGP


\textbf{} -  MoE
 SCX/EGP
 -  NLP MoE ''''shared expert
SCX/EGP 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{B. Knowledge Distillation for
MLIP}\label{b.-knowledge-distillation-for-mlip}

\subsubsection{B1. MACE → NEP }\label{b1.-mace-nep-ux84b8ux998f}

\textbf{} ``Constructing machine learning interatomic potentials
with minimum amount of ab initio data.'' npj Computational Materials,
s41524-026-02023-y (2026).

\textbf{}  DFT \textasciitilde200
 MLIP DFT


\textbf{} - \textbf{-} - Stage
1 \textbf{MACE-MP-0} 
\textasciitilde200  DFT  - Stage 2
MACE  \textbf{NEP}  -
LGPSLATPLYC

\textbf{} single-shotNEP
 MACE MD 

\textbf{ SCX/EGP } - \textbf{} ''
→ ''SCX/EGP '' →
'' SCX/EGP  1. \textbf{Gauge
} SCX/EGP  Gauge
 2.
\textbf{} SCX/EGP
→→

\textbf{SCX } - SCX/EGP
'''' - SCX/EGP
MACENequIPSevenNet


\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{B2. Ensemble Knowledge Distillation (EKD) for
MLIP}\label{b2.-ensemble-knowledge-distillation-ekd-for-mlip}

\textbf{} Matin et al.~``Ensemble Knowledge Distillation for
Machine Learning Interatomic Potentials.'' arXiv:2503.14293 (2025).

\textbf{}  QC


\textbf{} 1.  QC  \textbf{
MLIP}/ 2.
\textbf{}ensemble-averaged forces
3.  \textbf{QC  + } 4.  ANI-1ccx
 COMP6  SOTA

\textbf{ SCX/EGP } - \textbf{}
''''SCX/EGP
''''EKD


\textbf{SCX } - EKD SCX/EGP
 - EKD
SCX/EGP
''''

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{B3. Foundation Model
Distillation}\label{b3.-foundation-model-distillation}

\textbf{} Gardner et al.~``Distillation of atomistic foundation
models across architectures and chemical domains.'' arXiv:2506.10956
(2025).

\textbf{} distillation via synthetic
data

\textbf{} - large universal MLIP -
 -  \textgreater10xGNN→GNN
\textgreater100xGNN→ACE -


\textbf{ SCX/EGP } - \textbf{}
''→''SCX/EGP
''→''------
SCX/EGP 

\textbf{} - '''' SCX/EGP
 -
''''------ SCX/EGP


\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{B4. Teacher-Student Training for MLIPs (RSC
2025)}\label{b4.-teacher-student-training-for-mlips-rsc-2025}

\textbf{} Matin et al.~``Teacher-student training improves the
accuracy and efficiency of machine learning interatomic potentials.''
RSC Digital Discovery, 2025.

\textbf{}
-\textbf{}

\textbf{} - \textbf{HIPNN}Hierarchically Interacting
Particle Neural Network -
 -
 - 
\textbf{\textgreater2 }\textbf{\textless50\%
}\textbf{} - \textbf{}Pareto
dominance

\textbf{ SCX/EGP } - \textbf{}
\textbf{}SCX/EGP


\textbf{} - SCX/EGP
''''------
\textbf{EGP synergy}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{B5.
}\label{b5.-ux5176ux4ed6ux91cdux8981ux84b8ux998fux5de5ux4f5c}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1765}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1765}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1765}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4706}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 SCX/EGP 
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{SevenNet-Nano} (arXiv:2604.10887) & 2026 &  SevenNet-Omni
 Nano\textgreater10  &
 \\
\textbf{PFD } (Phys. Rev.~Materials, 2025) & 2025 &
Pre-train→Fine-tune→Distill \textasciitilde100 DFT
\textasciitilde100x & PFD  \\
\textbf{ARK Distillation} (npj Comput. Mater., 2026) & 2026 &
 + 11.9x  &
 \\
\end{longtable}

\textbf{ARK Distillation } - Lim et al.~``Angular
relational knowledge distillation of machine learning interatomic
potentials for scalable catalyst exploration.'' npj Computational
Materials, 2026. - \textbf{angular relational
vectors} \textbf{} PES  - 58
 11.6 GPU-hours59.4 CPU-years - SCX/EGP
''''
PES 

\textbf{PFD } - Wang, Gao et al.~``Pre-training,
fine-tuning, and distillation (PFD): Automatically generating machine
learning force fields from universal models.'' Phys. Rev.~Materials,
2025. - DPA-2 →  → DeePMD  \textasciitilde100  DFT
 - SCX/EGP  PFD PFD ''''SCX/EGP
''''

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{C. Expert Merging / Gauge / Energy
Alignment}\label{c.-expert-merging-gauge-energy-alignment}

\subsubsection{C1. Model Merging (Model Soup, Git Re-Basin,
)}\label{c1.-model-merging-model-soup-git-re-basin-ux7ebfux6027ux6a21ux5f0fux8fdeux63a5}

\textbf{}


\textbf{}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Model Soups} (Wortsman et al., ICML 2022) & 2022 &
/ \\
\textbf{Git Re-Basin} (Ainsworth et al., ICLR 2023) & 2023 &
permutation alignment
basin \\
\textbf{ZipIt!} (ICLR 2024) & 2024 &
 \\
\textbf{AME} (Lee \& Ngo, NeurIPS 2025) & 2025 &  soup
''Amortized Model
Ensembling`` \\
\textbf{Soup-of-Experts} (ICML 2025) & 2025 &
 \\
\textbf{LLM Souping} (Meta, 2025) & 2025 & 
LLM SoCE \\
\end{longtable}

\textbf{ SCX/EGP } - \textbf{} 
SCX/EGP '''' -
\textbf{}weight-space
averaging - SCX/EGP
\textbf{}prediction-space ensembling
- \textbf{}SCX/EGP
\textbf{}

\textbf{SCX } - \textbf{}
SCX/EGP  MACE + NEP + SevenNet  -
\textbf{} ''''
MLIP SCX/EGP  - \textbf{}
SCX/EGP 

\textbf{} - \textbf{}
MLIP  - ''MLIP
''------

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{C2. MLIP Gauge Freedom  Species Energy
Shift}\label{c2.-mlip-gauge-freedom-ux4e0e-species-energy-shift}

\textbf{} ``Gauge dependence of total energies and
cross-functional transfer learning.'' arXiv:2504.05565 (2025).

\textbf{}  MLIP  \textbf{Gauge gauge
freedom} DFT
''Gauge''
MLIP 

\textbf{} - ''
AtomRefspecies energy shift`` - \textbf{}
\textbf{ refit AtomRef  GNN}
 AtomRef  Pearson  0.09  0.93
- Method 4refit AtomRef + train GNN MAE = 17
meV/atom GNN  26 meV/atom

\textbf{ SCX/EGP } - \textbf{} Gauge freedom 
SCX/EGP  \textbf{Gauge } MLIP
OUTCAR ''energy
without entropy'' MACE  eV - SCX/EGP  Gauge
GAU

\textbf{SCX } -  AtomRef
SCX/EGP  Gauge
\textbf{}------
- SCX/EGP  Gauge
\textbf{} DOF 

\textbf{} - ''formation energy''
Gauge-invariant SCX/EGP  formation energy
 total energy  -
\textbf{ Gauge }------SCX/EGP
 - GGA/GGA+U →
r²SCANSCX/EGP  Gauge SPME → MACE →
NEP

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{C3. Species Energy Shift
}\label{c3.-species-energy-shift-ux6807ux51c6ux505aux6cd5}

 MLIP 

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
 &  &  \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
CHGNet & AtomRef + GNN &  +  \\
M3GNet & AtomRef + GNN &  \\
NequIP & AtomRef (subtract per-atom energy) &  \\
CACE & AtomRef + GNN & Composible Atomic Cluster Expansion \\
MACE-MP-0 &  &  MP  \\
SevenNet &  offset &  \\
\end{longtable}

\textbf{}------SCX/EGP
 Gauge \textbf{}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{D. Expert Compilation / GovernanceSCX
}\label{d.-expert-compilation-governancescx-ux6838ux5fc3ux521bux65b0ux533a}

 SCX/EGP
\textbf{}

\subsubsection{D1. ``''Expert Normalization Before
Distillation}\label{d1.-ux84b8ux998fux524dux7684ux4e13ux5bb6ux89c4ux8303ux5316expert-normalization-before-distillation}

\textbf{}

\begin{itemize}
\tightlist
\item
  MACE→NEP, SevenNet-Nano, PFD, ARK, EKD
  
\item
  
  EKD
\item
  \textbf{''''}
\item
  SCX/EGP  Gauge → Route → Compile
  ''''''''
\end{itemize}

\textbf{SCX } -  SCX/EGP
''''SCX


\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{D2. Expert Domain
Certification}\label{d2.-expert-domain-certification}

\textbf{ MLIP }

\begin{itemize}
\tightlist
\item
   AI ''AI ''AIGP
  / MLIP 
\item
  ``What Makes An Expert?'' (arXiv:2411.00179, 2024) 
  ML ''''
\item
  \textbf{''MLIP ''}
\item
  SCX/EGP  ECR
  
  SCX 
\end{itemize}

\textbf{SCX } - ''MLIP Expert
Domain Certification'' -  conformal prediction


\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{D3. Multi-Expert Conflict
Arbitration}\label{d3.-multi-expert-conflict-arbitrationux591aux4e13ux5bb6ux51b2ux7a81ux4ef2ux88c1}

\textbf{ ML  MLIP }

 ML 

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{King Solomon Protocol} (2026) &  &
 \\
\textbf{HCDF} (2026) &  &
→→ \\
\textbf{Cognitive Hives} (2026) &  &
// \\
\textbf{LLMediator / Mediator} (2025) & LLM  &
 +  MoE \\
\textbf{AEGIS-Nexus} (2026) &  &
 \\
\textbf{Multi-Expert Fusion Networks} survey &  &
top-k  \\
\end{longtable}

\textbf{MLIP }

\begin{itemize}
\tightlist
\item
  '''' ML-MIX, 
  MoE''''
\item
  SCX/EGP  Route
  RTE------
\end{itemize}

\textbf{SCX } -  ML  HCDF
LLMediator  MLIP  -
 EGP 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{D4.
}\label{d4.-ux591aux4e13ux5bb6ux6027ux80fdux7684ux8f93ux5165ux6761ux4ef6ux4f9dux8d56}

\textbf{}

\begin{itemize}
\tightlist
\item
  MoE
  \textbf{}''''
\item
  Active learning  ALIP
  ''''
\item
  DeepMD MoE
  ''''------
\item
  \textbf{''''expert performance
  profile over input conditions}
\end{itemize}

\textbf{SCX } - ECR
 -


\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{E.
}\label{e.-ux7adeux4e89ux683cux5c40ux603bux7ed3ux4e0eux6218ux7565ux5efaux8bae}

\subsubsection{E1.
}\label{e1.-ux521bux65b0ux7ef4ux5ea6ux5bf9ux6bd4ux77e9ux9635}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.2857}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1429}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1429}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1429}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1429}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1429}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
DeepMD MoE (2603)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Allegro MoE (2604)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering

\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MLIP 
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
SCX/EGP
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
 MoE & Y & Y & N & N & N \\
/ & N & N & Y & Y & \textbf{Y} \\
 & N & N & N & N & \textbf{Y} \\
Gauge  & N & Y() & N & N & \textbf{Y} \\
 & Y() & Y() & N & N & \textbf{Y} \\
 & N & N & N & N & \textbf{Y} \\
 & N & N & N & N & \textbf{Y} \\
 & N & N & N & N & \textbf{Y} \\
 & N & N & Y & N & \textbf{Y} \\
\end{longtable}

\textbf{SCX/EGP ''''}
 SCX/EGP 
SCX/EGP 

\subsubsection{E2.
}\label{e2.-ux7adeux4e89ux98ceux9669ux7b49ux7ea7ux8bc4ux4f30}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2381}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4762}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2857}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{} & DeepMD MoE (2603.07977) & DeepMD
 \\
\textbf{} & Allegro MoE (2604.26143) &
'''''''' EGP  \\
\textbf{} & ARK  (npj 2026) &  ARK
'''' \\
\textbf{} & PFD  (PRM 2025) &  PFD
'''' EGP  Compile  \\
\textbf{} &  ( ML) &
 \\
\textbf{} & Foundation  (2506.10956) &
→ \\
\end{longtable}

\subsubsection{E3. SCX/EGP
}\label{e3.-scxegp-ux5fc5ux987bux4e3bux52a8ux5360ux9886ux7684ux7a7aux767d}



\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Prerequisite Normalization}

  \begin{itemize}
  \tightlist
  \item
     SCX/EGP 
  \item
    \textbf{}
  \end{itemize}
\item
  \textbf{Expert Domain Certification}

  \begin{itemize}
  \tightlist
  \item
    ''''
  \item
     conformal prediction 
  \item
     VASP
    
  \end{itemize}
\item
  \textbf{Cross-Architecture Compilation}

  \begin{itemize}
  \tightlist
  \item
     SCX/EGP  MACE + NEP + SevenNet + NequIP 
  \item
    ------DeepMD
    MoE 
  \end{itemize}
\item
  \textbf{Synergistic Distillation}

  \begin{itemize}
  \tightlist
  \item
     EGP 
  \item
    ''EGP  \textgreater{}  \textgreater{}
    ''
  \end{itemize}
\item
  \textbf{Input-Conditioned Benchmark}

  \begin{itemize}
  \tightlist
  \item
    ''''
  \item
     NEB 
  \item
    ''''
  \end{itemize}
\end{enumerate}

\subsubsection{E4. }\label{e4.-ux5efaux8baeux884cux52a8ux9879}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3529}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3529}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
P0 & SCX  &  arXiv  SCX/EGP ''Expert
Compilation'' \\
P0 &  EGP  &  VASP + MACE + NEP 
Gauge→Route→Compile  \\
P1 & Gauge  &  MLIP Gauge freedom
 \\
P1 &  vs.~DeepMD MoE  &  OMat24 
DeepMD MoE  \\
P2 &  &  Expert Domain Certification  \\
P2 &  &  EGP  Route  \\
\end{longtable}

\subsubsection{E5.
/}\label{e5.-ux7adeux4e89ux65b9ux6cd5ux80ccux540eux7684ux56e2ux961fux7ec4ux7ec7}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2857}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4762}}
  >{\centering\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2381}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
 SCX 
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{DP Technology / AIS} (, ) & DeepMD MoE (2603.07977)
& ------ MLIP  \\
\textbf{Boris Kozinsky } (Harvard) & Allegro/NequIP, MoE (2604.26143)
& ------ \\
\textbf{InstaDeep} & mlip JAX  (2605.22698) &
------ \\
\textbf{LANL} (Sakib Matin ) & EKD, Teacher-student RSC &
------ \\
\textbf{Volker Deringer } (Oxford) & Foundation distillation
(2506.10956) & ------ \\
\textbf{Seoul National Univ.} (Jeong Woo Han ) & ARK Distillation &
------ \\
\textbf{USTC} () & PFD  (2502.20809) & ------PFD
'''' EGP  \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{}\label{ux53c2ux8003ux6587ux732e}

\subsubsection{A. MoE in MLIP}\label{a.-moe-in-mlip}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Liu, Y. et al.~``Scaling Machine Learning Interatomic Potentials with
  Mixtures of Experts.'' arXiv:2603.07977 (2026).
  https://arxiv.org/abs/2603.07977
\item
  Nascimento, G.M. et al.~``Mixture of Experts Framework in Machine
  Learning Interatomic Potentials for Atomistic Simulations.''
  arXiv:2604.26143 (2026). https://arxiv.org/abs/2604.26143
\item
  Shazeer, N. et al.~``Outrageously Large Neural Networks: The
  Sparsely-Gated Mixture-of-Experts Layer.'' ICLR 2017.
  https://arxiv.org/abs/1701.06538
\item
  Fedus, W., Zoph, B. \& Shazeer, N. ``Switch Transformers: Scaling to
  Trillion Parameter Models with Simple and Efficient Sparsity.'' JMLR
  2022. https://arxiv.org/abs/2101.03961
\item
  Birks, T., Swinburne, T. \& Kermode, J. ``Efficient and accurate
  spatial mixing of machine learned interatomic potentials.'' npj
  Computational Materials (2025).
\item
  Nascimento, G.M. \& Descoteaux, B. ``Mixture-of-Experts Framework in
  MLIP for Atomistic Simulations.'' (2026).
\end{enumerate}

\subsubsection{B. Knowledge Distillation for
MLIP}\label{b.-knowledge-distillation-for-mlip-1}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  ``Constructing machine learning interatomic potentials with minimum
  amount of ab initio data.'' npj Computational Materials,
  s41524-026-02023-y (2026).
  https://www.nature.com/articles/s41524-026-02023-y
\item
  Matin, S. et al.~``Ensemble Knowledge Distillation for Machine
  Learning Interatomic Potentials.'' arXiv:2503.14293 (2025).
  https://arxiv.org/abs/2503.14293
\item
  Gardner, J.L.A. et al.~``Distillation of atomistic foundation models
  across architectures and chemical domains.'' arXiv:2506.10956 (2025).
  https://arxiv.org/abs/2506.10956
\item
  Matin, S. et al.~``Teacher-student training improves the accuracy and
  efficiency of machine learning interatomic potentials.'' RSC Digital
  Discovery (2025).
  https://pubs.rsc.org/en/content/articlehtml/2025/dd/d5dd00085h
\item
  Kim, J. et al.~``A Lightweight Universal Machine-Learning Interatomic
  Potential via Knowledge Distillation for Scalable Atomistic
  Simulations (SevenNet-Nano).'' arXiv:2604.10887 (2026).
  https://arxiv.org/abs/2604.10887
\item
  Wang, R., Gao, Y. et al.~``Pre-training, fine-tuning, and distillation
  (PFD): Automatically generating machine learning force fields from
  universal models.'' Phys. Rev.~Materials 9, 113802 (2025).
  https://arxiv.org/abs/2502.20809
\item
  Lim, H. et al.~``Angular relational knowledge distillation of machine
  learning interatomic potentials for scalable catalyst exploration.''
  npj Computational Materials 12, 1 (2026).
  https://link.springer.com/article/10.1038/s41524-026-02062-5
\item
  ``Data-Efficient and Fast Machine Learning Molecular Dynamics through
  Integrated Active Learning and Knowledge Distillation.'' ChemRxiv
  (2026).
\end{enumerate}

\subsubsection{C. Expert Merging / Gauge
Alignment}\label{c.-expert-merging-gauge-alignment}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{14}
\tightlist
\item
  ``Gauge dependence of total energies and cross-functional transfer
  learning.'' arXiv:2504.05565 (2025). https://arxiv.org/abs/2504.05565
\item
  Wortsman, M. et al.~``Model soups: averaging weights of multiple
  fine-tuned models improves accuracy without increasing inference
  time.'' ICML 2022.
\item
  Ainsworth, S. et al.~``Git Re-Basin: Merging Models modulo Permutation
  Symmetries.'' ICLR 2023.
\item
  Lee, S.H. \& Ngo, R. ``On Defining Neural Averaging.'' NeurIPS 2025.
  https://arxiv.org/abs/2508.14832
\item
  Ablin, P. et al.~``Soup-of-Experts.'' ICML 2025.
\item
  Franke et al.~``Hierarchical Re-Basin.'' ESANN 2024.
\end{enumerate}

\subsubsection{D. Governance / Arbitration (General
ML)}\label{d.-governance-arbitration-general-ml}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{20}
\tightlist
\item
  Grassini Grimaldi, A. ``The King Solomon Protocol: A Framework for
  Human Supervised Arbitration in Multi-Model AI Systems.'' (2026).
\item
  Bie et al.~``Knowledge-guided policy arbitration: A hierarchical
  cognitive framework for safety-critical decision-making.'' Expert
  Systems with Applications (2026).
\item
  Hettiarachchi. ``From Monolithic Models to Cognitive Hives: A
  Framework for Multimodal Reasoning Systems.'' JCSTS (2026).
\item
  ``Beyond Monoliths: Expert Orchestration for More Capable, Democratic,
  and Safe Large Language Models.'' arXiv:2506.00051 (2025).
\item
  ``Toward Calibrated Mixture-of-Experts Under Distribution Shift.''
  arXiv:2606.20544 (2026).
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
 2026-06-25 arXivNature
PortfolioRSCAPSSpringer SCX/EGP
'''' DeepMD 
Kozinsky 
\end{quote}

\end{document}
