% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Distillation De-Virus: Mathematical Roots and Multi-Scenario Generalization}\label{distillation-de-virus-mathematical-roots-and-multi-scenario-generalization}

\begin{quote}
Date: 2026-06-26 \textbar{} SCX v4.0 \textbar{}
From ``Data Poisoning Defense'' to ``Distillation De-Virus'': the mathematical foundations
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Problem Definition: Why Is Distillation ``Poisoned''?}\label{problem-definition-why-is-distillation-poisoned}

\subsubsection{1.1 Standard Knowledge Distillation Review}\label{standard-knowledge-distillation-review}

The framework of Hinton et al.~(2015):

\textbf{Teacher} $f_T: X \to Y$, trained on training set $D_{\text{train}} = \{(x_i, y_i^*)\}$.

\textbf{Distillation}: Teacher generates soft or hard labels for a set of unlabeled data $\{x_i^{\text{distill}}\}$:

\begin{verbatim}
y_i^distill = f_T(x_i^distill)
\end{verbatim}

\textbf{Student} $f_S: X \to Y$, trained on $D_{\text{distill}} = \{(x_i^{\text{distill}}, y_i^{\text{distill}})\}$.

\subsubsection{1.2 The Mathematical Essence of the Problem}\label{the-mathematical-essence-of-the-problem}

The core problem is: \textbf{the Teacher is not perfect}. Define the Teacher's state-conditioned risk:

\begin{verbatim}
R_T(s) = E_{x~P(·|s)} [ ℓ(f_T(x), f*(x)) ]
\end{verbatim}

where $f^*(x)$ is the true label (unknown), and $s$ is the state.

\textbf{Core observation}: $R_T(s)$ varies greatly with $s$. In AlN v3 training:

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
State $s$ (configuration region) & $R_T(s)$ (force RMSE, eV/Å) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Phonon region & 0.008 \\
EOS near-equilibrium region & 0.015 \\
Elastic deformation region & 0.011 \\
\textbf{Thermal 1800K region} & \textbf{0.077} \\
\textbf{MLMD stress10 region} & \textbf{0.036} \\
\end{longtable}

The Teacher's error in the thermal region is \textbf{9.6$\times$} that of the phonon region.

\textbf{Distillation Poisoning Mechanism}:

\begin{verbatim}
Teacher generates distillation data in thermal 1800K region
    → Distillation labels y_distill = f_T(x) = f*(x) + δ_T(s)  [δ_T is teacher bias in state s]
    → Student fits: f_S(x) ≈ y_distill = f*(x) + δ_T(s)
    → Student error in thermal region ≥ δ_T(s)
    → Student "inherits" systematic bias from Teacher
\end{verbatim}

The difference from VASP data poisoning:

\begin{verbatim}
VASP Poisoning: f*(x) itself is wrong (DFT labels have problems)
Distillation Poisoning: f*(x) is correct (DFT labels are right), but Teacher did not correctly learn f*(x)
\end{verbatim}

\textbf{Commonality}: Both are state-conditioned---poisoning only occurs in specific configuration regions.

\subsubsection{1.3 Why Distillation Poisoning Is Harder to Diagnose Than VASP Poisoning}\label{why-distillation-poisoning-is-harder-to-diagnose}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
VASP Data Poisoning
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Distillation Data Poisoning
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Diagnosis signal} & $f_{\max} > 5$ eV/Å (directly observable) & No direct signal \\
\textbf{Label source} & DFT (known reliability) & Teacher (unknown reliability) \\
\textbf{Interpretability} & Yes---wrong configuration → wrong forces & \textbf{No}---distillation outputs look ``normal'' \\
\textbf{Prior knowledge needed} & Know that physically $f_{\max}$ should be $< 1$ eV/Å & Need to know how reliable Teacher is in each region \\
\textbf{SCX's role} & Automatically detect anomalous $f_{\max}$ & Use Teacher's residual map on training set to infer \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Mathematical Analysis of Existing Methods}\label{mathematical-analysis-of-existing-methods}

\subsubsection{2.1 Hinton KD (2015)---No Purification, Full Distillation}\label{hinton-kd-2015-no-purification}

\textbf{Approach}: $T \to S$, all distillation samples treated equally.

\textbf{Implicit assumption}: Teacher is equally reliable across all input regions.

\textbf{Mathematically}: This is equivalent to assuming $\text{Var}[\ell(f_T(x), f^*(x))] \approx 0$, i.e., the variance of teacher error is negligible. But in MLIP scenarios, Teacher error across different configuration regions can differ by an order of magnitude---this assumption clearly does not hold.

\subsubsection{2.2 Confidence-Based Filtering}\label{confidence-based-filtering}

\textbf{Approach}: Use Teacher's output entropy $H(f_T(x))$ or maximum softmax value as ``confidence,'' filtering low-confidence distillation samples.

\textbf{Mathematical form}:

\begin{verbatim}
D_clean = {x_i | max_j f_T^{(j)}(x_i) > τ_confidence}
\end{verbatim}

\textbf{Why it fails in MLIP scenarios}:

Teacher (ACE linear model) outputs energy $E$ and forces $F$, not a probability distribution. There is no ``confidence'' concept---$E$ and $F$ are deterministic outputs. \textbf{Confidence filtering is directly inapplicable.}

Even in classification scenarios: high confidence $\neq$ low error. A model can be highly confident about a wrong prediction (overconfidence). Confidence does not reflect state-conditioned reliability structure.

\subsubsection{2.3 Ensemble Disagreement Filtering}\label{ensemble-disagreement-filtering}

\textbf{Approach}: Train $M$ Teachers (different initializations/data subsets), use output variance as uncertainty:

\begin{verbatim}
σ²_ens(x) = Var[f_T1(x), f_T2(x), ..., f_TM(x)]
D_clean = {x_i | σ²_ens(x_i) < τ_ens}
\end{verbatim}

\textbf{Advantages}: More reliable than confidence; captures model uncertainty across regions.

\textbf{Limitations}: Requires training $M$ Teachers → high cost ($M$ ACE trainings = $M \times 40$ epochs in MLIP). Ensemble variance reflects ``consistency among multiple Teachers,'' not ``error between Teacher and true labels.'' If all Teachers make the same systematic error in the same region (common---due to architectural limitations), ensemble variance will be low but actual error large.

\subsubsection{2.4 Data Shapley / LOO Values}\label{data-shapley-loo-values}

\textbf{Approach}: Compute Shapley values or LOO error for each training sample; retain high-value samples, discard low-value ones.

\textbf{Mathematically}: Shapley value for the $i$-th sample computes ``the marginal contribution of this sample to model performance.'' But this is a \textbf{global quantity}---it does not distinguish contributions across different states.

\textbf{Limitations}: Like confidence, Shapley values contain no state-conditioned information. A sample contributing to the phonon region and one contributing to the thermal region may have identical Shapley values, but Teacher reliability in the thermal region is completely different.

\subsubsection{2.5 Fundamental Deficiency of All Methods}\label{fundamental-deficiency}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1875}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.3438}}
  >{\centering\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1562}}
  >{\centering\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1562}}
  >{\centering\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1562}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Method
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Signal needed
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MLIP applicable?
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Distinguishes Teacher regional reliability?
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Distinguishes state conditions?
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Confidence filtering & softmax/logits & ❌ (ACE has no probabilistic output) & ❌ & ❌ \\
Ensemble disagreement & $M$-model variance & ⚠️ (feasible but costly) & ⚠️ Partial & ❌ \\
Data Shapley & All-subset retraining & ❌ (computationally infeasible) & ❌ & ❌ \\
\textbf{SCX De-Virus} & \textbf{Teacher's residual on $D_{\text{train}}$ + Encoder} & ✅ & ✅ & ✅ \\
\end{longtable}

\textbf{All existing methods fail to distinguish states}---they treat Teacher reliability as a homogeneous problem solvable at the sample level (rather than state level). SCX's core contribution is pointing out that this assumption is wrong.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. SCX Distillation De-Virus Mathematical Framework}\label{scx-distillation-de-virus-mathematical-framework}

\subsubsection{3.1 Core Idea}\label{core-idea}

\textbf{Not judging distillation data quality sample by sample, but judging Teacher reliability at the state level.}

Intuition: Sample-level judgment: Is this distillation frame's force label correct? → Unanswerable, because there is no ground truth. State-level judgment: What is the overall reliability of the Teacher in the ``thermal 1800K configuration region''? → Answerable, because Teacher already shows high error on thermal test frames from the original training set.

\subsubsection{3.2 Formal Definition}\label{formal-definition}

\textbf{Known} (from Teacher training): $D_{\text{train}} = \{(x_i, y_i^*)\}$ --- original training set with DFT labels. Encoder $\varphi: \mathcal{X} \to \mathbb{R}^d$ --- State Encoder. Teacher $f_T$ already trained.

\textbf{Computable} (no additional DFT needed): Teacher's residual on each sample in $D_{\text{train}}$: $r_i = \ell(f_T(x_i), y_i^*)$. State discovery: cluster with $\varphi$ → state partition $\mathcal{S} = \{s_1, \ldots, s_K\}$. Teacher reliability per state:

\begin{verbatim}
R_T(s_k) = (1/|s_k|) Σ_{x_i∈s_k} ℓ(f_T(x_i), y_i^*)
\end{verbatim}

This is the \textbf{Teacher Reliability Map}---essentially a map of ``which configuration regions the Teacher is unreliable in.''

\textbf{Distillation data assessment}: $D_{\text{distill}} = \{(x_j^{\text{distill}}, f_T(x_j^{\text{distill}}))\}$. For each $x_j^{\text{distill}}$, encode $\varphi(x_j^{\text{distill}})$, assign to nearest state $s_k$. If $R_T(s_k) > \tau$, mark $x_j^{\text{distill}}$ as ``virus frame.''

\textbf{Filtered distillation}:

\begin{verbatim}
D_clean = {x_j^distill | R_T(state(x_j^distill)) ≤ τ}
\end{verbatim}

Train Student $f_S$ on $D_{\text{clean}}$.

\subsubsection{3.3 Mathematical Guarantee (Heuristic)}\label{mathematical-guarantee}

\textbf{Claim}: Let Teacher bias in state $s$ be $\delta_T(s) = |\mathbb{E}[f_T(x) - f^*(x) \mid x \in s]|$. Student trained on filtered distillation set satisfies, for expected bias in state $s$:

\begin{verbatim}
E[|f_S(x) - f*(x)| | x∈s] ≤ δ_S_clean(s) + λ · max(0, δ_T(s) - τ)
\end{verbatim}

where $\delta_{S,\text{clean}}(s)$ is Student's intrinsic learning error on clean data, and the second term is the leakage of teacher bias.

\textbf{Interpretation}: By setting $\tau$ to filter high-bias states, we can control the propagation of Teacher's systematic error to the Student. Stricter filtering (smaller $\tau$) means less leakage but fewer distillation samples.

\textbf{Trade-off}:

\begin{verbatim}
V_s(τ) = |D_clean(τ)| / |D_distill|  — retention ratio
Q_s(τ) = 1 - (δ_T(s) - τ)_+ / δ_T(s)  — quality improvement ratio
\end{verbatim}

The optimal $\tau$ depends on: distribution of Teacher reliability across states, coverage requirements of distillation data, and Student sensitivity to data volume.

\subsubsection{3.4 Why Two-Layer Descriptors Are Especially Critical Here}\label{why-two-layer-descriptors-are-critical}

Single-layer Encoder (e.g., MLIPEncoder 12-dim): Coarse granularity, 50\% of frames squeezed into one state. Cannot distinguish ``EOS near-equilibrium frames'' from ``thermal 1800K frames'' (they may be grouped into the same coarse state). → One state mixes Teacher-reliable and Teacher-unreliable sub-regions. → Filtering either ``wrongly kills'' reliable frames or ``lets through'' unreliable frames.

Two-layer Encoder (Layer 1 + ErrorDrivenEncoder): Layer 1 coarse clustering: large-region partitioning. Layer 2 error-driven fine clustering: within each Layer 1 cluster, re-cluster using feature dimensions correlated with Teacher error. → Separates Teacher-unreliable sub-regions from reliable ones. → Precise filtering: only discard sub-states where Teacher is genuinely unreliable.

\textbf{This is the unique value of the two-layer method in distillation de-virus---it can separate ``Teacher's reliable regions'' and ``Teacher's unreliable regions'' within the same coarse region.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Distillation De-Virus Across Different Scenarios}\label{distillation-de-virus-scenarios}

\subsubsection{4.1 Scenario 1: MLIP Potential Distillation (ACE → ACE/NEP)}\label{scenario-1-mlip-potential-distillation}

\textbf{Distillation approach}: Teacher ACE potential computes single-point energy + forces on many configurations → generates (configuration, energy, force) data → trains Student NEP / lightweight ACE.

\textbf{Poisoning risk}: Teacher's high error in extreme regions (thermal 1800K / MLMD stress10) propagates to Student.

\textbf{SCX Solution}: 1. Teacher trained on $D_{\text{train}}$ (534 frames). 2. Compute Teacher per-state force RMSE from $D_{\text{train}}$ → Reliability Map. 3. Distillation generates 5000 frames (Teacher performs single-point calculations on ASE-rattled configurations). 4. SCX Encoder encodes distillation frames → maps to states. 5. Filter distillation frames from thermal 1800K and MLMD stress10 regions. 6. Train Student on clean distillation set.

\textbf{Encoder choice}: \texttt{MLIPEncoder} (12-dim handcrafted descriptors), refined with two-layer method.

\subsubsection{4.2 Scenario 2: Image Model Distillation (ResNet-50 → MobileNet)}\label{scenario-2-image-model-distillation}

\textbf{Distillation approach}: Teacher ResNet-50 infers on unlabeled images → (image, soft label) → trains MobileNet.

\textbf{Poisoning risk}: Teacher recognition errors in specific image subspaces (e.g., low light, occlusion, blur) propagate to Student.

\textbf{SCX Solution}: 1. Teacher trained on CIFAR/ImageNet training set. 2. Compute Teacher per-state classification error from validation set → Reliability Map. 3. Distill many unlabeled images. 4. \texttt{VisionEncoder} (CNN backbone) encodes → maps to states. 5. Filter distillation frames from Teacher high-error states (e.g., blurry image regions). 6. Train Student on clean distillation set.

\textbf{Encoder choice}: \texttt{VisionEncoder} (simple\_cnn/resnet backbone).

\textbf{Key difference from confidence filtering}: Confidence filtering only looks at Teacher's softmax output (``how confident is Teacher itself?''). SCX looks at Teacher's actual prediction error on similar images (``does Teacher really err a lot in this region?'').

\subsubsection{4.3 Scenario 3: LLM Distillation (GPT → Small Model)}\label{scenario-3-llm-distillation}

\textbf{Distillation approach}: Teacher LLM generates many QA pairs → (prompt, response) → trains Student small model.

\textbf{Poisoning risk}: Teacher hallucinations/errors in specific semantic subspaces (e.g., reasoning problems, code generation) are amplified and propagated through distillation data.

\textbf{SCX Solution}: 1. Teacher evaluated on benchmark datasets. 2. Use evaluation results to build Teacher per-topic reliability → Reliability Map. 3. Distillation generates many QA pairs. 4. \texttt{LLMEncoder} (BERT/GPT embedding) encodes prompts → maps to semantic states. 5. Filter distillation QA pairs from Teacher high-error semantic states (e.g., multi-step reasoning). 6. Train Student on clean distillation set.

\textbf{Encoder choice}: Needs implementation of new \texttt{LLMEncoder}---use sentence embeddings (e.g., BERT) to encode prompts as vectors, cluster in semantic space.

\textbf{Unique value}: This is a scenario where even confidence filtering struggles---LLMs are often highly confident about wrong answers (hallucination with high confidence). SCX does not rely on Teacher's self-assessment; it characterizes reliability using actual performance on external benchmarks.

\subsubsection{4.4 Scenario 4: Simulation Fidelity Distillation (Fine-mesh FEM → Coarse-mesh + Surrogate)}\label{scenario-4-simulation-fidelity-distillation}

\textbf{Distillation approach}: Fine-mesh FEM samples parameter space → (parameter, response) → trains reduced-order model / coarse-mesh correction term.

\textbf{Poisoning risk}: When fine-mesh FEM's numerical accuracy degrades in specific parameter regions (e.g., near resonance, near bifurcation), distillation data carries numerical error.

\textbf{SCX Solution}: 1. Fine FEM evaluated on verification points. 2. Use verification error to build FEM per-region reliability → Reliability Map. 3. FEM densely samples parameter space. 4. \texttt{TabularEncoder} (simulation parameter vector) encodes → maps to parameter states. 5. Filter simulation points from FEM high-error parameter regions (e.g., near-bifurcation regions). 6. Train Student reduced-order model on clean simulation set.

\textbf{Encoder choice}: \texttt{TabularEncoder} (simulation parameter vector, e.g., Re, Ma, load amplitude...).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Relationship to Existing SCX Work}\label{relationship-to-existing-scx-work}

\begin{verbatim}
Unified view of SCX's four capabilities:

┌─────────────────────────────────────────────────────────┐
│           SCX = State-Conditioned Data Value Judgment Framework        │
├─────────────────────────────────────────────────────────┤
│                                                          │
│  ① Data Poisoning Defense                               │
│     Judge: fmax (directly observable from VASP)          │
│     Output: Flag bad VASP frames                         │
│     AlN v3 Verification: F1=0.585, 100% hit rate         │
│                                                          │
│  ② Redundancy Compression                               │
│     Judge: Current model error in state s                │
│     Output: Low-error high-density regions compressible  │
│     AlN v3: phonon region 50-80% compressible            │
│                                                          │
│  ③ Distillation De-Virus ← New                          │
│     Judge: Teacher reliability in state s (from D_train) │
│     Output: Filter distillation data from Teacher-unreliable states │
│     Principle: Generalization of ①—from VASP bad frames  │
│               to Teacher bad predictions                │
│                                                          │
│  ④ Expert Routing                                       │
│     Judge: Reliability differences among multiple experts│
│            in state s                                    │
│     Output: Route each state to the most reliable expert │
│     Principle: Generalization of ③—not just filtering,   │
│               but also selecting the best               │
│                                                          │
└─────────────────────────────────────────────────────────┘
\end{verbatim}

\textbf{Unified mathematical core}: All four capabilities are built on estimating ``state-conditioned reliability.'' The only difference is the \textbf{source of the judgment signal}:

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Capability & Judgment Signal & Signal Source \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Data Poisoning Defense & $f_{\max}$ (VASP output) & DFT calculation \\
Redundancy Compression & Current model error & Training set \\
Distillation De-Virus & Teacher error on $D_{\text{train}}$ & Teacher training history \\
Expert Routing & Multi-expert reliability differences & Cross-expert comparison \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Literature Survey: Quality Control in Knowledge Distillation}\label{literature-survey}

\subsubsection{Existing Research Directions}\label{existing-research-directions}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3214}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4643}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Direction
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Representative Work
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Difference from SCX
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Confidence-based} & Balan et al.~(2022), ``Confidence-Aware KD'' & Confidence $\neq$ accuracy; inapplicable to deterministic models (e.g., ACE) \\
\textbf{Ensemble-based} & Malinin et al.~(2020), ``Ensemble Distillation'' & Requires $M$ Teachers; ensemble variance $\neq$ true error \\
\textbf{Data valuation} & Ghorbani \& Zou (2019), ``Data Shapley'' & Global quantity; does not distinguish state conditions \\
\textbf{Curriculum KD} & Jin et al.~(2019), ``Curriculum KD'' & Sorted by difficulty, not by Teacher reliability \\
\textbf{Contrastive KD} & Tian et al.~(2020), ``CRD'' & Contrastive learning improves representation, does not address local Teacher unreliability \\
\textbf{Online KD} & Zhang et al.~(2018), ``Deep Mutual Learning'' & Mutual learning, does not distinguish which region Teacher is strong in \\
\textbf{SCX De-Virus} & \textbf{This paper} & \textbf{State-conditioned Teacher reliability → filter unreliable regions → purify distillation data} \\
\end{longtable}

\subsubsection{Core Gap}\label{core-gap}

The common assumption of all existing KD quality control methods: \textbf{Teacher reliability can be judged at the sample level (per-sample)}. SCX's unique contribution is pointing out that this assumption is false---reliability is state-level (per-state), because Teacher error is systematically concentrated in specific configuration/feature regions.

\textbf{SCX is the first work to introduce the concept of ``state-conditioned expert reliability'' into knowledge distillation.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Implementation Roadmap}\label{implementation-roadmap}

\subsubsection{Phase 1: MLIP Verification (AlN v3, 2--3 days)}\label{phase-1-mlip-verification}

\begin{verbatim}
1. Train Teacher (Single ACE) on AlN v3 534 frames
2. Compute Teacher per-state force RMSE → Reliability Map
3. Teacher distills 5000 frames (single-point energy+forces on ASE rattled configurations)
4. SCX two-layer analysis of distillation data
5. Filter Teacher high-error states (thermal 1800K, MLMD stress10)
6. Train Student A (full distillation set) vs Student B (de-virus distillation set)
7. Compare force RMSE of both on thermal/MLMD test sets
\end{verbatim}

\textbf{Expectation}: Student B has significantly lower force RMSE in the thermal region than Student A.

\subsubsection{Phase 2: Image Verification (CIFAR-100, 1 week)}\label{phase-2-image-verification}

\begin{verbatim}
1. Teacher ResNet-50 → MobileNet-v2 distillation
2. SCX uses validation set to establish Teacher per-class/per-feature-region reliability
3. Filter distillation data from Teacher high-error regions
4. Compare: Full KD vs Confidence-filtered KD vs SCX-filtered KD
\end{verbatim}

\subsubsection{Phase 3: LLM Verification (Needs LLMEncoder Implementation)}\label{phase-3-llm-verification}

\subsubsection{Phase 4: Public Dataset Cleaning (Community Contribution)}\label{phase-4-public-dataset-cleaning}

\begin{verbatim}
MPTraj / OC20 / ColabFit → SCX two-layer analysis → SCX-cleaned → GitHub Release
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{8. Summary of Core Insights}\label{summary-of-core-insights}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Distillation poisoning is state-conditioned}---Teacher error is not uniformly distributed; it concentrates in specific configuration/feature regions. This is the fundamental reason SCX can solve the problem.
\item
  \textbf{All existing methods judge at the sample level}---confidence, ensemble variance, Shapley values are all per-sample. They do not leverage the information that ``Teacher is overall unreliable in state $s$.''
\item
  \textbf{SCX's unique value is elevating KD quality control from sample-level to state-level}---using Teacher's per-state reliability map from the training set instead of per-sample confidence for structured data filtering.
\item
  \textbf{Two-layer descriptors are the core of precise filtering}---single-layer coarse clustering mixes Teacher-reliable and unreliable sub-regions; two-layer refinement separates them in the error subspace.
\item
  \textbf{This is a paper-level contribution}---``State-Conditioned Knowledge Distillation'' can be an important subsection of SCX-Theory (Paper 2) or a standalone workshop paper.
\end{enumerate}

\end{document}
