% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Theorem 1: Multi-Expert Consistency Guarantees for Label Noise
Detection}\label{theorem-1-multi-expert-consistency-guarantees-for-label-noise-detection}

\begin{quote}
\textbf{Core Claim}: when \(M\)
 trained on disjointdata subsets independently trained experts on a sample'sconsistency score \(C(x)\)
exceedsthreshold \(\theta\) time, shouldsampleisLabelnoise confidencedegreewithexponentialrateconverges to 1. 

\textbf{Revision History}: 2026-06-27 --- Lemma 3 Proof Reconstruction (newincrease A6
balanceerrordistributionAssumptions); Chernoff appendix KL direction correction; mainTheorem Statementupdateis 
(A1)-(A6). two bugs fixed per Verification Report. 
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1 Theorem Statement}\label{ux5b9aux7406ux9648ux8ff0}

\subsubsection{1.1 Notation and Setup
(Setup)}\label{ux7b26ux53f7ux4e0eux8bbeux5b9a-setup}

Set up the following objects: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Symbol/Concept
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Meaning
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\mathcal{X}\) & Input Space, may measure \\
\(\mathcal{Y}\) & Label Space, \(|\mathcal{Y}| = K\) (separate class)or
\(\mathcal{Y} \subseteq \mathbb{R}^d\) (time return) \\
\(\mathcal{D}\) & Data Distribution, \((X, Y) \sim \mathcal{D}\) \\
\(f^*: \mathcal{X} \to \mathcal{Y}\) & truemarkNotefunction (not know Oracle) \\
\(\{f_m\}_{m=1}^M\) & \(M\)
 expertmodel, \(f_m: \mathcal{X} \to \mathcal{Y}\) \\
\(\Pi = \{s_1, \dots, s_K\}\) & Status plan separate, \(\mathcal{X}\) may measure separate cut \\
\(\ell: \mathcal{Y} \times \mathcal{Y} \to [0, B]\) & has world loss function \\
\(\tau > 0\) & expertincorrectthreshold \\
\(\theta \in (0, 1)\) & noise detectionthreshold \\
\end{longtable}

\textbf{Labelnoisemodel}: foreach type this \((x, y^*)\) (where \(y^* = f^*(x)\)
is trueLabel), observetoLabel \(y\) generatesuch aslower: 

\[y = \begin{cases}
y^* & \text{withprobability } 1 - \eta \\
\text{Uniform}(\mathcal{Y} \setminus \{y^*\}) & \text{withprobability } \eta
\end{cases}\]

where \(\eta \in (0, 1/2)\) is globalnoise rate, noise matter file and \(x\)
and has trainingdataindependent. 

\textbf{consistency score}: tosample
\((x, y)\), Definitionexpertincorrect indicate show variableandconsistency score: 

\[e_m(x, y) = \mathbf{1}\{\ell(f_m(x), y) > \tau\}, \quad
C(x) = \frac{1}{M} \sum_{m=1}^M e_m(x, y)\]

\textbf{detection rule then}: type this \((x, y)\) isMarked as noiseif and only if
\(C(x) > \theta\). 

\subsubsection{1.2 Assumptions (Assumptions)}\label{ux5047ux8bbe-assumptions}

\textbf{(A1) disjointtraining set }: \(M\) expertin \(M\)
 disjointindependentsamedistributionsubsetuppertraining: 

\[D_m \sim \mathcal{D}^{n_m}, \quad D_m \cap D_{m'} = \varnothing, \quad D_m \perp D_{m'} \text{ for } m \neq m'\]

\textbf{(A2) clear clean dataupperConditionsindependent}: to any agree clear clean sample
\((x, y)\) (\(y = y^*\)), todetermine \(x\) Conditionslower, incorrect indicate show variable
\(\{e_m(x, y)\}_{m=1}^M\) isConditionsindependent. 

\emph{reasonablenessDescriptionandlimitation}: A2
is a kind\textbf{structurenessAssumptions}, byexpertexperimentdesign (disjointtraining set 
A1、independentinitialization -ization) argumentation, but\textbf{A1 and not contain contain 
A2}. i.e. make training set complete fulldisjoint, atsimilarityData Distributionuppertrainingexpertmayatdistributionoutersampleupperproduces\textbf{mutualcloseincorrect} (originates fromcommon share return accept bias place, non-common share trainingdata). becausethis
A2 in SCX flow water line datamiddle\textbf{notmay via verify check verify }------tocontinuousInput Space
\(\mathcal{X}\), each \(x\) inhas finite data set middlemostmany out present a time, joint distribution
\(P(e_1, \dots, e_M \mid x)\) unobservable. \textgreater{} \textgreater{}
\textbf{when A2 is violate counter time} (actual actual middle often see): use 
\(M_{\text{eff}} = M/(1 + (M-1)\bar{\rho})\) replace has concentration inequality in the 
\(M\), where \(\bar{\rho}\)
is expertincorrect indicate markon average become tomutualcloseness. to classical type deepdegree set become, \(\bar{\rho} \approx 0.1\)--\(0.3\). Recommendationat retain place verify set upperestimate
\(\bar{\rho}\), and using \(M_{\text{eff}}\) obtain to preserve guard guarantee. detail fineAnalysis participate see
\texttt{01\_symbol\_system.md} §12.5. 

\textbf{(A3) has world loss}:
\(\ell(a, b) \in [0, B], \; \forall a, b \in \mathcal{Y}\), where
\(B < \infty\). 

\textbf{(A4) uniform uniform independentnoise}: Labelflipped matter file and \(x\) and has \(D_m\)
independent. noiseLabelin \(\mathcal{Y} \setminus \{y^*\}\) upperuniform uniform distribution. 

\textbf{(A5) Statussameness (Status plan separatesufficientness)}: Status plan separate \(\Pi\)
makesateach Status \(s\)
inner, expert clear clean dataincorrect rate approximateuniform uniform. form -ization ground, existenceatStatus level often number 
\(\{\mu_s\}_{s \in \mathcal{S}}\) makes: 

\[\sup_{x \in s} \, \mathbb{E}[C(X) \mid \text{clean}, X = x] \leq \mu_s, \quad \forall s \in \mathcal{S}\]

i.e.Status \(s\) inner clear clean sampleon averageexpertincorrect rate is \(\mu_s\) consistentupper boundcontrols. 

\textbf{(A6) balanceerrordistribution}:
expertincorrectprobabilityat has incorrectCategoryuppercannot over degree set middle. existenceat often number 
\(C_{\text{bal}} \geq 1\) makesto any agree Status \(s\) and \(x \in s\): 

\[\max_{c \neq y^*} \mu_c(x) \leq C_{\text{bal}} \cdot \frac{\mu_s}{K-1}, \quad \mu_c(x) = \frac{1}{M} \sum_{m=1}^M \mathbb{P}(f_m(x) = c \mid x)\]

\(C_{\text{bal}} = 1\)
toshould complete fulluniform uniform incorrectdistribution (each incorrectCategory etc. probability); \(C_{\text{bal}} = 2\)
is classical type preserve guard select. thisAssumptionsmay check verify andatpracticemiddle several -ish self- action satisfies------atdisjointdry clean datauppertrainingexpertnothas reason bywillincorrect set middleto special determineCategory. 

\subsubsection{1.3 Key Lemmas
(Mean Separation)}\label{ux5173ux952eux5f15ux7406-ux5747ux503cux5206ux79bb}

\textbf{Lemma 1 (Mean Separation).} inAssumptions (A1)-(A5) lower (A6
does not affectexpect), to any agree \(x \in s\): 

\textbf{(clear clean sample)}
\[\mathbb{E}[C(X) \mid \text{clean}, X = x] \leq \mu_s\]

\textbf{(noisesample)}
\[\mathbb{E}[C(X) \mid \text{noise}, X = x] = 1 - \frac{1}{K-1} \cdot \mathbb{E}[C(X) \mid \text{clean}, X = x] \geq 1 - \frac{\mu_s}{K-1}\]

becausethis, when \(\mu_s < \frac{K-1}{K}\) time, existencein \(\theta\) makes: 

\[\mathbb{E}[C \mid \text{clean}] < \theta < \mathbb{E}[C \mid \text{noise}]\]

andseparation gapis: 

\[\Delta_s(\theta) = \min\left(\theta - \mu_s, \; 1 - \frac{\mu_s}{K-1} - \theta\right)\]

whenselectOptimal threshold
\(\theta_s^* = \frac{1}{2}\left(1 + \mu_s \cdot \frac{K-2}{K-1}\right)\)
time, separation gapmaximum -ization (in \(C_{\text{bal}}=1\) reason think caselower): 

\[\Delta_s^* = \frac{1}{2}\left(1 - \mu_s \cdot \frac{K}{K-1}\right)\]

when \(C_{\text{bal}} > 1\) time, Optimal thresholdto \(\mu_s\)
directionmicro- adjust with supplement compensate non-uniform uniform errordistribution. \(\theta_s^*\) Exact table reach formula must seek solve 
\(\theta - \mu_s = 1 - C_{\text{bal}} \cdot \mu_s/(K-1) - \theta\). 

\textbf{Proof.} clear clean sample world by (A5) directto out. tonoisesample: 

\[\begin{aligned}
\mathbb{E}[C \mid \text{noise}, X = x] 
&= \frac{1}{M} \sum_{m=1}^M \mathbb{P}\bigl(\ell(f_m(x), y) > \tau \;\big|\; \text{noise}, x\bigr) \\
&= \frac{1}{M} \sum_{m=1}^M \mathbb{P}\bigl(f_m(x) \neq y \;\big|\; \text{noise}, x\bigr) \quad (\text{to } 0\text{-}1 \text{ loss}) \\
&= 1 - \frac{1}{M} \sum_{m=1}^M \mathbb{P}\bigl(f_m(x) = y \;\big|\; \text{noise}, x\bigr)
\end{aligned}\]

by (A4), noiseLabel \(y\) in \(\mathcal{Y} \setminus \{y^*\}\)
upperuniform uniform distribution, andis independent of has expert: 

\[\begin{aligned}
\mathbb{P}(f_m(x) = y \mid \text{noise}, x) 
&= \sum_{c \neq y^*} \mathbb{P}(y = c \mid \text{noise}) \cdot \mathbb{P}(f_m(x) = c \mid x) \\
&= \frac{1}{K-1} \sum_{c \neq y^*} \mathbb{P}(f_m(x) = c \mid x) \\
&= \frac{1}{K-1} \cdot \mathbb{P}(f_m(x) \neq y^* \mid x) \\
&= \frac{1}{K-1} \cdot \mathbb{E}[e_m \mid \text{clean}, x]
\end{aligned}\]

becausethis: 

\[\begin{aligned}
\mathbb{E}[C \mid \text{noise}, x] 
&= 1 - \frac{1}{K-1} \cdot \frac{1}{M} \sum_{m=1}^M \mathbb{E}[e_m \mid \text{clean}, x] \\
&= 1 - \frac{1}{K-1} \cdot \mathbb{E}[C \mid \text{clean}, x] \\
&\geq 1 - \frac{\mu_s}{K-1} \quad \text{by (A5)}
\end{aligned}\]

separation gapexistenceatnessby \(\mu_s < \frac{K-1}{K}\) guarantee, thistime
\(1 - \frac{\mu_s}{K-1} > \mu_s\). \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsubsection{1.4 Main Theorem (Main
Result)}\label{ux4e3bux5b9aux7406-main-result}

\textbf{Theorem 1 (SCX noise detection guarantee).} setupAssumptions (A1)-(A6) holds, and
\(M \geq 1\) (tofew a expert), \(K \geq 2\) (tofewtwo LabelCategory, withguarantee
\(K-1 > 0\)). set \(\rho_s = \mathbb{P}(X \in s)\)
is Statusprobability. to any agree threshold \(\theta\) satisfies
\(\mu_s < \theta < 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1}\) Status
\(s\), DefinitionStatus level separation gap: 

\[\Delta_s = \min\left(\theta - \mu_s,\; 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} - \theta\right) > 0\]

when \(C_{\text{bal}} = 1\) (most optimal situation shape)time, \(\Delta_s\) restore complex original begin Definition; when
\(C_{\text{bal}} > 1\)
time, noise side between gap collect narrow (becauseis incorrectmayis NOT complete fulluniform uniform distribution). 

then SCX noise detection device F1 obtain separatesatisfieswithlowerlower bound: 

\[\text{F1} \;\geq\; 1 - \sum_{s \in \mathcal{S}} \rho_s \cdot \Bigl[ \exp\!\bigl(-2M\Delta_s^2\bigr) \;+\; \frac{1-\eta}{\eta} \cdot \exp\!\bigl(-2M\Delta_s^2\bigr) \Bigr]\]

orequivalent ground: 

\[\text{F1} \;\geq\; 1 - \frac{1}{\eta} \sum_{s \in \mathcal{S}} \rho_s \cdot \exp\!\bigl(-2M\Delta_s^2\bigr)\]

where \(\eta\) is globalnoise rate. 

\textbf{more tight Chernoff form}: tomutualsameSetup, make using Chernoff world generation replace 
Hoeffding can obtain more tight exponentialrate: 

\[\text{F1} \;\geq\; 1 - \frac{1}{\eta} \sum_{s \in \mathcal{S}} \rho_s \cdot \Bigl[ \exp\!\bigl(-M \cdot \text{KL}(\theta \,\|\, \mu_s)\bigr) \;+\; \frac{1-\eta}{\eta} \cdot \exp\!\bigl(-M \cdot \text{KL}\!\left(\theta \;\Big\|\; 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1}\right)\bigr) \Bigr]\]

where
\(\text{KL}(p \,\|\, q) = p \log\frac{p}{q} + (1-p)\log\frac{1-p}{1-q}\). Chernoff
 world in \(\mu_s\) approaches \(\theta\) or
\(1 - C_{\text{bal}} \cdot \mu_s/(K-1)\) approaches \(\theta\) timeexplicit notable is superior to
Hoeffding (may tight 2-5 times). 

\begin{quote}
\textbf{2026-06-27 correction}: original version Chernoff use
\(\text{KL}(1-\theta \| 1 - \mu_s/(K-1))\), KL directionhas error, andnot accept enter 
\(C_{\text{bal}}\). present alreadyto 
\(\text{KL}(\theta \| 1 - C_{\text{bal}} \cdot \mu_s/(K-1))\). 
\end{quote}

\textbf{keyCorollary}: when \(M \to \infty\) time, to has satisfies
\(\mu_s < \frac{K-1}{K}\) Status, F1 withexponentialrateconverges to 1: 

\[\text{F1} \;=\; 1 - \mathcal{O}_P\!\left(\frac{1}{\eta} \cdot e^{-2M\Delta_{\min}^2}\right)\]

where \(\Delta_{\min} = \min_{s \in \mathcal{S}} \Delta_s\)
is most difference Statusseparation gap. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2 Proof}\label{ux8bc1ux660e}

\subsubsection{2.1 clear clean datafalse positive rate (False Positive
Rate)}\label{ux6e05ux6d01ux6570ux636eux7684ux5047ux9633ux6027ux7387-false-positive-rate}

\textbf{Lemma 2 (FPR upper bound).} to any agree Status \(s \in \mathcal{S}\) satisfies
\(\mu_s < \theta\), clear clean sampleby error markis noiseprobabilitysatisfies: 

\[\mathbb{P}(C > \theta \mid \text{clean}, X \in s) \leq \exp\!\bigl(-2M(\theta - \mu_s)^2\bigr)\]

\textbf{Proof.} fixed \(x \in s\), by (A5) know 
\(\mathbb{E}[C \mid \text{clean}, X = x] \leq \mu_s\). by
(A1)-(A2), todetermine \(x\) time \(\{e_m\}\) Conditionsindependent. by (A3) know 
\(e_m \in [0, 1]\). to \(C = \frac{1}{M}\sum_m e_m\) shouldusing Hoeffding
inequality (Pollard, 1990): 

\[\mathbb{P}(C - \mathbb{E}[C] > \theta - \mu_s \mid \text{clean}, x) \leq \exp\!\bigl(-2M(\theta - \mu_s)^2\bigr)\]

byat \(\theta > \mu_s\), \((\theta - \mu_s) > 0\). should world to any agree 
\(x \in s\) holds, becausethis: 

\[\begin{aligned}
\mathbb{P}(C > \theta \mid \text{clean}, X \in s) 
&= \mathbb{E}_{X \mid s}\bigl[ \mathbb{P}(C > \theta \mid \text{clean}, X) \bigr] \\
&\leq \sup_{x \in s} \mathbb{P}(C > \theta \mid \text{clean}, x) \\
&\leq \exp\!\bigl(-2M(\theta - \mu_s)^2\bigr)
\end{aligned}\]

thisi.e. must. \(\square\)

\subsubsection{2.2 noisedatatrue positive rate (True Positive
Rate)}\label{ux566aux58f0ux6570ux636eux7684ux771fux9633ux6027ux7387-true-positive-rate}

\textbf{newincreaseAssumptions (A6) --- balanceerrordistribution}: to any agree Status \(s\) andsample
\(x \in s\), expertincorrectprobabilityat has incorrectCategoryuppercannot over degree set middle. form -ization ground, existenceat often number 
\(C_{\text{bal}} \geq 1\) makes: 

\[\max_{c \neq y^*} \mu_c(x) \leq C_{\text{bal}} \cdot \frac{\mu_s}{K-1}, \quad \text{where } \mu_c(x) = \frac{1}{M} \sum_{m=1}^M \mathbb{P}(f_m(x) = c \mid x)\]

\emph{reasonableness}: expertatdisjointdry clean datasubsetuppertraining, nothas reason bywillincorrect set middleto special determineincorrectCategory. uniform uniform noisemodellower, each incorrectCategoryprobability sky however convention is 
\(\mu_s/(K-1)\). \(C_{\text{bal}} = 2\)
alreadyis preserve guard select. thisAssumptionsmay check verify: at actual actual dataupperestimate
\(\max_c \mu_c(x) / (\mu_s/(K-1))\). 

\textbf{Lemma 3 (TPR lower bound).} inAssumptions (A1)-(A6) lower, to any agree Status
\(s \in \mathcal{S}\) satisfies
\(\theta < 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1}\), noisesamplebycorrectdetectionprobabilitysatisfies: 

\[\mathbb{P}(C > \theta \mid \text{noise}, X \in s) \geq 1 - \exp\!\left(-2M\left(1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} - \theta\right)^2\right)\]

\textbf{Proof}: fixed \(x \in s\). todeterminenoiseLabel
\(y = c\) (\(c \neq y^*\)), for 0-1 loss, expert \(m\) incorrect indicate show is 
\(e_m = \mathbf{1}\{f_m(x) \neq c\}\). byat \(f_m\) only-dependence
\(D_m\) (A1), todetermine \(x\) time \(\{f_m(x)\}\) Conditionsindependent (A2), becausethis
\(\{e_m\}\) intodetermine \((x, c)\) lowerConditionsindependent. 

each \(e_m\) expectis 
\(\mathbb{E}[e_m \mid x, c] = 1 - \mathbb{P}(f_m(x) = c \mid x) = 1 - \mu_{c,m}(x)\). becausethis: 

\[\mathbb{E}[C \mid x, c] = 1 - \mu_c(x) \geq 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} \quad \text{ (by A6)}\]

byat \(e_m \in [0, 1]\) Conditionsindependent, shouldusing Hoeffding inequality: 

\[\begin{aligned}
\mathbb{P}(C \leq \theta \mid x, c)
&= \mathbb{P}\!\left(\frac{1}{M} \sum_{m=1}^M e_m \leq \theta \;\Big|\; x, c\right) \\
&\leq \exp\!\left(-2M\left(1 - \mu_c(x) - \theta\right)^2\right) \\
&\leq \exp\!\left(-2M\left(1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} - \theta\right)^2\right)
\end{aligned}\]

where Hoeffding gap Conditions \(1 - \mu_c(x) > \theta\) byTheoremAssumptions
\(\theta < 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1}\) and A6 guarantee: 
\(1 - \mu_c(x) \geq 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} > \theta\). 

tonoiseLabel \(c\) takeon average (uniform uniform distribution over \(K-1\) class): 

\[\begin{aligned}
\mathbb{P}(C \leq \theta \mid \text{noise}, x)
&= \frac{1}{K-1} \sum_{c \neq y^*} \mathbb{P}(C \leq \theta \mid x, c) \\
&\leq \exp\!\left(-2M\left(1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} - \theta\right)^2\right)
\end{aligned}\]

becausethis
\(\mathbb{P}(C > \theta \mid \text{noise}, x) \geq 1 - \exp(-2M(1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1} - \theta)^2)\). to
\(x \in s\) takeexpecti.e. obtain. \(\square\)

\textbf{Note}: when \(C_{\text{bal}} = 1\) (complete fulluniform uniform incorrectdistribution)time, Lemma
3 restore complex original begin Proposition in the Optimal bound. \(C_{\text{bal}}\)
mayat actual actual dataupperestimate, and used for calibrate alignment detectionthreshold \(\theta\). 

\subsubsection{2.3 F1 lower bound}\label{f1-ux4e0bux754c}

 present atcombine Lemma 2 and Lemma 3 Derivation F1 lower bound. 

setdetection rule thenis \(R(x) = \mathbf{1}\{C(x) > \theta\}\). Definition: 

\[\begin{aligned}
\text{TPR}_s &= \mathbb{P}(R = 1 \mid \text{noise}, X \in s) \geq 1 - \exp(-2M(1 - \mu_s/(K-1) - \theta)^2) \\
\text{FPR}_s &= \mathbb{P}(R = 1 \mid \text{clean}, X \in s) \leq \exp(-2M(\theta - \mu_s)^2)
\end{aligned}\]

by (A4), noise matter file and \(X\) independent, becausethis
\(\mathbb{P}(\text{noise} \mid X \in s) = \eta\) to has Statusholds. whole
TPR and FPR is: 

\[\begin{aligned}
\text{TPR} &= \sum_s \rho_s \cdot \text{TPR}_s, \quad
\text{FPR} = \sum_s \rho_s \cdot \text{FPR}_s
\end{aligned}\]

F1 obtain separateis: 

\[\begin{aligned}
\text{F1} &= \frac{2 \cdot \text{TP} }{2 \cdot \text{TP} + \text{FP} + \text{FN}} \\
&= \frac{2\eta \cdot \text{TPR}}{2\eta \cdot \text{TPR} + (1-\eta) \cdot \text{FPR} + \eta \cdot (1 - \text{TPR})} \\
&= \frac{2\eta \cdot \text{TPR}}{\eta(1 + \text{TPR}) + (1-\eta) \cdot \text{FPR}}
\end{aligned}\]

 generation enter TPR \(\geq 1 - \delta_1\) and FPR \(\leq \delta_2\), where: 

\[\begin{aligned}
\delta_1 &= \sum_s \rho_s \cdot \exp(-2M(1 - C_{\text{bal}} \cdot \mu_s/(K-1) - \theta)^2) \\
\delta_2 &= \sum_s \rho_s \cdot \exp(-2M(\theta - \mu_s)^2)
\end{aligned}\]

 obtain to: 

\[\begin{aligned}
\text{F1} &\geq \frac{2\eta(1 - \delta_1)}{\eta(2 - \delta_1) + (1-\eta)\delta_2} \\
&= 1 - \frac{\eta\delta_1 + (1-\eta)\delta_2}{\eta(2 - \delta_1) + (1-\eta)\delta_2} \\
&\geq 1 - \frac{\eta\delta_1 + (1-\eta)\delta_2}{\eta} \quad (\text{becauseseparate parent } \geq \eta) \\
&= 1 - \delta_1 - \frac{1-\eta}{\eta}\delta_2
\end{aligned}\]

 generation enter \(\delta_1, \delta_2\) table reach formula, and Note agree when
\(\Delta_s = \min(\theta - \mu_s, \; 1 - C_{\text{bal}} \cdot \mu_s/(K-1) - \theta)\)
timehas \(\exp(-2M(\theta - \mu_s)^2) \leq \exp(-2M\Delta_s^2)\) and
\(\exp(-2M(1 - C_{\text{bal}} \cdot \mu_s/(K-1) - \theta)^2) \leq \exp(-2M\Delta_s^2)\), becausethis: 

\[ \text{F1} \geq 1 - \sum_s \rho_s \left[\exp(-2M\Delta_s^2) + \frac{1-\eta}{\eta}\exp(-2M\Delta_s^2)\right] = 1 - \frac{1}{\eta}\sum_s \rho_s \exp(-2M\Delta_s^2) \]

QED. \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3 Corollary (Corollaries)}\label{ux63a8ux8bba-corollaries}

\subsubsection{3.1 Corollary 1: to called expert (Symmetric
Experts)}\label{ux63a8ux8bba-1ux5bf9ux79f0ux4e13ux5bb6-symmetric-experts}

when has expert has has mutualsame clear clean dataincorrect rate 
\(\varepsilon\) (called is ''to called expert'')time, \(\mu_s = \varepsilon\)
to has Statusconsistent. thistime: 

\[\Delta = \frac{1}{2}\left(1 - \varepsilon \cdot \frac{K}{K-1}\right)\]

and F1 lower boundsimplifiesis: 

\[\text{F1} \geq 1 - \frac{1}{\eta} \exp\!\left(-\frac{M}{2}\left(1 - \frac{\varepsilon K}{K-1}\right)^2\right)\]

when \(K = 2\) (two separate class)time, advance a step simplifiesis: 

\[\text{F1} \geq 1 - \frac{1}{\eta} \exp\!\left(-2M\left(\frac{1}{2} - \varepsilon\right)^2\right)\]

\textbf{meaning}: in two separate class to called expertscenariolower, detectionquality complete fullbyexpertincorrect rate 
\(\varepsilon\) and \(1/2\) difference distance verdict determine. \(\varepsilon\) eachfar from \(1/2\)
 a Standard difference (\(\approx 1/\sqrt{2M}\)), F1 lower bound raise liter a exponentialquantity level. 

\subsubsection{3.2 Corollary 2: Optimal thresholdselect (Optimal
Threshold)}\label{ux63a8ux8bba-2ux6700ux4f18ux9608ux503cux9009ux62e9-optimal-threshold}

totodeterminenoise rate \(\eta\)、Category number \(K\)、expert number \(M\) andmaximum clear clean incorrect rate 
\(\mu_{\max} = \max_s \mu_s\), Optimal detectionthreshold \(\theta^*\) shouldsatisfies: 

\[\theta^* = \arg\max_\theta \min_s \Delta_s(\theta) = \frac{1}{2}\left(1 + \mu_{\max} \cdot \frac{K-2}{K-1}\right)\]

thistimeminimumseparation gapis: 

\[\Delta_{\min}^* = \frac{1}{2}\left(1 - \mu_{\max} \cdot \frac{K}{K-1}\right)\]

 mustmostfewexpert number (reach to F1 \(\geq 1 - \varepsilon_0\))is: 

\[M \geq \frac{1}{2\Delta_{\min}^{*2}} \log\left(\frac{1}{\eta \varepsilon_0}\right)\]

\textbf{Proof}: direct solve 
\(1 - \frac{1}{\eta}e^{-2M\Delta^2} \geq 1 - \varepsilon_0\) obtain 
\(e^{-2M\Delta^2} \leq \eta\varepsilon_0\), takelogarithmici.e. obtain. If needsametimecontrolsmany Statuserror (connect combined world), separate factor in the 
\(1\) replaceis Status number \(|\mathcal{S}|\). \(\square\)

\subsubsection{3.3 Corollary 3: consistentmaydetectionSufficient Condition (Uniform
Detectability)}\label{ux63a8ux8bba-3ux4e00ux81f4ux53efux68c0ux6d4bux7684ux5145ux5206ux6761ux4ef6-uniform-detectability}

such as result existencein \(\delta > 0\) makesto has Status \(s\) has 
\(\mu_s \leq \frac{K-1}{K} - \delta\), thenset
\(\theta = \frac{1}{2}\) (fixed threshold), to has Statushas 
\(\Delta_s \geq \frac{\delta}{2}\), and: 

\[\text{F1} \geq 1 - \frac{1}{\eta} \exp\!\left(-\frac{M\delta^2}{2}\right)\]

\textbf{meaning}: onlymusteach Status clear clean on averageexpertincorrect rate consistentbelow \((K-1)/K\) (when
\(K=2\) timeis \(1/2\)), usefixed threshold \(\theta = 1/2\) i.e.mayobtainexponential level different 
F1 guarantee. 

\subsubsection{3.4 Corollary 4: has finite sample calibrate currently (Finite-Sample
Correction)}\label{ux63a8ux8bba-4ux6709ux9650ux6837ux672cux6821ux6b63-finite-sample-correction}

atpracticemiddle, \(\mu_s\) mustFrom has finite verifydataestimate. setupeach Status \(s\) has \(n_s\)
 clear clean verifysample. use 
\(\hat{\mu}_s = \frac{1}{M}\sum_m \hat{\varepsilon}_m(s)\) estimate
\(\mu_s\), where \(\hat{\varepsilon}_m(s)\) isexpert \(m\) inStatus \(s\)
upper via verify incorrect rate. 

by Hoeffding inequalityand connect combined world: to any agree \(\delta_0 \in (0, 1)\), withprobability
\(\geq 1 - \delta_0\): 

\[|\hat{\mu}_s - \mu_s| \leq B \sqrt{\frac{\log(2M/\delta_0)}{2n_s}}, \quad \forall m, s\]

setup calibrate currentlybackthresholdis 
\(\tilde{\mu}_s = \hat{\mu}_s + B\sqrt{\frac{\log(2M/\delta_0)}{2n_s}}\) (preserve guard upper bound). then
Theorem 1 Conclusionto calibrate currentlypost-\(\tilde{\mu}_s\) withprobability
\(\geq 1 - \delta_0\) holds, onlymustin \(\Delta_s\) middleusing \(\tilde{\mu}_s\)
 replace generation \(\mu_s\). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4 tight cause ness discuss theory 
(Tightness)}\label{ux7d27ux81f4ux6027ux8ba8ux8bba-tightness}

\subsubsection{4.1 what time tight cause }\label{ux4f55ux65f6ux7d27ux81f4}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{large M extremely finite }: when
 \(M \to \infty\), \(C \xrightarrow{\text{a.s.}} \mathbb{E}[C]\) (bylaw of large numbers), detection change become determinism: \(C > \theta\)
 if and only if \(\mathbb{E}[C] > \theta\). thistime F1 by \(\eta\) and \(\mu_s\)
 complete full verdict determine, Theoremlower bound tight cause. 
\item
 \textbf{Statussameness}: when (A5)
 rigorousholds (Statusinner clear clean incorrect rate complete fulluniform uniform)andexpert on called, Hoeffding world often number because factor 
 \(2\) isOptimal (by Hoeffding lower boundguarantee). 
\item
 \textbf{\(K \to \infty\)}: separation gap \(\Delta_s \to \frac{1}{2}\) (fixed 
 \(\mu_s\)), detectionasymptotic complete beautiful. 
\end{enumerate}

\subsubsection{4.2 what time loose relax }\label{ux4f55ux65f6ux677eux5f1b}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{small M}: when \(M \leq 3\) time, Hoeffding
 world relatively loose relax. thistimeapplicationExact two term formula distribution (andnon-
 Hoeffding)may obtain more tight world. via verify rule then: \(M \geq 10\) time Hoeffding
 world reasonable tight cause; \(M \leq 5\) timeuseExact two term formula accumulate accumulate distribution (Corollary 4.1). 
\item
 \textbf{\(\mu_s\) approaches \(\frac{K-1}{K}\)}: when
 \(\mu_s \uparrow (K-1)/K\) time
 \(\Delta_s \to 0\), exponential decaybecause factor eliminate lose. thistime mustexpert number 
 \(M \propto 1/\Delta_s^2\)
 divergence. thisisthisness: whenexpertatnew clear clean dataupperapproachesrandom guess measure time, noise detection must howevernotmay reliable. 
\item
 \textbf{weakCharacteristics retreat -ization }:
 whenStatus plan separateNonemethodseparate from noisemain guide region domain and clear clean main guide region domain time, (A5) does not hold. thistime
 Theorem 1 not adapt use, must borrow help cross-validation or external deployment (See Prop 6
 two layerStatusdiscover). thisis SCX
 frameworkalready know boundary (participate see\emph{weakCharacteristics lose effect }document). 
\item
 \textbf{noisenon-uniform uniform }: when \(\eta\) -dependence \(x\)
 time (\(P(\text{noise} \mid x)\) non- often number), F1 table reach formula must use 
 \(\eta_s = \mathbb{P}(\text{noise} \mid X \in s)\) replace generation global \(\eta\): 
\end{enumerate}

\[\text{F1} \geq 1 - \frac{1}{\min_s \eta_s} \sum_s \rho_s \exp(-2M\Delta_s^2)\]

in \(\eta_s \to 0\) extremeStatus, detectionNonemeaning------shouldStatusthis self nothas noise. 

\subsubsection{4.3
 tight cause nessto ratio experiment}\label{ux7d27ux81f4ux6027ux5bf9ux6bd4ux5b9eux9a8c}

to classical type value \(K=10\)、\(\eta=0.1\)、\(M=20\)、\(\mu_s=0.2\), Theorem 1
to out: 

\[\Delta_s = \frac{1}{2}(1 - 0.2 \cdot 10/9) \approx 0.389, \quad \text{F1} \geq 1 - 10 \cdot e^{-2\cdot 20 \cdot 0.389^2} \approx 1 - 10 \cdot e^{-6.05} > 0.976\]

\textbf{Note agree }: upper describe computeAssumptions
\(\mu_s = 0.2\), i.e.expertat clear clean dataupperincorrect rate is 20\%. at actual actual CIFAR-10
experimentmiddle (\(M=20\), \(\eta=0.1\)), SCX-Noise F1 actual measure is 
\(0.617\). this value far below \(0.976\) lower bound, Description CIFAR-10
experimentmiddle\textbf{Assumptionsdoes not satisfy}------mostmayReasonis: (i) expertonlyin 3 epoch 
CPU trainingbackuse, actual actual \(\mu_s \approx 0.45\) (andnon-falsedetermine \(0.2\)), thistime
\(\Delta_s \approx 0.25\), lower bound retreat -ization is \(\text{F1} \geq 0.18\) (and actual measure 
\(0.617 > 0.18\) consistent); (ii) (A5)
StatussamenessatlowCharacteristicslowernotrigorousholds. \textbf{ teach train }: Theoremlower boundto \(\mu_s\)
highdegreesensitive, practicemiddlemustatverify set upper alignment confirm estimate \(\mu_s\) post-againapplicationFormula. in AlN v3
full GPU trainingscenariomiddle, \(\mu_s\) canlowto
\(0.03\text{--}0.18\), lower bound obtain with tight cause (seecurrently text Table 2). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5 and Dawid-Skene
relationship}\label{ux4e0e-dawid-skene-ux7684ux5173ux7cfb}

\subsubsection{5.1 Dawid-Skene
model time consider }\label{dawid-skene-ux6a21ux578bux56deux987e}

Dawid and Skene (1979) raise out using EM
algorithmFrom manymarkNoteone (expert)independentmarkNotemiddleestimatetrueLabelandmarkNoteonemay reliable ness. coreAssumptionsiseach markNoteone
\(m\) has a \textbf{globalconfusion matrix}: 

\[\pi_m(c' \mid c) = \mathbb{P}(f_m(x) = c' \mid y^* = c), \quad \forall x \in \mathcal{X}\]

i.e.markNoteoneincorrect model formula not-dependenceconcrete type this \(x\)------thisis a global often number. 

mostfinalLabelestimatethroughadd weight votecomplete: 

\[\hat{y}_{\text{DS}} = \arg\max_c \sum_{m=1}^M w_m \cdot \mathbf{1}\{f_m(x) = c\}\]

whereweight \(w_m\) by \(\pi_m\)
toangle dollar element (accuracy) verdict determine, ormoregenerally, byconfusion matrix inverse verdict determine. 

\subsubsection{5.2 SCX
improve: StatusConditionsmay reliable ness}\label{scx-ux7684ux6539ux8fdbux72b6ux6001ux6761ux4ef6ux53efux9760ux6027}

SCX willconfusion matrix push broad is \textbf{StatusConditions}form: 

\[\pi_m(c' \mid c, s) = \mathbb{P}(f_m(x) = c' \mid y^* = c, x \in s)\]

SCX weightbecausethis-dependenceatStatus: 

\[w_m(x) = \frac{\exp(-\alpha \hat{R}_m(s(x)))}{\sum_{m'} \exp(-\alpha \hat{R}_{m'}(s(x)))}\]

where \(\hat{R}_m(s) = \sum_{c} \pi_m(c \mid c, s)\) is StatusConditionsriskestimate. 

\subsubsection{5.3 SCX rigorousis superior to Dawid-Skene
Conditions}\label{scx-ux4e25ux683cux4f18ux4e8e-dawid-skene-ux7684ux6761ux4ef6}

\textbf{Proposition 5.1 (SCX-DS comparison).} set \(R_{\text{DS}}\) is 
Dawid-Skene add weight expect 0-1 loss, \(R_{\text{SCX}}\) is SCX
StatusConditionsadd weight expectloss. then: 

\[R_{\text{SCX}} \leq R_{\text{DS}}\]

 etc. number holdsif and only iftoeach markNoteone \(m\) and has Status
\(s, s' \in \mathcal{S}\), confusion matrixatStatusbetweennot change: 

\[\pi_m(\cdot \mid \cdot, s) = \pi_m(\cdot \mid \cdot, s'), \quad \forall s, s' \in \mathcal{S}\]

i.e.markNoteonemay reliable nessnot-dependencesampleStatus. 

\textbf{Proof.} to any agree 
\(x \in s\), add weight vote Bayesian leaf -ese Optimal weightisStatusConditionsweight
\(w_m^*(s)\), it minimum -ization point state expectloss: 

\[w^*(s) = \arg\min_{w \in \Delta^{M-1}} \mathbb{E}[\ell(\text{vote}_w(x), y^*) \mid x \in s]\]

Dawid-Skene useglobalweight \(\bar{w}\) is often number: 

\[\bar{w} = \arg\min_{w} \mathbb{E}_{X, Y^*}[\mathbb{E}[\ell(\text{vote}_w(X), Y^*) \mid X]]\]

byitem by item point Optimal ness, toeach \(x \in s\): 

\[\mathbb{E}[\ell(\text{vote}_{w^*(s)}(x), y^*) \mid x \in s] \leq \mathbb{E}[\ell(\text{vote}_{\bar{w}}(x), y^*) \mid x \in s]\]

atStatusdistributionuppertakeexpecti.e. obtain \(R_{\text{SCX}} \leq R_{\text{DS}}\). 

 etc. number Conditionsbyitem by item point inequalityrigorousness guide out: onlyhas whenglobalweightalsoiseach StatusOptimal weighttime etc. number holds, i.e.
\(\bar{w} \in \arg\min_{w} \mathbb{E}[\ell(\text{vote}_w(x), y^*) \mid x \in s]\)
to has \(s\)
holds. thisrequireeachStatusOptimal weightmutualsame, is equivalent toconfusion matrixnot-dependenceatStatus. \(\square\)

\subsubsection{5.4
atnoise detection language environment lowerconcreteDifference}\label{ux5728ux566aux58f0ux68c0ux6d4bux8bedux5883ux4e0bux7684ux5177ux4f53ux5deeux5f02}

Theorem 1
detection rule thenestablishat''consistentness''upper------noisesampleto has expertall present present highincorrect. thisis
SCX framework special has advantage, and Dawid-Skene globalmay reliable nessestimatehas thisDifference: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.1667}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3611}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4722}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dawid-Skene
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
SCX (Theorem 1)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
may reliable ness & global often number \(\pi_m\) & StatusConditions \(\pi_m(s)\) \\
noise detection & -dependencemarkNoteoneconsistentness\emph{back verify } & -dependencemulti-expertincorrect\emph{consistentness} \\
 mustmarkNote & needmustmany (through often \(\geq 3\))markNoteone & needmustmany expertmodel \\
 resist Categorynotbalance & weak (globalconfusion matrix receive main guide Category match) &
strong (Status level estimatenot receive impact) \\
distinguish Noise vs Hard & cannot (onlyto out back verify probability) & can (consistency score \(C\)
 sky howeverseparate from twoone) \\
theoryguarantee & asymptoticconsistentness (\(M \to \infty\)) & exponentialconvergence (\(M\) and \(K\)
allaccelerates) \\
\end{longtable}

\subsubsection{5.5 Dawid-Skene action is SCX
 special example }\label{dawid-skene-ux4f5cux4e3a-scx-ux7684ux7279ux4f8b}

whenStatus plan separate retreat -ization to flat all plan separate
\(\mathcal{S} = \{\mathcal{X}\}\) (i.e. has sampleat a Statusmiddle)time: 

\begin{itemize}
\tightlist
\item
 SCX weight retreat -ization is 
 \(\bar{w}_m \propto \exp(-\alpha \hat{R}_m)\), i.e.global often number weight
\item
 consistency score \(C(x)\) retreat -ization is globalon averageincorrect rate 
\item
 Theorem 1 detectionguarantee retreat -ization is -dependence
 \(\mu = \frac{1}{M}\sum_m \varepsilon_m\) globalguarantee
\end{itemize}

thistime SCX noise detectionis equivalent to Dawid-Skene globalmay reliable nessadd weight Methodology. thisproves SCX
frameworkrigorous push broad ness. 

\subsubsection{5.6
improve magnitude degreequantity}\label{ux6539ux8fdbux5e45ux5ea6ux7684ux91cfux5316}

SCX mutualfor Dawid-Skene improve magnitude degreeandStatusbetweenmay reliable ness change differentdirectly related: 

\[\Delta R = R_{\text{DS}} - R_{\text{SCX}} = \sum_{s \in \mathcal{S}} \rho_s \cdot \mathcal{D}_s\]

where
\(\mathcal{D}_s = \sum_{m=1}^M (\bar{w}_m - w_m^*(s)) \cdot R_m(s)\)
isStatus \(s\) upperuseglobalweightandnon-StatusConditionsweight amount outerrisk. 

by Proposition 3 (StatusConditionsweightProposition)Theorem 3.1: 

\[\Delta R \geq \frac{1}{\alpha} \cdot I(s; w)\]

where \(I(s; w) = \mathbb{E}_s[\text{KL}(w(s) \| \bar{w})]\)
isStatusandweight'sbetweenmutualinformation. 

\textbf{Conclusion}: SCX rigorousis superior to Dawid-Skene if and only ifexpertmay reliable nessStatus change different
\(I(s; w) > 0\)------thisatpracticemiddle several -ish totalisholds. Theorem 1 detectionguaranteeonlyis
SCX wholeframeworkadvantageatnoise detection any task upper set middle overall present. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{reference}\label{ux53c2ux8003ux6587ux732e}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 Dawid, A. P., \& Skene, A. M. (1979). Maximum likelihood estimation of
 observer error-rates using the EM algorithm. \emph{Journal of the
 Royal Statistical Society: Series C (Applied Statistics)}, 28(1),
 20-28.
\item
 Hoeffding, W. (1963). Probability inequalities for sums of bounded
 random variables. \emph{Journal of the American Statistical
 Association}, 58(301), 13-30.
\item
 Pollard, D. (1990). Empirical Processes: Theory and Applications.
 \emph{NSF-CBMS Regional Conference Series in Probability and
 Statistics}, 2, i-86.
\item
 Bartlett, P. L., \& Mendelson, S. (2002). Rademacher and Gaussian
 complexities: Risk bounds and structural results. \emph{Journal of
 Machine Learning Research}, 3(Nov), 463-482.
\item
 SCX Framework Definitions.
 \texttt{../definitions/01\_state\_conditioned\_risk.md}.
\item
 Proposition 3: State-Conditioned Expert Weighting.
 \texttt{../propositions/03\_state\_conditioned\_weighting.md}.
\item
 Proposition 4: Compression Fidelity.
 \texttt{../propositions/04\_compression\_fidelity.md}.
\item
 data prevent middle poison. \texttt{../../knowledge/02\_/data prevent middle poison.md}.
\item
 data four separate class. \texttt{../../knowledge/02\_/data four separate class.md}.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{appendix A: Chernoff
 world Derivation}\label{ux9644ux5f55-achernoff-ux754cux7684ux63a8ux5bfc}

Lemma 2 and Lemma 3 canwith make using Chernoff world (andnon-
Hoeffding)obtainmore tight exponentialrate. 

to clear clean data, set \(\{e_m\}\) is independent Bernoulli variable, with mean 
\(\mu = \mathbb{E}[C \mid \text{clean}, x] \leq \mu_s\). to
\(\theta > \mu\): 

\[\begin{aligned}
\mathbb{P}(C > \theta \mid \text{clean}, x) 
&= \mathbb{P}\!\left(\frac{1}{M}\sum_m e_m > \theta\right) \\
&\leq \inf_{\lambda > 0} e^{-\lambda M\theta} \mathbb{E}\!\left[e^{\lambda\sum_m e_m}\right] \\
&= \inf_{\lambda > 0} e^{-\lambda M\theta} \prod_{m=1}^M (1 - \mu_m + \mu_m e^\lambda) \\
&\leq \exp\!\left(-M \cdot \text{KL}(\theta \,\|\, \mu_s)\right)
\end{aligned}\]

where
\(\text{KL}(p \,\|\, q) = p\log\frac{p}{q} + (1-p)\log\frac{1-p}{1-q}\), andmostback a step use connect combined world and
\(\mu_m \leq \mu_s\). 

tonoisedata, by Lemma 3 correctionProof (todetermine \(c\) Conditionslower
Hoeffding), Chernoff formis: 

\[\mathbb{P}(C \leq \theta \mid \text{noise}, x, c) \leq \exp\!\left(-M \cdot \text{KL}\!\left(\theta \;\Big\|\; 1 - \mu_c(x)\right)\right) \leq \exp\!\left(-M \cdot \text{KL}\!\left(\theta \;\Big\|\; 1 - C_{\text{bal}} \cdot \frac{\mu_s}{K-1}\right)\right)\]

Note agree KL direction: \(\text{KL}(\theta \| q)\) toshould Chernoff upper bound
\(\mathbb{P}(\hat{p} \leq \theta) \leq \exp(-n \cdot \text{KL}(\theta \| p))\), where
\(\hat{p}\) is \(p\) samplemean. this inside 
\(p = \mathbb{E}[C \mid \text{noise}] = 1 - \mu_c(x) \geq 1 - C_{\text{bal}} \cdot \mu_s/(K-1)\), we s 
bound \(\mathbb{P}(C \leq \theta)\) (lower tail), withth two parameteristrueexpect
\(p \approx 1 - C_{\text{bal}} \cdot \mu_s/(K-1)\). 

\begin{quote}
\textbf{2026-06-27 correction}: original version Chernoff appendixuse
\(\text{KL}(1-\theta \| 1 - \mu_s/(K-1))\), KL directionhas error. correctformis 
\(\text{KL}(\theta \| 1 - C_{\text{bal}} \cdot \mu_s/(K-1))\)------lower tail 
bound middleth a parameteristhreshold \(\theta\), th two parameteristrueexpect. 
\end{quote}

Chernoff world is superior to Hoeffding world process degreebywithlowerinequality carve picture (Hoeffding is
Chernoff a orderapproximate): 

\[\exp(-2M\Delta^2) \geq \exp(-M \cdot \text{KL}(\mu + \Delta \,\|\, \mu))\]

 etc. number onlywhen \(\Delta \to 0\)
timeasymptoticholds. at actual actual parameter paradigm surround (\(\Delta \approx 0.1\text{-}0.4\))inner, Chernoff
 world may tight \(2\text{-}5\) times. 

\begin{quote}
\textbf{DEFECT-06 same step (2026-06-28) --- Bahadur-Rao lattice point correction}:
whenexpertincorrectis Bernoulli distribution (lattice point cross degree \(h=1\))time, correct Bahadur-Rao
 exhibit openin Chernoff world middleintroducebecause factor \((1-e^{-\lambda^*})^{-1}\) replace generation 
\(1/\lambda^*\). correctionbackextremelysmallextremelylarge often number \(C_{\min}^{\text{(corr)}}\)
and original begin \(C_{\min}\) mutual difference convention 
\(5\)--\(12\%\), extremelysmallextremelylargeOptimal nessConclusionnot change (may reach world andlower boundacceptmutualsame lattice point correctionbecause factor). detail fineDerivationsee
\texttt{06\_fixed\_point\_convergence.md} §12 (Remark on Bahadur-Rao
Lattice Correction). 
\end{quote}

\subsection{appendix B: non-0-1
loss push broad }\label{ux9644ux5f55-bux975e-0-1-ux635fux5931ux7684ux63a8ux5e7f}

Theorem 1 to a like has world loss \(\ell \in [0, B]\) alsoholds, onlymustwill Lemma 1 in the 
\(e_m\) explainsis 
\(\mathbf{1}\{\ell(f_m(x), y) > \tau\}\). coreDerivationnot change, becauseis Lemma 1
only-dependenceConditionsexpectlinearnessandnoiseLabeluniform uniform nessAssumptions (A4), not-dependence 0-1
lossconcreteform. 

specifically, to time return any task (\(\mathcal{Y} \subseteq \mathbb{R}^d\)), \(K\)
mustheavynewexplains. Lemma 1 in the \(K\) in time return scenariolower retreat -ization is: 

\[ \mathbb{E}[C \mid \text{noise}, x] \geq 1 - \mathbb{E}[C \mid \text{clean}, x] \]

i.e. \(K \to \infty\) extremely finite case. thistimeseparation gapto any agree \(\mu_s < 1/2\)
is currently. thisisbecauseis incontinuousLabel Spacemiddle, expert exactly goodpredictiontonoiseLabelprobabilityis zero, make noise detectionmoreeasy. Theorem
1 itsremainderpartnot change. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Changelog}\label{changelog}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1935}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2258}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2581}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3226}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Issue
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Change
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Severity
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
2026-06-28 & B1 & \textbf{Fixed A2 self-contradiction.} Updated A2
rationale to acknowledge A1 does NOT guarantee A2. Added
\(M_{\text{eff}}\) degradation formula and cross-reference to
\texttt{01\_symbol\_system.md} §12.5. & BLOCKING \\
2026-06-28 & B6 & \textbf{Added \(M \geq 1\), \(K \geq 2\) to Theorem 1
preamble.} Explicitly excludes degenerate edge cases where consensus
score is undefined or \(K-1=0\). & BLOCKING \\
2026-06-28 & B10 & \textbf{Synced DEFECT-06 (Bahadur-Rao lattice
correction)} to Theorem 1 appendix. Added cross-reference to
\texttt{06\_fixed\_point\_convergence.md} §12. & BLOCKING \\
\end{longtable}

\end{document}
