% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Theorem 2: Weak Feature Failure Lower
Bound}\label{theorem-2-weak-feature-failure-lower-bound}

\begin{quote}
whenCharacteristicsrepresentation \(\phi(x)\)
 contains the trueStatusinformationinsufficienttime, based onconsistentnessnoise detectionMethodology (includes
SCX)Nonemethodis superior tolossbaseline. thisTheoremFrom information theoryangledegreequantifiesthis a boundaryConditions. 
\end{quote}

\textbf{Related Experiments}: DermaMNIST (SCX-Health), AlN v3 MLIP (12-dim
 hand work Characteristics) \textbf{Related Code}: \texttt{src/scx/state/discovery.py},
\texttt{src/scx/valuation/noise\_score.py},
\texttt{src/scx/expert/reliability.py} \textbf{Related Concepts}:
{[}{[}weakCharacteristics lose effect {]}{]}, {[}{[}State-Conditioned eXpertise{]}{]}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Table of Contents}\label{ux76eeux5f55}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \hyperref[1-ux7b26ux53f7ux4e0eux8bbeux5b9a]{Notation and Setup}
\item
 \hyperref[2-ux5b9aux4e49ux5f31ux7279ux5f81]{Definition: Weak Features}
\item
 \hyperref[3-ux5f15ux7406-1ux72b6ux6001ux4f30ux8ba1ux8befux5deeux4e0bux754c]{Lemma
 1: Statusestimateerrorlower bound}
\item
 \hyperref[4-ux5f15ux7406-2scx-ux4f30ux8ba1ux9000ux5316]{Lemma 2: SCX
 estimate retreat -ization }
\item
 \hyperref[5-ux5b9aux7406-2ux5f31ux7279ux5f81ux5931ux8d25ux4e0bux754c]{Theorem
 2: weakCharacteristicsfaillower bound}
\item
 \hyperref[6-ux63a8ux8bba]{Corollary}
\item
 \hyperref[7-ux5b9eux7528ux89e3ux8bfb]{Practical Interpretation}
\item
 \hyperref[8-ux4e0e-theorem-1-ux7684ux8054ux7cfb]{and Theorem 1 connection}
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Notation and Setup}\label{ux7b26ux53f7ux4e0eux8bbeux5b9a}

\subsubsection{1.1 Data Generation}\label{ux6570ux636eux751fux6210}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2727}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2727}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4545}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Symbol/Concept
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Meaning
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Value Space
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(X\) & input enter random variable & \(\mathcal{X} \subseteq \mathbb{R}^d\) \\
\(Y\) & Labelrandom variable & \(\mathcal{Y}\) \\
\(S = s(X)\) & \textbf{notobservation}trueStatus &
\(\mathcal{S} = \{1,\dots,K\}\) \\
\(\phi(X)\) & \textbf{observation}toCharacteristicsrepresentation &
\(\Phi \subseteq \mathbb{R}^{d_\phi}\) \\
\(Z\) & noise indicate show variable (\(1\)=noise, \(0\)=dry clean) & \(\{0,1\}\) \\
\(\{f_m\}_{m=1}^M\) & \(M\) expert & \(\mathcal{X} \to \mathcal{Y}\) \\
\(D_m = \ell(f_m(X), Y)\) & expert \(m\) loss & \(\mathbb{R}^+\) \\
\(\eta = P(Z = 1)\) & edge actual noise rate & \([0,1]\) \\
\end{longtable}

\subsubsection{1.2
Markov chainstructure}\label{ux9a6cux5c14ux53efux592bux94feux7ed3ux6784}

SCX frameworkAssumptionswithlowergenerateprocess: 

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Z \to S \to (X, Y) \to φ(X)}
\NormalTok{ \to \{f\_m(X)\}}
\end{Highlighting}
\end{Shaded}

i.e.: 1. noiseStatus \(Z\) verdict determinesampletrueStatus
\(S\) (noisefromsome some data region domain) 2. Status \(S\) verdict determinedata point \((X, Y)\)
distribution 3. \(\phi(X)\) isFrom \(X\) raise takeCharacteristics (determinismorrandomfunction) 4.
expertprediction \(\{f_m(X)\}\) is \(X\) function

\textbf{keyMarkov chain (dataat reason inequalityapplication)}: 

\[Z \to S \to X \to \phi(X)\]

bythismay obtain: 

\[I(\phi(X); Z) \leq I(X; Z) \leq I(S; Z) \leq H(Z) \leq \log 2\]

\subsubsection{1.3 SCX
noise detection flow process }\label{scx-ux566aux58f0ux68c0ux6d4bux6d41ux7a0b}

SCX noise detectionComplete flow process is: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{Statusdiscover}: in \(\{\phi(x_i)\}\)
 upper motion linesclusteralgorithm (k-means/spectrumcluster/GMM), obtain to \(K\) estimateStatus
 \(\hat{\mathcal{S}} = \{\hat{s}_1, \dots, \hat{s}_K\}\), with and Statusseparate match function
 \(\hat{s}: \mathcal{X} \to \hat{\mathcal{S}}\)
\item
 \textbf{may reliable nessestimate}: toeach estimateStatus \(\hat{s}\) andexpert \(f_m\), compute: 

 \[SCX_m(\hat{s}) = \frac{|\{x_i \in \hat{s}: \ell(f_m(x_i), y_i) < \tau\}|}{|\{x_i \in \hat{s}\}|}\]
\item
 \textbf{consistentnesscompute}: Status \(\hat{s}\) innerpartconsistentness: 

 \[C(\hat{s}) = \begin{cases}
 \frac{p_{\max} - 1/|\mathcal{Y}|}{1 - 1/|\mathcal{Y}|}, & \text{discreteLabel}\\
 \frac{1}{1 + \text{Var}(Y \mid X \in \hat{s})}, & \text{continuousLabel}
 \end{cases}\]
\item
 \textbf{noise score}: type this \(x\) noise score: 

 \[NS(x) = \underbrace{r(x)}_{\text{residual}} \cdot \underbrace{\frac{w_\rho}{\rho(\hat{s}(x)) + \varepsilon}}_{\text{densedegree term }} \cdot \underbrace{(1 - C(\hat{s}(x)))}_{\text{is inconsistent term }}\]

 where \(r(x)\) issampleresidual/loss, \(\rho(\hat{s})\) isestimateStatusProportion. 
\item
 \textbf{detection}: throughthreshold -ization \(NS(x)\) obtain tonoiseLabel
 \(\hat{z}(x) = \mathbf{1}\{NS(x) > t\}\). 
\end{enumerate}

\subsubsection{1.4
lossbaselinedetection device }\label{ux635fux5931ux57faux7ebfux68c0ux6d4bux5668}

Definitionnot-dependenceStatusinformationlossbaselinedetection device: 

\[\hat{z}_{\text{loss}}(x) = \mathbf{1}\{\max_m D_m(x) > t\}\]

shoulddetection device onlyuseexpertloss \(\{D_m\}\), does not use \(\phi\) orStatusinformation. note itsOptimal 
F1 is: 

\[F1_{\text{base}} = \max_{t \in \mathbb{R}^+} F1\left(\hat{z}_{\text{loss}}\right)\]

similarly, \(AUC_{\text{base}}\) and \(PRAUC_{\text{base}}\) separate different is its
ROC-AUC and PR-AUC. 

\subsubsection{1.5 randombaseline}\label{ux968fux673aux57faux7ebf}

for complete fulldoes not use any what datainformationrandomdetection device, itsOptimal performanceis: 

\textbf{Lemma 0 (randombaseline)}: setup \(\eta = P(Z = 1)\)
is edge actual noise rate. mostbadcaselowernotis superior towithlowerrandombaseline: 

\[\begin{aligned}
AUC_{\text{rand}} &= 0.5 \\
PRAUC_{\text{rand}} &= \eta \\
F1_{\text{rand}} &= \max_{q \in [0,1]} \frac{2\eta q}{\eta + q} = \frac{2\eta}{1 + \eta}
\end{aligned}\]

if and only if \(F1_{\text{base}} = F1_{\text{rand}}\)
time, lossbaseline retreat -ization is randombaseline------this emit life atnoise andexpertlossindependenttime (example such asuniform uniform Labelnoise). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Definition: Weak Features}\label{ux5b9aux4e49ux5f31ux7279ux5f81}

\textbf{Definition 1 (\(\delta\)-weakCharacteristics)}: called Characteristicsmap
\(\phi: \mathcal{X} \to \Phi\) regardingtrueStatusmap
\(s: \mathcal{X} \to \mathcal{S}\) is \(\delta\)-weak, If: 

\[I(\phi(X); s(X)) \leq \delta\]

where \(I(\cdot; \cdot)\) is Shannon mutualinformation (with nat is single). 

\textbf{explains}: - \(\delta = 0\): \(\phi(X)\) and \(S\) independent, \(\phi\)
 complete fulldoes not containStatusinformation - \(\delta\) verysmall: \(\phi\) several -ish does not containStatusinformation -
\(\delta\) verylarge (such as \(\delta \gg \log K\)): \(\phi\)
containssufficientdistinguishStatusinformation

\textbf{normalizationweakdegree}: byat
\(I(\phi; S) \leq H(S) \leq \log K\), Definitionnormalizationweakdegree: 

\[\varepsilon_\phi = \frac{\delta}{\log K} \in [0, 1]\]

\(\varepsilon_\phi = 1\)
toshould complete fullNoneinformation (mostweak), \(\varepsilon_\phi = 0\) toshould complete fullinformation (moststrong). 

\textbf{ replace generation Definition (Statusmaydistinguishness)}: when \(I(\phi; S)\)
 difficult withdirectcomputetime, may use withlowerequivalentDefinition: 

to any agree two trueStatus \(s_1, s_2 \in \mathcal{S}\), Definition
\(TV_{s_1,s_2} = TV(P_{\phi|S=s_1}, P_{\phi|S=s_2})\) is \(\phi\)
distributiontotal variation distance from. thenCharacteristicsweakdegreeis: 

\[\Delta_\phi = \max_{s_1 \neq s_2} TV(P_{\phi|S=s_1}, P_{\phi|S=s_2})\]

by Pinsker inequality: \(\Delta_\phi \leq \sqrt{\delta / 2}\). when
\(\Delta_\phi\) verysmalltime, fromis differenttrueStatus point in \(\phi\) spacemiddleNonecannot distinguish. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Lemma
1: Statusestimateerrorlower bound}\label{ux5f15ux7406-1ux72b6ux6001ux4f30ux8ba1ux8befux5deeux4e0bux754c}

\textbf{Lemma 1 (Statusestimateerrorlower bound------Fano inequality)}: setup \(\hat{S}\)
isbased on \(\phi(X)\) any agree trueStatusestimate device (i.e. SCX
Statusdiscoveralgorithm input out). If \(\phi\) is \(\delta\)-weak, then: 

\[P(\hat{S} \neq S) \geq \frac{H(S) - \delta - \log 2}{\log K}\]

where \(K = |\mathcal{S}|\) is trueStatus number. 

\textbf{\textgreater{} FIXED (2026-06-28, DEFECT-10):}
 has mutualinformation, entropy, logarithmic term uniformwith \textbf{nats (self-howeverlogarithmic, bottom \(e\))}
is single. \(\log 2 \approx 0.6931\) nats is Fano inequalitycorrection term (is equivalent to 1
bit). \(\log K\) also is self-howeverlogarithmic. when \(K=2\)
andStatusuniform uniform time, \(H(S) = \log 2\), thistimeseparate factor 
\(H(S) - \delta - \log 2 = -\delta \leq 0\), world is \textbf{ empty general (vacuous)}. thisis
Fano inequalityalready know limitationness: onlyatConditionsentropy \(H(S \mid \phi(X))\)
sufficientsmalltimeprovidehas meaninglower bound. fornon-uniform uniform Statusdistribution (such as \(K=10\)
uniform uniform, \(H(S) = \ln 10 \approx 2.303\) nats), world at
\(\delta < 2.303 - 0.693 = 1.610\) nats timeis currently. 

\textbf{\textgreater{} FIXED (2026-06-28, DEFECT-11):} Lemma 1
is applicable toFrom \textbf{single }observation \(\phi(X)\) estimateStatus \(S\). butat actual actual middle, SCX
throughto \textbf{ has \(n\) sample} \(\{\phi(x_i)\}_{i=1}^n\)
cluster come sametimeestimate has Status. manysampleclusterhas effect mutualinformationis 
\(I(\phi(X_1),\ldots,\phi(X_n); S_1,\ldots,S_n)\), byatclusterintroduce-dependenceness, should value may\textbf{is much larger than}
\(n \cdot \delta\). becausethis, Theorem 2 middleuse
\(\delta = I(\phi(X); S)\) (singlesamplemutualinformation) world is\textbf{ preserve guard (conservative)}------ it high estimate 
SCX retreat -ization process degree. correctmanysample world is: 
\[P(\hat{S} \neq S) \geq \frac{H(S) - \frac{1}{n}I(\Phi^n; S^n) - \frac{\log 2}{n}}{\log K}\]
byat Theorem 2 is SCX improve\textbf{upper bound} (Description SCX Nonemethod super the morebaselineexceeds
\(C_F\sqrt{\delta/2}\)), usesinglesample world is preserve guard andbecausethishas effect. \textbf{this a preserve guard ness agree taste -ing: whenclustermay benefit use manysampleinformationtime, SCX
atpracticemiddlemayis superior to Theorem 2 prediction. }

\textbf{Proof}: by Fano inequality: 

\[\begin{aligned}
P(\hat{S} \neq S) &\geq \frac{H(S \mid \phi(X)) - \log 2}{\log K} \\
&= \frac{H(S) - I(\phi(X); S) - \log 2}{\log K} \\
&= \frac{H(S) - \delta - \log 2}{\log K}
\end{aligned}\]

whereth a equality make using \(I(\phi; S) = H(S) - H(S \mid \phi)\). \(\square\)

\textbf{Corollary
1.1 (uniform uniform Status)}: If trueStatusapproximateuniform uniform distribution, \(H(S) \approx \log K\), then: 

\[P(\hat{S} \neq S) \geq 1 - \frac{\delta + \log 2}{\log K}\]

when \(\delta \to 0\) and \(K\)
fixed time: \(P(\hat{S} \neq S) \to 1 - \frac{\log 2}{\log K} > 0\). tolarge
\(K\) this a extremely finite approaches \(1\). 

\textbf{Corollary 1.2 (smallStatus number)}: when \(K = 2\) time: 

\[P(\hat{S} \neq S) \geq \frac{H(S) - \delta - \log 2}{\log 2} = H(S) - \delta - \log 2\]

byat \(H(S) \leq \log 2 = 1\) nat, when \(\delta = 0\) timelower boundis 
\(H(S) - \log 2\), mayis 
\(0\) (butthisisnotcurrentlythen -ization lower bound). two separate class timerequiremorestrongAnalysis. 

\textbf{Corollary 1.3 (estimateerrormutualinformationupper bound------existenceatness)}: Fano
inequality inverse (achievability bound)indicateexistenceatsome in \(\phi\) upper operation action estimate device 
\(\hat{S}\) makes: 

\[P(\hat{S} \neq S) \leq \frac{\delta + \log 2}{\log K}\]

butthisis a \textbf{existenceatness}Conclusion: maximum likelihoodestimate device (Bayes-optimal classifier)can reach tothis world, and does not guarantee any agree special determinealgorithm (such as
k-means)alsocan reach to. 

 actual actual middle, SCX make using k-means clusterStatusestimateerrorby k-means
objective functioncontrols, and through rate lose truetheory (rate-distortion theory)and
\(I(\phi; S)\) mutualclose connect. k-means errorupper bound a like is: 

\[P(\hat{S} \neq S) \leq O\left(\frac{\delta}{\log K}\right) + \varepsilon_{\text{k-means}}\]

where \(\varepsilon_{\text{k-means}}\) is k-means
approximateerror (becauseinitialization -ization , non-convexness,  cluster shape state non- sphere shape etc. because element). whendatain \(\phi\)
spacemiddlemayseparatenessgoodtime \(\varepsilon_{\text{k-means}}\) small; when \(\delta\)
verysmall (weakCharacteristics)timeStatusdiscoveratrootthisupperDifficulty, any what algorithm (includes k-means)allNonemethodbreakthrough
Fano lower bound. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Lemma 2: SCX
estimate retreat -ization }\label{ux5f15ux7406-2scx-ux4f30ux8ba1ux9000ux5316}

\textbf{Lemma 2 (SCX can reliable nessestimate retreat -ization)}: setup \(C(s)\) is trueStatus \(s\)
consistentness, \(C(\hat{s})\) is based on \(\phi\) clusterestimateStatus \(\hat{s}\)
consistentness. If \(\phi\) is \(\delta\)-weak, then: 

\[\mathbb{E}\left[|C(\hat{S}) - C(S)|\right] \leq 2 \cdot P(\hat{S} \neq S) + O\left(\frac{1}{\sqrt{n_{\min}}}\right)\]

where \(n_{\min} = \min_{s \in \mathcal{S}} n_s\)
is trueStatusminimumsample quantity. 

\textbf{Proof}: willexpectdecomposeis correctestimateandincorrectestimatetwopart: 

\[\begin{aligned}
\mathbb{E}\left[|C(\hat{S}) - C(S)|\right]
&= P(\hat{S} = S) \cdot \mathbb{E}[|C(\hat{S}) - C(S)| \mid \hat{S} = S] \\
&\quad + P(\hat{S} \neq S) \cdot \mathbb{E}[|C(\hat{S}) - C(S)| \mid \hat{S} \neq S]
\end{aligned}\]

\textbf{ situation shape 1 (\(\hat{S} = S\))}: whenStatuscorrectestimatetime, \(C(\hat{S})\) is
\(C(S)\) consistentestimate quantity. by Chebyshev inequalityand \(C(\cdot) \in [0, 1]\): 

\[\mathbb{E}[|C(\hat{S}) - C(S)| \mid \hat{S} = S] \leq \frac{\sigma_C}{\sqrt{n_s}} \leq O\left(\frac{1}{\sqrt{n_{\min}}}\right)\]

where \(\sigma_C\) is \(C\) estimateStandard difference. 

\textbf{ situation shape 2 (\(\hat{S} \neq S\))}: whenStatusestimateincorrecttime, \(C(\hat{S})\)
isis differenttrueStatusmixedcombinedconsistentness: 

\[C(\hat{S}) = \sum_{s \in \mathcal{S}} P(S = s \mid \hat{S}) \cdot C(s)\]

byat \(C(\cdot) \in [0, 1]\): 

\[|C(\hat{S}) - C(S)| \leq 1\]

becausethis
\(\mathbb{E}[|C(\hat{S}) - C(S)| \mid \hat{S} \neq S] \leq 2\) (more tight world: \(\leq 2\)
becauseis two value uniformin \([0,1]\)). 

\textbf{merge}: 

\[\mathbb{E}[|C(\hat{S}) - C(S)|] \leq P(\hat{S}=S) \cdot O(1/\sqrt{n_{\min}}) + P(\hat{S}\neq S) \cdot 2\]

\[\leq 2 \cdot P(\hat{S} \neq S) + O(1/\sqrt{n_{\min}})\]

\(\square\)

\textbf{Corollary 2.1 (retreat -ization toglobalconsistentness)}: byLemma 1 and 
\(\delta \to 0\), \(P(\hat{S} \neq S) \to 1 - \log 2 / \log K\). thistime: 

\[C(\hat{S}) \xrightarrow{p} \sum_{s \in \mathcal{S}} \rho(s) C(s) = \bar{C}\]

where \(\rho(s) = P(S = s)\) isStatus first verify probability, \(\bar{C}\)
isglobalon averageconsistentness. i.e.: whenCharacteristics complete fullNoneinformationtime, has estimateStatusconsistentness tend towardmutualsame value 
\(\bar{C}\). 

\textbf{Corollary 2.2 (SCX
noise score retreat -ization ------mustStatusbalanceAssumptions)}: atsame type Conditionslower, noise score retreat -ization is: 

\[NS(x) \to r(x) \cdot \frac{w_\rho}{\rho(\hat{s}(x)) + \varepsilon} \cdot (1 - \bar{C})\]

byat \(\bar{C}\) isglobal often number, evaluate separatesortby
\(r(x) / (\rho(\hat{s}(x)) + \varepsilon)\) verdict determine. 

\textbf{StatusbalanceAssumptions}: If estimateStatusapproximate\textbf{balance}, i.e.maximumStatusProportionandminimumStatusProportion's ratio has world: 

\[\frac{\max_{\hat{s}} \rho(\hat{s})}{\min_{\hat{s}} \rho(\hat{s})} \leq R\]

tosome often number \(R\), then \(\rho(\hat{s}(x))\) to has \(\hat{s}\) sameorder, atis: 

\[NS(x) \propto r(x)\]

i.e.noise score retreat -ization is residualmonotonicfunction, SCX detection device SCX degenerates to the loss baselinedetection device. 

If Statusextremelynotbalance (example such as a estimateStatus occupy 90\% sample), then \(\rho(\hat{s})\)
 term willleads tosortbias from purelosssort. thistime retreat -ization isapproximate, amount outerbias difference fromStatusProportionestimateerror, is 
\(O(1/\sqrt{n})\) order of magnitude. atpracticemiddle, maythroughcheck
\(\{\rho(\hat{s})\}_{\hat{s} \in \hat{\mathcal{S}}}\)
 change differentcoefficient come Judge balanceness: change differentcoefficient \(> 1\) timebalanceAssumptionsmaydoes not hold. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Theorem
2: weakCharacteristicsfaillower bound}\label{ux5b9aux7406-2ux5f31ux7279ux5f81ux5931ux8d25ux4e0bux754c}

\subsubsection{5.1 Theorem Statement}\label{ux5b9aux7406ux9648ux8ff0}

\textbf{Theorem 2 (weakCharacteristicsfaillower bound)}. setup \(\phi: \mathcal{X} \to \Phi\)
isregardingtrueStatus \(S\) \(\delta\)-weakCharacteristicsmap. set \(h_{\text{SCX}}\)
is inupper describe flow process lower motion action SCX
noise detection device. AssumptionsclusteralgorithmproducesestimateStatusapproximatebalance (\(\max_{\hat{s}} \rho(\hat{s}) / \min_{\hat{s}} \rho(\hat{s}) \leq R\)
tosome often number \(R\)), andclusteralgorithmthis self notintroduce super the more Fano lower bound amount outererror. then: 

\textbf{(a) AUC world }: 

\[AUC(h_{\text{SCX}}) \leq AUC_{\text{base}} + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)\]

where \(AUC_{\text{base}}\) islossbaselinedetection device ROC-AUC, \(\eta = P(Z=1)\)
is edge actual noise rate. 

\textbf{(b) PR-AUC world }: 

\[PRAUC(h_{\text{SCX}}) \leq PRAUC_{\text{base}} + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)\]

where \(PRAUC_{\text{base}}\) islossbaselinedetection device PR-AUC. 

\textbf{(c) F1 world }: existenceat through use often number \(C_F\) (depends onminimumaccuracy/recall, through often 
\(C_F \leq 2\), seelower direction Derivation), makes: 

\[F1(h_{\text{SCX}}) \leq F1_{\text{base}} + C_F \cdot \sqrt{\frac{\delta}{2}}\]

where \(F1_{\text{base}}\) islossbaselinedetection device atOptimal thresholdlower F1 separate number. 

\textbf{Note}: AUC and PR-AUC world in the \(\eta\)
-dependenceoriginates fromthistwo indicate markinvolvingConditionsprobability. AUC =
\(P(\text{score}_{\text{noise}} > \text{score}_{\text{clean}})\)
requireseparate different From \(Z=1\) (noise)and
\(Z=0\) (dry clean)middleindependentsampling, thisamplifytotal variation world (see 5.2 section Step
5). \(\eta\)
the moresmall (noisethe moredilutehas), world the morewide loose ------This reflects the inherent difficulty of detecting rare noise.F1
 world Noneneed \(\eta\) -dependence, becauseis F1 isjoint distribution \(P(\hat{z}, Z)\)
function, total variation world direct adapt use. 

\subsubsection{5.2 Proof}\label{ux8bc1ux660e}

\textbf{Step 1: Constructionassistsdistribution \(\tilde{P}\)}

Definitionassistsdistribution \(\tilde{P}\), where \(\phi\) and \(S\) isstrong system independent: 

\[\tilde{P}(\phi, S) = P(\phi) \cdot P(S)\]

i.e. \(\tilde{P}\) maintains \(P(\phi)\) and \(P(S)\) edge edge distributionnot change, but cut break 
\(\phi\) and \(S\) 'sbetween-dependence. \(\tilde{P}\) lower has its he Conditionsdistributionand \(P\)
mutualsame (todetermine \(S\) post-\((X,Y)\) distributionnot change). 

\textbf{Step 2: KL scatter degreeandmutualinformationrelationship}

\(P\) and \(\tilde{P}\) 'sbetween KL scatter degreecurrentlygoodis \(\phi\) and \(S\) mutualinformation: 

\[KL(P \parallel \tilde{P}) = \iint P(\phi, S) \log \frac{P(\phi, S)}{P(\phi)P(S)} \, d\phi \, dS = I(\phi(X); S) = \delta\]

by Pinsker inequality: 

\[TV(P, \tilde{P}) \leq \sqrt{\frac{KL(P \parallel \tilde{P})}{2}} = \sqrt{\frac{\delta}{2}}\]

where \(TV(P, \tilde{P}) = \sup_A |P(A) - \tilde{P}(A)|\) istotal variation distance from. 

\textbf{Step 3: predictiondistribution TV world }

set \(P_{\text{pred}}\) and \(\tilde{P}_{\text{pred}}\) separate different is \(P\) and
\(\tilde{P}\) lower \((\hat{z}_{\text{SCX}}(X), Z)\)
joint distribution. bydataat reason inequality (detectionseparate number is original begin datafunction): 

\[TV(P_{\text{pred}}, \tilde{P}_{\text{pred}}) \leq TV(P, \tilde{P}) \leq \sqrt{\frac{\delta}{2}}\]

\textbf{Step 4: Analysis \(\tilde{P}\) lower SCX linesis }

in \(\tilde{P}\) lower, \(\phi \perp S\), becausethis \(\phi\) does not contain any what regarding
\(S\) information. byCorollary 2.2, SCX detection device SCX degenerates to the loss baselinedetection device: 

\[NS_{\tilde{P}}(x) \propto \max_m \ell(f_m(x), y)\]

becausethisin \(\tilde{P}\) lower: 

\[AUC_{\tilde{P}}(h_{\text{SCX}}) = AUC_{\text{base}}, \quad PRAUC_{\tilde{P}}(h_{\text{SCX}}) = PRAUC_{\text{base}}, \quad F1_{\tilde{P}}(h_{\text{SCX}}) = F1_{\text{base}}\]

\textbf{Step 5: TV world to AUC/PRAUC/F1 turn -ization }

AUC canwith write is expectform: 

AUC = \(P(\text{score}_{\text{noise}} > \text{score}_{\text{clean}})\)
involving\textbf{two independentsample}: a fromnoiseConditionsdistribution
\(P(\text{score} \mid Z=1)\), a fromdry clean Conditionsdistribution
\(P(\text{score} \mid Z=0)\). becausethis AUC isConditionsdistribution's accumulate expect. 

bytotal variationmaymultiplyness, toConditionsdistribution's accumulate: 

\[\begin{aligned}
&TV\left(P(\cdot \mid Z=1) \times P(\cdot \mid Z=0),\; \tilde{P}(\cdot \mid Z=1) \times \tilde{P}(\cdot \mid Z=0)\right) \\
&\leq TV(P(\cdot \mid Z=1), \tilde{P}(\cdot \mid Z=1)) + TV(P(\cdot \mid Z=0), \tilde{P}(\cdot \mid Z=0))
\end{aligned}\]

 connect lower come will edge edge TV world turn shift toConditionsdistribution. to any agree matter file \(A\): 

\[\begin{aligned}
|P(A \mid Z=1) - \tilde{P}(A \mid Z=1)|
&= \frac{|P(A \cap \{Z=1\}) - \tilde{P}(A \cap \{Z=1\})|}{\eta} \\
&\leq \frac{TV(P, \tilde{P})}{\eta} \leq \frac{1}{\eta} \sqrt{\frac{\delta}{2}}
\end{aligned}\]

similarly, \(TV(P(\cdot \mid Z=0), \tilde{P}(\cdot \mid Z=0)) \leq \frac{1}{1-\eta} \sqrt{\frac{\delta}{2}}\). 

combined rise come obtain to: 

\[\begin{aligned}
|AUC_P(h) - AUC_{\tilde{P}}(h)|
&\leq TV(P(\cdot \mid Z=1), \tilde{P}(\cdot \mid Z=1)) + TV(P(\cdot \mid Z=0), \tilde{P}(\cdot \mid Z=0)) \\
&\leq \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)
\end{aligned}\]

similarly, PR-AUC ineach thresholdupperinvolvingConditionsprobability Precision and Recall, same a amplify
TV world adapt use (at accumulate separateupperapplicationmutualsameargumentation, each point amplifybecause factor mutualsame). 

for \textbf{F1}, Note agree to F1 canwithdirect write is joint distribution \(P(\hat{z}, Z)\)
function: 

\[F1 = \frac{2 \cdot TP}{2 \cdot TP + FP + FN}\]

where
\(TP = P(\hat{z}=1, Z=1)\), \(FP = P(\hat{z}=1, Z=0)\), \(FN = P(\hat{z}=0, Z=1)\)
uniformis connect combinedprobability. becausethis F1 \textbf{does not involveConditionssampling}, TV world direct adapt use. 

F1 regarding connect combinedprobability \((TP, FP, FN)\) is Lipschitz continuous. specifically, F1
gradientsatisfies: 
\[\left|\frac{\partial F1}{\partial TP}\right| = \frac{2(FP+FN)}{(2TP+FP+FN)^2} \leq \frac{2}{\min(TP, 1)}\]
similarly, \(\left|\frac{\partial F1}{\partial FP}\right|, \left|\frac{\partial F1}{\partial FN}\right| \leq 2 / \text{denominator}\). becausethis
Lipschitz often number \(C_F\) satisfies: 
\[C_F \leq \frac{2}{p_{\min}^2}, \quad p_{\min} = \min(2TP+FP+FN)\]

atdetection device currently often motion lines paradigm surround inner (accuracyandrecalluniform
\(\geq 0.1\)), \(p_{\min} \geq 0.2\), therefore
\(C_F \leq 2 / 0.04 = 50\)? in practicethis over at preserve guard. directcompute: when
\(TP \geq 0.05, FP \leq 0.45, FN \leq 0.05\) (toshould Precision
\(\geq 0.1\), Recall \(\geq 0.5\))time, through number value computemay obtain 
\(C_F \leq 2\). moregenerally: - when \(Precision, Recall \geq 0.1\)
time, \(C_F \leq 3\) - when \(Precision, Recall \geq 0.5\)
time, \(C_F \leq 1\)

See appendix C number value verify. atis: 

\[|F1_P(h) - F1_{\tilde{P}}(h)| \leq C_F \cdot TV(P_{\text{pred}}, \tilde{P}_{\text{pred}}) \leq C_F \cdot \sqrt{\frac{\delta}{2}}\]

\textbf{Step 6: merge}

\[\begin{aligned}
AUC_P(h_{\text{SCX}}) &\leq AUC_{\tilde{P}}(h_{\text{SCX}}) + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right) \\
&= AUC_{\text{base}} + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)
\end{aligned}\]

similarly: 

\[PRAUC_P(h_{\text{SCX}}) \leq PRAUC_{\text{base}} + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)\]

\[F1_P(h_{\text{SCX}}) \leq F1_{\text{base}} + C_F \cdot \sqrt{\frac{\delta}{2}}\]

\textbf{ special example }: when \(\eta = 0.5\) (noiseanddry clean sample etc. quantity)time, AUC/PR-AUC
amplifybecause factor is \(1/\eta + 1/(1-\eta) = 4\), world retreat -ization is 
\(4\sqrt{\delta/2} = 2\sqrt{2\delta}\). when \(\eta\) tend near 0 or 1
time, world divergence------this counter reflect a direct view matter actual: dilutehas Nonenoise detectionatinformationtheoreticallymore difficult, Theorem apply add convention bind morefew. \(\square\)

\subsubsection{5.3 TheoremInterpretation}\label{ux5b9aux7406ux89e3ux8bfb}

\textbf{to called ness}: shouldTheoremhas a to called lower bound: 

\[AUC(h_{\text{SCX}}) \geq AUC_{\text{base}} - \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right)\]

this agree taste -ing when \(\delta = 0\) time, SCX AUC \textbf{ exactly goodequals}lossbaseline
AUC. when \(\delta \to 0\) time, SCX \textbf{cannot}is superior tolossbaseline. when \(\eta\)
verysmalltimeshould world more loose ------buttake etc. Conditions (\(\delta=0\))not receive impact. 

\textbf{convergence rate }: atfixed \(\eta\) lowerconvergence rate is \(O(\sqrt{\delta})\), by
Pinsker inequalitymain guide. amount outer \(\eta\)
because factor ismultiplyness: noisethe moredilutehas, convergence often number the morelarge. 

\textbf{ often number }: \(C_F \cdot \sqrt{\delta/2}\) in the \(C_F\)
depends onaccuracy/recallminimum value: - when \(Precision, Recall \geq 0.1\)
time, \(C_F \leq 3\) - when \(Precision, Recall \geq 0.5\)
time, \(C_F \leq 1\)

\textbf{\(\eta\) Meaning}: when \(\eta\) verysmalltime (dilutehas Nonenoise), AUC and
PR-AUC world finite wide loose ------SCX needmustmorestrongCharacteristics (moresmall
\(\delta\))only thencanmay reliable ground acrossandlossbaseline difference distance. thisprovide a may operation action diagnose break: If test set in the noise rate 
\(\eta\) verylow (such as \(<0.1\)), SCX
mutual ratio lossbaselineadvantagerequiremoresufficientCharacteristicsinformation come support. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Corollary}\label{ux63a8ux8bba}

\subsubsection{\texorpdfstring{6.1 Corollary
1: complete full lose effect (\(\delta = 0\))}{6.1 Corollary 1: complete full lose effect (\textbackslash delta = 0)}}\label{ux63a8ux8bba-1ux5b8cux5168ux5931ux6548delta-0}

\textbf{Corollary 1 (complete full lose effect)}. If \(\phi(X)\) and \(S\) independent (i.e. \(\phi\)
 complete fullNoneinformation), then: 

\[AUC(h_{\text{SCX}}) = AUC_{\text{base}}, \quad F1(h_{\text{SCX}}) = F1_{\text{base}}\]

SCX
noise detection device SCX degenerates to the loss baselinedetection device. If noiseislossNoneinformation (example such asuniform uniform Labelnoise), then
\(F1_{\text{base}} = F1_{\text{rand}}\), SCX Nonemethod super the morerandom guess measure. 

\subsubsection{6.2 Corollary
2: lossNoneinformationnoise}\label{ux63a8ux8bba-2ux635fux5931ux65e0ux4fe1ux606fux566aux58f0}

\textbf{Corollary
2 (uniform uniform Labelnoiselower lose effect)}. setupnoiseis uniform uniform Labelnoise (Labelbyrandomflippedto any agree Category, and input enter 
\(x\) Noneclose). thenlossdistributionandnoise indicate show variableindependent: 

\[P(\ell \mid Z = 1) = P(\ell \mid Z = 0)\]

atthisConditionslower, lossbaselinedetection device retreat -ization is randombaseline: 

\[AUC_{\text{base}} = 0.5, \quad PRAUC_{\text{base}} = \eta, \quad F1_{\text{base}} = \frac{2\eta}{1 + \eta}\]

byTheorem 2, when \(\phi\) is \(\delta\)-weaktime: 

\[\begin{aligned}
AUC(h_{\text{SCX}}) &\leq 0.5 + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right) \\
PRAUC(h_{\text{SCX}}) &\leq \eta + \sqrt{\frac{\delta}{2}} \cdot \left(\frac{1}{\eta} + \frac{1}{1-\eta}\right) \\
F1(h_{\text{SCX}}) &\leq \frac{2\eta}{1 + \eta} + C_F\sqrt{\frac{\delta}{2}}
\end{aligned}\]

when \(\delta = 0\) time, SCX PR-AUC is \(\eta\) world reside, F1 is
\(2\eta/(1+\eta)\) world reside. 

\textbf{experimentverify (DermaMNIST)}: 

\begin{longtable}[]{@{}cccc@{}}
\toprule\noalign{}
noise rate \(\eta\) & PR-AUC randombaseline & SCX PR-AUC & Loss PR-AUC \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.1 & 0.100 & 0.101 & 0.105 \\
0.2 & 0.200 & 0.212 & 0.211 \\
0.4 & 0.400 & 0.396 & 0.395 \\
\end{longtable}

SCX and Loss baseline PR-AUC uniformapproachesrandombaseline \(\eta\), difference
\(\leq \sqrt{\delta/2} \cdot (1/\eta + 1/(1-\eta)) \approx 0.01-0.02\), andTheoremconsistent. 

\subsubsection{6.3 Corollary
3: minimuminformationthreshold}\label{ux63a8ux8bba-3ux6700ux5c0fux4fe1ux606fux9608ux503c}

\textbf{Corollary 3 (SCX work action mustminimummutualinformation)}. setupexpect SCX detection F1
tofewexceedsrandombaseline \(\Delta\) withupper (\(\Delta > 0\)). mustminimummutualinformationis: 

\[\delta_{\min} \geq 2\left(\frac{\Delta}{C_F}\right)^2\]

for AUC, consider \(\eta\) -dependence: 

\[\delta_{\min} \geq 2 \left(\Delta_{AUC} \cdot \eta(1-\eta)\right)^2\]

when \(\eta = 0.5\) time: \(\delta_{\min} \geq \Delta_{AUC}^2 / 8\). when
\(\eta\)
verysmalltime, should world extremelywide loose (Theorem apply add convention bind morefew), but actual actual detectionnot must canachievesthis a difference distance. 

\textbf{ show example }: If must SCX F1 exceeds F1\_base tofew \(0.05\) (i.e.
\(\Delta = 0.05\)), take \(C_F = 2\): 

\[\delta_{\min} \geq 2\left(\frac{0.05}{2}\right)^2 = 2 \cdot 0.000625 = 1.25 \times 10^{-3} \text{ nats}\]

this see similar verysmall, butNote agree \(\delta\) is \(\phi\) and \(S\)
'sbetweenmutualinformation, \(\delta\) value small agree taste \(\phi\)
 several -ish does not contain any what Statusinformation. mustwill F1 raise liter 0.1 withupper, require
\(\delta \geq 5 \times 10^{-3}\) nats. 

\textbf{ actual actual Interpretation}: byat \(\delta\) with nat is singleand
\(\log K \leq \delta\) upper bound, \(\delta \geq 10^{-3}\) toshould convention 
\(\exp(10^{-3}) \approx 1.001\) similar however ratio ------ agree taste \(\phi\)
onlymust carry carry extremelya small amount ofStatusinformationthencanproducesmaydetectionimprove. Theorem 2 in \(\delta\)
smalltimeprovide tight world, in \(\delta\) largetime world change loose, i.e. SCX
atstrongCharacteristicslowermaywithandshouldwhenlarge magnitude super the morebaseline. 

\subsubsection{6.4 Corollary
4: Status number impact}\label{ux63a8ux8bba-4ux72b6ux6001ux6570ux7684ux5f71ux54cd}

\textbf{Corollary 4 (Status number \(K\) andweakCharacteristics cross mutual)}. follow -ing trueStatus number \(K\)
increaselarge, maintainsmutualsame \(\delta\) need \(\phi\)
Characteristicsqualitymorehigh. specifically, normalizationweakdegree \(\varepsilon_\phi = \delta / \log K\)
 verdict determine SCX has effect ness: 

\begin{itemize}
\tightlist
\item
 \(\varepsilon_\phi \approx 1\): \(\phi\) several -ish does not containStatusinformation, SCX retreat -ization 
\item
 \(\varepsilon_\phi \approx 0\): \(\phi\) containssufficientStatusinformation, SCX
 cancanhas effect 
\item
 \(\varepsilon_\phi \approx 0.5\): SCX
 Statusdiscover convention has a halfcorrect rate, noise detectionparthas effect 
\end{itemize}

\textbf{ actual actual meaning}: for SCX Analysis, notshould blind goal increasesStatus number \(K\). morelarge \(K\)
mustmust \(\phi\) providemutualshouldmoremanyinformation (morelarge
\(\delta\)), whethertheneach estimateStatuswillmixedcombinedmany trueStatus, leads to
\(C(\hat{s}) \to \bar{C}\). 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Practical Interpretation}\label{ux5b9eux7528ux89e3ux8bfb}

\subsubsection{7.1
such as what diagnose break weakCharacteristics}\label{ux5982ux4f55ux8bcaux65adux5f31ux7279ux5f81}

at actual actual shouldusing SCX noise detectionfront, throughwithlower diagnose break checkCharacteristicsiswhethertooweak: 

\textbf{ diagnose break 1: Statusinnerconsistentness tend same check verify }

If has estimateStatusconsistentness \(C(\hat{s})\)
approximatemutual etc. (variancesmall), thenCharacteristicsmay over weak: 

\[\text{Var}(\{C(\hat{s})\}_{\hat{s} \in \hat{\mathcal{S}}}) < \tau_C\]

in DermaMNIST experimentmiddle, \(K=8\) StatusconsistentnessStandard difference 
\(\sigma_C \approx 0.03\), andCharacteristicsdimensional number \(d_\phi = 128\) SimpleCNN
toshould. 

\textbf{ diagnose break 2: randombaselinecomparison}

attraining set uppercomputepredictionlossbaseline \(AUC_{\text{base}}\) and
\(PRAUC_{\text{base}}\). If it s approachesrandombaseline (\(AUC \approx 0.5\),
\(PRAUC \approx \eta\)), thennoise detection any task this self very difficult, weakCharacteristics advance a step prevent SCX
 super the morethisbaseline. 

\textbf{ diagnose break 3: supervisesStatusmatches (such as result may use)}

If partdatahas trueStatusLabel (such as model propose /already know batch come source), compute \(\phi\)
clusterandtrueStatusadjust orchid virtue exponential (ARI): 

\[ARI(\hat{S}, S) < 0.1 \implies \text{Characteristicstooweak}\]

ARI \textless{} 0.1 representation \(\phi\) cluster several -ish does not containStatusinformation. 

\textbf{ diagnose break 4: mutualinformationestimate}

using Kraskov estimate device (kNN mutualinformationestimate)or MINE estimate
\(I(\phi(X); Y)\), and willitsand \(\log K\) comparison: 

\[\frac{\hat{I}(\phi(X); Y)}{\log K} < 0.2 \implies \text{Characteristicsmaytooweak}\]

\subsubsection{7.2
minimuminformationthreshold indicate south }\label{ux6700ux5c0fux4fe1ux606fux9608ux503cux6307ux5357}

based onTheorem 2 Corollary 3, providewithlower actual actual indicate south: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Recommendation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\varepsilon_\phi < 0.2\) & SCX cancanhas effect, continues make using \\
\(0.2 \leq \varepsilon_\phi \leq 0.5\) & SCX parthas effect, checkits he indicate mark \\
\(\varepsilon_\phi > 0.5\) & SCX notrecommend, addstrongCharacteristicsbackagain test \\
\(AUC_{\text{base}} < 0.55\) and \(\varepsilon_\phi > 0.5\) & SCX
noise detectionNotedeterminefail \\
\(AUC_{\text{base}} > 0.7\) & lossbaselinealreadyverystrong, SCX cancanassistsimprove \\
\(\eta < 0.1\) and \(\varepsilon_\phi > 0.3\) & dilutehas Nonenoiselower AUC world byamplify
\(\geq\) 10\(\times\), mustmorestrongCharacteristicsormoremanynoisesample come support \\
\(0.3 \leq \eta \leq 0.7\) & noise rate moderate, AUC/PR-AUC world Optimal (\(\eta=0.5\)
timeamplifybecause factor minimum, is 4) \\
\end{longtable}

where \(\varepsilon_\phi = I(\phi; S) / \log K\), \(\eta = P(Z=1)\)
is edge actual noise rate. 

\textbf{ special different for MLIP}: - 12
dimensional hand work describe symbol (CN, bondlong etc.): \(\varepsilon_\phi \approx 0.6-0.8\) (weak), toshould
AlN in the lose  - ACE describe symbol (100+
dimensional): \(\varepsilon_\phi \approx 0.1-0.3\) (sufficientstrong), toshould AlN
 in the succeed (F1=0.585)

\subsubsection{7.3 supplement rescue direction}\label{ux8865ux6551ux65b9ux5411}

when diagnose break indicateCharacteristics over weaktime: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 \textbf{strengthensCharacteristics}: usemorestrongdescribe symbol (ACE/SOAP/MACE
 etc.), oruseincorrect drive action Characteristicslearning (SCX twolayerdescribe symbol)
\item
 \textbf{decreasesStatus number }: reduce \(K\) withreduceto \(\delta\)
 must seek, make Statusestimatemoremay reliable 
\item
 \textbf{ change use globalMethodology}: when \(AUC_{\text{base}}\) still maytime, release discard SCX
 StatusConditions -ization, directuselossthresholddetection
\item
 \textbf{ any task heavyDefinition}: If noise detectionis impossible, change use its he Methodology (such as cross verifyconsistentness, time sequence different often detection)
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{8. and Theorem 1
connection}\label{ux4e0e-theorem-1-ux7684ux8054ux7cfb}

\subsubsection{8.1 Theorem 1 time consider }\label{theorem-1-ux56deux987e}

\textbf{Theorem 1 (pressure shrink preserve true, alreadyhas)}: setupStatus \(s\)
satisfieshas world loss, complexitydegreecontrols,  redundant remainder-sensitivenessclose connect andboundarypreservesAssumptions. SCX-Compress
with rule model \(n_s'\) and redundant remainderseparate number \(D(s)\) selectadd weight subset, withhighprobabilitysatisfies: 

\[\sup_{f \in \mathcal{F}} |R_S(f) - R_C(f)| \leq B \cdot (1 - D_{\text{eff}}(s)) \cdot \left(1 - \frac{n_s'}{N_s}\right) + O\left(B\sqrt{\frac{d}{n_s'} \cdot (1 - D_{\text{eff}}(s)) \cdot \left(1 - \frac{n_s'}{N_s}\right) \log\frac{1}{\delta}}\right)\]

where \(D_{\text{eff}}(s) = D(s) \cdot (1 - \text{Boundary}(s))\). 

\subsubsection{8.2
two Theoremmutual supplement ness}\label{ux4e24ux4e2aux5b9aux7406ux7684ux4e92ux8865ux6027}

Theorem 1 and Theorem 2 From currently counter two surface world determine SCX framework adapt use paradigm surround: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.1154}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4231}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.4615}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
dimension
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Theorem 1 (pressure shrink preserve true)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Theorem 2 (weakCharacteristicsfail)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{ time answer Problem} & SCX what time\textbf{can} work action ? & SCX
 what time\textbf{cannot} work action ? \\
\textbf{keyConditions} & good goodStatus plan separate + high redundant remainder & weakCharacteristics + lowmutualinformation \\
\textbf{core quantity} & has effect redundant remainder \(D_{\text{eff}}(s)\) & Characteristics-Statusmutualinformation
\(I(\phi; S)\) \\
\textbf{performanceguarantee} & error \(\varepsilon(s)\) upper bound & F1 does not exceedbaseline +
\(O(\sqrt{\delta})\) \\
\textbf{whenConditionssatisfiestime} & pressure shrink preserve truedegreehigh & SCX retreat -ization is globalMethodology \\
\textbf{whenConditions violate counter time} & pressure shrink Noneguarantee (mustmoremanysample) & SCX
may work action (world change loose) \\
\end{longtable}

\subsubsection{8.3 unified figure landscape }\label{ux7edfux4e00ux56feux666f}

two Theorem connect combined carve picture SCX CompletenessConditions: 

\[\text{SCX has effect ness} \approx \underbrace{I(\phi; S)}_{\text{Characteristicsinformation quantity }} - \underbrace{O(1/\sqrt{n_s})}_{\text{estimateerror}} - \underbrace{O(\sqrt{d/n_s'})}_{\text{generalization error}}\]

\begin{itemize}
\tightlist
\item
 \textbf{succeedConditions} (Theorem 1
 region domain): \(I(\phi; S) \gg O(1/\sqrt{n_s}) + O(\sqrt{d/n_s'})\), and
 \(D_{\text{eff}}(s)\) high
\item
 \textbf{failConditions} (Theorem 2
 region domain): \(I(\phi; S) \to 0\), None theory its he Conditionssuch as what, SCX
 noise detectionNonemethod super the morelossbaseline
\item
 \textbf{ over transition region }: \(I(\phi; S)\) middle etc., SCX parthas effect 
\end{itemize}

\subsubsection{8.4 to SCX
frameworkwholemeaning}\label{ux5bf9-scx-ux6846ux67b6ux7684ux6574ux4f53ux610fux4e49}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{necessarybutis insufficient}: Theorem 2 ProofstrongCharacteristicsis SCX
 noise detectionnecessaryConditions. Theorem 1 Proofhigh redundant remainderis SCX
 pressure shrink succeedSufficient Condition (atgoodCharacteristicslower). 
\item
 \textbf{Characteristics work process superior first }: two Theoremcommonsame indicate toward: atshouldusing SCX pre-shouldensure \(\phi\)
 carry carry sufficientStatusinformation. thisand SCX experimentmiddlediscoverconsistent: 

 \begin{itemize}
 \tightlist
 \item
 AlN (ACE, 100+ dimensional) \to Theorem 1 can use, Theorem 2 not tight 
 \item
 DermaMNIST (SimpleCNN, 128 dimensionalbutdistinguishdegree difference) \to Theorem 2 tight 
 \end{itemize}
\item
 \textbf{twolayerdescribe symbol strategy}: SCX twolayerdescribe symbol (Layer 1: physical/Domaindescribe symbol \to
 Layer 2: incorrect drive action clusterCharacteristics)currentlyisat raise liter \(I(\phi; S)\), will \(\delta\)
 From large change small. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{appendix A: Symbol/Concept table }\label{ux9644ux5f55-aux7b26ux53f7ux8868}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Symbol/Concept & Meaning & time out present \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\phi\) & Characteristicsmap & Definition 1 \\
\(\delta\) & mutualinformationupper bound & Definition 1 \\
\(K\) & trueStatus number & Lemma 1 \\
\(\eta\) & edge actual noise rate & Corollary 2 \\
\(D_m\) & expert \(m\) loss & th 1.4 section \\
\(C(s)\) & Statusconsistentness & th 1.3 section \\
\(NS(x)\) & noise score & th 1.3 section \\
\(\bar{C}\) & globalon averageconsistentness & Corollary 2.1 \\
\(\varepsilon_\phi\) & normalizationweakdegree & Definition 1 \\
\(F1_{\text{base}}\) & lossbaseline F1 & th 1.4 section \\
\(F1_{\text{rand}}\) & randombaseline F1 & th 1.5 section \\
\end{longtable}

\subsection{appendix
B: keyinequality}\label{ux9644ux5f55-bux5173ux952eux4e0dux7b49ux5f0f}

\textbf{Fano inequality} (Lemma 1): 

\[P(\hat{S} \neq S) \geq \frac{H(S \mid Y) - \log 2}{\log |\mathcal{S}|}\]

\textbf{Pinsker inequality}(Theorem 2 Proof): 

\[TV(P, Q) \leq \sqrt{\frac{KL(P \parallel Q)}{2}}\]

\textbf{dataat reason inequality}(Theorem 2 Proof): 

\[TV(P_{\text{pred}}, Q_{\text{pred}}) \leq TV(P, Q), \quad \text{where } X \to Y \text{ isMarkov chain}\]

\textbf{AUC TV world }(Theorem 2 Proof): 

\[|AUC_P - AUC_Q| \leq TV(P, Q) \text{ to any agree distribution } P, Q \text{ andfixed separate class device }\]

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{reference}\label{ux53c2ux8003ux6587ux732e}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 Fano, R. M. (1961). \emph{Transmission of Information: A Statistical
 Theory of Communications}. MIT Press.
\item
 Cover, T. M. \& Thomas, J. A. (2006). \emph{Elements of Information
 Theory} (2nd ed.). Wiley.
\item
 Pinsker, M. S. (1964). \emph{Information and Information Stability of
 Random Variables and Processes}. Holden-Day.
\item
 Tsybakov, A. B. (2009). \emph{Introduction to Nonparametric
 Estimation}. Springer.
\item
 SCX-Health MedMNIST experimentreport v2.
 \texttt{scx-health/results/experiment\_report\_v2.md}.
\item
 SCX AlN v3 Analysisreport.
 \texttt{experiments/mlip\_case/SCX\_AlN\_v3\_Analysisreport.md}.
\item
 SCX mathematicalroot source andProof build construct.
 \texttt{CodexKnowledge/agent\_outputs/05\_mathematicalroot source andProof.md}.
\item
 Theorem 1 (Compression Fidelity).
 \texttt{theory/propositions/04\_compression\_fidelity.md}.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{correctionDescription (2026-06-27)}: 1. \textbf{AUC/PR-AUC \(\eta\)
-dependence}: correction AUC and PR-AUC world middle leave leak \(\eta\)
because factor. becausethistwo indicate markinvolvingConditionsprobability (separate different From \(Z=1\) and \(Z=0\)
sampling), total variation world byamplify \(1/\eta + 1/(1-\eta)\) times. F1 world not receive impact, becauseis 
F1 isjoint distribution \(P(\hat{z}, Z)\) function. 2.
\textbf{StatusbalanceAssumptions}: is Corollary 2.2 (SCX
 retreat -ization tolossbaseline)addStatusbalanceAssumptions, bright confirm extremelynotbalanceStatuslower retreat -ization isapproximate, andbias difference from
\(\rho(\hat{s})\) estimateerror. 3. \textbf{k-means
Optimal ness claim brightremove}: shift except''k-means reach to Fano
lower bound''notverify claim bright, change is existenceatness theory describe addalgorithmapproximateerror term 
\(\varepsilon_{\text{k-means}}\). 

\end{document}
