% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
 12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
 \usepackage{unicode-math} % this also loads fontspec
 \defaultfontfeatures{Scale=MatchLowercase}
 \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
 % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
 \usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
 \usepackage{parskip}
 }{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
 \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
 hidelinks,
 pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Theorem 3: Unidentifiability of Noise vs.~Learnable
Difficulty}\label{theorem-3-unidentifiability-of-noise-vs.-learnable-difficulty}

\begin{quote}
\textbf{Core Claim}: onlyfrom observationsdata \((x, y, \{f_m(x)\})\)
Nonecannot distinguishLabelnoise from sample intrinsicDifficultydegree. SCX frameworkAssumptions (A1)-(A6)
is NOT arbitrarytechniqueConditions, andisbreakthiskindunidentifiabilityminimumSufficient Condition. 

\textbf{Revision History}: 2026-06-27 --- First version complete. 
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1 Theorem Statement (Theorem
Statement)}\label{ux5b9aux7406ux9648ux8ff0-theorem-statement}

\subsubsection{1.1 Notation and Setup}\label{ux7b26ux53f7ux4e0eux8bbeux5b9a}

 along using Theorem 1 Symbol/Conceptsystem, supplementwithlower note number: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
 >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Symbol/Concept
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Meaning
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\mathcal{X}\) & Input Space \\
\(\mathcal{Y}\) & Label Space, \(|\mathcal{Y}| = K\) \\
\(y^* \in \mathcal{Y}\) & trueLabel (latent atvariable, \textbf{unobservable}) \\
\(y \in \mathcal{Y}\) & Observed label \\
\(\{f_m\}_{m=1}^M\) & \(M\) expertmodel \\
\(\mathcal{D}_{\text{obs}}\) &
observationData Distribution, \((X, Y, \{f_m(X)\}) \sim \mathcal{D}_{\text{obs}}\) \\
\(\mathcal{S}\) & State Space \\
\(s: \mathcal{X} \to \mathcal{S}\) & Statusseparate match function \\
\(\eta(x) = \mathbb{P}(Y \neq Y^* \mid X = x)\) & input enter -dependencenoise rate \\
\end{longtable}

\textbf{Latent Variable Model}: Data generation is determined by the following latent process: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
 trueLabel \(y^*\) by \(x\) throughnot know Oracle \(f^*\)
 generate: \(y^* = f^*(x)\)
\item
 Observed label \(y\) cancanbecausenoisebias from \(y^*\)
\item
 expertprediction \(\{f_m(x)\}\) is \(x\) andtrainingdatafunction
\item
 Status \(s(x)\) is \(\mathcal{X}\) to \(\mathcal{S}\)
 map, describe input enter intrinsicCharacteristics
\end{enumerate}

\textbf{Observation Space}: The researcher can only access
\((x_i, y_i, f_1(x_i), \dots, f_M(x_i))\) joint distribution, Nonemethoddirectobservation
\(y^*\), \(s(x)\) ornoise indicate show variable. 

\subsubsection{1.2 Formal Statement}\label{ux5f62ux5f0fux5316ux9648ux8ff0}

\textbf{Theorem 3 (Noise-Difficulty Unidentifiability).} to any agree 
\(K \geq 2\) class separate class Problem, any agree \(M \geq 1\) expert, any agree has finite State Space
\(\mathcal{S}\), existenceattwo is differentData Generationprocess \(\mathcal{P}_1\) and
\(\mathcal{P}_2\) makes: 

\textbf{(i)} in \(\mathcal{P}_1\) lower, Observed label \(y\)
some some incorrectby\textbf{Labelnoise} (label noise)leads to: i.e. \(y \neq y^*\), and
\(y^*\) istrueNone error. 

\textbf{(ii)} in \(\mathcal{P}_2\) lower, has Observed label \(y\) equalstrueLabel
\(y^*\), butsome some sample has has \textbf{intrinsicDifficultydegree} (intrinsic
hardness): expertsystemness ground commit wrong and non-becauseis Labelincorrect, andisbecauseis samplethis self ambiguity. 

\textbf{(iii)} edge actual -Conditionsdecompose constant etc.: two processproduces complete fullmutualsameobservationjoint distribution: 

\[\mathcal{P}_1\bigl(x, y, \{f_m(x)\}_{m=1}^M\bigr) = \mathcal{P}_2\bigl(x, y, \{f_m(x)\}_{m=1}^M\bigr), \quad \forall (x, y, \{f_m\}) \in \mathcal{X} \times \mathcal{Y} \times \mathcal{Y}^M\]

\textbf{(iv)} becausethis, any agree algorithm
\(\mathcal{A}: (\mathcal{X} \times \mathcal{Y} \times \mathcal{Y}^M)^n \to \{0,1\}^n\) (will
\(n\)
 observationmapto''noise/non-noise''Assessment)Nonemethodsametimecorrectidentifiestwo world world in the noise: attofew a world world in the expectincorrect rate tofewis 
\(\eta\rho/2\) (where \(\eta\) is noise rate, \(\rho\)
is ambiguityStatusProportion). in particular, when \(\eta > 0\) and \(\rho > 0\)
timeshouldlower boundrigorousis currently, Prooftwo world world Nonemethodby complete beautiful distinguish. 

\begin{quote}
\textbf{Interpretation}: Theorem 3
indicate, atnot attach add amount outerAssumptionsConditionslower, ``this sampleLabeliswrong''and''this sampletoo difficult withexpertall do wrong''isobservationupperequivalentProposition. SCX
frameworkmustintroduceAssumptions (A1)-(A6) come breakthiskindequivalentness. 
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2 Proof (Proof)}\label{ux8bc1ux660e-proof}

\subsubsection{2.1 Construction Framework}\label{ux6784ux9020ux6846ux67b6}

adopt\textbf{Constructionness counter example method} (constructive counterexample). Constructiontwo world world 
\(\mathcal{P}_{\text{noise}}\) (noise world world)and
\(\mathcal{P}_{\text{hard}}\) (Difficulty world world), makes it s producesmutualsameobservationdistributionbuthas rootthisis differentbecause result explains. 

setup two separate class Problem \(K = 2\), Label Space \(\mathcal{Y} = \{0, 1\}\). setup
\(M \geq 1\) expert. Definitiontwo mutual repel Status \(s_1, s_2 \in \mathcal{S}\). 

\subsubsection{2.2 world world A: noise drive action 
(Noise-Driven)}\label{ux4e16ux754c-aux566aux58f0ux9a71ux52a8-noise-driven}

in \(\mathcal{P}_{\text{noise}}\) lower: 

\textbf{Status \(s_1\)} (``clean and easy''Status): -
sampleproportion: \(\mathbb{P}(X \in s_1) = \rho\), where \(\rho \in (0, 1)\) -
trueLabel constant determine: to has \(x \in s_1\), \(y^* = 0\) - noisemechanism: Labelwithprobability
\(\eta \in (0, 1/2)\) isflipped
\[\mathbb{P}(y \neq y^* \mid x \in s_1) = \eta\] -
expertaccuracy: to clear clean sample, has expertindependent ground with \(1 - \varepsilon_1\)
probabilitycorrectprediction \(y^*\)
\[\mathbb{P}(f_m(x) = y^* \mid x \in s_1, \text{clean}) = 1 - \varepsilon_1, \quad \varepsilon_1 \in (0, 1/2)\]

\textbf{Status \(s_2\)} (``Difficulty''Status): -
sampleproportion: \(\mathbb{P}(X \in s_2) = 1 - \rho\) - trueLabel constant determine: to has 
\(x \in s_2\), \(y^* = 0\) -
Nonenoise: \(\mathbb{P}(y \neq y^* \mid x \in s_2) = 0\) -
expertaccuracyrelativelylow (shouldStatustoexpert come say general time Difficulty): 
\[\mathbb{P}(f_m(x) = y^* \mid x \in s_2) = 1 - \varepsilon_2, \quad \varepsilon_1 < \varepsilon_2 \leq 1/2\]

\subsubsection{2.3 world world B: difficult degree drive action 
(Difficulty-Driven)}\label{ux4e16ux754c-bux96beux5ea6ux9a71ux52a8-difficulty-driven}

in \(\mathcal{P}_{\text{hard}}\) lower: 

\textbf{Status \(s_1\)} (`` contain difficult example ''Status): -
sampleproportion: \(\mathbb{P}(X \in s_1) = \rho\) (and world world A mutualsame) -
\textbf{ has Labeluniformis trueLabel} (Nonenoise): \(y = y^*\) to has type this -
buttrueLabelthis self has has ambiguity: 
\[\mathbb{P}(y^* = 0 \mid x \in s_1) = 1 - \eta, \quad \mathbb{P}(y^* = 1 \mid x \in s_1) = \eta\]
where \(\eta\) and world world A noise rate mutualsame -
expertConditionsaccuracy: todeterminetrueLabel, expertpredictionlinesis such aslower: - when \(y^* = 0\)
time: \(\mathbb{P}(f_m(x) = 0 \mid x \in s_1, y^* = 0) = 1 - \varepsilon_1\)
- when \(y^* = 1\)
time: \(\mathbb{P}(f_m(x) = 0 \mid x \in s_1, y^* = 1) = 1 - \varepsilon_1\)

i.e.experttotalis incline towardatpredictionCategory \(0\) (biastowardness), andtrueLabelNoneclose. becausethis: - when
\(y^* = 0\) time, expertaccuracyis \(1 - \varepsilon_1\) (table present good good) - when
\(y^* = 1\) time, expertaccuracyis \(\varepsilon_1\) (systemness ground commit wrong)

\textbf{Status \(s_2\)} (``Difficulty''Status): -
sampleproportion: \(\mathbb{P}(X \in s_2) = 1 - \rho\) (and world world A mutualsame) -
 has Labelis trueLabel: \(y = y^*\), and
\(\mathbb{P}(y^* = 0 \mid x \in s_2) = 1\) -
expertaccuracy: \(\mathbb{P}(f_m(x) = y^* \mid x \in s_2) = 1 - \varepsilon_2\) (and world world 
A mutualsame)

\subsubsection{2.4
World Equivalence Verification}\label{ux4e16ux754cux7b49ux4ef7ux6027ux9a8cux8bc1}

 present verifytwo world world observationjoint distribution constant etc.. observationdatais 
\((x, y, f_1(x), \dots, f_M(x))\), itsjoint distributionmaydecomposeis: 

\[\mathcal{P}(x, y, f_1, \dots, f_M) = \mathcal{P}(x) \cdot \mathcal{P}(y \mid x) \cdot \prod_{m=1}^M \mathcal{P}(f_m \mid x)\]

whereth two step make using Theorem 1 Assumptions (A1)-(A2) (expertattodetermine \(x\)
lowerConditionsindependent)------butNote agree thisat we s \textbf{does not require}shouldAssumptions general time holds, onlyisatConstructionmiddle make itsholdswithproducesmost simple clean counter example. 

\textbf{Step 1: edge actual distribution \(\mathcal{P}(x)\)}

two world world middle
\(\mathcal{P}(X \in s_1) = \rho\), \(\mathcal{P}(X \in s_2) = 1 - \rho\). constant etc.. 

\textbf{Step 2: ConditionsLabeldistribution \(\mathcal{P}(y \mid x)\)}

in \(\mathcal{P}_{\text{noise}}\) middle (world world A, Status \(s_1\)): 
\[\begin{aligned}
\mathcal{P}_{\text{noise}}(y = 0 \mid s_1) &= \mathbb{P}(\text{clean}) \cdot \mathbb{P}(y = y^* \mid \text{clean}) + \mathbb{P}(\text{noise}) \cdot \mathbb{P}(y = y^* \mid \text{noise}) \\
&= (1 - \eta) \cdot 1 + \eta \cdot 0 = 1 - \eta \\
\mathcal{P}_{\text{noise}}(y = 1 \mid s_1) &= (1 - \eta) \cdot 0 + \eta \cdot 1 = \eta
\end{aligned}\]

in \(\mathcal{P}_{\text{hard}}\) middle (world world B, Status \(s_1\)): 
\[\begin{aligned}
\mathcal{P}_{\text{hard}}(y = 0 \mid s_1) &= \mathbb{P}(y^* = 0 \mid s_1) = 1 - \eta \\
\mathcal{P}_{\text{hard}}(y = 1 \mid s_1) &= \mathbb{P}(y^* = 1 \mid s_1) = \eta
\end{aligned}\]

toStatus \(s_2\), two world world uniformNonenoise/ambiguity: 
\[\mathcal{P}(y = 0 \mid s_2) = 1, \quad \mathcal{P}(y = 1 \mid s_2) = 0\]

becausethisConditionsLabeldistributionattwo world world middle constant etc.. \(\checkmark\)

\textbf{Step 3: Conditionsexpertdistribution \(\mathcal{P}(f_m \mid x)\)}

in \(\mathcal{P}_{\text{noise}}\) middle (world world A, Status \(s_1\)): -
expertpredictiondistributionbytrueLabel \(y^*\) training/ learn obtain verdict determine, not-dependenceObserved label \(y\): 
\[\mathcal{P}_{\text{noise}}(f_m = 0 \mid s_1) = \mathbb{P}(f_m = y^* \mid s_1) = 1 - \varepsilon_1\]
\[\mathcal{P}_{\text{noise}}(f_m = 1 \mid s_1) = \varepsilon_1\]

in \(\mathcal{P}_{\text{hard}}\) middle (world world B, Status \(s_1\)): 
\[\begin{aligned}
\mathcal{P}_{\text{hard}}(f_m = 0 \mid s_1) &= \mathbb{P}(y^* = 0 \mid s_1) \cdot \mathbb{P}(f_m = 0 \mid s_1, y^* = 0) \\
&\quad + \mathbb{P}(y^* = 1 \mid s_1) \cdot \mathbb{P}(f_m = 0 \mid s_1, y^* = 1) \\
&= (1 - \eta) \cdot (1 - \varepsilon_1) + \eta \cdot (1 - \varepsilon_1) \\
&= 1 - \varepsilon_1
\end{aligned}\]

\[\begin{aligned}
\mathcal{P}_{\text{hard}}(f_m = 1 \mid s_1) &= (1 - \eta) \cdot \varepsilon_1 + \eta \cdot \varepsilon_1 = \varepsilon_1
\end{aligned}\]

toStatus \(s_2\)
two world world mutualsame: \(\mathcal{P}(f_m = 0 \mid s_2) = 1 - \varepsilon_2\), \(\mathcal{P}(f_m = 1 \mid s_2) = \varepsilon_2\). 

becausethisConditionsexpertdistributionattwo world world middle constant etc.. \(\checkmark\)

\textbf{Step 4: Conditionsindependentnessverify}

attwo world world middle, we s Constructionexpertattodetermine \(x\) (orequivalent ground, todetermine
\(s(x)\))lowerConditionsindependent. thisisbecauseis: - in world world A
middle, experttrainingdataindependent (byAssumptions A1 directto out) - in world world B
middle, expertpredictionerrorattodetermineStatusandtrueLabellowerindependent

becausethis \(\prod_{m=1}^M \mathcal{P}(f_m \mid x)\)
decomposeholds, andeachbecause factor attwo world world betweenmatches. \(\checkmark\)

\textbf{Step 5: joint distribution constant etc. }

 comprehensive combined Step 1-4 obtain: 

\[\mathcal{P}_{\text{noise}}\bigl(x, y, \{f_m\}\bigr) = \mathcal{P}_{\text{hard}}\bigl(x, y, \{f_m\}\bigr), \quad \forall (x, y, \{f_m\})\]

two world world observationjoint distribution complete fullconsistent. \(\square\) (QEDth (iii) part)

\subsubsection{2.5
Algorithm Unidentifiability}\label{ux7b97ux6cd5ux4e0dux53efux533aux5206ux6027}

byth (iii) part, two world world producesobservationupperindistinguishablejoint distribution. setup \(\mathcal{A}\)
is any agree algorithm, make using \(n\) observationsample
\(D_n = \{(x_i, y_i, \{f_m(x_i)\})\}_{i=1}^n\) input out each samplenoiseAssessment
\(z_i \in \{0, 1\}\). 

DefinitiontwokindAssessmentMeaning: - in \(\mathcal{P}_{\text{noise}}\) lower: \(z_i = 1\)
representation''shouldsampleLabelbyflipped (noise)'' - in \(\mathcal{P}_{\text{hard}}\)
lower: \(z_i = 1\) representation''shouldsampleLabelcorrectbutexpertsystem commit wrong (Difficulty)''

byat \(\mathcal{A}\) only-dependence \(D_n\), and \(D_n\)
attwo world world in the distribution complete fullmutualsame: 

\[\mathcal{L}_{\mathcal{P}_{\text{noise}}}(D_n) = \mathcal{L}_{\mathcal{P}_{\text{hard}}}(D_n)\]

where \(\mathcal{L}\) representationdistribution law. becausethis \(\mathcal{A}\)
 input out distributionattwo world world middlemutualsame: 

\[\mathcal{L}_{\mathcal{P}_{\text{noise}}}(\mathcal{A}(D_n)) = \mathcal{L}_{\mathcal{P}_{\text{hard}}}(\mathcal{A}(D_n))\]

butattwo world world middle, \textbf{correct''noise'' set is different}: -
\(\mathcal{P}_{\text{noise}}\) middletruecurrentlynoisesample set 
\(\mathcal{Z}_{\text{noise}}^* = \{i: y_i \neq y_i^*\} = \{i: x_i \in s_1, y_i = 1\}\)
- \(\mathcal{P}_{\text{hard}}\) middletruecurrentlynoisesample set 
\(\mathcal{Z}_{\text{hard}}^* = \varnothing\) (Nonenoise)

byat \(\mathcal{A}\) Nonecannot distinguishtwo world world, its input out Marked set \(\hat{\mathcal{Z}}\)
mustsametimeis applicable totwo world world. set \(a\) is algorithmatambiguitysubset
\(\{i: x_i \in s_1, y_i = 1\}\)
upperexpectMarked rate (should value attwo world world middlemutualsame, becauseis observationdistributionmutualsame). at world world A
middle, this some sampleuniformis noise (correctMarkedis \(1\)), incorrect rate is \((1-a)\); at world world B
middle, this some sampleuniformnon-noise (correctMarkedis \(0\)), incorrect rate is \(a\). becausethis: 

\[\max\bigl(\text{Error}_{\mathcal{P}_{\text{noise}}}(\mathcal{A}), \text{Error}_{\mathcal{P}_{\text{hard}}}(\mathcal{A})\bigr) \geq \frac{\text{Error}_{\text{noise}} + \text{Error}_{\text{hard}}}{2} \geq \frac{\eta\rho}{2}\]

where \(\eta\rho\) isambiguitysubsetatall type this in the proportion. shouldlower boundin \(a = 1/2\)
time reach to (algorithmatambiguitysampleupperis identical torandom guess measure). for classical type noise rate \(\eta = 0.1\)
and \(s_1\) proportion \(\rho = 0.5\), lower boundis 
\(0.025\)------althoughsmallbutrigorousis currently, provestwo world world not complete fullmaydistinguishness. determinenessunidentifiabilityConclusion (Nonemethod complete beautiful distinguish noise fromDifficulty)still holds. \(\square\)
(QEDth (iv) part)

\subsubsection{2.6 Extension: a like K
 class separate class }\label{ux6269ux5c55ux4e00ux822c-k-ux7c7bux5206ux7c7b}

upper describe Constructionmaydirect push broad to \(K > 2\) case. onlymustwillnoisemodel change is uniform uniform flipped (such as
Theorem 1 A4), and atDifficulty world world middlewill \(y^*\)
distributionsetup place is andnoise world world Observed labeldistributionconsistent. Complete push broad Proofseeappendix A. 

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3 Corollary
(Corollaries): which some Assumptionscanbreakunidentifiability? }\label{ux63a8ux8bba-corollariesux54eaux4e9bux5047ux8bbeux80fdux6253ux7834ux4e0dux53efux8bc6ux522bux6027}

Theorem 3
indicate, notadd amount outerAssumptionstimenoise andDifficultyisindistinguishable. withlowerCorollary list out \textbf{Sufficient Condition}------each Conditionssingle independent out present i.e.maybreakunidentifiability. 

\subsubsection{3.1 Corollary 1: anchor point Label (Anchor
Points)}\label{ux63a8ux8bba-1ux951aux70b9ux6807ux7b7e-anchor-points}

\textbf{Corollary 1 (Anchor Labels Break Unidentifiability).}
If tosome already know subset
\(\mathcal{X}_{\text{anchor}} \subset \mathcal{X}\), research investigate onecan obtain take\textbf{trueLabel}
\(y^*\) (such as person work complex kernel), thennoise andDifficultyin \(\mathcal{X}_{\text{anchor}}\)
uppermaydistinguish. advance a step ground, If \(\mathcal{X}_{\text{anchor}}\)
 cover cover has Status, thenglobalmaydistinguish. 

\textbf{Proof Idea}: at anchor point \(x \in \mathcal{X}_{\text{anchor}}\)
upper, comparisonObserved label \(y\) andtrueLabel \(y^*\): - If 
\(y \neq y^*\), thenshouldsample must however contain noise - If \(y = y^*\) butexpert
\(f_m(x) \neq y\), thenshouldsample must howeveris''Difficulty''

will anchor point upperexpertincorrect rate outer push tonon- anchor point sample, maybreakglobalunidentifiability. \(\square\)

\textbf{and SCX connection}: SCX ''cold enable action collaborate dispute '' (cold start
protocol)-dependenceata small amount of person work reviewsample action is anchor point come initialization -ization Status level noise rate estimate. 

\subsubsection{3.2 Corollary 2: already know experttraining domain (Known Training Domains ---
A1)}\label{ux63a8ux8bba-2ux5df2ux77e5ux4e13ux5bb6ux8badux7ec3ux57df-known-training-domains-a1}

\textbf{Corollary 2 (Known Disjoint Training Data Break
Unidentifiability).} If \(M\) expertin \(M\)
 \textbf{disjoint}datasubsetupperindependently trained (Assumptions
A1), and research investigate one know path this some training set constitutes, thennoise andDifficultymaydistinguish. 

\textbf{Proof Idea}: disjointtraining set agree taste -ing expertincorrect model formula is\textbf{independent } (todetermine
\(x\)). atnoisesampleupper, has expertincorrect rate sametimerises (becauseis has expertpredictionallfortrue
\(y^*\), andObserved label
\(y \neq y^*\)). atDifficultysampleupper, expertincorrectmaymutualmutualindependent (buteach expertincorrect rate this self morehigh). 

keyquantityDifference: set
\(\bar{e}(x) = \frac{1}{M}\sum_m \mathbf{1}\{f_m(x) \neq y\}\)
is on averageexpertincorrect rate. atnoisesampleupperConditionsdistributionis: 

\[\mathbb{E}[\bar{e}(x) \mid \text{noise}] = 1 - \frac{1 - \mathbb{E}[\bar{e}(x) \mid \text{clean}]}{K-1}\]

 (Theorem 1, Lemma
1). this relationshipatonlyhas hardness world world middleNonemethodsametimeholds. becausethis, check verify \(\bar{e}(x)\)
distributioniswhethersatisfies Lemma 1 relationship formula maywithdistinguishtwo world world. \(\square\)

\textbf{importantDescription}: Corollary 2
-dependenceat research investigate one\textbf{ confirm cut know path }training set disjoint. If training set relationshipnot know, thenNonemethodapplicationthis check verify. 

\subsubsection{3.3 Corollary 3: noiseis independent of input enter (Noise Independent of x ---
A4)}\label{ux63a8ux8bba-3ux566aux58f0ux72ecux7acbux4e8eux8f93ux5165-noise-independent-of-x-a4}

\textbf{Corollary 3 (Input-Independent Noise Rate Breaks
Unidentifiability).} If Labelnoise rate 
\(\eta(x) = \mathbb{P}(Y \neq Y^* \mid X = x)\) is often number (not-dependenceat
\(x\)), andDifficultyStatusexpertincorrect rate atStatusbetweenhas Difference, thennoise andDifficultymaydistinguish. 

\textbf{Proof Idea}: If \(\eta(x) = \eta\) is global often number, thenObserved label \(y\)
andexpertincorrect \(\{f_m(x) \neq y\}\)
'sbetweenmutualcloseness model formula atnoiseandDifficulty world world middlehas thisDifference. 

atnoise world world middle, \(\bar{e}(x)\)
expect value at clear clean sampleupperlow, atnoisesampleupperhigh, butnoise matter file and \(x\) independent, becausethis
\(\bar{e}(x)\) Conditionsdistributionnot follow \(x\) change -ization (edge actual upper), onlywhenwith \(y\)
is Conditionstimeonly thenexplicit present Difference. 

atDifficulty world world middle, \(y\) distributionmay follow \(x\)
 change -ization, andexpertincorrect rate also follow Status change -ization. through check verify \(\bar{e}(x)\) and \(y\)
iswhetheratentire Input Spaceupper has has constant determinejoint distribution model formula (i.e.noise world world ''uniform uniform noise''Characteristics), maywithdistinguishtwoone. \(\square\)

\subsubsection{3.4 Corollary 4: Statussameness (State Homogeneity ---
A5)}\label{ux63a8ux8bba-4ux72b6ux6001ux540cux8d28ux6027-state-homogeneity-a5}

\textbf{Corollary 4 (State Homogeneity Breaks Unidentifiability).}
If Status plan separate \(\mathcal{S}\) makesateach Status \(s\)
inner, clear clean sampleexpectexpertincorrect rate \(\mu_s\) is often number upper boundcontrols (Assumptions
A5), andStatus number \(|\mathcal{S}| \geq 2\) and andis differentStatusbetween \(\mu_s\)
has Difference, thennoise andDifficultymaydistinguish. 

\textbf{Proof Idea}: atTheorem 3 counter example Constructionmiddle, two world world Status \(s_1\) and
\(s_2\) inStatussamenessAssumptionslowerwillproduces contradiction contradiction \(\mu_s\)
estimate. specifically, atwithlowertwokindexplainsmiddletomany a kindcansatisfiesStatussamenessAssumptions: 

\begin{itemize}
\item
 \textbf{noiseexplains} (world world A): Status \(s_1\) clear clean incorrect rate is 
 \(\varepsilon_1\), butobservationto
 \(\bar{e}(s_1) = (1-\eta)\varepsilon_1 + \eta(1-\varepsilon_1)\)------thisisnoiseand clear clean mixedcombined. such as result 
 \(\eta\) isnot know, NonemethodatthisStatuslowerdistinguishnoiseandDifficultyseparate quantity. 
\item
 \textbf{Difficultyexplains} (world world B): Status \(s_1\) middle \(\eta\)
 ProportionsampletrueLabelis Category 1, expertatshould factor group overall upperaccuracyonlyis 
 \(\varepsilon_1\). thisis equivalent to recognize is Status \(s_1\) containstwo factor Status (Category 0 
 \(s_{1a}\) andCategory 1 \(s_{1b}\)), and it s expertincorrect rate is different. 
\end{itemize}

StatussamenessAssumptions (A5)requireeach Statusinner clear clean incorrect rate uniform uniform. such as result \(s_1\)
innerpartexpertaccuracyatis differentCategorysamplebetweenexplicit notable is different (world world B middleis 
\(1-\varepsilon_1\) vs \(\varepsilon_1\)), then (A5)
atthisStatusupperby violate counter. becausethis, such as result (A5) holds, world world B Constructiondoes not hold. \(\square\)

\subsubsection{3.5 Corollary 5: balanceerrordistribution (Balanced Error Distribution ---
A6)}\label{ux63a8ux8bba-5ux5e73ux8861ux8befux5deeux5206ux5e03-balanced-error-distribution-a6}

\textbf{Corollary 5 (Balanced Error Distribution Breaks
Unidentifiability).} If expertincorrectat has incorrectCategoryupperuniform uniform distribution (Assumptions
A6, \(C_{\text{bal}} = 1\)), thennoise andDifficultymaydistinguish. 

\textbf{Proof Idea}: atTheorem 3 world world B Constructionmiddle, expertpredictionbiastowardCategory 0. when
\(y^* = 1\) time, expertwith \(1 - \varepsilon_1\) highprobabilityprediction
0, this agree taste -ing atStatus \(s_1\) middle, \(f_m\) incorrect several -ish all set middleatCategory 1 \to
predictionis 0 directionupper. this violate anti-A6 balancerequire------expertincorrectextremelydegreenotbalance. 

becausethis, such as result already know expertincorrectisbalance (\(C_{\text{bal}} \approx 1\)), then world world 
B Constructionbyexcludes. \(\square\)

\subsubsection{3.6 Corollary 6: manyStatus cross stack (Multi-State
Overlap)}\label{ux63a8ux8bba-6ux591aux72b6ux6001ux4ea4ux53e0-multi-state-overlap}

\textbf{Corollary 6 (Differential Expert Accuracy Across States Breaks
Unidentifiability).} If existenceattofewtwo Status \(s_a, s_b \in \mathcal{S}\)
makesexpertin \(s_a\) upperaccuracy\textbf{above}in \(s_b\)
upperaccuracy, and and research investigate onecanmay reliable ground willsampleseparate match toStatus, thennoise andDifficultymaydistinguish. 

\textbf{Proof Idea}: Theorem 3 world world A inConstructionupper-dependenceatStatus \(s_1\)
sametimecontains clear clean andnoisesamplemixedcombined, butitsObserved labeldistributionandStatus \(s_2\)
is different. throughcomparisonis differentStatus'sbetweenconsistency scoredistribution, maywith check verify ''mixedcombined''Assumptionsiswhetherholds. 

specifically, setup \(\mu_s = \mathbb{E}[C \mid \text{clean}, X \in s]\). at
Theorem 1 Lemma 1 middle, noisesampleconsistency scoreexpectis 
\(1 - \mu_s/(K-1)\). If insome Statusmiddleobservationtoconsistency scoremean
\(\bar{C}_s > 1 - \hat{\mu}_s/(K-1)\) (where \(\hat{\mu}_s\)
isFrom lowseparate number estimate clear clean baseline), thenshouldStatusmiddlemaydoes not existnoise (i.e. world world B
 class type). \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4 and SCX Assumptionssystem (A1-A6)
connection}\label{ux4e0e-scx-ux5047ux8bbeux4f53ux7cfb-a1-a6-ux7684ux8054ux7cfb}

Theorem 3 and itsCorollaryreveal SCX Assumptions (A1)-(A6)
rootthis action use: it s is NOT any agree attach addtechniqueConditions, andis\textbf{breakunidentifiabilityminimumSufficient Condition}. 

\subsubsection{4.1
Assumptionsseparate class: which some Assumptionsatbreak what what }\label{ux5047ux8bbeux5206ux7c7bux54eaux4e9bux5047ux8bbeux5728ux6253ux7834ux4ec0ux4e48}

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1875}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1875}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3438}}
 >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2812}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Assumptions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Content
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
breakambiguity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
toshouldCorollary
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{A1} & disjointtraining set &
expertindependentness: ensurenoisesampleupper has expertsametime out wrong model formula anddry clean sampleis different & Corollary
2 \\
\textbf{A2} & clear clean Conditionsindependent & todetermine \(x\) lowerexpertincorrectindependentnessdecompose & Corollary 2
techniqueFoundation \\
\textbf{A3} & has world loss & techniquenessAssumptions (Hoeffding
inequality must), notdirectbreakunidentifiability & --- \\
\textbf{A4} & uniform uniform independentnoise & noise rate input enter Noneclosenessanduniform uniform distribution: excludes \(x\)
-dependencenoise rate leads to pseudo ''Difficulty'' model formula & Corollary 3 \\
\textbf{A5} & Statussameness &
Statusinner clear clean incorrect rate uniform uniform: excludes''intrinsicDifficultysample aggregate set atsame a Status''mayness &
Corollary 4 \\
\textbf{A6} & balanceerrordistribution &
expertincorrectat has incorrectCategoryupperuniform uniform: excludesexpertbiastowardnessleads to pseudo ''Difficulty'' model formula &
Corollary 5 \\
\end{longtable}

\subsubsection{4.2
minimumbreaking ness set combined}\label{ux6700ux5c0fux7834ux574fux6027ux96c6ux5408}

A1-A6 middle which some isbreakunidentifiabilitynecessaryAssumptions? Analysissuch aslower: 

\textbf{Theorem 3 anti- example resistsness}: Theorem 3 ConstructionusewithlowerConditions: 1. \(K = 2\)
separate class (most simple singlecase) 2. expertConditionsindependent (A1/2) 3. noise rate -dependenceatStatus (\(\eta\)
vs 0)------thisand A4 ''uniform uniform independentnoise''has finemicro- difference different 4.
Statusbetweenexpertincorrect rate is different (\(\varepsilon_1 \neq \varepsilon_2\))

mustexcludesTheorem 3 Construction, \textbf{tofewrequirewithlowerAssumptions's a }: - A1
\textbf{disjointtraining domain know recognize } (Corollary
2): If we s know path experttraining domain confirm actual disjoint, world world B Constructionmay violate expertindependentness -
A4 \textbf{uniform uniform noise} (Corollary 3): If noise rate global constant determine, world world A Constructionrequire
\(\eta\) in \(s_1\) and \(s_2\) middlemutualsame, force make heavynewdesign - A5
\textbf{Statussameness} (Corollary 4): If Statusinnerincorrect rate uniform uniform, world world B \(s_1\)
inner \(y^*\) -dependenceincorrect rate violate counter thisAssumptions - A6 \textbf{balanceerrordistribution} (Corollary
5): If expertincorrectuniform uniform distribution, world world B expertbiastowardnessbyexcludes

\textbf{becausethis, SCX breakunidentifiability mustminimumAssumptions set combinedis}: 
\[A_{\min} = \{A1, A4, A5\} \quad \text{or} \quad \{A1, A4, A6\} \quad \text{or} \quad \{A5, A6\} \text{ addupperStatus number } |\mathcal{S}| \geq 2\]

i.e.tofewrequire\textbf{training domain independentness} + \textbf{noiseuniform uniform ness} +
\textbf{ a kindincorrectuniform uniform nessConditions} (StatusinnerorCategorybetween), orone\textbf{Statussameness} +
\textbf{incorrectbalanceness} + \textbf{manyStatusstructure}. 

\subsubsection{4.3
Assumptions violate counter time retreat -ization }\label{ux5047ux8bbeux8fddux53cdux65f6ux7684ux9000ux5316}

 reason solve each AssumptionsRoleback, maywithestablish violate counter eachAssumptionstime retreat -ization model formula: 

\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3929}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2857}}
 >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3214}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
 violate counter Assumptions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 retreat -ization is 
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
 actual actual table present 
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
A1 (training domain heavy stack) & expertmutualclosenessleads to Theorem 1 exponential world eliminate lose & SCX noise detection
F1 falls, butstillmayis superior torandom \\
A4 (noise-dependence \(x\)) & noise rate andCharacteristics difficult degreecoupled & SCX
noiseestimatehas bias, butrelativesortmaypreserves \\
A5 (Statusdifferent) & Statusinner clear clean /noisemixedcombined & Theorem 1
not adapt use, requiremorefineStatus plan separate \\
A6 (incorrectnotuniform balance) & noise detectionbiastoward often seeincorrectCategory & false positive set middleat special determineCategory, F1
fallsbutmaystillmayusing \\
\textbf{all violate counter } & complete full retreat -ization & \textbf{Theorem 3
unidentifiability life effect, any what algorithmNonemethoddistinguish noise fromDifficulty} \\
\end{longtable}

\subsubsection{4.4 SCX
frameworksuch as what implicit formula usethis some Conditions}\label{scx-ux6846ux67b6ux5982ux4f55ux9690ux5f0fux4f7fux7528ux8fd9ux4e9bux6761ux4ef6}

at actual actual SCX applicationmiddle, A1-A6
 and non-totalisbright confirm check verify. frameworkthroughwithlowermechanism\textbf{implicit formula }usebreakunidentifiabilitystructure: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{cold enable action collaborate dispute }: require person work reviewa small amount of anchor point samplegenerateinitializationStatusLabel (is equivalent toCorollary
 1 anchor point). 
\item
 \textbf{Statusdiscover}: throughto \(\phi(X)\) cluster come plan separateStatus, implicit contain ground Assumptions
 A5------cluster test figure ateach cluster innerminimum -ization change different, make Statussamenessapproximateholds. 
\item
 \textbf{multi-expert aggregate combined}: usedisjointtrainingexpert (A1), thisrequireexperimentdesignorder segment bright confirm separate match trainingdata. 
\item
 \textbf{ system compute check verify }: check \(\hat{C}(s)\) via verify distributioniswhethersatisfies Lemma 1
 Mean Separationrelationship, If does not satisfythenMarkedis ''weakCharacteristics lose effect '' (Theorem 2). 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5
and present has text contribute connection}\label{ux4e0eux73b0ux6709ux6587ux732eux7684ux8054ux7cfb}

\subsubsection{5.1 measureerrormodel (Measurement Error
Models)}\label{ux6d4bux91cfux8befux5deeux6a21ux578b-measurement-error-models}

 system compute study measureerror text contribute middlehas a classicalConclusion: atclassicalmeasureerrormodel (classical
measurement error model)lower, If nohas \textbf{verifydata} (validation
data)or\textbf{toolvariable} (instrumental
variables), measureerrordistributionandtrue value distributionindistinguishable (Fuller, 1987; Carroll
et al., 2006). 

form -ization ground, tomodel \(Y = Y^* + U\), where \(U\) is measureerror: If onlyobservationto
\(Y\) and \(Y^*\) and \(U\) uniformnot know, thenhas None exhaustive manyto \((Y^*, U)\)
distributionsatisfiesobservationto \(Y\)
distribution. thisis\textbf{measureerrorunidentifiability}Standard result. 

Theorem 3
willthisresultFrom continuousmeasureerrorExtensiontoseparate class Labelnoise andexpertDifficultyto called Problemmiddle. key push broad is: notonlyhas 
\(Y\) and \(Y^*\) relationship, stillintroduceexpertprediction \(\{f_m(X)\}\)
 action is amount outerobservation. we s Proofi.e. make has this some amount outerinformation, unidentifiabilitystill holds------unlessto
\(\{f_m\}\) structure apply addAssumptions. 

\subsubsection{5.2
Labelnoisetheoryunidentifiability}\label{ux6807ux7b7eux566aux58f0ux7406ux8bbaux7684ux4e0dux53efux8bc6ux522bux6027}

Labelnoise text contribute in the classicalresultis\textbf{noise turn shift matrixunidentifiability} (Menon et al.,
2015; Patrini et al., 2017): 

to \(K\) class separate class, noise turn shift matrix \(T \in [0,1]^{K \times K}\) satisfies
\(T_{ij} = \mathbb{P}(\tilde{Y} = j \mid Y^* = i)\). atonlyobservationto
\((\tilde{Y}, X)\) Conditionslower, \(T\) andtrueConditionsdistribution
\(\mathbb{P}(Y^* \mid X)\) connect combinedisunidentifiable------unlessaddupperSufficient Condition (such as
\(T\) toangle occupy superior ness, existenceat anchor point , orexistenceatmany noise rate already know markNoteone). 

Theorem 3
isthisresulttoward\textbf{multi-expertscenario} push broad. atmulti-expertscenariomiddle, we s notonlyobservationto
\(\tilde{Y}\), stillobservationto
\(\{f_m(X)\}\), butunidentifiabilitystill holds. key hole observe is: expertthis self incorrect model formula mayfrom
\(Y^*\) Difficulty region domain, alsomayfrom \(Y^*\)
byflippedback pseudo model formula ------thistwooneatobservationupperequivalent. 

\subsubsection{5.3 Dawid-Skene
unidentifiability}\label{dawid-skene-ux7684ux4e0dux53efux8bc6ux522bux6027}

Dawid-Skene model (1979)middle, each markNoteone \(m\) has a confusion matrix
\(\pi_m\), trueLabel \(Y^*\)
is latent atvariable. Standard resultis: atto called ness convention bind (such asfixed a markNoteoneconfusion matrix)or anchor point markNoteone (somemarkNoteone complete fullmay reliable)Conditionslower, modelonly thenmayidentifies. 

Theorem 3 and Dawid-Skene unidentifiabilityrelationship: - Dawid-Skene
unidentifiabilitycloseNote\textbf{is differentmarkNoteoneconfusion matrix'sbetweenpermuteto called ness} (LabelflippedProblem)
- Theorem 3 closeNote\textbf{noise andDifficulty'sbetween language meaning indistinguishability}------thisis SCX
 special has a orderProblem - twoonecurrently cross: i.e. make solve verdict Dawid-Skene
LabelpermuteProblem (through anchor point markNoteone), Theorem 3 unidentifiabilitystillhoweverexistenceat

 exchange word 's, Dawid-Skene solve verdict is'' which markNoteonemay reliable ''Problem, and Theorem 3
revealis''sampleis what what notmay study ''Problem------backoneisfrontonefront raise (atexplainsexpertincorrect model formula 'sfront, must know incorrect come source). 

\subsubsection{5.4
because result infer view angle}\label{ux56e0ux679cux63a8ux65adux7684ux89c6ux89d2}

From because result inferangledegree, Theorem 3
mayby reason solve is a kind\textbf{because result structureunidentifiability} (causal structure
unidentifiability; Pearl, 2009): 

considertwo because result figure: 

\textbf{ figure A (noisebecause result figure)}: 

\begin{verbatim}
S \to Y* \to Y ← Noise
↓
X \to {f_m}
\end{verbatim}

where arrow head \(Y^* \to Y\) representationtrueLabel verdict determineObserved label, but \(Y\)
alsomaybyouterpartnoisedirectimpact. 

\textbf{ figure B (Difficultybecause result figure)}: 

\begin{verbatim}
S \to Y* (= Y)
↓
X \to {f_m}
\end{verbatim}

whereDifficultyStatus \(S\) sametimeimpacttrueLabeldistributionandexpertaccuracy. 

two because result figure generatemutualsameobservationdistribution
\((X, Y, \{f_m\})\), but contain contain is differentbecause result explains. onlyfrom observationsdataNonecannot distinguishthistwo figure ------thisis classical type \textbf{ horse -er may -ist equivalent class } (Markov
equivalence class)Problem. 

\subsubsection{5.5
mixedcombinedmodelandhas finite mixedcombinedmayidentifiesness}\label{ux6df7ux5408ux6a21ux578bux4e0eux6709ux9650ux6df7ux5408ux7684ux53efux8bc6ux522bux6027}

athas finite mixedcombinedmodel (finite mixture
models) text contribute middle, mayidentifiesness through often requireseparate quantity distribution has has special determineparameterform (such asexponential family), orrequireseparate quantity number has upper bound (Teicher,
1963; McLachlan \& Peel, 2000). 

Theorem 3 counter example mayby view is a \textbf{twoseparate quantity mixedcombinedmodel}unidentifiabilityexample: -
 world world A: separate quantity 1 (clear clean, occupy \(1-\eta\))andseparate quantity 2 (noise, occupy \(\eta\)) -
 world world B: separate quantity 1 (\(y^*=0\), occupy \(1-\eta\))andseparate quantity 2 (\(y^*=1\), occupy 
\(\eta\))

two mixedcombinedmodelproducesmutualsameobservationjoint distribution, butseparate quantity has is different language meaning explains. thisindicateatLabelnoisescenariomiddle, i.e. make separate quantity number already know (\(K=2\)), Standard mixedcombinedmodelmayidentifiesnessConditions (such asparameter -ization formalready know)stillhoweveris insufficient------stillrequiretoseparate quantity 'sbetween''expertprediction model formula '' apply add amount outerstructure convention bind. 

\subsubsection{5.6
minimum value andmaximum value: Assumptionsnecessarynessspectrum system }\label{ux6700ux5c0fux503cux4e0eux6700ux5927ux503cux5047ux8bbeux7684ux5fc5ux8981ux6027ux8c31ux7cfb}

 comprehensive combinedwithupper discuss theory, may obtain withlower\textbf{Assumptionsnecessarynessspectrum system }: 

\begin{verbatim}
mostweakAssumptions (unidentifiable)
├── NoneAssumptions \to Theorem 3 holds: complete fullunidentifiable
├── + anchor point Label (Corollary 1) \to in anchor point uppermayidentifies
├── + disjointtraining domain (A1, Corollary 2) \to throughincorrectindependentness check verify mayidentifies
├── + uniform uniform noise (A4, Corollary 3) \to throughincorrect rate input enter not change nessmayidentifies
├── + Statussameness (A5, Corollary 4) \to throughStatusinnerincorrect rate uniform uniform nessmayidentifies
├── + incorrectbalanceness (A6, Corollary 5) \to throughexpertincorrectCategorydistributionmayidentifies
└── A1 + A4 + A5 (or A1 + A4 + A6) \to complete fullmayidentifies (SCX Standard Assumptions)
moststrongAssumptions (complete fullmayidentifies)
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6 actual actual meaning}\label{ux5b9eux9645ux610fux4e49}

\subsubsection{6.1 to SCX
frameworkcurrentlywhenness -ization }\label{ux5bf9-scx-ux6846ux67b6ux7684ux6b63ux5f53ux6027ux5316}

Theorem 3 is SCX Assumptionssystemproviderootthisnesscurrentlywhenness: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{A1-A6
 is NOT over degreeAssumptions}: it s isbreaknotmayavoidunidentifiability mustminimum attach addstructure. any what claimscanatNoneAssumptionsConditionslowerdistinguish noise fromDifficultyMethodology, must what implicit formula containsequivalentAssumptions, must what existenceattheorydefect. 
\item
 \textbf{Theorems 1-3 commonsameconstitutes logic edit close environment }: 

 \begin{itemize}
 \tightlist
 \item
 Theorem 1: in A1-A6 lower, noisemaydetection (Positive Statement)
 \item
 Theorem 2: whenCharacteristicsweakensdetectioncanforce (Boundary Statement)
 \item
 Theorem 3: NoneAssumptionstimenoise andDifficultyindistinguishable (Necessity Statement)
 \end{itemize}
\item
 \textbf{cold enable action collaborate dispute theoryFoundation}: SCX
 must seek initializationorder segment person work reviewa small amount ofsample come estimateStatus level noise rate. Theorem 3
 Proofthisis\textbf{theory must must}andnon- work process proper collaborate ------nothas this anchor point, systemNonemethodbreakthroughunidentifiability. 
\end{enumerate}

\subsubsection{6.2
to compete dispute Methodology batch judge nessAnalysis}\label{ux5bf9ux7adeux4e89ux65b9ux6cd5ux7684ux6279ux5224ux6027ux5206ux6790}

Theorem 3 provideevaluateits he noise detectionMethodologyframework: 

\begin{itemize}
\item
 \textbf{loss-based Methodology} (such as MentorNet, Decoupling,
 Co-teaching): implicit formula Assumptions''highloss = noise'', i.e.in Theorem 3
 middleselect'' world world A''explains. whendata actual actual follow follow '' world world 
 B'' model formula time, this some MethodologywillwillDifficultysample error markis noise. 
\item
 \textbf{confidencelearningMethodology} (Confidence Learning; Northcutt et al.,
 2021): use compute number matrixestimatenoise turn shift matrix, implicit formula Assumptionsexpert (separate class device) compute number andnoise rate 'sbetweenexistenceatmaydecomposerelationship. at
 `` world world B'' middle, shoulddecompose lose effect. 
\item
 \textbf{many view figure Methodology} (Multi-view
 methods): -dependencemany independent view figure consistentness, is equivalent to Theorem 1 A1-A2
 Assumptions. Theorem 3 Proof, nothas this some Assumptionstimemany view figure consistentnessalsoNonemethoddistinguish noise fromDifficulty. 
\end{itemize}

\subsubsection{6.3
experimentdesignRecommendation}\label{ux5b9eux9a8cux8bbeux8ba1ux5efaux8bae}

based on Theorem 3, raise out withlowerexperimentdesign original then: 

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 \textbf{bright confirm Assumptions}: atpaperorexperimentreportmiddlebright confirm list out do Assumptions (such as'' we s Assumptionsnoise and
 x independent''). Theorem 3 agree taste -ing not list out AssumptionsthenisnotsatisfiesidentifiesConditions. 
\item
 \textbf{sensitivenessAnalysis}: check verify ConclusiontoAssumptions violate counter sensitiveness. example such as, If Statussameness slightly weak (A5
 notrigorousholds), noise detection F1 fallsmanyfew? 
\item
 \textbf{Assumptions check verify }: design system compute check verify come verifykeyAssumptions. example such as, using \(\chi^2\)
 check verify expertincorrectiswhetheruniform uniform distribution (check verify A6), orusing KS
 check verify Statusbetweenconsistency scoredistributionDifference (check verify A5). 
\item
 \textbf{manyMethodologycross-validation}: sametime motion lines SCX and loss-based
 Methodology, willtwooneMarkedis inconsistentsamplesubmit person work review------thiscurrentlyis SCX
 cold enable action collaborate dispute actual actual operation action direction formula. 
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{appendix A: K class push broad }\label{ux9644ux5f55-ak-ux7c7bux63a8ux5e7f}

willTheorem 3 push broad to \(K > 2\). adoptand two separate class is differentConstructionstrategy------at world world B
middle, expertbyConstructionis \textbf{ complete fullrandom} (notrelies on real Label), thismakesjoint distributionat any agree 
\(K\) lowermaintains constant etc.. 

\textbf{ world world A (noise)}: to has 
\(x \in s_1\), \(y^* = 0\). noisemechanism: withprobability \(\eta\) willLabeluniform uniform flippedto
\(\mathcal{Y} \setminus \{0\}\): 
\[\mathbb{P}(y = c \mid \text{noise}, x \in s_1) = \frac{1}{K-1}, \quad c \neq 0\]

expertaccuracy: \(\mathbb{P}(f_m(x) = y^* \mid x \in s_1) = 1 - \varepsilon_1\), incorrectuniform uniform distribution: 
\[\mathbb{P}(f_m = c \mid s_1) = \frac{\varepsilon_1}{K-1}, \quad c \neq 0\]

observationdistributionis: 
\[\mathcal{P}_{\text{noise}}(y = 0 \mid s_1) = 1 - \eta, \quad \mathcal{P}_{\text{noise}}(y = c \mid s_1) = \frac{\eta}{K-1}, \; c \neq 0\]
\[\mathcal{P}_{\text{noise}}(f_m = 0 \mid s_1) = 1 - \varepsilon_1, \quad \mathcal{P}_{\text{noise}}(f_m = c \mid s_1) = \frac{\varepsilon_1}{K-1}, \; c \neq 0\]

andbyat (A4) noise andexpertindependent, \(y\) and \(f_m\) intodetermine \(s_1\) lowerindependent. 

\textbf{ world world B (Difficulty)}: to
\(x \in s_1\), has Labeluniformis trueLabel (Nonenoise): 
\[\mathbb{P}(y^* = 0 \mid s_1) = 1 - \eta, \quad \mathbb{P}(y^* = c \mid s_1) = \frac{\eta}{K-1}, \; c \neq 0\]

\textbf{keyConstruction}: at world world B
middle, expert\textbf{notlearning}------itsprediction complete fullrandomandis independent oftrueLabel: 
\[\mathbb{P}(f_m = 0 \mid s_1) = 1 - \varepsilon_1, \quad \mathbb{P}(f_m = c \mid s_1) = \frac{\varepsilon_1}{K-1}, \; c \neq 0\]
and \(f_m \perp y^* \mid s_1\) (expertpredictionandtrueLabelindependent). 

atthisConstructionlower: - \(y\) edge actual distribution: two world world mutualsame (world world A noiseLabeldistribution =
 world world B trueLabeldistribution) - \(f_m\) edge actual distribution: two world world mutualsame (uniformis 
\(1-\varepsilon_1\) inCategory 0, \(\varepsilon_1/(K-1)\) inits he Category) -
\((y, f_m)\) joint distribution: two world world uniformis independentproduct (world world A middle \(y\) and \(f_m\)
because (A4) independent; world world B middle \(f_m\) ispurerandomthereforealsoand \(y\) independent)

becausethisto any agree 
\(K \geq 2\), \(\mathcal{P}_{\text{noise}}(x, y, \{f_m\}) = \mathcal{P}_{\text{hard}}(x, y, \{f_m\})\)
 constant holds. \(\square\)

\textbf{explains}: world world B
Constructionmiddle, ``Difficulty'' language meaning isextreme------shouldStatusmiddle\textbf{ has }sampleto has expertallsame type Difficulty (expertis identical torandom guess measure), None theory trueLabelis what what. thistoshould actual actual middle''Characteristics complete fullNoneinformation''extremely finite case (Theorem
2 middle
\(\delta = 0\)). althoughhoweverextreme, but it action is \textbf{existenceatness counter example }alreadysufficient: existenceattofew a with''Difficulty''is explains world world, itsobservationdistributionandwith''noise''is explains world world complete fullmutualsame, becausethistwooneNonemethodFrom observationdatamiddledistinguish. 

\begin{quote}
\textbf{FIXED (2026-06-28, DEFECT-12):} to \(K>2\)
Constructionuse\textbf{ complete fullrandomexpert} (\(f_m \perp y^* \mid s_1\))------thisis'' difficult degree'' a kindextreme/ retreat -ization form. for
\(K=2\), Constructionuse present actual form difficult degree (expertbiastowardCategory 0), but \(K>2\)
Constructionrequire complete fullrandomexpert come ensurejoint distributionat has Categoryuppermatches. this agree taste -ing: althoughhowever\textbf{ original thenupper}existenceatsomekind'' difficult degree''distributionandnoiseindistinguishable, butcanachievesthiskindindistinguishability'' difficult degree''Category\textbf{non- often narrow narrow}. present actual world world in the Difficultysample through often has has \textbf{structure -ization }expertincorrect model formula (such asexpertconsistentincorrect ground predictionsome special determineincorrectCategory), thisat attach add temperature andAssumptions (such as
A5 Statussamenessor A6 balanceerrordistribution)lowermayandnoisedistinguishopen come. \textbf{\(K=2\)
Construction has has moststrongpracticemutualcloseness}, becauseis it use present actual formexpertbiastowardness; \(K>2\)
Constructionprimary action is existenceatnessProof, itsextremeAssumptions (complete fullrandomexpert)at present actual 
MLIP/Medical Imaging/LLM scenariomiddle several -ish is impossible out present. 
\end{quote}

\begin{quote}
\textbf{2026-06-27 correction}: original version appendix A Constructionmiddle, world world B
expertattodeterminetrueLabellower has has non- flat all accuracy \(1-\varepsilon_1\), thisleads tofor
\(K>2\) timeexpert edge actual distributionand world world A
notmatches. correctionbackConstructionadopt'' complete fullrandomexpert'', guaranteeto any agree \(K \geq 2\)
joint distribution constant etc. ness. 
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{appendix B: Symbol/Concept table }\label{ux9644ux5f55-bux7b26ux53f7ux8868}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Symbol/Concept & Meaning & time out present \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\mathcal{P}_{\text{noise}}\) & noise world world Data Generationprocess & 2.2 \\
\(\mathcal{P}_{\text{hard}}\) & Difficulty world world Data Generationprocess & 2.3 \\
\(\eta\) & noise rate (world world A)/ Labelambiguity rate (world world B) & 2.2 \\
\(\varepsilon_1\) & Status \(s_1\) upperexpertincorrect rate & 2.2 \\
\(\varepsilon_2\) & Status \(s_2\) upperexpertincorrect rate & 2.2 \\
\(\rho\) & Status \(s_1\) edge actual probability & 2.2 \\
\(\mathcal{D}_{\text{obs}}\) & observationData Distribution & 1.1 \\
\(\mathcal{Z}^*\) & truenoisesample set & 2.5 \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{reference}\label{ux53c2ux8003ux6587ux732e}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
 Fuller, W. A. (1987). \emph{Measurement Error Models}. Wiley.
\item
 Carroll, R. J., Ruppert, D., Stefanski, L. A., \& Crainiceanu, C. M.
 (2006). \emph{Measurement Error in Nonlinear Models: A Modern
 Perspective} (2nd ed.). Chapman and Hall/CRC.
\item
 Menon, A. K., Van Rooyen, B., Ong, C. S., \& Williamson, R. C. (2015).
 Learning from corrupted binary labels via class-probability
 estimation. \emph{ICML}.
\item
 Patrini, G., Rozza, A., Menon, A. K., Nock, R., \& Qu, L. (2017).
 Making deep neural networks robust to label noise: A loss correction
 approach. \emph{CVPR}.
\item
 Dawid, A. P., \& Skene, A. M. (1979). Maximum likelihood estimation of
 observer error-rates using the EM algorithm. \emph{Journal of the
 Royal Statistical Society: Series C (Applied Statistics)}, 28(1),
 20-28.
\item
 Northcutt, C. G., Jiang, L., \& Chuang, I. L. (2021). Confident
 learning: Estimating uncertainty in dataset labels. \emph{Journal of
 Artificial Intelligence Research}, 70, 1373-1411.
\item
 Pearl, J. (2009). \emph{Causality: Models, Reasoning, and Inference}
 (2nd ed.). Cambridge University Press.
\item
 Teicher, H. (1963). Identifiability of finite mixtures. \emph{The
 Annals of Mathematical Statistics}, 34(4), 1265-1269.
\item
 McLachlan, G. J., \& Peel, D. (2000). \emph{Finite Mixture Models}.
 Wiley.
\item
 SCX Theorem 1: Multi-Expert Consistency Guarantees for Label Noise
 Detection. \texttt{./01\_noise\_detection\_guarantee.md}.
\item
 SCX Theorem 2: Weak Feature Failure Lower Bound.
 \texttt{./02\_weak\_feature\_failure.md}.
\item
 SCX Data Four-Classification.
 \texttt{../../knowledge/02\_/data four separate class.md}.
\item
 SCX Data Poisoning Defense.
 \texttt{../../knowledge/02\_/data prevent middle poison.md}.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{correctionDescription (2026-06-27)}: First version complete. Theorem 3 orientationis SCX
Assumptionssystem''necessaryness column ''------Proof A1-A6
A1-A6 are not arbitrary assumptions, andisbreakunidentifiability mustminimumstructureConditions. 

\end{document}
