% ============================================================================
%  Viewpoint 5 修正: O(d) 规范固定可通过 Cartan 分解线性化
%  Viewpoint 5 Correction: O(d) Gauge-Fixing is Linearizable via Cartan
%  SCX 规范理论 — 中英双语
%  SCX Gauge Theory — Chinese+English
% ============================================================================
\documentclass[11pt,a4paper]{article}
\usepackage[UTF8]{ctex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[margin=2.5cm]{geometry}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}

\theoremstyle{plain}
\newtheorem{theorem}{定理}
\newtheorem{proposition}[theorem]{命题}
\newtheorem{lemma}[theorem]{引理}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{定义}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{注记}
\newtheorem{correction}[theorem]{修正}

\newcommand{\R}{\mathbb{R}}
\newcommand{\SO}{\mathrm{SO}}
\newcommand{\so}{\mathfrak{so}}
\newcommand{\gauge}{g}
\newcommand{\gaugegroup}{\mathcal{G}}
\newcommand{\Skew}{\mathrm{Skew}}

\title{\textbf{SCX 规范理论观点5修正报告}\\[0.3cm]
\large O(d) 规范固定：Cartan线性化 vs.\ 非阿贝尔规范固定\\[0.2cm]
\large Viewpoint 5 Correction: Cartan Linearization vs.\ Full Non-Abelian Gauge-Fixing}
\author{SCX 审查委员会 \\ SCX Review Committee}
\date{2026年7月2日 \\ July 2, 2026}

\begin{document}
\maketitle

% ===================================================================
\section{原始声明 (被审查) \\ Original Claim (Under Review)}
% ===================================================================

\begin{quote}
\textbf{观点5 (原版):}

ACE 势能具有 O(3) 旋转对称性，使得规范群为 O(d) 而非平移群。
由于 O(d) 是非阿贝尔李群，规范固定需要\textbf{非阿贝尔规范固定}
(non-abelian gauge fixing) 技术，具体表现为 Riemannian Gauss-Newton
优化在流形上的应用。该过程具有二次收敛性。
\end{quote}

\begin{quote}
\textbf{Viewpoint 5 (Original):}

ACE potentials possess O(3) rotational symmetry, making the gauge group
O(d) rather than the translation group. Because O(d) is a non-abelian
Lie group, gauge fixing requires \textbf{non-abelian gauge fixing}
techniques, implemented as Riemannian Gauss-Newton optimization
on the manifold. This procedure has quadratic convergence.
\end{quote}

% ===================================================================
\section{审查判决 \\ Review Verdict}
% ===================================================================

\begin{center}
\fbox{\parbox{0.9\textwidth}{
\centering
\textbf{判决: 过度陈述 (OVERSTATED)}

原始声称在技术上有部分正确性，但表述严重过度：\\
(1) 将 O(d) 的非阿贝尔群论性质等同于需要完整的非阿贝尔规范固定机制；\\
(2) 忽略了 Cartan 分解可将问题线性化到李代数 $\so(d)$ 的关键事实；\\
(3) ``非阿贝尔规范固定'' 一词在物理学中特指 Yang-Mills 理论中的\\
\quad 鬼场、Gribov 副本等技术——这些\textbf{完全不适用于}本问题；\\
(4) Gauss-Newton 在残差非零时仅具有\textbf{线性收敛}（非二次）。
}}
\end{center}

\begin{center}
\fbox{\parbox{0.9\textwidth}{
\centering
\textbf{Verdict: OVERSTATED}

The original claim is partially correct in its technical content, but
its formulation is significantly overstated:\\
(1) It conflates the group-theoretic non-abelian property of O(d) with
the need for full non-abelian gauge-fixing machinery;\\
(2) It overlooks the critical fact that Cartan decomposition linearizes
the problem to the Lie algebra $\so(d)$;\\
(3) ``Non-abelian gauge fixing'' in physics specifically denotes Yang-Mills
techniques — ghost fields, Gribov copies, etc. — \textbf{none of which apply};\\
(4) Gauss-Newton has only \textbf{linear convergence} (not quadratic) when
residuals are non-zero.
}}
\end{center}

% ===================================================================
\section{为何原始声明过度 \\ Why the Original Claim Is Overstated}
% ===================================================================

\subsection{O(d) 的非阿贝尔性 ≠ 需要非阿贝尔规范固定}
\subsection*{O(d) Being Non-Abelian $\neq$ Requiring Non-Abelian Gauge-Fixing}

\noindent
\textbf{中文.}
O(d) 作为一个群确实是非阿贝尔的：对于 $R_1, R_2 \in \mathrm{O}(d)$，一般有
$R_1 R_2 \neq R_2 R_1$。然而，\textbf{群的代数性质与解决规范固定问题的算法复杂度
是两个不同层面的问题}。问题的关键在于：O(d) 作为李群，其李代数 $\so(d)$
是一个\textbf{向量空间}（维度 $\frac{d(d-1)}{2}$），而 Cartan 分解（指数映射的逆，
即对数映射）提供了从群流形到该向量空间的局部微分同胚。

\medskip
\noindent
\textbf{English.}
O(d) is indeed non-abelian as a group: for $R_1, R_2 \in \mathrm{O}(d)$,
generally $R_1 R_2 \neq R_2 R_1$. However, \textbf{the algebraic property of the
group and the algorithmic complexity of solving the gauge-fixing problem
are distinct concerns}. The crucial fact is: O(d) as a Lie group has Lie
algebra $\so(d)$, which is a \textbf{vector space} (of dimension
$\frac{d(d-1)}{2}$), and Cartan decomposition — the inverse of the
exponential map, i.e., the logarithm — provides a local diffeomorphism
from the group manifold to this vector space.

\subsection{``非阿贝尔规范固定'' 的物理学含义}
\subsection*{What ``Non-Abelian Gauge Fixing'' Means in Physics}

\noindent
\textbf{中文.}
在量子场论中，``非阿贝尔规范固定''（如 Yang-Mills 理论中的 Faddeev-Popov 方法）
涉及以下技术要素：

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{鬼场 (Ghost Fields):} 引入反对易的标量场以抵消非物理的规范自由度；
  \item \textbf{Gribov 副本 (Gribov Copies):} 规范固定条件的非唯一性问题—
        同一物理构型可能对应多个规范固定解；
  \item \textbf{BRST 对称性:} 保证鬼场与规范场之间的超对称结构；
  \item \textbf{路径积分测度:} 需要 Jacobian 行列式（Faddeev-Popov 行列式）来修正积分测度。
\end{enumerate}

\textbf{这些问题均不适用于 ACE 势能的 O(d) 规范固定问题。} ACE 中的规范固定
是一个有限维度的参数估计问题，而非无穷维度的量子场论路径积分问题。

\medskip
\noindent
\textbf{English.}
In quantum field theory, ``non-abelian gauge fixing'' (e.g., the
Faddeev-Popov procedure in Yang-Mills theory) involves:

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{Ghost Fields:} Anti-commuting scalar fields introduced to
        cancel unphysical gauge degrees of freedom;
  \item \textbf{Gribov Copies:} Non-uniqueness of gauge-fixing conditions —
        the same physical configuration may correspond to multiple gauge-fixed solutions;
  \item \textbf{BRST Symmetry:} A super-symmetry structure relating ghosts
        and gauge fields;
  \item \textbf{Path-Integral Measure:} A Jacobian determinant (Faddeev-Popov
        determinant) is required to correct the integration measure.
\end{enumerate}

\textbf{None of these issues apply to the O(d) gauge-fixing problem for ACE
potentials.} The ACE gauge-fixing is a finite-dimensional parameter estimation
problem, not an infinite-dimensional quantum field theory path-integral problem.

\subsection{Gauss-Newton 的收敛阶}
\subsection*{Convergence Order of Gauss-Newton}

\noindent
\textbf{中文.}
原始声明声称 Gauss-Newton 具有二次收敛性。这是对标准数值优化理论的误读：

\begin{itemize}
  \item Gauss-Newton 的\textbf{局部二次收敛}仅在\textbf{残差为零}（零残差问题，
        即问题在解处精确满足）时成立；
  \item 当残差非零时（这是 ACE 规范固定的典型情况——不存在完美消除所有规范
        冗余的精确解），Gauss-Newton 仅具有\textbf{线性收敛}；
  \item 线性收敛速率由问题的条件数决定：$\|x_{k+1} - x^*\| \leq \rho \|x_k - x^*\|$，
        其中 $\rho < 1$ 但 $\rho$ 可能接近 1。
\end{itemize}

\medskip
\noindent
\textbf{English.}
The original claim asserts quadratic convergence for Gauss-Newton. This
misreads standard numerical optimization theory:

\begin{itemize}
  \item Gauss-Newton exhibits \textbf{local quadratic convergence} only when
        \textbf{residuals are zero} (zero-residual problem, where the problem
        is exactly satisfied at the solution);
  \item When residuals are non-zero (the typical case for ACE gauge-fixing —
        there is no exact solution that perfectly eliminates all gauge
        redundancy), Gauss-Newton has only \textbf{linear convergence};
  \item The linear convergence rate is governed by the problem's condition
        number: $\|x_{k+1} - x^*\| \leq \rho \|x_k - x^*\|$, where $\rho < 1$
        but $\rho$ may be close to 1.
\end{itemize}

% ===================================================================
\section{修正后的正确表述 \\ Corrected Claim}
% ===================================================================

\begin{correction}[观点5 修正版]
\label{cor:viewpoint5}

ACE 势能具有 O(3) 旋转对称性，使得规范群为 O(d) 而非平移群。
虽然 O(d) 作为群是非阿贝尔的，但规范固定问题可以通过 \textbf{Cartan 分解}
线性化：将群元素写为 $g_m = g_0 \cdot \exp(\xi_m)$，其中
$\xi_m \in \so(d)$ 属于李代数。由于 $\so(d)$ 是一个维度为
$\frac{d(d-1)}{2}$ 的向量空间，问题退化为该向量空间上的\textbf{标准线性最小二乘问题}。

在对数映射有效的范围内（即接近恒等变换时），这一线性化仅增加
$\mathcal{O}(d^2)$ 的计算开销。完整的 Riemannian Gauss-Newton
仅在远离解时需要，且在残差非零时具有\textbf{线性收敛}（非二次收敛）。

\bigskip
\noindent
\textit{ACE potentials possess O(3) rotational symmetry, making the gauge
group O(d) rather than the translation group. While O(d) is non-abelian as
a group, the gauge-fixing problem can be linearized via \textbf{Cartan
decomposition}: writing $g_m = g_0 \cdot \exp(\xi_m)$ with
$\xi_m \in \so(d)$ in the Lie algebra reduces it to a
\textbf{standard linear least-squares problem} on a vector space of
dimension $\frac{d(d-1)}{2}$.

For applications where the log map is valid (near the identity), this
linearization adds only $\mathcal{O}(d^2)$ overhead. The full Riemannian
Gauss-Newton is needed only when far from the solution, and converges
\textbf{linearly} (not quadratically) when residuals are non-zero.}
\end{correction}

% ===================================================================
\section{数学论证 \\ Mathematical Justification}
% ===================================================================

\subsection{Cartan 分解与对数映射}
\subsection*{Cartan Decomposition and the Logarithm Map}

\begin{definition}[Cartan 分解]
\label{def:cartan}

设 $G = \mathrm{O}(d)$ 为 $d$ 维正交群，其李代数为：
\begin{equation}
  \so(d) = \{ \Omega \in \R^{d \times d} : \Omega^\top = -\Omega \},
\end{equation}
即所有 $d \times d$ 反对称矩阵的集合。$\so(d)$ 是一个维度为
$\frac{d(d-1)}{2}$ 的\textbf{向量空间}。

\textbf{Cartan 分解}表明：每个 $g \in \mathrm{O}(d)$ 在恒等元附近
可唯一地写为：
\begin{equation}
  g = \exp(\xi), \quad \xi \in \so(d),
\end{equation}
其中 $\exp: \so(d) \to \mathrm{O}(d)$ 是矩阵指数映射。其逆映射
$\log: \mathrm{O}(d) \to \so(d)$（矩阵对数）在 $\mathrm{O}(d)$ 上
除 $\det(g) = -1$ 且 $g$ 有特征值 $-1$ 的测度零子集外处处有定义。
\end{definition}

\bigskip
\noindent
\textbf{Definition (Cartan Decomposition).}
Let $G = \mathrm{O}(d)$ be the $d$-dimensional orthogonal group with Lie algebra:
\begin{equation}
  \so(d) = \{ \Omega \in \R^{d \times d} : \Omega^\top = -\Omega \},
\end{equation}
the set of all $d \times d$ skew-symmetric matrices. $\so(d)$ is a
\textbf{vector space} of dimension $\frac{d(d-1)}{2}$.

\textbf{Cartan decomposition} states: every $g \in \mathrm{O}(d)$ near the
identity can be uniquely written as:
\begin{equation}
  g = \exp(\xi), \quad \xi \in \so(d),
\end{equation}
where $\exp: \so(d) \to \mathrm{O}(d)$ is the matrix exponential. Its
inverse $\log: \mathrm{O}(d) \to \so(d)$ (the matrix logarithm) is
defined everywhere on $\mathrm{O}(d)$ except a measure-zero subset
where $\det(g) = -1$ and $g$ has eigenvalue $-1$.

% -------------------------------------------------------------------
\subsection{规范固定问题的线性化}
\subsection*{Linearization of the Gauge-Fixing Problem}

\noindent
\textbf{中文.}
考虑 ACE 势能中的规范固定问题。设 $\{g_m\}_{m=1}^M \subset \mathrm{O}(d)$
为 $M$ 个专家的规范变换（旋转矩阵）。规范固定目标为最小化某个损失函数
$\mathcal{L}(\{g_m\})$，该函数度量旋转后的势能之间的不一致性。

\textbf{线性化步骤}：

\begin{enumerate}[label=\textbf{步骤 \arabic*}:, leftmargin=*]
  \item 选取参考旋转 $g_0 \in \mathrm{O}(d)$（如当前估计）。
  \item 将每个 $g_m$ 参数化为：
        \begin{equation}\label{eq:param}
          g_m = g_0 \cdot \exp(\xi_m), \quad \xi_m \in \so(d).
        \end{equation}
  \item 将反对称矩阵 $\xi_m$ 向量化。$\so(d)$ 同构于 $\R^{d(d-1)/2}$，
        基为 $\{E_{ij} = e_i e_j^\top - e_j e_i^\top\}_{1 \leq i < j \leq d}$：
        \begin{equation}\label{eq:basis}
          \xi_m = \sum_{1 \leq i < j \leq d} \omega_m^{ij} \cdot E_{ij},
          \quad \omega_m^{ij} \in \R.
        \end{equation}
  \item 将损失函数 $\mathcal{L}$ 关于 $\{\omega_m^{ij}\}$ 展开至一阶：
        \begin{equation}\label{eq:expand}
          \mathcal{L}(\{\omega_m^{ij}\}) \approx
          \mathcal{L}_0 + \mathbf{J} \cdot \bm{\omega} + \frac{1}{2} \bm{\omega}^\top \mathbf{H} \bm{\omega},
        \end{equation}
        其中 $\bm{\omega} \in \R^{M \cdot d(d-1)/2}$ 为所有参数的拼接向量，
        $\mathbf{J}$ 为梯度（Jacobian），$\mathbf{H}$ 为近似 Hessian。
  \item 求解线性系统 $\mathbf{H} \cdot \bm{\omega}^* = -\mathbf{J}^\top$ 得到
        Newton 步，然后通过 $\exp$ 映射回群流形。
\end{enumerate}

\textbf{关键洞见：} 步骤 2--5 的所有计算都在\textbf{向量空间}
$\R^{M \cdot d(d-1)/2}$ 上进行，而非在弯曲的群流形 $\mathrm{O}(d)^M$ 上。
这就是 Cartan 分解的威力——将非阿贝尔群上的优化问题转化为向量空间上的
标准线性最小二乘问题。

\medskip
\noindent
\textbf{English.}
Consider the gauge-fixing problem in ACE potentials. Let
$\{g_m\}_{m=1}^M \subset \mathrm{O}(d)$ be the gauge transformations
(rotation matrices) of $M$ experts. The gauge-fixing objective is to
minimize a loss function $\mathcal{L}(\{g_m\})$ measuring inconsistency
among rotated potentials.

\textbf{Linearization procedure}:

\begin{enumerate}[label=\textbf{Step \arabic*}:, leftmargin=*]
  \item Choose a reference rotation $g_0 \in \mathrm{O}(d)$ (e.g., the current estimate).
  \item Parameterize each $g_m$ as:
        \begin{equation}
          g_m = g_0 \cdot \exp(\xi_m), \quad \xi_m \in \so(d).
        \end{equation}
  \item Vectorize the skew-symmetric matrix $\xi_m$. $\so(d)$ is isomorphic to
        $\R^{d(d-1)/2}$ with basis
        $\{E_{ij} = e_i e_j^\top - e_j e_i^\top\}_{1 \leq i < j \leq d}$:
        \begin{equation}
          \xi_m = \sum_{1 \leq i < j \leq d} \omega_m^{ij} \cdot E_{ij},
          \quad \omega_m^{ij} \in \R.
        \end{equation}
  \item Expand the loss $\mathcal{L}$ to first order in $\{\omega_m^{ij}\}$:
        \begin{equation}
          \mathcal{L}(\{\omega_m^{ij}\}) \approx
          \mathcal{L}_0 + \mathbf{J} \cdot \bm{\omega} + \frac{1}{2} \bm{\omega}^\top \mathbf{H} \bm{\omega},
        \end{equation}
        where $\bm{\omega} \in \R^{M \cdot d(d-1)/2}$ concatenates all parameters,
        $\mathbf{J}$ is the gradient (Jacobian), and $\mathbf{H}$ is the approximate Hessian.
  \item Solve the linear system $\mathbf{H} \cdot \bm{\omega}^* = -\mathbf{J}^\top$
        for the Newton step, then map back to the group manifold via $\exp$.
\end{enumerate}

\textbf{Key insight:} All computations in Steps 2--5 take place in the
\textbf{vector space} $\R^{M \cdot d(d-1)/2}$, not on the curved group
manifold $\mathrm{O}(d)^M$. This is the power of Cartan decomposition —
transforming an optimization problem on a non-abelian group into a standard
linear least-squares problem on a vector space.

% -------------------------------------------------------------------
\subsection{复杂度分析}
\subsection*{Complexity Analysis}

\begin{proposition}[线性化的计算开销]
\label{prop:complexity}

对于 $M$ 个专家和 $d$ 维旋转群：
\begin{itemize}
  \item 参数总数：$M \cdot \frac{d(d-1)}{2}$。
  \item 每次迭代的矩阵指数 $\exp(\xi_m)$ 计算开销：$\mathcal{O}(d^3)$（通过
        Rodrigues 公式可优化至 $\mathcal{O}(d^2)$，尤其对于 $d=3$ 即 SO(3)）。
  \item Jacobian 计算：通过自动微分或解析梯度，$\mathcal{O}(M \cdot d^3)$。
  \item 线性方程组求解：$\mathcal{O}((M d^2)^3)$ 对于稠密直接求解，
        或 $\mathcal{O}(M d^2 \cdot \kappa)$ 对于共轭梯度法。
\end{itemize}

对于实际场景 ($d=3$, $M \lesssim 100$)，总开销为 $\mathcal{O}(M \cdot 27)$
次浮点运算——完全可行，且远低于完整的 Riemannian 优化。
\end{proposition}

\bigskip
\noindent
\textbf{Proposition (Computational Overhead of Linearization).}
For $M$ experts and $d$-dimensional rotation group:
\begin{itemize}
  \item Total parameters: $M \cdot \frac{d(d-1)}{2}$.
  \item Matrix exponential $\exp(\xi_m)$ per iteration: $\mathcal{O}(d^3)$
        (reducible to $\mathcal{O}(d^2)$ via Rodrigues formula, especially
        for $d=3$ i.e., SO(3)).
  \item Jacobian computation: $\mathcal{O}(M \cdot d^3)$ via automatic
        differentiation or analytic gradients.
  \item Linear system solve: $\mathcal{O}((M d^2)^3)$ for dense direct
        solver, or $\mathcal{O}(M d^2 \cdot \kappa)$ for conjugate gradient.
\end{itemize}

For practical settings ($d=3$, $M \lesssim 100$), total overhead is
$\mathcal{O}(M \cdot 27)$ floating-point operations — entirely feasible
and far cheaper than full Riemannian optimization.

% ===================================================================
\section{Riemannian Gauss-Newton 的适用条件}
\section*{When Riemannian Gauss-Newton IS Needed}

\noindent
\textbf{中文.}
Cartan 线性化并非万能。对数映射 $\log: \mathrm{O}(d) \to \so(d)$ 的有效
范围受限于以下条件：

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{接近恒等元:} 当 $g_m$ 远离恒等变换时（如旋转角 $> \pi/2$），
        $\log(g_m)$ 可能不存在或产生较大的线性化误差；
  \item \textbf{反射分量:} $\mathrm{O}(d)$ 包含 $\det = -1$ 的反射分支，
        对数映射在该分支上存在奇点；
  \item \textbf{大规模全局优化:} 当需要在整个 $\mathrm{O}(d)$ 上搜索全局
        最优时，Riemannian 优化（如 Riemannian trust-region 方法）更为稳健。
\end{enumerate}

\textbf{在这些情况下，Riemannian Gauss-Newton 是合适的工具}——但其收敛阶仍为
线性（当残差非零时），且其每次迭代开销显著高于线性化方法
（需计算 Riemannian 联络、平行移动和测地线）。

\medskip
\noindent
\textbf{English.}
Cartan linearization is not a panacea. The logarithm map
$\log: \mathrm{O}(d) \to \so(d)$ is valid only under:

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{Proximity to identity:} When $g_m$ is far from the identity
        (e.g., rotation angle $> \pi/2$), $\log(g_m)$ may not exist or may
        incur large linearization error;
  \item \textbf{Reflection components:} $\mathrm{O}(d)$ includes the
        $\det = -1$ reflection branch, where the logarithm has singularities;
  \item \textbf{Large-scale global optimization:} When searching for a global
        optimum over all of $\mathrm{O}(d)$, Riemannian optimization (e.g.,
        Riemannian trust-region methods) is more robust.
\end{enumerate}

\textbf{In these cases, Riemannian Gauss-Newton is the appropriate tool} —
but its convergence remains linear (when residuals are non-zero),
and its per-iteration cost far exceeds that of the linearized method
(requiring computation of Riemannian connections, parallel transport,
and geodesics).

% ===================================================================
\section{修正前后对比 \\ Before-and-After Comparison}
% ===================================================================

\begin{table}[h]
\centering
\caption{观点5修正前后对比 \\ Comparison Before and After Correction}
\begin{tabular}{@{}p{4.5cm}p{5cm}p{5cm}@{}}
\toprule
\textbf{要素\newline Aspect} &
\textbf{原始声明\newline Original} &
\textbf{修正后\newline Corrected} \tabularnewline
\midrule

规范固定方法\newline Gauge-fixing method &
``非阿贝尔规范固定''\newline ``Non-abelian gauge fixing'' &
Cartan 分解 $\to$ 向量空间线性最小二乘\newline Cartan decomp. $\to$ vector-space linear LS \tabularnewline
\midrule

数学机制\newline Mathematical mechanism &
Yang-Mills 式（鬼场、Gribov）\newline Yang-Mills style (ghosts, Gribov) &
对数映射 $\log: \mathrm{O}(d) \to \so(d)$\newline Logarithm map $\log: \mathrm{O}(d) \to \so(d)$ \tabularnewline
\midrule

计算复杂度\newline Computational complexity &
未明确，暗示高开销\newline Unspecified, implied high &
$\mathcal{O}(d^2)$ 附加开销，接近恒等元时极低\newline $\mathcal{O}(d^2)$ overhead, near-zero near identity \tabularnewline
\midrule

Gauss-Newton 收敛\newline Gauss-Newton convergence &
声称二次收敛\newline Claimed quadratic &
\textbf{线性收敛}（残差非零时）\newline \textbf{Linear convergence} (non-zero residuals) \tabularnewline
\midrule

适用范围\newline Scope of applicability &
未限定\newline Unqualified &
接近恒等元时用线性化；远离时用 Riemannian\newline Linearized near identity; Riemannian when far \tabularnewline
\bottomrule
\end{tabular}
\end{table}

% ===================================================================
\section{总结 \\ Summary}
% ===================================================================

\noindent
\textbf{中文总结:}

观点5的核心物理直觉——ACE 势能具有 O(3) 旋转对称性，规范群为 O(d)——是\textbf{正确的}。
但由此断言需要``非阿贝尔规范固定''是\textbf{过度陈述}，原因有三：

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{Cartan 分解的存在性:} 对数映射 $\log: \mathrm{O}(d) \to \so(d)$
        将群流形上的问题线性化到向量空间 $\so(d) \cong \R^{d(d-1)/2}$，
        使其退化为标准线性最小二乘——无需鬼场、Gribov 副本或 BRST 对称性。
  \item \textbf{物理学语义的误用:} ``非阿贝尔规范固定''在量子场论中有精确的
        技术含义（Faddeev-Popov 方法），这些技术完全不适用于 ACE 势能的
        有限维参数估计问题。
  \item \textbf{收敛阶的错误声称:} Gauss-Newton 的二次收敛仅在零残差情况下
        成立。ACE 规范固定问题具有非零残差，因此收敛仅为线性。
\end{enumerate}

修正后的表述准确反映了问题的真实数学结构：一个可通过 Cartan 分解线性化的
非阿贝尔对称性，其求解复杂度为 $\mathcal{O}(d^2)$，仅在远离解时需要完整的
Riemannian 优化。

\bigskip
\noindent
\textbf{English Summary:}

The core physical intuition of Viewpoint 5 — that ACE potentials have O(3)
rotational symmetry, making the gauge group O(d) — is \textbf{correct}.
However, concluding that this requires ``non-abelian gauge fixing'' is
\textbf{overstated} for three reasons:

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{Existence of Cartan decomposition:} The logarithm map
        $\log: \mathrm{O}(d) \to \so(d)$ linearizes the problem on the
        group manifold to the vector space $\so(d) \cong \R^{d(d-1)/2}$,
        reducing it to standard linear least squares — no ghost fields,
        Gribov copies, or BRST symmetry required.
  \item \textbf{Misuse of physics terminology:} ``Non-abelian gauge fixing''
        has a precise technical meaning in quantum field theory
        (Faddeev-Popov procedure), and none of that machinery applies to
        the finite-dimensional parameter estimation problem of ACE potentials.
  \item \textbf{Incorrect convergence claim:} Gauss-Newton has quadratic
        convergence only in the zero-residual case. The ACE gauge-fixing
        problem has non-zero residuals, so convergence is only linear.
\end{enumerate}

The corrected formulation accurately reflects the true mathematical structure:
a non-abelian symmetry that can be linearized via Cartan decomposition, with
$\mathcal{O}(d^2)$ solution complexity, requiring full Riemannian optimization
only when far from the solution.

% ===================================================================
\section*{致谢 \\ Acknowledgments}
% ===================================================================

\noindent
感谢审稿人对 O(d) 规范群与李代数之间关系的深刻洞察，以及对 Gauss-Newton
收敛理论的准确提醒。本修正基于数值优化、李群理论和量子场论中规范固定的
标准结果。

\bigskip
\noindent
\textit{We thank the reviewer for the penetrating insight into the relationship
between the O(d) gauge group and its Lie algebra, and for the accurate reminder
about Gauss-Newton convergence theory. This correction is based on standard
results in numerical optimization, Lie group theory, and gauge fixing in
quantum field theory.}

\end{document}
