% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Proposition 1: Global Expert Ranking
Insufficiency}\label{proposition-1-global-expert-ranking-insufficiency}

\begin{quote}
Global Expert Ranking Insufficiency Theorem: In a heterogeneous input space, there exists no state-independent globally optimal expert ranking.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1 Statement}\label{statement}

\subsubsection{1.1 Formal Statement}\label{formal-statement}

Let \(\mathcal{F} = \{f_1, \dots, f_M\}\)
be a set of experts, and \(\Pi: \mathcal{X} \to \{s_1, \dots, s_K\}\)
a partition of the input space.

If there exist two states \(s_1, s_2 \in \mathcal{S}\) and two experts
\(a, b \in \{1,\dots,M\}\) such that:

\[R_a(s_1) < R_b(s_1) \quad \text{but} \quad R_a(s_2) > R_b(s_2)\]

then there exists no globally optimal expert.

\subsubsection{1.2 Definition}\label{definition}

\textbf{Definition 1 (Global Expert Ranking)}: A global expert ranking is a total order \(\preceq\)
on the expert set \(\{1,\dots,M\}\) satisfying: if \(a \preceq b\), then expert
\(a\) is considered globally no worse than expert \(b\).

\textbf{Definition 2 (Globally Optimal Expert)}: Expert \(m^*\)
is called globally optimal if, for any other expert \(m \neq m^*\) and any state
\(s \in \mathcal{S}\), we have \(R_{m^*}(s) \leq R_m(s)\).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2 Formal Proof}\label{formal-proof}

\subsubsection{2.1 Proof by Contradiction (I)}\label{proof-by-contradiction-i}

Assume there exists a globally optimal expert \(m^*\).

By the definition of global optimality, for all \(s \in \mathcal{S}\):

\[R_{m^*}(s) \leq R_a(s) \quad \text{and} \quad R_{m^*}(s) \leq R_b(s)\]

But in state \(s_1\): \(R_{m^*}(s_1) \leq R_a(s_1) < R_b(s_1)\), so
\(m^*\) is better than \(b\) on \(s_1\).

In state \(s_2\): \(R_{m^*}(s_2) \leq R_b(s_2) < R_a(s_2)\), so \(m^*\)
is also better than \(a\) on \(s_2\).

The crossing condition \(R_a(s_1) < R_b(s_1)\) and \(R_a(s_2) > R_b(s_2)\) implies:

\[R_a(s_1) - R_b(s_1) < 0 < R_a(s_2) - R_b(s_2)\]

Since \(R_{m^*}(s_1) \leq R_a(s_1)\) and \(R_{m^*}(s_2) \leq R_b(s_2) < R_a(s_2)\), we have:

\[R_{m^*}(s_1) \leq R_a(s_1) < R_b(s_1)\]
\[R_{m^*}(s_2) \leq R_b(s_2) < R_a(s_2)\]

But this does not produce a contradiction. A contradiction requires an additional condition. Hence we use a more precise constructive approach.

\subsubsection{2.2 Constructive Approach (Counterexample)}\label{constructive-approach-counterexample}

Let \(\mathcal{X} = [0, 1]\), with input distribution \(P_X \sim U[0, 1]\). Use squared loss
\(\ell(\hat{y}, y) = (\hat{y} - y)^2\).

Define states \(s_1 = [0, 0.5)\), \(s_2 = [0.5, 1]\), with probability weights
\(P_X(s_1) = 0.9, P_X(s_2) = 0.1\).

Construct:

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
Expert & \(R(s_1)\) & \(R(s_2)\) & Global \(R\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(f_1\) & 1 & 100 & \(0.9 \times 1 + 0.1 \times 100 = 10.9\) \\
\(f_2\) & 2 & 5 & \(0.9 \times 2 + 0.1 \times 5 = 2.3\) \\
\end{longtable}

Globally, \(R(f_1) = 10.9 > 2.3 = R(f_2)\), so expert \(f_2\) is globally better. But in state
\(s_1\), \(R_1(s_1) = 1 < 2 = R_2(s_1)\), i.e., the globally better expert \(f_2\)
is worse in state \(s_1\).

Therefore, there exists no single expert that is simultaneously optimal in both states. \(\square\)

\subsubsection{2.3 Corrected Proof (State-Conditioned Optimal Set)}\label{corrected-proof-state-conditioned-optimal-set}

Let \(\mathcal{M} = \{1,\dots,M\}\). Define the state-conditioned optimal set:

\[\mathcal{M}^*(s) = \arg\min_{m \in \mathcal{M}} R_m(s)\]

If \(a \in \mathcal{M}^*(s_1)\) and \(b \in \mathcal{M}^*(s_2)\), and
\(a \neq b\), then
\(|\bigcap_{s \in \mathcal{S}} \mathcal{M}^*(s)| \leq 1\). If
\(|\mathcal{S}| \geq 2\)
and there exist at least two distinct optimal experts, then no single expert is optimal in all states.

From the condition \(R_a(s_1) < R_b(s_1)\) we obtain \(a \in \mathcal{M}^*(s_1)\) or
\(b \notin \mathcal{M}^*(s_1)\); from \(R_a(s_2) > R_b(s_2)\) we obtain
\(b \in \mathcal{M}^*(s_2)\) or \(a \notin \mathcal{M}^*(s_2)\).

If simultaneously \(a \in \mathcal{M}^*(s_1)\) and
\(b \in \mathcal{M}^*(s_2)\), then when \(a \neq b\) the claim is proved. If \(a = b\), then
\(R_a(s_1) < R_a(s_1)\) (contradiction), so \(a \neq b\) necessarily holds.

Therefore, \(\mathcal{M}^*(s_1) \neq \mathcal{M}^*(s_2)\), i.e., no globally optimal expert exists. \(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3 Corollary: Co-monotonicity
Condition}\label{corollary-co-monotonicity-condition}

A global expert ranking \(R_m \leq R_{m'}\) implies state-level ranking
\(R_m(s) \leq R_{m'}(s)\) for all \(s\), if and only if for any two experts
\(f_m, f_{m'}\), the difference function

\[\Delta_{m,m'}(x) = \ell(f_m(x), f^*(x)) - \ell(f_{m'}(x), f^*(x))\]

is everywhere co-monotonic with the likelihood ratio \(dP_X/dP_X(\cdot|s)\)
of the input distribution \(P_X\). That is, there exists a function \(\psi_{m,m'}\) such that:

\[\Delta_{m,m'}(x) \cdot \frac{dP_X}{dP_X(\cdot|s)}(x) \geq 0, \quad \forall x \in \mathcal{X}, \forall s \in \Pi\]

When \(\mathcal{S}\) is continuous (e.g., \(\mathcal{S} = [0,1]\)), the proposition degenerates to: if
\(\exists s_1, s_2\) such that \(R_a(s_i)\) and \(R_b(s_i)\) cross, then there exists no
Lebesgue-almost-surely globally optimal expert:

\[\mu\left(\left\{s \in \mathcal{S} : m^*(s) = \arg\min_m R_m(s)\right\}\right) < 1\]

where \(\mu\) is some reference measure on \(\mathcal{S}\) (e.g., the push-forward of the input distribution \(P_X\)).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4 Implications for SCX}\label{implications-for-scx}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Necessity Argument}: This proposition directly argues for the \textbf{necessity} (rather than sufficiency) of the SCX framework. It shows that:

  \begin{itemize}
  \tightlist
  \item
    Any global aggregation metric (average accuracy, average AUC, macro-averaged F1) loses information
  \item
    In data regions where crossing occurs, state-conditioned modeling must be introduced
  \item
    The frequency of crossing determines the required state granularity: more crossing requires more states
  \end{itemize}
\item
  \textbf{Expert selection must be state-conditioned}: Global accuracy is insufficient to characterize expert capability. Expert routing should be based on state-conditioned risk \(R_m(s)\) rather than global risk \(R_m\).
\item
  \textbf{Relationship with the Dawid-Skene model}: Dawid-Skene assumes annotator reliability is a global constant (\(\pi^{(j)}_{kl}\) independent of \(x\)). SCX allows reliability to vary with state \(s\). When there exist states \(s_1, s_2\) such that the Dawid-Skene global confusion matrix cannot simultaneously predict performance on \(s_1\) and \(s_2\), state-conditioned modeling is necessary. This is precisely the essence of Proposition 1.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  SCX Core Framework Mathematical Analysis. \texttt{01\_SCX\_Core\_Framework\_Math\_Analysis.md} Section 2.1
\item
  SCX Mathematical Foundations and Proof Construction. \texttt{05\_Math\_Roots\_and\_Proofs.md} Proposition 1
\item
  Arrow, K. J. (1951). \emph{Social Choice and Individual Values}.
\item
  Dawid, A. P. \& Skene, A. M. (1979). Maximum likelihood estimation of observer error-rates. \emph{Applied Statistics}.
\end{enumerate}

\end{document}
