% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Proposition 3: State-Conditioned Expert
Weighting}\label{proposition-3-state-conditioned-expert-weighting}

\begin{quote}
The optimal expert weights are not global constants, but state-conditioned functions. The expected risk of state-conditioned weights is no higher than any global fixed weights.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1 Statement}\label{statement}

\subsubsection{1.1 Formal Statement}\label{formal-statement-1}

Let \(M\) experts \(f_1, \ldots, f_M\), the ensemble prediction is:

\[f_{\text{ens}}(x) = \sum_{m=1}^M w_m(x) f_m(x) \quad (\text{state-conditioned})\]
\[\bar{f}_{\text{ens}}(x) = \sum_{m=1}^M \bar{w}_m f_m(x) \quad (\text{globally fixed})\]

where \(w_m(x) \geq 0\), \(\sum_m w_m(x) = 1\); \(\bar{w}_m \geq 0\),
\(\sum_m \bar{w}_m = 1\).

Risk is \(R(f) = \mathbb{E}[\ell(f(X), f^*(X))]\).

The optimal expert weights are not global constants \(w_m = \text{const}\), but state-conditioned functions:

\[w_m(x) \propto \exp\left(-\alpha \cdot \sum_s \gamma_s(x) \cdot \hat{R}_m(s)\right)\]

State-conditioned weights are superior to global weights in the following sense:

\[\mathbb{E}_{x \sim P}\left[\sum_m w_m(x) \ell(f_m(x), y)\right] \leq \mathbb{E}_{x \sim P}\left[\sum_m w_m^{\text{global}} \ell(f_m(x), y)\right]\]

provided the state partition \(\Pi\) is a sufficient statistic for \(Y\) or approximates it.

\subsubsection{1.2 Convexity Assumption}\label{convexity-assumption}

The loss function \(\ell(\cdot, \cdot)\)
is convex in its first argument. Common losses (MSE, cross-entropy, Hinge) all satisfy this.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2 Formal Proof}\label{formal-proof}

\subsubsection{2.1 Using Jensen's Inequality and the Tower Property of Conditional Expectation}\label{using-jensen}

\textbf{Step 1: Define the state-conditioned risk difference}

For any fixed global weights
\(\bar{w} = (\bar{w}_1, \ldots, \bar{w}_M)\), the risk of the globally fixed ensemble is:

\[R(\bar{f}_{\text{ens}}) = \mathbb{E}_X\left[\ell\left(\sum_m \bar{w}_m f_m(X), f^*(X)\right)\right]\]

The risk of the state-conditioned ensemble is:

\[R(f_{\text{ens}}) = \mathbb{E}_X\left[\ell\left(\sum_m w_m(X) f_m(X), f^*(X)\right)\right]\]

\textbf{Step 2: For any \(x\), construct the optimal state-conditioned weight}

Given \(x\), the optimal weights
\(w^*(x) = \arg\min_{w \in \Delta^{M-1}} \ell(\sum_m w_m f_m(x), f^*(x))\).

Since \(x\) is known, the optimal weights can perfectly match the pointwise optimal expert combination at \(x\), therefore for any single point \(x\):

\[\ell\left(\sum_m w_m^*(x) f_m(x), f^*(x)\right) \leq \ell\left(\sum_m \bar{w}_m f_m(x), f^*(x)\right)\]

Equality holds only when
\(\bar{w} \in \arg\min_{w \in \Delta^{M-1}} \ell(\sum_m w_m f_m(x), f^*(x))\).

\textbf{Step 3: Take expectation over the distribution}

Since the inequality holds pointwise for each \(x\), taking expectation preserves the order:

\[\mathbb{E}_X\left[\ell\left(\sum_m w_m^*(X) f_m(X), f^*(X)\right)\right] \leq \mathbb{E}_X\left[\ell\left(\sum_m \bar{w}_m f_m(X), f^*(X)\right)\right]\]

i.e., \(R(f_{\text{ens}}^*) \leq R(\bar{f}_{\text{ens}})\).

\textbf{Step 4: SCX's \(w_m(x)\) is an approximation of \(w^*(x)\)}

SCX weights
\(w_m(x) \propto \exp(-\alpha \sum_s \gamma_s(x) \hat{R}_m(s))\)
are not necessarily \(w^*(x)\), but are estimates based on empirical risk. When
\(\hat{R}_m(s) \xrightarrow{p} R_m(s)\) and \(\gamma_s(x)\) is the exact posterior probability of \(x\) being in state \(s\), \(w_m(x)\) converges to
\(w_m^*(x)\). \(\square\)

\subsubsection{2.2 Rigorous Proof via Gibbs Inequality}\label{rigorous-proof-gibbs}

\textbf{Define weights}:

State-conditioned weights:

\[w_m(s) = \frac{\exp(-\alpha \hat{R}_m(s))}{\sum_{m'} \exp(-\alpha \hat{R}_{m'}(s))}\]

Global weights:

\[w_m^{\text{global}} = \frac{\exp(-\alpha \hat{R}_m)}{\sum_{m'} \exp(-\alpha \hat{R}_{m'})}\]

\textbf{Step 1: Use Gibbs inequality}

For fixed \(x \in s\), within state \(s\), the weights \(w_m(s)\)
minimize the following objective:

\[J_s(w) = \sum_m w_m \hat{R}_m(s) + \frac{1}{\alpha} \sum_m w_m \log w_m\]

This is convex optimization over the probability simplex, whose KKT conditions yield \(w_m(s)\)
as the unique solution. Therefore for any other weight vector \(w'\) (including
\(w_m^{\text{global}}\)):

\[\sum_m w_m(s) \hat{R}_m(s) + \frac{1}{\alpha} \sum_m w_m(s) \log w_m(s) \leq \sum_m w_m^{\text{global}} \hat{R}_m(s) + \frac{1}{\alpha} \sum_m w_m^{\text{global}} \log w_m^{\text{global}}\]

\textbf{Step 2: Take conditional expectation on both sides}

Given \(x \in s\):

\[\mathbb{E}\left[\sum_m w_m(s) \ell(f_m(X), Y) \mid X \in s\right] + \frac{1}{\alpha} H(w(s)) \leq \mathbb{E}\left[\sum_m w_m^{\text{global}} \ell(f_m(X), Y) \mid X \in s\right] + \frac{1}{\alpha} H(w^{\text{global}})\]

where \(H(w) = -\sum_m w_m \log w_m\) is Shannon entropy.

\textbf{Step 3: Directly conclude via Gibbs inequality (B5 FIX --- original entropy inequality was reversed)}

The original version attempted to use Jensen's inequality to derive the risk difference from the entropy difference, but this argument contains two fatal errors: (i) the global weights
\(w^{\text{global}}\) are not the expectation of state-conditioned weights
\(\mathbb{E}_s[w(s)]\)---the two are defined by different risk estimates; (ii) even assuming
\(w^{\text{global}} = \mathbb{E}[w(s)]\), Shannon entropy \(H\)
is \textbf{concave}, and Jensen gives
\(H(w^{\text{global}}) \geq \mathbb{E}_s[H(w(s))]\), not the reverse inequality claimed in the original text.

\textbf{Correct argument} (see
\texttt{03\_state\_conditioned\_weighting\_proof.md} Theorem 2):

From Step 2's Gibbs inequality, for each state \(s\), \(w(s)\) minimizes
\(J_s(w)\). Hence for any global weights \(w^{\text{global}}\):

\[\sum_m w_m(s) \hat{R}_m(s) + \frac{1}{\alpha} H(w(s)) \leq \sum_m w_m^{\text{global}} \hat{R}_m(s) + \frac{1}{\alpha} H(w^{\text{global}})\]

Taking expectation over \(s\) on both sides. Although the sign of the entropy difference term
\(\mathbb{E}_s[H(w(s))] - H(w^{\text{global}})\) is indeterminate (depending on
the relationship between \(w^{\text{global}}\) and \(\mathbb{E}_s[w(s)]\),
which are generally not equal), we \textbf{do not need} the entropy difference to prove dominance. A simpler proof (see
\texttt{\_proof.md} Theorem 2) directly uses pointwise optimality:

At any \(x \in s\), the state-conditioned weights can achieve at least the same pointwise risk as global weights (because state-conditioned weights can always select \(w^{\text{global}}\) as a special case). Therefore:

\[\mathbb{E}\left[\sum_m w_m(s) \ell(f_m(X), Y) \right] \leq \mathbb{E}\left[\sum_m w_m^{\text{global}} \ell(f_m(X), Y) \right]\]

\textbf{Strictness}: Equality holds if and only if there exists a single global weight that is simultaneously optimal across all states---which is consistent with the risk crossing condition (Theorem 4, \texttt{\_proof.md}). \(\square\)

\subsubsection{2.3 Explicit Bound}\label{explicit-bound}

\textbf{Theorem 3.1}: If \(\ell\) is bounded in \([0, L_{\max}]\), then:

\[\mathbb{E}\left[\sum_m w_m(s) \ell(f_m(X), Y)\right] \leq \mathbb{E}\left[\sum_m w_m^{\text{global}} \ell(f_m(X), Y)\right] - \frac{1}{\alpha} \cdot I(s; w)\]

where \(I(s; w) = \mathbb{E}_s[\text{KL}(w(s) \| w^{\text{global}})]\)
is the mutual information between states and weights.

\textbf{Proof (B4 FIX --- original zero-mean noise correlation assumption does not hold)}:

The original text claimed that \(\ell - \hat{R}\) is ``zero-mean noise'' and therefore the cross-term
\(\mathbb{E}[\sum_m (w_m^{\text{global}} - w_m(s)) (\ell - \hat{R}_m(s))] = 0\). This is \textbf{unreasonable}: the weight difference
\((w^{\text{global}} - w(s))\) is constructed precisely from differences in \(\hat{R}_m(s)\),
so it is structurally correlated with \((\ell - \hat{R}_m(s))\),
and its expectation is not necessarily zero.

\textbf{Correct proof} (using the standard form of Gibbs inequality):

By the KKT conditions of Gibbs inequality, \(w(s)\) minimizes
\(J_s(w) = \sum_m w_m \hat{R}_m(s) + \frac{1}{\alpha} \sum_m w_m \log w_m\). Therefore for any weight vector
\(w'\) (including \(w^{\text{global}}\)):

\[\sum_m w_m(s) \hat{R}_m(s) + \frac{1}{\alpha} H(w(s)) \leq \sum_m w_m^{\text{global}} \hat{R}_m(s) + \frac{1}{\alpha} H(w^{\text{global}})\]

Rearranging and taking expectation over \(s\), we obtain:

\[\mathbb{E}_s\!\left[\sum_m (w_m^{\text{global}} - w_m(s)) \hat{R}_m(s)\right] \geq \frac{1}{\alpha} \mathbb{E}_s[I(s; w)]\]

where \(I(s; w) = \mathbb{E}_s[\text{KL}(w(s) \| w^{\text{global}})]\)
is the state-weight mutual information. This inequality gives a lower bound directly from Gibbs inequality---without assuming zero-mean noise.

\textbf{However, we cannot claim} that this implies the actual expected risk difference
\(\mathbb{E}[\sum_m (w_m^{\text{global}} - w_m(s)) \ell(f_m, Y)]\) is lower-bounded by
\(I(s;w)/\alpha\). This is because the gap term between \(\ell\) and \(\hat{R}\)
\(\mathbb{E}[\sum_m (w_m^{\text{global}} - w_m(s))(\ell - \hat{R}_m(s))]\)
may be negative, potentially shrinking or even reversing the inequality.

\textbf{Correct formulation of Theorem 3.1 (revised)}: If \(\ell\) is bounded in \([0, L_{\max}]\)
and the empirical risk estimates \(\hat{R}_m(s)\) satisfy
\(\max_{m,s} |\hat{R}_m(s) - R_m(s)| \leq \varepsilon_R\) (uniform convergence), then:

\[\mathbb{E}\!\left[\sum_m w_m(s) \ell(f_m, Y)\right] \leq \mathbb{E}\!\left[\sum_m w_m^{\text{global}} \ell(f_m, Y)\right] - \frac{1}{\alpha} I(s; w) + 2\varepsilon_R\]

This upper bound (not lower bound) shows that the advantage of state-conditioned weights is related to
\(I(s;w)/\alpha\) when empirical risk estimates are accurate (\(\varepsilon_R \to 0\)).
The penalty term \(2\varepsilon_R\) is the finite-sample cost.

\textbf{Rigorous proof} (see \texttt{\_proof.md} Theorem 2). \(\square\)

This bound shows that the advantage of state-conditioned weights is positively correlated with the state-weight mutual information. In the asymptotic limit (\(n_s \to \infty\),
\(\varepsilon_R \to 0\)), \(I(s; w) > 0\) implies strict superiority.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3 Connection to MoE Gating}\label{connection-to-moe-gating}

\subsubsection{3.1 Comparison with Shazeer et al.~(2017) Sparse-Gated MoE}\label{comparison-moe}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Dimension & Traditional MoE & SCX Expert Routing \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Gating input & Raw features \(x\) & State assignment \(\gamma_s(x)\) \\
Gating parameters & Learnable \(W_g\) & Estimated \(\hat{R}_m(s)\) \\
Training method & End-to-end backpropagation & Two-stage (estimate risk first, then route) \\
Sparsity & hard/top-k & Exponential soft weighting \\
Generalization guarantee & Empirical risk minimization & State-conditioned PAC bounds \\
\end{longtable}

SCX's routing is based not on \(x\) itself, but on \textbf{state-conditioned risk}
\(\sum_s \gamma_s(x) R_m(s)\). This is a \textbf{meta-learning variant} of MoE
gating: the gating decision does not directly learn
a mapping \(x \to \text{expert}\), but indirectly determines routing by estimating conditional risk.

\subsubsection{3.2 Load Balancing}\label{load-balancing}

Traditional MoE requires auxiliary loss to balance expert usage:

\[\mathcal{L}_{\text{balance}} = \alpha \cdot CV(\text{expert\_load})^2\]

SCX's state-conditioned weights
\(w_m(x) \propto \exp(-\alpha \sum_s \gamma_s(s) \hat{R}_m(s))\)
naturally achieve load balancing: if an expert performs poorly across all states, its weight decays exponentially; different experts win in different states, automatically distributing load.

\subsubsection{3.3 Relationship to Jacob et al.~(1991) Adaptive Mixture of Experts}\label{relation-jacob}

SCX's expert routing \(m^*(x) = \arg\min_m \sum_s \gamma_s(x) R_m(s)\)
can be seen as a hard-routing version of Jacob's gating network (hard routing vs.~soft routing).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4 Extreme Cases}\label{extreme-cases}

\subsubsection{4.1 Degeneration to Global Weights}\label{degeneration-global}

When
\(\mathcal{S} = \{\text{all}\}\) (only one state), \(\gamma_s(x) \equiv 1\), \(w_m(x) = \frac{\exp(-\alpha \hat{R}_m)}{\sum_{m'} \exp(-\alpha \hat{R}_{m'})}\), at this point
\(w_m(x)\) is constant, and the proposition degenerates to the trivial case \(\Delta R = 0\).

\subsubsection{4.2 Achieving the Theoretical Upper Bound}\label{achieving-upper-bound}

When each \(x\)
forms its own independent state (\(\mathcal{S} = \mathcal{X}\)), \(\gamma_s(x) = \delta_{x,s}\), \(w_m(x) = \frac{\exp(-\alpha \hat{R}_m(x))}{\sum_{m'} \exp(-\alpha \hat{R}_{m'}(x))}\), at this point the theoretical upper bound is achieved, and \(\Delta R\)
is maximized.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5 Implications for SCX}\label{implications-for-scx}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Unifies the weight formulas for expert routing, ensemble, and distillation}: SCX's
  weight formula
  \(w_m(x) \propto \exp(-\alpha \sum_s \gamma_s(x) \hat{R}_m(s))\)
  provides a unified mathematical framework.
\item
  \textbf{PAC-Bayes interpretation}: SCX's
  state-conditioned weights can be viewed as a \textbf{state-dependent posterior}
  \(\mathcal{Q}_x\), whose KL divergence penalty relative to a uniform prior is controlled by the \(\alpha\)
  parameter, achieving state-level PAC-Bayes bounds:
  \[\mathbb{E}_{h \sim \mathcal{Q}_s}[R(h|s)] \leq \mathbb{E}_{h \sim \mathcal{Q}_s}[\hat{R}(h|s)] + \sqrt{\frac{\text{KL}(\mathcal{Q}_s \| \mathcal{P}_s) + \log(n_s/\delta)}{2n_s - 1}}\]
\item
  \textbf{Relationship to the Information Bottleneck}: SCX's
  weights are a variational approximation of expert selection uncertainty, interpretable as compressing expert selection information at the state level while maximizing prediction fidelity.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references-prop3}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  SCX Core Framework Mathematical Analysis. \texttt{01\_SCX\_Core\_Framework\_Math\_Analysis.md} Section
  2.3
\item
  SCX Mathematical Foundations and Proof Construction. \texttt{05\_Mathematical\_Roots\_and\_Proofs.md} Proposition 3
\item
  Jacobs, R. A., et al.~(1991). Adaptive mixtures of local experts.
  \emph{Neural Computation}.
\item
  Shazeer, N., et al.~(2017). Outrageously large neural networks: The
  sparsely-gated mixture-of-experts layer. \emph{ICLR}.
\item
  McAllester, D. A. (1999). PAC-Bayesian model averaging. \emph{COLT}.
\end{enumerate}

\end{document}
