pdfoutput=1
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\pdfoutput=1
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{A2 (Conditional Independence) --- Vulnerability Analysis and Fix}\label{a2-conditional-independence-vulnerability-analysis-and-fix}

\begin{quote}
\textbf{Trigger}: Hostile review pointed out that A2 is ``almost certainly false, the Achilles' heel of the entire framework.'' \textbf{Conclusion}: Diagnosis correct, but fixable. Replace with weak-correlation concentration inequality, degrade the decay rate.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Problem Diagnosis}\label{problem-diagnosis}

Theorem 1's proof chain:

\begin{verbatim}
A2 (conditional independence) → e_m independent → Hoeffding concentration → exp(-2MΔ²) exponential rate
\end{verbatim}

If expert errors have conditional correlation \(\rho > 0\) (given state \(s\)): - CNN: all experts err together on blurry images - MLIP: all experts err together on transition-state configurations - Effective independent expert count: \(M_{\text{eff}} \approx M / (1 + (M-1)\rho)\)

At \(\rho = 0.3\), \(M=12\) degrades to \(M_{\text{eff}} = 2.8\) --- the exponential gain nearly vanishes.

\textbf{CIFAR-10 experimental evidence}: Theoretical lower bound F1≥0.976 (assumed independence), measured F1=0.617. The gap cannot be explained by \(\mu_s\) estimation error alone.

\textbf{AlN experimental evidence}: Global bound F1≥0.87 matches empirical value. But this only holds when \(\rho \lesssim 0.02\) --- an extremely weak assumption that requires verification.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Fix Strategy}\label{fix-strategy}

\subsubsection{2.1 Weak-Correlation Hoeffding Bound}\label{weak-correlation-hoeffding-bound}

Let \(\rho_{ij}\) be the error correlation between experts \(i\) and \(j\). With average correlation \(\bar{\rho} = \frac{1}{M(M-1)}\sum_{i\neq j}\rho_{ij}\):

\begin{equation}
\text{Var}\left(\frac{1}{M}\sum_{i=1}^M e_i\right) = \frac{\sigma^2}{M}\left[1 + (M-1)\bar{\rho}\right]
\end{equation}

This is an exact formula for the variance under correlation. The variance inflation factor is \(1 + (M-1)\bar{\rho}\).

\textbf{Crucial distinction}: This variance inflation is exact for the \textbf{variance}, but plugging it into Hoeffding gives a \textbf{heuristic} bound, not a rigorous theorem. See \texttt{a2\_rigorous\_analysis.tex} for the rigorous treatment.

\textbf{Heuristic fix}: Replace \(M\) with \(M_{\text{eff}} = M / (1 + (M-1)\bar{\rho})\) in the Hoeffding bound:

\begin{equation}
F_1 \geq 1 - \frac{1}{\eta}\exp\left(-\frac{2M\Delta^2}{1 + (M-1)\bar{\rho}}\right)
\end{equation}

\textbf{Limitation of the heuristic}: The MGF factorization required by Hoeffding only holds under independence. For correlated variables, this is an approximation widely used in survey sampling but not a theorem.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. How to Estimate \(\bar{\rho}\)}\label{how-to-estimate-rho}

\subsubsection{3.1 Paired error correlation on training set}\label{paired-error-correlation}

For each pair of experts \((i, j)\), compute their error correlation on the training set:

\begin{equation}
\hat{\rho}_{ij} = \text{Corr}(\mathbf{1}\{f_i(x) \neq y\}, \mathbf{1}\{f_j(x) \neq y\})
\end{equation}

Then average: \(\hat{\bar{\rho}} = \frac{1}{M(M-1)}\sum_{i\neq j}\hat{\rho}_{ij}\).

\subsubsection{3.2 Bootstrap confidence intervals}\label{bootstrap-confidence-intervals}

Use bootstrap resampling of training samples to obtain confidence intervals for \(\bar{\rho}\). The effective \(M_{\text{eff}}\) can be reported with uncertainty: \(M_{\text{eff}} \in [M_{\text{eff}}^{\text{low}}, M_{\text{eff}}^{\text{high}}]\) at 95\% confidence.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. When is A2 ``Good Enough''?}\label{when-is-a2-good-enough}

A2 does not need to hold perfectly. It needs to hold well enough that \(M_{\text{eff}}\) remains large. The practical threshold:

\begin{itemize}
\item If \(\bar{\rho} < 0.1\): \(M=10\) gives \(M_{\text{eff}} \approx 5.3\) — still exponential gain
\item If \(\bar{\rho} < 0.05\): \(M=10\) gives \(M_{\text{eff}} \approx 6.9\) — strong gain
\item If \(\bar{\rho} > 0.3\): \(M=10\) gives \(M_{\text{eff}} \approx 2.7\) — gain marginal
\end{itemize}

\textbf{The honest question}: What is the typical \(\bar{\rho}\) in real SCX deployments? For heterogeneous experts (CNN + RandomForest + XGBoost), \(\bar{\rho}\) is likely low (<0.2). For homogeneous experts (all CNNs with different initializations), \(\bar{\rho}\) may be high (>0.4). SCX must report \(\bar{\rho}\) estimates alongside \(M_{\text{eff}}\).

\subsection{5. Relationship to the Rigorous Analysis}\label{relationship-to-rigorous}

This document presents the \textbf{heuristic fix} (variance inflation plugged into Hoeffding). \texttt{a2\_rigorous\_analysis.tex} provides the \textbf{rigorous treatment}:

1. The Hoeffding bound with \(M_{\text{eff}}\) is \textbf{not} a theorem for correlated variables.
2. The relationship \(M_{\text{eff}} = M/(1+(M-1)\bar{\rho})\) is exact for the variance but not for the tail probability.
3. Rigorous alternatives: Azuma-Hoeffding for martingale difference sequences, or Bernstein inequalities for weak dependence.

\end{document}
