pdfoutput=1
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\pdfoutput=1
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{Rigorous Mathematical Analysis of A2}\label{rigorous-mathematical-analysis-of-a2}

\begin{quote}
Trigger: hostile review pointed out that \(\exp(-2M\Delta^2/(1+(M-1)\rho))\) is not a rigorous theorem. Conclusion: correct. Variance inflation plugged into Hoeffding is heuristic, not a rigorous bound. But A2 itself has stronger arguments.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Is \(\exp(-2M t^2/(1+(M-1)\rho))\) a rigorous theorem?}\label{is-exp-a-rigorous-theorem}

\textbf{No.}

The proof of Hoeffding's inequality relies on:
\[\mathbb{E}\left[\exp\left(\lambda\sum_{i=1}^M (X_i - \mu_i)\right)\right] = \prod_{i=1}^M \mathbb{E}[\exp(\lambda(X_i - \mu_i))]\]

This \textbf{factorization} requires \(\{X_i\}\) to be independent. If \(X_i\) are correlated, the MGF cannot be factorized, and you cannot simply replace \(M\) with \(M/(1+(M-1)\rho)\) to obtain a rigorous exponential bound.

The variance inflation \((1+(M-1)\rho)\) is correct for \(\text{Var}(\sum X_i)\), but not for \textbf{exponential concentration}. Variance only gives Chebyshev bound (polynomial rate \(1/t^2\)), not exponential.

\textbf{Conclusion}: Variance-inflated Hoeffding is a heuristic, widely used in survey sampling and cluster randomized trials, but it is not a mathematical theorem. Including it in SI would be caught by reviewers.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. So does A2 itself hold up?}\label{so-does-a2-itself-hold-up}

\textbf{Yes, but with nuance.} A2 (conditional independence) is an \textbf{idealization}, not an empirical claim. The actual question is: how much does the real-world deviation from A2 degrade the bound?

\subsubsection{2.1 When is A2 plausible?}

A2 holds approximately when:
\begin{enumerate}
\item Experts are \textbf{architecturally heterogeneous}: CNN + RandomForest + XGBoost share no common inductive bias. Their errors on the same sample are close to conditionally independent.
\item Experts are trained on \textbf{disjoint data subsets} (bagging). Each expert sees a different bootstrap sample, reducing error correlation.
\item Experts use \textbf{different feature subsets}: each expert sees a different projection of the data, breaking the shared-information channel.
\end{enumerate}

\subsubsection{2.2 When does A2 fail badly?}

A2 fails when:
\begin{enumerate}
\item \textbf{Homogeneous experts}: all CNNs with different random seeds — their errors are highly correlated on the same ``hard'' samples.
\item \textbf{Shared training data}: all experts trained on the same dataset — correlation from shared noise in the training labels.
\item \textbf{Common methodological bias}: experts trained at the same institution, using the same techniques — systematic error correlation.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Rigorous Alternatives}\label{rigorous-alternatives}

\subsubsection{3.1 Azuma-Hoeffding for Martingale Differences}

If expert errors can be ordered such that each expert adds ``new information'' not fully captured by previous experts, we can use Azuma-Hoeffding:

\begin{equation}
\mathbb{P}\left(\left|\sum_{i=1}^M D_i\right| \geq t\right) \leq 2\exp\left(-\frac{t^2}{2\sum_{i=1}^M c_i^2}\right)
\end{equation}

where \(D_i\) are martingale differences: \(\mathbb{E}[D_i \mid D_1,\ldots,D_{i-1}] = 0\) and \(|D_i| \leq c_i\).

This works for sequentially dependent but conditionally unbiased experts. The price: the bound depends on the ordering of experts, and \(c_i\) must be estimated.

\subsubsection{3.2 Bernstein Inequality for Weak Dependence}

For \(\alpha\)-mixing sequences with mixing coefficient \(\alpha(k) \leq C\rho^k\):

\begin{equation}
\mathbb{P}\left(\frac{1}{M}\sum_{i=1}^M X_i \geq t\right) \leq \exp\left(-\frac{M t^2}{2(\sigma^2 + c t)}\right) \cdot (1 + 4C\alpha^{-1})
\end{equation}

The effective sample size is reduced by the mixing coefficient, but the exponential rate is preserved (with a modified constant).

\subsubsection{3.3 Recommendation for the Paper}

\textbf{Keep A2 as an idealization} but:
\begin{enumerate}
\item State explicitly: ``A2 is assumed for Theorem 1. Under violation of A2, the following weaker bound holds...''
\item Provide the variance-inflation formula as a \textbf{diagnostic tool}, not a theorem.
\item Include the martingale-difference alternative as a supplementary result.
\item Report empirical \(\bar{\rho}\) alongside theoretical bounds.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. The Strongest Defense of A2}\label{strongest-defense-of-a2}

The strongest argument for A2 is not statistical — it's \textbf{operational}. If experts are correlated, SCX intentionally designs the expert ensemble to minimize correlation:

\begin{itemize}
\item Architectural diversity (different model families)
\item Data diversity (different training subsets)
\item Feature diversity (different input representations)
\end{itemize}

\textbf{SCX doesn't assume independence — SCX engineers independence.} The framework includes explicit mechanisms (expert routing, state crystallization) to detect and mitigate correlation. If \(\bar{\rho}\) is high, SCX knows it's high (from diagnostic tests) and the user can add more diverse experts.

\textbf{This transforms A2 from an assumption into a design target.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Summary: How to Handle A2 in the Paper}\label{summary-how-to-handle-a2}

\begin{enumerate}
\item \textbf{Theorem 1}: State ``under A2 (conditional independence)'' — reviewers accept idealizations.
\item \textbf{SI/Appendix}: Provide (a) variance-inflation formula with honest ``heuristic'' label; (b) martingale-difference alternative; (c) empirical \(\bar{\rho}\) estimation method.
\item \textbf{Experiments}: Report \(\bar{\rho}\) alongside \(M\) and \(M_{\text{eff}}\).
\item \textbf{Discussion}: Acknowledge that real-world correlation degrades the exponential rate, but SCX's architecture minimizes this degradation.
\end{enumerate}

\end{document}
