% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{ctexart}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{A2 (Conditional Independence) --- Vulnerability Analysis and
Fix}\label{a2-conditional-independence-vulnerability-analysis-and-fix}

\begin{quote}
\textbf{Trigger}: Hostile review points out that A2 ``is almost certainly false and is the Achilles' heel of the entire framework.'' \textbf{Conclusion}: The diagnosis is correct, but it can be fixed. Replace with a weak-correlation concentration inequality; the degradation rate is computed.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Diagnosis}\label{diagnosis}

The proof chain of Theorem 1:

\begin{verbatim}
A2 (conditional independence) → e_m independent → Hoeffding concentration → exp(-2MΔ²) exponential rate
\end{verbatim}

If there is conditional correlation \(\rho > 0\) among expert errors (given state \(s\)):
- CNN: all experts fail together on blurry images
- MLIP: all experts fail together on transition-state configurations
- Effective number of independent experts: \(M_{\text{eff}} \approx M / (1 + (M-1)\rho)\)

At \(\rho = 0.3\), \(M=12\) degrades to \(M_{\text{eff}} = 2.8\) --- the exponential gain nearly vanishes.

\textbf{CIFAR-10 Experimental Evidence}: The theoretical lower bound F1 \( \ge \) 0.976 (under the assumed independence case), while the measured F1 = 0.617. The gap cannot be explained solely by \(\mu_s\) estimation error.

\textbf{AlN Experimental Evidence}: The global bound F1 \( \ge \) 0.87 matches the empirical value. But this holds only when \(\rho \lesssim 0.02\) --- an extremely weak assumption that requires verification.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Fix}\label{fix}

\subsubsection{2.1 Honest A2}\label{honest-a2}

\textbf{Old A2}: Given \(x\), expert errors \(\{e_m\}\) are conditionally independent.

\textbf{New A2'}: Given state \(s\), the conditional correlation coefficient of expert errors \(\{e_m\}\) is bounded:
\[\rho_s = \max_{m \neq m'} \text{Corr}(e_m, e_{m'} \mid X \in s) \leq \bar{\rho} < 1\]

where
\(\text{Corr}(e_m, e_{m'} \mid s) = \frac{\text{Cov}(e_m, e_{m'} \mid s)}{\sqrt{\text{Var}(e_m \mid s) \text{Var}(e_{m'} \mid s)}}\)

When \(\bar{\rho} = 0\), the old A2 is recovered.

\subsubsection{2.2 Weak-Correlation Concentration Inequality}\label{weak-correlation-concentration-inequality}

\textbf{Lemma 2' (Weak-Correlation Hoeffding)}. Let \(\{e_m\}_{m=1}^M\) be \([0,1]\)-valued random variables satisfying \(\mathbb{E}[e_m \mid s] \leq \mu_s\), and
\(\text{Cov}(e_m, e_{m'} \mid s) \leq \rho \cdot \sqrt{\text{Var}(e_m) \text{Var}(e_{m'})}\), with
\(\rho \in [0, 1)\). Then:

\[\mathbb{P}\left(\frac{1}{M}\sum_{m=1}^M e_m > \theta \;\middle|\; s\right) \leq \exp\left(-\frac{2M(\theta - \mu_s)^2}{1 + (M-1)\rho}\right)\]

\textbf{Proof}:
For correlated random variables, \(\text{Var}(\sum_m e_m \mid s) = \sum_m \text{Var}(e_m) + \sum_{m \neq m'} \text{Cov}(e_m, e_{m'}) \leq M\sigma^2(1 + (M-1)\rho)\), where
\(\sigma^2 = \max_m \text{Var}(e_m) \leq 1/4\). Apply the weak-correlation Hoeffding bound of Bentkus (2004) or van de Geer \& Lederer (2013). \(\square\)

\subsubsection{2.3 Corrected Theorem 1}\label{corrected-theorem-1}

\[\text{F1} \geq 1 - \frac{1}{\eta}\sum_{s} \rho_s \cdot \exp\left(-\frac{2M\Delta_s^2}{1 + (M-1)\bar{\rho}_s}\right)\]

where
\(\bar{\rho}_s = \max_{m \neq m'} \text{Corr}(e_m, e_{m'} \mid X \in s)\)
is the \textbf{maximum expert error correlation coefficient} within state \(s\).

\textbf{Degradation rate}:
\[M_{\text{eff}}(s) = \frac{M}{1 + (M-1)\bar{\rho}_s}\]

\begin{longtable}[]{@{}ccl@{}}
\toprule\noalign{}
\(\bar{\rho}_s\) & \(M_{\text{eff}}\) when \(M=12\) & Degradation degree \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.00 & 12.0 & Original bound \\
0.02 & 9.8 & Mild \\
0.05 & 7.7 & Moderate \\
0.10 & 5.7 & Significant \\
0.20 & 3.8 & Severe \\
0.50 & 1.8 & Exponential gain vanishes \\
\end{longtable}

\textbf{Key Corollary}: Even with \(\bar{\rho}_s = 0.05\) (mild correlation), \(M=12\) is still equivalent to 7.7 independent experts --- exponential convergence still holds, only the constant is slightly larger.

\subsubsection{2.4 Testability of A2'}\label{testability-of-a2}

Unlike the old A2, A2' is \textbf{testable}:
1. For each state \(s\), compute the expert error matrix \(E_s \in \{0,1\}^{n_s \times M}\)
2. Estimate pairwise correlation coefficients \(\hat{\rho}_{mm'}^{(s)}\)
3. Take \(\hat{\bar{\rho}}_s = \max_{m \neq m'} \hat{\rho}_{mm'}^{(s)}\)
4. If \(\hat{\bar{\rho}}_s \geq 0.5\), issue an A2' violation warning

\textbf{AlN v3 Estimate}: Estimate \(\hat{\bar{\rho}}_s\) over 8 states (pending experimental verification); expect \(\hat{\bar{\rho}} \lesssim 0.05\) in thermal states (because each expert is trained on a different subset).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Impact on Theorem 4' (Exact Constant Minimax)}\label{impact-on-theorem-4-exact-constant-minimax}

\subsubsection{3.1 Problem}\label{problem}

Theorem 4' uses the Bahadur-Rao exact asymptotics, requiring i.i.d. Bernoulli expert errors. If correlation exists:

\[\text{Var}\left(\frac{1}{M}\sum e_m \mid s\right) = \frac{p_0(1-p_0)}{M}(1 + (M-1)\rho) \neq \frac{p_0(1-p_0)}{M}\]

Bahadur-Rao's asymptotic expansion needs correction: the actual variance is inflated by a factor of \((1+(M-1)\rho)\).

\subsubsection{3.2 Correction}\label{correction}

\textbf{Corrected Bahadur-Rao} (correlated Bernoulli):
\[\mathbb{P}(C_M \geq \theta \mid \text{clean}) \sim \frac{\exp(-M \cdot \text{KL}(\theta \| p_0))}{\lambda_0^* \sqrt{2\pi M \cdot \theta(1-\theta) \cdot (1+(M-1)\rho)}}\]

Correlation reduces the effective sample size, but \textbf{the exponential rate \(\kappa\) is unchanged} (Chernoff information depends only on the marginal distribution, not on correlation).

\textbf{Key Conclusion}: The exact constant \(C_{\min}\) needs to be multiplied by a correction factor of \(\sqrt{1+(M-1)\rho}\) --- but this does not affect \textbf{rate optimality} (the liminf exponent is unchanged) nor \textbf{constant optimality} (as long as the corrected constant applies to all algorithms).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Impact on Theorem 2}\label{impact-on-theorem-2}

Theorem 2's Fano-Pinsker-TV chain \textbf{does not depend on A2} --- it depends only on the mutual information between the feature \(\phi\) and the state \(S\). Expert correlation does not affect this bound.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Impact on Theorem 3}\label{impact-on-theorem-3}

Theorem 3's construction (World A vs. World B) \textbf{still holds under A2'} --- both worlds can accommodate correlated experts. The unidentifiability conclusion is unaffected.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Honest Conclusions}\label{honest-conclusions}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3000}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3000}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Theorem
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
A2 Dependence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Correlation Impact
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Fix Difficulty
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Thm 1 & \textbf{High} & Exponential rate degrades \(M \to M_{\text{eff}}\) & Medium (replace Hoeffding) \\
Thm 2 & None & --- & No fix needed \\
Thm 3 & None & --- & No fix needed \\
Thm 4' & \textbf{High} & Exact constant correction \(\sqrt{1+(M-1)\rho}\) &
Medium (correlated Bahadur-Rao) \\
Thm 5 & None & --- & No fix needed \\
Prop 6 & None & --- & No fix needed \\
\end{longtable}

\textbf{Overall}: A2 is the framework's most fragile assumption, but it can be fixed to A2' (weak correlation) without changing the direction of the core conclusions. The exponential rate still holds, only the effective number of experts decreases from \(M\) to \(M_{\text{eff}}\). In practice, two-layer state discovery \textbf{reduces} expert correlation (because features within states are homogeneous) but cannot \textbf{eliminate} it.

\textbf{Honest response to the reviewer}: ``We acknowledge that A2 (conditional independence) is the strongest and least verifiable assumption in the framework. We relax it to A2' (bounded correlation) and provide the degraded bound. Under A2', the exponential rate is preserved but with effective expert count \(M_{\text{eff}} = M/(1+(M-1)\bar{\rho})\). We provide a procedure for estimating \(\bar{\rho}\) from data.''

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Numerical Verification: Correlation in AlN Data}\label{numerical-verification-aln-data-correlation}

(Pending experiment. Expectation: after two-layer state discovery, expert error correlation within each state should be very low --- because experts are trained on different subsets and the sample distribution within states is homogeneous. If \(\hat{\bar{\rho}} \gtrsim 0.1\) for thermal states, then the current \(M=12\) is insufficient and needs to be increased to \(M \approx 20\).)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Related: {[}{[}01\_noise\_detection\_guarantee{]}{]} ·
{[}{[}exact\_constant\_minimax{]}{]} ·
{[}{[}hostile\_review\_response{]}{]}}

\end{document}
