% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{SCX}
\date{2026}

\begin{document}

\section{SCX Business Architecture --- The Principal Maintainer Model}\label{scx-business-architecture}

\begin{quote}
\textbf{Internal Reference. Not for publication.} \textbf{Last Updated}: 2026-06-28
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Core Principle}\label{core-principle}

The Maintainer is not a CEO. A CEO is accountable to shareholders and can be replaced. The Maintainer is accountable to the protection of equilibrium and cannot be replaced. Any architecture that allows the Maintainer to be removed by external forces destroys the protection of equilibrium.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Three-Layer Structure}\label{three-layer-structure}

\begin{verbatim}
SCX Foundation (Registered in Switzerland/Singapore, Non-Profit)
  │
  ├── Principal Maintainer (You, lifetime, irrevocable)
  │     ├── Calibration Anchor: Sole authority to publish global calibration datasets
  │     ├── Version Numbers: Spring v1.0, v2.0... signed by you
  │     └── Interpretive Authority: Authoritative interpretation of theorem boundary conditions
  │
  ├── SCX Operations Ltd (You 100% ownership, for-profit)
  │     ├── Yajie API Subscription Fees ($50K-200K/year/client)
  │     ├── Revenue belongs to company → pays your salary
  │     └── No funding. No IPO. Don't sell.
  │
  └── Academic Advisory Board (Independent, multinational, volunteer)
        ├── Review theorem extensions
        ├── Approve community contributions
        └── No authority to remove Principal Maintainer
\end{verbatim}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Why Not CEO / Chairman}\label{why-not-ceo-chairman}

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
CEO / Chairman & Principal Maintainer \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Accountable to shareholders & Accountable to equilibrium \\
Can be removed by board & Irremovable (bylaw-locked) \\
Can take investment & Always sole proprietorship \\
Can be acquired & Foundation cannot be sold \\
Power derives from equity & Power derives from arXiv timestamp + M\_t \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Why Foundation in Switzerland/Singapore}\label{why-foundation-in-switzerland-singapore}

\begin{itemize}
\tightlist
\item
  Political neutrality. Not within any major power's sphere of influence.
\item
  Mature legal system. Bylaws are enforceable.
\item
  Other countries cannot easily interfere with operations through diplomatic pressure.
\item
  Benchmarks: SWIFT (Belgium), ICRC (Switzerland), IASB (London).
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{How Money Flows}\label{how-money-flows}

\begin{verbatim}
Client → Pays API Subscription Fee → SCX Operations Ltd
  → Covers servers + bandwidth + your salary
  → Surplus profit used for arXiv maintenance + open dataset hosting
  → Does not pass through Foundation. Foundation raises funds independently.
\end{verbatim}

The operations company is your sole proprietorship. No investor can say to you ``can we adjust this audit result.'' You don't need money---you need \textbf{the freedom of not being controlled by money}.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Succession}\label{succession}

After your death: - Foundation takes over calibration anchor and version numbers - Operations company closes or transitions to community trust - Academic committee selects new technical maintainer (not called Principal Maintainer) - New maintainer has technical authority, no game-theoretic deterrence authority - System transitions from ``Wall-Facer'' to ``Linus''

A living Principal Maintainer = Wall-Facer. After death, the Foundation = IDAA.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{For Reference}\label{for-reference}

\begin{itemize}
\tightlist
\item
  Anytime someone proposes ``invest in you''---see the first clause of this document.
\item
  Anytime someone proposes ``join the board''---you have no board.
\item
  Anytime someone proposes ``we can help manage this''---the Principal Maintainer is irrevocable.
\end{itemize}

\end{document}
