\pdfoutput=1
% ══════════════════════════════════════════════════════════════════════
% Viewpoint 4 Corrigendum: Cercis-Fisher is Conditional Equivalence, NOT Natural Emergence
%
% This is a standalone corrigendum to be inserted after Section 4 (Information-Geometric
% Bulk-Boundary Correspondence) of gauge_formalized.tex, or read alongside it.
%
% Based on formal peer review (docs/reviews/GAUGE_5VIEWPOINTS_FINAL.md,
% docs/reviews/GAUGE_VIEWPOINTS_REVIEW.md).
% ══════════════════════════════════════════════════════════════════════

\documentclass{article}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage[margin=2.5cm]{geometry}

% ─── Math notation (consistent with gauge_formalized.tex) ───
\newcommand{\GeoDist}{\mathrm{GeoDist}}
\newcommand{\KL}{\mathrm{KL}}
\newcommand{\Fisher}{\mathcal{I}}
\newcommand{\cercis}{\mathcal{C}}
\newcommand{\cercisnorm}{\bar{\cercis}}
\newcommand{\Situs}{\mathcal{S}}
\newcommand{\ExpFam}{\mathcal{E}}
\newcommand{\Od}{\mathrm{O}(d)}
\newcommand{\SOthree}{\mathrm{SO}(3)}
\newcommand{\ZBetti}{\beta_1}
\newcommand{\Lone}{L_1}
\newcommand{\edgs}{\mathcal{E}}
\newcommand{\verts}{\mathcal{V}}
\newcommand{\face}{\mathcal{F}}
\newcommand{\grph}{\mathcal{G}}
\newcommand{\Gflatquot}{\mathcal{M}_{\mathrm{flat}}}
\newcommand{\triv}{\mathbf{1}}
\newcommand{\doneT}{d_1^{\mathsf{T}}}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Log}{Log}
\DeclareMathOperator{\Ad}{Ad}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Stab}{Stab}
\newcommand{\E}{\mathbb{E}}
\DeclareMathOperator{\dist}{dist}

\title{\textbf{Viewpoint 4 Corrigendum} \\
\large Cercis-Fisher Equivalence: Conditional Equivalence, Not Natural Emergence}
\author{SCX}
\date{\today}

\begin{document}
\maketitle

% ══════════════════════════════════════════════════════════════════════
\section{Review Verdict}
% ══════════════════════════════════════════════════════════════════════

\begin{center}
\begin{tabular}{|p{6cm}|p{8cm}|}
\hline
\textbf{Original Claim} & \textbf{Verdict} \\
\hline
Cercis${}^2 \propto$ average KL divergence $=$ squared Fisher geodesic length.
``Naturally emerges'' from information geometry.
& ⚠️ Mathematically correct but ``natural emergence'' is overstated \\
\hline
\end{tabular}
\end{center}

\subsection{Specific Issues}

\begin{enumerate}[label=(\arabic*)]
    \item \textbf{Not ``average KL'' but ``minimum KL''.}
    Cercis is the \emph{minimum} KL divergence to the pure-gauge submanifold:
    $\cercis \approx 2 \cdot \min_{h} \KL(P_A \| P_{d_0 h})$,
    not an ``average KL divergence.''

    \item \textbf{Not ``natural emergence'' but conditional equivalence.}
    The $\GeoDist^2 \leftrightarrow 2\KL$ equivalence depends on 8 strong assumptions (see \S\ref{sec:assumptions}).
    If any single assumption fails, the equivalence breaks. SCX data is extremely unlikely to form an exponential family.

    \item \textbf{Complete breakdown outside exponential families.}
    For general statistical manifolds, the difference between $\GeoDist^2$ and $2\KL$ is controlled
    by the Amari-Chentsov third-order tensor. This tensor is generically non-zero for SCX data,
    so the equivalence degrades.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{The Eight Assumptions}
\label{sec:assumptions}
% ══════════════════════════════════════════════════════════════════════

For $\GeoDist(p,q)^2 = 2\KL(p\|q) + O(\|p-q\|^3)$ to hold rigorously,
the following eight assumptions must be simultaneously satisfied.
Failure of any single one degrades the equality.

\vspace{1em}

\begin{enumerate}[label=\textbf{H\arabic*:}, leftmargin=*, itemsep=0.8em]

    \item \textbf{Exponential Family Assumption.} \\
    The output distribution family $\mathcal{P} = \{p(x \mid \theta)\}$ must be an exponential family,
    i.e., there exists a sufficient statistic $T(x)$ and cumulant generating function $\psi(\theta)$ such that
    $p(x \mid \theta) = h(x) \exp(\langle\theta, T(x)\rangle - \psi(\theta))$.
    Only under this condition is the Fisher information matrix the Hessian of $\psi$,
    and KL divergence equals Bregman divergence (Lemma~4.3 / Lemma~\ref{lem:kl_bregman} in the original context).
    \\[0.3em]
    \textbf{Consequence of failure:} Outside exponential families, $\KL(\theta_1\|\theta_2)$ is no longer a Bregman divergence of $\psi$,
    and the second-order term of the Taylor expansion is no longer $g_{ij}\Delta\theta^i\Delta\theta^j$.
    The difference between $\GeoDist^2$ and $2\KL$ is controlled by the Amari-Chentsov third-order tensor (Open Problem~6.2 / Open Problem~\ref{prob:non_exp_fam}).

    \item \textbf{Small-Distance / Local Approximation Assumption.} \\
    $\theta_p$ and $\theta_q$ must be sufficiently close so that third-order
    and higher remainders $O(\|\theta_q - \theta_p\|^3)$ are negligible. All derivations
    truncate the Taylor expansion at second order.
    \\[0.3em]
    \textbf{Consequence of failure:} For large discrepancies (common in cross-domain SCX experts),
    third-order error terms dominate and $\GeoDist^2 \approx 2\KL$ no longer holds.

    \item \textbf{$C^2$ Differentiability Assumption.} \\
    The map $\theta \mapsto p(x \mid \theta)$ must be $C^2$ for almost all $x$.
    This is required for second-order Taylor expansion and for the Fisher matrix as a second derivative.
    \\[0.3em]
    \textbf{Consequence of failure:} If the density has non-differentiable points (e.g., piecewise
    distributions from ReLU activations), the Fisher metric is undefined at those points and
    geodesics are not uniquely determined.

    \item \textbf{Integration-Differentiation Interchangeability Assumption.} \\
    The interchange $\partial_\theta \int p(x|\theta) d\mu(x) = \int \partial_\theta p(x|\theta) d\mu(x)$
    must hold. This is the standard regularity condition ensuring $\E[\partial_\theta \log p] = 0$, which
    underpins the Fisher information definition.
    \\[0.3em]
    \textbf{Consequence of failure:} If the support depends on $\theta$ (e.g., uniform distributions),
    the standard Fisher information form breaks down.

    \item \textbf{Fisher Full-Rank Assumption.} \\
    The Fisher information matrix $\Fisher(\theta)$ must be positive-definite
    (full rank) at every point. This is required for $(\Theta, g)$ to be a genuine Riemannian
    manifold — a degenerate metric does not define unique geodesics.
    \\[0.3em]
    \textbf{Consequence of failure:} If parameters are functionally dependent (e.g., normalization
    constraints causing effective dimensionality reduction), the Fisher matrix is singular,
    geodesics may not exist or be unique, and $\GeoDist$ is undefined.

    \item \textbf{Small-Connection Approximation Assumption.} \\
    Edge assignments $a_e$ and gauge transformations $h_v$ must be within the
    small-connection limit ($\max_e \dist_{\Od}(A_e, I_d) < r_0$, the matrix logarithm convergence radius).
    Only then does non-abelian gauge-fixing reduce to linear Hodge theory on $\mathfrak{so}(d)$.
    \\[0.3em]
    \textbf{Consequence of failure:} Under large rotations (e.g., $90^\circ$ frame misalignment
    between cross-modal experts), nonlinear Cartan terms cannot be neglected, and the relationship
    between $\GeoDist$ and Euclidean distance breaks.

    \item \textbf{Euclidean Approximation / Degenerate Fisher Assumption.} \\
    For the Euclidean Cercis definition ($\|r\|^2 = \|r_{\text{harm}}\|^2 + \|r_{\text{coexact}}\|^2$)
    to coincide with the Fisher definition, the Fisher metric must degenerate to Euclidean:
    $\Fisher_{ij}(\theta) = \delta_{ij}$. This means all directions on the information manifold
    have equal weight and zero curvature.
    \\[0.3em]
    \textbf{Consequence of failure:} Real SCX parameter spaces almost certainly have non-trivial
    Fisher curvature. A non-negligible curvature correction exists between Euclidean Cercis and
    Fisher geodesic distance.

    \item \textbf{Distribution Compactness / Bounded Support Assumption.} \\
    All expert output distributions must be sufficiently concentrated on a compact
    subset of $\Theta$, so that the local Taylor expansion is uniformly valid across all
    boundary observations and gauge bulk states.
    \\[0.3em]
    \textbf{Consequence of failure:} Heavy-tailed distributions (e.g., Student's-t) may place
    some observations outside the local validity domain, causing uncontrolled growth of
    third-order error terms on individual edges.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{Corrected Claim}
% ══════════════════════════════════════════════════════════════════════

\subsection{English Correction}

\textbf{Original claim (overstated, to be removed):}
\begin{quote}
``The Cercis Score is the geodesic distance under the Fisher information metric.
Under the exponential family assumption, Cercis${}^2 \propto$ mean KL divergence
$=$ squared Fisher geodesic length. The audit score is not artificially defined —
it naturally emerges from the information geometry of the data manifold.''
\end{quote}

\textbf{Corrected claim (conditional equivalence):}
\begin{quote}
\textbf{Under the exponential family assumption and local approximation,} the Cercis Score
equals twice the \textbf{minimum} KL divergence from the observed edge distribution to the
pure-gauge submanifold, which in turn equals (to second order) the squared Fisher geodesic
distance. This provides an information-geometric \textbf{interpretation} — not a ``natural emergence''
but a \textbf{conditional equivalence} that holds when and only when the Situs manifold
admits an exponential family structure.
\textbf{For non-exponential-family data, the relationship degrades by the Amari-Chentsov
tensor (Open Problem~6.2).}
\end{quote}

Explicit formulation:

\vspace{0.5em}
\fbox{%
\begin{minipage}{0.94\textwidth}
\vspace{0.8em}
\noindent \textbf{Theorem (Information-Geometric Conditional Equivalence of Cercis, Revised)}
\label{thm:cercis_conditional_equivalence_en}

\textbf{Assumptions:} All 8 assumptions (H1--H8) hold simultaneously.

\textbf{Conclusion:}
\[
\cercis = \min_{d_0 h \in \im(d_0)} \GeoDist(P_A,\; P_{d_0 h})^2
       = 2 \min_{d_0 h \in \im(d_0)} \KL(P_A \| P_{d_0 h}) + O(\|\mathfrak{a}\|^3).
\]

Cercis measures the minimal Fisher information distance from the observed expert edge
distributions to the ``pure-gauge bulk states'' — the hypothetical world where all
experts can be perfectly aligned.

\textbf{Limitations:}
\begin{enumerate}[label=(\roman*)]
    \item The equivalence is rigorous within exponential families (H1); outside them,
    the difference between $\GeoDist^2$ and $2\KL$ is controlled by the Amari-Chentsov
    third-order tensor $T_{ijk} = \E[\partial_i \log p \cdot \partial_j \log p \cdot \partial_k \log p]$,
    which is generically non-zero for SCX data (Open Problem~\ref{prob:non_exp_fam}).
    \item The equivalence holds under local approximation (H2); for large discrepancies,
    third-order terms dominate.
    \item ``Natural emergence'' is an incorrect characterization — this is a
    \emph{constructed equivalence} under strong assumptions, not an intrinsic
    geometric inevitability of the data.
\end{enumerate}
\vspace{0.5em}
\end{minipage}%
}
\vspace{0.5em}

% ══════════════════════════════════════════════════════════════════════
\section{Five Breakdown Points}
% ══════════════════════════════════════════════════════════════════════

Below, we summarize five principal ways the equivalence can break on real SCX data:

\vspace{0.5em}

\begin{enumerate}[label=\textbf{B\arabic*:}, leftmargin=*, itemsep=0.8em]

    \item \textbf{Non-Exponential-Family Failure (Most Severe).} \\
    SCX expert outputs (MoE routing weights, softmax scores, attention distributions)
    are extremely unlikely to form an exponential family. Exponential families require specific
    algebraic structure (linear sufficient statistics $T(x)$) that deep learning outputs do not
    generally satisfy. The document itself lists this as Open Problem~6.2 (\ref{prob:non_exp_fam}).

    \item \textbf{Confusing ``min KL'' with ``average KL''.} \\
    Cercis solves $\min_h \KL(P_A \| P_{d_0 h})$ — a minimization problem, not an average of
    KL divergences over all expert pairs. These are mathematically distinct.

    \item \textbf{Hidden Euclidean Approximation Requirement.} \\
    The equivalence between Euclidean-norm Cercis and Fisher geodesic distance requires
    $\Fisher_{ij} = \delta_{ij}$ (H7) — meaning the information manifold must be everywhere flat,
    which is extremely unlikely on real SCX parameter spaces.

    \item \textbf{Intrinsic Correction in Theorem~4.2 Proof.} \\
    Choosing the wrong submanifold loses the harmonic component. The correct definition is:
    the bulk submanifold is parameterized by exact connections (pure gauge) $\im(d_0)$,
    not harmonic bulk states $\ker(\Lone)$. Theorem~\ref{thm:cercis_bulk_boundary} identifies
    and corrects this issue in its proof.

    \item \textbf{Locality Constraint.} \\
    All results hold only when $\|\theta_q - \theta_p\|$ is sufficiently small.
    When inter-configuration expert differences are large (e.g., different domains, different
    training stages), the third-order error term $O(\|\Delta\theta\|^3)$ is non-negligible.

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{Recommended Revisions}
% ══════════════════════════════════════════════════════════════════════

\subsection{Changes Needed in gauge\_formalized.tex}

\begin{enumerate}[label=(\arabic*)]

    \item \textbf{Abstract (lines 234--238):}
    Remove ``natural emergence'' language. Change
    \begin{quote}
        ``We prove that under this family, Fisher geodesic length and KL divergence satisfy local equivalence...''
    \end{quote}
    to:
    \begin{quote}
        ``We prove that under the exponential family assumption and local approximation, a conditional equivalence relation holds between Fisher geodesic length and KL divergence...''
    \end{quote}
    Modify the English abstract similarly.

    \item \textbf{Section~\ref{sec:info} title:}
    Change
    \begin{quote}
        ``Fisher Geodesic--KL Equivalence under Exponential Families''
    \end{quote}
    to:
    \begin{quote}
        ``Conditional Fisher Geodesic--KL Equivalence under Exponential Families''
    \end{quote}

    \item \textbf{Theorem~\ref{thm:cercis_bulk_boundary}:}
    Add an explicit list of assumptions before the theorem statement, referencing H1--H8 above. Add a limiting remark at the end of the theorem.

    \item \textbf{Theorem~\ref{thm:unified} (Unified Structure Theorem):}
    Change the phrasing from
    \begin{quote}
        ``Cercis is reformulated as the Fisher geodesic distance from boundary to pure-gauge bulk states, locally equivalent to $2\KL$ under exponential families''
    \end{quote}
    to:
    \begin{quote}
        ``Under the exponential family assumption and local approximation, Cercis can be reinterpreted as the Fisher geodesic distance from boundary to pure-gauge bulk states,
        conditionally equivalent to $2\KL$ (to second order). Behavior outside exponential families is handled by Open Problem~6.2.''
    \end{quote}

    \item \textbf{Comparison table (lines 1529--1532):}
    For
    \begin{quote}
        ``Cercis definition: $\min_{d_0 h} \GeoDist(P_A, P_{d_0 h})^2$'' and
        ``Cercis probabilistic interpretation: $\approx 2\min_{d_0 h} \KL(P_A \| P_{d_0 h})$''
    \end{quote}
    add a footnote or parenthetical note: ``(requires exponential family assumption; degrades outside exponential families)''

    \item \textbf{Section~\ref{sec:comparison} ``Rigor of Information Geometry'' paragraph (lines 1560--1564):}
    Change
    \begin{quote}
        ``Its application to SCX's $\Situs$ manifold in this paper is a natural generalization...''
    \end{quote}
    to:
    \begin{quote}
        ``Its application to SCX's $\Situs$ manifold in this paper is a conditional generalization under specific assumptions.
        When these assumptions are not satisfied (especially the non-exponential-family case, Open Problem~6.2), the equivalence degrades via the Amari-Chentsov tensor.''
    \end{quote}

    \item \textbf{Conclusion item 3 (lines 1678--1684):}
    Change
    \begin{quote}
        ``proved the local equivalence of Fisher geodesic length and KL divergence under exponential families... established an information-geometric interpretation of the Cercis Score''
    \end{quote}
    to:
    \begin{quote}
        ``proved the conditional equivalence of Fisher geodesic length and KL divergence under the exponential family assumption... established a conditional information-geometric interpretation of the Cercis Score, explicitly noting the 8 prerequisite assumptions and failure conditions.''
    \end{quote}

\end{enumerate}

% ══════════════════════════════════════════════════════════════════════
\section{Cross-Reference to Open Problem 6.2}
% ══════════════════════════════════════════════════════════════════════

This corrigendum is directly linked to Open Problem~\ref{prob:non_exp_fam}
(``Fisher-KL Equivalence Outside Exponential Families''), which states:

\begin{quote}
For general SCX output distributions (potentially non-exponential-family),
the relationship between Fisher geodesic distance and KL divergence is controlled
by the Amari-Chentsov third-order tensor $T_{ijk} = \E[\partial_i \log p \cdot \partial_j \log p \cdot \partial_k \log p]$.
Bounds on this tensor for empirical SCX output distributions are needed.
\end{quote}

\vspace{1em}
\noindent\rule{\textwidth}{0.5pt}
\vspace{0.5em}

\noindent {\small \textbf{Review basis:} This corrigendum is based on:
\texttt{docs/reviews/GAUGE\_5VIEWPOINTS\_FINAL.md} (verdict: ``overstated'') and
\texttt{docs/reviews/GAUGE\_VIEWPOINTS\_REVIEW.md} (detailed analysis).
The verdict on Viewpoint~4 is: $\checkmark$ mathematically correct relationship but
``natural emergence'' is severely overstated — downgrade to ``conditional equivalence''.}

\end{document}
